### 2022-03-18 text/Manuscript.Rmd
MS file: updated Yaml

---
title: "Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates UV avoidance in zooplankton"
author: "Kei Jokura, Nobuo Ueda, Martin Gühmann, Gáspár Jékely"
date: "*Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4 4QD, UK*"
output:
  pdf_document: 
      latex_engine: xelatex
      template: style.tex
  html_document: 
    theme: lumen
geometry: left=0.8in,right=0.8in,top=0.8in,bottom=0.5in
mainfont: Arial
documentclass: article
classoption:
  - onecolumn
linestretch: 1.2
fontsize: 10pt
bibliography: references.bib
csl: elife.csl

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Abstract

Nitric oxide is... 
Here we show...

## Introduction

add a reference:

*Platynereis dumerilii* is a... [@ozpolat2021]

## Results

Including Figures

You can add your figures into the rendered document (to define the path, go up one dir from /text and go into /figures):

![Figure 1. Legend goes here](../figures/Figure1.png){width=80%}

## Materials and Methods

## Acknowledgements

This work was funded by the Wellcome Trust (214337/Z/18/Z).

## References {.unnumbered}

::: {#refs}
:::

### 2022-05-06 text/Manuscript.Rmd
manuscript_test

@@
 csl: elife.csl
-
 ---
@@
 
-Nitric oxide is... 
-Here we show...
+Nitric oxide is... Here we show...
 
@@
 
+**Nitric oxide synthase is expressed in neurons in the cPRC circuit** First we performed a nitric oxide synthase (NOS) homology search on Platynereis transcriptome data and identified one NOS gene.
+We examined the gene expression pattern of NOS using in situ hybridization chain reaction (HCR).
+In the 2-day-old larvae, the NOS gene was strongly expressed in four cells, one pair on the dorsal side and one pair on the ventral side (Fig.
+1A).
+NOS cells were labelled using 12 kbp upstream of the transcription start point of the NOS gene and NOS cells were identified as four interneurons in the ciliary-type photoreceptor cells (cPRCs) circuit from Platynereis reconstruction data (Fig.
+1B).
+We named these NOS interneurons, INNOS (INNOSdl, dr, vl, vr) (Fig.
+1C).
+The cPRC circuit has been reported as a neural circuit involved in c-opsin1 mediated UV avoidance behaviour (Verasztó et al., 2017).
+INNOS were presynaptic from the cPRCs and postsynaptic to the RGW interneurons in the cPRC circuit (Fig.
+1D).
+To determine whether NO is involved in UV avoidance behaviour, we generated mutant lines with the CRISPR/Cas-9 system and examined their behavioural responses to UV light.
+
+**NOS mutant delay from UV avoidance** We created two mutant lines with different deletions in the third exon of the NOS gene (NOSΔ11/Δ11 and Δ23/Δ23) (Fig.
+2A).
+Compared to wild-type larvae, homozygous NOS mutant larvae have a defective UV-avoidance response (Fig.
+2B).
+Whereas wild-type larvae showed downward swimming behaviour within 30 seconds in response to lateral UV light (UV-avoidance), the NOS mutant larvae showed a delayed onset of this behaviour and a reduced downward swimming speed (Fig.
+2B).
+These results are similar to previously reported responses of UV-avoidance behaviour in c-opsin1 mutants (Verasztó et al., 2017), suggesting that NO is involved in UV-avoidance behaviour via the cPRC circuit.
+We therefore searched for target molecules for NO.
+
+**GCs with NIT (nitrite/nitrate sensing) domains (NIT-GCs) are expressed in cPRC** Generally, NO can act via cyclic guanosine monophosphate (cGMP) as a second messenger.
+In this signalling pathway, NO binds to the heme group of soluble guanylate cyclase (sGC), a member of the guanylate cyclase family with a CYC domain (PF00211), to increase cGMP.
+However, no NO targets were found in cells that highly express sGC within the cPRC circuit.
+Recently, Moroz and others reported that GC with NIT (nitrite/nitrate sensing) domains (PF08376) (NIT-GC), identified as a sensor for nitrate and nitrite, the oxides of NO, is relatively widely conserved in metazoans (Moroz et al., 2020).
+To search for NIT-GCs in Platynereis, genes with the CYC domains were searched using the transcriptome and genome sequences of 45 metazoan species and choanoflagellate.
+By sequence-similarity searches, multiple sequence alignments and cluster analyses, a cluster of membrane-bound (guanylate cyclase-coupled receptor) type was found and a phylogenetic tree was constructed with soluble type (sGC) as an outgroup.
+As in the phylogenetic analysis of Moroz et al, we found a group of GCs with NIT domains that included many metazoans such as placozoa, cnidaria, ecdysozoa, echinoderms and lophotrochozoa, and in Platynereis 15 NIT-GCs were found (Fig.
+3A).
+To find the candidate NIT-GCs involved in the cPRC circuit among the 15 NIT-GCs, we used previously published spatially mapped single-cell transcriptome data (Achim et al., 2015; Williams et al., 2017) and Two NIT-GCs were found to be highly expressed in the cPRC, and we analysed their expression using in situ HCR as NIT-GC1 and NIT-GC2.
+As a result, strong signals for both were observed specifically in the vicinity of the cPRCs in 2-day-old larvae (Fig.
+3B).
+Both c-opsin1 and MLD pNP (pedal peptide neuropeptide precursor 2), already identified as cPRC markers, and NIT-GC1 and NIT-GC2 were found to be co-expressed.
+They were co-expressed c-opsin1 and MLD pNP (pedal peptide neuropeptide precursor 2), already identified as cPRC markers (Fig.
+3C) (Arendt et al., 2004; Williams et al., 2017).
+To analyse localisations at the protein level, immunostaining was performed using polyclonal antibodies against NIT-GC1 and NIT-GC2.
+Interestingly, NIT-GC1 was found to localise to the neuronal axon and NIT-GC2 to the ciliary region of the cPRC (Fig.
+3D and E).
+Furthermore, co-immunostaining with NOS antibodies revealed that NOS and NIT-GC1 were localised in close proximity in the neurosecretory plexus (Fig.
+3F).
+These results suggest that NIT-GC1 may function as a major target for NO.
+
+**NIT-GC1 produces cGMP by nitric oxide** NIT-GC1 has a highly conserved cyclase domain and activation of NIT-GC1 is expected to produce cGMP.
+Recently, a synthetic protein called Green cGull was developed as a cGMP indicator that can be used in cultured cell systems (Matsuda et al., 2017).
+We co-expressed Green cGull and NIT-GC1 in cultured cells COS-7 (monkey kidney) and tested the intracellular dynamics of cGMP in the activation of NIT-GC1.
+First, S-nitroso-N-acetyl-DL-penicillamine (SNAP; NO donor), 8-Br-cGMP (8-Bromoguanosine 3',5'-cyclic monophosphate; membrane-permeable cGMP analogue, positive control) and DMSO (negative control) were applied to COS-7 cells expressing Green cGull only.
+8-Br-cGMP increased fluorescence intensity by more than sixfold, while little changes were detected for SNAP and DMSO (Fig.
+4A).
+Nest, the two kind of genes, Green cGull and NIT-GC1, were co-expressed using the 2A self-cleaving peptide and the same solutions were applied.
+As a result, a clear increase in cGMP fluorescence intensity was detected in SNAP compared to DMSO or expression of Green cGull only, although not as high as the increase in 8-Br-cGMP (Fig.
+4B).
+Finally, expression of a mutant form of NIT-GC1 with deletion of the NIT domain significantly suppressed the increase in cGMP (Note, we need re-experiment!, Fig.
+4C).
+These results indicate that NIT-GC1 is activated by NO to produce cGMP.
+They also show that the NIT domain is required for the activation of NIT-GC1.
+
+**NO induces depolarisation of the cPRCs via NIT-GC1** To determine the function of NO in the cPRC circuit, the intracellular dynamics of neurons in the cPRC circuit were investigated, focusing on NIT-GC1.
+First, we investigated the changes in NO during UV stimulation of the cPRC using DAF-FM, a fluorescent probe that can detect NO, in order to determine whether NO is produced by UV light.
+By injecting DAF-FM directly into zygotes, INNOS cells could be identified under live imaging.
+To visualise the location of cPRCs, the calcium indicator RGECO1a was also injected together, as cPRCs maintain a higher resting calcium levels compared to other cells.
+Direct UV stimulation to the cPRCs increased the fluorescence intensity of DAF-FM compared to stimulation of non-cPRC region (Fig.
+5A).
+Next, we did calcium imaging of cells in the cPRC circuit during UV stimulation.
+We locally stimulated the cPRC cilia in wild-type with a 405 nm laser (UV) and detected a slight increase followed by a strong decrease in calcium levels (hyperpolarisation) (Fig.
+5B).
+Turning off the UV after or during the decrease caused a strong increase (depolarisation) and a return to the same calcium level as before (Fig.
+5B).
+These responses were similar to those reported previously (Verasztó et al., 2017).
+We performed a similar stimulation using NOS mutants, which hyperpolarised during the UV stimulation, but no depolarising responses were observed when UV was stopped (Fig.
+5C).
+Similarly, in NIT-GC1 knockdown larvae using the morpholinos, no depolarisation after the hyperpolarisation response were observed (Fig.
+5D).
+Interestingly, no hyperpolarising response during UV stimulation was found to occur in NIT-GC2 knockdown larvae.
+And we compared the changes of the INNOS calcium levels identified by their position in relation to cPRCs during UV stimulation in wild-type and in NOS mutants.
+In both cases, a depolarising response was observed at the same time as the hyperpolarising response of the cPRCs (Fig.
+5E).
+Furthermore, to perform calcium imaging of INRGWs in UV stimulation, INRGWs were identified by immunostaining with RGWa antibodies after calcium imaging (Fig.
+5F).
+It has been reported that the INRGW cells form many large synapses with synaptic vesicles on two serotonergic ciliomotor neurons (Ser-h1) that are known to play a role in the regulation of ciliary beating (Verasztó et al., 2017) and the activity of the INRGW cells increased the negative correlation between the activity patterns of serotonergic neurons and the MC neuron (Williams et al., 2018).
+The results showed that depolarisation of INNOS occurs at the same time as hyperpolarisation of cPRC in the wild type.
+Identified INRGW cells were found to cause depolarisation at the same time as cPRC hyperpolarisation in the wild type (Fig.
+5G).
+It was also found that in NOS mutants where cPRC hyperpolarisation does not occur, the INRGW depolarisation response also does not occur (Fig.
+5H).
+
 Including Figures
@@
 
-![Figure 1. Legend goes here](../figures/Figure1.png){width=80%}
+![Figure 1. Legend goes here](../figures/Figure1.png){width="80%"}
 

### 2022-05-09 text/Manuscript.Rmd
redults

@@
 
-**Nitric oxide synthase is expressed in neurons in the cPRC circuit** First we performed a nitric oxide synthase (NOS) homology search on Platynereis transcriptome data and identified one NOS gene.
+**Nitric oxide synthase is expressed in neurons in the cPRC circuit**
+
+First we performed a nitric oxide synthase (NOS) homology search on Platynereis transcriptome data and identified one NOS gene.
 We examined the gene expression pattern of NOS using in situ hybridization chain reaction (HCR).
@@
 
-**NOS mutant delay from UV avoidance** We created two mutant lines with different deletions in the third exon of the NOS gene (NOSΔ11/Δ11 and Δ23/Δ23) (Fig.
+**NOS mutant delay from UV avoidance**
+
+We created two mutant lines with different deletions in the third exon of the NOS gene (NOSΔ11/Δ11 and Δ23/Δ23) (Fig.
 2A).
@@
 
-**GCs with NIT (nitrite/nitrate sensing) domains (NIT-GCs) are expressed in cPRC** Generally, NO can act via cyclic guanosine monophosphate (cGMP) as a second messenger.
+**GCs with NIT (nitrite/nitrate sensing) domains (NIT-GCs) are expressed in cPRC**
+
+Generally, NO can act via cyclic guanosine monophosphate (cGMP) as a second messenger.
 In this signalling pathway, NO binds to the heme group of soluble guanylate cyclase (sGC), a member of the guanylate cyclase family with a CYC domain (PF00211), to increase cGMP.
@@
 
-**NIT-GC1 produces cGMP by nitric oxide** NIT-GC1 has a highly conserved cyclase domain and activation of NIT-GC1 is expected to produce cGMP.
+**NIT-GC1 produces cGMP by nitric oxide**
+
+NIT-GC1 has a highly conserved cyclase domain and activation of NIT-GC1 is expected to produce cGMP.
 Recently, a synthetic protein called Green cGull was developed as a cGMP indicator that can be used in cultured cell systems (Matsuda et al., 2017).
@@
 
-**NO induces depolarisation of the cPRCs via NIT-GC1** To determine the function of NO in the cPRC circuit, the intracellular dynamics of neurons in the cPRC circuit were investigated, focusing on NIT-GC1.
+**NO induces depolarisation of the cPRCs via NIT-GC1**
+
+To determine the function of NO in the cPRC circuit, the intracellular dynamics of neurons in the cPRC circuit were investigated, focusing on NIT-GC1.
 First, we investigated the changes in NO during UV stimulation of the cPRC using DAF-FM, a fluorescent probe that can detect NO, in order to determine whether NO is produced by UV light.
@@
 
+## Discussion
+
+In the present study, we explored the role of NO in the UV avoidance behaviour of Platynereis.
+We demonstrated that there are interneurons expressing NOS in the cPRC circuit and that NO is required for UV avoidance in behavioural experiments using NOS mutants.
+We identified the axon-localised type NIT-GC1 and the cilia-localised type NIT-GC2 expressed in the cPRCs as targets of NO and showed from cultured cell experiments that these produce cGMP in response to NO.
+We have shown by calcium imaging that INNOSs are activated in response to the hyperpolarising response of the cPRCs, that the NO produced feeds back to the cPRC via NIT-GC1 and induces a depolarising response in the cPRC, and that the depolarising response of the cPRC is involved in activating downstream INRGW.
+Lastly, we confirmed that RGWamide induces the downward swimming behaviour of larvae.
+
 ## Materials and Methods
+
+**CRISPR-Cas9 Design and Microinjection**
+
+Before designing the small guide RNA (sgRNA) for the sgRNA:Cas9 nuclease, splice sites and polymorphic sites in our laboratory culture were identified to avoid them.
+The sgRNA targeted the third exon of Platynereis dumerilii NOS (target site: 5'-GGGCAATACTGGCTCCACTC-3').
+The sgRNA was assembled from two annealed oligonucleotides (5'-TAGGGCAATACTGGCTCCACTC-3', 5'-AAACGAGTGGAGCCAGTATTGC-3') forming overhangs for cloning into a BsaI site of the plasmid pDR27456 (Hwang et al.
+2013)(42250, Addgene), which contains next to the BsaI site a tracrRNA sequence.
+The plasmid was then used to PCR amplify DNA (primers: T7, 5'-AAAAGCACCGACTCGGTGCC-3') for synthesizing the sgRNA.
+The DNA was purified with the QIAquick PCR Purification Kit (Qiagen).
+From the DNA, the sgRNA was synthesized with the MEGAshortscript Kit (Thermo Fisher Scientific) and was purified with the MEGAclear Kit (Thermo Fisher Scientific).
+Cas9-mRNA was transcribed, capped, and polyA-tailed with the mMessage mMachine Kit and the Poly(A) Tailing Kit (both Thermo Fisher Scientific) from a plasmid (pUC57-T7-RPP2-Cas9) containing the Cas9 ORF fused to 169 base pair 5' UTR from the Platynereis dumerilii 60S acidic ribosomal protein P2.
+The sgRNA (18 ng/ml) and the Cas9-mRNA (180 ng/µl) were coinjected into fertilized eggs of Platynereis dumerilii wild-type parents according to an established injection procedure (Conzelmann et al.
+2013).
+The eggs were kept at 18°C for 45 min before injection and were injected at 14.5°C.
+The injected individuals were kept at 18°C for 5 to 8 days in 6-well- plates (Nunc multidish no.
+150239, Thermo Scientific) and then cultured at 22°C until sexual maturity.
+The mature worms were crossed to wild-type worms and the progeny was genotyped, resulting in two founder lines, which were bred to homozygosity.
+
+**NOS sequencing and genotyping**
+
+For genotyping of the NOS locus, genomic DNA was isolated from single larvae, groups of 6-20 larvae, or from the tails of adult worms.
+The DNA was amplified by PCR (primers: 5'-GGTTCATTGGTTTCGATAACATTGCGG-3', 5'-CAGAGTCGATCAGTCTGCATATCTCCA-3') with the dilution protocol of the Phusion Human Specimen Direct PCR Kit (Thermo Scientific).
+The PCR product was sequenced directly with a nested sequencing primer (5'-GGTGCTCTCCCGGGTACACAA-3').
+A mixture of wild-type and deletion alleles in a sample gave double peaks in the sequencing chromatograms, with the relative height of the double peaks reflecting the relative allele ratio in the sample.
+
+**Vertical column setup for pharmacology**
+
+Freely swimming 2-day-old larvae were recorded in 5 adjacent vertical columns (each 31 mm x 10 mm x 144 mm water height) with a DMK 21BF03 camera (The Imaging Source) at 30 frames/s.
+The columns were illuminated laterally with red (633 nm) light-emitting diodes (LEDs).
+This light cannot be detected by 2-day-old larvae.
+The larvae were treated with NOS inhibitors and NO donors.
+The NOS inhibitors were L-NAME, SMIS, and AGH.
+The NO donors were SNAP and NONOate.
+The larvae were treated with different concentrations in adjacent columns.
+The concentrations for the NOS inhibitors were control, 10 mM, 1 mM, 0.1 mM, and 0.01 mM; for SNAP were control, DMSO control, 1 mM, 0.1 mM, and 0.01 mM; and for NONOate were control 1 mM, 0.1 mM, and 0.01 mM.
+The larvae were recorded for 30 s then the drugs were added.
+30 s, 1 min, 1 h, and 2 h later, the larvae were recorded again for 30 s.
+The larvae were tracked and their vertical displacement was calculated.
+Scripts are available at <https://github.com/JekelyLab/NOS>.
+
+**Vertical column setup for measuring photoresponses**
+
+Photoresponses of larvae of different ages were assayed in a vertical Plexiglas column (31 mm x 10 mm x 160 mm water height).
+The column was illuminated from top with light from a monochromator (Polychrome II, Till Photonics).
+The monochromator was controlled by AxioVision 4.8.2.0 (Carl Zeiss MicroImaging GmbH) via analog voltage.
+The light passed a collimator lens (LAG-65.0-53.0-C with MgF2 Coating, CVI Melles Griot) before entering the column.
+The column was illuminated from both sides with light-emitting diodes (LEDs).
+The LEDs on each side were grouped into two strips.
+One strip contained UV (395 nm) LEDs (SMB1W-395, Roithner Lasertechnik) and the other infrared (810 nm) LEDs (SMB1W-810NR-I, Roithner Lasertechnik).
+The UV LEDs were run at 4 V to stimulate the larvae in the column from the side.
+The infrared LEDs were run at 8 V (overvoltage) to illuminate the larvae for the camera (DMK 22BUC03, The Imaging Source), which recorded videos at 15 frames per second and was controlled by IC Capture (The Imaging Source).
+
+**Comparing behavior of wildtype and NOS-knockout 3-day-old larvae**
+
+To compare the behavior of wildtype and NOS-knockout larvae at 3 days in the vertical column, the larvae were mixed and left in the dark for 5 min.
+The larvae were recorded for 1 min in the dark followed by exposure to collimated cyan (480 nm) light from the top of the column for 2 min, then 2 min darkness, and finally collimated UV (395 nm) light from the top of the column for 2 min.
+Stimulus light was provided by the monochromator (Polychrome II, Till Photonics).
+Scripts are available at <https://github.com/JekelyLab/NOS>.
+
+**In situ HCR**
+
+Larvae were fixed and treated with Proteinase K, according to the conventional WMISH protocol (Tessmar-Raible et al., 2005), with fixation in 4% paraformaldehyde/ PTW (PBS with 0.05% Tween20) for 2 hr at room temperature, and Proteinase K treatment in 100 µg/ml Proteinase K/ PTW for 3 min (Tessmar-Raible et al., 2005).
+Specifically, for the HCR protocol, samples were processed in 1.5 ml tubes.
+Probe hybridization buffer, probe wash buffer, amplification buffer, and fluorescent HCR hairpins were purchased from Molecular Instruments (Los Angeles, USA).
+Hairpins associated with the b2 initiator sequence were labeled with Alexa Fluor 647, and the hairpins associated with the b3 initiator sequence were labeled with Alexa Fluor 546.
+To design probes for HCR, we used custom software (Kuehn et al., 2021) to create 20 DNA oligo probe pairs specific to P.
+dumerilii NOS, NIT-GC1, NIT-GC2, RYa-pNP (GenBank accession: JF811330.1), c-opsin1 (GenBank accession: AY692353.1), CNGAα (GenBank accession: KM199644.1), and MLD/pedal2-pNP (GenBank accession: KF515945.1).
+The NOS, NIT-GC1 and NIT-GC2 probes were designed to be associated with the b2 initiator sequence, while the RYa-pNP, c-opsin1, CNGAα and MLD/pedal2-pNP probes were designed to be associated with the b3 initiator sequence.
+For the detection stage, samples were pre-hybridized in 200 µl of probe hybridization buffer for 1 hr at 37°C, and then incubated in 250 µl hybridization buffer containing probe oligos (4 pmol/ml) overnight at 37°C.
+To remove excess probe, samples were washed 4× with 1 ml hybridization wash buffer for 15 min at 37°C, and subsequently 2× in 1 ml 5× SSCT (5× SSC with 0.1% Tween20) for 5 min at room temperature.
+For the amplification stage, samples were pre-incubated with 100 µl of amplification buffer for 30 min, room temperature, and then incubated with 150 µl amplification buffer containing fluorescently labeled hairpins (40nM concentration (2ul of 3uM stock in 150ul amplification buffer, snap-cooled as described; Choi et al., 2018) overnight in the dark at 25°C.
+To remove excess hairpins, samples were washed in 1 ml 5× SSCT at room temperature, twice for 5 min, twice for 30 min, and once for 5 min.
+During the first 30 min wash, samples were counterstained with DAPI (Cat.
+#40043, Biotium, USA).
+
+**Immunohistochemistry**
+
+Whole-mount immunostaining of 2 day old Platynereis larvae fixed with 4% paraformaldehyde were carried out using primary antibodies raised against NIT-GC1, NIT-GC2, NOS, RYamide neuropeptide, RGWamide neuropeptide in rabbit, plus a commercial antibody raised against acetylated tubulin in mouse (Sigma T7451).
+The synthetic peptides contained an N-terminal Cys that was used for coupling during purification.
+Antibodies were affinity purified from sera as previously described (Conzelmann and Jékely 2012).
+Immunostainings were carried out as previously described (Conzelmann and Jékely 2012).
+
+**Calcium imaging**
+
+For calcium imaging, 49--55 hpf larvae were used.
+Experiments were performed at room temperature and larvae were immobilised by being embedded in 2.5% agarose filtered artificial seawater between a slide and coverslip spaced with adhesive tape.
+GCaMP6s mRNA (1 mg/ml) was injected into zygotes as described previously (Randel et al., 2014).
+Larvae were imaged on a Zeiss LSM 880 with Airyscan (with a C-Apochromat 63X/1.2 Corr - water) with a frame rate of 1.88 frame/sec and an image size of 512 x 512 pixels.
+The larvae were stimulated in a region of interest (a circle with ??
+pixel diameter) with 405 nm lasers controlled by the Bleaching mode.
+The imaging laser had a similar intensity than the stimulus laser but covered an area that was ??
+times larger than the stimulus ROI.
 

### 2022-05-09 text/Manuscript.Rmd
Method

@@
 
+**Cell culture experiment**
+
+Green cGull was used for the cGMP assay (Matsuda et al., 2017).
+A full-length Pdum-NIT-GC1 and -NIT-GC2 coding sequences were amplified by PCR starting from a Platynereis dumerilii cDNA library and cloned into the pcDNA3.1(+) vector using the T2A self-cleaving sequence.
+Cos-7 cells with low expression of endogenous soluble guanylate cyclase were used as cultured cells for gene expression.
+This cell line was purchased from Angio-proteomie (CAT no.
+cAP-0203).
+The Cos-7 cells were maintained at 37 °C in 35mm dishes (Nunc™ Glass Bottom Dishes) containing 3 mL of DMEM, high glucose glutamax medium (Thermo; Cat.
+No.
+10566016) supplemented with 10% fetal bovine serum (Thermo; Cat.
+No.
+10082147).
+Upon reaching confluency of approximately 85%, we transfected the cells with the plasmid containing Green cGull-T2A-NITGC1.
+Transfections were carried out with 150 ng of each plasmid and 0.3 μl of the transfection Lipofectamine 3000 Reagent (invitrogen; Cat.
+No.
+???).
+Two days post-transfection, we removed the culture medium and substituted it for fresh DMEM-medium.
+For single-wavelength imaging experiments, cells in 35-mm dishes were washed twice and imaged in modified Ringer's buffer (140 mM NaCl, 3.5 mM KCl, 0.5mM NaH2PO4 , 0.5mM S-3 MgSO4 , 1.5 mM CaCl2 , 10 mM HEPES, 2 mM NaHCO3 and 5 mM glucose).
+Dishes were mounted on a stage heated at 37 °C and imaging was performed using an inverted microscope (LSM880, Zeiss) equipped with an oil-immersion objective lens (UApo/340, 40, NA = xx).
+Images were acquired using a xenon lamp, 460--495 nm excitation filter, 505-nm dichroic mirror and 510-- 550-nm emission filter (????).
+8-br-cGMP and S-Nitroso-N-acetyl-D, L-penicillamine (SNAP) were purchased from Sigma-Aldrich (St.
+Louis, MO, USA) and Cayman Chemical (Ann Arbor, MI, USA).
+The exposure time of the EM-CCD camera was controlled by the ZEN software (Zeiss).
+Images were acquired every 15 s for 10 min and stimulation was initiated 2 min after starting image acquisition.
+Imaging data analysis was performed using ImageJ (National Institutes of Health, Bethesda, MD, USA).
+
 ## Acknowledgements

### 2022-05-18 text/Manuscript.Rmd
Figure update

@@
 
+![Figure 1. Legend goes here](../figures/Figure1.png){width="100%"}
+
 **NOS mutant delay from UV avoidance**
@@
 We therefore searched for target molecules for NO.
+
+![Figure 2. Legend goes here](../figures/Figure2.png){width="100%"}
 
@@
 These results suggest that NIT-GC1 may function as a major target for NO.
+
+![Figure 3. Legend goes here](../figures/Figure3.png){width="100%"}
 
@@
 5H).
-
-Including Figures
-
-You can add your figures into the rendered document (to define the path, go up one dir from /text and go into /figures):
-
-![Figure 1. Legend goes here](../figures/Figure1.png){width="80%"}
 

### 2022-05-20 code/220519_NITGC1_cGMP_analysis.R
Figure4 add

@@ -1,0 +1,405 @@
+
+#Clear R's Brain
+rm(list = ls())
+
+library(dplyr)
+library(ggplot2)
+library(tidyverse)
+library(scales)
+library(plotly)
+library(ggplotgui)
+library(tidyr)
+library(esquisse)
+library(ggraptR)
+library(readr)
+
+
+
+
+
+#NIT-GC1_analysis imaging graph
+NITGC1_analysis <- read_csv("data/220519_GcG-T2A-NITGC1_norm.csv")
+
+
+
+
+#抽出:expressionとsolutionからデータを抽出する
+GcG_NIT1_cGMP <- subset(NITGC1_analysis, subset = expression == "GcG-NIT1" 
+                        & solution == "cGMP")
+
+#すべてのデータを確認してみる  
+ggplot(GcG_NIT1_cGMP, aes(x=time, y=intensity)) + 
+  geom_point(size = 1)+
+  geom_line(aes(color=cell), size = 0.8, alpha = 0.5) 
+
+#tapply関数は第1引数に処理対象データ、第2引数に分類データ、第3引数に処理したい関数
+GcG_NIT1_cGMP_mean <- data.frame(time = GcG_NIT1_cGMP$time, 
+                                number = tapply(GcG_NIT1_cGMP$intensity, GcG_NIT1_cGMP$time, length), 
+                                mean = tapply(GcG_NIT1_cGMP$intensity, GcG_NIT1_cGMP$time, mean))
+
+#sd
+GcG_NIT1_cGMP_mean$sd <- tapply(GcG_NIT1_cGMP$intensity, GcG_NIT1_cGMP$time, sd)
+#se
+GcG_NIT1_cGMP_mean$se <- GcG_NIT1_cGMP_mean$sd/sqrt(GcG_NIT1_cGMP_mean$n-1)
+
+#95% confidence interval
+GcG_NIT1_cGMP_mean$CI_lower <- 
+  GcG_NIT1_cGMP_mean$mean + qt((1-0.95)/2, df=GcG_NIT1_cGMP_mean$n-1) * GcG_NIT1_cGMP_mean$se
+GcG_NIT1_cGMP_mean$CI_upper <- 
+  GcG_NIT1_cGMP_mean$mean - qt((1-0.95)/2, df=GcG_NIT1_cGMP_mean$n-1) * GcG_NIT1_cGMP_mean$se
+
+#graph
+ggplot(GcG_NIT1_cGMP_mean) +
[…]

### 2022-05-20 text/Manuscript.Rmd
Figure4 add
Figure5 add

@@
 knitr::opts_chunk$set(echo = TRUE)
+
 ```
@@
 Recently, a synthetic protein called Green cGull was developed as a cGMP indicator that can be used in cultured cell systems (Matsuda et al., 2017).
-We co-expressed Green cGull and NIT-GC1 in cultured cells COS-7 (monkey kidney) and tested the intracellular dynamics of cGMP in the activation of NIT-GC1.
+We co-expressed Green cGull and NIT-GC1 in cultured cells COS-7 (monkey kidney) and tested the intracellular dynamics of cGMP in the activation of NIT-GC1 (Fig.
+4A).
 First, S-nitroso-N-acetyl-DL-penicillamine (SNAP; NO donor), 8-Br-cGMP (8-Bromoguanosine 3',5'-cyclic monophosphate; membrane-permeable cGMP analogue, positive control) and DMSO (negative control) were applied to COS-7 cells expressing Green cGull only.
 8-Br-cGMP increased fluorescence intensity by more than sixfold, while little changes were detected for SNAP and DMSO (Fig.
-4A).
-Nest, the two kind of genes, Green cGull and NIT-GC1, were co-expressed using the 2A self-cleaving peptide and the same solutions were applied.
+S4A).
+Next, the two kind of genes, Green cGull and NIT-GC1, were co-expressed using the 2A self-cleaving peptide and the same solutions were applied.
 As a result, a clear increase in cGMP fluorescence intensity was detected in SNAP compared to DMSO or expression of Green cGull only, although not as high as the increase in 8-Br-cGMP (Fig.
-4B).
+4B, C).
 Finally, expression of a mutant form of NIT-GC1 with deletion of the NIT domain significantly suppressed the increase in cGMP (Note, we need re-experiment!, Fig.
-4C).
+4D).
 These results indicate that NIT-GC1 is activated by NO to produce cGMP.
 They also show that the NIT domain is required for the activation of NIT-GC1.
+
+![Figure 4. Legend goes here](../figures/Figure4.png){width="100%"}
 

### 2022-05-20 code/220413_WTvsNOS11.R
data resource changed

@@ -14,7 +14,7 @@ library(readr)
 library(patchwork)
 
 
-WTvsNOS11_cPRC_INNOS <- read_csv("C:/Users/klamd/Downloads/211129_WTvsNOS11_cPRC_INNOS.csv")
+WTvsNOS11_cPRC_INNOS <- read_csv("data/211129_WTvsNOS11_cPRC_INNOS(2).csv")
 
 WT_cPRC <- WTvsNOS11_cPRC_INNOS %>% select(frame, starts_with("WT_cPRC"))
 NOS11_cPRC <- WTvsNOS11_cPRC_INNOS %>% select(frame, starts_with("NOS11_cPRC"))

### 2022-05-20 code/220414_WTvsNOS23_INRGWa.R

@@ -14,7 +14,7 @@ library(readr)
 library(patchwork)
 
 
-WTvsNOS23_cPRC_INRGWa <- read_csv("C:/Users/klamd/Downloads/211209_WTvsNOS23_cPRC_INRGWa.csv")
+WTvsNOS23_cPRC_INRGWa <- read_csv("data/211209_WTvsNOS23_cPRC_INRGWa(2).csv")
 
 #抽出
 WT_cPRC <- WTvsNOS23_cPRC_INRGWa %>% select(frame, starts_with("WT_cPRC"))

### 2022-05-27 text/Manuscript.Rmd
Add Discussion

@@
 
+![Figure 5. Legend goes here](../figures/Figure5.png){width="100%"}
+
 ## Discussion
@@
 Lastly, we confirmed that RGWamide induces the downward swimming behaviour of larvae.
+
+cPRCs expressing UV-absorbing c-opsin1 have high resting calcium levels and show hyperpolarising and subsequent depolarising responses to UV light (405 nm) (Verasztó et al., 2018).
+In the present study, we have shown that the cells that strongly synapse with cPRCs are interneurons that specifically express NOS (INNOS).
+The fluorescence intensity of DAF-FM, which specifically detected INNOS, showed a transient tonic-like increase dependent on UV stimulation of the cPRCs.
+The graph shows a decrease in intensity after UV light is switched off, but the response of DAF-FM to NO is irreversible and this decrease is due to breaching by the relatively strong excitation light.
+Attempts to measure changes in the intensity of the calcium indicator RGECO1a simultaneously with DAF-FM imaging did not lead to detection of changes in fluorescence intensity compared to GCaMP6s and could only visualise the cPRCs.
+In calcium imaging using GCaMP6s, INNOS identified by the positional relationship between the cPRC and INNOS and the labelling of RYa depolarised in clear synchrony with the hyperpolarising response of the cPRCs.
+This tonic response may involve dopaminergic nerves based on single-cell data (Fig) (Puopolo et al., 2001, Zhang et al., 2007, Randel et al., 2014).
+In the mammalian retina, it is known that signals from bipolar photoreceptor cells are transmitted to downstream NOS-expressing amacrine cells (NOAC), where NO is produced (Jacoby et al., 2018, Neuron).
+These suggest that UV stimulation of cPRCs causes the production of NO via a depolarising response of INNOS.
+
+Our results suggest that NIT-GC1, which is specifically expressed in cPRCs, functions as a target for NO produced by INNOS and contributes to the regulation of the cPRC circuit transmission pathway via a depolarising response by increasing cGMP.
+In the mammalian retina, NO is an essential signalling molecule and one of the most well-studied (Cudeiro and Rivadulla, 1999).
+It has been reported that NO produced in one of the retinal cells modulates cGMP-mediated neurotransmission by activating presynaptic neural sGCs (general target of NO) through retrograde signalling (Vielma et al., 2014; Wei et al., 2012).
+In Platynereis cPRC, NIT-GC1 was also identified as having the property of producing cGMP in a NO-dependent manner.
+Furthermore, it is known that CNGAα, one of the cyclic nucleotide-gated channels, is specifically highly expressed in cPRC (Tosches et al., 2014), and in INNOS-cPRC signalling, NO is implicated in the neuromodulation of the cPRC by retrograde signalling.
+NO is a molecule with a very short half-life, and when functioning as a signalling molecule, NOS and sGC are known to localise very close together at synapses between neurons (Burette et al., 2002; Garthwaite., 2016).
+In our study, we found 12 sGCs in Platynereis, but none expressed in the cPRC.
+Localisation analysis of NOS and NIT-GC1 by antibody staining showed that the two proteins localised very close together in the anterior neurosecretory plexus region.
+Calcium imaging results using NOS mutant and NIT-GC1 knockdown larvae suggest that NIT-GC1 activation induces a depolarising response in the cPRC.
+
+Another type of interneuron that strongly synapses with the cPRC is the INRGWa, which synapses directly onto the ciliomotor neuron and is therefore considered to be a very important neuron in the control of larval behaviour (Verasztó et al., 2018, Williams et al., 2018).
+The present results suggest that the depolarising response of the cPRC due to NO feedback induces INRGW activation.
+RGWamide is also a neuropeptide that induces larval downward swimming quite strongly (unpublished?
+need to show data?), and the delayed downward swimming of NOS mutant larvae in response to UV stimulation may be responsible for the failure of INRGW activation to occur.
+In addition, in many marine invertebrates, NO has been reported to be involved in many aspects of settlement behaviour and metamorphosis induction (1, 2, 3, 4, 5, 6, 7, 8, 9, 10).
+The downward swimming behaviour of Platynereis larvae is the first step in the onset of settlement behaviour and metamorphosis.
+These suggest that downward swimming due to NO activation of INRGWs may trigger the onset of settlement behaviour and metamorphosis (Conzelmann et al., 2013, Williams et al., 2015).
+These suggest that NO feedback signalling may act as a trigger for downward swimming via activation of INRGWs, which induces the initiation of settlement behaviour and metamorphosis.
+
+An outstanding question is how the NO~NIT-GC response alters neuronal dynamics.
+NO is a single molecule and, due to its short half-life, cannot be transmitted over long distances and has a limited signalling repertoire.
+It has therefore been considered unsuitable for generating diversity in signalling compared to other signalling molecules such as neuropeptides and classical neurotransmitters (Jekely, 2021).
+However, our results in the present study indicate that NO-induced activation of NIT-GC1 is involved in the regulation of neuronal circuits (cPRC pathway).
+The NIT domain was originally found in bacteria and the existence of GCs with NIT domains was first reported by Moroz et al in 2020 (Shu et al., 2003; Moroz et al., 2020).
+It has been proposed that NIT domains in bacteria regulate cellular functions in response to changes in extracellular and intracellular nitrate and/or nitrite concentrations (Camargo et al., 2007).
+In fact, NO is a free radical molecule, with a half-life of 5 msec in vivo, and is readily converted to nitrate and nitrite (Santos et al., 2011; Garthwaite., 2016; Möller et al., 2019).
+And nitrate and nitrite have been reported to accumulate in cells and tissues with high NOS activity in placozoans and cnidarians (Moroz et al., 2004; Moroz et al., 2020).
+It is possible that the NIT domains of NIT-GC1 identified in this study may also be used for nitrate/nitrite sensing, as in bacteria.
+Furthermore, if they show different sensitivities to NO, nitrite and nitrate, it is possible that the temporal variation in their respective half-lives could give rise to a repertoire of activation timings (Lundberg  et al., 2011).
+In addition, we have found at least 15 NIT-GCs in the Platynereis transcriptome data, of which NIT-GC2 is highly expressed in the cPRC, in addition to NIT-GC1.
+Interestingly, NIT-GC2 showed a very different localisation pattern to NIT-GC1, and furthermore, the results of calcium imaging of cPRC in response to UV stimulation using NIT-GC2 knockdown larvae were the different dynamics (such that hyperpolarisation was suppressed) with NIT-GC1 knockdown.
+It was not clear from the present analysis whether this difference was due to differences in the localisation or the property of the NIT-GCs.
+However, the diversification and spatially different localisation patterns of NIT-GCs may increase their signalling repertoire.
+We also believe that the appearance of NIT-GCs was one of the means to generate the diversity of signalling pathways using NO during the early stages of nervous system evolution in metazoans, since sponges and ctenophores, where NO signalling is present, have no NIT-GCs at all, as many as 12 have been found in placozoans, and NIT-GCs are functional in the “proto-eye” circuit of Platynereis.
 

### 2022-05-27 code/scRNAseq_Clustered_DotPlots.R
Add figure6

@@ -1,42 +1,0 @@
-#Clear R's Brain
-rm(list = ls())
-
-library(tidyverse)
-library(ggdendro)
-library(patchwork)
-library(ggtree)
-library(cowplot)
-
-
-gene_cluster <- read_csv('E:/NOS project/Single-cell Transcriptome/R Scripts/Seurat_test/220319_scRNAseq_forR.csv')
-
-gene_cluster %>%
-  ggplot(aes(x=cluster, y = Gene, color = count, size = `% Expressing`)) + 
-  geom_point()  
-
-
-gene_cluster %>%
-  filter(count > 0, `% Expressing` > 1) %>% 
-  ggplot(aes(x=cluster, y = Gene, color = count, size = `% Expressing`)) + 
-  geom_point() + 
-  scale_color_viridis_c(name = 'log10') + 
-  cowplot::theme_cowplot() + 
-  theme(axis.line  = element_blank()) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  ylab('') +
-  theme(axis.ticks = element_blank())
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-

### 2022-05-27 text/Manuscript.Rmd
Add figure6
Add references

@@
       latex_engine: xelatex
-      template: style.tex
+      template: "default"
   html_document: 
@@
 
-cPRCs expressing UV-absorbing c-opsin1 have high resting calcium levels and show hyperpolarising and subsequent depolarising responses to UV light (405 nm) (Verasztó et al., 2018).
+cPRCs expressing UV-absorbing c-opsin1 have high resting calcium levels and show hyperpolarising and subsequent depolarising responses to UV light (405 nm) [@Verasztó2018].
 In the present study, we have shown that the cells that strongly synapse with cPRCs are interneurons that specifically express NOS (INNOS).
@@
 In calcium imaging using GCaMP6s, INNOS identified by the positional relationship between the cPRC and INNOS and the labelling of RYa depolarised in clear synchrony with the hyperpolarising response of the cPRCs.
-This tonic response may involve dopaminergic nerves based on single-cell data (Fig) (Puopolo et al., 2001, Zhang et al., 2007, Randel et al., 2014).
-In the mammalian retina, it is known that signals from bipolar photoreceptor cells are transmitted to downstream NOS-expressing amacrine cells (NOAC), where NO is produced (Jacoby et al., 2018, Neuron).
+This tonic response may involve dopaminergic nerves based on single-cell data (Fig) [@puopolo2001][@zhang2007][@Randel2014a].
+In the mammalian retina, it is known that signals from bipolar photoreceptor cells are transmitted to downstream NOS-expressing amacrine cells (NOAC), where NO is produced [@jacoby2018].
 These suggest that UV stimulation of cPRCs causes the production of NO via a depolarising response of INNOS.
@@
 Our results suggest that NIT-GC1, which is specifically expressed in cPRCs, functions as a target for NO produced by INNOS and contributes to the regulation of the cPRC circuit transmission pathway via a depolarising response by increasing cGMP.
-In the mammalian retina, NO is an essential signalling molecule and one of the most well-studied (Cudeiro and Rivadulla, 1999).
-It has been reported that NO produced in one of the retinal cells modulates cGMP-mediated neurotransmission by activating presynaptic neural sGCs (general target of NO) through retrograde signalling (Vielma et al., 2014; Wei et al., 2012).
+In the mammalian retina, NO is an essential signalling molecule and one of the most well-studied [@Cudeiro1999].
+It has been reported that NO produced in one of the retinal cells modulates cGMP-mediated neurotransmission by activating presynaptic neural sGCs (general target of NO) through retrograde signalling [@Vielma2014][@Wei2012].
 In Platynereis cPRC, NIT-GC1 was also identified as having the property of producing cGMP in a NO-dependent manner.
-Furthermore, it is known that CNGAα, one of the cyclic nucleotide-gated channels, is specifically highly expressed in cPRC (Tosches et al., 2014), and in INNOS-cPRC signalling, NO is implicated in the neuromodulation of the cPRC by retrograde signalling.
-NO is a molecule with a very short half-life, and when functioning as a signalling molecule, NOS and sGC are known to localise very close together at synapses between neurons (Burette et al., 2002; Garthwaite., 2016).
+Furthermore, it is known that CNGAα, one of the cyclic nucleotide-gated channels, is specifically highly expressed in cPRC [@Tosches2014], and in INNOS-cPRC signalling, NO is implicated in the neuromodulation of the cPRC by retrograde signalling.
+NO is a molecule with a very short half-life, and when functioning as a signalling molecule, NOS and sGC are known to localise very close together at synapses between neurons [@burette2002][@garthwaite2015a].
 In our study, we found 12 sGCs in Platynereis, but none expressed in the cPRC.
@@
 
-Another type of interneuron that strongly synapses with the cPRC is the INRGWa, which synapses directly onto the ciliomotor neuron and is therefore considered to be a very important neuron in the control of larval behaviour (Verasztó et al., 2018, Williams et al., 2018).
+Another type of interneuron that strongly synapses with the cPRC is the INRGWa, which synapses directly onto the ciliomotor neuron and is therefore considered to be a very important neuron in the control of larval behaviour [@Verasztó2018][@williams2017].
 The present results suggest that the depolarising response of the cPRC due to NO feedback induces INRGW activation.
@@
 need to show data?), and the delayed downward swimming of NOS mutant larvae in response to UV stimulation may be responsible for the failure of INRGW activation to occur.
-In addition, in many marine invertebrates, NO has been reported to be involved in many aspects of settlement behaviour and metamorphosis induction (1, 2, 3, 4, 5, 6, 7, 8, 9, 10).
+In addition, in many marine invertebrates, NO has been reported to be involved in many aspects of settlement behaviour and metamorphosis induction [@leise2001; @bishop2003; @bishop2007; @pechenik2007; @castellano2014; @ueda2013; @ueda2014; @ueda2016; @yang2018; @zhu2020; @song2021].
 The downward swimming behaviour of Platynereis larvae is the first step in the onset of settlement behaviour and metamorphosis.
-These suggest that downward swimming due to NO activation of INRGWs may trigger the onset of settlement behaviour and metamorphosis (Conzelmann et al., 2013, Williams et al., 2015).
+These suggest that downward swimming due to NO activation of INRGWs may trigger the onset of settlement behaviour and metamorphosis [@Conzelmann2013][@Williams2015].
 These suggest that NO feedback signalling may act as a trigger for downward swimming via activation of INRGWs, which induces the initiation of settlement behaviour and metamorphosis.
 
-An outstanding question is how the NO~NIT-GC response alters neuronal dynamics.
+An outstanding question is how the NO\~NIT-GC response alters neuronal dynamics.
 NO is a single molecule and, due to its short half-life, cannot be transmitted over long distances and has a limited signalling repertoire.
-It has therefore been considered unsuitable for generating diversity in signalling compared to other signalling molecules such as neuropeptides and classical neurotransmitters (Jekely, 2021).
+It has therefore been considered unsuitable for generating diversity in signalling compared to other signalling molecules such as neuropeptides and classical neurotransmitters [@Jékely2021].
 However, our results in the present study indicate that NO-induced activation of NIT-GC1 is involved in the regulation of neuronal circuits (cPRC pathway).
-The NIT domain was originally found in bacteria and the existence of GCs with NIT domains was first reported by Moroz et al in 2020 (Shu et al., 2003; Moroz et al., 2020).
-It has been proposed that NIT domains in bacteria regulate cellular functions in response to changes in extracellular and intracellular nitrate and/or nitrite concentrations (Camargo et al., 2007).
-In fact, NO is a free radical molecule, with a half-life of 5 msec in vivo, and is readily converted to nitrate and nitrite (Santos et al., 2011; Garthwaite., 2016; Möller et al., 2019).
-And nitrate and nitrite have been reported to accumulate in cells and tissues with high NOS activity in placozoans and cnidarians (Moroz et al., 2004; Moroz et al., 2020).
+The NIT domain was originally found in bacteria and the existence of GCs with NIT domains was first reported by Moroz et al in 2020 [@Shu2003; @Moroz2020].
+It has been proposed that NIT domains in bacteria regulate cellular functions in response to changes in extracellular and intracellular nitrate and/or nitrite concentrations [@camargo2007].
+In fact, NO is a free radical molecule, with a half-life of 5 msec in vivo, and is readily converted to nitrate and nitrite [@santos2011; @möller2019; @garthwaite2015a].
+And nitrate and nitrite have been reported to accumulate in cells and tissues with high NOS activity in placozoans and cnidarians [@Moroz2004; @Moroz2020].
 It is possible that the NIT domains of NIT-GC1 identified in this study may also be used for nitrate/nitrite sensing, as in bacteria.
-Furthermore, if they show different sensitivities to NO, nitrite and nitrate, it is possible that the temporal variation in their respective half-lives could give rise to a repertoire of activation timings (Lundberg  et al., 2011).
+Furthermore, if they show different sensitivities to NO, nitrite and nitrate, it is possible that the temporal variation in their respective half-lives could give rise to a repertoire of activation timings [@lundberg2011].
 In addition, we have found at least 15 NIT-GCs in the Platynereis transcriptome data, of which NIT-GC2 is highly expressed in the cPRC, in addition to NIT-GC1.
@@
 However, the diversification and spatially different localisation patterns of NIT-GCs may increase their signalling repertoire.
-We also believe that the appearance of NIT-GCs was one of the means to generate the diversity of signalling pathways using NO during the early stages of nervous system evolution in metazoans, since sponges and ctenophores, where NO signalling is present, have no NIT-GCs at all, as many as 12 have been found in placozoans, and NIT-GCs are functional in the “proto-eye” circuit of Platynereis.
+We also believe that the appearance of NIT-GCs was one of the means to generate the diversity of signalling pathways using NO during the early stages of nervous system evolution in metazoans, since sponges and ctenophores, where NO signalling is present, have no NIT-GCs at all, as many as 12 have been found in placozoans, and NIT-GCs are functional in the "proto-eye" circuit of Platynereis.
+
+![Figure 6. Legend goes here](../figures/Figure6.png){width="100%"}
 
@@
 Cas9-mRNA was transcribed, capped, and polyA-tailed with the mMessage mMachine Kit and the Poly(A) Tailing Kit (both Thermo Fisher Scientific) from a plasmid (pUC57-T7-RPP2-Cas9) containing the Cas9 ORF fused to 169 base pair 5' UTR from the Platynereis dumerilii 60S acidic ribosomal protein P2.
-The sgRNA (18 ng/ml) and the Cas9-mRNA (180 ng/µl) were coinjected into fertilized eggs of Platynereis dumerilii wild-type parents according to an established injection procedure (Conzelmann et al.
-2013).
+The sgRNA (18 ng/ml) and the Cas9-mRNA (180 ng/µl) were coinjected into fertilized eggs of Platynereis dumerilii wild-type parents according to an established injection procedure [@Conzelmann2013a].
 The eggs were kept at 18°C for 45 min before injection and were injected at 14.5°C.
@@
 
-Larvae were fixed and treated with Proteinase K, according to the conventional WMISH protocol (Tessmar-Raible et al., 2005), with fixation in 4% paraformaldehyde/ PTW (PBS with 0.05% Tween20) for 2 hr at room temperature, and Proteinase K treatment in 100 µg/ml Proteinase K/ PTW for 3 min (Tessmar-Raible et al., 2005).
+Larvae were fixed and treated with Proteinase K, according to the conventional WMISH protocol [@Tessmar-Raible2005], with fixation in 4% paraformaldehyde/ PTW (PBS with 0.05% Tween20) for 2 hr at room temperature, and Proteinase K treatment in 100 µg/ml Proteinase K/ PTW for 3 min [@Tessmar-Raible2005].
 Specifically, for the HCR protocol, samples were processed in 1.5 ml tubes.
@@
 Hairpins associated with the b2 initiator sequence were labeled with Alexa Fluor 647, and the hairpins associated with the b3 initiator sequence were labeled with Alexa Fluor 546.
-To design probes for HCR, we used custom software (Kuehn et al., 2021) to create 20 DNA oligo probe pairs specific to P.
+To design probes for HCR, we used custom software [@Kuehn2022] to create 20 DNA oligo probe pairs specific to P.
 dumerilii NOS, NIT-GC1, NIT-GC2, RYa-pNP (GenBank accession: JF811330.1), c-opsin1 (GenBank accession: AY692353.1), CNGAα (GenBank accession: KM199644.1), and MLD/pedal2-pNP (GenBank accession: KF515945.1).
@@
 To remove excess probe, samples were washed 4× with 1 ml hybridization wash buffer for 15 min at 37°C, and subsequently 2× in 1 ml 5× SSCT (5× SSC with 0.1% Tween20) for 5 min at room temperature.
-For the amplification stage, samples were pre-incubated with 100 µl of amplification buffer for 30 min, room temperature, and then incubated with 150 µl amplification buffer containing fluorescently labeled hairpins (40nM concentration (2ul of 3uM stock in 150ul amplification buffer, snap-cooled as described; Choi et al., 2018) overnight in the dark at 25°C.
+For the amplification stage, samples were pre-incubated with 100 µl of amplification buffer for 30 min, room temperature, and then incubated with 150 µl amplification buffer containing fluorescently labeled hairpins (40nM concentration (2ul of 3uM stock in 150ul amplification buffer, snap-cooled as described; [@Choi2018]) overnight in the dark at 25°C.
 To remove excess hairpins, samples were washed in 1 ml 5× SSCT at room temperature, twice for 5 min, twice for 30 min, and once for 5 min.
@@
 The synthetic peptides contained an N-terminal Cys that was used for coupling during purification.
-Antibodies were affinity purified from sera as previously described (Conzelmann and Jékely 2012).
-Immunostainings were carried out as previously described (Conzelmann and Jékely 2012).
+Antibodies were affinity purified from sera as previously described [@Conzelmann2012].
+Immunostainings were carried out as previously described [@Conzelmann2012].
 
@@
 
-Green cGull was used for the cGMP assay (Matsuda et al., 2017).
+Green cGull was used for the cGMP assay [@Matsuda2017].
 A full-length Pdum-NIT-GC1 and -NIT-GC2 coding sequences were amplified by PCR starting from a Platynereis dumerilii cDNA library and cloned into the pcDNA3.1(+) vector using the T2A self-cleaving sequence.

### 2022-05-27 text/Manuscript.Rmd
update bib file

@@
 3A).
-To find the candidate NIT-GCs involved in the cPRC circuit among the 15 NIT-GCs, we used previously published spatially mapped single-cell transcriptome data (Achim et al., 2015; Williams et al., 2017) and Two NIT-GCs were found to be highly expressed in the cPRC, and we analysed their expression using in situ HCR as NIT-GC1 and NIT-GC2.
+To find the candidate NIT-GCs involved in the cPRC circuit among the 15 NIT-GCs, we used previously published spatially mapped single-cell transcriptome data [@Achim2015][@williams2017] and Two NIT-GCs were found to be highly expressed in the cPRC, and we analysed their expression using in situ HCR as NIT-GC1 and NIT-GC2.
 As a result, strong signals for both were observed specifically in the vicinity of the cPRCs in 2-day-old larvae (Fig.

### 2022-05-28 text/Manuscript.Rmd

@@
 output:
+  html_document: 
+    theme: lumen
   pdf_document: 
@@
       template: "default"
-  html_document: 
-    theme: lumen
 geometry: left=0.8in,right=0.8in,top=0.8in,bottom=0.5in
@@
 
-![Figure 1. Legend goes here](../figures/Figure1.png){width="100%"}
+![**Figure 1.
+Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** (A and B) Immunostaining with anti-HA antibody against HA tagged palmitoylated-tdTomato expressed under the upstream of the start site of NOS (magenta).
+Co-staining with acetylated α-tubulin antibody as marker for cilia and axonal scaffold (ac α-tub: green).
+INNOSs with the entire cell body and axon labelled are shown with white arrowheads and with only the cell body labelled are shown with black arrowheads.
+Views of larvae at 2-day-old (A) and 3-day-old (B) from the anterior side, respectively.
+(C) Three types of neurons (cPRC, INNOS and INRGWa) in the cPRC circuit reconstructed in the whole-body transmission electron microscopy (ssTEM) dataset of a 3-day-old larva.
+The INNOS on the ventral side labelled with (A) and (B) are shown in purple.
+Other neurons are shown in light grey.
+Axons and dendrites appear as lines and cell body positions are represented by spheres.
+(D and E) Expression analysis of the NOS gene (magenta) using in situ HCR.
+Anterior view of the larva at 2-day-old.
+(D) Co-stained image using acetylated α-tubulin antibody (ac α-tub: green).
+(E) Co-expression analysis image with the RYamide proneuropeptide gene (RYa-pNP: green), which is highly expressed in INNOS.
+White arrowheads indicate co-expressed INNOS.
+Black arrowheads indicate larval eyes.
+White dashed lines indicate the outer frame of the larva.
+(F) Wiring diagram of the cPRC circuit.
+Hexagons represent neuronal groups (numbers indicated in square brackets), arrows represent synaptic connections (numbers indicated in number of synapses); NOS and RGW indicate INNOS and INRGW-dc1 respectively.
+Ciliomotor neuron shows two serotonergic neurons (Ser-h1).](../figures/Figure1.png){width="100%"}
 
@@
 
-![Figure 2. Legend goes here](../figures/Figure2.png){width="100%"}
+![**Figure 2.
+Larvae with inhibited NOS have shorter ciliary arrest times and a delayed response to UV-avoidance.** (A) Percentage of time in which cilia are closed in DMSO-treated larvae (control) vs.
+larvae exposed to NOS inhibitor.
+(B) Top: The domain structure of NOS.
+Bottom: Close-up region showing the genomic locus of NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9.
+The generated mutants (NOSΔ11, NOSΔ23) are also shown.
+Pink indicates target sites.
+Yellow shows PAM sequences, red shows stop codons.
+(C and D) Vertical swimming in wild type and mutant larvae at 2-day-old (C) and 3-day-old (D) stimulated with blue (480 nm) and UV (395 nm) light from above.
+the data are shown in 30 s bins.](../figures/Figure2.png){width="100%"}
 
@@
 
-![Figure 3. Legend goes here](../figures/Figure3.png){width="100%"}
+![**Figure 3.
+cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.** (A) Phylogenetic tree of guanylate cyclase by maximum likelihood (ML).
+Guanylate cyclase-coupled receptor and soluble guanylate cyclase as outgroups.
+Guanylate cyclases with NIT domains are found in most animal phyla except Porifera, Ctenophora, Urochordata and Chordata.
+Tree robustness was tested with 1000 replicates of ultrafast bootstrap.
+Species names are as follows: Ap (Aiptasia pallida), Bb (Branchiostoma belcheri), Cv (Crassostrea virnigica), Dm (Drosophila melanogaster), Dp (Daphnia pulex), Hh (Hoilungia hongkongensis), Lg (Lottia gigantea), Pc (Priapulus caudatus), Pd (Platynereis dumerilii), Sk (Saccoglossus kowalevskii), Sp (Strongylocentrotus purpuratus), Ta (Trichoplax adhaerens), Tc (Tribolium castaneum).
+(B-D) Expression analysis of the NIT-GC1 gene (magenta) using in situ HCR.
+White arrowheads indicate cells expressing NIT-GC1.
+(B) Co-staining image using acetylated α-tubulin antibody (ac α-tub: green).
+(C) Co-expression analysis image of the c-opsin1 gene (green).
+Black arrowheads indicate cells expressing c-opsin1 differently from cPRC.
+(D) Co-expression analysis image of the MLD-pedal2 amide proneuropeptide gene (MLD-pNP: green).
+Black arrowheads indicate cells different from cPRC.
+The white dashed lines in (C) and (D) indicate the outer frame of the larvae.
+(E, F) Localisation analysis using NIT-GC1 and NIT-GC2 antibodies.
+Green shows co-staining with acetylated α-tubulin antibody (ac α-tub).
+(G) Co-localisation analysis using NIT-GC1 and NOS antibodies.
+(B-G) Anterior view of the 2-day-old larvae.](../figures/Figure3.png){width="100%"}
 
@@
 
-![Figure 4. Legend goes here](../figures/Figure4.png){width="100%"}
+![**Figure 4.
+NIT-GC1 is induced by NO to produce cGMP.** (A) Top: The domain structure of NIT-GC1.
+Grey squares indicate the transmembrane domain.
+Bottom shows a schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control.
+Green cGull shows cGMP-dependent green fluorescence.
+(B) The intensity changes in cGMP fluorescence over time for the three conditions are shown.
+Gray traces show the optical responses to individual intensities (n = 6).
+The individual intensities are normalized (ΔF/F0).
+The dark line shows the mean value and the band the standard error range of the mean.
+The respective solution was acted on 2 min after the start of imaging (grey bars).
+(C) The mean intensity (Fmean/F0) after acting on the solution.
+Multiple comparisons were made using the Tukey honestly significant difference test.](../figures/Figure4.png){width="100%"}
 
@@
 
-![Figure 6. Legend goes here](../figures/Figure6.png){width="100%"}
+![**Figure 6.
+Gene expressions and signalling mechanisms between cPRC, INNOS and INRGWa.** (A) Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq.
+The size of the dots is expressed in proportion to the percentage of cells expressing that gene relative to all cells.
+The colours represent the normal logarithm of the number of transcripts in the cells expressing the gene.
+(B) Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback.](../figures/Figure6.png){width="100%"}
 

### 2022-05-29 text/Manuscript.Rmd

@@
 output:
-  html_document: 
-    theme: lumen
   pdf_document: 
@@
       template: "default"
+  html_document: 
+    theme: lumen
 geometry: left=0.8in,right=0.8in,top=0.8in,bottom=0.5in
@@
 
-![Figure 5. Legend goes here](../figures/Figure5.png){width="100%"}
+![**Figure 5.
+NO produced by UV stimulation to cPRCs induces a depolarising response in cPRC via NIT-GC1.** (A) An INNOS visualised by DAF-FM is shown (a white arrowhead).
+A black arrowhead indicates the position of the adult eye.
+The white dashed line indicates the outer frame of the larva.
+All subsequent images are of 2-day-old larvae.
+(B) Dual-colour imaging of DAF-FM (green) and RGECO1a (magenta) allows simultaneous visualisation of the INNOS and cPRCs.
+(C) The intensity of DAF-FM in INNOS changes over time when cPRC (top) and other cells (NC) (bottom) are stimulated with UV (405 nm).
+Hereafter, light purple squares indicate the timing of UV stimulation, gray traces show the optical responses to individual intensities, each intensities are normalized (ΔF/F0) and the dark lines and the bands show the mean value and the standard error range of the mean.
+(D) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in WT (purple) and NOS mutant (NOSΔ23) (gray) larvae.
+(E) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in control (non-injection) and NIT-GC1 knockdown by morpholino (NITGC1-MO1 and 2) larvae.
+(F) Co-expression analysis image of the cPRC marker gene (NIT-GC1: magenta) and the INNOS marker gene (RYa-pNP: green) using in situ HCR.
+White arrowheads indicate INNOS strongly expressing RYa-pNP.
+(G) INNOS (blue dashed line) identified from calcium imaging using GCaMP6s by their position in relation to the cPRC (purple dashed line).
+(H) Comparison of changes over time in GCaMP6s intensities in cPRC (purple) and INNOS (blue) when cPRC is stimulated with UV in WT (left) and NOS mutant (NOSΔ11: right) larvae.
+(I) Schematic diagram of the immunostaining procedure after calcium imaging.
+(J) The position of the cPRC (purple dashed line) in calcium imaging with GCaMP6s.
+Cells around the cPRC are numbered according to the position of the nuclei.
+(K) Immunostaining image using RGWamide (RGWa) and serotonin (Ser) antibodies (red) and DAPI (light blue) in the same region as (J).
+INRGW (green dashed line) and Ser-h1 (black arrowheads) identified from the numbered nuclear positions are shown.
+(L) Comparison of changes over time in GCaMP6s intensities in cPRC (purple) and INRGW (green) when the cPRC are stimulated with UV in WT (left) and NOS mutant (NOSΔ23: right) larvae.](../figures/Figure5.png){width="100%"}
 

### 2022-06-01 text/Manuscript.Rmd

@@
 
-add a reference:
-
-*Platynereis dumerilii* is a... [@ozpolat2021]
+The ability to perceive changes in the environment and adapt behaviour accordingly is essential for the survival of all living organisms.
+Avoidance from UV radiation, which damages DNA essential for survival, is proposed to be one of the behaviours that existed before the abundant development of Earth's oxygen to block UV radiation in the atmosphere (Pittendrigh).
+In the ocean, almost all zooplankton have a pronounced rhythmic behaviour called Diel vertical migration (DVM), which consists of upwards towards the surface at dusk and descend back to deeper water before dawn.
+Avoidance behaviour to escape harmful UV radiation during the day is closely linked to the control of vertical migration in the ocean.
+In metamorphosis, a life history strategy widely found in metazoans, downward swimming towards the seafloor is known to be associated with larval settlement and the initiation of metamorphosis, a dramatic body change from the larval stage, and UV-avoidance behaviour allows dispersal and the acquisition of different ecological niches.
+However, little is known about the molecular neural circuit mechanisms in UV avoidance behaviour in many animals, including zooplankton.
+In neural circuits in general light response, the activation of various receptors by excitatory and inhibitory neurotransmitters such as glutamate, GABA and dopamine is known in various species.
+In addition to this, it has long been known that in nitric oxide (NO), as a neuromodulatory molecule, it is also important as a modulator of visual activity.
+NO is a small free-radical molecule with diffusible and changeable characteristics, and thus works as one of the neural signalling molecules.
+The function of NO in the neural circuitry of vision has been well studied in vertebrates.
+In the mammalian retina, NO is an essential signalling molecule in all retinal cells.
+NO produced by nitric oxide synthase (NOS) in amacrine cells, one of the retinal cells, activates soluble guanylate cyclase (sGC) expressed in cone bipolar cells, which produces cGMP and regulates downstream signalling.
+On the other hand, the role of NO in zooplankton has been reported in many animals, including sponges, cnidarians, molluscs, arthropods, annelids, echinoderms and ascidians, as a molecule that controls settlement and metamorphosis initiation.
+We therefore hypothesise that NO-mediated neural signalling may be involved in the photoreceptor circuits that control UV avoidance behaviour.
+To test this hypothesis we used the annelid *Platynereis dumerilii*, which conserved vertebrate-type ciliary photoreceptors (cPRCs) [@Arendt2004].
+The cPRC expresses c-opsin1, which absorbs UV radiation, is involved in melatonin production and regulates circadian swimming behaviour [@Tosches2014; @tsukamoto2017].
+Furthermore, we previously reported abnormal UV avoidance behaviour in c-opsin1 knockout larvae [@Verasztó2018].
+These suggest that Platynereis larvae are a highly suitable model for elucidating the neural circuit mechanisms of NO-induced UV avoidance behaviour, and the accessibility of genome editing, behavioural assays, calcium imaging and whole-body connectomics could be very useful for study at the molecular level[@ozpolat2021].
 

### 2022-06-06 text/Manuscript.Rmd
Figs ms updates

@@
 Hexagons represent neuronal groups (numbers indicated in square brackets), arrows represent synaptic connections (numbers indicated in number of synapses); NOS and RGW indicate INNOS and INRGW-dc1 respectively.
-Ciliomotor neuron shows two serotonergic neurons (Ser-h1).](../figures/Figure1.png){width="100%"}
+Ciliomotor neuron shows two serotonergic neurons (Ser-h1).](../figures/Fig1.png){width="100%"}
 
@@
 (C and D) Vertical swimming in wild type and mutant larvae at 2-day-old (C) and 3-day-old (D) stimulated with blue (480 nm) and UV (395 nm) light from above.
-the data are shown in 30 s bins.](../figures/Figure2.png){width="100%"}
+the data are shown in 30 s bins.](../figures/Fig2.png){width="100%"}
 
@@
 (G) Co-localisation analysis using NIT-GC1 and NOS antibodies.
-(B-G) Anterior view of the 2-day-old larvae.](../figures/Figure3.png){width="100%"}
+(B-G) Anterior view of the 2-day-old larvae.](../figures/Fig3.png){width="100%"}
 
@@
 (C) The mean intensity (Fmean/F0) after acting on the solution.
-Multiple comparisons were made using the Tukey honestly significant difference test.](../figures/Figure4.png){width="100%"}
+Multiple comparisons were made using the Tukey honestly significant difference test.](../figures/Fig4.png){width="100%"}
 
@@
 INRGW (green dashed line) and Ser-h1 (black arrowheads) identified from the numbered nuclear positions are shown.
-(L) Comparison of changes over time in GCaMP6s intensities in cPRC (purple) and INRGW (green) when the cPRC are stimulated with UV in WT (left) and NOS mutant (NOSΔ23: right) larvae.](../figures/Figure5.png){width="100%"}
+(L) Comparison of changes over time in GCaMP6s intensities in cPRC (purple) and INRGW (green) when the cPRC are stimulated with UV in WT (left) and NOS mutant (NOSΔ23: right) larvae.](../figures/Fig5.png){width="100%"}
 

### 2022-06-06 code/Fig1.R

@@ -233,8 +233,8 @@ A#B#C
 D#E#F
 "
 
-Fig1 <- panel_SEM + panel_INNOS_ventr + panel_INNOS_ant +
-  panel_NOS_HCR + panel_NOS3d + panel_Network +
+Fig1 <- panel_SEM + 
+  panel_NOS_HCR + panel_NOS3d + panel_INNOS_ventr + panel_INNOS_ant + panel_Network +
   patchwork::plot_layout(design = layout, heights = c(1,0.02,1), widths = c(1,0.02,1,0.02,1)) + #we can change the heights of the rows in our layout (widths also can be defined)
   patchwork::plot_annotation(tag_levels = "A") &  #we can change this to 'a' for small caps or 'i' or '1'
   ggplot2::theme(plot.tag = element_text(size = 12, 

### 2022-06-06 code/Fig6.R

@@ -32,22 +32,22 @@ NIT2_MO_tb%>%
 
 # plot NITGC1_analysis ----------------------------------------------------
 NITGC1_analysis %>%
-  count(solution)
+  count(expression)
 
 NITGC1_analysis %>%
-  filter(solution == "SNAP") %>%
+  filter(solution == "SNAP" & expression == "GcG-NIT1") %>%
 #cell names are not unique and the same across expression, fix by adding new column  
   mutate(cell_expression = paste(cell, expression, sep = "")) %>%
   ggplot(aes(x = time, y = intensity, color = expression)) +
-  geom_line(aes(group = cell_expression), size = 0.4, alpha = 0.2) +
+  geom_line(aes(group = cell_expression), size = 0.4, alpha = 0.2, 
+            show.legend = FALSE) +
   geom_smooth(level = 0.99, size = 1, span = 0.2, 
-              method = "loess") +
-  theme_void()  +
+              method = "loess", show.legend = FALSE) +
+  theme_minimal()  +
   theme(legend.title = element_blank(), legend.text.align=0,
         legend.text = element_text(size=10)) +
   #Specify colours and legend labels
-  scale_color_manual(values=c(Okabe_Ito[6], Okabe_Ito[2]),
-                     labels = c('GcG','GcG + NIT-GC1')) +
+  scale_color_manual(values=c(Okabe_Ito[2])) +
   scale_y_continuous(breaks=seq(1,1.15,length=4),limits = c(0.95, 1.15))+
   scale_x_continuous(breaks=seq(0,10,length=6),limits = c(0, 10))+
   annotate("segment", x=2, xend=10, y=1.135, yend=1.135, size=2, color = "gray")+
@@ -57,7 +57,32 @@ NITGC1_analysis %>%
 # save plot ---------------------------------------------------------------
 
 ggsave("pictures/GcG-NIT-GC1-SNAP.png", limitsize = FALSE, 
-       units = c("px"), width = 1600, height = 600, bg='white')  
+       units = c("px"), width = 800, height = 600, bg='white')  
+
+# plot NITGC1_analysis ----------------------------------------------------
+NITGC1_analysis %>%
+  count(expression)
+
+NITGC1_analysis %>%
+  filter(solution == "SNAP" & expression == "GcG") %>%
[…]

### 2022-06-12 text/Manuscript.Rmd
MS: results section 1 edits

@@
 
-**Nitric oxide synthase is expressed in neurons in the cPRC circuit**
+**Nitric oxide synthase is expressed in interneurons of the UV-avoidance circuit**
 
-First we performed a nitric oxide synthase (NOS) homology search on Platynereis transcriptome data and identified one NOS gene.
-We examined the gene expression pattern of NOS using in situ hybridization chain reaction (HCR).
-In the 2-day-old larvae, the NOS gene was strongly expressed in four cells, one pair on the dorsal side and one pair on the ventral side (Fig.
+We identified a single *nitric oxide synthase* (*NOS*) gene in the *Platynereis dumerilii* genome and transcriptome data.
+Phylogenetic analysis of NOS proteins indicate that the *Platynereis* NOS belongs to an orthology group of metazoan NOS sequences [need to add tree].
+To characterise the expression pattern of *NOS* we used in situ hybridization chain reaction (HCR) and transient transgenesis.
+In two-day-old larvae, we detected *NOS* expression in four cells in the apical organ region (Fig.
 1A).
-NOS cells were labelled using 12 kbp upstream of the transcription start point of the NOS gene and NOS cells were identified as four interneurons in the ciliary-type photoreceptor cells (cPRCs) circuit from Platynereis reconstruction data (Fig.
+The same four cells cells were labelled with a reporter construct driving palmitoylated-tdTomato by a 12 kbp upstream genomic region of the *NOS* gene.
+This reporter also revealed the axonal projections of the NOS-expressing neurons.
+The position and axonal projections of the four *NOS*-expressing cells allowed us to identify the same cells as four interneurons in the ciliary-type photoreceptor cells (cPRCs) circuit from Platynereis reconstruction data (Fig.
 1B).
@@
 (C) The mean intensity (Fmean/F0) after acting on the solution.
-Multiple comparisons were made using the Tukey honestly significant difference test.](../figures/Fig4.png){width="100%"}
+Multiple comparisons were made using the Tukey honestly significant difference test.](../figures/Fig6.png){width="100%"}
 

### 2022-06-12 text/Manuscript.Rmd
MS: results 1 edits

@@
 We identified a single *nitric oxide synthase* (*NOS*) gene in the *Platynereis dumerilii* genome and transcriptome data.
-Phylogenetic analysis of NOS proteins indicate that the *Platynereis* NOS belongs to an orthology group of metazoan NOS sequences [need to add tree].
+Phylogenetic analysis of NOS proteins indicate that the *Platynereis* NOS belongs to an orthology group of metazoan NOS sequences [@Kei - need to add tree].
 To characterise the expression pattern of *NOS* we used in situ hybridization chain reaction (HCR) and transient transgenesis.
 In two-day-old larvae, we detected *NOS* expression in four cells in the apical organ region (Fig.
-1A).
+1A) [@Kei - 3d HCR and transgene?].
 The same four cells cells were labelled with a reporter construct driving palmitoylated-tdTomato by a 12 kbp upstream genomic region of the *NOS* gene.
 This reporter also revealed the axonal projections of the NOS-expressing neurons.
-The position and axonal projections of the four *NOS*-expressing cells allowed us to identify the same cells as four interneurons in the ciliary-type photoreceptor cells (cPRCs) circuit from Platynereis reconstruction data (Fig.
-1B).
-We named these NOS interneurons, INNOS (INNOSdl, dr, vl, vr) (Fig.
-1C).
-The cPRC circuit has been reported as a neural circuit involved in c-opsin1 mediated UV avoidance behaviour (Verasztó et al., 2017).
-INNOS were presynaptic from the cPRCs and postsynaptic to the RGW interneurons in the cPRC circuit (Fig.
-1D).
+The position and axonal projections of the four *NOS*-expressing cells allowed us to identify the same cells in the three-day-old connectome data as four interneurons (INNOS) in the ciliary photoreceptor-cell (cPRC) circuit (Figure 1D-E).
+The cPRCs express the UV-absorbing c-opsin1 and mediate a UV avoidance behaviour in the larvae [@Verasztó2018] and modulates locomotion and maturation in adult worms [@Veedin2021].
+INNOS cells are postsynaptic to the cPRCs and presynaptic to the RGW interneurons in the cPRC circuit (Fig.
+1F).
+
 To determine whether NO is involved in UV avoidance behaviour, we generated mutant lines with the CRISPR/Cas-9 system and examined their behavioural responses to UV light.

### 2022-06-12 text/Manuscript.Rmd
MS and bib edits

--- 2022-06-12 text/Manuscript.Rmd
@@
 However, little is known about the molecular neural circuit mechanisms in UV avoidance behaviour in many animals, including zooplankton.
+The cPRCs express the UV-absorbing c-opsin1 and mediate UV avoidance in the larvae [@veraszto2018].
+In adult worms, UV light and c-opsin1 modulate locomotion and maturation [@Veedin2021].
 In neural circuits in general light response, the activation of various receptors by excitatory and inhibitory neurotransmitters such as glutamate, GABA and dopamine is known in various species.
@@
 The cPRC expresses c-opsin1, which absorbs UV radiation, is involved in melatonin production and regulates circadian swimming behaviour [@Tosches2014; @tsukamoto2017].
-Furthermore, we previously reported abnormal UV avoidance behaviour in c-opsin1 knockout larvae [@Verasztó2018].
+Furthermore, we previously reported abnormal UV avoidance behaviour in c-opsin1 knockout larvae [@Veraszto2018].
 These suggest that Platynereis larvae are a highly suitable model for elucidating the neural circuit mechanisms of NO-induced UV avoidance behaviour, and the accessibility of genome editing, behavioural assays, calcium imaging and whole-body connectomics could be very useful for study at the molecular level[@ozpolat2021].
@@
 We identified a single *nitric oxide synthase* (*NOS*) gene in the *Platynereis dumerilii* genome and transcriptome data.
-Phylogenetic analysis of NOS proteins indicate that the *Platynereis* NOS belongs to an orthology group of metazoan NOS sequences [need to add tree].
+Phylogenetic analysis of NOS proteins indicate that the *Platynereis* NOS belongs to an orthology group of metazoan NOS sequences (#Kei - need to add tree).
 To characterise the expression pattern of *NOS* we used in situ hybridization chain reaction (HCR) and transient transgenesis.
 In two-day-old larvae, we detected *NOS* expression in four cells in the apical organ region (Fig.
-1A).
+1A) (#Kei - 3d HCR and transgene?).
 The same four cells cells were labelled with a reporter construct driving palmitoylated-tdTomato by a 12 kbp upstream genomic region of the *NOS* gene.
 This reporter also revealed the axonal projections of the NOS-expressing neurons.
-The position and axonal projections of the four *NOS*-expressing cells allowed us to identify the same cells as four interneurons in the ciliary-type photoreceptor cells (cPRCs) circuit from Platynereis reconstruction data (Fig.
-1B).
-We named these NOS interneurons, INNOS (INNOSdl, dr, vl, vr) (Fig.
-1C).
-The cPRC circuit has been reported as a neural circuit involved in c-opsin1 mediated UV avoidance behaviour (Verasztó et al., 2017).
-INNOS were presynaptic from the cPRCs and postsynaptic to the RGW interneurons in the cPRC circuit (Fig.
-1D).
-To determine whether NO is involved in UV avoidance behaviour, we generated mutant lines with the CRISPR/Cas-9 system and examined their behavioural responses to UV light.
+The position and morphology of the four *NOS*-expressing cells allowed us to identify the same cells in the three-day-old whole-body volume EM data [@williams2017; @veraszto2020] as four interneurons (INNOS)(Figure 1D-E).
+The INNOS cells are postsynaptic to the UV-sensory cPRCs and presynaptic to the INRGW interneurons, which are also cPRC targets (Fig.
+1F).
 
@@
 
-**NOS mutant delay from UV avoidance**
+**Nitric oxide produced during UV/violet stimulation gates the output of the cPRC circuit**
 
-We created two mutant lines with different deletions in the third exon of the NOS gene (NOSΔ11/Δ11 and Δ23/Δ23) (Fig.
-2A).
-Compared to wild-type larvae, homozygous NOS mutant larvae have a defective UV-avoidance response (Fig.
-2B).
-Whereas wild-type larvae showed downward swimming behaviour within 30 seconds in response to lateral UV light (UV-avoidance), the NOS mutant larvae showed a delayed onset of this behaviour and a reduced downward swimming speed (Fig.
-2B).
-These results are similar to previously reported responses of UV-avoidance behaviour in c-opsin1 mutants (Verasztó et al., 2017), suggesting that NO is involved in UV-avoidance behaviour via the cPRC circuit.
-We therefore searched for target molecules for NO.
+The specific expression of NOS in interneurons directly postsynaptic to the cPRCs suggests that NO signalling is involved in the UV response.
+To test if NO is produced during UV/violet responses, we injected zygotes with the fluorescent NO-reporter DAF-FM and imaged two-day-old larvae during the local stimulation of the cPRC cilia by 405 nm light.
+Following light stimulation, we detected an increase in DAF-FM fluorescence in the region of the INNOS cells.
+(#Kei - are there time course data that could be shown?).
+We next analysed the dynamics of the cPRC circuit following light stimulation.
+As we have shown previously [@Veraszto2018], a 30-sec illumination of the cPRC ciliary region lead to a decrease in cPRC calcium levels, likely due to the hyperpolarisation of the cPRCs.
+After 30 min, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus - a response that may involve depolarisation.
+This activation phase ocurres after the end of the stimulation and may be due to a slower neuroendocrine feedback [@Veraszto2018].
+We also quantified in more detail the responses of the INNOS and INRGW interneurons under a UV/violet stimulus to the cPRCs.
+To unambiguously identify the cells from which we recorded calcium signals, we developed an on-slide staining method to directly fix, immunostain and image the same larva that we imaged previously.
+We used the cell-specific neuropeptide markers anti-RYamide (INNOS) and anti-RGWa (INRGW) to mark the interneurons.
+Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision.
+INNOS cells showed increase calcium signals during the inhibition phase of the cPRCs.
+In contrast, INRGW cells were only active during the second activation phase of cPRCs.\
+To determine whether NO is involved in the observed circuit dynamics, we generated two *Platynereis* *NOS* knockout lines with the CRISPR/Cas9 method.
+We recovered two deletions (NOSΔ11/Δ11 and Δ23/Δ23), both leading to an early stop codon and thus likely representing null alleles (Figure 3A).
+We could establish a homozygous line for both mutations indicating that NOS is not an essential gene.
+Calcium imaging in *NOS*-mutant larvae revealed perturbed cPRC responses and circuit activation.
+In cPRCs, we could detect the inhibition phase but this was not followed by activation and calcium levels remained low.
+We did not detect a difference in the response of the INNOS cells between wild-type and *NOS*-mutant larvae.
+In contrast, INRGW cells did not activate in *NOS*-mutant larvae, suggesting that they will not initiate a circuit output (to Ser-h1 cells and the ciliary band).
+
+**Nitric oxide signalling mediates UV-avoidance behaviour** Our calcium imaging data suggest that *NOS*-mutant larvae will have defective UV-avoidance behaviour.
+To test this, we recorded the trajectories of freely swimming larvae in vertical columns, illuminated from two lateral sides with xy nm violet light.
+As previously shown, wild-type larvae swim downward within 30 sec following UV/violet light stimulation.
+In contrast, homozygous *NOS*-mutant larvae showed a strongly diminished UV-avoidance response (Figure 3B).
+This phenotype is similar to the defective UV-avoidance of *c-opsin1* mutant larvae.
 
@@
 
-**GCs with NIT (nitrite/nitrate sensing) domains (NIT-GCs) are expressed in cPRC**
+**Two unconventional guanylyl cyclases (NIT-GC) are expressed in the cPRCs**
 
@@
 5B).
-These responses were similar to those reported previously (Verasztó et al., 2017).
+These responses were similar to those reported previously [@Veraszto2018].
 We performed a similar stimulation using NOS mutants, which hyperpolarised during the UV stimulation, but no depolarising responses were observed when UV was stopped (Fig.
@@
 5F).
-It has been reported that the INRGW cells form many large synapses with synaptic vesicles on two serotonergic ciliomotor neurons (Ser-h1) that are known to play a role in the regulation of ciliary beating (Verasztó et al., 2017) and the activity of the INRGW cells increased the negative correlation between the activity patterns of serotonergic neurons and the MC neuron (Williams et al., 2018).
+It has been reported that the INRGW cells form many large synapses with synaptic vesicles on two serotonergic ciliomotor neurons (Ser-h1) that are known to play a role in the regulation of ciliary beating and the activity of the INRGW cells increased the negative correlation between the activity patterns of serotonergic neurons and the MC neuron (Williams et al., 2018).
 The results showed that depolarisation of INNOS occurs at the same time as hyperpolarisation of cPRC in the wild type.
@@
 
-cPRCs expressing UV-absorbing c-opsin1 have high resting calcium levels and show hyperpolarising and subsequent depolarising responses to UV light (405 nm) [@Verasztó2018].
+cPRCs expressing UV-absorbing c-opsin1 have high resting calcium levels and show hyperpolarising and subsequent depolarising responses to UV light (405 nm) [@Veraszto2018].
 In the present study, we have shown that the cells that strongly synapse with cPRCs are interneurons that specifically express NOS (INNOS).
@@
 
-Another type of interneuron that strongly synapses with the cPRC is the INRGWa, which synapses directly onto the ciliomotor neuron and is therefore considered to be a very important neuron in the control of larval behaviour [@Verasztó2018][@williams2017].
+Another type of interneuron that strongly synapses with the cPRC is the INRGWa, which synapses directly onto the ciliomotor neuron and is therefore considered to be a very important neuron in the control of larval behaviour [@Veraszto2018][@williams2017].
 The present results suggest that the depolarising response of the cPRC due to NO feedback induces INRGW activation.

### 2022-06-14 text/Manuscript.Rmd
MS: added abstract, added comments with Markdown syntax

@@
 ---
-title: "Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates UV avoidance in zooplankton"
+title: Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates UV avoidance
+  in zooplankton
 author: "Kei Jokura, Nobuo Ueda, Martin Gühmann, Gáspár Jékely"
-date: "*Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4 4QD, UK*"
+date: '*Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4
+  4QD, UK*'
 output:
@@
   pdf_document:
-      latex_engine: xelatex
-      template: "default"
+    latex_engine: lualatex
+    template: default
+  word_document: default
 geometry: left=0.8in,right=0.8in,top=0.8in,bottom=0.5in
 mainfont: Arial
+monofont: Times New Roman
 documentclass: article
-classoption:
-  - onecolumn
+classoption: onecolumn
 linestretch: 1.2
@@
 
-Nitric oxide is... Here we show...
+Nitric oxide (NO) generated by nitric-oxide synthase (NOS) is a key regulator of animal physiology.
+In nervous systems, NO can modulate circuits, but its effects on neural dynamics and behaviour remain unclear.
+Here we show how NO signalling shapes circuit dynamics to orchestrate light-avoidance behaviour.
+We studied UV-avoidance mediated by brain ciliary photoreceptor cells (cPRCs) in larval *Platynereis dumerilii*, a marine annelid.
+We found NOS specifically expressed in interneurons (INNOS) postsynaptic to cPRCs.
+Stimulation of cPRCs by violet light leads to cPRC inhibition but concomitant INNOS activation and NO production.
+NO feeds back to cPRCs and triggers delayed cPRC activation through an unconventional guanylyl cyclase.
+This results in the activation of a second population of postsynaptic interneurons (INRGW).
+In *NOS* knockout larvae, NO feedback and INRGW activation do not occur and avoidance behaviour is defective.
+Our results reveal how NO-mediated neuroendocrine signalling gates a synaptic circuit to initiate light-avoidance behaviour.
 
@@
 However, little is known about the molecular neural circuit mechanisms in UV avoidance behaviour in many animals, including zooplankton.
-The cPRCs express the UV-absorbing c-opsin1 and mediate UV avoidance in the larvae [@veraszto2018].
-In adult worms, UV light and c-opsin1 modulate locomotion and maturation [@Veedin2021].
 In neural circuits in general light response, the activation of various receptors by excitatory and inhibitory neurotransmitters such as glutamate, GABA and dopamine is known in various species.
@@
 NO produced by nitric oxide synthase (NOS) in amacrine cells, one of the retinal cells, activates soluble guanylate cyclase (sGC) expressed in cone bipolar cells, which produces cGMP and regulates downstream signalling.
-On the other hand, the role of NO in zooplankton has been reported in many animals, including sponges, cnidarians, molluscs, arthropods, annelids, echinoderms and ascidians, as a molecule that controls settlement and metamorphosis initiation.
+On the other hand, the role of NO in zooplankton has been reported in many animals, including sponges, cnidarians, molluscs, arthropods, annelids, echinoderms and ascidians, as a molecule that controls settlement and metamorphosis <!---# I think metamorhosis is less relevant for the paper --> initiation.
 We therefore hypothesise that NO-mediated neural signalling may be involved in the photoreceptor circuits that control UV avoidance behaviour.
@@
 Furthermore, we previously reported abnormal UV avoidance behaviour in c-opsin1 knockout larvae [@Veraszto2018].
+The cPRCs express the UV-absorbing c-opsin1 and mediate UV avoidance in the larvae [@Veraszto2018].
+In adult worms, UV light and c-opsin1 modulate locomotion and maturation [@Veedin2021].
 These suggest that Platynereis larvae are a highly suitable model for elucidating the neural circuit mechanisms of NO-induced UV avoidance behaviour, and the accessibility of genome editing, behavioural assays, calcium imaging and whole-body connectomics could be very useful for study at the molecular level[@ozpolat2021].
@@
 We identified a single *nitric oxide synthase* (*NOS*) gene in the *Platynereis dumerilii* genome and transcriptome data.
-Phylogenetic analysis of NOS proteins indicate that the *Platynereis* NOS belongs to an orthology group of metazoan NOS sequences (#Kei - need to add tree).
+Phylogenetic analysis of NOS proteins indicate that the *Platynereis* NOS belongs to an orthology group of metazoan NOS sequences <!---# need to add a NOS tree as supplement -->.
 To characterise the expression pattern of *NOS* we used in situ hybridization chain reaction (HCR) and transient transgenesis.
-In two-day-old larvae, we detected *NOS* expression in four cells in the apical organ region (Fig.
-1A) (#Kei - 3d HCR and transgene?).
-The same four cells cells were labelled with a reporter construct driving palmitoylated-tdTomato by a 12 kbp upstream genomic region of the *NOS* gene.
+In two-day-old larvae, we detected *NOS* expression in two cells in the apical organ region (Fig.
+1A) <!---# 3d HCR and transgene?
+-->.
+The same two cells cells were labelled with a reporter construct driving palmitoylated-tdTomato by a 12 kbp upstream genomic region of the *NOS* gene.
 This reporter also revealed the axonal projections of the NOS-expressing neurons.
+In three-day-old larvae...
+<!---# need 3d HCR and transgene data as well -->
+
 The position and morphology of the four *NOS*-expressing cells allowed us to identify the same cells in the three-day-old whole-body volume EM data [@williams2017; @veraszto2020] as four interneurons (INNOS)(Figure 1D-E).
@@
 
-**Nitric oxide produced during UV/violet stimulation gates the output of the cPRC circuit**
-
-The specific expression of NOS in interneurons directly postsynaptic to the cPRCs suggests that NO signalling is involved in the UV response.
-To test if NO is produced during UV/violet responses, we injected zygotes with the fluorescent NO-reporter DAF-FM and imaged two-day-old larvae during the local stimulation of the cPRC cilia by 405 nm light.
-Following light stimulation, we detected an increase in DAF-FM fluorescence in the region of the INNOS cells.
-(#Kei - are there time course data that could be shown?).
-We next analysed the dynamics of the cPRC circuit following light stimulation.
-As we have shown previously [@Veraszto2018], a 30-sec illumination of the cPRC ciliary region lead to a decrease in cPRC calcium levels, likely due to the hyperpolarisation of the cPRCs.
-After 30 min, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus - a response that may involve depolarisation.
-This activation phase ocurres after the end of the stimulation and may be due to a slower neuroendocrine feedback [@Veraszto2018].
-We also quantified in more detail the responses of the INNOS and INRGW interneurons under a UV/violet stimulus to the cPRCs.
-To unambiguously identify the cells from which we recorded calcium signals, we developed an on-slide staining method to directly fix, immunostain and image the same larva that we imaged previously.
-We used the cell-specific neuropeptide markers anti-RYamide (INNOS) and anti-RGWa (INRGW) to mark the interneurons.
-Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision.
-INNOS cells showed increase calcium signals during the inhibition phase of the cPRCs.
-In contrast, INRGW cells were only active during the second activation phase of cPRCs.\
-To determine whether NO is involved in the observed circuit dynamics, we generated two *Platynereis* *NOS* knockout lines with the CRISPR/Cas9 method.
+**Nitric oxide signalling mediates UV-avoidance behaviour** The specific expression of NOS in interneurons directly postsynaptic to the cPRCs suggests that NO signalling is involved in the UV response.
+To test this, we generated two *Platynereis* *NOS* knockout lines with the CRISPR/Cas9 method.
 We recovered two deletions (NOSΔ11/Δ11 and Δ23/Δ23), both leading to an early stop codon and thus likely representing null alleles (Figure 3A).
 We could establish a homozygous line for both mutations indicating that NOS is not an essential gene.
-Calcium imaging in *NOS*-mutant larvae revealed perturbed cPRC responses and circuit activation.
-In cPRCs, we could detect the inhibition phase but this was not followed by activation and calcium levels remained low.
-We did not detect a difference in the response of the INNOS cells between wild-type and *NOS*-mutant larvae.
-In contrast, INRGW cells did not activate in *NOS*-mutant larvae, suggesting that they will not initiate a circuit output (to Ser-h1 cells and the ciliary band).
-
-**Nitric oxide signalling mediates UV-avoidance behaviour** Our calcium imaging data suggest that *NOS*-mutant larvae will have defective UV-avoidance behaviour.
-To test this, we recorded the trajectories of freely swimming larvae in vertical columns, illuminated from two lateral sides with xy nm violet light.
+To quantify UV avoidance, we recorded the trajectories of freely swimming larvae in vertical columns, illuminated laterally from two opposite sides with xy nm violet light.
 As previously shown, wild-type larvae swim downward within 30 sec following UV/violet light stimulation.
 In contrast, homozygous *NOS*-mutant larvae showed a strongly diminished UV-avoidance response (Figure 3B).
-This phenotype is similar to the defective UV-avoidance of *c-opsin1* mutant larvae.
+This phenotype is similar to the defective UV-avoidance of *c-opsin1* mutant larvae and reveals a requirement for NOS in UV-avoidance behaviour.
 
@@
 (C and D) Vertical swimming in wild type and mutant larvae at 2-day-old (C) and 3-day-old (D) stimulated with blue (480 nm) and UV (395 nm) light from above.
-the data are shown in 30 s bins.](../figures/Fig2.png){width="100%"}
+the data are shown in 30 s bins.](../figures/Fig2.png){alt="Figure 2.
+Larvae with inhibited NOS have shorter ciliary arrest times and a delayed response to UV-avoidance.
+(A) Percentage of time in which cilia are closed in DMSO-treated larvae (control) vs.
+larvae exposed to NOS inhibitor.
+(B) Top: The domain structure of NOS.
+Bottom: Close-up region showing the genomic locus of NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9.
+The generated mutants (NOSΔ11, NOSΔ23) are also shown.
+Pink indicates target sites.
+Yellow shows PAM sequences, red shows stop codons.
+(C and D) Vertical swimming in wild type and mutant larvae at 2-day-old (C) and 3-day-old (D) stimulated with blue (480 nm) and UV (395 nm) light from above.
+the data are shown in 30 s bins." width="100%"}
+
+**Nitric oxide produced during UV/violet stimulation gates the output of the cPRC circuit**
+
+To test if NO is produced during UV/violet responses, we injected zygotes with the fluorescent NO-reporter DAF-FM and imaged two-day-old larvae during the local stimulation of the cPRC cilia by 405 nm light.
+Following light stimulation, we detected an increase in DAF-FM fluorescence in the region of the INNOS cells.
+(#Kei - are there time course data that could be shown?).
+We next analysed the dynamics of the cPRC circuit following light stimulation.
+As we have shown previously [@Veraszto2018], a 30-sec illumination of the cPRC ciliary region lead to a decrease in cPRC calcium levels, likely due to the hyperpolarisation of the cPRCs.
+After 30 min, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus - a response that may involve depolarisation.
+This activation phase occurs after the end of the stimulation and may be due to a slower neuroendocrine feedback [@Veraszto2018].
+To determine whether NO mediates such a feedback, we measured calcium signals in *NOS*-mutant larvae.
+While we detected the inhibition phase in cPRC in *NOS*-mutants, this was not followed by activation and calcium levels remained low.
+These results revealed an essential requirement for NOS signalling in the second phase of the cPRC response.
+
+We next quantified the responses of the INNOS and INRGW interneurons following a UV/violet stimulus to the cPRCs.
+To unambiguously identify the cells from which we recorded calcium signals, we developed an on-slide immunostaining method.
+We used the cell-specific neuropeptide markers anti-RYamide (INNOS) and anti-RGWa (INRGW) [@Conzelmann2011] to mark the interneurons.
+Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision (Figure 2 - figure supplement ).
+In both wild type and *NOS*-mutant larvae, INNOS cells showed increase calcium signals during the inhibition phase of the cPRCs.
+In contrast, calcium signals increased in INRGW cells in wild-type larvae during the second activation phase of cPRCs, a response that was lacking in *NOS*-mutants.
+This suggests that there is no circuit output to the Ser-h1 cells and the ciliary band in *NOS*-mutants, consistent with the observed behavioural defects.
 
@@
 
-Generally, NO can act via cyclic guanosine monophosphate (cGMP) as a second messenger.
+NO generally acts via cyclic guanosine monophosphate (cGMP) as a second messenger.
 In this signalling pathway, NO binds to the heme group of soluble guanylate cyclase (sGC), a member of the guanylate cyclase family with a CYC domain (PF00211), to increase cGMP.

### 2022-08-06 text/Manuscript.Rmd
NIT1_deletion

@@
   in zooplankton
-author: "Kei Jokura, Nobuo Ueda, Martin Gühmann, Gáspár Jékely"
+author: "Kei Jokura, Nobuo Ueda, Martin Gühmann, Luis Alfonso Yañez-Guerra, Gáspár Jékely"
 date: '*Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4
@@
 output:
-  html_document:
-    theme: lumen
   pdf_document:
@@
     template: default
+  html_document:
+    theme: lumen
   word_document: default
@@
 NO produced by nitric oxide synthase (NOS) in amacrine cells, one of the retinal cells, activates soluble guanylate cyclase (sGC) expressed in cone bipolar cells, which produces cGMP and regulates downstream signalling.
-On the other hand, the role of NO in zooplankton has been reported in many animals, including sponges, cnidarians, molluscs, arthropods, annelids, echinoderms and ascidians, as a molecule that controls settlement and metamorphosis <!---# I think metamorhosis is less relevant for the paper --> initiation.
 We therefore hypothesise that NO-mediated neural signalling may be involved in the photoreceptor circuits that control UV avoidance behaviour.

### 2022-08-06 text/NOS_slide.Rmd
NIT1_deletion

---
title: "NOS paper"
subtitle: "Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates UV avoidance in zooplankton"
author: "Kei Jokura, Nobuo Ueda, Martin Gühmann, Luis Alfonso Yañez-Guerra, Gáspár Jékely"
institute: "*Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4
  4QD, UK*"
date: "2022/08/03 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

background-image: "pictures/INNOS_Catmaid.png"

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

???

Image credit: [Wikimedia Commons](https://commons.wikimedia.org/wiki/File:Sharingan_triple.svg)

---
class: center, middle

# xaringan

### /ʃaː.'riŋ.ɡan/

---
class: inverse, center, middle

# Get Started

---

# Hello World

Install the **xaringan** package from [Github](https://github.com/yihui/xaringan):

```{r eval=FALSE, tidy=FALSE}
remotes::install_github("yihui/xaringan")
```

--

You are recommended to use the [RStudio IDE](https://www.rstudio.com/products/rstudio/), but you do not have to.

- Create a new R Markdown document from the menu `File -> New File -> R Markdown -> From Template -> Ninja Presentation`;<sup>1</sup>

--

- Click the `Knit` button to compile it;

--

- or use the [RStudio Addin](https://rstudio.github.io/rstudioaddins/)<sup>2</sup> "Infinite Moon Reader" to live preview the slides (every time you update and save the Rmd document, the slides will be automatically reloaded in RStudio Viewer.

.footnote[
[1] 中文用户请看[这份教程](https://slides.yihui.org/xaringan/zh-CN.html)

[2] See [#2](https://github.com/yihui/xaringan/issues/2) if you do not see the template or addin in RStudio.
]

---
background-image: url(`r xaringan:::karl`)
background-position: 50% 50%
class: center, bottom, inverse

# You only live once!

---

# Hello Ninja

As a presentation ninja, you certainly should not be satisfied by the "Hello World" example. You need to understand more about two things:

1. The [remark.js](https://remarkjs.com) library;

1. The **xaringan** package;

Basically **xaringan** injected the chakra of R Markdown (minus Pandoc) into **remark.js**. The slides are rendered by remark.js in the web browser, and the Markdown source needed by remark.js is generated from R Markdown (**knitr**).

---

# remark.js

You can see an introduction of remark.js from [its homepage](https://remarkjs.com). You should read the [remark.js Wiki](https://github.com/gnab/remark/wiki) at least once to know how to

- create a new slide (Markdown syntax<sup>*</sup> and slide properties);

- format a slide (e.g. text alignment);

- configure the slideshow;

- and use the presentation (keyboard shortcuts).

It is important to be familiar with remark.js before you can understand the options in **xaringan**.

.footnote[[*] It is different with Pandoc's Markdown! It is limited but should be enough for presentation purposes. Come on... You do not need a slide for the Table of Contents! Well, the Markdown support in remark.js [may be improved](https://github.com/gnab/remark/issues/142) in the future.]

---
background-image: url(`r xaringan:::karl`)
background-size: cover
class: center, bottom, inverse

# I was so happy to have discovered remark.js!

---
class: inverse, middle, center

# Using xaringan

---

# xaringan

Provides an R Markdown output format `xaringan::moon_reader` as a wrapper for remark.js, and you can use it in the YAML metadata, e.g.

```yaml
---
title: "A Cool Presentation"
output:
  xaringan::moon_reader:
    yolo: true
    nature:
      autoplay: 30000
---
```

See the help page `?xaringan::moon_reader` for all possible options that you can use.

---

# remark.js vs xaringan

Some differences between using remark.js (left) and using **xaringan** (right):

.pull-left[
1. Start with a boilerplate HTML file;

1. Plain Markdown;

1. Write JavaScript to autoplay slides;

1. Manually configure MathJax;

1. Highlight code with `*`;

1. Edit Markdown source and refresh browser to see updated slides;
]

.pull-right[
1. Start with an R Markdown document;

1. R Markdown (can embed R/other code chunks);

1. Provide an option `autoplay`;

1. MathJax just works;<sup>*</sup>

1. Highlight code with `{{}}`;

1. The RStudio addin "Infinite Moon Reader" automatically refreshes slides on changes;
]

.footnote[[*] Not really. See next page.]

---

# Math Expressions

You can write LaTeX math expressions inside a pair of dollar signs, e.g. &#36;\alpha+\beta$ renders $\alpha+\beta$. You can use the display style with double dollar signs:

```
$$\bar{X}=\frac{1}{n}\sum_{i=1}^nX_i$$
```

$$\bar{X}=\frac{1}{n}\sum_{i=1}^nX_i$$

Limitations:

1. The source code of a LaTeX math expression must be in one line, unless it is inside a pair of double dollar signs, in which case the starting `$$` must appear in the very beginning of a line, followed immediately by a non-space character, and the ending `$$` must be at the end of a line, led by a non-space character;

1. There should not be spaces after the opening `$` or before the closing `$`.

1. Math does not work on the title slide (see [#61](https://github.com/yihui/xaringan/issues/61) for a workaround).

---

# R Code

```{r comment='#'}
# a boring regression
fit = lm(dist ~ 1 + speed, data = cars)
coef(summary(fit))
dojutsu = c('地爆天星', '天照', '加具土命', '神威', '須佐能乎', '無限月読')
grep('天', dojutsu, value = TRUE)
```

---

# R Plots

```{r cars, fig.height=4, dev='svg'}
par(mar = c(4, 4, 1, .1))
plot(cars, pch = 19, col = 'darkgray', las = 1)
abline(fit, lwd = 2)
```

---

# Tables

If you want to generate a table, make sure it is in the HTML format (instead of Markdown or other formats), e.g.,

```{r}
knitr::kable(head(iris), format = 'html')
```

---

# HTML Widgets

I have not thoroughly tested HTML widgets against **xaringan**. Some may work well, and some may not. It is a little tricky.

Similarly, the Shiny mode (`runtime: shiny`) does not work. I might get these issues fixed in the future, but these are not of high priority to me. I never turn my presentation into a Shiny app. When I need to demonstrate more complicated examples, I just launch them separately. It is convenient to share slides with other people when they are plain HTML/JS applications.

See the next page for two HTML widgets.

---

```{r out.width='100%', fig.height=6, eval=require('leaflet')}
library(leaflet)
leaflet() %>% addTiles() %>% setView(-93.65, 42.0285, zoom = 17)
```

---

```{r eval=require('DT'), tidy=FALSE}
DT::datatable(
  head(iris, 10),
  fillContainer = FALSE, options = list(pageLength = 8)
)
```

---

# Some Tips

- Do not forget to try the `yolo` option of `xaringan::moon_reader`.

    ```yaml
    output:
      xaringan::moon_reader:
        yolo: true
    ```

---

# Some Tips

- Slides can be automatically played if you set the `autoplay` option under `nature`, e.g. go to the next slide every 30 seconds in a lightning talk:

    ```yaml
    output:
      xaringan::moon_reader:
        nature:
          autoplay: 30000
    ```

- If you want to restart the play after it reaches the last slide, you may set the sub-option `loop` to TRUE, e.g.,

    ```yaml
    output:
      xaringan::moon_reader:
        nature:
          autoplay:
            interval: 30000
            loop: true
    ```

---

# Some Tips

- A countdown timer can be added to every page of the slides using the `countdown` option under `nature`, e.g. if you want to spend one minute on every page when you give the talk, you can set:

    ```yaml
    output:
      xaringan::moon_reader:
        nature:
          countdown: 60000
    ```

    Then you will see a timer counting down from `01:00`, to `00:59`, `00:58`, ... When the time is out, the timer will continue but the time turns red.

---

# Some Tips

- The title slide is created automatically by **xaringan**, but it is just another remark.js slide added before your other slides.

    The title slide is set to `class: center, middle, inverse, title-slide` by default. You can change the classes applied to the title slide with the `titleSlideClass` option of `nature` (`title-slide` is always applied).

    ```yaml
    output:
      xaringan::moon_reader:
        nature:
          titleSlideClass: [top, left, inverse]
    ```

--

- If you'd like to create your own title slide, disable **xaringan**'s title slide with the `seal = FALSE` option of `moon_reader`.

    ```yaml
    output:
      xaringan::moon_reader:
        seal: false
    ```

---

# Some Tips

- There are several ways to build incremental slides. See [this presentation](https://slides.yihui.org/xaringan/incremental.html) for examples.

- The option `highlightLines: true` of `nature` will highlight code lines that start with `*`, or are wrapped in `{{ }}`, or have trailing comments `#<<`;

    ```yaml
    output:
      xaringan::moon_reader:
        nature:
          highlightLines: true
    ```

    See examples on the next page.

---

# Some Tips

.pull-left[
An example using a leading `*`:

    ```r
    if (TRUE) {
    ** message("Very important!")
    }
    ```
Output:
```r
if (TRUE) {
* message("Very important!")
}
```

This is invalid R code, so it is a plain fenced code block that is not executed.
]

.pull-right[
An example using `{{}}`:

````
`r ''````{r tidy=FALSE}
if (TRUE) {
*{{ message("Very important!") }}
}
```
````
Output:
```{r tidy=FALSE}
if (TRUE) {
{{ message("Very important!") }}
}
```

It is valid R code so you can run it. Note that `{{}}` can wrap an R expression of multiple lines.
]

---

# Some Tips

An example of using the trailing comment `#<<` to highlight lines:

````markdown
`r ''````{r tidy=FALSE}
library(ggplot2)
ggplot(mtcars) + 
  aes(mpg, disp) + 
  geom_point() +   #<<
  geom_smooth()    #<<
```
````

Output:

```{r tidy=FALSE, eval=FALSE}
library(ggplot2)
ggplot(mtcars) + 
  aes(mpg, disp) + 
  geom_point() +   #<<
  geom_smooth()    #<<
```

---

# Some Tips

When you enable line-highlighting, you can also use the chunk option `highlight.output` to highlight specific lines of the text output from a code chunk. For example, `highlight.output = TRUE` means highlighting all lines, and `highlight.output = c(1, 3)` means highlighting the first and third line.

````md
`r ''````{r, highlight.output=c(1, 3)}
head(iris)
```
````

```{r, highlight.output=c(1, 3), echo=FALSE}
head(iris)
```

Question: what does `highlight.output = c(TRUE, FALSE)` mean? (Hint: think about R's recycling of vectors)

---

# Some Tips

- To make slides work offline, you need to download a copy of remark.js in advance, because **xaringan** uses the online version by default (see the help page `?xaringan::moon_reader`).

- You can use `xaringan::summon_remark()` to download the latest or a specified version of remark.js. By default, it is downloaded to `libs/remark-latest.min.js`.

- Then change the `chakra` option in YAML to point to this file, e.g.

    ```yaml
    output:
      xaringan::moon_reader:
        chakra: libs/remark-latest.min.js
    ```

- If you used Google fonts in slides (the default theme uses _Yanone Kaffeesatz_, _Droid Serif_, and _Source Code Pro_), they won't work offline unless you download or install them locally. The Heroku app [google-webfonts-helper](https://google-webfonts-helper.herokuapp.com/fonts) can help you download fonts and generate the necessary CSS.

---

# Macros

- remark.js [allows users to define custom macros](https://github.com/yihui/xaringan/issues/80) (JS functions) that can be applied to Markdown text using the syntax `![:macroName arg1, arg2, ...]` or `![:macroName arg1, arg2, ...](this)`. For example, before remark.js initializes the slides, you can define a macro named `scale`:

    ```js
    remark.macros.scale = function (percentage) {
      var url = this;
      return '<img src="' + url + '" style="width: ' + percentage + '" />';
    };
    ```

    Then the Markdown text

    ```markdown
    ![:scale 50%](image.jpg)
    ```

    will be translated to

    ```html
    <img src="image.jpg" style="width: 50%" />
    ```

---

# Macros (continued)

- To insert macros in **xaringan** slides, you can use the option `beforeInit` under the option `nature`, e.g.,

    ```yaml
    output:
      xaringan::moon_reader:
        nature:
          beforeInit: "macros.js"
    ```

    You save your remark.js macros in the file `macros.js`.

- The `beforeInit` option can be used to insert arbitrary JS code before `remark.create()`. Inserting macros is just one of its possible applications.

---

# CSS

Among all options in `xaringan::moon_reader`, the most challenging but perhaps also the most rewarding one is `css`, because it allows you to customize the appearance of your slides using any CSS rules or hacks you know.

You can see the default CSS file [here](https://github.com/yihui/xaringan/blob/master/inst/rmarkdown/templates/xaringan/resources/default.css). You can completely replace it with your own CSS files, or define new rules to override the default. See the help page `?xaringan::moon_reader` for more information.

---

# CSS

For example, suppose you want to change the font for code from the default "Source Code Pro" to "Ubuntu Mono". You can create a CSS file named, say, `ubuntu-mono.css`:

```css
@import url(https://fonts.googleapis.com/css?family=Ubuntu+Mono:400,700,400italic);

.remark-code, .remark-inline-code { font-family: 'Ubuntu Mono'; }
```

Then set the `css` option in the YAML metadata:

```yaml
output:
  xaringan::moon_reader:
    css: ["default", "ubuntu-mono.css"]
```

Here I assume `ubuntu-mono.css` is under the same directory as your Rmd.

See [yihui/xaringan#83](https://github.com/yihui/xaringan/issues/83) for an example of using the [Fira Code](https://github.com/tonsky/FiraCode) font, which supports ligatures in program code.

---

# CSS (with Sass)

**xaringan** also supports Sass support via **rmarkdown**. Suppose you want to use the same color for different elements, e.g., first heading and bold text. You can create a `.scss` file, say `mytheme.scss`, using the [sass](https://sass-lang.com/) syntax with variables:

```scss
$mycolor: #ff0000; 
.remark-slide-content > h1 { color: $mycolor; }
.remark-slide-content strong { color: $mycolor; }
```

Then set the `css` option in the YAML metadata using this file placed under the same directory as your Rmd:

```yaml
output:
  xaringan::moon_reader:
    css: ["default", "mytheme.scss"]
```

This requires **rmarkdown** >= 2.8 and the [**sass**](https://rstudio.github.io/sass/) package. You can learn more about **rmarkdown** and **sass** support in [this blog post](https://blog.rstudio.com/2021/04/15/2021-spring-rmd-news/#sass-and-scss-support-for-html-based-output) and in [**sass** overview vignette](https://rstudio.github.io/sass/articles/sass.html).

---

# Themes

Don't want to learn CSS? Okay, you can use some user-contributed themes. A theme typically consists of two CSS files `foo.css` and `foo-fonts.css`, where `foo` is the theme name. Below are some existing themes:

```{r, R.options=list(width = 70)}
names(xaringan:::list_css())
```

---

# Themes

To use a theme, you can specify the `css` option as an array of CSS filenames (without the `.css` extensions), e.g.,

```yaml
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
```

If you want to contribute a theme to **xaringan**, please read [this blog post](https://yihui.org/en/2017/10/xaringan-themes).

---
class: inverse, middle, center
background-image: url(https://upload.wikimedia.org/wikipedia/commons/3/39/Naruto_Shiki_Fujin.svg)
background-size: contain

# Naruto

---
background-image: url(https://upload.wikimedia.org/wikipedia/commons/b/be/Sharingan_triple.svg)
background-size: 100px
background-position: 90% 8%

# Sharingan

The R package name **xaringan** was derived<sup>1</sup> from **Sharingan**, a dōjutsu in the Japanese anime _Naruto_ with two abilities:

- the "Eye of Insight"

- the "Eye of Hypnotism"

I think a presentation is basically a way to communicate insights to the audience, and a great presentation may even "hypnotize" the audience.<sup>2,3</sup>

.footnote[
[1] In Chinese, the pronounciation of _X_ is _Sh_ /ʃ/ (as in _shrimp_). Now you should have a better idea of how to pronounce my last name _Xie_.

[2] By comparison, bad presentations only put the audience to sleep.

[3] Personally I find that setting background images for slides is a killer feature of remark.js. It is an effective way to bring visual impact into your presentations.
]

---

# Naruto terminology

The **xaringan** package borrowed a few terms from Naruto, such as

- [Sharingan](https://naruto.fandom.com/wiki/Sharingan) (写輪眼; the package name)

- The [moon reader](https://naruto.fandom.com/wiki/Moon_Reader) (月読; an attractive R Markdown output format)

- [Chakra](https://naruto.fandom.com/wiki/Chakra) (查克拉; the path to the remark.js library, which is the power to drive the presentation)

- [Nature transformation](https://naruto.fandom.com/wiki/Nature_Transformation) (性質変化; transform the chakra by setting different options)

- The [infinite moon reader](https://naruto.fandom.com/wiki/Infinite_Tsukuyomi) (無限月読; start a local web server to continuously serve your slides)

- The [summoning technique](https://naruto.fandom.com/wiki/Summoning_Technique) (download remark.js from the web)

You can click the links to know more about them if you want. The jutsu "Moon Reader" may seem a little evil, but that does not mean your slides are evil.

---

class: center

# Hand seals (印)

Press `h` or `?` to see the possible ninjutsu you can use in remark.js.

![](https://upload.wikimedia.org/wikipedia/commons/7/7e/Mudra-Naruto-KageBunshin.svg)

---

class: center, middle

# Thanks!

Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan).

The chakra comes from [remark.js](https://remarkjs.com), [**knitr**](https://yihui.org/knitr/), and [R Markdown](https://rmarkdown.rstudio.com).

### 2022-08-06 code/220519_NITGC1_cGMP_analysis.R

@@ -315,8 +315,66 @@ GcG_NIT1_DMSO_PNG <- ggplot(GcG_NIT1_DMSO_mean) +
 
 
 #
+mutNIT1_SNAP <- read_csv("D:/NOS project/20220802_cGMP assay/R/normal_vs_deletion_mutNIT1_SNAP.csv")
   
+mutNIT1_SNAP_df <- mutNIT1_SNAP %>% 
+    pivot_longer(cols = c("cell1", "cell2", "cell3", "cell4", "cell5", "cell6", "cell7", "cell8", 
+                          "cell9", "cell10", "cell11", "cell12", "cell13", "cell14", "cell15", "cell16", "cell17", "cell18"), 
+                 names_to = "cell", 
+                 values_to = "intensity")
+  
+mutNIT1_SNAP_mean <- data.frame(mean = tapply(mutNIT1_SNAP_df$intensity, mutNIT1_SNAP_df$time, mean))
+  
+  
+#tapply
+mutNIT1_SNAP_mean <- data.frame(time = tapply(mutNIT1_SNAP_df$time, mutNIT1_SNAP_df$time, mean),
+                                  number = tapply(mutNIT1_SNAP_df$intensity, mutNIT1_SNAP_df$time, length),
+                                  mean = tapply(mutNIT1_SNAP_df$intensity, mutNIT1_SNAP_df$time, mean))
+  
+#sd
+mutNIT1_SNAP_mean$sd <- tapply(mutNIT1_SNAP_df$intensity, mutNIT1_SNAP_df$time, sd)
+#se
+mutNIT1_SNAP_mean$se <- mutNIT1_SNAP_mean$sd/sqrt(mutNIT1_SNAP_mean$n-1)
+  
+  
+#95% confidence interval
+mutNIT1_SNAP_mean$CI_lower <- 
+    mutNIT1_SNAP_mean$mean + qt((1-0.95)/2, df=mutNIT1_SNAP_mean$n-1) * mutNIT1_SNAP_mean$se
+mutNIT1_SNAP_mean$CI_upper <- 
+    mutNIT1_SNAP_mean$mean - qt((1-0.95)/2, df=mutNIT1_SNAP_mean$n-1) * mutNIT1_SNAP_mean$se  
  
+
+#graph
+mutNIT1_SNAP_PNG <- ggplot(mutNIT1_SNAP_mean) +
+  aes(x = time, y = mean) +
[…]

### 2022-08-06 code/220725_closure_time_fromNobuoData.R

@@ -1,0 +1,54 @@
+
+#Clear R's Brain
+rm(list = ls())
+
+library(dplyr)
+library(ggplot2)
+library(tidyverse)
+library(scales)
+library(plotly)
+library(ggplotgui)
+library(tidyr)
+library(esquisse)
+library(ggraptR)
+library(readr)
+library(multcomp)
+
+#closure time for boxplot
+closure_analysis <- read_csv("data/closure_time_inhibitor_donor.csv")
+
+#Check all data  
+ggplot(closure_analysis) +
+  aes(x=condition, y = second, fill = condition) +
+  geom_boxplot(shape = "circle") +
+  geom_jitter(size = 3, alpha = 0.25, width = 0, colour = 'black') +
+  scale_fill_manual(values = c(`Control` = "gray80",
+                               `SNAP0.1` = "steelblue3",
+                               `SNAP0.01` = "steelblue2",
+                               `L-NAME1` = "lightcoral")
+  ) +
+  scale_y_continuous(breaks=seq(0,60,length=4),limits = c(0, 60))+
+  labs(x = "Conditions", y = "Ciliary closure time (sec)") +
+  scale_x_discrete(labels = c("L-NAME1"="NOS inhibitor","SNAP0.01"="10 µM \n NO donor","SNAP0.1"="100 µM \n NO donor"))+
+  theme_classic() +
+  theme(
+    legend.position = "none",
+    axis.title.y = element_text(size = 14L,face = "bold"),
+    axis.title.x = element_text(size = 14L,face = "bold"),
+    axis.text.y = element_text(size = 14),
+    axis.text.x = element_text(size = 14)
+    )
+
+
+#statistical test
+SG_data <- lm(second ~ condition, data = closure_analysis) 
+SG_res1<-aov(second ~ condition, d = closure_analysis)
+summary(SG_res1)
+
+conditionx=factor(closure_analysis$condition)
+secondx=closure_analysis$second
+summary(glht(aov(secondx~conditionx),linfct=mcp(conditionx="Dunnett")))
+
+
+
+

### 2022-08-06 code/Fig2.R

@@ -13,7 +13,7 @@ library(magick)
 
 #read png convert to image panel
 
-panelA <- ggdraw() + draw_image(readPNG("pictures/closure period.png"))
+panelA <- ggdraw() + draw_image(readPNG("pictures/2dpf_ciliary closure_phamacological.png"))
 panelB <- ggdraw() + draw_image(readPNG("pictures/NOS-Architecture.png"))
 panelC <- ggdraw() + draw_image(readPNG("pictures/2dpf_wt_vs_NOS_mutants_for_discussion_1.png"))
 panelD <- ggdraw() + draw_image(readPNG("pictures/3dpf_wt_vs_NOS_mutants_for_discussion.png"))

### 2022-08-06 code/Fig6.R

@@ -112,6 +112,36 @@ NITGC1_analysis %>%
 ggsave("pictures/GcG-NIT-GC1-DMSO.png", limitsize = FALSE, 
        units = c("px"), width = 800, height = 600, bg='white')  
 
+
+
+# plot NITGC1_analysis----------------------------------------------------------
+
+NITGC1_analysis %>%
+  filter(expression == "GcG-mutNIT1") %>%
+  #cell names are not unique and the same across expression, fix by adding new column  
+  mutate(cell_expression = paste(cell, expression, sep = "")) %>%
+  ggplot(aes(x = time, y = intensity, color = expression)) +
+  geom_line(aes(group = cell_expression), size = 0.4, alpha = 0.2, 
+            show.legend = FALSE) +
+  geom_smooth(level = 0.99, size = 1, span = 0.2, 
+              method = "loess", show.legend = FALSE) +
+  theme_minimal()  +
+  theme(legend.title = element_blank(), legend.text.align=0,
+        legend.text = element_text(size=10)) +
+  #Specify colours and legend labels
+  scale_color_manual(values=c('grey50'),
+                     labels = c('GcG + mutNIT-GC1')) +
+  scale_y_continuous(breaks=seq(1,1.15,length=4),limits = c(0.95, 1.15))+
+  scale_x_continuous(breaks=seq(0,10,length=6),limits = c(0, 10))+
+  annotate("segment", x=2, xend=10, y=1.135, yend=1.135, size=2.5, color = "gray")+
+  annotate("text", x=2.5, y=1.15, size=4, label="SNAP")+
+  annotate("segment", x=0, xend=10, y=1, yend=1, size=0.5, linetype="dashed")
+
+# save plot ---------------------------------------------------------------
+
+ggsave("pictures/GcG-mutNIT-GC1-SNAP.png", limitsize = FALSE, 
+       units = c("px"), width = 800, height = 600, bg='white')  
+
 # plot NIT1 morphant data with geom_smooth ----------------------------------------------
 
 NIT1_MO_tb %>%
[…]

### 2022-08-06 code/cGMP_assay_deletion.R

@@ -1,0 +1,288 @@
+
+#Clear R's Brain
+rm(list = ls())
+
+library(dplyr)
+library(ggplot2)
+library(tidyverse)
+library(scales)
+library(plotly)
+library(ggplotgui)
+library(tidyr)
+library(esquisse)
+library(ggraptR)
+library(readr)
+
+mutNIT1_cGMP <- read_csv("D:/NOS project/20220802_cGMP assay/R/normal_vs_deletion_mutNIT1_cGMP.csv")
+
+mutNIT1_cGMP_df <- mutNIT1_cGMP %>% 
+  pivot_longer(cols = c("cell1", "cell2", "cell3", "cell4", "cell5", "cell6", "cell7", "cell8", 
+                        "cell9", "cell10"), 
+               names_to = "cell", 
+               values_to = "intensity")
+
+mutNIT1_cGMP_mean <- data.frame(mean = tapply(mutNIT1_cGMP_df$intensity, mutNIT1_cGMP_df$time, mean))
+
+
+#tapply
+mutNIT1_cGMP_mean <- data.frame(time = tapply(mutNIT1_cGMP_df$time, mutNIT1_cGMP_df$time, mean),
+                                number = tapply(mutNIT1_cGMP_df$intensity, mutNIT1_cGMP_df$time, length),
+                                mean = tapply(mutNIT1_cGMP_df$intensity, mutNIT1_cGMP_df$time, mean))
+
+#sd
+mutNIT1_cGMP_mean$sd <- tapply(mutNIT1_cGMP_df$intensity, mutNIT1_cGMP_df$time, sd)
+#se
+mutNIT1_cGMP_mean$se <- mutNIT1_cGMP_mean$sd/sqrt(mutNIT1_cGMP_mean$n-1)
+
+
+#95% confidence interval
+mutNIT1_cGMP_mean$CI_lower <- 
+  mutNIT1_cGMP_mean$mean + qt((1-0.95)/2, df=mutNIT1_cGMP_mean$n-1) * mutNIT1_cGMP_mean$se
+mutNIT1_cGMP_mean$CI_upper <- 
+  mutNIT1_cGMP_mean$mean - qt((1-0.95)/2, df=mutNIT1_cGMP_mean$n-1) * mutNIT1_cGMP_mean$se
+
+
+#graph
+ggplot(mutNIT1_cGMP_mean) +
+  aes(x = time, y = mean) +
[…]

### 2022-08-19 text/Manuscript.Rmd
delete worng files

@@
 
-![**Figure 5.
+![**Figure 4.
 NO produced by UV stimulation to cPRCs induces a depolarising response in cPRC via NIT-GC1.** (A) An INNOS visualised by DAF-FM is shown (a white arrowhead).
@@
 INRGW (green dashed line) and Ser-h1 (black arrowheads) identified from the numbered nuclear positions are shown.
-(L) Comparison of changes over time in GCaMP6s intensities in cPRC (purple) and INRGW (green) when the cPRC are stimulated with UV in WT (left) and NOS mutant (NOSΔ23: right) larvae.](../figures/Fig5.png){width="100%"}
+(L) Comparison of changes over time in GCaMP6s intensities in cPRC (purple) and INRGW (green) when the cPRC are stimulated with UV in WT (left) and NOS mutant (NOSΔ23: right) larvae.](../figures/Fig4.png){width="100%"}
 
@@
 
-![**Figure 6.
+![**Figure 5.
 Gene expressions and signalling mechanisms between cPRC, INNOS and INRGWa.** (A) Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq.
@@
 The colours represent the normal logarithm of the number of transcripts in the cells expressing the gene.
-(B) Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback.](../figures/Figure6.png){width="100%"}
+(B) Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback.](../figures/Fig5.png){width="100%"}
 

### 2022-08-19 text/Manuscript.Rmd
Add suplements

@@
 ![**Figure 3.
-cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.** (A) Phylogenetic tree of guanylate cyclase by maximum likelihood (ML).
-Guanylate cyclase-coupled receptor and soluble guanylate cyclase as outgroups.
-Guanylate cyclases with NIT domains are found in most animal phyla except Porifera, Ctenophora, Urochordata and Chordata.
-Tree robustness was tested with 1000 replicates of ultrafast bootstrap.
-Species names are as follows: Ap (Aiptasia pallida), Bb (Branchiostoma belcheri), Cv (Crassostrea virnigica), Dm (Drosophila melanogaster), Dp (Daphnia pulex), Hh (Hoilungia hongkongensis), Lg (Lottia gigantea), Pc (Priapulus caudatus), Pd (Platynereis dumerilii), Sk (Saccoglossus kowalevskii), Sp (Strongylocentrotus purpuratus), Ta (Trichoplax adhaerens), Tc (Tribolium castaneum).
-(B-D) Expression analysis of the NIT-GC1 gene (magenta) using in situ HCR.
+cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.**  (A-D) Expression analysis of the NIT-GC1 gene (magenta) using in situ HCR.
 White arrowheads indicate cells expressing NIT-GC1.
@@
 (G) Co-localisation analysis using NIT-GC1 and NOS antibodies.
-(B-G) Anterior view of the 2-day-old larvae.](../figures/Fig3.png){width="100%"}
+(B-G) Anterior view of the 2-day-old larvae.NIT-GC1 is induced by NO to produce cGMP.** (A) Top: The domain structure of NIT-GC1.
+Grey squares indicate the transmembrane domain.
+Bottom shows a schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control.
+Green cGull shows cGMP-dependent green fluorescence.
+(B) The intensity changes in cGMP fluorescence over time for the three conditions are shown.
+Gray traces show the optical responses to individual intensities (n = 6).
+The individual intensities are normalized (ΔF/F0).
+The dark line shows the mean value and the band the standard error range of the mean.
+The respective solution was acted on 2 min after the start of imaging (grey bars).
+(C) The mean intensity (Fmean/F0) after acting on the solution.
+Multiple comparisons were made using the Tukey honestly significant difference test.](../figures/Fig3.png){width="100%"}
 
@@
 They also show that the NIT domain is required for the activation of NIT-GC1.
-
-![**Figure 4.
-NIT-GC1 is induced by NO to produce cGMP.** (A) Top: The domain structure of NIT-GC1.
-Grey squares indicate the transmembrane domain.
-Bottom shows a schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control.
-Green cGull shows cGMP-dependent green fluorescence.
-(B) The intensity changes in cGMP fluorescence over time for the three conditions are shown.
-Gray traces show the optical responses to individual intensities (n = 6).
-The individual intensities are normalized (ΔF/F0).
-The dark line shows the mean value and the band the standard error range of the mean.
-The respective solution was acted on 2 min after the start of imaging (grey bars).
-(C) The mean intensity (Fmean/F0) after acting on the solution.
-Multiple comparisons were made using the Tukey honestly significant difference test.](../figures/Fig6.png){width="100%"}
 

### 2022-08-19 text/Supplements.Rmd
Add suplements

![**Figure 1—figure supplement 1. (A) Phylogenetic tree of NOS by maximum likelihood (ML). Tree robustness was tested with 1000 replicates of ultrafast bootstrap. Species names are as follows: Ap (Aiptasia pallida), Bb (Branchiostoma belcheri), Cv (Crassostrea virnigica), Dm (Drosophila melanogaster), Dp (Daphnia pulex), Hh (Hoilungia hongkongensis), Lg (Lottia gigantea), Pc (Priapulus caudatus), Pd (Platynereis dumerilii), Sk (Saccoglossus kowalevskii), Sp (Strongylocentrotus purpuratus), Ta (Trichoplax adhaerens), Tc (Tribolium castaneum).](../figures/Fig1_sup1.png){width="100%"}
![**Figure 3—figure supplement 1. (A) Phylogenetic tree of guanylate cyclase by maximum likelihood (ML). Guanylate cyclase-coupled receptor and soluble guanylate cyclase as outgroups. Guanylate cyclases with NIT domains are found in most animal phyla except Porifera, Ctenophora, Urochordata and Chordata. Tree robustness was tested with 1000 replicates of ultrafast bootstrap. Species names are as follows: Ap (Aiptasia pallida), Bb (Branchiostoma belcheri), Cv (Crassostrea virnigica), Dm (Drosophila melanogaster), Dp (Daphnia pulex), Hh (Hoilungia hongkongensis), Lg (Lottia gigantea), Pc (Priapulus caudatus), Pd (Platynereis dumerilii), Sk (Saccoglossus kowalevskii), Sp (Strongylocentrotus purpuratus), Ta (Trichoplax adhaerens), Tc (Tribolium castaneum).](../figures/Fig3_sup1.png){width="100%"}

### 2022-08-19 code/Fig1_sup1.R

@@ -1,0 +1,27 @@
+#Code to assemble Fig1 of the Jokura et al Platynereis NOS paper
+#2022 June - Kei Jokura, Gaspar Jekely
+
+# source packages ---------------------------------------------------------
+source("code/Packages_Catmaid_connection.R")
+
+# assemble figure ---------------------------------------------------------
+panel_A <- ggdraw() + draw_image(readPNG("pictures/FINALSeq_NOS_cdhit080_one_Lopho_2nd.fasta.alngapp.fasta.treefile.png"))
+
+#combine panels into Figure and save final figure as pdf and png
+#panels of different sizes
+layout <- "
+A
+"
+
+Fig1_sup1 <- panel_A + 
+  patchwork::plot_layout(design = layout) + #we can change the heights of the rows in our layout (widths also can be defined)
+  patchwork::plot_annotation(tag_levels = "A") &  #we can change this to 'a' for small caps or 'i' or '1'
+  ggplot2::theme(plot.tag = element_text(size = 12, face='plain', color='black')) #or 'bold', 'italic'
+
+Fig1_sup1
+
+ggsave("figures/Fig1_sup1.png", limitsize = FALSE, 
+       units = c("px"), Fig1_sup1, width = 2350, height = 1700, bg='white')  
+
+ggsave("figures/Fig1_sup1.pdf", limitsize = FALSE, 
+       units = c("px"), Fig1_sup1, width = 2350, height = 1700)  

### 2022-08-19 code/Fig3_sup1.R

@@ -1,0 +1,26 @@
+#Code to assemble Fig1 of the Jokura et al Platynereis NOS paper
+#2022 June - Kei Jokura, Gaspar Jekely
+
+
+
+# assemble figure ---------------------------------------------------------
+panel_A <- ggdraw() + draw_image(readPNG("pictures/NITGC_allseq_mbGC_sGC_CD80_plus_Lopho.fasta.alngapp.fasta.treefile.png"))
+
+#combine panels into Figure and save final figure as pdf and png
+#panels of different sizes
+layout <- "
+A
+"
+
+Fig3_sup1 <- panel_A + 
+  patchwork::plot_layout(design = layout) + #we can change the heights of the rows in our layout (widths also can be defined)
+  patchwork::plot_annotation(tag_levels = "A") &  #we can change this to 'a' for small caps or 'i' or '1'
+  ggplot2::theme(plot.tag = element_text(size = 12, face='plain', color='black')) #or 'bold', 'italic'
+
+Fig3_sup1
+
+ggsave("figures/Fig3_sup1.png", limitsize = FALSE, 
+       units = c("px"), Fig3_sup1, width = 2350, height = 1700, bg='white')  
+
+ggsave("figures/Fig3_sup1.pdf", limitsize = FALSE, 
+       units = c("px"), Fig3_sup1, width = 2350, height = 1700)  

### 2022-08-21 text/Manuscript.Rmd
Upload Supplement Figures

@@
 This results in the activation of a second population of postsynaptic interneurons (INRGW).
+It was also suggested that activation of INRGWs activates serotonergic ciliomotor neurons (Ser-h1), which are postsynaptic and involved in increasing ciliary beat frequency.
 In *NOS* knockout larvae, NO feedback and INRGW activation do not occur and avoidance behaviour is defective.
@@
 We identified a single *nitric oxide synthase* (*NOS*) gene in the *Platynereis dumerilii* genome and transcriptome data.
-Phylogenetic analysis of NOS proteins indicate that the *Platynereis* NOS belongs to an orthology group of metazoan NOS sequences <!---# need to add a NOS tree as supplement -->.
+Phylogenetic analysis of NOS proteins indicate that the *Platynereis* NOS belongs to an orthology group of metazoan NOS sequences (Fig.
+1---figure supplement 1).
 To characterise the expression pattern of *NOS* we used in situ hybridization chain reaction (HCR) and transient transgenesis.
 In two-day-old larvae, we detected *NOS* expression in two cells in the apical organ region (Fig.
-1A) <!---# 3d HCR and transgene?
--->.
+1A) <!---# 3d HCR, Kei will put them -->.
 The same two cells cells were labelled with a reporter construct driving palmitoylated-tdTomato by a 12 kbp upstream genomic region of the *NOS* gene.
@@
 In three-day-old larvae...
-<!---# need 3d HCR and transgene data as well -->
+<!---# need 3d HCR , Kei will put them -->
 
@@
 ![**Figure 1.
-Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** (A and B) Immunostaining with anti-HA antibody against HA tagged palmitoylated-tdTomato expressed under the upstream of the start site of NOS (magenta).
+Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** (A) SEM image of a 72 hpf Platynereis larva.(B-D) Expression analysis of the NOS gene (magenta) using in situ HCR.
+Co-stained image using acetylated α-tubulin antibody (ac α-tub: green).
+Anterior view of the larva at 54 and 72 hpf.
+(E) Immunostaining with anti-HA antibody against HA tagged palmitoylated-tdTomato expressed under the upstream of the start site of NOS (magenta).
 Co-staining with acetylated α-tubulin antibody as marker for cilia and axonal scaffold (ac α-tub: green).
 INNOSs with the entire cell body and axon labelled are shown with white arrowheads and with only the cell body labelled are shown with black arrowheads.
-Views of larvae at 2-day-old (A) and 3-day-old (B) from the anterior side, respectively.
-(C) Three types of neurons (cPRC, INNOS and INRGWa) in the cPRC circuit reconstructed in the whole-body transmission electron microscopy (ssTEM) dataset of a 3-day-old larva.
-The INNOS on the ventral side labelled with (A) and (B) are shown in purple.
-Other neurons are shown in light grey.
+Views of larvae at 72 hpf from the anterior side.
+(F and G) Five types of neurons (cPRC, INNOS, INRGWa, Ser-h1 and MC) in the cPRC circuit reconstructed in the whole-body transmission electron microscopy (ssTEM) dataset of a 72 hpf larva.
+The INNOS on the ventral side labelled with (E) are shown in purple.
 Axons and dendrites appear as lines and cell body positions are represented by spheres.
-(D and E) Expression analysis of the NOS gene (magenta) using in situ HCR.
-Anterior view of the larva at 2-day-old.
-(D) Co-stained image using acetylated α-tubulin antibody (ac α-tub: green).
-(E) Co-expression analysis image with the RYamide proneuropeptide gene (RYa-pNP: green), which is highly expressed in INNOS.
-White arrowheads indicate co-expressed INNOS.
-Black arrowheads indicate larval eyes.
-White dashed lines indicate the outer frame of the larva.
-(F) Wiring diagram of the cPRC circuit.
-Hexagons represent neuronal groups (numbers indicated in square brackets), arrows represent synaptic connections (numbers indicated in number of synapses); NOS and RGW indicate INNOS and INRGW-dc1 respectively.
+(H) Wiring diagram of the cPRC circuit.
+<!---# This graph widget does not look very good.
+We need to edit this a bit more.
+--> Hexagons represent neuronal groups (numbers indicated in square brackets), arrows represent synaptic connections (numbers indicated in number of synapses); NOS and RGW indicate INNOS and INRGW-dc1 respectively.
 Ciliomotor neuron shows two serotonergic neurons (Ser-h1).](../figures/Fig1.png){width="100%"}
@@
 ![**Figure 2.
-Larvae with inhibited NOS have shorter ciliary arrest times and a delayed response to UV-avoidance.** (A) Percentage of time in which cilia are closed in DMSO-treated larvae (control) vs.
-larvae exposed to NOS inhibitor.
-(B) Top: The domain structure of NOS.
+Larvae with inhibited NOS have shorter ciliary arrest times and a delayed response to UV-avoidance.** (A) Top: The domain structure of NOS.
 Bottom: Close-up region showing the genomic locus of NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9.
@@
 Yellow shows PAM sequences, red shows stop codons.
-(C and D) Vertical swimming in wild type and mutant larvae at 2-day-old (C) and 3-day-old (D) stimulated with blue (480 nm) and UV (395 nm) light from above.
+(B and C) Vertical swimming in wild type and mutant larvae at 2-day-old (B) and 3-day-old (C) stimulated with blue (480 nm) and UV (395 nm) light from above.
 the data are shown in 30 s bins.](../figures/Fig2.png){alt="Figure 2.
@@
 the data are shown in 30 s bins." width="100%"}
-
-**Nitric oxide produced during UV/violet stimulation gates the output of the cPRC circuit**
-
-To test if NO is produced during UV/violet responses, we injected zygotes with the fluorescent NO-reporter DAF-FM and imaged two-day-old larvae during the local stimulation of the cPRC cilia by 405 nm light.
-Following light stimulation, we detected an increase in DAF-FM fluorescence in the region of the INNOS cells.
-(#Kei - are there time course data that could be shown?).
-We next analysed the dynamics of the cPRC circuit following light stimulation.
-As we have shown previously [@Veraszto2018], a 30-sec illumination of the cPRC ciliary region lead to a decrease in cPRC calcium levels, likely due to the hyperpolarisation of the cPRCs.
-After 30 min, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus - a response that may involve depolarisation.
-This activation phase occurs after the end of the stimulation and may be due to a slower neuroendocrine feedback [@Veraszto2018].
-To determine whether NO mediates such a feedback, we measured calcium signals in *NOS*-mutant larvae.
-While we detected the inhibition phase in cPRC in *NOS*-mutants, this was not followed by activation and calcium levels remained low.
-These results revealed an essential requirement for NOS signalling in the second phase of the cPRC response.
-
-We next quantified the responses of the INNOS and INRGW interneurons following a UV/violet stimulus to the cPRCs.
-To unambiguously identify the cells from which we recorded calcium signals, we developed an on-slide immunostaining method.
-We used the cell-specific neuropeptide markers anti-RYamide (INNOS) and anti-RGWa (INRGW) [@Conzelmann2011] to mark the interneurons.
-Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision (Figure 2 - figure supplement ).
-In both wild type and *NOS*-mutant larvae, INNOS cells showed increase calcium signals during the inhibition phase of the cPRCs.
-In contrast, calcium signals increased in INRGW cells in wild-type larvae during the second activation phase of cPRCs, a response that was lacking in *NOS*-mutants.
-This suggests that there is no circuit output to the Ser-h1 cells and the ciliary band in *NOS*-mutants, consistent with the observed behavioural defects.
 
@@
 
-![**Figure 3.
-cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.**  (A-D) Expression analysis of the NIT-GC1 gene (magenta) using in situ HCR.
-White arrowheads indicate cells expressing NIT-GC1.
-(B) Co-staining image using acetylated α-tubulin antibody (ac α-tub: green).
-(C) Co-expression analysis image of the c-opsin1 gene (green).
-Black arrowheads indicate cells expressing c-opsin1 differently from cPRC.
-(D) Co-expression analysis image of the MLD-pedal2 amide proneuropeptide gene (MLD-pNP: green).
-Black arrowheads indicate cells different from cPRC.
-The white dashed lines in (C) and (D) indicate the outer frame of the larvae.
-(E, F) Localisation analysis using NIT-GC1 and NIT-GC2 antibodies.
-Green shows co-staining with acetylated α-tubulin antibody (ac α-tub).
-(G) Co-localisation analysis using NIT-GC1 and NOS antibodies.
-(B-G) Anterior view of the 2-day-old larvae.NIT-GC1 is induced by NO to produce cGMP.** (A) Top: The domain structure of NIT-GC1.
-Grey squares indicate the transmembrane domain.
-Bottom shows a schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control.
-Green cGull shows cGMP-dependent green fluorescence.
-(B) The intensity changes in cGMP fluorescence over time for the three conditions are shown.
-Gray traces show the optical responses to individual intensities (n = 6).
-The individual intensities are normalized (ΔF/F0).
-The dark line shows the mean value and the band the standard error range of the mean.
-The respective solution was acted on 2 min after the start of imaging (grey bars).
-(C) The mean intensity (Fmean/F0) after acting on the solution.
-Multiple comparisons were made using the Tukey honestly significant difference test.](../figures/Fig3.png){width="100%"}
-
 **NIT-GC1 produces cGMP by nitric oxide**
@@
 They also show that the NIT domain is required for the activation of NIT-GC1.
+
+![**Figure 3.
+cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.** (A) Expression analysis of the NIT-GC1 gene (magenta) using in situ HCR.
+White arrowheads indicate cells expressing NIT-GC1.Co-staining image using acetylated α-tubulin antibody (ac α-tub: green).
+(B) Co-expression analysis image of the c-opsin1 gene (green).
+Black arrowheads indicate cells expressing c-opsin1 differently from cPRC.
+(C and D) Localisation analysis using NIT-GC1 and NIT-GC2 antibodies.
+Green shows co-staining with acetylated α-tubulin antibody (ac α-tub).
+(E and F) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in control (non-injection) and NIT-GC1 and NIT-GC2 knockdown by morpholino (MO1 and 2) larvae.
+(G-J) The intensity changes in cGMP fluorescence over time for the four conditions are shown.
+Gray traces show the optical responses to individual intensities (n > 6).
+The individual intensities are normalized (ΔF/F0).
+The dark line shows the mean value and the band the standard error range of the mean.
+The respective solution was acted on 2 min after the start of imaging (grey bars).](../figures/Fig3.png){width="100%"}
+
+**Nitric oxide produced during UV/violet stimulation gates the output of the cPRC circuit**
+
+To test if NO is produced during UV/violet responses, we injected zygotes with the fluorescent NO-reporter DAF-FM and imaged two-day-old larvae during the local stimulation of the cPRC cilia by 405 nm light.
+Following light stimulation, we detected an increase in DAF-FM fluorescence in the region of the INNOS cells.
+We next analysed the dynamics of the cPRC circuit following light stimulation.
+As we have shown previously [@Veraszto2018], a 30-sec illumination of the cPRC ciliary region lead to a decrease in cPRC calcium levels, likely due to the hyperpolarisation of the cPRCs.
+After 30 min, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus - a response that may involve depolarisation.
+This activation phase occurs after the end of the stimulation and may be due to a slower neuroendocrine feedback [@Veraszto2018].
+To determine whether NO mediates such a feedback, we measured calcium signals in *NOS*-mutant larvae.
+While we detected the inhibition phase in cPRC in *NOS*-mutants, this was not followed by activation and calcium levels remained low.
+These results revealed an essential requirement for NOS signalling in the second phase of the cPRC response.
+
+We next quantified the responses of the INNOS and INRGW interneurons following a UV/violet stimulus to the cPRCs.
+To unambiguously identify the cells from which we recorded calcium signals, we developed an on-slide immunostaining method.
+We used the cell-specific neuropeptide markers anti-RYamide (INNOS) and anti-RGWa (INRGW) [@Conzelmann2011] to mark the interneurons.
+Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision (Figure 4---figure supplement 1).
+In both wild type and *NOS*-mutant larvae, INNOS cells showed increase calcium signals during the inhibition phase of the cPRCs.
+In contrast, calcium signals increased in INRGW cells in wild-type larvae during the second activation phase of cPRCs, a response that was lacking in *NOS*-mutants.
+This suggests that there is no circuit output to the Ser-h1 cells and the ciliary band in *NOS*-mutants, consistent with the observed behavioural defects.
 
@@
 (H) Comparison of changes over time in GCaMP6s intensities in cPRC (purple) and INNOS (blue) when cPRC is stimulated with UV in WT (left) and NOS mutant (NOSΔ11: right) larvae.
-(I) Schematic diagram of the immunostaining procedure after calcium imaging.
-(J) The position of the cPRC (purple dashed line) in calcium imaging with GCaMP6s.
-Cells around the cPRC are numbered according to the position of the nuclei.
-(K) Immunostaining image using RGWamide (RGWa) and serotonin (Ser) antibodies (red) and DAPI (light blue) in the same region as (J).
-INRGW (green dashed line) and Ser-h1 (black arrowheads) identified from the numbered nuclear positions are shown.
 (L) Comparison of changes over time in GCaMP6s intensities in cPRC (purple) and INRGW (green) when the cPRC are stimulated with UV in WT (left) and NOS mutant (NOSΔ23: right) larvae.](../figures/Fig4.png){width="100%"}

### 2022-08-21 text/Supplements.Rmd
Upload Supplement Figures

@@
-![**Figure 1—figure supplement 1.
-(A) Phylogenetic tree of NOS by maximum likelihood (ML).
+---
+title: Figure supplements
+author: "Kei Jokura, Nobuo Ueda, Martin Gühmann, Luis Alfonso Yañez-Guerra, Gáspár Jékely"
+date: '*Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4
+  4QD, UK*'
+output:
+  pdf_document:
+    latex_engine: lualatex
+    template: default
+  html_document:
+    theme: lumen
+  word_document: default
+geometry: left=0.8in,right=0.8in,top=0.8in,bottom=0.5in
+mainfont: Arial
+monofont: Times New Roman
+documentclass: article
+classoption: onecolumn
+linestretch: 1.2
+fontsize: 10pt
+bibliography: references.bib
+csl: elife.csl
+---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+
+```
+
+![**Figure 1---figure supplement 1.** (A) Phylogenetic tree of NOS by maximum likelihood (ML).
 Tree robustness was tested with 1000 replicates of ultrafast bootstrap.
 Species names are as follows: Ap (Aiptasia pallida), Bb (Branchiostoma belcheri), Cv (Crassostrea virnigica), Dm (Drosophila melanogaster), Dp (Daphnia pulex), Hh (Hoilungia hongkongensis), Lg (Lottia gigantea), Pc (Priapulus caudatus), Pd (Platynereis dumerilii), Sk (Saccoglossus kowalevskii), Sp (Strongylocentrotus purpuratus), Ta (Trichoplax adhaerens), Tc (Tribolium castaneum).](../figures/Fig1_sup1.png){width="100%"}
-![**Figure 3—figure supplement 1.
-(A) Phylogenetic tree of guanylate cyclase by maximum likelihood (ML).
+
+![**Figure 1---figure supplement 2.** (A) Immunostaining with NOS antibodies.(B) Co-expression analysis image with the RYamide proneuropeptide gene (RYa-pNP: green), which is highly expressed in INNOS.
+White arrowheads indicate co-expressed INNOS.
+Black arrowheads indicate larval eyes.
+White dashed lines indicate the outer frame of the larva.](../figures/Fig1_sup2.png){width="100%"}
+
+![\*\*Figure 3---figure supplement 1.\*\* (A) Phylogenetic tree of guanylate cyclase by maximum likelihood (ML).
 Guanylate cyclase-coupled receptor and soluble guanylate cyclase as outgroups.
@@
 Species names are as follows: Ap (Aiptasia pallida), Bb (Branchiostoma belcheri), Cv (Crassostrea virnigica), Dm (Drosophila melanogaster), Dp (Daphnia pulex), Hh (Hoilungia hongkongensis), Lg (Lottia gigantea), Pc (Priapulus caudatus), Pd (Platynereis dumerilii), Sk (Saccoglossus kowalevskii), Sp (Strongylocentrotus purpuratus), Ta (Trichoplax adhaerens), Tc (Tribolium castaneum).](../figures/Fig3_sup1.png){width="100%"}
+
+![**Figure 3---figure supplement 2.** (A) Co-expression analysis image of the MLD-pedal2 amide proneuropeptide gene (MLD-pNP: green).
+Black arrowheads indicate cells different from cPRC.
+(B) Co-localisation analysis using NIT-GC1 and NOS antibodies.](../figures/Fig3_sup2.png){width="100%"}
+
+![**Figure 3---figure supplement 3.** (A) Comparison of amino acid sequences of regions of possible NO binding in NIT-GC1 and NIT-GC2.
+(B) 3D stuructures of NIT-GC1 and NIT-GC2 (C)](../figures/Fig3_sup3.png){width="100%"}
+
+![**Figure 3---figure supplement 4.** (A and B) Left: The domain structure of NIT-GC1.
+Grey squares indicate the transmembrane domain.
+Right shows a schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control.
+Green cGull shows cGMP-dependent green fluorescence.](../figures/Fig3_sup4.png){width="100%"}
+
+![**Figure 4---figure supplement 1.** (A) Schematic diagram of the immunostaining procedure after calcium imaging.
+(B) The position of the cPRC (purple dashed line) in calcium imaging with GCaMP6s.
+Cells around the cPRC are numbered according to the position of the nuclei.
+(C) Immunostaining image using RGWamide (RGWa) and serotonin (Ser) antibodies (red) and DAPI (light blue) in the same region as (B).
+INRGW (green dashed line) and Ser-h1 (black arrowheads) identified from the numbered nuclear positions are shown.
+.](../figures/Fig4_sup1.png){width="100%"}

### 2022-08-21 code/Fig1.R

@@ -218,65 +218,68 @@ panel_SEM <- ggdraw() +
   draw_image(SEM) +
   draw_line(x = c(0.1, 0.2785), y = c(0.07, 0.07), color = "black", size = 0.5) +
   draw_label(expression(paste("50 ", mu, "m")), x = 0.2, y = 0.1, size = 10) +
-  draw_label("dorsal view", x = 0.2, y = 0.99, size = 10)
+  draw_label("dosal view", x = 0.3, y = 0.99, size = 10)
 
 panel_NOS2d <- ggdraw() + draw_image(readPNG("pictures/NOS-promotor_2d_acTub_XXum.png"))
+
 panel_NOS3d <- ggdraw() + draw_image(readPNG("pictures/NOS-promotor_3d_acTub_XXum_crop.png")) +
-  draw_label("NOSp::palmy-3xH-TdTomato", x = 0.45, y = 0.99, size = 10) +
+  draw_label("NOSp::palmi-tdTomato", x = 0.45, y = 0.99, size = 10) +
   draw_label("INNOS", x = 0.55, y = 0.75, color='#CC79A7', size = 12, fontface='bold') +
   draw_label("cPRC", x = 0.3, y = 0.85, color='white',size = 12,fontface='bold') +
-  draw_label("*", x = 0.55, y = 0.32, color='white',size = 18,fontface='bold')
+  draw_label("*", x = 0.55, y = 0.32, color='white',size = 18,fontface='bold') +
+  draw_line(x = c(0.05, 0.407), y = c(0.1, 0.1), color = "white", size = 0.5) +
+  draw_label(expression(paste("20 ", mu, "m")), x = 0.23, y = 0.13, color = "white", size = 10)
+
 panel_INNOS_ant <- ggdraw() + draw_image(readPNG("pictures/INNOS_Catmaid.png")) +
-  draw_label("anterior view", x = 0.3, y = 0.99, size = 10)  +
+  draw_label("anterior view", x = 0.35, y = 0.99, size = 10)  +
   draw_label("*", x = 0.51, y = 0.205, color='black',size = 18,fontface='bold')
 panel_INNOS_ant
+
 panel_INNOS_ventr <- ggdraw() + draw_image(readPNG("pictures/INNOS_Catmaid_ventr.png")) +
-  draw_label("ventral view", x = 0.25, y = 0.99, size = 10)
+  draw_label("ventral view", x = 0.3, y = 0.99, size = 10)
 
[…]

### 2022-08-21 code/Fig1_sup2.R

@@ -5,23 +5,24 @@
 source("code/Packages_Catmaid_connection.R")
 
 # assemble figure ---------------------------------------------------------
-panel_A <- ggdraw() + draw_image(readPNG("pictures/FINALSeq_NOS_cdhit080_one_Lopho_2nd.fasta.alngapp.fasta.treefile.png"))
+panel_A <- ggdraw() + draw_image(readPNG("pictures/IHC_54_AP_NOS_actub_71.46um.png"))
+panel_B <- ggdraw() + draw_image(readPNG("pictures/HCR_52_AP_NOS_RYa_101.29um.png"))
 
 #combine panels into Figure and save final figure as pdf and png
 #panels of different sizes
 layout <- "
-A
+AB
 "
 
-Fig1_sup1 <- panel_A + 
+Fig1_sup2 <- panel_A + panel_B +
   patchwork::plot_layout(design = layout) + #we can change the heights of the rows in our layout (widths also can be defined)
   patchwork::plot_annotation(tag_levels = "A") +  #we can change this to 'a' for small caps or 'i' or '1'
   ggplot2::theme(plot.tag = element_text(size = 12, face='plain', color='black')) #or 'bold', 'italic'
 
-Fig1_sup1
+Fig1_sup2
 
-ggsave("figures/Fig1_sup1.png", limitsize = FALSE, 
-       units = c("px"), Fig1_sup1, width = 2350, height = 1700, bg='white')  
+ggsave("figures/Fig1_sup2.png", limitsize = FALSE, 
+       units = c("px"), Fig1_sup2, width = 2350, height = 1700, bg='white')  
 
-ggsave("figures/Fig1_sup1.pdf", limitsize = FALSE, 
-       units = c("px"), Fig1_sup1, width = 2350, height = 1700)  
+ggsave("figures/Fig1_sup2.pdf", limitsize = FALSE, 
+       units = c("px"), Fig1_sup2, width = 2350, height = 1700)  

### 2022-08-21 code/Fig2.R

@@ -1,36 +1,34 @@
-rm(list = ls(all.names = TRUE)) #will clear all objects includes hidden objects.
-gc() #free up memory and report the memory usage.
+#Code to assemble Fig2 of the Jokura et al Platynereis NOS paper
+#2022 June - Kei Jokura, Gaspar Jekely
 
-# load some packages
-library(tidyverse)
-library(cowplot)
-library(png)
-library(patchwork)
-library(magick)
-
+# source packages ---------------------------------------------------------
+source("code/Packages_to_load.R")
 
 # assemble figure ---------------------------------------------------------
 
 #read png convert to image panel
 
-panelA <- ggdraw() + draw_image(readPNG("pictures/2dpf_ciliary closure_phamacological.png"))
-panelB <- ggdraw() + draw_image(readPNG("pictures/NOS-Architecture.png"))
-panelC <- ggdraw() + draw_image(readPNG("pictures/2dpf_wt_vs_NOS_mutants_for_discussion_1.png"))
-panelD <- ggdraw() + draw_image(readPNG("pictures/3dpf_wt_vs_NOS_mutants_for_discussion.png"))
-
+panel_Architecture <- ggdraw() + draw_image(readPNG("pictures/NOS-Architecture.png"))
+panel_2dpf <- ggdraw() + draw_image(readPNG("pictures/2dpf_wt_vs_NOS_mutants_for_discussion_1.png"))
+panel_3dpf <- ggdraw() + draw_image(readPNG("pictures/3dpf_wt_vs_NOS_mutants_for_discussion.png"))
 
 #combine panels into Figure and save final figure as pdf and png
-Fig2 <- plot_grid(panelA,panelB,panelC,panelD,
-                  ncol=2,
-                  rel_widths = c(1, 1, 1, 1),
-                  labels=c("A","B","C","D"),
-                  label_size = 18, label_y = 1, label_x = 0,
-                  label_fontfamily = "sans", label_fontface = "plain") + 
-  theme(plot.margin = unit(c(1, 1, 1, 1), units = "pt"))
+#panels of different sizes
+layout <- "
+ABC
+"
 
-Fig2
+Fig2 <- panel_Architecture + panel_2dpf + panel_3dpf +
+  patchwork::plot_layout(design = layout) + #we can change the heights of the rows in our layout (widths also can be defined)
[…]

### 2022-08-21 code/Fig3_sup2.R

@@ -1,4 +1,4 @@
-#Code to assemble Fig3 sup1 of the Jokura et al Platynereis NOS paper
+#Code to assemble Fig3 sup2 of the Jokura et al Platynereis NOS paper
 #2022 June - Kei Jokura, Gaspar Jekely
 
 rm(list = ls(all.names = TRUE)) #will clear all objects includes hidden objects.
@@ -12,23 +12,24 @@ library(patchwork)
 library(magick)
 
 # assemble figure ---------------------------------------------------------
-panel_A <- ggdraw() + draw_image(readPNG("pictures/NITGC_allseq_mbGC_sGC_CD80_plus_Lopho.fasta.alngapp.fasta.treefile.png"))
+panel_MLD <- ggdraw() + draw_image(readPNG("pictures/HCR_51_AP_NITGC1_MLD_112.55um.png"))
+panel_NIT1andNOS <- ggdraw() + draw_image(readPNG("pictures/IHC_55_NIT1_NOS-RAT1_acTub_AP_2_Airyscan Processing-Orthogonal Projection-11_c1-3_71.46um.png"))
 
 #combine panels into Figure and save final figure as pdf and png
 #panels of different sizes
 layout <- "
-A
+AB
 "
 
-Fig3_sup1 <- panel_A + 
+Fig3_sup2 <- panel_MLD + panel_NIT1andNOS +
   patchwork::plot_layout(design = layout) + #we can change the heights of the rows in our layout (widths also can be defined)
   patchwork::plot_annotation(tag_levels = "A") +  #we can change this to 'a' for small caps or 'i' or '1'
   ggplot2::theme(plot.tag = element_text(size = 12, face='plain', color='black')) #or 'bold', 'italic'
 
-Fig3_sup1
+Fig3_sup2
 
-ggsave("figures/Fig3_sup1.png", limitsize = FALSE, 
-       units = c("px"), Fig3_sup1, width = 2350, height = 1700, bg='white')  
+ggsave("figures/Fig3_sup2.png", limitsize = FALSE, 
+       units = c("px"), Fig3_sup2, width = 2400, height = 1500, bg='white')  
 
-ggsave("figures/Fig3_sup1.pdf", limitsize = FALSE, 
-       units = c("px"), Fig3_sup1, width = 2350, height = 1700)  
+ggsave("figures/Fig3_sup2.pdf", limitsize = FALSE, 
+       units = c("px"), Fig3_sup2, width = 2350, height = 1700)  

### 2022-08-21 code/Fig3_sup3.R

@@ -1,4 +1,4 @@
-#Code to assemble Fig3 sup1 of the Jokura et al Platynereis NOS paper
+#Code to assemble Fig3 sup3 of the Jokura et al Platynereis NOS paper
 #2022 June - Kei Jokura, Gaspar Jekely
 
 rm(list = ls(all.names = TRUE)) #will clear all objects includes hidden objects.
@@ -12,7 +12,7 @@ library(patchwork)
 library(magick)
 
 # assemble figure ---------------------------------------------------------
-panel_A <- ggdraw() + draw_image(readPNG("pictures/NITGC_allseq_mbGC_sGC_CD80_plus_Lopho.fasta.alngapp.fasta.treefile.png"))
+panel_3D <- ggdraw() + draw_image(readPNG("pictures/NITdomain_3D.png"))
 
 #combine panels into Figure and save final figure as pdf and png
 #panels of different sizes
@@ -20,15 +20,14 @@ layout <- "
 A
 "
 
-Fig3_sup1 <- panel_A + 
+Fig3_sup3 <- panel_3D + 
   patchwork::plot_layout(design = layout) + #we can change the heights of the rows in our layout (widths also can be defined)
-  patchwork::plot_annotation(tag_levels = "A") +  #we can change this to 'a' for small caps or 'i' or '1'
   ggplot2::theme(plot.tag = element_text(size = 12, face='plain', color='black')) #or 'bold', 'italic'
 
-Fig3_sup1
+Fig3_sup3
 
-ggsave("figures/Fig3_sup1.png", limitsize = FALSE, 
-       units = c("px"), Fig3_sup1, width = 2350, height = 1700, bg='white')  
+ggsave("figures/Fig3_sup3.png", limitsize = FALSE, 
+       units = c("px"), Fig3_sup3, width = 2400, height = 1600, bg='white')  
 
 ggsave("figures/Fig3_sup1.pdf", limitsize = FALSE, 
        units = c("px"), Fig3_sup1, width = 2350, height = 1700)  

### 2022-08-21 code/Fig3_sup4.R

@@ -1,4 +1,4 @@
-#Code to assemble Fig3 sup1 of the Jokura et al Platynereis NOS paper
+#Code to assemble Fig3 sup4 of the Jokura et al Platynereis NOS paper
 #2022 June - Kei Jokura, Gaspar Jekely
 
 rm(list = ls(all.names = TRUE)) #will clear all objects includes hidden objects.
@@ -12,23 +12,24 @@ library(patchwork)
 library(magick)
 
 # assemble figure ---------------------------------------------------------
-panel_A <- ggdraw() + draw_image(readPNG("pictures/NITGC_allseq_mbGC_sGC_CD80_plus_Lopho.fasta.alngapp.fasta.treefile.png"))
+panel_A <- ggdraw() + draw_image(readPNG("pictures/NITGC_domain_structure.png"))
+panel_B <- ggdraw() + draw_image(readPNG("pictures/NITGC&GreencGull_expression.png"))
 
 #combine panels into Figure and save final figure as pdf and png
 #panels of different sizes
 layout <- "
-A
+ABC
 "
 
-Fig3_sup1 <- panel_A + 
-  patchwork::plot_layout(design = layout) + #we can change the heights of the rows in our layout (widths also can be defined)
-  patchwork::plot_annotation(tag_levels = "A") +  #we can change this to 'a' for small caps or 'i' or '1'
+Fig3_sup4 <- panel_A + plot_spacer()+ panel_B + 
+  patchwork::plot_layout(design = layout, widths = c(1, 0.05, 1)) + #we can change the heights of the rows in our layout (widths also can be defined)
+  patchwork::plot_annotation(tag_levels = "A") +
   ggplot2::theme(plot.tag = element_text(size = 12, face='plain', color='black')) #or 'bold', 'italic'
 
-Fig3_sup1
+Fig3_sup4
 
-ggsave("figures/Fig3_sup1.png", limitsize = FALSE, 
-       units = c("px"), Fig3_sup1, width = 2350, height = 1700, bg='white')  
+ggsave("figures/Fig3_sup4.png", limitsize = FALSE, 
+       units = c("px"), Fig3_sup4, width = 2400, height = 1000, bg='white')  
 
 ggsave("figures/Fig3_sup1.pdf", limitsize = FALSE, 
        units = c("px"), Fig3_sup1, width = 2350, height = 1700)  

### 2022-08-21 code/Fig4.R

@@ -1,4 +1,4 @@
-#Code to assemble Fig5 of the Jokura et al Platynereis NOS paper
+#Code to assemble Fig4 of the Jokura et al Platynereis NOS paper
 #2022 June - Kei Jokura, Gaspar Jekely
 
 # source packages ---------------------------------------------------------
@@ -374,22 +374,22 @@ panel_NIT2MO_INNOS <- ggdraw() + draw_image(readPNG("pictures/NIT2MO_INNOS.png")
 #combine panels into Figure and save final figure as pdf and png
 #panels of different sizes
 layout <- "
-A#B#C
-D#E#F
-G#H#I
+ABC
+DEF
+GHI
 "
 
 Fig4 <- panel_DAFFM + panel_cPRC_NOS11 + panel_cPRC_NOS23 + 
   panel_INNOS + panel_INRGW + panel_NIT2MO_INNOS + panel_NIT2MO_INRGW + panel_Ser + panel_MC +
-  patchwork::plot_layout(design = layout, widths = c(2.5, 0.5)) + #we can change the heights of the rows in our layout (widths also can be defined)
-  patchwork::plot_annotation(tag_levels = 'A') &  #we can change this to 'a' for small caps or 'i' or '1'
+  patchwork::plot_layout(design = layout) + #we can change the heights of the rows in our layout (widths also can be defined)
+  patchwork::plot_annotation(tag_levels = 'A') +  #we can change this to 'a' for small caps or 'i' or '1'
   ggplot2::theme(plot.tag = element_text(size = 12, 
           face='plain')) #or 'bold', 'italic'
 
 Fig4
 
 ggsave("figures/Fig4.png", limitsize = FALSE, 
-       units = c("px"), Fig4, width = 2400, height = 2400, bg='white')  
+       units = c("px"), Fig4, width = 2400, height = 1400, bg='white')  
 
 ggsave("figures/Fig4.pdf", limitsize = FALSE, 
        units = c("px"), Fig4, width = 2350, height = 1700)  

### 2022-08-21 code/Fig4_sup1.R

@@ -1,0 +1,26 @@
+#Code to assemble Fig4 sup1 of the Jokura et al Platynereis NOS paper
+#2022 June - Kei Jokura, Gaspar Jekely
+
+# source packages ---------------------------------------------------------
+source("code/Packages_Catmaid_connection.R")
+
+# assemble figure ---------------------------------------------------------
+panel_A <- ggdraw() + draw_image(readPNG("pictures/Diagram of agarose embedding immunostaining.png"))
+panel_B <- ggdraw() + draw_image(readPNG("pictures/n54hpf_WT_e-1_GCaMP6s.png"))
+panel_C <- ggdraw() + draw_image(readPNG("pictures/n54hpf_WT_e-1_RGWa.png"))
+
+#combine panels into Figure and save final figure as pdf and png
+#panels of different sizes
+
+Fig4_sup1 <- panel_A - (panel_B + panel_C) +
+  patchwork::plot_layout(ncol = 1, heights = c(1,2.5)) + #we can change the heights of the rows in our layout (widths also can be defined)
+  patchwork::plot_annotation(tag_levels = "A") +  #we can change this to 'a' for small caps or 'i' or '1'
+  ggplot2::theme(plot.tag = element_text(size = 12, face='plain', color='black')) #or 'bold', 'italic'
+
+Fig4_sup1
+
+ggsave("figures/Fig4_sup1.png", limitsize = FALSE, 
+       units = c("px"), Fig4_sup1, width = 2350, height = 1700, bg='white')  
+
+ggsave("figures/Fig4_sup1.pdf", limitsize = FALSE, 
+       units = c("px"), Fig4_sup1, width = 2350, height = 1700)  

### 2022-08-21 code/Fig5.R

@@ -24,14 +24,14 @@ AB
 
 Fig5 <- panel_A + panel_B + 
   patchwork::plot_layout(design = layout, widths = c(1, 2.5)) + #we can change the heights of the rows in our layout (widths also can be defined)
-  patchwork::plot_annotation(tag_levels = 'A') &  #we can change this to 'a' for small caps or 'i' or '1'
+  patchwork::plot_annotation(tag_levels = 'A') +  #we can change this to 'a' for small caps or 'i' or '1'
   ggplot2::theme(plot.tag = element_text(size = 12, 
                                          face='plain')) #or 'bold', 'italic'
 
 Fig5
 
 ggsave("figures/Fig5.png", limitsize = FALSE, 
-       units = c("px"), Fig5, width = 2400, height = 2400, bg='white')  
+       units = c("px"), Fig5, width = 2400, height = 1400, bg='white')  
 
 ggsave("figures/Fig5.pdf", limitsize = FALSE, 
        units = c("px"), Fig5, width = 2350, height = 1700)  

### 2022-08-22 text/Supplements.Rmd
delete some figures

@@
 
-![**Figure 4---figure supplement 1.** (A) Schematic diagram of the immunostaining procedure after calcium imaging.
+![**Figure 4---figure supplement 1.** (A) An INNOS visualised by DAF-FM is shown (a white arrowhead).
+A black arrowhead indicates the position of the adult eye.
+The white dashed line indicates the outer frame of the larva.
+All subsequent images are of 2-day-old larvae.
+(B) Dual-colour imaging of DAF-FM (green) and RGECO1a (magenta) allows simultaneous visualisation of the INNOS and cPRCs.
+(C) Enlarged view of (B).
+cPRCs and INNOS are indicated by arrows.](../figures/Fig4_sup1.png){width="100%"}
+
+![**Figure 4---figure supplement 2.** (A) Schematic diagram of the immunostaining procedure after calcium imaging.
 (B) The position of the cPRC (purple dashed line) in calcium imaging with GCaMP6s.
@@
 (C) Immunostaining image using RGWamide (RGWa) and serotonin (Ser) antibodies (red) and DAPI (light blue) in the same region as (B).
-INRGW (green dashed line) and Ser-h1 (black arrowheads) identified from the numbered nuclear positions are shown.
-.](../figures/Fig4_sup1.png){width="100%"}
+INRGW (green dashed line) and Ser-h1 (black arrowheads) identified from the numbered nuclear positions are shown.](../figures/Fig4_sup2.png){width="100%"}

### 2022-08-22 code/Fig3_sup4.R

@@ -21,7 +21,7 @@ layout <- "
 ABC
 "
 
-Fig3_sup4 <- panel_A + plot_spacer()+ panel_B + 
+Fig3_sup4 <- panel_A + plot_spacer() + panel_B + 
   patchwork::plot_layout(design = layout, widths = c(1, 0.05, 1)) + #we can change the heights of the rows in our layout (widths also can be defined)
   patchwork::plot_annotation(tag_levels = "A") +
   ggplot2::theme(plot.tag = element_text(size = 12, face='plain', color='black')) #or 'bold', 'italic'

### 2022-08-22 code/Fig4_sup1.R

@@ -5,22 +5,29 @@
 source("code/Packages_Catmaid_connection.R")
 
 # assemble figure ---------------------------------------------------------
-panel_A <- ggdraw() + draw_image(readPNG("pictures/Diagram of agarose embedding immunostaining.png"))
-panel_B <- ggdraw() + draw_image(readPNG("pictures/n54hpf_WT_e-1_GCaMP6s.png"))
-panel_C <- ggdraw() + draw_image(readPNG("pictures/n54hpf_WT_e-1_RGWa.png"))
+panel_A <- ggdraw() + draw_image(readPNG("pictures/55hpf_DAF-FM_134.95um.png")) +
+  draw_line(x = c(0.05, 0.42), y = c(0.1, 0.1), color = "white", size = 0.5) +
+  draw_label(expression(paste("50 ", mu, "m")), x = 0.23, y = 0.13, color = "white", size = 10)
+
+panel_B <- ggdraw() + draw_image(readPNG("pictures/55hpf_DAF-FM_RGECO_134.95um.png"))
+
+panel_C <- ggdraw() + draw_image(readPNG("pictures/55hpf_DAF-FM_RGECO_highmag_56.04um.png")) +
+  draw_line(x = c(0.05, 0.407), y = c(0.1, 0.1), color = "white", size = 0.5) +
+  draw_label(expression(paste("20 ", mu, "m")), x = 0.23, y = 0.13, color = "white", size = 10)
 
 #combine panels into Figure and save final figure as pdf and png
 #panels of different sizes
+layout <- "
+ABCDE
+"
 
-Fig4_sup1 <- panel_A - (panel_B + panel_C) +
-  patchwork::plot_layout(ncol = 1, heights = c(1,2.5)) + #we can change the heights of the rows in our layout (widths also can be defined)
+Fig4_sup1 <- panel_A + plot_spacer() +panel_B + plot_spacer() +panel_C +
+  patchwork::plot_layout(design = layout, widths = c(1, 0.05, 1, 0.05, 1)) + #we can change the heights of the rows in our layout (widths also can be defined)
   patchwork::plot_annotation(tag_levels = "A") +  #we can change this to 'a' for small caps or 'i' or '1'
   ggplot2::theme(plot.tag = element_text(size = 12, face='plain', color='black')) #or 'bold', 'italic'
 
-Fig4_sup1
-
 ggsave("figures/Fig4_sup1.png", limitsize = FALSE, 
[…]

### 2022-08-22 code/Fig4_sup2.R

@@ -1,0 +1,26 @@
+#Code to assemble Fig4 sup2 of the Jokura et al Platynereis NOS paper
+#2022 June - Kei Jokura, Gaspar Jekely
+
+# source packages ---------------------------------------------------------
+source("code/Packages_Catmaid_connection.R")
+
+# assemble figure ---------------------------------------------------------
+panel_A <- ggdraw() + draw_image(readPNG("pictures/Diagram of agarose embedding immunostaining.png"))
+panel_B <- ggdraw() + draw_image(readPNG("pictures/n54hpf_WT_e-1_GCaMP6s.png"))
+panel_C <- ggdraw() + draw_image(readPNG("pictures/n54hpf_WT_e-1_RGWa.png"))
+panel_D <- ggdraw() + draw_image(readPNG("pictures/n52hpf_WT_a-2_ROI60_PD1.54_power1_pin216.9_frame_201-241_401-441_co_cPRC.png"))
+panel_E <- ggdraw() + draw_image(readPNG("pictures/n52hpf_WT_a-2_ROI60_PD1.54_power1_pin216.9_frame_201-241_401-441_co_NOS.png"))
+
+#combine panels into Figure and save final figure as pdf and png
+#panels of different sizes
+
+Fig4_sup2 <- {((panel_A / (panel_B|panel_C))/(panel_D|panel_E))} +
+  patchwork::plot_layout(nrow = 3, heights = c(1,2.25, 2.25)) + #we can change the heights of the rows in our layout (widths also can be defined)
+  patchwork::plot_annotation(tag_levels = "A") +  #we can change this to 'a' for small caps or 'i' or '1'
+  ggplot2::theme(plot.tag = element_text(size = 12, face='plain', color='black')) #or 'bold', 'italic'
+
+ggsave("figures/Fig4_sup2.png", limitsize = FALSE, 
+       units = c("px"), Fig4_sup2, width = 2400, height = 3200, bg='white')  
+
+ggsave("figures/Fig4_sup2.pdf", limitsize = FALSE, 
+       units = c("px"), Fig4_sup2, width = 2350, height = 1700)  

### 2022-09-12 text/Manuscript.Rmd
Upload Manuscript Figure
new behavior analysis

@@
 (G-J) The intensity changes in cGMP fluorescence over time for the four conditions are shown.
-Gray traces show the optical responses to individual intensities (n > 6).
+Gray traces show the optical responses to individual intensities (n \> 6).
 The individual intensities are normalized (ΔF/F0).
@@
 ![**Figure 4.
-NO produced by UV stimulation to cPRCs induces a depolarising response in cPRC via NIT-GC1.** (A) An INNOS visualised by DAF-FM is shown (a white arrowhead).
-A black arrowhead indicates the position of the adult eye.
-The white dashed line indicates the outer frame of the larva.
-All subsequent images are of 2-day-old larvae.
-(B) Dual-colour imaging of DAF-FM (green) and RGECO1a (magenta) allows simultaneous visualisation of the INNOS and cPRCs.
-(C) The intensity of DAF-FM in INNOS changes over time when cPRC (top) and other cells (NC) (bottom) are stimulated with UV (405 nm).
+NO produced by UV stimulation to cPRCs induces a depolarising response in cPRC via NIT-GC1.** (A) The intensity of DAF-FM in INNOS changes over time when cPRC (top) and other cells (NC) (bottom) are stimulated with UV (405 nm).
 Hereafter, light purple squares indicate the timing of UV stimulation, gray traces show the optical responses to individual intensities, each intensities are normalized (ΔF/F0) and the dark lines and the bands show the mean value and the standard error range of the mean.
-(D) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in WT (purple) and NOS mutant (NOSΔ23) (gray) larvae.
-(E) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in control (non-injection) and NIT-GC1 knockdown by morpholino (NITGC1-MO1 and 2) larvae.
-(F) Co-expression analysis image of the cPRC marker gene (NIT-GC1: magenta) and the INNOS marker gene (RYa-pNP: green) using in situ HCR.
+(B) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in WT (purple) and NOS mutant (NOSΔ23) (gray) larvae.
+(C) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in control (non-injection) and NIT-GC1 knockdown by morpholino (NITGC1-MO1 and 2) larvae.
+(D) Co-expression analysis image of the cPRC marker gene (NIT-GC1: magenta) and the INNOS marker gene (RYa-pNP: green) using in situ HCR.
 White arrowheads indicate INNOS strongly expressing RYa-pNP.
-(G) INNOS (blue dashed line) identified from calcium imaging using GCaMP6s by their position in relation to the cPRC (purple dashed line).
-(H) Comparison of changes over time in GCaMP6s intensities in cPRC (purple) and INNOS (blue) when cPRC is stimulated with UV in WT (left) and NOS mutant (NOSΔ11: right) larvae.
-(L) Comparison of changes over time in GCaMP6s intensities in cPRC (purple) and INRGW (green) when the cPRC are stimulated with UV in WT (left) and NOS mutant (NOSΔ23: right) larvae.](../figures/Fig4.png){width="100%"}
+(E) INNOS (blue dashed line) identified from calcium imaging using GCaMP6s by their position in relation to the cPRC (purple dashed line).
+(F) Comparison of changes over time in GCaMP6s intensities in cPRC (purple) and INNOS (blue) when cPRC is stimulated with UV in WT (left) and NOS mutant (NOSΔ11: right) larvae.
+(G) Comparison of changes over time in GCaMP6s intensities in cPRC (purple) and INRGW (green) when the cPRC are stimulated with UV in WT (left) and NOS mutant (NOSΔ23: right) larvae.
+(H)(I)](../figures/Fig4.png){width="100%"}
 

### 2022-09-12 code/220609_vertical_displacement_2d_WTvsNOS.R

@@ -25,17 +25,16 @@ WTvsNOS_2d <- bind_rows(WT_2d_tidy_phenotype, NOS_2d_tidy_phenotype) %>%
   mutate(phenotype = factor(phenotype, levels = c("WT", "NOS")))
 
 WTvsNOS_2d %>%
-  filter(!(condition %in% c("dark_30", "dark_60", "dark_90", "dark_120", "dark_150", "dark_180", 
-                            "dark_210", "dark_240", "dark_270", "dark_300", "sideUV_30", "sideUV_60", "sideUV_90", "sideUV_120")))%>%
   ggplot() +
   aes(x = reorder(condition, time), y = displacement, fill = phenotype) +
   geom_boxplot() +
   theme_minimal()+
   theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   labs(x = "30 sec bins", y = "Vertical displacement (mm/s)")+
-  scale_y_continuous(limits = c(-1, 1))
+  scale_y_continuous(limits = c(-1, 1.2))
 
-
+ggsave("pictures/vertical_displacement_2dpf_WTvsNOSmix.png", limitsize = FALSE, 
+       units = c("px"), width = 3000, height = 2000, bg='white')
 
 
 

### 2022-09-12 code/220610_vertical_displacement_3d_WTvsNOS.R

@@ -25,8 +25,6 @@ WTvsNOS_3d <- bind_rows(WT_3d_tidy_phenotype, NOS_3d_tidy_phenotype) %>%
   mutate(phenotype = factor(phenotype, levels = c("WT", "NOS")))
 
 WTvsNOS_3d %>%
-  filter(!(condition %in% c("dark_30", "dark_60", "dark_90", "dark_120", "dark_150", "dark_180", 
-                            "dark_210", "dark_240", "dark_270", "dark_300", "sideUV_30", "sideUV_60", "sideUV_90", "sideUV_120")))%>%
   ggplot() +
   aes(x = reorder(condition, time), y = displacement, fill = phenotype) +
   geom_boxplot() +
@@ -34,6 +32,9 @@ WTvsNOS_3d %>%
   theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   labs(x = "30 sec bins", y = "Vertical displacement (mm/s)")+
   scale_y_continuous(limits = c(-1, 1))
+
+ggsave("pictures/vertical_displacement_3dpf_WTvsNOSmix.png", limitsize = FALSE, 
+       units = c("px"), width = 3000, height = 2000, bg='white')
 
 
 

### 2022-09-12 code/220825_tracking raw data_originating and extracting 75.R

@@ -1,0 +1,768 @@
+rm(list = ls(all.names = TRUE)) #will clear all objects includes hidden objects.
+gc() #free up memory and report the memory usage.
+
+# load some packages
+library(tidyverse)
+library(cowplot)
+library(png)
+library(patchwork)
+library(magick)
+library(ggplot2)
+library(esquisse)
+library(readr)
+
+df_suvb5 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/NOS/NOS_0703_01/raw data/NOS_0703_01_sUV_before_1-5.csv", skip = 3)
+df_suv5 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/NOS/NOS_0703_01/raw data/NOS_0703_01_sUV_1-5.csv", skip = 3)
+df_suv10 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/NOS/NOS_0703_01/raw data/NOS_0703_01_sUV_6-10.csv", skip = 3)
+df_suv15 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/NOS/NOS_0703_01/raw data/NOS_0703_01_sUV_11-15.csv", skip = 3)
+df_suv20 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/NOS/NOS_0703_01/raw data/NOS_0703_01_sUV_16-20.csv", skip = 3)
+df_suv25 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/NOS/NOS_0703_01/raw data/NOS_0703_01_sUV_21-25.csv", skip = 3)
+df_suv30 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/NOS/NOS_0703_01/raw data/NOS_0703_01_sUV_26-30.csv", skip = 3)
+df_suv35 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/NOS/NOS_0703_01/raw data/NOS_0703_01_sUV_31-35.csv", skip = 3)
+df_suv40 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/NOS/NOS_0703_01/raw data/NOS_0703_01_sUV_36-40.csv", skip = 3)
[…]

### 2022-09-12 code/220830_vertical_displacement_NOS.R

@@ -1,0 +1,393 @@
+rm(list = ls(all.names = TRUE)) #will clear all objects includes hidden objects.
+gc() #free up memory and report the memory usage.
+
+# load some packages
+library(tidyverse)
+library(cowplot)
+library(png)
+library(patchwork)
+library(magick)
+library(ggplot2)
+library(esquisse)
+library(readr)
+
+
+df_suv_NOS_0623_00_b5 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/NOS/NOS_0623_00/NOS_0623_00_sUV_vertical_before_1-5.csv")
+df_suv_NOS_0623_00_5 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/NOS/NOS_0623_00/NOS_0623_00_sUV_vertical_1-5.csv")
+df_suv_NOS_0623_00_10 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/NOS/NOS_0623_00/NOS_0623_00_sUV_vertical_6-10.csv")
+df_suv_NOS_0623_00_15 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/NOS/NOS_0623_00/NOS_0623_00_sUV_vertical_11-15.csv")
+df_suv_NOS_0623_00_20 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/NOS/NOS_0623_00/NOS_0623_00_sUV_vertical_16-20.csv")
+df_suv_NOS_0623_00_25 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/NOS/NOS_0623_00/NOS_0623_00_sUV_vertical_21-25.csv")
+df_suv_NOS_0623_00_30 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/NOS/NOS_0623_00/NOS_0623_00_sUV_vertical_26-30.csv")
+df_suv_NOS_0623_00_35 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/NOS/NOS_0623_00/NOS_0623_00_sUV_vertical_31-35.csv")
+df_suv_NOS_0623_00_40 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/NOS/NOS_0623_00/NOS_0623_00_sUV_vertical_36-40.csv")
[…]

### 2022-09-12 code/220830_vertical_displacement_WT.R

@@ -1,0 +1,359 @@
+rm(list = ls(all.names = TRUE)) #will clear all objects includes hidden objects.
+gc() #free up memory and report the memory usage.
+
+# load some packages
+library(tidyverse)
+library(cowplot)
+library(png)
+library(patchwork)
+library(magick)
+library(ggplot2)
+library(esquisse)
+library(readr)
+
+
+
+
+df_suv_WT_0701_08_b5 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/WT/WT_0701_08/WT_0701_08_sUV_vertical_before_1-5.csv", col_select = c(nor_x, nor_y))
+df_suv_WT_0701_08_5 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/WT/WT_0701_08/WT_0701_08_sUV_vertical_1-5.csv", col_select = c(nor_x, nor_y))
+df_suv_WT_0701_08_10 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/WT/WT_0701_08/WT_0701_08_sUV_vertical_6-10.csv", col_select = c(nor_x, nor_y))
+df_suv_WT_0701_08_15 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/WT/WT_0701_08/WT_0701_08_sUV_vertical_11-15.csv", col_select = c(nor_x, nor_y))
+df_suv_WT_0701_08_20 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/WT/WT_0701_08/WT_0701_08_sUV_vertical_16-20.csv", col_select = c(nor_x, nor_y))
+df_suv_WT_0701_08_25 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/WT/WT_0701_08/WT_0701_08_sUV_vertical_21-25.csv", col_select = c(nor_x, nor_y))
+df_suv_WT_0701_08_30 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/WT/WT_0701_08/WT_0701_08_sUV_vertical_26-30.csv", col_select = c(nor_x, nor_y))
+df_suv_WT_0701_08_35 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/WT/WT_0701_08/WT_0701_08_sUV_vertical_31-35.csv", col_select = c(nor_x, nor_y))
[…]

### 2022-09-12 code/220831_vertical_displacement_mean.R

@@ -1,0 +1,1042 @@
+rm(list = ls(all.names = TRUE)) #will clear all objects includes hidden objects.
+gc() #free up memory and report the memory usage.
+
+# load some packages
+library(tidyverse)
+library(cowplot)
+library(png)
+library(patchwork)
+library(magick)
+library(ggplot2)
+library(esquisse)
+library(readr)
+
+df_suv_WT_0701_08_b5 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/WT/WT_0701_08/WT_0701_08_sUV_vertical_before_1-5.csv")
+df_suv_WT_0701_08_5 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/WT/WT_0701_08/WT_0701_08_sUV_vertical_1-5.csv")
+df_suv_WT_0701_08_10 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/WT/WT_0701_08/WT_0701_08_sUV_vertical_6-10.csv")
+df_suv_WT_0701_08_15 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/WT/WT_0701_08/WT_0701_08_sUV_vertical_11-15.csv")
+df_suv_WT_0701_08_20 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/WT/WT_0701_08/WT_0701_08_sUV_vertical_16-20.csv")
+df_suv_WT_0701_08_25 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/WT/WT_0701_08/WT_0701_08_sUV_vertical_21-25.csv")
+df_suv_WT_0701_08_30 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/WT/WT_0701_08/WT_0701_08_sUV_vertical_26-30.csv")
+df_suv_WT_0701_08_35 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/WT/WT_0701_08/WT_0701_08_sUV_vertical_31-35.csv")
+df_suv_WT_0701_08_40 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/WT/WT_0701_08/WT_0701_08_sUV_vertical_36-40.csv")
[…]

### 2022-09-12 code/220901_tracking_for circular histogram.R

@@ -1,0 +1,923 @@
+rm(list = ls(all.names = TRUE)) #will clear all objects includes hidden objects.
+gc() #free up memory and report the memory usage.
+
+# load some packages
+library(tidyverse)
+library(cowplot)
+library(png)
+library(patchwork)
+library(magick)
+library(ggplot2)
+library(esquisse)
+library(REdaS)
+library(readr)
+
+df_suvb5 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/WT/sUV_WT_b5.csv")
+df_suv5 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/WT/sUV_WT_5.csv")
+df_suv10 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/WT/sUV_WT_10.csv")
+df_suv15 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/WT/sUV_WT_15.csv")
+df_suv20 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/WT/sUV_WT_20.csv")
+df_suv25 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/WT/sUV_WT_25.csv")
+df_suv30 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/WT/sUV_WT_30.csv")
+df_suv35 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/WT/sUV_WT_35.csv")
+df_suv40 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/WT/sUV_WT_40.csv")
+df_suv45 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/WT/sUV_WT_45.csv")
+df_suv50 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/WT/sUV_WT_50.csv")
+df_suv55 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/WT/sUV_WT_55.csv")
+df_suv60 <- read_csv("C:/Users/klamd/Documents/Martin behaviour/WT/sUV_WT_60.csv")
+
+
+
+df_suvb5_first <- df_suvb5 %>% filter(nor_x > 0, nor_y > 0)
+df_suvb5_second <- df_suvb5 %>% filter(nor_x < 0, nor_y > 0)
+df_suvb5_third <- df_suvb5 %>% filter(nor_x < 0, nor_y < 0)
[…]

### 2022-09-12 code/220909_Track_features.R

@@ -1,0 +1,1117 @@
+rm(list = ls(all.names = TRUE)) #will clear all objects includes hidden objects.
+gc() #free up memory and report the memory usage.
+
+# load some packages
+library(tidyverse)
+library(cowplot)
+library(png)
+library(patchwork)
+library(magick)
+library(ggplot2)
+library(esquisse)
+library(readr)
+
+tf_suv_WT_0701_08_b5 <- read_csv("~/Martin behaviour/WT/WT_0701_08/tracking parameters/WT_0701_08_sUV_before_1-5_t.csv", 
+                                 col_names = FALSE, skip = 4)
+tf_suv_WT_0701_08_5 <- read_csv("~/Martin behaviour/WT/WT_0701_08/tracking parameters/WT_0701_08_sUV_1-5_t.csv", 
+                                 col_names = FALSE, skip = 4)
+tf_suv_WT_0701_08_10 <- read_csv("~/Martin behaviour/WT/WT_0701_08/tracking parameters/WT_0701_08_sUV_6-10_t.csv", 
+                                col_names = FALSE, skip = 4)
+tf_suv_WT_0701_08_15 <- read_csv("~/Martin behaviour/WT/WT_0701_08/tracking parameters/WT_0701_08_sUV_11-15_t.csv", 
+                                col_names = FALSE, skip = 4)
+tf_suv_WT_0701_08_20 <- read_csv("~/Martin behaviour/WT/WT_0701_08/tracking parameters/WT_0701_08_sUV_16-20_t.csv", 
+                                col_names = FALSE, skip = 4)
+tf_suv_WT_0701_08_25 <- read_csv("~/Martin behaviour/WT/WT_0701_08/tracking parameters/WT_0701_08_sUV_21-25_t.csv", 
+                                col_names = FALSE, skip = 4)
+tf_suv_WT_0701_08_30 <- read_csv("~/Martin behaviour/WT/WT_0701_08/tracking parameters/WT_0701_08_sUV_26-30_t.csv", 
+                                col_names = FALSE, skip = 4)
+tf_suv_WT_0701_08_35 <- read_csv("~/Martin behaviour/WT/WT_0701_08/tracking parameters/WT_0701_08_sUV_31-35_t.csv", 
+                                col_names = FALSE, skip = 4)
+tf_suv_WT_0701_08_40 <- read_csv("~/Martin behaviour/WT/WT_0701_08/tracking parameters/WT_0701_08_sUV_36-40_t.csv", 
+                                col_names = FALSE, skip = 4)
[…]

### 2022-09-12 code/Fig2.R

@@ -9,8 +9,8 @@ source("code/Packages_to_load.R")
 #read png convert to image panel
 
 panel_Architecture <- ggdraw() + draw_image(readPNG("pictures/NOS-Architecture.png"))
-panel_2dpf <- ggdraw() + draw_image(readPNG("pictures/2dpf_wt_vs_NOS_mutants_for_discussion_1.png"))
-panel_3dpf <- ggdraw() + draw_image(readPNG("pictures/3dpf_wt_vs_NOS_mutants_for_discussion.png"))
+panel_2dpf <- ggdraw() + draw_image(readPNG("pictures/vertical_displacement_2dpf_WTvsNOSmix.png"))
+panel_3dpf <- ggdraw() + draw_image(readPNG("pictures/vertical_displacement_3dpf_WTvsNOSmix.png"))
 
 #combine panels into Figure and save final figure as pdf and png
 #panels of different sizes

### 2022-09-30 code/Fig1.R
Change of content

@@ -245,14 +245,14 @@ panel_NOS2d_ant_HCR <- ggdraw() + draw_image(readPNG("pictures/HCR-IHC_51_AP_NOS
   draw_line(x = c(0.05, 0.407), y = c(0.1, 0.1), color = "white", size = 0.5) +
   draw_label(expression(paste("20 ", mu, "m")), x = 0.23, y = 0.13, color = "white", size = 10)
 
-panel_NOS3d_ant_HCR <- ggdraw() + draw_image(readPNG("pictures/HCR-IHC_72_AP_NOS_actub_118.08um.png")) +
+panel_NOS3d_ant_HCR <- ggdraw() + draw_image(readPNG("pictures/HCR_72_AP_NOS_92.02um.png")) +
   draw_label("HCR in situ", x = 0.3, y = 0.99, size = 10) +
   draw_label("INNOS", x = 0.55, y = 0.67, color='#CC79A7', size = 12, fontface='bold') +
   draw_label("cPRC", x = 0.3, y = 0.85, color='white',size = 12, fontface='bold')  +
   draw_line(x = c(0.05, 0.407), y = c(0.1, 0.1), color = "white", size = 0.5) +
   draw_label(expression(paste("20 ", mu, "m")), x = 0.23, y = 0.13, color = "white", size = 10)
 
-panel_NOS3d_ventr_HCR <- ggdraw() + draw_image(readPNG("pictures/HCR-IHC_72_DV_NOS_actub_112.55um.png")) +
+panel_NOS3d_ventr_HCR <- ggdraw() + draw_image(readPNG("pictures/HCR_72_DV_NOS_71.46um.png")) +
   draw_label("HCR in situ", x = 0.3, y = 0.99, size = 10) +
   draw_label("INNOS", x = 0.55, y = 0.67, color='#CC79A7', size = 12, fontface='bold') +
   draw_label("cPRC", x = 0.3, y = 0.85, color='white',size = 12, fontface='bold')  +
@@ -271,7 +271,7 @@ ABCD
 EFGH
 "
 
-Fig1 <- panel_SEM + panel_NOS3d_ventr_HCR + panel_NOS2d_ant_HCR + panel_NOS3d_ant_HCR +  
+Fig1 <- panel_SEM + panel_NOS3d_ventr_HCR + panel_NOS3d_ant_HCR + panel_NOS2d_ant_HCR +  
   panel_NOS3d + panel_INNOS_ventr + panel_INNOS_ant + panel_Network +
[…]

### 2022-09-30 code/Fig1_sup2.R

@@ -1,28 +1,0 @@
-#Code to assemble Fig1 of the Jokura et al Platynereis NOS paper
-#2022 June - Kei Jokura, Gaspar Jekely
-
-# source packages ---------------------------------------------------------
-source("code/Packages_Catmaid_connection.R")
-
-# assemble figure ---------------------------------------------------------
-panel_A <- ggdraw() + draw_image(readPNG("pictures/IHC_54_AP_NOS_actub_71.46um.png"))
-panel_B <- ggdraw() + draw_image(readPNG("pictures/HCR_52_AP_NOS_RYa_101.29um.png"))
-
-#combine panels into Figure and save final figure as pdf and png
-#panels of different sizes
-layout <- "
-AB
-"
-
-Fig1_sup2 <- panel_A + panel_B +
-  patchwork::plot_layout(design = layout) + #we can change the heights of the rows in our layout (widths also can be defined)
-  patchwork::plot_annotation(tag_levels = "A") +  #we can change this to 'a' for small caps or 'i' or '1'
-  ggplot2::theme(plot.tag = element_text(size = 12, face='plain', color='black')) #or 'bold', 'italic'
-
-Fig1_sup2
-
-ggsave("figures/Fig1_sup2.png", limitsize = FALSE, 
-       units = c("px"), Fig1_sup2, width = 2350, height = 1700, bg='white')  
-
-ggsave("figures/Fig1_sup2.pdf", limitsize = FALSE, 
-       units = c("px"), Fig1_sup2, width = 2350, height = 1700)  

### 2022-09-30 code/Fig2.R

@@ -32,7 +32,7 @@ df_NOS <- sapply(files, read_csv, simplify = FALSE)%>%
 
 
 
-# tidying the data -----------------------------------------------------------------------
+# tidying Martin data (30 sec bins, vertical displacement)------------------------------------------------
 WT_2d_tidy <- gather(WT_2d, key="WT", value="displacement", -time, -condition)
 WT_2d_tidy_phenotype <- gather(WT_2d_tidy, key="phenotype", value="batch", -time, -condition, -displacement)
 
@@ -54,17 +54,31 @@ WTvsNOS_3d <- bind_rows(WT_3d_tidy_phenotype, NOS_3d_tidy_phenotype) %>%
   mutate(phenotype = factor(phenotype, levels = c("WT", "NOS")))
 
 
-
 # plot Martin data (2d larvae, 30 sec bins, vertical displacement)-------------------------------
 
 WTvsNOS_2d %>%
   ggplot() +
+  annotate("rect", xmin="sideUV_30", xmax="sideUV_120", ymin=-Inf, ymax=Inf,alpha=0.1, fill="blue") +
+  annotate("rect", xmin="top488_30", xmax="top488_120", ymin=-Inf, ymax=Inf,alpha=0.1, fill="deepskyblue") +
+  annotate("rect", xmin="top395_30", xmax="top395_120", ymin=-Inf, ymax=Inf,alpha=0.1, fill="blue") +
   aes(x = reorder(condition, time), y = displacement, fill = phenotype) +
   geom_boxplot() +
   theme_minimal()+
   theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   labs(x = "30 sec bins", y = "Vertical displacement (mm/s)")+
-  scale_y_continuous(limits = c(-1, 1.2))
+  scale_y_continuous(limits = c(-1, 1.2)) +
+  scale_x_discrete(limit=c("dark_30", "dark_60", "dark_90", "dark_120", 
+                             "dark_150", "dark_180", "dark_210", "dark_240", 
+                             "sideUV_30", "sideUV_60", "sideUV_90", "sideUV_120", 
+                             "dark1_30", "dark1_60", "dark1_90", "dark1_120", 
+                             "top488_30", "top488_60", "top488_90", "top488_120", 
+                             "dark2_30", "dark2_60", "dark2_90", "dark2_120", 
+                             "top395_30", "top395_60", "top395_90", "top395_120", 
[…]

### 2022-09-30 code/Fig3.R

@@ -214,7 +214,8 @@ panel_HCR_NIT1 <- ggdraw() + draw_image(readPNG("pictures/HCR-IHC_51_AP_NITGC1_a
 
 panel_HCR_NIT1_cOps <- ggdraw() + draw_image(readPNG("pictures/HCR_RT28_AP_NITGC1_c-opsin1_112.55um.png")) +
   draw_label("HCR in situ", x = 0.3, y = 0.99, size = 11) +
-  draw_label("NIT-GC1", x = 0.15, y = 0.08, color='#CC79A7', size = 12, fontface='bold')
+  draw_label("NIT-GC1", x = 0.15, y = 0.08, color='#CC79A7', size = 12, fontface='bold')+
+  draw_label("c-opsin1", x = 0.85, y = 0.08, color='green', size = 12, fontface='bold')
 
 panel_IHC_NIT1 <- ggdraw() + draw_image(readPNG("pictures/IHC_55_AP_NITGC1_actub_58.47um.png")) +
   draw_label("IHC", x = 0.2, y = 0.99, size = 11) +
@@ -227,8 +228,6 @@ panel_IHC_NIT2 <- ggdraw() + draw_image(readPNG("pictures/IHC_55_AP_NITGC2_actub
   draw_label("cPRC", x = 0.22, y = 0.8, color='white',size = 12, fontface='bold') +
   draw_label("anti-NIT-GC2", x = 0.22, y = 0.08, color='#CC79A7', size = 12, fontface='bold') +
   draw_label("acTub", x = 0.85, y = 0.08, color='green', size = 12, fontface='bold')
-
-panel_GcG_scheme <- ggdraw() + draw_image(readPNG("pictures/NITGC1_cGMPassay_2.png"))
 
 panel_GcG_NIT_DMSO <- ggdraw() + draw_image(readPNG("pictures/GcG-NIT-GC1-DMSO.png")) +
   draw_label("GcG + NIT-GC1", x = 0.35, y = 0.9, size = 11)
@@ -244,25 +243,25 @@ panel_GcG_mutNIT_SNAP <- ggdraw() + draw_image(readPNG("pictures/GcG-mutNIT-GC1-
 layout <- "
 A#B#C#D
 #######
-EEE#FFF
+E#F#G#H
 #######
-H#I#J#K
+III#JJJ
 "
 
[…]

### 2022-09-30 code/Fig3_sup1.R

@@ -12,7 +12,7 @@ library(patchwork)
 library(magick)
 
 # assemble figure ---------------------------------------------------------
-panel_A <- ggdraw() + draw_image(readPNG("pictures/NITGC_allseq_mbGC_sGC_CD80_plus_Lopho.fasta.alngapp.fasta.treefile.png"))
+panel_A <- ggdraw() + draw_image(readPNG("pictures/FINALSeq_NOS_cdhit080_one_Lopho_3rd.fasta.alngapp.fasta.png"))
 
 #combine panels into Figure and save final figure as pdf and png
 #panels of different sizes
@@ -21,11 +21,8 @@ A
 "
 
 Fig3_sup1 <- panel_A + 
-  patchwork::plot_layout(design = layout) + #we can change the heights of the rows in our layout (widths also can be defined)
-  patchwork::plot_annotation(tag_levels = "A") +  #we can change this to 'a' for small caps or 'i' or '1'
-  ggplot2::theme(plot.tag = element_text(size = 12, face='plain', color='black')) #or 'bold', 'italic'
+  patchwork::plot_layout(design = layout) #we can change the heights of the rows in our layout (widths also can be defined)
 
-Fig3_sup1
 
 ggsave("figures/Fig3_sup1.png", limitsize = FALSE, 
        units = c("px"), Fig3_sup1, width = 2350, height = 1700, bg='white')  

### 2022-09-30 code/Fig3_sup2.R

@@ -12,24 +12,47 @@ library(patchwork)
 library(magick)
 
 # assemble figure ---------------------------------------------------------
-panel_MLD <- ggdraw() + draw_image(readPNG("pictures/HCR_51_AP_NITGC1_MLD_112.55um.png"))
-panel_NIT1andNOS <- ggdraw() + draw_image(readPNG("pictures/IHC_55_NIT1_NOS-RAT1_acTub_AP_2_Airyscan Processing-Orthogonal Projection-11_c1-3_71.46um.png"))
+panel_MLD <- ggdraw() + draw_image(readPNG("pictures/HCR_51_DV_NITGC1_MLD_DAPI_112.55µm.png"))+
+  draw_label("HCR in situ", x = 0.3, y = 0.99, size = 10) +
+  draw_label("NIT1", x = 0.55, y = 0.67, color='magenta', size = 12, fontface='bold') +
+  draw_label("MLD", x = 0.3, y = 0.85, color='green',size = 12, fontface='bold')  +
+  draw_line(x = c(0.05, 0.407), y = c(0.1, 0.1), color = "white", size = 0.5) +
+  draw_label(expression(paste("20 ", mu, "m")), x = 0.23, y = 0.13, color = "white", size = 10)
+
+panel_NIT2 <- ggdraw() + draw_image(readPNG("pictures/2d_647-NIT-GC2_546-c-opsin1_AP_N_106.62um.png"))+
+  draw_label("HCR in situ", x = 0.3, y = 0.99, size = 10) +
+  draw_label("NIT2", x = 0.55, y = 0.67, color='magenta', size = 12, fontface='bold') + 
+  draw_line(x = c(0.05, 0.407), y = c(0.1, 0.1), color = "white", size = 0.5) +
+  draw_label(expression(paste("20 ", mu, "m")), x = 0.23, y = 0.13, color = "white", size = 10)
+
+panel_cop1 <- ggdraw() + draw_image(readPNG("pictures/2d_647-NIT-GC2_546-c-opsin1_AP_c_106.62um.png"))+
+  draw_label("HCR in situ", x = 0.3, y = 0.99, size = 10) +
+  draw_label("c-opsin1", x = 0.3, y = 0.85, color='green',size = 12, fontface='bold')  +
+  draw_line(x = c(0.05, 0.407), y = c(0.1, 0.1), color = "white", size = 0.5) +
+  draw_label(expression(paste("20 ", mu, "m")), x = 0.23, y = 0.13, color = "white", size = 10)
+
[…]

### 2022-09-30 code/Fig3_sup3.R

@@ -12,22 +12,43 @@ library(patchwork)
 library(magick)
 
 # assemble figure ---------------------------------------------------------
-panel_3D <- ggdraw() + draw_image(readPNG("pictures/NITdomain_3D.png"))
+panel_NIT1 <- ggdraw() + draw_image(readPNG("pictures/IHC_55_NIT1_NOS_acTub_AP_1_58.87um.png"))+
+  draw_label("IHC", x = 0.3, y = 0.99, size = 10) +
+  draw_label("NIT1", x = 0.55, y = 0.67, color='magenta', size = 12, fontface='bold') +
+  draw_label("ac-tub", x = 0.3, y = 0.85, color='blue',size = 12, fontface='bold') +
+  draw_line(x = c(0.05, 0.407), y = c(0.1, 0.1), color = "white", size = 0.5) +
+  draw_label(expression(paste("20 ", mu, "m")), x = 0.23, y = 0.13, color = "white", size = 10)
+
+panel_NOS <- ggdraw() + draw_image(readPNG("pictures/IHC_55_NIT1_NOS_acTub_AP_2_58.87um.png"))+
+  draw_label("IHC", x = 0.3, y = 0.99, size = 10) +
+  draw_label("NOS", x = 0.7, y = 0.45, color='green',size = 12, fontface='bold') +
+  draw_label("ac-tub", x = 0.3, y = 0.85, color='blue',size = 12, fontface='bold') +
+  draw_line(x = c(0.05, 0.407), y = c(0.1, 0.1), color = "white", size = 0.5) +
+  draw_label(expression(paste("20 ", mu, "m")), x = 0.23, y = 0.13, color = "white", size = 10)
+
+panel_NIT1_NOS <- ggdraw() + draw_image(readPNG("pictures/IHC_55_NIT1_NOS_acTub_AP_3_58.87um.png"))+
+  draw_label("IHC", x = 0.3, y = 0.99, size = 10) +
+  draw_label("NIT1", x = 0.55, y = 0.67, color='magenta', size = 12, fontface='bold') +
+  draw_label("NOS", x = 0.7, y = 0.45, color='green',size = 12, fontface='bold') +
+  draw_label("ac-tub", x = 0.3, y = 0.85, color='blue',size = 12, fontface='bold') +
+  draw_line(x = c(0.05, 0.407), y = c(0.1, 0.1), color = "white", size = 0.5) +
+  draw_label(expression(paste("20 ", mu, "m")), x = 0.23, y = 0.13, color = "white", size = 10)
+
 
[…]

### 2022-09-30 text/Manuscript.Rmd
Put supplement

@@
 Avoidance behaviour to escape harmful UV radiation during the day is closely linked to the control of vertical migration in the ocean.
-In metamorphosis, a life history strategy widely found in metazoans, downward swimming towards the seafloor is known to be associated with larval settlement and the initiation of metamorphosis, a dramatic body change from the larval stage, and UV-avoidance behaviour allows dispersal and the acquisition of different ecological niches.
 However, little is known about the molecular neural circuit mechanisms in UV avoidance behaviour in many animals, including zooplankton.
@@
 We identified a single *nitric oxide synthase* (*NOS*) gene in the *Platynereis dumerilii* genome and transcriptome data.
-Phylogenetic analysis of NOS proteins indicate that the *Platynereis* NOS belongs to an orthology group of metazoan NOS sequences (Fig.
-1---figure supplement 1).
+Phylogenetic analysis of NOS proteins indicate that the *Platynereis* NOS belongs to an orthology group of metazoan NOS sequences (Figure 1---figure supplement 1).
 To characterise the expression pattern of *NOS* we used in situ hybridization chain reaction (HCR) and transient transgenesis.
-In two-day-old larvae, we detected *NOS* expression in two cells in the apical organ region (Fig.
-1A) <!---# 3d HCR, Kei will put them -->.
-The same two cells cells were labelled with a reporter construct driving palmitoylated-tdTomato by a 12 kbp upstream genomic region of the *NOS* gene.
+In 2 and 3-day-old larvae, we detected *NOS* expression in two or four cells in the apical organ region (Figure 1A-D).
+The same four cells were labelled with a reporter construct driving palmitoylated-tdTomato by a 12 kbp upstream genomic region of the *NOS* gene in 3-day-old larvae(Figure 1E).
 This reporter also revealed the axonal projections of the NOS-expressing neurons.
-In three-day-old larvae...
-<!---# need 3d HCR , Kei will put them -->
-
-The position and morphology of the four *NOS*-expressing cells allowed us to identify the same cells in the three-day-old whole-body volume EM data [@williams2017; @veraszto2020] as four interneurons (INNOS)(Figure 1D-E).
-The INNOS cells are postsynaptic to the UV-sensory cPRCs and presynaptic to the INRGW interneurons, which are also cPRC targets (Fig.
-1F).
+The position and morphology of the four *NOS*-expressing cells allowed us to identify the same cells in the 3-day-old whole-body volume EM data [@williams2017; @veraszto2020] as four interneurons (INNOS)(Figure 1F and G).
+The INNOS cells are postsynaptic to the UV-sensory cPRCs and presynaptic to the INRGW interneurons, which are also cPRC targets (Figure 1H).
 
 ![**Figure 1.
-Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** (A) SEM image of a 72 hpf Platynereis larva.(B-D) Expression analysis of the NOS gene (magenta) using in situ HCR.
-Co-stained image using acetylated α-tubulin antibody (ac α-tub: green).
-Anterior view of the larva at 54 and 72 hpf.
+Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** (A) SEM image of a 3-day-old Platynereis larva.(B-D) Expression analysis of the NOS gene (magenta) using in situ HCR.
+(D) Co-stained image using acetylated α-tubulin antibody (ac α-tub: green).
+(B) Dorsal view of the larva at 3-day-old.
+(C and D) Anterior view of the larva at 2 (D) and 3-day-old (C).
 (E) Immunostaining with anti-HA antibody against HA tagged palmitoylated-tdTomato expressed under the upstream of the start site of NOS (magenta).
@@
 INNOSs with the entire cell body and axon labelled are shown with white arrowheads and with only the cell body labelled are shown with black arrowheads.
-Views of larvae at 72 hpf from the anterior side.
-(F and G) Five types of neurons (cPRC, INNOS, INRGWa, Ser-h1 and MC) in the cPRC circuit reconstructed in the whole-body transmission electron microscopy (ssTEM) dataset of a 72 hpf larva.
+Views of larvae at 3-day-old from the anterior side.
+(F and G) Five types of neurons (cPRC, INNOS, INRGWa, Ser-h1 and MC) in the cPRC circuit reconstructed in the whole-body transmission electron microscopy (ssTEM) dataset of a 3-day-old larva.
 The INNOS on the ventral side labelled with (E) are shown in purple.
@@
 We need to edit this a bit more.
---> Hexagons represent neuronal groups (numbers indicated in square brackets), arrows represent synaptic connections (numbers indicated in number of synapses); NOS and RGW indicate INNOS and INRGW-dc1 respectively.
-Ciliomotor neuron shows two serotonergic neurons (Ser-h1).](../figures/Fig1.png){width="100%"}
+--> ](../figures/Fig1.png){width="100%"}
 
@@
 To test this, we generated two *Platynereis* *NOS* knockout lines with the CRISPR/Cas9 method.
-We recovered two deletions (NOSΔ11/Δ11 and Δ23/Δ23), both leading to an early stop codon and thus likely representing null alleles (Figure 3A).
+We recovered two deletions (NOSΔ11/Δ11 and Δ23/Δ23), both leading to an early stop codon and thus likely representing null alleles (Figure 2A).
 We could establish a homozygous line for both mutations indicating that NOS is not an essential gene.
-To quantify UV avoidance, we recorded the trajectories of freely swimming larvae in vertical columns, illuminated laterally from two opposite sides with xy nm violet light.
-As previously shown, wild-type larvae swim downward within 30 sec following UV/violet light stimulation.
-In contrast, homozygous *NOS*-mutant larvae showed a strongly diminished UV-avoidance response (Figure 3B).
-This phenotype is similar to the defective UV-avoidance of *c-opsin1* mutant larvae and reveals a requirement for NOS in UV-avoidance behaviour.
+To quantify UV avoidance, we recorded the trajectories of freely swimming larvae in vertical columns, illuminated laterally from two opposite sides with 395 nm violet light.
+As previously shown, wild-type larvae swim downward within 30 sec following UV/violet light stimulation [@veraszto2018].
+In contrast, homozygous *NOS*-mutant larvae showed a strongly diminished UV-avoidance response (Figure 2B and C).
+This phenotype is similar to the defective UV-avoidance of *c-opsin1* mutant larvae and reveals a requirement for NOS in UV-avoidance behaviour [@veraszto2018].
+
+To analyse the behaviour of larvae before and after UV stimulation in more detail, the temporal changes in some parameters were compared using tracking of individual larvae of the WT and mutant (Figure 2D).
+The results showed that in the WT there was a rapid and strong increase in downward swimming approximately 15 s after UV stimulation, whereas this did not occur in the mutant (Figure 2E).
+Interestingly, there was already a difference in swimming speed between the WT and the mutant before UV stimulation (in the dark) (Figure 2F).
+After UV stimulation, the swimming speed of the WT increased and reached a maximum after about 15 s.
+In contrast, the swimming speed of the mutant did not increase after about 15 s.
+In contrast, the mutants showed a slight increase in speed after UV stimulation, but swimming speed was not maintained and returned to the speed before UV stimulation.
+To investigate the role of NO on UV avoidance behaviour, receptors for NO were explored.
 
@@
 Yellow shows PAM sequences, red shows stop codons.
-(B and C) Vertical swimming in wild type and mutant larvae at 2-day-old (B) and 3-day-old (C) stimulated with blue (480 nm) and UV (395 nm) light from above.
-the data are shown in 30 s bins.](../figures/Fig2.png){alt="Figure 2.
-Larvae with inhibited NOS have shorter ciliary arrest times and a delayed response to UV-avoidance.
-(A) Percentage of time in which cilia are closed in DMSO-treated larvae (control) vs.
-larvae exposed to NOS inhibitor.
-(B) Top: The domain structure of NOS.
-Bottom: Close-up region showing the genomic locus of NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9.
-The generated mutants (NOSΔ11, NOSΔ23) are also shown.
-Pink indicates target sites.
-Yellow shows PAM sequences, red shows stop codons.
-(C and D) Vertical swimming in wild type and mutant larvae at 2-day-old (C) and 3-day-old (D) stimulated with blue (480 nm) and UV (395 nm) light from above.
-the data are shown in 30 s bins." width="100%"}
+(B and C) Vertical swimming in wild type and mutant larvae at 2-day-old (B) and 3-day-old (C) stimulated with blue (488 nm) and UV (395 nm) light from above.
+The data are shown in 30 s bins.
+(D) Overlaid trajectories for WT (left, n=20) and NOS mutant (right, n=31) 3-day-old larvae.
+Grey shows the trajectories before UV stimulation and purple shows the trajectories after UV stimulation.
+(E) The temporal changes in the vertical position of the WT and mutant larvae before and after UV stimulation are shown.
+The starting points of each larval trajectory are set to 0.
+(F) The temporal changes in the distance traveled of the WT and mutant larvae before and after UV stimulation are shown.
+(E and F) After UV stimulation is indicated by purple squares.](../figures/Fig2.png){width="100%"}
 
@@
 However, no NO targets were found in cells that highly express sGC within the cPRC circuit.
-Recently, Moroz and others reported that GC with NIT (nitrite/nitrate sensing) domains (PF08376) (NIT-GC), identified as a sensor for nitrate and nitrite, the oxides of NO, is relatively widely conserved in metazoans (Moroz et al., 2020).
+Recently, Moroz and others reported that GC with NIT (nitrite/nitrate sensing) domains (PF08376) (NIT-GC), identified as a sensor for nitrate and nitrite, the oxides of NO, is relatively widely conserved in metazoans [@Moroz2020].
 To search for NIT-GCs in Platynereis, genes with the CYC domains were searched using the transcriptome and genome sequences of 45 metazoan species and choanoflagellate.
 By sequence-similarity searches, multiple sequence alignments and cluster analyses, a cluster of membrane-bound (guanylate cyclase-coupled receptor) type was found and a phylogenetic tree was constructed with soluble type (sGC) as an outgroup.
-As in the phylogenetic analysis of Moroz et al, we found a group of GCs with NIT domains that included many metazoans such as placozoa, cnidaria, ecdysozoa, echinoderms and lophotrochozoa, and in Platynereis 15 NIT-GCs were found (Fig.
-3A).
-To find the candidate NIT-GCs involved in the cPRC circuit among the 15 NIT-GCs, we used previously published spatially mapped single-cell transcriptome data [@Achim2015][@williams2017] and Two NIT-GCs were found to be highly expressed in the cPRC, and we analysed their expression using in situ HCR as NIT-GC1 and NIT-GC2.
-As a result, strong signals for both were observed specifically in the vicinity of the cPRCs in 2-day-old larvae (Fig.
-3B).
-Both c-opsin1 and MLD pNP (pedal peptide neuropeptide precursor 2), already identified as cPRC markers, and NIT-GC1 and NIT-GC2 were found to be co-expressed.
-They were co-expressed c-opsin1 and MLD pNP (pedal peptide neuropeptide precursor 2), already identified as cPRC markers (Fig.
-3C) (Arendt et al., 2004; Williams et al., 2017).
+As in the phylogenetic analysis of Moroz et al, we found a group of GCs with NIT domains that included many metazoans such as placozoa, cnidaria, ecdysozoa, echinoderms and lophotrochozoa, and in Platynereis 15 NIT-GCs were found (Figure 3---figure supplement 1).
+Using previously published spatially mapped single-cell transcriptome data  [@Achim2015; @williams2017] to find candidate NIT-GCs involved in the cPRC circuit among 15 NIT-GCs, two NIT-GCs (NIT-GC1 and NIT-GC2) were found to be highly expressed in the cPRC (Figure 3---figure supplement 1).
+Expression analysis with c-opsin1 and MLD pNPs (pedal peptide neuropeptide precursor 2) [@Arendt2004; @williams2017], which have already been identified as cPRC markers using in situ HCR, showed that both NIT-GC1 and NIT-GC2 are strongly and specifically expressed in cPRC (Figure 3A and B, Figure 3---figure supplement 2).
 To analyse localisations at the protein level, immunostaining was performed using polyclonal antibodies against NIT-GC1 and NIT-GC2.
-Interestingly, NIT-GC1 was found to localise to the neuronal axon and NIT-GC2 to the ciliary region of the cPRC (Fig.
-3D and E).
-Furthermore, co-immunostaining with NOS antibodies revealed that NOS and NIT-GC1 were localised in close proximity in the neurosecretory plexus (Fig.
-3F).
+Interestingly, NIT-GC1 was found to localise to the neuronal axon and NIT-GC2 to the ciliary region of the cPRC (Figure 3C and D).
+Furthermore, co-immunostaining with NOS antibodies revealed that NOS and NIT-GC1 were localised in close proximity in the neurosecretory plexus (Figure 3---figure supplement 3).
 These results suggest that NIT-GC1 may function as a major target for NO.
@@
 NIT-GC1 has a highly conserved cyclase domain and activation of NIT-GC1 is expected to produce cGMP.
-Recently, a synthetic protein called Green cGull was developed as a cGMP indicator that can be used in cultured cell systems (Matsuda et al., 2017).
-We co-expressed Green cGull and NIT-GC1 in cultured cells COS-7 (monkey kidney) and tested the intracellular dynamics of cGMP in the activation of NIT-GC1 (Fig.
-4A).
-First, S-nitroso-N-acetyl-DL-penicillamine (SNAP; NO donor), 8-Br-cGMP (8-Bromoguanosine 3',5'-cyclic monophosphate; membrane-permeable cGMP analogue, positive control) and DMSO (negative control) were applied to COS-7 cells expressing Green cGull only.
-8-Br-cGMP increased fluorescence intensity by more than sixfold, while little changes were detected for SNAP and DMSO (Fig.
-S4A).
-Next, the two kind of genes, Green cGull and NIT-GC1, were co-expressed using the 2A self-cleaving peptide and the same solutions were applied.
-As a result, a clear increase in cGMP fluorescence intensity was detected in SNAP compared to DMSO or expression of Green cGull only, although not as high as the increase in 8-Br-cGMP (Fig.
-4B, C).
-Finally, expression of a mutant form of NIT-GC1 with deletion of the NIT domain significantly suppressed the increase in cGMP (Note, we need re-experiment!, Fig.
-4D).
+Recently, a synthetic protein called Green cGull was developed as a cGMP indicator that can be used in cultured cell systems [@Matsuda2017].
+We co-expressed Green cGull and NIT-GC1 using the 2A self-cleaving peptide in cultured cells COS-7 (monkey kidney) and tested the intracellular dynamics of cGMP in the activation of NIT-GC1 (Figure 3---figure supplement 4).
+When the NO donor SNAP was applied to the co-expression cells, the fluorescence intensity in the cells was significantly increased compared to when NIT-GC1 was not expressed (Figure 3E and F).
+The increase in fluorescence intensity also did not occur when DMSO was applied (Figure 3G).
+To confirm whether this increase in fluorescence intensity is NIT domain-dependent response of NIT-GC1, mutant form of NIT-GC1 without the NIT domain was expressed and SNAP was applied (Figure 3---figure supplement 4A).
+As a result, the increase in fluorescence intensity was significantly reduced  (Figure 3H).
 These results indicate that NIT-GC1 is activated by NO to produce cGMP.

### 2022-10-03 text/Manuscript.Rmd
Figure update

@@
 We need to edit this a bit more.
---> ](../figures/Fig1.png){width="100%"}
+-->](../figures/Fig1.png){width="100%"}
 
@@
 Furthermore, co-immunostaining with NOS antibodies revealed that NOS and NIT-GC1 were localised in close proximity in the neurosecretory plexus (Figure 3---figure supplement 3).
-These results suggest that NIT-GC1 may function as a major target for NO.
+These two localisation differences suggest the existence of different intracellular signalling mechanisms in the ciliary and neuroaxonal regions within the cPRC.
+
+**NIT-GC1 is required for the depolarising response of cPRC to UV stimulation**
+
+To characterise the role of NIT-GCs within the cPRC during UV stimulation in more detail, the calcium sensor GCaMP6s were ubiquitously expressed in larvae in which NIT-GCs were knocked down using morpholino to examine changes in calcium levels in the cPRC.
+Local stimulation of the ciliated part of the cPRC with 405 nm light causes a transient decrease in calcium levels in the cPRC (hyperpolarisation), followed by an increase in calcium levels (depolarisation) when the light is switched off [@veraszto2018].
+In NIT-GC1 morphant larvae, 405 nm light stimulation caused a decrease in calcium levels in the cPRC, but the increase in calcium levels that occurred after the light stimulation was stopped was strongly suppressed (Figure 3E).
+This result suggests that NIT-GC1 mediates depolarising responses after UV stimulation of cPRC and may be involved in extracellular signalling.
+Interestingly, in NIT-GC2 morphant larvae, an increase in calcium levels occurred immediately after light stimulation (Figure 3F).
+The different responses of cPRC upon knockdown of NIT-GC1 and NIT-GC2 suggest that there may be different intracellular signalling mechanisms associated with their respective localisations.
+We focused our analysis on NIT-GC1 because it was localised to cPRC neuroaxons within the neuroplexas region and is involved in extracellular signalling after UV stimulation.
+It is generally known that depolarisation in the PRC involves elevated cGMPs in the photoreceptors.
+We therefore used an in vitro system in which cultured cells expressed NIT-GC1 to assess the cGMP activity of NIT-GC1.
 
@@
 We co-expressed Green cGull and NIT-GC1 using the 2A self-cleaving peptide in cultured cells COS-7 (monkey kidney) and tested the intracellular dynamics of cGMP in the activation of NIT-GC1 (Figure 3---figure supplement 4).
-When the NO donor SNAP was applied to the co-expression cells, the fluorescence intensity in the cells was significantly increased compared to when NIT-GC1 was not expressed (Figure 3E and F).
-The increase in fluorescence intensity also did not occur when DMSO was applied (Figure 3G).
-To confirm whether this increase in fluorescence intensity is NIT domain-dependent response of NIT-GC1, mutant form of NIT-GC1 without the NIT domain was expressed and SNAP was applied (Figure 3---figure supplement 4A).
-As a result, the increase in fluorescence intensity was significantly reduced  (Figure 3H).
-These results indicate that NIT-GC1 is activated by NO to produce cGMP.
-They also show that the NIT domain is required for the activation of NIT-GC1.
+When the NO donor SNAP was applied to the NIT-GC1 and Green cGull co-expression cells, the intracellular cGMP levels was significantly increased compared to when NIT-GC1 was not expressed (Figure 3G and H).
+The increase in fluorescence intensity also did not occur when DMSO was applied (Figure 3I).
+To confirm whether this increase in intracellular cGMP levels is NIT domain-dependent response of NIT-GC1, mutant form of NIT-GC1 without the NIT domain was expressed and SNAP was applied (Figure 3---figure supplement 4A).
+As a result, the increase in fluorescence intensity was significantly reduced (Figure 3J).
+These results indicate that NIT-GC1 produces cGMP via the NIT domain.
+In bacteria, it has been proposed that the NIT domain regulates cellular functions in response to changes in extracellular and intracellular nitrate and nitrite concentrations.
+Our results suggest that NIT-GC1 may be sensitive to NO.
+This could be a possible feedback pathway in which NIT-GC1 expressed in cPRC is activated by NO produced by IN-NOS.
+In order to further explore the role of NO in the cPRC circuit, we therefore compared calcium imaging of postsynaptic cells (IN-NOS, IN-RGW, Ser-h1, MC) in the cPRC circuit during UV stimulation of cPRC between WT and NOS mutant larvae.
 
@@
 
-To test if NO is produced during UV/violet responses, we injected zygotes with the fluorescent NO-reporter DAF-FM and imaged two-day-old larvae during the local stimulation of the cPRC cilia by 405 nm light.
+To test if NO is produced during UV/violet responses, we injected zygotes with the fluorescent NO-reporter DAF-FM and imaged two-day-old larvae during the local stimulation of the cPRC cilia by 405 nm light (Figure 4---figure supplement 1).
 Following light stimulation, we detected an increase in DAF-FM fluorescence in the region of the INNOS cells.
 We next analysed the dynamics of the cPRC circuit following light stimulation.
-As we have shown previously [@Veraszto2018], a 30-sec illumination of the cPRC ciliary region lead to a decrease in cPRC calcium levels, likely due to the hyperpolarisation of the cPRCs.
-After 30 min, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus - a response that may involve depolarisation.
+As we have shown previously [@Veraszto2018], a 20-sec illumination of the cPRC ciliary region lead to a decrease in cPRC calcium levels, likely due to the hyperpolarisation of the cPRCs.
+After 20 sec, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus - a response that may involve depolarisation.
 This activation phase occurs after the end of the stimulation and may be due to a slower neuroendocrine feedback [@Veraszto2018].
@@
 We used the cell-specific neuropeptide markers anti-RYamide (INNOS) and anti-RGWa (INRGW) [@Conzelmann2011] to mark the interneurons.
-Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision (Figure 4---figure supplement 1).
+Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision (Figure 4---figure supplement 2).
 In both wild type and *NOS*-mutant larvae, INNOS cells showed increase calcium signals during the inhibition phase of the cPRCs.
@@
 This suggests that there is no circuit output to the Ser-h1 cells and the ciliary band in *NOS*-mutants, consistent with the observed behavioural defects.
-
-**NO induces depolarisation of the cPRCs via NIT-GC1**
-
-To determine the function of NO in the cPRC circuit, the intracellular dynamics of neurons in the cPRC circuit were investigated, focusing on NIT-GC1.
-First, we investigated the changes in NO during UV stimulation of the cPRC using DAF-FM, a fluorescent probe that can detect NO, in order to determine whether NO is produced by UV light.
-By injecting DAF-FM directly into zygotes, INNOS cells could be identified under live imaging.
-To visualise the location of cPRCs, the calcium indicator RGECO1a was also injected together, as cPRCs maintain a higher resting calcium levels compared to other cells.
-Direct UV stimulation to the cPRCs increased the fluorescence intensity of DAF-FM compared to stimulation of non-cPRC region (Fig.
-5A).
-Next, we did calcium imaging of cells in the cPRC circuit during UV stimulation.
-We locally stimulated the cPRC cilia in wild-type with a 405 nm laser (UV) and detected a slight increase followed by a strong decrease in calcium levels (hyperpolarisation) (Fig.
-5B).
-Turning off the UV after or during the decrease caused a strong increase (depolarisation) and a return to the same calcium level as before (Fig.
-5B).
-These responses were similar to those reported previously [@Veraszto2018].
-We performed a similar stimulation using NOS mutants, which hyperpolarised during the UV stimulation, but no depolarising responses were observed when UV was stopped (Fig.
-5C).
-Similarly, in NIT-GC1 knockdown larvae using the morpholinos, no depolarisation after the hyperpolarisation response were observed (Fig.
-5D).
-Interestingly, no hyperpolarising response during UV stimulation was found to occur in NIT-GC2 knockdown larvae.
-And we compared the changes of the INNOS calcium levels identified by their position in relation to cPRCs during UV stimulation in wild-type and in NOS mutants.
-In both cases, a depolarising response was observed at the same time as the hyperpolarising response of the cPRCs (Fig.
-5E).
-Furthermore, to perform calcium imaging of INRGWs in UV stimulation, INRGWs were identified by immunostaining with RGWa antibodies after calcium imaging (Fig.
-5F).
-It has been reported that the INRGW cells form many large synapses with synaptic vesicles on two serotonergic ciliomotor neurons (Ser-h1) that are known to play a role in the regulation of ciliary beating and the activity of the INRGW cells increased the negative correlation between the activity patterns of serotonergic neurons and the MC neuron (Williams et al., 2018).
-The results showed that depolarisation of INNOS occurs at the same time as hyperpolarisation of cPRC in the wild type.
-Identified INRGW cells were found to cause depolarisation at the same time as cPRC hyperpolarisation in the wild type (Fig.
-5G).
-It was also found that in NOS mutants where cPRC hyperpolarisation does not occur, the INRGW depolarisation response also does not occur (Fig.
-5H).
 
@@
 (G) Comparison of changes over time in GCaMP6s intensities in cPRC (purple) and INRGW (green) when the cPRC are stimulated with UV in WT (left) and NOS mutant (NOSΔ23: right) larvae.
-(H)(I)](../figures/Fig4.png){width="100%"}
+(H)(I)(J)](../figures/Fig4.png){width="100%"}
 

### 2022-10-03 code/Fig3.R

@@ -243,15 +243,15 @@ panel_GcG_mutNIT_SNAP <- ggdraw() + draw_image(readPNG("pictures/GcG-mutNIT-GC1-
 layout <- "
 A#B#C#D
 #######
-E#F#G#H
+EEE#FFF
 #######
-III#JJJ
+G#H#I#J
 "
 
-Fig3 <- panel_HCR_NIT1 + panel_HCR_NIT1_cOps + panel_IHC_NIT1 + panel_IHC_NIT2 + 
-  panel_GcG_NIT_SNAP + panel_GcG_SNAP + panel_GcG_NIT_DMSO + panel_GcG_mutNIT_SNAP +
+Fig3 <- panel_HCR_NIT1 + panel_HCR_NIT1_cOps + panel_IHC_NIT1 + panel_IHC_NIT2 +
   panel_cPRC_NIT1_MO + panel_cPRC_NIT2_MO +
-  patchwork::plot_layout(design = layout, heights = c(0.8, 0.02, 0.8, 0.02, 0.6),
+  panel_GcG_NIT_SNAP + panel_GcG_SNAP + panel_GcG_NIT_DMSO + panel_GcG_mutNIT_SNAP +
+  patchwork::plot_layout(design = layout, heights = c(0.9, 0.02, 0.7, 0.02, 0.85),
                          widths = c(1,0.02,1,0.02,1,0.02,1)) + #we can change the heights of the rows in our layout (widths also can be defined)
   patchwork::plot_annotation(tag_levels = 'A') +  #we can change this to 'a' for small caps or 'i' or '1'
   ggplot2::theme(plot.tag = element_text(size = 12,face='bold')) #or 'bold', 'plain', 'italic'

### 2022-10-03 code/Fig3_sup2.R

@@ -12,7 +12,19 @@ library(patchwork)
 library(magick)
 
 # assemble figure ---------------------------------------------------------
-panel_MLD <- ggdraw() + draw_image(readPNG("pictures/HCR_51_DV_NITGC1_MLD_DAPI_112.55µm.png"))+
+panel_NIT1 <- ggdraw() + draw_image(readPNG("pictures/HCR_51_DV_NITGC1_MLD_DAPI_1_112.55µm.png"))+
+  draw_label("HCR in situ", x = 0.3, y = 0.99, size = 10) +
+  draw_label("NIT1", x = 0.55, y = 0.67, color='magenta', size = 12, fontface='bold') +
+  draw_line(x = c(0.05, 0.407), y = c(0.1, 0.1), color = "white", size = 0.5) +
+  draw_label(expression(paste("20 ", mu, "m")), x = 0.23, y = 0.13, color = "white", size = 10)
+
+panel_MLD <- ggdraw() + draw_image(readPNG("pictures/HCR_51_DV_NITGC1_MLD_DAPI_2_112.55µm.png"))+
+  draw_label("HCR in situ", x = 0.3, y = 0.99, size = 10) +
+  draw_label("MLD", x = 0.3, y = 0.85, color='green',size = 12, fontface='bold')  +
+  draw_line(x = c(0.05, 0.407), y = c(0.1, 0.1), color = "white", size = 0.5) +
+  draw_label(expression(paste("20 ", mu, "m")), x = 0.23, y = 0.13, color = "white", size = 10)
+
+panel_NIT1_MLD <- ggdraw() + draw_image(readPNG("pictures/HCR_51_DV_NITGC1_MLD_DAPI_112.55µm.png"))+
   draw_label("HCR in situ", x = 0.3, y = 0.99, size = 10) +
   draw_label("NIT1", x = 0.55, y = 0.67, color='magenta', size = 12, fontface='bold') +
   draw_label("MLD", x = 0.3, y = 0.85, color='green',size = 12, fontface='bold')  +
@@ -42,17 +54,19 @@ panel_NIT2_cop1 <- ggdraw() + draw_image(readPNG("pictures/2d_647-NIT-GC2_546-c-
 #combine panels into Figure and save final figure as pdf and png
 #panels of different sizes
 layout <- "
-A#B#C#D
+A#B#C
+#####
+D#E#F
 "
 
-Fig3_sup2 <- panel_MLD + panel_NIT2 + panel_cop1 + panel_NIT2_cop1 +
[…]

### 2022-10-03 code/Fig4.R

@@ -89,12 +89,10 @@ DAFFM
 min(DAFFM$intensity)
 
 DAFFM %>%
-  ggplot(aes(x=frame,y=intensity,color=stimuli)) +
-  geom_smooth(level = 0.99, size = 1, span = 0.02, 
-              method = "loess") +
+  ggplot(aes(x=frame,y=intensity,color = stimuli)) +
+  annotate("rect", xmin=51, xmax=90, ymin=-Inf, ymax=Inf, alpha=0.1, fill="blue") +
   geom_line(aes(group = sample), size=0.2, alpha=0.2) +
-  annotate("rect", xmin=51, xmax=90, ymin=-Inf, ymax=Inf, 
-           alpha=0.1, fill="blue") +
+  geom_smooth(aes(fill = stimuli), level = 0.95, size = 1, span = 0.05, method = "loess") +
   annotate("segment", x=20, xend=40, y=1.2, yend=1.2, size=1)+
   annotate("segment", x=20, xend=20, y=1.2, yend=1.3, size=1)+
   annotate("text", x=29, y=1.15, label="10 sec", size=3)+
@@ -105,9 +103,10 @@ DAFFM %>%
   theme_void()  +
   theme(legend.title = element_blank(), legend.text.align=0,
         legend.text = element_text(size=12)) +
-  #Specify colours and legend labels
-  scale_color_manual(values=c(Okabe_Ito[6], Okabe_Ito[2]),
-                     labels = c(expression('cPRC stim.'), "ctr stim." ))
+  scale_fill_manual(values=c(Okabe_Ito[3], dark2[8]),
+                    labels = c("cPRC stim.", "ctr stim."))+
+  scale_color_manual(values=c(Okabe_Ito[3], dark2[8]),
+                     labels = c("cPRC stim.", "ctr stim."))
 
 # save plot ---------------------------------------------------------------
 
@@ -116,13 +115,16 @@ ggsave("pictures/DAFFM.png", limitsize = FALSE,
 
 # plot NOS23 cPRC data with geom_smooth ----------------------------------------------
 
+WT_cPRC_tidy$genotype <- factor(WT_cPRC_tidy$genotype, 
+                                      levels=c("WT_cPRC", "NOS23_cPRC"))
+
 WT_cPRC_tidy %>%
   ggplot(aes(x=frame,y=intensity,color=genotype)) +
-  geom_smooth(level = 0.99, size = 1, span = 0.02, 
[…]

### 2022-10-03 code/Fig4_sup1.R

@@ -5,29 +5,38 @@
 source("code/Packages_Catmaid_connection.R")
 
 # assemble figure ---------------------------------------------------------
-panel_A <- ggdraw() + draw_image(readPNG("pictures/55hpf_DAF-FM_134.95um.png")) +
+panel_DAF <- ggdraw() + draw_image(readPNG("pictures/55hpf_DAF-FM_134.95um.png")) +
+  draw_label("DAF-FM", x = 0.15, y = 0.9, color='green',size = 12, fontface='bold') +
   draw_line(x = c(0.05, 0.42), y = c(0.1, 0.1), color = "white", size = 0.5) +
   draw_label(expression(paste("50 ", mu, "m")), x = 0.23, y = 0.13, color = "white", size = 10)
 
-panel_B <- ggdraw() + draw_image(readPNG("pictures/55hpf_DAF-FM_RGECO_134.95um.png"))
+panel_DAF_RGECO <- ggdraw() + draw_image(readPNG("pictures/55hpf_DAF-FM_RGECO_134.95um.png")) +
+  draw_label("RGECO1a", x = 0.5, y = 0.9, color='magenta', size = 12, fontface='bold') +
+  draw_label("DAF-FM", x = 0.15, y = 0.9, color='green',size = 12, fontface='bold') +
+  draw_line(x = c(0.05, 0.42), y = c(0.1, 0.1), color = "white", size = 0.5) +
+  draw_label(expression(paste("50 ", mu, "m")), x = 0.23, y = 0.13, color = "white", size = 10)
 
-panel_C <- ggdraw() + draw_image(readPNG("pictures/55hpf_DAF-FM_RGECO_highmag_56.04um.png")) +
+panel_enlarge <- ggdraw() + draw_image(readPNG("pictures/55hpf_DAF-FM_RGECO_highmag_56.04um.png")) +
+  draw_label("RGECO1a", x = 0.5, y = 0.9, color='magenta', size = 12, fontface='bold') +
+  draw_label("cPRC", x = 0.7, y = 0.67, color='magenta', size = 12, fontface='bold') +
+  draw_label("DAF-FM", x = 0.15, y = 0.9, color='green',size = 12, fontface='bold') +
+  draw_label("IN-NOS", x = 0.7, y = 0.45, color='white',size = 12, fontface='bold') +
   draw_line(x = c(0.05, 0.407), y = c(0.1, 0.1), color = "white", size = 0.5) +
[…]

### 2022-10-03 code/Fig4_sup2.R

@@ -5,22 +5,39 @@
 source("code/Packages_Catmaid_connection.R")
 
 # assemble figure ---------------------------------------------------------
-panel_A <- ggdraw() + draw_image(readPNG("pictures/Diagram of agarose embedding immunostaining.png"))
-panel_B <- ggdraw() + draw_image(readPNG("pictures/n54hpf_WT_e-1_GCaMP6s.png"))
-panel_C <- ggdraw() + draw_image(readPNG("pictures/n54hpf_WT_e-1_RGWa.png"))
-panel_D <- ggdraw() + draw_image(readPNG("pictures/n52hpf_WT_a-2_ROI60_PD1.54_power1_pin216.9_frame_201-241_401-441_co_cPRC.png"))
-panel_E <- ggdraw() + draw_image(readPNG("pictures/n52hpf_WT_a-2_ROI60_PD1.54_power1_pin216.9_frame_201-241_401-441_co_NOS.png"))
+panel_method <- ggdraw() + draw_image(readPNG("pictures/Diagram of agarose embedding immunostaining.png"))
+
+panel_calcium <- ggdraw() + draw_image(readPNG("pictures/n54hpf_WT_e-1_GCaMP6s.png"))+
+  draw_label("calcium imaging", x = 0.3, y = 0.99, size = 10) +
+  draw_label("GCaMP6s", x = 0.18, y = 0.9, color='white', size = 12, fontface='bold') +
+  draw_line(x = c(0.5, 0.95), y = c(0.1, 0.1), color = "white", size = 0.5) +
+  draw_label(expression(paste("20 ", mu, "m")), x = 0.725, y = 0.13, color = "white", size = 10)
+
+panel_immunostain <- ggdraw() + draw_image(readPNG("pictures/n54hpf_WT_e-1_RGWa.png"))+
+  draw_label("confocal imaging", x = 0.3, y = 0.99, size = 10) +
+  draw_label("nucleus", x = 0.15, y = 0.9, color='cyan', size = 12, fontface='bold') +
+  draw_label("postsynaptic cells", x = 0.31, y = 0.83, color='red',size = 12, fontface='bold') +
+  draw_label("Ser-h1", x = 0.9, y = 0.75, color='white', size = 9, fontface='plain') +
+  draw_label("IN-RGW", x = 0.35, y = 0.35, color='white',size = 9, fontface='plain') +
+  draw_line(x = c(0.5, 0.95), y = c(0.1, 0.1), color = "white", size = 0.5) +
[…]

### 2022-10-03 code/Packages_to_load.R

@@ -35,5 +35,8 @@ Okabe_Ito <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00",
 pie(rep(1,8), col=Okabe_Ito, Okabe_Ito, main='Okabe Ito')
 blues <- brewer.pal(9, 'Blues')
 bluepurple <- brewer.pal(9, 'BuPu')
-oranges <- brewer.pal(9, 'YlOrRd')
+yelloworangered <- brewer.pal(9, 'YlOrRd')
+set2 <- brewer.pal(8, 'Set2')
+paired <- brewer.pal(12, 'Paired')
+dark2 <- brewer.pal(8, 'Dark2')
 

### 2022-10-03 text/Manuscript.Rmd
Add correlation data

@@
 To test if NO is produced during UV/violet responses, we injected zygotes with the fluorescent NO-reporter DAF-FM and imaged two-day-old larvae during the local stimulation of the cPRC cilia by 405 nm light (Figure 4---figure supplement 1).
-Following light stimulation, we detected an increase in DAF-FM fluorescence in the region of the INNOS cells.
+Following light stimulation, we detected an increase in DAF-FM fluorescence in the region of the INNOS cells (Figure 4A).
 We next analysed the dynamics of the cPRC circuit following light stimulation.
 As we have shown previously [@Veraszto2018], a 20-sec illumination of the cPRC ciliary region lead to a decrease in cPRC calcium levels, likely due to the hyperpolarisation of the cPRCs.
-After 20 sec, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus - a response that may involve depolarisation.
+After 20-sec, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus - a response that may involve depolarisation.
 This activation phase occurs after the end of the stimulation and may be due to a slower neuroendocrine feedback [@Veraszto2018].
 To determine whether NO mediates such a feedback, we measured calcium signals in *NOS*-mutant larvae.
-While we detected the inhibition phase in cPRC in *NOS*-mutants, this was not followed by activation and calcium levels remained low.
+While we detected the inhibition phase in cPRC in *NOS*-mutants, this was not followed by activation and calcium levels remained low (Figure 4B and C).
 These results revealed an essential requirement for NOS signalling in the second phase of the cPRC response.
-
 We next quantified the responses of the INNOS and INRGW interneurons following a UV/violet stimulus to the cPRCs.
@@
 Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision (Figure 4---figure supplement 2).
-In both wild type and *NOS*-mutant larvae, INNOS cells showed increase calcium signals during the inhibition phase of the cPRCs.
-In contrast, calcium signals increased in INRGW cells in wild-type larvae during the second activation phase of cPRCs, a response that was lacking in *NOS*-mutants.
-This suggests that there is no circuit output to the Ser-h1 cells and the ciliary band in *NOS*-mutants, consistent with the observed behavioural defects.
+In both wild type and *NOS*-mutant larvae, INNOS cells showed increase calcium signals during the inhibition phase of the cPRCs (Figure 4D and E).
+In addition, In-NOS depolarisation was suppressed in NIT-GC2 molphants where cPRC hyperpolarisation was suppressed.
+This indicates that cPRC hyperpolarisation is required for IN-NOS depolarisation (Figure 4F).
+In contrast, calcium signals increased in INRGW cells in wild-type larvae during the second activation phase of cPRCs, a response that was lacking in *NOS*-mutants (Figure 4G and H).
+This suggests that there is no circuit output to the Ser-h1 cells and the ciliary band in *NOS*-mutants, consistent with the observed behavioural defects(Figure 4I).
 

### 2022-10-03 code/Fig4.R

@@ -430,26 +430,28 @@ panel_NIT2MO_INRGW <- ggdraw() + draw_image(readPNG("pictures/NIT2MO_INRGW.png")
 panel_NIT2MO_INNOS <- ggdraw() + draw_image(readPNG("pictures/NIT2MO_INNOS.png"))
 
 panel_correlation1 <- ggdraw() + draw_image(readPNG("pictures/n54hpf_WT_e-1_ROI60_PD1.54_power2-8_pin216.9_frame_101-141_401-441_co1-1_crop.png")) +
-  draw_label("cPRC", x = 0.4, y = 0.6, size = 11) +
-  draw_label("IN-NOS", x = 0.35, y = 0.35, size = 11) +
-  draw_label("IN-RGWa", x = 0.7, y = 0.28, size = 11)
+  draw_label("cPRC", x = 0.4, y = 0.6, size = 9) +
+  draw_label("IN-NOS", x = 0.35, y = 0.35, size = 9) +
+  draw_label("IN-RGWa", x = 0.7, y = 0.28, size = 9) +
+  draw_label("Ser-h1", x = 0.8, y = 0.5, size = 9)
 
 panel_correlation2 <- ggdraw() + draw_image(readPNG("pictures/n54hpf_WT_e-1_ROI60_PD1.54_power2-8_pin216.9_frame_101-141_401-441_co1-2_crop.png")) +
-  draw_label("cPRC", x = 0.4, y = 0.6, size = 11) +
-  draw_label("IN-NOS", x = 0.35, y = 0.35, size = 11) +
-  draw_label("IN-RGWa", x = 0.7, y = 0.28, size = 11)
+  draw_label("cPRC", x = 0.4, y = 0.6, size = 9) +
+  draw_label("IN-NOS", x = 0.35, y = 0.35, size = 9) +
+  draw_label("IN-RGWa", x = 0.7, y = 0.28, size = 9) +
+  draw_label("Ser-h1", x = 0.8, y = 0.5, size = 9)
 
 #combine panels into Figure and save final figure as pdf and png
 #panels of different sizes
 layout <- "
-ABC#
-DEF#
-GHIJ
+ABC
+DEF
+GHI
 "
 
 Fig4 <- panel_DAFFM + panel_cPRC_NOS11 + panel_cPRC_NOS23 + 
   panel_correlation2 + panel_INNOS + panel_NIT2MO_INNOS + 
[…]

### 2022-10-04 text/Manuscript.Rmd
figure 4 legend

@@
   in zooplankton
-author: "Kei Jokura, Nobuo Ueda, Martin Gühmann, Luis Alfonso Yañez-Guerra, Gáspár Jékely"
-date: '*Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4
-  4QD, UK*'
+author: "Kei Jokura, Nobuo Ueda, Martin Gühmann, Luis Alfonso Yañez-Guerra, Kyle C. A. Wedgwood, Piotr Słowiński, Gáspár Jékely"
+date: "*Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4
+  4QD, UK*"
 output:
@@
 ![**Figure 4.
-NO produced by UV stimulation to cPRCs induces a depolarising response in cPRC via NIT-GC1.** (A) The intensity of DAF-FM in INNOS changes over time when cPRC (top) and other cells (NC) (bottom) are stimulated with UV (405 nm).
-Hereafter, light purple squares indicate the timing of UV stimulation, gray traces show the optical responses to individual intensities, each intensities are normalized (ΔF/F0) and the dark lines and the bands show the mean value and the standard error range of the mean.
-(B) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in WT (purple) and NOS mutant (NOSΔ23) (gray) larvae.
-(C) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in control (non-injection) and NIT-GC1 knockdown by morpholino (NITGC1-MO1 and 2) larvae.
-(D) Co-expression analysis image of the cPRC marker gene (NIT-GC1: magenta) and the INNOS marker gene (RYa-pNP: green) using in situ HCR.
-White arrowheads indicate INNOS strongly expressing RYa-pNP.
-(E) INNOS (blue dashed line) identified from calcium imaging using GCaMP6s by their position in relation to the cPRC (purple dashed line).
-(F) Comparison of changes over time in GCaMP6s intensities in cPRC (purple) and INNOS (blue) when cPRC is stimulated with UV in WT (left) and NOS mutant (NOSΔ11: right) larvae.
-(G) Comparison of changes over time in GCaMP6s intensities in cPRC (purple) and INRGW (green) when the cPRC are stimulated with UV in WT (left) and NOS mutant (NOSΔ23: right) larvae.
-(H)(I)(J)](../figures/Fig4.png){width="100%"}
+NO produced by UV stimulation to cPRCs induces a depolarising response in cPRC via NIT-GC1.** (A) The intensity of DAF-FM in INNOS changes over time when cPRC (green) and other cells (NC) (gray) are stimulated with UV/violet (405 nm).
+Hereafter, light purple squares indicate the timing of 405 nm light stimulation, the transparent lines show the optical responses to individual intensities, each intensities are normalized (ΔF/F0) and the dark trace lines and the shaded bands show the mean value and the standard error range of the mean.
+(B and C) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in WT (dark yellow) and NOS mutant (B: NOSΔ11/Δ11, B: NOSΔ23/Δ23) (light yellow) larvae.
+(D and G) Correlation (Pearson’s r) map of neuronal activity of the cPRC, IN-NOS, IN-RGWa and Ser-h1 neurons.
+Correlation values were calculated relative to the cPRC (D) or IN-NOS (G).
+(E) Changes over time in GCaMP6s intensities in IN-NOS when cPRC is stimulated with UV in WT (dark purple) and NOS mutant (light purple) larvae.
+(F) Changes over time in GCaMP6s intensities in IN-NOS when cPRC is stimulated with UV in NIT-GC2 morphants larvae.
+(H) Changes over time in GCaMP6s intensities in IN-RGWa when cPRC is stimulated with UV in WT (dark blue) and NOS mutant (light blue) larvae.
+(I) Changes over time in GCaMP6s intensities in Ser-h1 when cPRC is stimulated with UV in WT (dark red) and NOS mutant (light red) larvae.](../figures/Fig4.png){width="100%"}
 

### 2022-10-04 code/Fig3.R

@@ -30,7 +30,7 @@ NIT1_MO_tb%>%
 NIT2_MO_tb%>%
   count(morphant)
 
-# plot NITGC1_analysis ----------------------------------------------------
+# plot NITGC1_analysis cGMP analysis in cell culture-------------------------------------------------
 NITGC1_analysis %>%
   count(expression)
 
@@ -59,7 +59,7 @@ NITGC1_analysis %>%
 ggsave("pictures/GcG-NIT-GC1-SNAP.png", limitsize = FALSE, 
        units = c("px"), width = 800, height = 600, bg='white')  
 
-# plot NITGC1_analysis ----------------------------------------------------
+# plot NITGC1_analysis cGMP analysis in cell culture ----------------------------------------------------
 NITGC1_analysis %>%
   count(expression)
 
@@ -84,7 +84,7 @@ NITGC1_analysis %>%
 ggsave("pictures/GcG-SNAP.png", limitsize = FALSE, 
        units = c("px"), width = 800, height = 600, bg='white')  
 
-# plot NITGC1_analysis ----------------------------------------------------
+# plot NITGC1_analysis cGMP analysis in cell culture ----------------------------------------------------
 
 NITGC1_analysis %>%
   filter(solution == "DMSO") %>%
@@ -114,7 +114,7 @@ ggsave("pictures/GcG-NIT-GC1-DMSO.png", limitsize = FALSE,
 
 
 
-# plot NITGC1_analysis----------------------------------------------------------
+# plot NITGC1_analysis cGMP analysis in cell culture----------------------------------------------------------
 
 NITGC1_analysis %>%
   filter(expression == "GcG-mutNIT1") %>%
@@ -142,15 +142,18 @@ NITGC1_analysis %>%
 ggsave("pictures/GcG-mutNIT-GC1-SNAP.png", limitsize = FALSE, 
        units = c("px"), width = 800, height = 600, bg='white')  
 
-# plot NIT1 morphant data with geom_smooth ----------------------------------------------
+# plot NIT1 morphant calcium imaging data with geom_smooth ----------------------------------------------
+
+NIT1_MO_tb$morphant <- factor(NIT1_MO_tb$morphant, 
+                                levels=c("non-MO", "NIT1-MO1", "NIT1-MO2"))
 
 NIT1_MO_tb %>%
[…]

### 2022-10-04 code/Fig4.R

@@ -86,7 +86,11 @@ WTvsNOS23_cPRC_INRGWa_tb%>%
 DAFFM %>%
   count(stimuli)
 DAFFM
+
+
+max(DAFFM$intensity)
 min(DAFFM$intensity)
+
 
 DAFFM %>%
   ggplot(aes(x=frame,y=intensity,color = stimuli)) +
@@ -96,10 +100,10 @@ DAFFM %>%
   annotate("segment", x=20, xend=40, y=1.2, yend=1.2, size=1)+
   annotate("segment", x=20, xend=20, y=1.2, yend=1.3, size=1)+
   annotate("text", x=29, y=1.15, label="10 sec", size=3)+
-  annotate("text", x=23, y=1.35, label="0.2 ΔF/F0", size=3)+
-  annotate("text", x=70, y=1.4, label="405 nm", size=4)+
-  annotate("text", x=160, y=1.4, label="neuropil DAF-FM", size=5)+
-  ylim(0.6,1.4)+
+  annotate("text", x=23, y=1.31, label="0.2 ΔF/F0", size=3)+
+  annotate("text", x=70, y=1.32, label="405 nm", size=4)+
+  annotate("text", x=160, y=1.32, label="neuropil DAF-FM", size=5)+
+  ylim(0.68,1.32)+
   theme_void()  +
   theme(legend.title = element_blank(), legend.text.align=0,
         legend.text = element_text(size=12)) +
@@ -164,7 +168,7 @@ WTvsNOS23_cPRC_INRGWa_tb %>%
   annotate("text", x=29, y=1.25, label="10 sec", size=3)+
   annotate("text", x=23, y=1.45, label="0.2 ΔF/F0", size=3)+
   annotate("text", x=70, y=1.7, label="405 nm", size=4)+
-  annotate("text", x=160, y=1.7, label="INRGW", size=5)+
+  annotate("text", x=160, y=1.7, label="IN-RGWa", size=5)+
   ylim(0.75,1.7)+
   theme_void()  +
   theme(legend.title = element_blank(), legend.text.align=0,
@@ -234,7 +238,7 @@ WTvsNOS11_cPRC_INNOS_tb %>%
   annotate("text", x=29, y=1.45, label="10 sec", size=3)+
   annotate("text", x=23, y=1.85, label="0.2 ΔF/F0", size=3)+
   annotate("text", x=70, y=3.12, label="405 nm", size=4)+
-  annotate("text", x=160, y=3.12, label="INNOS", size=5)+
[…]

### 2022-10-04 code/Fig5.R

@@ -13,22 +13,22 @@ library(magick)
 
 #read png convert to image panel
 
-panel_A <- ggdraw() + draw_image(readPNG("pictures/scRNAseq_forR_4.png"))
-panel_B <- ggdraw() + draw_image(readPNG("pictures/graph-2_2.png"))
+panel_scRNA_dotplot <- ggdraw() + draw_image(readPNG("pictures/scRNAseq_forR_4.png"))
+panel_model <- ggdraw() + draw_image(readPNG("pictures/model.png"))
+panel_hackathon1 <- ggdraw() + draw_image(readPNG("pictures/hackathon1.png"))
+panel_hackathon2 <- ggdraw() + draw_image(readPNG("pictures/hackathon2.png"))
 
 #combine panels into Figure and save final figure as pdf and png
 #panels of different sizes
 layout <- "
-AB
+ABB
+ACD
 "
 
-Fig5 <- panel_A + panel_B + 
-  patchwork::plot_layout(design = layout, widths = c(1, 2.5)) + #we can change the heights of the rows in our layout (widths also can be defined)
+Fig5 <- panel_scRNA_dotplot + panel_model + panel_hackathon1 + panel_hackathon2 +
+  patchwork::plot_layout(design = layout, widths = c(1, 1.25, 1.25), heights = c(1, 0.5)) + #we can change the heights of the rows in our layout (widths also can be defined)
   patchwork::plot_annotation(tag_levels = 'A') +  #we can change this to 'a' for small caps or 'i' or '1'
-  ggplot2::theme(plot.tag = element_text(size = 12, 
-                                         face='plain')) #or 'bold', 'italic'
-
-Fig5
+  ggplot2::theme(plot.tag = element_text(size = 12, face='bold')) #or 'bold', 'italic'
 
 ggsave("figures/Fig5.png", limitsize = FALSE, 
        units = c("px"), Fig5, width = 2400, height = 1400, bg='white')  

### 2022-10-22 text/Manuscript.Rmd
Edit all figures 4

@@
   in zooplankton
-author: "Kei Jokura, Nobuo Ueda, Martin Gühmann, Luis Alfonso Yañez-Guerra, Kyle C. A. Wedgwood, Piotr Słowiński, Gáspár Jékely"
+author: "*Kei Jokura, Nobuo Ueda, Martin Gühmann, Luis Alfonso Yañez-Guerra, Kyle C. A. Wedgwood, Piotr Słowiński, Gáspár Jékely*"
 date: "*Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4
@@
 Axons and dendrites appear as lines and cell body positions are represented by spheres.
-(H) Wiring diagram of the cPRC circuit.
-<!---# This graph widget does not look very good.
-We need to edit this a bit more.
--->](../figures/Fig1.png){width="100%"}
+(H) Wiring diagram of the cPRC circuit.](../figures/Fig1.png){width="100%"}
 
@@
 (B and C) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in WT (dark yellow) and NOS mutant (B: NOSΔ11/Δ11, B: NOSΔ23/Δ23) (light yellow) larvae.
-(D and G) Correlation (Pearson’s r) map of neuronal activity of the cPRC, IN-NOS, IN-RGWa and Ser-h1 neurons.
+(D and G) Correlation (Pearson's r) map of neuronal activity of the cPRC, IN-NOS, IN-RGWa and Ser-h1 neurons.
 Correlation values were calculated relative to the cPRC (D) or IN-NOS (G).

### 2022-10-22 text/Supplements.Rmd
Put supplement
Edit all figures 4

@@
 title: Figure supplements
-author: "Kei Jokura, Nobuo Ueda, Martin Gühmann, Luis Alfonso Yañez-Guerra, Gáspár Jékely"
+author: "Kei Jokura, Nobuo Ueda, Martin Gühmann, Luis Alfonso Yañez-Guerra, Kyle C. A. Wedgwood, Piotr Słowiński, Gáspár Jékely"
 date: '*Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4
@@
 
-![\*\*Figure 3---figure supplement 1.\*\* (A) Phylogenetic tree of guanylate cyclase by maximum likelihood (ML).
+![**Figure 3---figure supplement 1.** (A) Phylogenetic tree of guanylate cyclase by maximum likelihood (ML).
 Guanylate cyclase-coupled receptor and soluble guanylate cyclase as outgroups.

### 2022-11-23 text/Manuscript.Rmd
MS -editing results section until NIT-GC immunos

@@
   in zooplankton
-author: "*Kei Jokura, Nobuo Ueda, Martin Gühmann, Luis Alfonso Yañez-Guerra, Kyle C. A. Wedgwood, Piotr Słowiński, Gáspár Jékely*"
+author: "*Kei Jokura, Nobuo Ueda, Martin Gühmann, Luis Alfonso Yañez-Guerra, Piotr Słowiński, Kyle C. A. Wedgwood, Gáspár Jékely*"
 date: "*Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4
@@
 Nitric oxide (NO) generated by nitric-oxide synthase (NOS) is a key regulator of animal physiology.
-In nervous systems, NO can modulate circuits, but its effects on neural dynamics and behaviour remain unclear.
-Here we show how NO signalling shapes circuit dynamics to orchestrate light-avoidance behaviour.
-We studied UV-avoidance mediated by brain ciliary photoreceptor cells (cPRCs) in larval *Platynereis dumerilii*, a marine annelid.
+Here we uncover a role for NO in shaping circuit dynamics to orchestrate light-avoidance behaviour.
+We studied UV-avoidance mediated by brain ciliary photoreceptors (cPRCs) in larval *Platynereis dumerilii*, a marine annelid.
 We found NOS specifically expressed in interneurons (INNOS) postsynaptic to cPRCs.
 Stimulation of cPRCs by violet light leads to cPRC inhibition but concomitant INNOS activation and NO production.
-NO feeds back to cPRCs and triggers delayed cPRC activation through an unconventional guanylyl cyclase.
-This results in the activation of a second population of postsynaptic interneurons (INRGW).
-It was also suggested that activation of INRGWs activates serotonergic ciliomotor neurons (Ser-h1), which are postsynaptic and involved in increasing ciliary beat frequency.
-In *NOS* knockout larvae, NO feedback and INRGW activation do not occur and avoidance behaviour is defective.
+NO feeds back to cPRCs and triggers their delayed activation through an unconventional guanylyl cyclase.
+This results in the activation of projection interneurons and the inhibition of serotonergic ciliomotor neurons.
+In *NOS* mutants, NO feedback and projection-neuron activation do not occur and avoidance behaviour is defective.
+By mathematical modelling, we could capture the circuit dynamics in both wild-type and mutant larvae.
 Our results reveal how NO-mediated neuroendocrine signalling gates a synaptic circuit to initiate light-avoidance behaviour.
@@
 We identified a single *nitric oxide synthase* (*NOS*) gene in the *Platynereis dumerilii* genome and transcriptome data.
-Phylogenetic analysis of NOS proteins indicate that the *Platynereis* NOS belongs to an orthology group of metazoan NOS sequences (Figure 1---figure supplement 1).
+Phylogenetic analysis of NOS proteins indicate that *Platynereis* NOS belongs to an orthology group of metazoan NOS sequences (Figure 1---figure supplement 1).
 To characterise the expression pattern of *NOS* we used in situ hybridization chain reaction (HCR) and transient transgenesis.
-In 2 and 3-day-old larvae, we detected *NOS* expression in two or four cells in the apical organ region (Figure 1A-D).
-The same four cells were labelled with a reporter construct driving palmitoylated-tdTomato by a 12 kbp upstream genomic region of the *NOS* gene in 3-day-old larvae(Figure 1E).
+In two- and three-day-old larvae, we detected *NOS* expression in four cells (two of them weakly expressing) in the apical organ region (Figure 1A-D).
+The same four cells were labelled with a reporter construct driving palmitoylated-tdTomato by a 12 kbp upstream genomic region of the *NOS* gene (Figure 1E).
 This reporter also revealed the axonal projections of the NOS-expressing neurons.
-The position and morphology of the four *NOS*-expressing cells allowed us to identify the same cells in the 3-day-old whole-body volume EM data [@williams2017; @veraszto2020] as four interneurons (INNOS)(Figure 1F and G).
-The INNOS cells are postsynaptic to the UV-sensory cPRCs and presynaptic to the INRGW interneurons, which are also cPRC targets (Figure 1H).
+The position and morphology of the four *NOS*-expressing cells allowed us to identify the same cells in the three-day-old whole-body volume EM data [@williams2017; @veraszto2020] as four interneurons (INNOS)(Figure 1F and G).
+The INNOS cells are postsynaptic to the UV-sensory cPRCs and presynaptic to the INRGW interneurons, which are also cPRC targets (Figure 1C, D).
+INRGW neurons synapse on the head serotonergic ciliomotor neurons (Ser-h1), which synapse on the prototroch ciliary band and the cholinergic MC ciliomotor neurons (Figure 1C, D).
+We also detected *NOS* expression by HCR in the region of the visual eyes and the pigmented eyespots.
 
@@
 
-**Nitric oxide signalling mediates UV-avoidance behaviour** The specific expression of NOS in interneurons directly postsynaptic to the cPRCs suggests that NO signalling is involved in the UV response.
+**Nitric oxide signalling mediates UV-avoidance behaviour**
+
+The specific expression of NOS in interneurons directly postsynaptic to the cPRCs suggests that NO signalling is involved in the UV response.
 To test this, we generated two *Platynereis* *NOS* knockout lines with the CRISPR/Cas9 method.
@@
 To quantify UV avoidance, we recorded the trajectories of freely swimming larvae in vertical columns, illuminated laterally from two opposite sides with 395 nm violet light.
-As previously shown, wild-type larvae swim downward within 30 sec following UV/violet light stimulation [@veraszto2018].
+We also tested phototaxis, by exposing larvae to 480 nm directional and collimated light from the top of the column.
+As previously shown, wild-type larvae swim downward within 30 sec following non-directional UV/violet light stimulation [@veraszto2018].
 In contrast, homozygous *NOS*-mutant larvae showed a strongly diminished UV-avoidance response (Figure 2B and C).
 This phenotype is similar to the defective UV-avoidance of *c-opsin1* mutant larvae and reveals a requirement for NOS in UV-avoidance behaviour [@veraszto2018].
+*NOS*-mutant larvae also showed reduced phototactic behaviour, suggesting a function in the visual eyes mediating phototaxis [@Randel2014a], where *NOS* is also expressed (Figure 1).
+To test if the observed defects are due to a role of NOS in neuronal signalling or circuit developmental, we also tested larvae exposed to a NOS inhibitor, L-NAME.
+Larvae incubated for xy min in 0.1 mM or 1 mM L-NAME showed defective UV avoidance, but normal phototaxis.
+Overall, our results indicate an acute requirement for NOS signalling in UV-avoidance and a possible indirect, developmental role in the visual system.
 
@@
 
-NO generally acts via cyclic guanosine monophosphate (cGMP) as a second messenger.
-In this signalling pathway, NO binds to the heme group of soluble guanylate cyclase (sGC), a member of the guanylate cyclase family with a CYC domain (PF00211), to increase cGMP.
-However, no NO targets were found in cells that highly express sGC within the cPRC circuit.
-Recently, Moroz and others reported that GC with NIT (nitrite/nitrate sensing) domains (PF08376) (NIT-GC), identified as a sensor for nitrate and nitrite, the oxides of NO, is relatively widely conserved in metazoans [@Moroz2020].
-To search for NIT-GCs in Platynereis, genes with the CYC domains were searched using the transcriptome and genome sequences of 45 metazoan species and choanoflagellate.
-By sequence-similarity searches, multiple sequence alignments and cluster analyses, a cluster of membrane-bound (guanylate cyclase-coupled receptor) type was found and a phylogenetic tree was constructed with soluble type (sGC) as an outgroup.
-As in the phylogenetic analysis of Moroz et al, we found a group of GCs with NIT domains that included many metazoans such as placozoa, cnidaria, ecdysozoa, echinoderms and lophotrochozoa, and in Platynereis 15 NIT-GCs were found (Figure 3---figure supplement 1).
-Using previously published spatially mapped single-cell transcriptome data [@Achim2015; @williams2017] to find candidate NIT-GCs involved in the cPRC circuit among 15 NIT-GCs, two NIT-GCs (NIT-GC1 and NIT-GC2) were found to be highly expressed in the cPRC (Figure 3---figure supplement 1).
-Expression analysis with c-opsin1 and MLD pNPs (pedal peptide neuropeptide precursor 2) [@Arendt2004; @williams2017], which have already been identified as cPRC markers using in situ HCR, showed that both NIT-GC1 and NIT-GC2 are strongly and specifically expressed in cPRC (Figure 3A and B, Figure 3---figure supplement 2).
-To analyse localisations at the protein level, immunostaining was performed using polyclonal antibodies against NIT-GC1 and NIT-GC2.
-Interestingly, NIT-GC1 was found to localise to the neuronal axon and NIT-GC2 to the ciliary region of the cPRC (Figure 3C and D).
-Furthermore, co-immunostaining with NOS antibodies revealed that NOS and NIT-GC1 were localised in close proximity in the neurosecretory plexus (Figure 3---figure supplement 3).
-These two localisation differences suggest the existence of different intracellular signalling mechanisms in the ciliary and neuroaxonal regions within the cPRC.
+NO generally acts via soluble guanylate cyclases (sGC), belonging to the guanylate cyclase family with a CYC domain (PF00211).
+NO binding to the heme group of sGC leads to increased cyclic guanosine monophosphate (cGMP) production.
+Analysis of sGCs in *Platynereis* indicated that these genes are not expressed in any of the cells of the cPRC circuit ([@Veraszto2017] and data not shown).
+Recently, Moroz and coworkers reported an atypical but widely conserved family of guanylyl cyclases with a NIT (nitrite/nitrate sensing) domain (PF08376) (NIT-GC) as potential mediators of NO signalling [@Moroz2020].
+To search for NIT-GCs in _Platynereis_, we searched transcriptome resources and retrieved 15 potential NIT-GC homologs.
+To analyse the relationship of these sequences to metazoan NIT-GCs, we retrieved protein sequences with a CYC domain from the transcriptome and genome databases of 45 metazoan species and [**how many?] choanoflagellate.
+We carried out cluster analysis and did phylogenetic reconstruction on a group of membrane-bound guanylyl cyclases with sGCs as an outgroup.
+In agreement with Moroz et al.
+[@Moroz2020],we found a group of GCs with NIT domains with representatives in placozoans, cnidarians, some ecdysozoans, echinoderms, and lophotrochozoans.
+The 15 _Platynereis_ sequences were parts of several major and ancient clades of the tree (Figure 4---figure supplement 1).
+To characterise the expression of NIT-GCs, we using previously published spatially mapped single-cell transcriptome data [@Achim2015; @williams2017].
+Among the 15 NIT-GCs, four showed high and specific expression in the cPRCs and one was expressed in the INNOS cells (Figure 3---figure supplement 1).
+In the single-cell data, we could identify the cPRCs by the specific expression of c-opsin1 and a pedal-peptide2 neuropeptide precursor (MLD proneuropeptide) [@Arendt2004; @williams2017], which have previously been described as cPRC markers (Figure 3A and B, Figure 3---figure supplement 2).
+The INNOS cells were identified by NOS expression and their spatial mapping.
+We decided to focus on two NIT-GCs expressed in the cPRCs and with a full-length sequences, NIT-GC1 and NIT-GC2.
+To confirm the single-cell data, we first carried out HCR with probes for _NIT-GC1_ and _NIT-GC2_ mRNA.
+Both genes showed very clear and specific expression in the four cPRCs, as confirmed by co-labeling with an acetylated tubulin antibody and with an HCR probe against _MLD proneuropeptide_.
+To analyse the subcellular localisation of NIT-GC1 and NIT-GC2 at the protein level, we raised and affinity-purified polyclonal antibodies against a specific peptide sequence from both proteins.
+In immunostainings, we found that NIT-GC1 was localise to the region corresponding to the axonal projections of the cPRCs in the anterior nerve plexus.
+Co-immunostaining with the rabbit NIT-GC1 and the rat NOS antibodies revealed that both proteins were localised in close proximity in the neurosecretory plexus (Figure 3---figure supplement 3).
+In contrast, NIT-GC2 strongly and specifically labelled the large sensory ciliary region of the cPRCs (Figure 3C and D).
+These different subcellular localisations suggest that the two NIT-GCs are involved in different intracellular signalling processes in the ciliary and axonal regions of the cPRCs.
 

### 2022-12-02 text/Manuscript.Rmd
Figure_pdf

--- 2022-10-22 text/Manuscript.Rmd
@@
 ![**Figure 1.
-Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** (A) SEM image of a 3-day-old Platynereis larva.(B-D) Expression analysis of the NOS gene (magenta) using in situ HCR.
-(D) Co-stained image using acetylated α-tubulin antibody (ac α-tub: green).
-(B) Dorsal view of the larva at 3-day-old.
-(C and D) Anterior view of the larva at 2 (D) and 3-day-old (C).
-(E) Immunostaining with anti-HA antibody against HA tagged palmitoylated-tdTomato expressed under the upstream of the start site of NOS (magenta).
-Co-staining with acetylated α-tubulin antibody as marker for cilia and axonal scaffold (ac α-tub: green).
-INNOSs with the entire cell body and axon labelled are shown with white arrowheads and with only the cell body labelled are shown with black arrowheads.
-Views of larvae at 3-day-old from the anterior side.
-(F and G) Five types of neurons (cPRC, INNOS, INRGWa, Ser-h1 and MC) in the cPRC circuit reconstructed in the whole-body transmission electron microscopy (ssTEM) dataset of a 3-day-old larva.
-The INNOS on the ventral side labelled with (E) are shown in purple.
+Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** (A) SEM image of a 3-day-old Platynereis larva.
+(B and C) Five types of neurons (cPRC, INNOS, INRGWa, Ser-h1 and MC) in the cPRC circuit reconstructed in the whole-body transmission electron microscopy (ssTEM) dataset of a 3-day-old larva.
+The INNOS on the ventral side labelled with (C) are shown in purple.
 Axons and dendrites appear as lines and cell body positions are represented by spheres.
-(H) Wiring diagram of the cPRC circuit.](../figures/Fig1.png){width="100%"}
+(D) Wiring diagram of the cPRC circuit.
+(E-G) Expression analysis of the NOS gene (magenta) using in situ HCR.
+(E) Dorsal view of the larva at 3-day-old.
+(F and G) Anterior view of the larva at 3 (F) and 2-day-old (G).
+(F) Insert: showing two NOS cells on each side.
+(G) Co-stained image using acetylated α-tubulin antibody (acTub: green).
+(H) Immunostaining with anti-HA antibody against HA tagged palmitoylated-tdTomato expressed under the upstream of the start site of NOS (NOSp::palmi-3xHA: magenta).
+Co-staining with acetylated α-tubulin antibody as marker for cilia of cPRC and axonal scaffold (acTub: green).
+Views of larvae at 3-day-old from the anterior side.](../figures/Fig1.png){width="100%"}
 
@@
 
-![**Figure 2.
-Larvae with inhibited NOS have shorter ciliary arrest times and a delayed response to UV-avoidance.** (A) Top: The domain structure of NOS.
+![**Figure 2. Larvae with inhibited NOS have shorter ciliary arrest times and a delayed response to UV-avoidance.** 
+(A) Vertical swimming in larvae treated with NOS inhibitors, L-NAME at 3-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
+The data are shown in 30 s bins.
+(B) Top: The domain and Exon/Intron structure of NOS.
 Bottom: Close-up region showing the genomic locus of NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9.
@@
 Pink indicates target sites.
-Yellow shows PAM sequences, red shows stop codons.
-(B and C) Vertical swimming in wild type and mutant larvae at 2-day-old (B) and 3-day-old (C) stimulated with blue (488 nm) and UV (395 nm) light from above.
+Gray shows PAM sequences, red shows stop codons.
+(C and D) Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23) larvae at 2-day-old (C) and 3-day-old (D) stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
 The data are shown in 30 s bins.
-(D) Overlaid trajectories for WT (left, n=20) and NOS mutant (right, n=31) 3-day-old larvae.
+
+(E and F) Overlaid trajectories for WT (left, n=20) and NOS mutant (right, n=31) 3-day-old larvae.
 Grey shows the trajectories before UV stimulation and purple shows the trajectories after UV stimulation.
-(E) The temporal changes in the vertical position of the WT and mutant larvae before and after UV stimulation are shown.
-The starting points of each larval trajectory are set to 0.
+
+(E) The temporal changes in the vertical position of the WT and mutant larvae before and after UV stimulation are shown. The starting points of each larval trajectory are set to 0.
+
 (F) The temporal changes in the distance traveled of the WT and mutant larvae before and after UV stimulation are shown.
+
 (E and F) After UV stimulation is indicated by purple squares.](../figures/Fig2.png){width="100%"}
+
+**Nitric oxide produced during UV/violet stimulation gates the output of the cPRC circuit**
+
+To test if NO is produced during UV/violet responses, we injected zygotes with the fluorescent NO-reporter DAF-FM and imaged two-day-old larvae during the local stimulation of the cPRC cilia by 405 nm light (Figure 4---figure supplement 1).
+Following light stimulation, we detected an increase in DAF-FM fluorescence in the region of the INNOS cells (Figure 4A).
+We next analysed the dynamics of the cPRC circuit following light stimulation.
+As we have shown previously [@Veraszto2018], a 20-sec illumination of the cPRC ciliary region lead to a decrease in cPRC calcium levels, likely due to the hyperpolarisation of the cPRCs.
+After 20-sec, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus - a response that may involve depolarisation.
+This activation phase occurs after the end of the stimulation and may be due to a slower neuroendocrine feedback [@Veraszto2018].
+To determine whether NO mediates such a feedback, we measured calcium signals in *NOS*-mutant larvae.
+While we detected the inhibition phase in cPRC in *NOS*-mutants, this was not followed by activation and calcium levels remained low (Figure 4B and C).
+These results revealed an essential requirement for NOS signalling in the second phase of the cPRC response.
+
+![**Figure 3.
+NO produced by UV stimulation to cPRCs induces a depolarising response in cPRC.** (A) A NO-positive cell visualised by DAF-FM is shown (dashline).
+The white dashed line indicates the outer frame of the larva.
+All subsequent images are of 2-day-old larvae.
+(B) Intensity changes before and during 405 nm light stimulation in a cell visualised by DAF-FM.
+(C) The intensity of DAF-FM changes over time when cPRC (green) and other cells (NC) (gray) are stimulated with UV/violet (405 nm) light.
+Purple squares indicate the timing of 405 nm light stimulation, the transparent lines show the optical responses to individual intensities, each intensities are normalized (ΔF/F0) and the dark trace lines and the shaded bands show the mean value and the standard error range of the mean.
+(D and E) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in WT (dark yellow) and NOS mutant (D: NOSΔ11/Δ11, E: NOSΔ23/Δ23) (light yellow) larvae.](../figures/Fig3.png){width="100%"}
 
@@
 
-![**Figure 3.
+![**Figure 4.
 cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.** (A) Expression analysis of the NIT-GC1 gene (magenta) using in situ HCR.
@@
 Green shows co-staining with acetylated α-tubulin antibody (ac α-tub).
-(E and F) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in control (non-injection) and NIT-GC1 and NIT-GC2 knockdown by morpholino (MO1 and 2) larvae.
-(G-J) The intensity changes in cGMP fluorescence over time for the four conditions are shown.
+
+(E-G) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in control (non-injection) and NIT-GC1 and NIT-GC2 knockdown by morpholino (MO1 and 2) larvae.
+
+(H) The domain structure of NIT-GC1. Grey squares indicate the transmembrane domain. 
+(I) A schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control.
+Green cGull shows cGMP-dependent green fluorescence.
+(J-M) The intensity changes in cGMP fluorescence over time for the four conditions are shown.
 Gray traces show the optical responses to individual intensities (n \> 6).
@@
 The dark line shows the mean value and the band the standard error range of the mean.
-The respective solution was acted on 2 min after the start of imaging (grey bars).](../figures/Fig3.png){width="100%"}
+The respective solution was acted on 2 min after the start of imaging (grey bars).](../figures/Fig4.png){width="100%"}
 
-**Nitric oxide produced during UV/violet stimulation gates the output of the cPRC circuit**
-
-To test if NO is produced during UV/violet responses, we injected zygotes with the fluorescent NO-reporter DAF-FM and imaged two-day-old larvae during the local stimulation of the cPRC cilia by 405 nm light (Figure 4---figure supplement 1).
-Following light stimulation, we detected an increase in DAF-FM fluorescence in the region of the INNOS cells (Figure 4A).
-We next analysed the dynamics of the cPRC circuit following light stimulation.
-As we have shown previously [@Veraszto2018], a 20-sec illumination of the cPRC ciliary region lead to a decrease in cPRC calcium levels, likely due to the hyperpolarisation of the cPRCs.
-After 20-sec, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus - a response that may involve depolarisation.
-This activation phase occurs after the end of the stimulation and may be due to a slower neuroendocrine feedback [@Veraszto2018].
-To determine whether NO mediates such a feedback, we measured calcium signals in *NOS*-mutant larvae.
-While we detected the inhibition phase in cPRC in *NOS*-mutants, this was not followed by activation and calcium levels remained low (Figure 4B and C).
-These results revealed an essential requirement for NOS signalling in the second phase of the cPRC response.
 We next quantified the responses of the INNOS and INRGW interneurons following a UV/violet stimulus to the cPRCs.
@@
 
-![**Figure 4.
-NO produced by UV stimulation to cPRCs induces a depolarising response in cPRC via NIT-GC1.** (A) The intensity of DAF-FM in INNOS changes over time when cPRC (green) and other cells (NC) (gray) are stimulated with UV/violet (405 nm).
-Hereafter, light purple squares indicate the timing of 405 nm light stimulation, the transparent lines show the optical responses to individual intensities, each intensities are normalized (ΔF/F0) and the dark trace lines and the shaded bands show the mean value and the standard error range of the mean.
-(B and C) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in WT (dark yellow) and NOS mutant (B: NOSΔ11/Δ11, B: NOSΔ23/Δ23) (light yellow) larvae.
-(D and G) Correlation (Pearson's r) map of neuronal activity of the cPRC, IN-NOS, IN-RGWa and Ser-h1 neurons.
-Correlation values were calculated relative to the cPRC (D) or IN-NOS (G).
+![**Figure 5.
+NO produced by UV stimulation to cPRCs induces a depolarising response in cPRC via NIT-GC1.** (A) Schematic diagram of the immunostaining procedure after calcium imaging.
+(B and D) The position of the cPRC in calcium imaging with GCaMP6s.
+Cells around the cPRC are numbered according to the position of the nuclei.
+(C and E) Immunostaining image using RYamide (C) or RGWamide and 5-HT antibodies (red) (E) and DAPI (cyan) in the same region as (B) or (D).
+IN-NOS, IN-RGWa and Ser-h1 identified from the numbered nuclear positions are shown.
+(F) Correlation (Pearson's r) map of neuronal activity of the cPRC, IN-NOS, IN-RGWa and Ser-h1 neurons.
+
 (E) Changes over time in GCaMP6s intensities in IN-NOS when cPRC is stimulated with UV in WT (dark purple) and NOS mutant (light purple) larvae.
@@
 (H) Changes over time in GCaMP6s intensities in IN-RGWa when cPRC is stimulated with UV in WT (dark blue) and NOS mutant (light blue) larvae.
-(I) Changes over time in GCaMP6s intensities in Ser-h1 when cPRC is stimulated with UV in WT (dark red) and NOS mutant (light red) larvae.](../figures/Fig4.png){width="100%"}
+(I) Changes over time in GCaMP6s intensities in Ser-h1 when cPRC is stimulated with UV in WT (dark red) and NOS mutant (light red) larvae.](../figures/Fig5.png){width="100%"}
 
@@
 
+**NIT-GC Identification and Phylogenetic Analysis**
+To identify GPR173 receptors, we obtained a database of vertebrate SREB sequences, including GPR173, GPR85 and GPR27 from (Breton et al.
+2021).
+From these sequences, we produced a Hidden Markov Model (HMM) and used this to mine the 49 species investigated.
+HMM models were run in HMMR3 with an e-value of 1e−15.
+The same SREB sequences were used to carry out similarity-based searches using BlastP with an e-value cutoff of 1e−15.
+We merged these two databases and ran CD-Hit (Fu et al.
+2012) to eliminate redundant sequences (at a 99% threshold).
+To identify the sequences that are closely related to the GPR173 sequences, we ran a cluster-based analysis in CLANS (Frickey and Lupas 2004).The CLANS analysis is available as supplementary fig.
+S4, Supplementary Material online.
+To identify clusters, we used the convex-clustering option with 100 jack-knife replicates.
+The SREB receptors are extremely well conserved and form an easily recognizable cluster.
+To analyze the phylogeny of SREB receptors, the cluster containing these receptors together with monoaminergic receptors were parsed and used for tree building.
+We aligned the sequences with MAFFT version 7, with the iterative refinement method E-INS-i.
+Alignments were trimmed with TrimAl in gappy-out mode (Capella-Gutiérrez et al.
+2009).
+To calculate maximum-likelihood trees, we used IQ-tree2 with the LG+G4 model.
+To calculate branch support, we ran 1,000 replicates with the aLRT-SH-like and aBayes methods (Minh et al.
+2020).
+The sequences used for the phylogenetic analysis are available in supplementary XX, Supplementary Material online, the trimmed alignment is available in supplementary XX, Supplementary Material online.
+The raw nexus tree of the SREB receptors is available in supplementary file XX, Supplementary Material online.
+
+**Single-cell analysis**
+I used Achim et al 2015 for the single-cell data. In fact, in Liz's 2018 paper, she used 107 cells as neurons by removing duplicates from Achim's single-cell data, so I used those cells.
+Since the raw data were read count data, I normalized them to TPM using Python.
+After that we converted them to log10. From the sum of the expression levels in 107 cells for each gene, I calculated the percentage in each cell.
+I chose the appropriate genes for now. For each cell, I identified them with marker genes.
+After created the data in Python, plotted it using R dot plots.
+
+RPKM? TPM?
+RPKM calculates the total number of reads per million bp and then divides by the length of each gene, so it is not possible to compare between samples.
+Instead, TPM first divides by the length of the gene and then divides by the total number of reads per million bases, which allows for more accurate comparisons between samples.
+In this case I wanted to compare between samples, so I used TPM.
+In fact, TPM is becoming the mainstream method for many single-cell analyses.
+The total TPM of each gene between the samples was used to calculate the percentage of expressed genes.
+
 **Comparing behavior of wildtype and NOS-knockout 3-day-old larvae**

### 2022-12-02 text/Supplements.Rmd
Figure_pdf

@@
 
-![**Figure 1---figure supplement 1.** (A) Phylogenetic tree of NOS by maximum likelihood (ML).
-Tree robustness was tested with 1000 replicates of ultrafast bootstrap.
-Species names are as follows: Ap (Aiptasia pallida), Bb (Branchiostoma belcheri), Cv (Crassostrea virnigica), Dm (Drosophila melanogaster), Dp (Daphnia pulex), Hh (Hoilungia hongkongensis), Lg (Lottia gigantea), Pc (Priapulus caudatus), Pd (Platynereis dumerilii), Sk (Saccoglossus kowalevskii), Sp (Strongylocentrotus purpuratus), Ta (Trichoplax adhaerens), Tc (Tribolium castaneum).](../figures/Fig1_sup1.png){width="100%"}
+![**Figure 1---figure supplement 1.** Phylogenetic tree of NOS by maximum likelihood.
+Tree robustness was tested with 1000 replicates of ultrafast bootstrap with the aLRT-SH-like and aBayes methods.](../figures/Fig1_sup1.png){width="100%"}
 
-![**Figure 1---figure supplement 2.** (A) Immunostaining with NOS antibodies.(B) Co-expression analysis image with the RYamide proneuropeptide gene (RYa-pNP: green), which is highly expressed in INNOS.
+![**Figure 4---figure supplement 1.** Phylogenetic tree of guanylate cyclase by maximum likelihood (ML).
+Guanylate cyclase-coupled receptor and soluble guanylate cyclases (sGC) as outgroups.
+Guanylate cyclases with NIT domains are found in most animal phyla except Porifera, Ctenophora, Urochordata and Chordata.](../figures/Fig4_sup1.png){width="100%"}
+
+![**Figure 4---figure supplement 2.** (A) Immunostaining with NOS antibodies.
+(B) Co-expression analysis image with the RYamide proneuropeptide gene (RYa-pNP: green), which is highly expressed in INNOS.
 White arrowheads indicate co-expressed INNOS.
 Black arrowheads indicate larval eyes.
-White dashed lines indicate the outer frame of the larva.](../figures/Fig1_sup2.png){width="100%"}
-
-![**Figure 3---figure supplement 1.** (A) Phylogenetic tree of guanylate cyclase by maximum likelihood (ML).
-Guanylate cyclase-coupled receptor and soluble guanylate cyclase as outgroups.
-Guanylate cyclases with NIT domains are found in most animal phyla except Porifera, Ctenophora, Urochordata and Chordata.
-Tree robustness was tested with 1000 replicates of ultrafast bootstrap.
-Species names are as follows: Ap (Aiptasia pallida), Bb (Branchiostoma belcheri), Cv (Crassostrea virnigica), Dm (Drosophila melanogaster), Dp (Daphnia pulex), Hh (Hoilungia hongkongensis), Lg (Lottia gigantea), Pc (Priapulus caudatus), Pd (Platynereis dumerilii), Sk (Saccoglossus kowalevskii), Sp (Strongylocentrotus purpuratus), Ta (Trichoplax adhaerens), Tc (Tribolium castaneum).](../figures/Fig3_sup1.png){width="100%"}
-
-![**Figure 3---figure supplement 2.** (A) Co-expression analysis image of the MLD-pedal2 amide proneuropeptide gene (MLD-pNP: green).
+White dashed lines indicate the outer frame of the larva.
+(A) Co-expression analysis image of the MLD-pedal2 amide proneuropeptide gene (MLD-pNP: green).
 Black arrowheads indicate cells different from cPRC.
-(B) Co-localisation analysis using NIT-GC1 and NOS antibodies.](../figures/Fig3_sup2.png){width="100%"}
-
-![**Figure 3---figure supplement 3.** (A) Comparison of amino acid sequences of regions of possible NO binding in NIT-GC1 and NIT-GC2.
-(B) 3D stuructures of NIT-GC1 and NIT-GC2 (C)](../figures/Fig3_sup3.png){width="100%"}
-
-![**Figure 3---figure supplement 4.** (A and B) Left: The domain structure of NIT-GC1.
-Grey squares indicate the transmembrane domain.
-Right shows a schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control.
-Green cGull shows cGMP-dependent green fluorescence.](../figures/Fig3_sup4.png){width="100%"}
-
-![**Figure 4---figure supplement 1.** (A) An INNOS visualised by DAF-FM is shown (a white arrowhead).
-A black arrowhead indicates the position of the adult eye.
-The white dashed line indicates the outer frame of the larva.
-All subsequent images are of 2-day-old larvae.
-(B) Dual-colour imaging of DAF-FM (green) and RGECO1a (magenta) allows simultaneous visualisation of the INNOS and cPRCs.
-(C) Enlarged view of (B).
-cPRCs and INNOS are indicated by arrows.](../figures/Fig4_sup1.png){width="100%"}
-
-![**Figure 4---figure supplement 2.** (A) Schematic diagram of the immunostaining procedure after calcium imaging.
-(B) The position of the cPRC (purple dashed line) in calcium imaging with GCaMP6s.
-Cells around the cPRC are numbered according to the position of the nuclei.
-(C) Immunostaining image using RGWamide (RGWa) and serotonin (Ser) antibodies (red) and DAPI (light blue) in the same region as (B).
-INRGW (green dashed line) and Ser-h1 (black arrowheads) identified from the numbered nuclear positions are shown.](../figures/Fig4_sup2.png){width="100%"}
+(B) Co-localisation analysis using NIT-GC1 and NOS antibodies.](../figures/Fig4_sup2.png){width="100%"}

### 2023-01-01 text/Manuscript.Rmd
merge

@@
 
-![**Figure 2. Larvae with inhibited NOS have shorter ciliary arrest times and a delayed response to UV-avoidance.** 
-(A) Vertical swimming in larvae treated with NOS inhibitors, L-NAME at 3-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
+![**Figure 2.
+Larvae with inhibited NOS have shorter ciliary arrest times and a delayed response to UV-avoidance.** (A) Vertical swimming in larvae treated with NOS inhibitors, L-NAME at 3-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
 The data are shown in 30 s bins.
@@
 Recently, Moroz and coworkers reported an atypical but widely conserved family of guanylyl cyclases with a NIT (nitrite/nitrate sensing) domain (PF08376) (NIT-GC) as potential mediators of NO signalling [@Moroz2020].
-To search for NIT-GCs in _Platynereis_, we searched transcriptome resources and retrieved 15 potential NIT-GC homologs.
-To analyse the relationship of these sequences to metazoan NIT-GCs, we retrieved protein sequences with a CYC domain from the transcriptome and genome databases of 45 metazoan species and [**how many?] choanoflagellate.
+To search for NIT-GCs in *Platynereis*, we searched transcriptome resources and retrieved 15 potential NIT-GC homologs.
+To analyse the relationship of these sequences to metazoan NIT-GCs, we retrieved protein sequences with a CYC domain from the transcriptome and genome databases of 45 metazoan species and [\*\*how many?] choanoflagellate.
 We carried out cluster analysis and did phylogenetic reconstruction on a group of membrane-bound guanylyl cyclases with sGCs as an outgroup.
@@
 [@Moroz2020],we found a group of GCs with NIT domains with representatives in placozoans, cnidarians, some ecdysozoans, echinoderms, and lophotrochozoans.
-The 15 _Platynereis_ sequences were parts of several major and ancient clades of the tree (Figure 4---figure supplement 1).
+The 15 *Platynereis* sequences were parts of several major and ancient clades of the tree (Figure 4---figure supplement 1).
 To characterise the expression of NIT-GCs, we using previously published spatially mapped single-cell transcriptome data [@Achim2015; @williams2017].
@@
 We decided to focus on two NIT-GCs expressed in the cPRCs and with a full-length sequences, NIT-GC1 and NIT-GC2.
-To confirm the single-cell data, we first carried out HCR with probes for _NIT-GC1_ and _NIT-GC2_ mRNA.
-Both genes showed very clear and specific expression in the four cPRCs, as confirmed by co-labeling with an acetylated tubulin antibody and with an HCR probe against _MLD proneuropeptide_.
+To confirm the single-cell data, we first carried out HCR with probes for *NIT-GC1* and *NIT-GC2* mRNA.
+Both genes showed very clear and specific expression in the four cPRCs, as confirmed by co-labeling with an acetylated tubulin antibody and with an HCR probe against *MLD proneuropeptide*.
 To analyse the subcellular localisation of NIT-GC1 and NIT-GC2 at the protein level, we raised and affinity-purified polyclonal antibodies against a specific peptide sequence from both proteins.
@@
 
-**NIT-GC Identification and Phylogenetic Analysis**
-To identify GPR173 receptors, we obtained a database of vertebrate SREB sequences, including GPR173, GPR85 and GPR27 from (Breton et al.
+**NIT-GC Identification and Phylogenetic Analysis** To identify GPR173 receptors, we obtained a database of vertebrate SREB sequences, including GPR173, GPR85 and GPR27 from (Breton et al.
 2021).
@@
 
-**Single-cell analysis**
-I used Achim et al 2015 for the single-cell data. In fact, in Liz's 2018 paper, she used 107 cells as neurons by removing duplicates from Achim's single-cell data, so I used those cells.
+**Single-cell analysis** I used Achim et al 2015 for the single-cell data.
+In fact, in Liz's 2018 paper, she used 107 cells as neurons by removing duplicates from Achim's single-cell data, so I used those cells.
 Since the raw data were read count data, I normalized them to TPM using Python.
-After that we converted them to log10. From the sum of the expression levels in 107 cells for each gene, I calculated the percentage in each cell.
-I chose the appropriate genes for now. For each cell, I identified them with marker genes.
+After that we converted them to log10.
+From the sum of the expression levels in 107 cells for each gene, I calculated the percentage in each cell.
+I chose the appropriate genes for now.
+For each cell, I identified them with marker genes.
 After created the data in Python, plotted it using R dot plots.
 
-RPKM? TPM?
+RPKM?
+TPM?
 RPKM calculates the total number of reads per million bp and then divides by the length of each gene, so it is not possible to compare between samples.

### 2023-01-01 code/Fig1.R

@@ -257,6 +257,7 @@ panel_Network <- ggdraw() + draw_image(readPNG("pictures/visNetwork_INNOS_2.png"
 
 
 arrow_fluo <- data.frame(x1 = 0.95, x2 = 0.95, y1 = 0.75, y2 = 0.85)
+
 panel_NOS3d_ventr_HCR <- ggdraw() + draw_image(readPNG("pictures/HCR_72_DV_NOS_90um.png")) +
   draw_label("in situ HCR", x = 0.3, y = 0.99, size = 10) +
   draw_label("NOS", x = 0.12, y = 0.9, color="magenta", size = 11, fontface="italic") +
@@ -278,11 +279,11 @@ panel_NOS3d_ventr_HCR <- ggdraw() + draw_image(readPNG("pictures/HCR_72_DV_NOS_9
   draw_line(x = c(0.77, 0.65), y = c(0.43, 0.35), color = "white", size = 0.2)
 
 
-panel_NOS3d_ant_HCR <- ggdraw() + draw_image(readPNG("pictures/HCR_72_AP_NOS_92.02um.png")) +
+panel_NOS3d_ant_HCR <- ggdraw() + draw_image(readPNG("pictures/HCR_72_AP_NOS_89.25um.png")) +
   draw_label("in situ HCR", x = 0.3, y = 0.99, size = 10) +
   draw_label("NOS", x = 0.12, y = 0.9, color="magenta", size = 11, fontface="italic") +
   draw_label("DAPI", x = 0.32, y = 0.9, color="cyan", size = 11, fontface="plain") +
-  draw_line(x = c(0.04, 0.221), y = c(0.08, 0.08), color = "white", size = 0.5) +
+  draw_line(x = c(0.04, 0.264), y = c(0.08, 0.08), color = "white", size = 0.5) +
   draw_label("D", x = 0.95, y = 0.88, size = 6, color = "white") +
   draw_image(readPNG("pictures/HCR_48_DV_NOS_15.68um_insert.png"), 
              scale = 0.35, x = 1, hjust = 1, halign = 1, valign = 0.12) +
@@ -315,37 +316,60 @@ panel_NOS2d_ant_HCR <- ggdraw() + draw_image(readPNG("pictures/HCR-IHC_51_AP_NOS
 panel_NOS3d <- ggdraw() + draw_image(readPNG("pictures/NOS-promotor_3d_acTub_XXum_crop.png")) +
[…]

### 2023-01-01 code/Fig1_sup2.R

@@ -1,0 +1,79 @@
+#Code to assemble Fig1 sup2 of the Jokura et al Platynereis NOS paper
+#2022 June - Kei Jokura, Gaspar Jekely
+
+rm(list = ls(all.names = TRUE)) #will clear all objects includes hidden objects.
+gc() #free up memory and report the memory usage.
+
+# load some packages
+library(tidyverse)
+library(cowplot)
+library(png)
+library(patchwork)
+library(magick)
+
+# assemble figure ---------------------------------------------------------
+arrow_fluo <- data.frame(x1 = 0.95, x2 = 0.95, y1 = 0.75, y2 = 0.85)
+panel_NOS3d_ventr_HCR <- ggdraw() + draw_image(readPNG("pictures/HCR_72_DV_NOS_90um.png")) +
+  draw_label("in situ HCR", x = 0.3, y = 0.99, size = 10) +
+  draw_label("NOS", x = 0.12, y = 0.9, color="magenta", size = 11, fontface="italic") +
+  draw_label("DAPI", x = 0.32, y = 0.9, color="cyan", size = 11, fontface="plain") +
+  draw_line(x = c(0.04, 0.26), y = c(0.08, 0.08), color = "white", size = 0.5) +
+  draw_label(expression(paste("20 ", mu, "m")), x = 0.15, y = 0.11, color = "white", size = 8) +
+  draw_label("A", x = 0.95, y = 0.88, size = 6, color = "white") +
+  draw_label("P", x = 0.95, y = 0.72, size = 6, color = "white") +
+  geom_segment(aes(x = x1, y = y1, xend = x2, yend = y2), data = arrow_fluo, color = "white", 
+               arrow = arrow(ends = "both", type = "closed", length = unit(0.1,"cm")),
+               lineend = "butt",
+               linejoin = "mitre",
+               arrow.fill = "white", size = 0.2) +
+  draw_label("eyespots", x = 0.55, y = 0.68, size = 6, color = "white") +
+  draw_line(x = c(0.3, 0.45), y = c(0.63, 0.67), color = "white", size = 0.2) +
+  draw_line(x = c(0.77, 0.66), y = c(0.65, 0.67), color = "white", size = 0.2) +
+  draw_label("adult eyes", x = 0.55, y = 0.34, size = 6, color = "white") +
[…]

### 2023-01-01 code/Fig2.R

@@ -4,465 +4,87 @@
 # source packages ---------------------------------------------------------
 source("code/Packages_to_load.R")
 
-# read data----------------------------------------------------------------
+# read data ---------------------------------------------------------------
 
-{
-WT_2d <- read_csv("data/2017-08-29_YDisplacement_2d_NNxNN_UV-Side-480nm-395nm-Good.csv")
-NOS11x11_2d <- read_csv("data/2017-08-29_YDisplacement_2d_nn11xnn11_UV-Side-480nm-395nm-Good.csv")
-NOS23x23_2d <- read_csv("data/2017-08-29_YDisplacement_2d_nn23xnn23_UV-Side-480nm-395nm-Good.csv")
-NOS11x23_2d <- read_csv("data/2017-08-29_YDisplacement_2d_nn11xnn23_UV-Side-480nm-395nm-Good.csv")
-
-WT_3d <- read_csv("data/2017-08-29_YDisplacement_3d_NNxNN_UV-Side-480nm-395nm-Good.csv")
-NOS11x11_3d <- read_csv("data/2017-08-29_YDisplacement_3d_nn11xnn11_UV-Side-480nm-395nm-Good.csv")
-NOS23x23_3d <- read_csv("data/2017-08-29_YDisplacement_3d_nn23xnn23_UV-Side-480nm-395nm-Good.csv")
-NOS11x23_3d <- read_csv("data/2017-08-29_YDisplacement_3d_nn11xnn23_UV-Side-480nm-395nm-Good.csv")
-
-L_NAME_00mM_3d <- read_csv("data/2017-08-29_YDisplacement_3d_0mM_L-NAME_UV-Side-480nm-395nm-Good.csv")
-L_NAME_01mM_3d <- read_csv("data/2017-08-29_YDisplacement_3d_0.1mM_L-NAME_UV-Side-480nm-395nm-Good.csv")
-L_NAME_1mM_3d <- read_csv("data/2017-08-29_YDisplacement_3d_1mM_L-NAME_UV-Side-480nm-395nm-Good.csv")
-
-#Import trajectory data focusing on sideUV, 10s before and 30 s after
-
-df_2d_nor <- read_csv("data/221011_sUV_tracking/sUV_tracking_2d.csv")
-df_3d_nor <- read_csv("data/221011_sUV_tracking/sUV_tracking_3d.csv")
-<<<<<<< HEAD
-<<<<<<< HEAD
-
-=======
[…]

### 2023-01-01 code/Fig3.R

@@ -4,184 +4,476 @@
 # source packages ---------------------------------------------------------
 source("code/Packages_to_load.R")
 
-# read data ---------------------------------------------------------------
+# read data----------------------------------------------------------------
 
-DAFFM <- read_csv("data/220213_DAF-FM_long.csv")
+{
+WT_2d <- read_csv("data/2017-08-29_YDisplacement_2d_NNxNN_UV-Side-480nm-395nm-Good.csv")
+NOS11x11_2d <- read_csv("data/2017-08-29_YDisplacement_2d_nn11xnn11_UV-Side-480nm-395nm-Good.csv")
+NOS23x23_2d <- read_csv("data/2017-08-29_YDisplacement_2d_nn23xnn23_UV-Side-480nm-395nm-Good.csv")
+NOS11x23_2d <- read_csv("data/2017-08-29_YDisplacement_2d_nn11xnn23_UV-Side-480nm-395nm-Good.csv")
 
-WTvsNOS23 <- read_csv("data/211121_WTvsNOS23_cPRC.csv")
-WTvsNOS11_cPRC_INNOS <- read_csv("data/211129_WTvsNOS11_cPRC_INNOS (2).csv")
+WT_3d <- read_csv("data/2017-08-29_YDisplacement_3d_NNxNN_UV-Side-480nm-395nm-Good.csv")
+NOS11x11_3d <- read_csv("data/2017-08-29_YDisplacement_3d_nn11xnn11_UV-Side-480nm-395nm-Good.csv")
+NOS23x23_3d <- read_csv("data/2017-08-29_YDisplacement_3d_nn23xnn23_UV-Side-480nm-395nm-Good.csv")
+NOS11x23_3d <- read_csv("data/2017-08-29_YDisplacement_3d_nn11xnn23_UV-Side-480nm-395nm-Good.csv")
 
-# tidying the data -----------------------------------------------------------
+L_NAME_00mM_3d <- read_csv("data/2017-08-29_YDisplacement_3d_0mM_L-NAME_UV-Side-480nm-395nm-Good.csv")
+L_NAME_01mM_3d <- read_csv("data/2017-08-29_YDisplacement_3d_0.1mM_L-NAME_UV-Side-480nm-395nm-Good.csv")
+L_NAME_1mM_3d <- read_csv("data/2017-08-29_YDisplacement_3d_1mM_L-NAME_UV-Side-480nm-395nm-Good.csv")
 
-WT_cPRC_tidy <- WTvsNOS23 %>% 
[…]

### 2023-01-01 code/Fig4.R

@@ -6,12 +6,31 @@ source("code/Packages_to_load.R")
 
 # read data ---------------------------------------------------------------
 
+WTvsNOS23 <- read_csv("data/211121_WTvsNOS23_cPRC.csv")
+WTvsNOS11_cPRC_INNOS <- read_csv("data/211129_WTvsNOS11_cPRC_INNOS (2).csv")
+
 NIT1_MO <- read_csv("data/220214_non-MOvsNIT1-MO1and2.csv")
 NIT2_MO <- read_csv("data/220522_NIT2-MO1and2.csv")
 
 NITGC1_analysis <- read_csv("data/220519_GcG-T2A-NITGC1_norm.csv")
 
 # tidying the data -----------------------------------------------------------
+
+
+WT_cPRC_tidy <- WTvsNOS23 %>% 
+  pivot_longer(cols = -c("frame"),
+               names_to = "genotype",
+               values_to = "intensity") %>%
+  separate(col = c("genotype"), 
+           into = c("genotype", "sample"), sep = "\\...")
+
+WTvsNOS11_cPRC_INNOS_tb <- WTvsNOS11_cPRC_INNOS %>% 
+  pivot_longer(cols = -c("frame"),
+               names_to = "genotype",
+               values_to = "intensity") %>%
+  separate(col = c("genotype"), into = c("genotype", "sample"), sep = "\\...")
+
+
 
 NIT1_MO_tb <- NIT1_MO %>% 
   pivot_longer(cols = -c("frame"),
@@ -30,6 +49,86 @@ NIT1_MO_tb%>%
   count(morphant)
 NIT2_MO_tb%>%
   count(morphant)
+
+
+
+
+
+
+# plot NOS23 cPRC data with geom_smooth ----------------------------------------------
+
+WT_cPRC_tidy$genotype <- factor(WT_cPRC_tidy$genotype, 
+                                levels=c("WT_cPRC", "NOS23_cPRC"))
+
+WT_cPRC_tidy %>%
+  ggplot(aes(x=frame,y=intensity,color=genotype)) +
+  annotate("rect", xmin=51, xmax=90, ymin=-Inf, ymax=Inf, 
+           alpha=0.1, fill="blue") +
+  geom_line(aes(group = sample), size=0.2, alpha=0.2) +
+  geom_smooth(aes(fill = genotype), level = 0.99, size = 1, span = 0.03, 
[…]

### 2023-01-01 code/Fig5.R

@@ -360,30 +360,32 @@ panel_MC <- ggdraw() + draw_image(readPNG("pictures/WTvsNOS11_MC.png"))
 
 layout <- "
 AAAAAABBBCCCFFFFFF
+AAAAAA######FFFFFF
 AAAAAADDDEEEFFFFFF
 ##################
-GGGGGGHHHHHHIIIIII
-JJJJJJKKKKKKLLLLLL"
+GGGGGGIIIIIIKKKKKK
+HHHHHHJJJJJJLLLLLL"
 
 Fig5 <- panel_method + 
   panel_Ca_NOS + panel_IHC_NOS +
   panel_Ca_RGW + panel_IHC_RGW + 
   panel_correlation +
-  panel_cPRC + panel_INNOS + panel_INRGW + 
-  panel_Ser + panel_NIT2MO_INNOS + panel_NIT2MO_INRGW  +  
+  panel_INNOS + panel_NIT2MO_INNOS + 
+  panel_INRGW + panel_NIT2MO_INRGW +
+  panel_Ser + panel_MC +
   patchwork::plot_layout(design = layout, 
-                         heights = c(0.5, 0.5, 0.02, 1, 1)) + #we can change the heights of the rows in our layout (widths also can be defined)
+                         heights = c(0.4, 0.02, 0.4, 0.05, 0.8, 0.8)) + #we can change the heights of the rows in our layout (widths also can be defined)
   patchwork::plot_annotation(tag_levels = 'A') &  #we can change this to 'a' for small caps or 'i' or '1'
   ggplot2::theme(plot.tag = element_text(size = 12, 
-          face='plain')) #or 'bold', 'italic'
-
-ggsave("figures/Fig5.pdf", limitsize = FALSE, 
-       units = c("px"), Fig5, width = 3200, height = 2300)  
+          face='bold')) #or 'bold', 'italic'
+ 
 
 ggsave("figures/Fig5.png", limitsize = FALSE, 
-       units = c("px"), Fig5, width = 3200, height = 2300, bg='white')  
+       units = c("px"), Fig5, width = 3200, height = 2500, bg='white')  
 
 
+ggsave("figures/Fig5.pdf", limitsize = FALSE, 
+       units = c("px"), Fig5, width = 3200, height = 2500) 
 
 
 

### 2023-01-04 text/Manuscript.Rmd
edit a part of manuscripts

@@
 ![**Figure 1.
-Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** (A) SEM image of a 3-day-old Platynereis larva.
-(B and C) Five types of neurons (cPRC, INNOS, INRGWa, Ser-h1 and MC) in the cPRC circuit reconstructed in the whole-body transmission electron microscopy (ssTEM) dataset of a 3-day-old larva.
+Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** **(A)** SEM image of a 3-day-old Platynereis larva.
+**(B and C)** Five types of neurons (cPRC, INNOS, INRGWa, Ser-h1 and MC) in the cPRC circuit reconstructed in the whole-body transmission electron microscopy (ssTEM) dataset of a 3-day-old larva.
+Ciliated cells labelled with (B) are shown in grey.
 The INNOS on the ventral side labelled with (C) are shown in purple.
 Axons and dendrites appear as lines and cell body positions are represented by spheres.
-(D) Wiring diagram of the cPRC circuit.
-(E-G) Expression analysis of the NOS gene (magenta) using in situ HCR.
-(E) Dorsal view of the larva at 3-day-old.
-(F and G) Anterior view of the larva at 3 (F) and 2-day-old (G).
-(F) Insert: showing two NOS cells on each side.
-(G) Co-stained image using acetylated α-tubulin antibody (acTub: green).
-(H) Immunostaining with anti-HA antibody against HA tagged palmitoylated-tdTomato expressed under the upstream of the start site of NOS (NOSp::palmi-3xHA: magenta).
+**(D)** Expression analysis of the NOS gene (magenta) using in situ HCR.
+Anterior view of the larva at 2-day-old.
+Co-stained image using acetylated α-tubulin antibody (acTub: green).
+**(E)** Immunostaining with anti-HA antibody against HA tagged palmitoylated-tdTomato expressed under the upstream of the start site of NOS (NOSp::palmi-3xHA: magenta).
 Co-staining with acetylated α-tubulin antibody as marker for cilia of cPRC and axonal scaffold (acTub: green).
-Views of larvae at 3-day-old from the anterior side.](../figures/Fig1.png){width="100%"}
+Views of larvae at 3-day-old from the anterior side.
+**(F)** Wiring diagram of the cPRC circuit.
+Hexagons and arrows represent neuron groups and their synaptic connections, respectively.
+Numbers inside hexagons indicate the number of neurons grouped in each node.
+Interactions with less than five synapses were filtered out in this display.
+Arrow line thickness (synapse weight) is equal to the common logarithm of the number of synapses (scale is shown on the bottom left of the panel).](../figures/Fig1.png){width="100%"}
 
@@
 The data are shown in 30 s bins.
-
 (E and F) Overlaid trajectories for WT (left, n=20) and NOS mutant (right, n=31) 3-day-old larvae.
 Grey shows the trajectories before UV stimulation and purple shows the trajectories after UV stimulation.
-
-(E) The temporal changes in the vertical position of the WT and mutant larvae before and after UV stimulation are shown. The starting points of each larval trajectory are set to 0.
-
-(F) The temporal changes in the distance traveled of the WT and mutant larvae before and after UV stimulation are shown.
-
-(E and F) After UV stimulation is indicated by purple squares.](../figures/Fig2.png){width="100%"}
+(E) The temporal changes in the vertical position of the WT and mutant larvae before and after UV stimulation are shown.
+The starting points of each larval trajectory are set to 0.(F) The temporal changes in the distance traveled of the WT and mutant larvae before and after UV stimulation are shown.(E and F) After UV stimulation is indicated by purple squares.](../figures/Fig2.png){width="100%"}
 
@@
 (C and D) Localisation analysis using NIT-GC1 and NIT-GC2 antibodies.
-Green shows co-staining with acetylated α-tubulin antibody (ac α-tub).
-
-(E-G) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in control (non-injection) and NIT-GC1 and NIT-GC2 knockdown by morpholino (MO1 and 2) larvae.
-
-(H) The domain structure of NIT-GC1. Grey squares indicate the transmembrane domain.
-(I) A schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control.
+Green shows co-staining with acetylated α-tubulin antibody (ac α-tub).(E-G) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in control (non-injection) and NIT-GC1 and NIT-GC2 knockdown by morpholino (MO1 and 2) larvae.(H) The domain structure of NIT-GC1.
+Grey squares indicate the transmembrane domain.(I) A schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control.
 Green cGull shows cGMP-dependent green fluorescence.
@@
 IN-NOS, IN-RGWa and Ser-h1 identified from the numbered nuclear positions are shown.
-(F) Correlation (Pearson's r) map of neuronal activity of the cPRC, IN-NOS, IN-RGWa and Ser-h1 neurons.
-
-(E) Changes over time in GCaMP6s intensities in IN-NOS when cPRC is stimulated with UV in WT (dark purple) and NOS mutant (light purple) larvae.
+(F) Correlation (Pearson's r) map of neuronal activity of the cPRC, IN-NOS, IN-RGWa and Ser-h1 neurons.(E) Changes over time in GCaMP6s intensities in IN-NOS when cPRC is stimulated with UV in WT (dark purple) and NOS mutant (light purple) larvae.
 (F) Changes over time in GCaMP6s intensities in IN-NOS when cPRC is stimulated with UV in NIT-GC2 morphants larvae.
@@
 
-![**Figure 5.
+![**Figure 6.
 Gene expressions and signalling mechanisms between cPRC, INNOS and INRGWa.** (A) Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq.
@@
 The colours represent the normal logarithm of the number of transcripts in the cells expressing the gene.
-(B) Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback.](../figures/Fig5.png){width="100%"}
+(B) Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback.](../figures/Fig6.png){width="100%"}
 

### 2023-01-08 text/Manuscript.Rmd
edit text

@@
 Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** **(A)** SEM image of a 3-day-old Platynereis larva.
-**(B and C)** Five types of neurons (cPRC, INNOS, INRGWa, Ser-h1 and MC) in the cPRC circuit reconstructed in the whole-body transmission electron microscopy (ssTEM) dataset of a 3-day-old larva.
+**(B, C)** Five types of neurons (cPRC, INNOS, INRGWa, Ser-h1 and MC) in the cPRC circuit reconstructed in the whole-body transmission electron microscopy (ssTEM) dataset of a 3-day-old larva.
 Ciliated cells labelled with (B) are shown in grey.
@@
 Views of larvae at 3-day-old from the anterior side.
+Insert: One of the reconstracted INNOS is shown from the anterior side.
 **(F)** Wiring diagram of the cPRC circuit.
-Hexagons and arrows represent neuron groups and their synaptic connections, respectively.
-Numbers inside hexagons indicate the number of neurons grouped in each node.
-Interactions with less than five synapses were filtered out in this display.
-Arrow line thickness (synapse weight) is equal to the common logarithm of the number of synapses (scale is shown on the bottom left of the panel).](../figures/Fig1.png){width="100%"}
+Hexagons and arrows represent cell groups and their synaptic connections, respectively.
+Numbers inside hexagons indicate the number of cells grouped in each node.
+Arrow line thickness (synapse weight) is equal to the number of synapses.](../figures/Fig1.png){width="100%"}
 
@@
 ![**Figure 2.
-Larvae with inhibited NOS have shorter ciliary arrest times and a delayed response to UV-avoidance.** (A) Vertical swimming in larvae treated with NOS inhibitors, L-NAME at 3-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
-The data are shown in 30 s bins.
-(B) Top: The domain and Exon/Intron structure of NOS.
-Bottom: Close-up region showing the genomic locus of NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9.
-The generated mutants (NOSΔ11, NOSΔ23) are also shown.
-Pink indicates target sites.
-Gray shows PAM sequences, red shows stop codons.
-(C and D) Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23) larvae at 2-day-old (C) and 3-day-old (D) stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
-The data are shown in 30 s bins.
-(E and F) Overlaid trajectories for WT (left, n=20) and NOS mutant (right, n=31) 3-day-old larvae.
-Grey shows the trajectories before UV stimulation and purple shows the trajectories after UV stimulation.
-(E) The temporal changes in the vertical position of the WT and mutant larvae before and after UV stimulation are shown.
-The starting points of each larval trajectory are set to 0.(F) The temporal changes in the distance traveled of the WT and mutant larvae before and after UV stimulation are shown.(E and F) After UV stimulation is indicated by purple squares.](../figures/Fig2.png){width="100%"}
+NO produced by UV stimulation to cPRCs.** **(A)** A NO-positive cell visualised by DAF-FM (green) is shown (dashline).
+The white line indicates the outer frame of the larva.
+All subsequent images are of 2-day-old larvae.
+**(B)** Intensity changes before (upper) and during 405 nm light stimulation (lower) in a cell visualised by DAF-FM.
+**(C)** The intensity of DAF-FM changes over time when cPRC (green) and other cells (ctr: gray) are stimulated with UV/violet (405 nm) light.
+Purple squares indicate the timing of 405 nm light stimulation, the transparent lines show the optical responses to individual intensities, each intensities are normalized (ΔF/F0) and the dark trace lines and the shaded bands show the mean value and the standard error range of the mean.](../figures/Fig2.png){width="100%"}
 
@@
 ![**Figure 3.
-NO produced by UV stimulation to cPRCs induces a depolarising response in cPRC.** (A) A NO-positive cell visualised by DAF-FM is shown (dashline).
-The white dashed line indicates the outer frame of the larva.
-All subsequent images are of 2-day-old larvae.
-(B) Intensity changes before and during 405 nm light stimulation in a cell visualised by DAF-FM.
-(C) The intensity of DAF-FM changes over time when cPRC (green) and other cells (NC) (gray) are stimulated with UV/violet (405 nm) light.
-Purple squares indicate the timing of 405 nm light stimulation, the transparent lines show the optical responses to individual intensities, each intensities are normalized (ΔF/F0) and the dark trace lines and the shaded bands show the mean value and the standard error range of the mean.
-(D and E) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in WT (dark yellow) and NOS mutant (D: NOSΔ11/Δ11, E: NOSΔ23/Δ23) (light yellow) larvae.](../figures/Fig3.png){width="100%"}
+Larvae with inhibited NOS have a delayed response to UV-avoidance.** **(A)** Top: The domain and Exon/Intron structure of NOS.
+Bottom: Close-up region showing the genomic locus of NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9.
+The generated mutants (NOSΔ11, NOSΔ23) are also shown.
+Pink indicates target sites.
+Gray shows PAM sequences, red shows stop codons.
+**(B)** Overlaid trajectories for WT (left, n=20) and NOS mutant (right, n=31) 3-day-old larvae.
+Grey shows the trajectories before UV stimulation and purple shows the trajectories after UV stimulation.
+**(C)** The temporal changes in the vertical position of the WT and mutant larvae before and after UV stimulation are shown.
+The starting points of each larval trajectory are set to 0.
+After UV stimulation is indicated by purple squares.
+**(D)** Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23) larvae at 3-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
+The data are shown in 30 s bins.
+**(E)** Vertical swimming in larvae treated with NOS inhibitors, L-NAME at 3-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
+The data are shown in 30 s bins.
+.](../figures/Fig3.png){width="100%"}
 
@@
 ![**Figure 4.
-cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.** (A) Expression analysis of the NIT-GC1 gene (magenta) using in situ HCR.
+cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.** (D and E) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in WT (dark yellow) and NOS mutant (D: NOSΔ11/Δ11, E: NOSΔ23/Δ23) (light yellow) larvae.
+(A) Expression analysis of the NIT-GC1 gene (magenta) using in situ HCR.
 White arrowheads indicate cells expressing NIT-GC1.Co-staining image using acetylated α-tubulin antibody (ac α-tub: green).

### 2023-01-08 text/Supplements.Rmd
edit a part of manuscripts
edit text

@@
 
+![**Figure 1---figure supplement 2.** Expression analysis of the NOS gene (magenta) using in situ HCR using the larva at 3-day-old in dorsal view (A) and anterior view (B).
+(B) Insert: showing two NOS cells on each side.](../figures/Fig1_sup2.png){width="100%"}
+
+![**Figure 3---figure supplement 1.** Expression analysis of the NOS gene (magenta) using in situ HCR using the larva at 3-day-old in dorsal view (A) and anterior view (B).
+(B) Insert: showing two NOS cells on each side.(F) The temporal changes in the distance traveled of the WT and mutant larvae before and after UV stimulation are shown.(E and F) After UV stimulation is indicated by purple.](../figures/Fig3_sup1.png){width="100%"}
+
 ![**Figure 4---figure supplement 1.** Phylogenetic tree of guanylate cyclase by maximum likelihood (ML).

### 2023-01-08 code/Fig3.R

@@ -121,7 +121,7 @@ WTvsNOS_2d %>%
   theme_plot +
   labs(fill = "genotype", x = "30 sec bins", y = "Vertical displacement (mm/s)")+
   scale_y_continuous(limits = c(-0.6, 0.65)) +
-  scale_x_discrete(labels=c("30", "60", "90","120", 
+  scale_x_discrete(labels=c("30", "60", "90", 
                             "30", "60", "90","120",
                             "30", "60", "90","120",
                             "30", "60", "90","120",
@@ -172,8 +172,8 @@ WTvsNOS_3d %>%
   theme_plot +
   labs(fill = "genotype", x = "30 sec bins", y = "Vertical displacement (mm/s)")+
   scale_y_continuous(limits = c(-0.81, 0.5)) +
-  scale_x_discrete(labels=c("30", "60", "90","120", 
-                           "30", "60", "90","120",
+  scale_x_discrete(labels=c("30", "60", "90", 
+                           "30", "60", "90",
                            "30", "60", "90","120",
                            "30", "60", "90","120",
                            "30", "60", "90","120",

### 2023-01-10 text/Manuscript.Rmd
edit figure legends

@@
 **(E)** Vertical swimming in larvae treated with NOS inhibitors, L-NAME at 3-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
-The data are shown in 30 s bins.
-.](../figures/Fig3.png){width="100%"}
+The data are shown in 30 s bins.](../figures/Fig3.png){width="100%"}
 
@@
 ![**Figure 4.
-cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.** (D and E) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in WT (dark yellow) and NOS mutant (D: NOSΔ11/Δ11, E: NOSΔ23/Δ23) (light yellow) larvae.
-(A) Expression analysis of the NIT-GC1 gene (magenta) using in situ HCR.
-White arrowheads indicate cells expressing NIT-GC1.Co-staining image using acetylated α-tubulin antibody (ac α-tub: green).
-(B) Co-expression analysis image of the c-opsin1 gene (green).
-Black arrowheads indicate cells expressing c-opsin1 differently from cPRC.
-(C and D) Localisation analysis using NIT-GC1 and NIT-GC2 antibodies.
-Green shows co-staining with acetylated α-tubulin antibody (ac α-tub).(E-G) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in control (non-injection) and NIT-GC1 and NIT-GC2 knockdown by morpholino (MO1 and 2) larvae.(H) The domain structure of NIT-GC1.
-Grey squares indicate the transmembrane domain.(I) A schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control.
+cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.** **(A, B)** Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in WT (dark yellow) and NOS mutant (A: NOSΔ11/Δ11, B: NOSΔ23/Δ23) (light yellow) larvae.
+**(C, D)** Expression analysis of the NIT-GC1 (C) and NIT-GC2 (D) gene (magenta) using in situ HCR.
+Co-staining image using acetylated α-tubulin antibody (ac α-tub: green).
+**(E, F)** Localisation analysis using NIT-GC1 and NIT-GC2 antibodies.
+Green shows co-staining with acetylated α-tubulin antibody (ac α-tub).
+**(G, H)** Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in NIT-GC1 (G) and NIT-GC2 (H) knockdown by morpholino (MO1 and 2) larvae.
+**(I)** The domain structure of NIT-GC1.
+Grey squares indicate the transmembrane domain.
+**(J)** A schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control.
 Green cGull shows cGMP-dependent green fluorescence.
-(J-M) The intensity changes in cGMP fluorescence over time for the four conditions are shown.
+**(K-N)** The intensity changes in cGMP fluorescence over time for the four conditions are shown.
 Gray traces show the optical responses to individual intensities (n \> 6).
@@
 In contrast, calcium signals increased in INRGW cells in wild-type larvae during the second activation phase of cPRCs, a response that was lacking in *NOS*-mutants (Figure 4G and H).
-This suggests that there is no circuit output to the Ser-h1 cells and the ciliary band in *NOS*-mutants, consistent with the observed behavioural defects(Figure 4I).
+This suggests that there is no circuit output to the Ser-h1 cells and the ciliary band in *NOS*-mutants, consistent with the observed behavioural defects (Figure 4I).
 

### 2023-01-10 text/Supplements.Rmd
edit figure legends

--- 2022-12-02 text/Supplements.Rmd
@@
 
+![**Figure 1---figure supplement 2.** Expression analysis of the NOS gene (magenta) using in situ HCR using the larva at 3-day-old in dorsal view (A) and anterior view (B).
+(B) Insert: showing two NOS cells on each side.](../figures/Fig1_sup2.png){width="100%"}
+
+![**Figure 3---figure supplement 1.** **(A)** Overlaid trajectories for WT (left, n=20) and NOS mutant (right, n=31) 2-day-old larvae.
+Grey shows the trajectories before UV stimulation and purple shows the trajectories after UV stimulation.
+**(B)** The temporal changes in the vertical position of the WT and mutant 2-days-old larvae before and after UV stimulation are shown.
+The starting points of each larval trajectory are set to 0.
+After UV stimulation is indicated by purple squares.
+**(C)** Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23) larvae at 2-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
+The data are shown in 30 s bins.
+**(D, E)** The temporal changes in the distance traveled of the WT and mutant in 2-day (D) and 3-day-old (E) larvae before and after UV stimulation are shown.](../figures/Fig3_sup1.png){width="100%"}
+
 ![**Figure 4---figure supplement 1.** Phylogenetic tree of guanylate cyclase by maximum likelihood (ML).
@@
 
-![**Figure 4---figure supplement 2.** (A) Immunostaining with NOS antibodies.
-(B) Co-expression analysis image with the RYamide proneuropeptide gene (RYa-pNP: green), which is highly expressed in INNOS.
-White arrowheads indicate co-expressed INNOS.
-Black arrowheads indicate larval eyes.
-White dashed lines indicate the outer frame of the larva.
-(A) Co-expression analysis image of the MLD-pedal2 amide proneuropeptide gene (MLD-pNP: green).
-Black arrowheads indicate cells different from cPRC.
-(B) Co-localisation analysis using NIT-GC1 and NOS antibodies.](../figures/Fig4_sup2.png){width="100%"}
+![**Figure 4---figure supplement 2.** **(A)** Co-expression analysis image of the NIT-GC1 (magenta) and MLD-pedal2 amide proneuropeptide gene (MLD: green).
+Anterior view of the larva at 2-day-old.
+**(B, C)** Expression analysis of the NIT-GC2 gene (magenta) using in situ HCR.
+Anterior (B) and posterior (C) views of the larva at 3-day-old.
+**(D, E)** Localisation analysis for NIT-GC1 (B) and NIT-GC2 (C) morphant using NIT-GC1 and NIT-GC2 antibodies.
+Green shows co-staining with acetylated α-tubulin antibody (acTub).
+Anterior view of the larva at 2-day-old.
+**(F)** Co-localisation analysis using NIT-GC1 (magenta) and NOS (green) antibodies.
+Anterior view of the larva at 2-day-old.](../figures/Fig4_sup2.png){width="100%"}

### 2023-01-10 text/Manuscript.Rmd
Figure legend_2

@@
 ![**Figure 5.
-NO produced by UV stimulation to cPRCs induces a depolarising response in cPRC via NIT-GC1.** (A) Schematic diagram of the immunostaining procedure after calcium imaging.
-(B and D) The position of the cPRC in calcium imaging with GCaMP6s.
+NO produced by UV stimulation to cPRCs induces a depolarising response in cPRC via NIT-GC1.** **(A)** Schematic diagram of the immunostaining procedure after calcium imaging.
+**(B, D)** The position of the cPRC in calcium imaging with GCaMP6s.
 Cells around the cPRC are numbered according to the position of the nuclei.
-(C and E) Immunostaining image using RYamide (C) or RGWamide and 5-HT antibodies (red) (E) and DAPI (cyan) in the same region as (B) or (D).
+**(C, E)** Immunostaining images using RYamide (C) or RGWamide and 5-HT antibodies (E) and DAPI (cyan) in the same region as (B) or (D).
 IN-NOS, IN-RGWa and Ser-h1 identified from the numbered nuclear positions are shown.
-(F) Correlation (Pearson's r) map of neuronal activity of the cPRC, IN-NOS, IN-RGWa and Ser-h1 neurons.(E) Changes over time in GCaMP6s intensities in IN-NOS when cPRC is stimulated with UV in WT (dark purple) and NOS mutant (light purple) larvae.
-(F) Changes over time in GCaMP6s intensities in IN-NOS when cPRC is stimulated with UV in NIT-GC2 morphants larvae.
-(H) Changes over time in GCaMP6s intensities in IN-RGWa when cPRC is stimulated with UV in WT (dark blue) and NOS mutant (light blue) larvae.
-(I) Changes over time in GCaMP6s intensities in Ser-h1 when cPRC is stimulated with UV in WT (dark red) and NOS mutant (light red) larvae.](../figures/Fig5.png){width="100%"}
+**(F)** Correlation (Pearson's r) map of neuronal activity of the cPRC, IN-NOS, IN-RGWa and Ser-h1 neurons.
+**(G)** Changes over time in GCaMP6s intensities in IN-NOS when cPRC is stimulated with UV in WT (dark purple) and NOSΔ11/Δ11 mutant (light purple) larvae.
+**(H)** Changes over time in GCaMP6s intensities in IN-NOS when cPRC is stimulated with UV in NIT-GC2 morphants (MO1 and MO2: pink and light pink) larvae.
+**(I)** Changes over time in GCaMP6s intensities in IN-RGWa when cPRC is stimulated with UV in WT (dark blue) and NOSΔ23/Δ23 mutant (light blue) larvae.
+**(J)** Changes over time in GCaMP6s intensities in IN-RGWa when cPRC is stimulated with UV in NIT-GC2 morphants (MO1 and MO2: blue and light blue).
+**(K, L)** Changes over time in GCaMP6s intensities in Ser-h1 and MC when cPRC is stimulated with UV in WT (Ser-h1 and MC: dark red and brown) and NOSΔ11/Δ11 mutant (Ser-h1 and MC: light red and brown) larvae.](../figures/Fig5.png){width="100%"}
 
@@
 ![**Figure 6.
-Gene expressions and signalling mechanisms between cPRC, INNOS and INRGWa.** (A) Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq.
+Gene expressions and signalling mechanisms between cPRC, INNOS and INRGWa.** **(A)** Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq.
 The size of the dots is expressed in proportion to the percentage of cells expressing that gene relative to all cells.
 The colours represent the normal logarithm of the number of transcripts in the cells expressing the gene.
-(B) Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback.](../figures/Fig6.png){width="100%"}
+**(B)** Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback.](../figures/Fig6.png){width="100%"}
 

### 2023-01-10 text/Supplements.Rmd
Figure legend_2

@@
 Guanylate cyclase-coupled receptor and soluble guanylate cyclases (sGC) as outgroups.
-Guanylate cyclases with NIT domains are found in most animal phyla except Porifera, Ctenophora, Urochordata and Chordata.](../figures/Fig4_sup1.png){width="100%"}
+Guanylate cyclases with NIT domains are found in most animal phyla except Porifera, Ctenophora, Urochordata and Chordata.
+Dot plot of Platynereis NIT-GC genes (columns) expressed in cPRC, INNOS and INRGWa (rows) using single cell RNA-Seq.
+The size of the dots is expressed in proportion to the percentage of cells expressing that gene relative to all cells.
+The colours represent the normal logarithm of the number of transcripts in the cells expressing the gene.](../figures/Fig4_sup1.png){width="100%"}
 

### 2023-01-10 text/Manuscript.Rmd
Fig5 edit

@@
 
-**CRISPR-Cas9 Design and Microinjection**
-
-Before designing the small guide RNA (sgRNA) for the sgRNA:Cas9 nuclease, splice sites and polymorphic sites in our laboratory culture were identified to avoid them.
+**CRISPR-Cas9 Design and Microinjection** Before designing the small guide RNA (sgRNA) for the sgRNA:Cas9 nuclease, splice sites and polymorphic sites in our laboratory culture were identified to avoid them.
 The sgRNA targeted the third exon of Platynereis dumerilii NOS (target site: 5'-GGGCAATACTGGCTCCACTC-3').
@@
 
-**NOS sequencing and genotyping**
-
-For genotyping of the NOS locus, genomic DNA was isolated from single larvae, groups of 6-20 larvae, or from the tails of adult worms.
+**NOS sequencing and genotyping** For genotyping of the NOS locus, genomic DNA was isolated from single larvae, groups of 6-20 larvae, or from the tails of adult worms.
 The DNA was amplified by PCR (primers: 5'-GGTTCATTGGTTTCGATAACATTGCGG-3', 5'-CAGAGTCGATCAGTCTGCATATCTCCA-3') with the dilution protocol of the Phusion Human Specimen Direct PCR Kit (Thermo Scientific).
@@
 
-**Vertical column setup for pharmacology**
-
-Freely swimming 2-day-old larvae were recorded in 5 adjacent vertical columns (each 31 mm x 10 mm x 144 mm water height) with a DMK 21BF03 camera (The Imaging Source) at 30 frames/s.
+**Vertical column setup for pharmacology** Freely swimming 2-day-old larvae were recorded in 5 adjacent vertical columns (each 31 mm x 10 mm x 144 mm water height) with a DMK 21BF03 camera (The Imaging Source) at 30 frames/s.
 The columns were illuminated laterally with red (633 nm) light-emitting diodes (LEDs).
@@
 
-**Vertical column setup for measuring photoresponses**
-
-Photoresponses of larvae of different ages were assayed in a vertical Plexiglas column (31 mm x 10 mm x 160 mm water height).
+**Vertical column setup for measuring photoresponses** Photoresponses of larvae of different ages were assayed in a vertical Plexiglas column (31 mm x 10 mm x 160 mm water height).
 The column was illuminated from top with light from a monochromator (Polychrome II, Till Photonics).
@@
 
-**Comparing behavior of wildtype and NOS-knockout 3-day-old larvae**
-
-To compare the behavior of wildtype and NOS-knockout larvae at 3 days in the vertical column, the larvae were mixed and left in the dark for 5 min.
+**Comparing behavior of wildtype and NOS-knockout 3-day-old larvae** To compare the behavior of wildtype and NOS-knockout larvae at 3 days in the vertical column, the larvae were mixed and left in the dark for 5 min.
 The larvae were recorded for 1 min in the dark followed by exposure to collimated cyan (480 nm) light from the top of the column for 2 min, then 2 min darkness, and finally collimated UV (395 nm) light from the top of the column for 2 min.
@@
 
-**In situ HCR**
-
-Larvae were fixed and treated with Proteinase K, according to the conventional WMISH protocol [@Tessmar-Raible2005], with fixation in 4% paraformaldehyde/ PTW (PBS with 0.05% Tween20) for 2 hr at room temperature, and Proteinase K treatment in 100 µg/ml Proteinase K/ PTW for 3 min [@Tessmar-Raible2005].
+**In situ HCR** Larvae were fixed and treated with Proteinase K, according to the conventional WMISH protocol [@Tessmar-Raible2005], with fixation in 4% paraformaldehyde/ PTW (PBS with 0.05% Tween20) for 2 hr at room temperature, and Proteinase K treatment in 100 µg/ml Proteinase K/ PTW for 3 min [@Tessmar-Raible2005].
 Specifically, for the HCR protocol, samples were processed in 1.5 ml tubes.
@@
 
-**Immunohistochemistry**
-
-Whole-mount immunostaining of 2 day old Platynereis larvae fixed with 4% paraformaldehyde were carried out using primary antibodies raised against NIT-GC1, NIT-GC2, NOS, RYamide neuropeptide, RGWamide neuropeptide in rabbit, plus a commercial antibody raised against acetylated tubulin in mouse (Sigma T7451).
+**Immunohistochemistry** Whole-mount immunostaining of 2 day old Platynereis larvae fixed with 4% paraformaldehyde were carried out using primary antibodies raised against NIT-GC1, NIT-GC2, NOS, RYamide neuropeptide, RGWamide neuropeptide in rabbit, plus a commercial antibody raised against acetylated tubulin in mouse (Sigma T7451).
 The synthetic peptides contained an N-terminal Cys that was used for coupling during purification.
@@
 
-**Calcium imaging**
-
-For calcium imaging, 49--55 hpf larvae were used.
+**Calcium imaging** For calcium imaging, 49--55 hpf larvae were used.
 Experiments were performed at room temperature and larvae were immobilised by being embedded in 2.5% agarose filtered artificial seawater between a slide and coverslip spaced with adhesive tape.
@@
 
-**Cell culture experiment**
-
-Green cGull was used for the cGMP assay [@Matsuda2017].
+**Cell culture experiment** Green cGull was used for the cGMP assay [@Matsuda2017].
 A full-length Pdum-NIT-GC1 and -NIT-GC2 coding sequences were amplified by PCR starting from a Platynereis dumerilii cDNA library and cloned into the pcDNA3.1(+) vector using the T2A self-cleaving sequence.

### 2023-01-10 code/Fig5.R

@@ -321,8 +321,8 @@ panel_IHC_NOS <- ggdraw() + draw_image(readPNG("pictures/n51hpf_WT_e-1_IHC_point
   draw_label("RYa", x = 0.15, y = 0.92, hjust = 0, color='red',size = 10, fontface='bold') +
   draw_label("DAPI", x = 0.15, y = 0.82, hjust = 0, color='cyan', size = 10, fontface='bold') +
 #  draw_line(x = c(0.05, 0.45), y = c(0.1, 0.1), color = "white", size = 0.5) +
-  draw_label("IN-NOS", x = 0.45, y = 0.35, color='white', size = 10, fontface='plain') +
-  draw_line(x = c(0.32, 0.45), y = c(0.5, 0.4), color = "white", size = 0.4)
+  draw_label("IN-NOS", x = 0.3, y = 0.3, color='white', size = 10, fontface='plain') +
+  draw_line(x = c(0.34, 0.3), y = c(0.54, 0.35), color = "white", size = 0.4)
 
 panel_Ca_RGW <- ggdraw() + draw_image(readPNG("pictures/n54hpf_WT_e-1_GCaMP6s.png")) +
   draw_label("GCaMP6s", x = 0.15, y = 0.92, hjust = 0, color='white', size = 10, fontface='bold') +
@@ -334,10 +334,10 @@ panel_IHC_RGW <- ggdraw() + draw_image(readPNG("pictures/n54hpf_WT_e-1_RGWa.png"
   draw_label("RGWa, 5-HT", x = 0.15, y = 0.92, hjust = 0, color='red',size = 10, fontface='bold') +
   draw_label("DAPI", x = 0.15, y = 0.82, hjust = 0, color='cyan', size = 10, fontface='bold') +
 #  draw_line(x = c(0.05, 0.45), y = c(0.1, 0.1), color = "white", size = 0.5) +
-  draw_label("IN-RGWa", x = 0.2, y = 0.45, color='white', size = 10, fontface='plain') +
-  draw_line(x = c(0.2, 0.44), y = c(0.4, 0.3), color = "white", size = 0.4) +
-  draw_label("Ser-h1", x = 0.86, y = 0.55, color='white', size = 10, fontface='plain') +
-  draw_line(x = c(0.85, 0.87), y = c(0.75, 0.59), color = "white", size = 0.4)
[…]

### 2023-01-13 text/Manuscript.Rmd
edit model

--- 2023-01-10 text/Manuscript.Rmd
@@
 We also tested phototaxis, by exposing larvae to 480 nm directional and collimated light from the top of the column.
-As previously shown, wild-type larvae swim downward within 30 sec following non-directional UV/violet light stimulation [@veraszto2018].
+As previously shown, wild-type larvae swim downward within 30 sec following non-directional UV/violet light stimulation [@Veraszto2018].
 In contrast, homozygous *NOS*-mutant larvae showed a strongly diminished UV-avoidance response (Figure 2B and C).
-This phenotype is similar to the defective UV-avoidance of *c-opsin1* mutant larvae and reveals a requirement for NOS in UV-avoidance behaviour [@veraszto2018].
+This phenotype is similar to the defective UV-avoidance of *c-opsin1* mutant larvae and reveals a requirement for NOS in UV-avoidance behaviour [@Veraszto2018].
 *NOS*-mutant larvae also showed reduced phototactic behaviour, suggesting a function in the visual eyes mediating phototaxis [@Randel2014a], where *NOS* is also expressed (Figure 1).
@@
 To characterise the role of NIT-GCs within the cPRC during UV stimulation in more detail, the calcium sensor GCaMP6s were ubiquitously expressed in larvae in which NIT-GCs were knocked down using morpholino to examine changes in calcium levels in the cPRC.
-Local stimulation of the ciliated part of the cPRC with 405 nm light causes a transient decrease in calcium levels in the cPRC (hyperpolarisation), followed by an increase in calcium levels (depolarisation) when the light is switched off [@veraszto2018].
+Local stimulation of the ciliated part of the cPRC with 405 nm light causes a transient decrease in calcium levels in the cPRC (hyperpolarisation), followed by an increase in calcium levels (depolarisation) when the light is switched off [@Veraszto2018].
 In NIT-GC1 morphant larvae, 405 nm light stimulation caused a decrease in calcium levels in the cPRC, but the increase in calcium levels that occurred after the light stimulation was stopped was strongly suppressed (Figure 3E).
@@
 
-**NIT-GC Identification and Phylogenetic Analysis** To identify GPR173 receptors, we obtained a database of vertebrate SREB sequences, including GPR173, GPR85 and GPR27 from (Breton et al.
+**NIT-GC Identification and Phylogenetic Analysis**
+
+To identify GPR173 receptors, we obtained a database of vertebrate SREB sequences, including GPR173, GPR85 and GPR27 from (Breton et al.
 2021).
@@
 
-**Single-cell analysis** I used Achim et al 2015 for the single-cell data.
+**Single-cell analysis**
+
+I used Achim et al 2015 for the single-cell data.
 In fact, in Liz's 2018 paper, she used 107 cells as neurons by removing duplicates from Achim's single-cell data, so I used those cells.
@@
 After created the data in Python, plotted it using R dot plots.
-
 RPKM?

### 2023-01-24 text/Manuscript.Rmd
edit all figures

@@
 In the mammalian retina, NO is an essential signalling molecule in all retinal cells.
-NO produced by nitric oxide synthase (NOS) in amacrine cells, one of the retinal cells, activates soluble guanylate cyclase (sGC) expressed in cone bipolar cells, which produces cGMP and regulates downstream signalling.
+NO produced by nitric oxide synthase (NOS) in amacrine cells, one of the retinal cells, activates *soluble guanylate cyclase (sGC)* expressed in cone bipolar cells, which produces cGMP and regulates downstream signalling.
 We therefore hypothesise that NO-mediated neural signalling may be involved in the photoreceptor circuits that control UV avoidance behaviour.
-To test this hypothesis we used the annelid *Platynereis dumerilii*, which conserved vertebrate-type ciliary photoreceptors (cPRCs) [@Arendt2004].
-The cPRC expresses c-opsin1, which absorbs UV radiation, is involved in melatonin production and regulates circadian swimming behaviour [@Tosches2014; @tsukamoto2017].
-Furthermore, we previously reported abnormal UV avoidance behaviour in c-opsin1 knockout larvae [@Veraszto2018].
+To test this hypothesis we used the annelid *Platynereis dumerilii* **(Figure 1A)**, which conserved vertebrate-type ciliary photoreceptors (cPRCs) [@Arendt2004].
+The cPRC expresses *c-opsin1*, which absorbs UV radiation, is involved in melatonin production and regulates circadian swimming behaviour [@Tosches2014; @tsukamoto2017].
+Furthermore, we previously reported abnormal UV avoidance behaviour in *c-opsin1* knockout larvae [@Veraszto2018].
 The cPRCs express the UV-absorbing c-opsin1 and mediate UV avoidance in the larvae [@Veraszto2018].
@@
 We identified a single *nitric oxide synthase* (*NOS*) gene in the *Platynereis dumerilii* genome and transcriptome data.
-Phylogenetic analysis of NOS proteins indicate that *Platynereis* NOS belongs to an orthology group of metazoan NOS sequences (Figure 1---figure supplement 1).
+Phylogenetic analysis of NOS proteins indicate that *Platynereis* NOS belongs to an orthology group of metazoan NOS sequences **(Figure 1---figure supplement 1)**.
 To characterise the expression pattern of *NOS* we used in situ hybridization chain reaction (HCR) and transient transgenesis.
-In two- and three-day-old larvae, we detected *NOS* expression in four cells (two of them weakly expressing) in the apical organ region (Figure 1A-D).
-The same four cells were labelled with a reporter construct driving palmitoylated-tdTomato by a 12 kbp upstream genomic region of the *NOS* gene (Figure 1E).
+In two- and three-day-old larvae, we detected *NOS* expression in four cells (two of them weakly expressing) in the apical organ region **(Figure 1D and Figure 1---figure supplement 2)**.
+The same four cells were labelled with a reporter construct driving palmitoylated-tdTomato by a 12 kbp upstream genomic region of the *NOS* gene **(Figure 1E)**.
 This reporter also revealed the axonal projections of the NOS-expressing neurons.
-The position and morphology of the four *NOS*-expressing cells allowed us to identify the same cells in the three-day-old whole-body volume EM data [@williams2017; @veraszto2020] as four interneurons (INNOS)(Figure 1F and G).
-The INNOS cells are postsynaptic to the UV-sensory cPRCs and presynaptic to the INRGW interneurons, which are also cPRC targets (Figure 1C, D).
-INRGW neurons synapse on the head serotonergic ciliomotor neurons (Ser-h1), which synapse on the prototroch ciliary band and the cholinergic MC ciliomotor neurons (Figure 1C, D).
-We also detected *NOS* expression by HCR in the region of the visual eyes and the pigmented eyespots.
+The position and morphology of the four *NOS*-expressing cells allowed us to identify the same cells in the three-day-old whole-body volume EM data [@williams2017; @veraszto2020] as four interneurons (INNOS) **(Figure 1B,C)**.
+The INNOS cells are postsynaptic to the UV-sensory cPRCs and presynaptic to the INRGW interneurons, which are also cPRC targets **(Figure 1C,F)**.
+INRGW neurons synapse on the head serotonergic ciliomotor neurons (Ser-h1), which synapse on the prototroch ciliary band and the cholinergic MC ciliomotor neurons **(Figure 1C,F)**.
+We also detected *NOS* expression by HCR in the region of the visual eyes (adult eyes) and the pigmented eyespots **(Figure 1---figure supplement 1)**.
 
 ![**Figure 1.
-Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** **(A)** SEM image of a 3-day-old Platynereis larva.
-**(B, C)** Five types of neurons (cPRC, INNOS, INRGWa, Ser-h1 and MC) in the cPRC circuit reconstructed in the whole-body transmission electron microscopy (ssTEM) dataset of a 3-day-old larva.
+Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** **(A)** SEM image of a three-day-old Platynereis larva.
+**(B, C)** Five types of neurons (cPRC, INNOS, INRGW, Ser-h1 and MC) in the cPRC circuit reconstructed in the whole-body transmission electron microscopy (ssTEM) dataset of a three-day-old larva.
 Ciliated cells labelled with (B) are shown in grey.
@@
 **(D)** Expression analysis of the NOS gene (magenta) using in situ HCR.
-Anterior view of the larva at 2-day-old.
+Anterior view of the larva at two-day-old.
 Co-stained image using acetylated α-tubulin antibody (acTub: green).
@@
 Co-staining with acetylated α-tubulin antibody as marker for cilia of cPRC and axonal scaffold (acTub: green).
-Views of larvae at 3-day-old from the anterior side.
+Views of larvae at three-day-old from the anterior side.
 Insert: One of the reconstracted INNOS is shown from the anterior side.
@@
 
-**Nitric oxide signalling mediates UV-avoidance behaviour**
+**Nitric oxide produced during UV/violet stimulation gates the output of the cPRC circuit**
 
-The specific expression of NOS in interneurons directly postsynaptic to the cPRCs suggests that NO signalling is involved in the UV response.
-To test this, we generated two *Platynereis* *NOS* knockout lines with the CRISPR/Cas9 method.
-We recovered two deletions (NOSΔ11/Δ11 and Δ23/Δ23), both leading to an early stop codon and thus likely representing null alleles (Figure 2A).
-We could establish a homozygous line for both mutations indicating that NOS is not an essential gene.
-To quantify UV avoidance, we recorded the trajectories of freely swimming larvae in vertical columns, illuminated laterally from two opposite sides with 395 nm violet light.
-We also tested phototaxis, by exposing larvae to 480 nm directional and collimated light from the top of the column.
-As previously shown, wild-type larvae swim downward within 30 sec following non-directional UV/violet light stimulation [@Veraszto2018].
-In contrast, homozygous *NOS*-mutant larvae showed a strongly diminished UV-avoidance response (Figure 2B and C).
-This phenotype is similar to the defective UV-avoidance of *c-opsin1* mutant larvae and reveals a requirement for NOS in UV-avoidance behaviour [@Veraszto2018].
-*NOS*-mutant larvae also showed reduced phototactic behaviour, suggesting a function in the visual eyes mediating phototaxis [@Randel2014a], where *NOS* is also expressed (Figure 1).
-To test if the observed defects are due to a role of NOS in neuronal signalling or circuit developmental, we also tested larvae exposed to a NOS inhibitor, L-NAME.
-Larvae incubated for xy min in 0.1 mM or 1 mM L-NAME showed defective UV avoidance, but normal phototaxis.
-Overall, our results indicate an acute requirement for NOS signalling in UV-avoidance and a possible indirect, developmental role in the visual system.
-
-To analyse the behaviour of larvae before and after UV stimulation in more detail, the temporal changes in some parameters were compared using tracking of individual larvae of the WT and mutant (Figure 2D).
-The results showed that in the WT there was a rapid and strong increase in downward swimming approximately 15 s after UV stimulation, whereas this did not occur in the mutant (Figure 2E).
-Interestingly, there was already a difference in swimming speed between the WT and the mutant before UV stimulation (in the dark) (Figure 2F).
-After UV stimulation, the swimming speed of the WT increased and reached a maximum after about 15 s.
-In contrast, the swimming speed of the mutant did not increase after about 15 s.
-In contrast, the mutants showed a slight increase in speed after UV stimulation, but swimming speed was not maintained and returned to the speed before UV stimulation.
-To investigate the role of NO on UV avoidance behaviour, receptors for NO were explored.
+The specific expression of *NOS* in interneurons directly postsynaptic to the cPRCs suggests that NO signalling is involved in the UV response.
+To test if NO is produced during UV/violet responses, we injected zygotes with the fluorescent NO-reporter DAF-FM and imaged two-day-old larvae during the local stimulation of the cPRC cilia by 405 nm light.
+Following light stimulation, we detected an increase in DAF-FM fluorescence in the region of the INNOS cells **(Figure 2)**.
 
 ![**Figure 2.
-NO produced by UV stimulation to cPRCs.** **(A)** A NO-positive cell visualised by DAF-FM (green) is shown (dashline).
+NO produced by UV/violet stimulation to cPRCs.** **(A)** A NO-positive cell visualised by DAF-FM (green) is shown (dashline).
 The white line indicates the outer frame of the larva.
-All subsequent images are of 2-day-old larvae.
+All subsequent images are of two-day-old larvae.
 **(B)** Intensity changes before (upper) and during 405 nm light stimulation (lower) in a cell visualised by DAF-FM.
@@
 
-**Nitric oxide produced during UV/violet stimulation gates the output of the cPRC circuit**
+**Nitric oxide signalling mediates UV-avoidance behaviour**
 
-To test if NO is produced during UV/violet responses, we injected zygotes with the fluorescent NO-reporter DAF-FM and imaged two-day-old larvae during the local stimulation of the cPRC cilia by 405 nm light (Figure 4---figure supplement 1).
-Following light stimulation, we detected an increase in DAF-FM fluorescence in the region of the INNOS cells (Figure 4A).
-We next analysed the dynamics of the cPRC circuit following light stimulation.
-As we have shown previously [@Veraszto2018], a 20-sec illumination of the cPRC ciliary region lead to a decrease in cPRC calcium levels, likely due to the hyperpolarisation of the cPRCs.
-After 20-sec, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus - a response that may involve depolarisation.
-This activation phase occurs after the end of the stimulation and may be due to a slower neuroendocrine feedback [@Veraszto2018].
-To determine whether NO mediates such a feedback, we measured calcium signals in *NOS*-mutant larvae.
-While we detected the inhibition phase in cPRC in *NOS*-mutants, this was not followed by activation and calcium levels remained low (Figure 4B and C).
-These results revealed an essential requirement for NOS signalling in the second phase of the cPRC response.
+To test this, we generated two *Platynereis* *NOS* knockout lines with the CRISPR/Cas9 method.
+We recovered two deletions (NOSΔ11/Δ11 and Δ23/Δ23), both leading to an early stop codon and thus likely representing null alleles **(Figure 3A)**.
+We could establish a homozygous line for both mutations indicating that NOS is not an essential gene.
+To quantify UV avoidance, we recorded the trajectories of freely swimming larvae in vertical columns, illuminated laterally from two opposite sides with 395 nm UV/violet light.
+To analyse the behaviour of larvae before and after UV stimulation, the temporal changes in some parameters were compared using tracking of individual larvae of the WT and mutant **(Figure 3B,C and Figure 3---figure supplement 1A-D)**.
+The results showed that in the WT there was a rapid and strong increase in downward swimming approximately 15 s after UV stimulation, whereas this did not occur in the mutant **(Figure 3C and Figure 3---figure supplement 1B)**.
+Interestingly, there was already a difference in swimming speed between the WT and the mutant before UV stimulation (in the dark) **(Figure 3---figure supplement 1C,D)**.
+After UV stimulation, the swimming speed of the WT increased and reached a maximum after about 20 s.
+In contrast, the mutants showed a slight increase in speed after UV stimulation, but swimming speed was not maintained and returned to the speed before UV stimulation.
+We also tested phototaxis, by exposing larvae to 480 nm directional and collimated light from the top of the column.
+As previously shown, wild-type larvae swim downward within 30 sec following non-directional UV/violet light stimulation [@Veraszto2018].
+In contrast, homozygous *NOS*-mutant larvae showed a strongly diminished UV-avoidance response **(Figure 3D and Figure 3---figure supplement 1E)**.
+This phenotype is similar to the defective UV-avoidance of *c-opsin1* mutant larvae and reveals a requirement for NOS in UV-avoidance behaviour [@Veraszto2018].
+*NOS*-mutant larvae also showed reduced phototactic behaviour, suggesting a function in the visual eyes mediating phototaxis [@Randel2014a], where *NOS* is also expressed **(Figure 1---figure supplement 1A,B)**.
+To test if the observed defects are due to a role of NOS in neuronal signalling or circuit developmental, we also tested larvae exposed to a NOS inhibitor, L-NAME.
+Larvae incubated for 5 min in 0.1 mM or 1 mM L-NAME showed defective UV avoidance, but normal phototaxis **(Figure 3E)**.
+Overall, our results indicate an acute requirement for NOS signalling in UV-avoidance and a possible indirect, developmental role in the visual system.
+To investigate the role of NO on UV avoidance behaviour, receptors for NO were explored.
 
@@
 Gray shows PAM sequences, red shows stop codons.
-**(B)** Overlaid trajectories for WT (left, n=20) and NOS mutant (right, n=31) 3-day-old larvae.
-Grey shows the trajectories before UV stimulation and purple shows the trajectories after UV stimulation.
-**(C)** The temporal changes in the vertical position of the WT and mutant larvae before and after UV stimulation are shown.
+**(B)** Overlaid trajectories for WT (n=32) and NOS mutant (NOSΔ11/Δ11, n=26 and NOSΔ23/Δ23, n=47) at three-day-old larvae.
+0 sec as the starting point.
+After 10 sec, UV (395 nm) stimulation from the side.
+**(C)** The temporal changes in the vertical position of the WT and mutant larvae before (dark) and after UV stimulation (side 395 nm) are shown.
 The starting points of each larval trajectory are set to 0.
-After UV stimulation is indicated by purple squares.
-**(D)** Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23) larvae at 3-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
+**(D)** Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23) larvae at three-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
 The data are shown in 30 s bins.
-**(E)** Vertical swimming in larvae treated with NOS inhibitors, L-NAME at 3-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
+**(E)** Vertical swimming in larvae treated with NOS inhibitors, L-NAME at three-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
 The data are shown in 30 s bins.](../figures/Fig3.png){width="100%"}
 
-**Two unconventional guanylyl cyclases (NIT-GC) are expressed in the cPRCs**
+**Two unconventional guanylyl cyclases are expressed in the cPRCs**
 
@@
 NO binding to the heme group of sGC leads to increased cyclic guanosine monophosphate (cGMP) production.
-Analysis of sGCs in *Platynereis* indicated that these genes are not expressed in any of the cells of the cPRC circuit ([@Veraszto2017] and data not shown).
+Analysis of sGCs in *Platynereis* indicated that these genes are not expressed in any of the cells of the cPRC circuit [@Veraszto2017].
 Recently, Moroz and coworkers reported an atypical but widely conserved family of guanylyl cyclases with a NIT (nitrite/nitrate sensing) domain (PF08376) (NIT-GC) as potential mediators of NO signalling [@Moroz2020].
-To search for NIT-GCs in *Platynereis*, we searched transcriptome resources and retrieved 15 potential NIT-GC homologs.
-To analyse the relationship of these sequences to metazoan NIT-GCs, we retrieved protein sequences with a CYC domain from the transcriptome and genome databases of 45 metazoan species and [\*\*how many?] choanoflagellate.
+To search for NIT-GCs in *Platynereis*, we searched transcriptome resources and retrieved 15 potential NIT-GC homologs **(Figure 4---figure supplement 1)**.
+To analyse the relationship of these sequences to metazoan NIT-GCs, we retrieved protein sequences with a CYC domain from the transcriptome and genome databases of 45 metazoan species and 2 choanoflagellate species.
 We carried out cluster analysis and did phylogenetic reconstruction on a group of membrane-bound guanylyl cyclases with sGCs as an outgroup.
@@
 [@Moroz2020],we found a group of GCs with NIT domains with representatives in placozoans, cnidarians, some ecdysozoans, echinoderms, and lophotrochozoans.
-The 15 *Platynereis* sequences were parts of several major and ancient clades of the tree (Figure 4---figure supplement 1).
+The 15 *Platynereis* sequences were parts of several major and ancient clades of the tree **(Figure 4---figure supplement 1)**.
 To characterise the expression of NIT-GCs, we using previously published spatially mapped single-cell transcriptome data [@Achim2015; @williams2017].
-Among the 15 NIT-GCs, four showed high and specific expression in the cPRCs and one was expressed in the INNOS cells (Figure 3---figure supplement 1).
-In the single-cell data, we could identify the cPRCs by the specific expression of c-opsin1 and a pedal-peptide2 neuropeptide precursor (MLD proneuropeptide) [@Arendt2004; @williams2017], which have previously been described as cPRC markers (Figure 3A and B, Figure 3---figure supplement 2).
+Among the 15 NIT-GCs, two showed high and specific expression in the cPRCs and one was expressed in the INNOS cells **(Figure 4---figure supplement 1)**.
+In the single-cell data, we could identify the cPRCs by the specific expression of c-opsin1 and a pedal-peptide2 neuropeptide precursor (MLD proneuropeptide) [@Arendt2004; @williams2017], which have previously been described as cPRC markers **(Figure 4---figure supplement 2A)**.
 The INNOS cells were identified by NOS expression and their spatial mapping.
 We decided to focus on two NIT-GCs expressed in the cPRCs and with a full-length sequences, NIT-GC1 and NIT-GC2.
-To confirm the single-cell data, we first carried out HCR with probes for *NIT-GC1* and *NIT-GC2* mRNA.
-Both genes showed very clear and specific expression in the four cPRCs, as confirmed by co-labeling with an acetylated tubulin antibody and with an HCR probe against *MLD proneuropeptide*.
+To confirm the single-cell data, we first carried out in situ HCR with probes for *NIT-GC1* and *NIT-GC2* mRNA.
+Both genes showed very clear and specific expression in the four cPRCs, as confirmed by co-labeling with an acetylated α-tubulin antibody and with an HCR probe against *MLD proneuropeptide* **(Figure 4A,B and Figure 4---figure supplement 2A-C)**.
 To analyse the subcellular localisation of NIT-GC1 and NIT-GC2 at the protein level, we raised and affinity-purified polyclonal antibodies against a specific peptide sequence from both proteins.
-In immunostainings, we found that NIT-GC1 was localise to the region corresponding to the axonal projections of the cPRCs in the anterior nerve plexus.
-Co-immunostaining with the rabbit NIT-GC1 and the rat NOS antibodies revealed that both proteins were localised in close proximity in the neurosecretory plexus (Figure 3---figure supplement 3).
-In contrast, NIT-GC2 strongly and specifically labelled the large sensory ciliary region of the cPRCs (Figure 3C and D).
+In immunostainings, we found that NIT-GC1 was localise to the region corresponding to the axonal projections of the cPRCs in the anterior nerve plexus **(Figure 4C)**.
+Co-immunostaining with the rabbit NIT-GC1 and the rat NOS antibodies revealed that both proteins were localised in close proximity in the neurosecretory plexus **(Figure 4---figure supplement 2D)**.
+In contrast, NIT-GC2 strongly and specifically labelled the large sensory ciliary region of the cPRCs **(Figure 4D)**.
 These different subcellular localisations suggest that the two NIT-GCs are involved in different intracellular signalling processes in the ciliary and axonal regions of the cPRCs.
@@
 
-To characterise the role of NIT-GCs within the cPRC during UV stimulation in more detail, the calcium sensor GCaMP6s were ubiquitously expressed in larvae in which NIT-GCs were knocked down using morpholino to examine changes in calcium levels in the cPRC.
-Local stimulation of the ciliated part of the cPRC with 405 nm light causes a transient decrease in calcium levels in the cPRC (hyperpolarisation), followed by an increase in calcium levels (depolarisation) when the light is switched off [@Veraszto2018].
-In NIT-GC1 morphant larvae, 405 nm light stimulation caused a decrease in calcium levels in the cPRC, but the increase in calcium levels that occurred after the light stimulation was stopped was strongly suppressed (Figure 3E).
-This result suggests that NIT-GC1 mediates depolarising responses after UV stimulation of cPRC and may be involved in extracellular signalling.
-Interestingly, in NIT-GC2 morphant larvae, an increase in calcium levels occurred immediately after light stimulation (Figure 3F).
+To investigate how NIT-GCs-mediated NO signalling is involved in cPRC activation, the calcium sensor GCaMP6s was ubiquitously expressed in larvae and calcium imaging in the cPRC during UV/violet stimulation.
+As we have shown previously, a 20-sec local stimulation of the ciliated part of the cPRC with 405 nm light lead to a transient decrease in cPRC calcium levels, likely due to the hyperpolarisation of the cPRCs [@Veraszto2018].
+After 20-sec, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus - a response that may involve depolarisation.
+This activation phase occurs after the end of the stimulation and may be due to a slower neuroendocrine feedback [@Veraszto2018].
+First, to determine whether NO mediates such a feedback, we measured calcium levels in *NOS*-mutant larvae.
+While we detected the inhibition phase in cPRC in *NOS*-mutants, this was not followed by activation and calcium levels remained low **(Figure 4E,F)**.
+These results revealed an essential requirement for NO signalling in the second phase of the cPRC response.
+Next, to characterise the role of NIT-GCs in changes of calcium levels in cPRC, calcium imaging was performed when each *NIT-GC* gene was knocked down using morpholino **(Figure 4---figure supplement 2E,F)**.
+In NIT-GC1 morphant larvae, 405 nm light stimulation caused a decrease in calcium levels in the cPRC, but the increase in calcium levels that occurred after the light stimulation was stopped was strongly suppressed **(Figure 4G)**.
+This result is similar to the response of NOS mutants in cPRC, suggesting that NIT-GC1 mediates depolarising responses after UV stimulation of cPRC by NO and may be involved in extracellular signalling.
+Interestingly, in NIT-GC2 morphant larvae, an increase in calcium levels occurred immediately after light stimulation **(Figure 4H)**.
 The different responses of cPRC upon knockdown of NIT-GC1 and NIT-GC2 suggest that there may be different intracellular signalling mechanisms associated with their respective localisations.
@@
 
-**NIT-GC1 produces cGMP by nitric oxide**
-
-NIT-GC1 has a highly conserved cyclase domain and activation of NIT-GC1 is expected to produce cGMP.
-Recently, a synthetic protein called Green cGull was developed as a cGMP indicator that can be used in cultured cell systems [@Matsuda2017].
-We co-expressed Green cGull and NIT-GC1 using the 2A self-cleaving peptide in cultured cells COS-7 (monkey kidney) and tested the intracellular dynamics of cGMP in the activation of NIT-GC1 (Figure 3---figure supplement 4).
-When the NO donor SNAP was applied to the NIT-GC1 and Green cGull co-expression cells, the intracellular cGMP levels was significantly increased compared to when NIT-GC1 was not expressed (Figure 3G and H).
-The increase in fluorescence intensity also did not occur when DMSO was applied (Figure 3I).
-To confirm whether this increase in intracellular cGMP levels is NIT domain-dependent response of NIT-GC1, mutant form of NIT-GC1 without the NIT domain was expressed and SNAP was applied (Figure 3---figure supplement 4A).
-As a result, the increase in fluorescence intensity was significantly reduced (Figure 3J).
-These results indicate that NIT-GC1 produces cGMP via the NIT domain.
-In bacteria, it has been proposed that the NIT domain regulates cellular functions in response to changes in extracellular and intracellular nitrate and nitrite concentrations.
-Our results suggest that NIT-GC1 may be sensitive to NO.
-This could be a possible feedback pathway in which NIT-GC1 expressed in cPRC is activated by NO produced by IN-NOS.
-In order to further explore the role of NO in the cPRC circuit, we therefore compared calcium imaging of postsynaptic cells (IN-NOS, IN-RGW, Ser-h1, MC) in the cPRC circuit during UV stimulation of cPRC between WT and NOS mutant larvae.
-
 ![**Figure 4.
-cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.** **(A, B)** Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in WT (dark yellow) and NOS mutant (A: NOSΔ11/Δ11, B: NOSΔ23/Δ23) (light yellow) larvae.
-**(C, D)** Expression analysis of the NIT-GC1 (C) and NIT-GC2 (D) gene (magenta) using in situ HCR.
+cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.** **(A, B)** Expression analysis of the NIT-GC1 (C) and NIT-GC2 (D) gene (magenta) using in situ HCR.
 Co-staining image using acetylated α-tubulin antibody (ac α-tub: green).
-**(E, F)** Localisation analysis using NIT-GC1 and NIT-GC2 antibodies.
+**(C, D)** Localisation analysis using NIT-GC1 and NIT-GC2 antibodies.
 Green shows co-staining with acetylated α-tubulin antibody (ac α-tub).
-**(G, H)** Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in NIT-GC1 (G) and NIT-GC2 (H) knockdown by morpholino (MO1 and 2) larvae.
+**(E, F)** Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in WT (dark yellow) and NOS mutant (E: NOSΔ11/Δ11, F: NOSΔ23/Δ23) (light pink) larvae.
+**(G, H)** Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in NIT-GC1 (G: NIT1) and NIT-GC2 (H: NIT2) knockdown by morpholino (MO1 and 2) larvae.
 **(I)** The domain structure of NIT-GC1.
-Grey squares indicate the transmembrane domain.
+Below shows the mutant NIT-GC1 (ΔNIT-GC1) with deletion of the NIT domain.
+Grey squares indicate the transmembrane region (TM).
 **(J)** A schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control.
@@
 
-We next quantified the responses of the INNOS and INRGW interneurons following a UV/violet stimulus to the cPRCs.
-To unambiguously identify the cells from which we recorded calcium signals, we developed an on-slide immunostaining method.
-We used the cell-specific neuropeptide markers anti-RYamide (INNOS) and anti-RGWa (INRGW) [@Conzelmann2011] to mark the interneurons.
-Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision (Figure 4---figure supplement 2).
-In both wild type and *NOS*-mutant larvae, INNOS cells showed increase calcium signals during the inhibition phase of the cPRCs (Figure 4D and E).
-In addition, In-NOS depolarisation was suppressed in NIT-GC2 molphants where cPRC hyperpolarisation was suppressed.
-This indicates that cPRC hyperpolarisation is required for IN-NOS depolarisation (Figure 4F).
-In contrast, calcium signals increased in INRGW cells in wild-type larvae during the second activation phase of cPRCs, a response that was lacking in *NOS*-mutants (Figure 4G and H).
-This suggests that there is no circuit output to the Ser-h1 cells and the ciliary band in *NOS*-mutants, consistent with the observed behavioural defects (Figure 4I).
+**NIT-GC1 produces cGMP by nitric oxide**
+
+NIT-GC1 has a highly conserved cyclase domain and activation of NIT-GC1 is expected to produce cGMP **(Figure 4I)**.
+Recently, a synthetic protein called Green cGull was developed as a cGMP indicator that can be used in cultured cell systems [@Matsuda2017].
+We co-expressed Green cGull and NIT-GC1 using the 2A self-cleaving peptide in cultured cells COS-7 (monkey kidney) and tested the intracellular dynamics of cGMP in the activation of NIT-GC1 **(Figure 4J)**.
+When the NO donor SNAP was applied to the NIT-GC1 and Green cGull co-expression cells, the intracellular cGMP levels was significantly increased compared to when NIT-GC1 was not expressed **(Figure 4K,L)**.
+The increase in fluorescence intensity also did not occur when DMSO was applied **(Figure 4M)**.
+To confirm whether this increase in intracellular cGMP levels is NIT domain-dependent response of NIT-GC1, mutant form of NIT-GC1 without the NIT domain was expressed and SNAP was applied **(Figure 4I)**.
+As a result, the increase in fluorescence intensity was significantly reduced **(Figure 4N)**.
+These results indicate that NIT-GC1 produces cGMP via the NIT domain.
+In bacteria, it has been proposed that the NIT domain regulates cellular functions in response to changes in extracellular and intracellular nitrate and nitrite concentrations.
+Our results suggest that NIT-GC1 may be sensitive to NO.
+This could be a possible feedback pathway in which NIT-GC1 expressed in cPRC is activated by NO produced by INNOS.
+In order to further explore the role of NO in the cPRC circuit, we therefore compared calcium imaging of postsynaptic cells (INNOS, INRGW, Ser-h1 and MC) in the cPRC circuit during UV stimulation of cPRC between WT and NOS mutant larvae.
+
+**Depolarisation of cPRC is required for activation of downstream interneurons and ciliomotor neurons**
+
+We next quantified the responses of the INNOS and INRGW following a UV/violet stimulus to the cPRCs.
+To unambiguously identify the cells from which we recorded calcium signals, we developed an on-slide immunostaining method **(Figure 5---figure supplement 1)**.
+We used the cell-specific neuropeptide markers using anti-RYamide (INNOS) and anti-RGWamide (INRGW) antibodies [@Conzelmann2011] to mark the interneurons.
+Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision **(Figure 5A,B)**.
+In both wild-type and *NOS*-mutant larvae, INNOS cells showed increase calcium signals during the inhibition phase of the cPRCs **(Figure 5D)**.
+In addition, INNOS depolarisation was suppressed in NIT-GC2 molphants where cPRC hyperpolarisation was suppressed **(Figure 5E)**.
+This indicates that cPRC hyperpolarisation is required for INNOS depolarisation.
+In contrast, calcium signals increased in INRGW cells in wild-type larvae during the second activation phase of cPRCs, a response that was lacking in *NOS*-mutants **(Figure 5F,G)**.
+This suggests that there is no circuit output to the Ser-h1 and MC cells and the ciliary band in *NOS*-mutants, consistent with the observed behavioural defects **(Figure 5H,I)**.
 
 ![**Figure 5.
-NO produced by UV stimulation to cPRCs induces a depolarising response in cPRC via NIT-GC1.** **(A)** Schematic diagram of the immunostaining procedure after calcium imaging.
-**(B, D)** The position of the cPRC in calcium imaging with GCaMP6s.
-Cells around the cPRC are numbered according to the position of the nuclei.
-**(C, E)** Immunostaining images using RYamide (C) or RGWamide and 5-HT antibodies (E) and DAPI (cyan) in the same region as (B) or (D).
-IN-NOS, IN-RGWa and Ser-h1 identified from the numbered nuclear positions are shown.
-**(F)** Correlation (Pearson's r) map of neuronal activity of the cPRC, IN-NOS, IN-RGWa and Ser-h1 neurons.
-**(G)** Changes over time in GCaMP6s intensities in IN-NOS when cPRC is stimulated with UV in WT (dark purple) and NOSΔ11/Δ11 mutant (light purple) larvae.
-**(H)** Changes over time in GCaMP6s intensities in IN-NOS when cPRC is stimulated with UV in NIT-GC2 morphants (MO1 and MO2: pink and light pink) larvae.
-**(I)** Changes over time in GCaMP6s intensities in IN-RGWa when cPRC is stimulated with UV in WT (dark blue) and NOSΔ23/Δ23 mutant (light blue) larvae.
-**(J)** Changes over time in GCaMP6s intensities in IN-RGWa when cPRC is stimulated with UV in NIT-GC2 morphants (MO1 and MO2: blue and light blue).
-**(K, L)** Changes over time in GCaMP6s intensities in Ser-h1 and MC when cPRC is stimulated with UV in WT (Ser-h1 and MC: dark red and brown) and NOSΔ11/Δ11 mutant (Ser-h1 and MC: light red and brown) larvae.](../figures/Fig5.png){width="100%"}
+NO produced by UV stimulation to cPRCs induces a depolarising response in cPRC via NIT-GC1.** **(A, B)** Each left figure shows a calcium imaging using GCaMP6s.
+The respective right figures show immunostaining images of the same positions in the respective left figures, (A) shows RYamide antibodies, (B) shows RGWamide and serotonin (5-HT) antibodies in red and DAPI in cyan.
+Asterisks indicate the nuclei of the cPRC, respectively, and the nuclei of the cells that were identified from the arrangement of the surrounding cells are indicated by numbers.
+From the position of the cells, the calcium imaging data identify INNOS in (A) and INRGW and Ser-h1 in (B), respectively.
+**(C)** Correlation (Pearson's r) map of neuronal activity of the INNOS, INRGW, Ser-h1 and MC neurons.
+**(D)** Changes over time in GCaMP6s intensities in INNOS when cPRC is stimulated with UV in WT (dark purple) and NOSΔ11/Δ11 mutant (light purple) larvae.
+**(E)** Changes over time in GCaMP6s intensities in INNOS when cPRC is stimulated with UV in NIT-GC2 morphants (MO1 and MO2: pink and light pink) larvae.
+**(F)** Changes over time in GCaMP6s intensities in INRGW when cPRC is stimulated with UV in WT (dark blue) and NOSΔ23/Δ23 mutant (light blue) larvae.
+**(G)** Changes over time in GCaMP6s intensities in INRGW when cPRC is stimulated with UV in NIT-GC2 morphants (MO1 and MO2: blue and light blue).
+**(H, I)** Changes over time in GCaMP6s intensities in Ser-h1 and MC when cPRC is stimulated with UV in WT (Ser-h1 and MC: dark red and brown) and NOSΔ11/Δ11 mutant (Ser-h1 and MC: light red and brown) larvae.](../figures/Fig5.png){width="100%"}
 

### 2023-01-24 code/Fig1_sup2.R

@@ -26,12 +26,12 @@ panel_NOS3d_ventr_HCR <- ggdraw() + draw_image(readPNG("pictures/HCR_72_DV_NOS_9
                lineend = "butt",
                linejoin = "mitre",
                arrow.fill = "white", size = 0.2) +
-  draw_label("eyespots", x = 0.55, y = 0.68, size = 6, color = "white") +
-  draw_line(x = c(0.3, 0.45), y = c(0.63, 0.67), color = "white", size = 0.2) +
-  draw_line(x = c(0.77, 0.66), y = c(0.65, 0.67), color = "white", size = 0.2) +
-  draw_label("adult eyes", x = 0.55, y = 0.34, size = 6, color = "white") +
-  draw_line(x = c(0.26, 0.44), y = c(0.43, 0.35), color = "white", size = 0.2) +
-  draw_line(x = c(0.77, 0.65), y = c(0.43, 0.35), color = "white", size = 0.2)
+  draw_label("eyespots", x = 0.55, y = 0.68, size = 5, color = "white") +
+  draw_line(x = c(0.3, 0.45), y = c(0.63, 0.67), color = "white", size = 0.1) +
+  draw_line(x = c(0.77, 0.66), y = c(0.65, 0.67), color = "white", size = 0.1) +
+  draw_label("adult eyes", x = 0.55, y = 0.34, size = 5, color = "white") +
+  draw_line(x = c(0.26, 0.44), y = c(0.43, 0.35), color = "white", size = 0.1) +
+  draw_line(x = c(0.77, 0.65), y = c(0.43, 0.35), color = "white", size = 0.1)
 
 
 panel_NOS3d_ant_HCR <- ggdraw() + draw_image(readPNG("pictures/HCR_72_AP_NOS_89.25um.png")) +
@@ -40,22 +40,20 @@ panel_NOS3d_ant_HCR <- ggdraw() + draw_image(readPNG("pictures/HCR_72_AP_NOS_89.
   draw_label("DAPI", x = 0.32, y = 0.9, color="cyan", size = 11, fontface="plain") +
   draw_line(x = c(0.04, 0.264), y = c(0.08, 0.08), color = "white", size = 0.5) +
   draw_label("D", x = 0.95, y = 0.88, size = 6, color = "white") +
-  draw_image(readPNG("pictures/HCR_48_DV_NOS_15.68um_insert.png"), 
-             scale = 0.35, x = 1, hjust = 1, halign = 1, valign = 0.12) +
[…]

### 2023-01-24 code/Fig3.R

@@ -94,20 +94,20 @@ WTvsNOS_2d %>%
   filter(genotype %in% c("WT", "NOS11x11", "NOS23x23")) %>%
   ggplot() +
   annotate("rect", xmin=0.5, xmax=3.5, ymin=-Inf, ymax=Inf,alpha=0.4, fill="black") +
-  annotate("text", x=2, y=-0.75, label="dark", color="black", size = 6, fontface="plain") +
+  annotate("text", x=2, y=-0.46, label="dark", color="black", size = 5.5, fontface="plain") +
   annotate("rect", xmin=3.5, xmax=7.5, ymin=-Inf, ymax=Inf,alpha=0.1, fill="blue") +
-  annotate("text", x=5.5, y=-0.5, label="side 395nm", color="purple", size = 6, fontface="plain") +
-  annotate("text", x=5.5, y=0.65, label="UV avoidance", color="black", size = 5, fontface="plain") +
+  annotate("text", x=5.5, y=-0.46, label="side 395nm", color="purple", size = 5.5, fontface="plain") +
+  annotate("text", x=5.5, y=0.56, label="UV avoidance", color="black", size = 4.5, fontface="plain") +
   annotate("rect", xmin=7.5, xmax=11.5, ymin=-Inf, ymax=Inf,alpha=0.4, fill="black") +
-  annotate("text", x=9.5, y=-0.75, label="dark", color="black", size = 6, fontface="plain") +
+  annotate("text", x=9.5, y=-0.46, label="dark", color="black", size = 5.5, fontface="plain") +
   annotate("rect", xmin=11.5, xmax=15.5, ymin=-Inf, ymax=Inf,alpha=0.1, fill="deepskyblue") +
-  annotate("text", x=13.5, y=-0.5, label="top 480nm", color="deepskyblue3", size = 6, fontface="plain") +
-  annotate("text", x=13.5, y=0.65, label="phototaxis", color="black", size = 5, fontface="plain") +
+  annotate("text", x=13.5, y=-0.46, label="top 480nm", color="deepskyblue3", size = 5.5, fontface="plain") +
+  annotate("text", x=13.5, y=0.56, label="phototaxis", color="black", size = 4.5, fontface="plain") +
   annotate("rect", xmin=15.5, xmax=19.5, ymin=-Inf, ymax=Inf,alpha=0.4, fill="black") +
[…]

### 2023-01-24 code/Fig4.R

@@ -61,27 +61,28 @@ WT_cPRC_tidy$genotype <- factor(WT_cPRC_tidy$genotype,
                                 levels=c("WT_cPRC", "NOS23_cPRC"))
 
 WT_cPRC_tidy %>%
-  ggplot(aes(x=frame,y=intensity,color=genotype)) +
-  annotate("rect", xmin=51, xmax=90, ymin=-Inf, ymax=Inf, 
+  ggplot() +
+  annotate("rect", xmin=51, xmax=93, ymin=-Inf, ymax=Inf, 
            alpha=0.1, fill="blue") +
+  aes(x=frame,y=intensity,color=genotype) +
   geom_line(aes(group = sample), size=0.2, alpha=0.2) +
   geom_smooth(aes(fill = genotype), level = 0.99, size = 1, span = 0.03, 
-              method = "loess") +
-  annotate("segment", x=20, xend=40, y=1.3, yend=1.3, size=1)+
-  annotate("segment", x=20, xend=20, y=1.3, yend=1.4, size=1)+
-  annotate("text", x=29, y=1.25, label="10 sec", size=3)+
-  annotate("text", x=23, y=1.45, label="0.2 ΔF/F0", size=3)+
-  annotate("text", x=70, y=1.7, label="405 nm", size=4)+
+              method = "loess", formula = y ~ x) +
+  annotate("segment", x=20, xend=40, y=1.4, yend=1.4, size=1)+
+  annotate("segment", x=20, xend=20, y=1.4, yend=1.65, size=1)+
+  annotate("text", x=29, y=1.35, label="10 sec", size=3)+
+  annotate("text", x=35, y=1.53, label="0.5 ΔF/F0", size=3)+
+  annotate("text", x=72, y=1.7, label="405 nm", color="purple", size=4)+
   annotate("text", x=160, y=1.7, label="cPRC", size=5)+
   ylim(0.2,1.7)+
   theme_void()  +
-  theme(legend.title = element_blank(), legend.text.align=0,
+  theme(legend.title = element_blank(), 
+        legend.text.align = 0,
         legend.text = element_text(size=12)) +
-  scale_color_manual(values=c(Okabe_Ito[6], paired[7]),
-                     labels = c("wt", expression('NOS'^'Δ23/Δ23'))) +
-  scale_fill_manual(values=c(Okabe_Ito[6], paired[7]),
[…]

### 2023-01-24 code/Fig4_sup2.R

@@ -51,13 +51,17 @@ panel_72_DV_NIT2 <- ggdraw() + draw_image(readPNG("pictures/HCR_72_DV_NIT_180.08
 
 
 panel_NIT1MO_NIT1 <- ggdraw() + draw_image(readPNG("pictures/IHC_51_AP_NIT1MO_NITGC1_actub_56.82um.png")) +
-  draw_label("IHC: NIT-GC1 morpholino", x = 0.4, y = 0.96, size = 10) +
+  draw_label("IHC:", x = 0.15, y = 0.96, size = 10) +
+  draw_label(expression(paste(italic("NIT-GC1"))), x = 0.32, y = 0.96, size = 10) +
+  draw_label("morpholino", x = 0.57, y = 0.96, size = 10) +
   draw_label("NIT-GC1", x = 0.12, y = 0.87, color="magenta", size = 10, fontface="plain") +
   draw_label("acTub", x = 0.33, y = 0.87, color="green", size = 10, fontface="plain") +
   draw_line(x = c(0.04, 0.39), y = c(0.11, 0.11), color = "white", size = 0.5)
 
 panel_NIT2MO_NIT2 <- ggdraw() + draw_image(readPNG("pictures/IHC_57_AP_NIT2MO_NITGC2_actub_57um.png")) +
-  draw_label("IHC: NIT-GC2 morpholino", x = 0.4, y = 0.96, size = 10) +
+  draw_label("IHC:", x = 0.15, y = 0.96, size = 10) +
+  draw_label(expression(paste(italic("NIT-GC2"))), x = 0.32, y = 0.96, size = 10) +
+  draw_label("morpholino", x = 0.57, y = 0.96, size = 10) +
   draw_label("NIT-GC2", x = 0.12, y = 0.87, color="magenta", size = 10, fontface="plain") +
   draw_label("acTub", x = 0.33, y = 0.87, color="green", size = 10, fontface="plain") +
   draw_line(x = c(0.04, 0.39), y = c(0.11, 0.11), color = "white", size = 0.5)
@@ -76,17 +80,15 @@ panel_NIT1_NOS <- ggdraw() + draw_image(readPNG("pictures/IHC_55_NIT1_NOS_acTub_
 #panels of different sizes
 layout <- "
 A#B#C
-#####
 D#E#F
 "
 
[…]

### 2023-01-24 code/Fig5.R

@@ -71,17 +71,18 @@ WTvsNOS23_cPRC_INRGWa_tb$genotype <- factor(WTvsNOS23_cPRC_INRGWa_tb$genotype,
 
 WTvsNOS23_cPRC_INRGWa_tb %>%
   filter(genotype == "NOS23_cPRC" | genotype == "WT_cPRC") %>%
-  ggplot(aes(x=frame,y=intensity,color=genotype)) +
-  annotate("rect", xmin=51, xmax=90, ymin=-Inf, ymax=Inf, 
+  ggplot() +
+  annotate("rect", xmin=51, xmax=93, ymin=-Inf, ymax=Inf, 
            alpha=0.1, fill="blue") +
+  aes(x=frame,y=intensity,color=genotype) +
   geom_line(aes(group = sample), size=0.2, alpha=0.2) +
   geom_smooth(aes(fill = genotype), level = 0.99, size = 1, span = 0.03, 
               method = "loess") +
-  annotate("segment", x=20, xend=40, y=1.3, yend=1.3, size=1)+
-  annotate("segment", x=20, xend=20, y=1.3, yend=1.4, size=1)+
-  annotate("text", x=29, y=1.25, label="10 sec", size=3)+
-  annotate("text", x=23, y=1.45, label="0.2 ΔF/F0", size=3)+
-  annotate("text", x=70, y=1.7, label="405 nm", size=4)+
+  annotate("segment", x=15, xend=35, y=1.3, yend=1.3, size=1)+
+  annotate("segment", x=15, xend=15, y=1.3, yend=1.5, size=1)+
+  annotate("text", x=24, y=1.25, label="10 sec", size=3)+
+  annotate("text", x=31, y=1.4, label="0.4 ΔF/F0", size=3)+
+  annotate("text", x=72, y=1.7, label="405 nm", color="purple", size=4)+
   annotate("text", x=160, y=1.7, label="cPRC", size=5)+
   ylim(0.4,1.7)+
   theme_void()  +
@@ -89,14 +90,14 @@ WTvsNOS23_cPRC_INRGWa_tb %>%
         legend.text = element_text(size=12)) +
   #Specify colours and legend labels
   scale_color_manual(values=c(Okabe_Ito[6], paired[7]),
-                     labels = c("wt", expression('NOS'^'Δ23/Δ23'))) +
+                     labels = c("WT", expression('NOS'^'Δ23/Δ23'))) +
[…]

### 2023-01-24 code/Fig5_sup1.R

@@ -1,0 +1,25 @@
+#Code to assemble Fig5 sup1 of the Jokura et al Platynereis NOS paper
+#2022 June - Kei Jokura, Gaspar Jekely
+
+# source packages ---------------------------------------------------------
+source("code/Packages_to_load.R")
+
+
+# assemble figure ---------------------------------------------------------
+panel_method <- ggdraw() + draw_image(readPNG("pictures/Diagram of agarose embedding immunostaining.png"))
+
+#combine panels into Figure and save final figure as pdf and png
+#panels of different sizes
+layout <- "
+A
+"
+
+Fig5_sup1 <- panel_method + 
+  patchwork::plot_layout(design = layout) #we can change the heights of the rows in our layout (widths also can be defined)
+
+
+ggsave("figures/Fig5_sup1.png", limitsize = FALSE, 
+       units = c("px"), Fig5_sup1, width = 1857, height = 970, bg='white')  
+
+ggsave("figures/Fig5_sup1.pdf", limitsize = FALSE, 
+       units = c("px"), Fig5_sup1, width = 1857, height = 970)  

### 2023-01-24 code/Fig6.R

@@ -21,8 +21,8 @@ AB
 Fig6 <- panel_scRNA_dotplot + panel_model +
   patchwork::plot_layout(design = layout, 
                          widths = c(1, 1)) + #we can change the heights of the rows in our layout (widths also can be defined)
-  patchwork::plot_annotation(tag_levels = 'A') +  #we can change this to 'a' for small caps or 'i' or '1'
-  ggplot2::theme(plot.tag = element_text(size = 12, face='bold')) #or 'bold', 'italic'
+  patchwork::plot_annotation(tag_levels = 'A') &  #we can change this to 'a' for small caps or 'i' or '1'
+  ggplot2::theme(plot.tag = element_text(size = 12, face='plain')) #or 'plain', 'italic'
 
 ggsave("figures/Fig6.png", limitsize = FALSE, 
        units = c("px"), Fig6, width = 1800, height = 1000, bg='white')  

### 2023-01-24 text/Manuscript.Rmd
edit Methods

@@
 In calcium imaging using GCaMP6s, INNOS identified by the positional relationship between the cPRC and INNOS and the labelling of RYa depolarised in clear synchrony with the hyperpolarising response of the cPRCs.
-This tonic response may involve dopaminergic nerves based on single-cell data (Fig) [@puopolo2001][@zhang2007][@Randel2014a].
+This tonic response may involve dopaminergic nerves based on single-cell data (Figure 6A) [@puopolo2001][@zhang2007][@Randel2014a].
 In the mammalian retina, it is known that signals from bipolar photoreceptor cells are transmitted to downstream NOS-expressing amacrine cells (NOAC), where NO is produced [@jacoby2018].
@@
 
-To identify GPR173 receptors, we obtained a database of vertebrate SREB sequences, including GPR173, GPR85 and GPR27 from (Breton et al.
-2021).
-From these sequences, we produced a Hidden Markov Model (HMM) and used this to mine the 49 species investigated.
+To identify NIT-GCs, we obtained a "seed" database of Adenylate and Guanylate cyclase catalytic domain in Pfam database, PF00211.
+From these sequences, we produced a Hidden Markov Model (HMM) and used this to mine the 45 metazoan species, 2 choanoflagellate species and 2 filasterea species investigated.
 HMM models were run in HMMR3 with an e-value of 1e−15.
-The same SREB sequences were used to carry out similarity-based searches using BlastP with an e-value cutoff of 1e−15.
-We merged these two databases and ran CD-Hit (Fu et al.
-2012) to eliminate redundant sequences (at a 99% threshold).
-To identify the sequences that are closely related to the GPR173 sequences, we ran a cluster-based analysis in CLANS (Frickey and Lupas 2004).The CLANS analysis is available as supplementary fig.
-S4, Supplementary Material online.
+We ran CD-Hit (Fu et al.
+2012) to eliminate redundant sequences (at a 80% threshold).
+The CLANS analysis is available as **Supplementary File 1**.
 To identify clusters, we used the convex-clustering option with 100 jack-knife replicates.
-The SREB receptors are extremely well conserved and form an easily recognizable cluster.
-To analyze the phylogeny of SREB receptors, the cluster containing these receptors together with monoaminergic receptors were parsed and used for tree building.
+The NIT-GCs are extremely well conserved in membrane-bound guanylate cyclases and form an easily recognizable cluster.
+To analyze the phylogeny of NIT-GCs, the cluster containing these GCs together with membrane-bound guanylate cyclases were parsed and used for tree building.
 We aligned the sequences with MAFFT version 7, with the iterative refinement method E-INS-i.
@@
 2020).
-The sequences used for the phylogenetic analysis are available in supplementary XX, Supplementary Material online, the trimmed alignment is available in supplementary XX, Supplementary Material online.
-The raw nexus tree of the SREB receptors is available in supplementary file XX, Supplementary Material online.
+The sequences used for the phylogenetic analysis are available in **Supplementary File 2**, the trimmed alignment is available in **Supplementary File 3**.
 

### 2023-01-24 code/Fig1.R

@@ -370,6 +370,6 @@ ggsave("figures/Fig1.png", limitsize = FALSE,
 
 
 ggsave("figures/Fig1.pdf", limitsize = FALSE, 
-       units = c("px"), Fig1, width = 2400, height = 1350) 
+       units = c("px"), Fig1, width = 1800, height = 1350) 
 
 

### 2023-01-24 text/Manuscript.Rmd
edit materials

@@
 
-**Vertical column setup for pharmacology**
-
-Freely swimming 2-day-old larvae were recorded in 5 adjacent vertical columns (each 31 mm x 10 mm x 144 mm water height) with a DMK 21BF03 camera (The Imaging Source) at 30 frames/s.
-The columns were illuminated laterally with red (633 nm) light-emitting diodes (LEDs).
-This light cannot be detected by 2-day-old larvae.
-The larvae were treated with NOS inhibitors and NO donors.
-The NOS inhibitors were L-NAME, SMIS, and AGH.
-The NO donors were SNAP and NONOate.
-The larvae were treated with different concentrations in adjacent columns.
-The concentrations for the NOS inhibitors were control, 10 mM, 1 mM, 0.1 mM, and 0.01 mM; for SNAP were control, DMSO control, 1 mM, 0.1 mM, and 0.01 mM; and for NONOate were control 1 mM, 0.1 mM, and 0.01 mM.
-The larvae were recorded for 30 s then the drugs were added.
-30 s, 1 min, 1 h, and 2 h later, the larvae were recorded again for 30 s.
-The larvae were tracked and their vertical displacement was calculated.
-Scripts are available at <https://github.com/JekelyLab/NOS>.
-
 **Vertical column setup for measuring photoresponses**
@@
 
+**Comparing behavior of wildtype and NOS-knockout 3-day-old larvae**
+
+To compare the behavior of wildtype and NOS-knockout larvae at 3 days in the vertical column, the larvae were mixed and left in the dark for 5 min.
+The larvae were treated with NOS inhibitors for pharmacology.
+The NOS inhibitors were L-NAME.
+The larvae were treated with different concentrations in adjacent columns.
+The concentrations for the NOS inhibitors were control, 1 mM, 0.1 mM.
+The larvae were recorded for 1 min in the dark followed by exposure to collimated cyan (480 nm) light from the top of the column for 2 min, then 2 min darkness, and finally collimated UV (395 nm) light from the top of the column for 2 min.
+Stimulus light was provided by the monochromator (Polychrome II, Till Photonics).
+Scripts are available at <https://github.com/JekelyLab/NOS>.
+
+**NOS Identification and Phylogenetic Analysis**
+
+To identify NOS, we obtained a "seed" database of oxygenase domain in Pfam database, PF02898.
+From these sequences, we produced a Hidden Markov Model (HMM) and used this to mine the 47 metazoan species, 2 choanoflagellate species and 2 filasterea species investigated.
+HMM models were run in HMMR3 with an e-value of 1e−15.
+We ran CD-Hit [@Fu2012] to eliminate redundant sequences (at a 80% threshold).
+We aligned the sequences with MAFFT version 7, with the iterative refinement method E-INS-i.
+Alignments were trimmed with TrimAl in gappy-out mode [@Capella-Gutiérrez2009].
+To calculate maximum-likelihood trees, we used IQ-tree2 with the LG+G4 model.
+To calculate branch support, we ran 1,000 replicates with the aLRT-SH-like and aBayes methods [@Minh2020].
+The sequences used for the phylogenetic analysis are available in **Supplementary File 2**, the trimmed alignment is available in **Supplementary File 3**.
+
 **NIT-GC Identification and Phylogenetic Analysis**
@@
 HMM models were run in HMMR3 with an e-value of 1e−15.
-We ran CD-Hit (Fu et al.
-2012) to eliminate redundant sequences (at a 80% threshold).
+We ran CD-Hit [@Fu2012] to eliminate redundant sequences (at a 80% threshold).
 The CLANS analysis is available as **Supplementary File 1**.
@@
 We aligned the sequences with MAFFT version 7, with the iterative refinement method E-INS-i.
-Alignments were trimmed with TrimAl in gappy-out mode (Capella-Gutiérrez et al.
-2009).
+Alignments were trimmed with TrimAl in gappy-out mode [@Capella-Gutiérrez2009].
 To calculate maximum-likelihood trees, we used IQ-tree2 with the LG+G4 model.
-To calculate branch support, we ran 1,000 replicates with the aLRT-SH-like and aBayes methods (Minh et al.
-2020).
+To calculate branch support, we ran 1,000 replicates with the aLRT-SH-like and aBayes methods [@Minh2020].
 The sequences used for the phylogenetic analysis are available in **Supplementary File 2**, the trimmed alignment is available in **Supplementary File 3**.
@@
 
-I used Achim et al 2015 for the single-cell data.
-In fact, in Liz's 2018 paper, she used 107 cells as neurons by removing duplicates from Achim's single-cell data, so I used those cells.
-Since the raw data were read count data, I normalized them to TPM using Python.
+We used Achim et al.
+for the single-cell data [@Achim2015].
+In fact, in Williams et al., they used 107 cells as neurons by removing duplicates from Achim et al.
+single-cell data, so we used those cells [@williams2017].
+Since the raw data were read count data, we normalized them to TPM using Python.
 After that we converted them to log10.
 From the sum of the expression levels in 107 cells for each gene, I calculated the percentage in each cell.
-I chose the appropriate genes for now.
-For each cell, I identified them with marker genes.
+We chose the appropriate genes for now.
+For each cell, we identified them with marker genes.
 After created the data in Python, plotted it using R dot plots.
-RPKM?
-TPM?
 RPKM calculates the total number of reads per million bp and then divides by the length of each gene, so it is not possible to compare between samples.
 Instead, TPM first divides by the length of the gene and then divides by the total number of reads per million bases, which allows for more accurate comparisons between samples.
-In this case I wanted to compare between samples, so I used TPM.
+In this case we wanted to compare between samples, so we used TPM.
 In fact, TPM is becoming the mainstream method for many single-cell analyses.
 The total TPM of each gene between the samples was used to calculate the percentage of expressed genes.
-
-**Comparing behavior of wildtype and NOS-knockout 3-day-old larvae**
-
-To compare the behavior of wildtype and NOS-knockout larvae at 3 days in the vertical column, the larvae were mixed and left in the dark for 5 min.
-The larvae were recorded for 1 min in the dark followed by exposure to collimated cyan (480 nm) light from the top of the column for 2 min, then 2 min darkness, and finally collimated UV (395 nm) light from the top of the column for 2 min.
-Stimulus light was provided by the monochromator (Polychrome II, Till Photonics).
-Scripts are available at <https://github.com/JekelyLab/NOS>.
 
@@
 Immunostainings were carried out as previously described [@Conzelmann2012].
+The NOS promoter (fragment sizes: 12 Kb) was amplified and cloned upstream of 3xHA- Palmi-tdTomato.
+Larvae injected with promoter constructs (ca.
+250 ng/ml) were analysed for reporter expression at 3 days post fertilization using an AxioImager Z.1 fluorescence wide-field microscope (Carl Zeiss GmbH, Jena) and immediately fixed for immunostainings.
+The protocol followed for immunostaining of HA-tagged reporters was recently described [@Veraszto2017].
+Specimens were imaged with a LSM 780 NLO Confocal Microscope (Zeiss, Jena).
 
@@
 Experiments were performed at room temperature and larvae were immobilised by being embedded in 2.5% agarose filtered artificial seawater between a slide and coverslip spaced with adhesive tape.
-GCaMP6s mRNA (1 mg/ml) was injected into zygotes as described previously (Randel et al., 2014).
+GCaMP6s mRNA (1 mg/ml) was injected into zygotes as described previously [@Randel2014a].
 Larvae were imaged on a Zeiss LSM 880 with Airyscan (with a C-Apochromat 63X/1.2 Corr - water) with a frame rate of 1.88 frame/sec and an image size of 512 x 512 pixels.
@@
 pixel diameter) with 405 nm lasers controlled by the Bleaching mode.
-The imaging laser had a similar intensity than the stimulus laser but covered an area that was ??
-times larger than the stimulus ROI.
+The imaging laser had a similar intensity than the stimulus laser but covered an area that was 10 times larger than the stimulus ROI.
 
@@
 No.
-???).
+L3000001).
 Two days post-transfection, we removed the culture medium and substituted it for fresh DMEM-medium.
@@
 Images were acquired using a xenon lamp, 460--495 nm excitation filter, 505-nm dichroic mirror and 510-- 550-nm emission filter (????).
-8-br-cGMP and S-Nitroso-N-acetyl-D, L-penicillamine (SNAP) were purchased from Sigma-Aldrich (St.
-Louis, MO, USA) and Cayman Chemical (Ann Arbor, MI, USA).
+S-Nitroso-N-acetyl-D, L-penicillamine (SNAP) was purchased from Sigma-Aldrich (St.
+Louis, MO, USA) .
 The exposure time of the EM-CCD camera was controlled by the ZEN software (Zeiss).

### 2023-05-05 text/Manuscript.Rmd
Fig6 update

@@
 
+![**Figure 6.
+Gene expressions and signalling mechanisms between cPRC, INNOS and INRGWa.** **(C)** Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq.
+The size of the dots is expressed in proportion to the percentage of cells expressing that gene relative to all cells.
+The colours represent the normal logarithm of the number of transcripts in the cells expressing the gene.
+**(D)** Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback.](../figures/Fig6.png){width="100%"}
+
 cPRCs expressing UV-absorbing c-opsin1 have high resting calcium levels and show hyperpolarising and subsequent depolarising responses to UV light (405 nm) [@Veraszto2018].
@@
 In calcium imaging using GCaMP6s, INNOS identified by the positional relationship between the cPRC and INNOS and the labelling of RYa depolarised in clear synchrony with the hyperpolarising response of the cPRCs.
-This tonic response may involve dopaminergic nerves based on single-cell data (Figure 6A) [@puopolo2001][@zhang2007][@Randel2014a].
+This tonic response may involve dopaminergic nerves based on single-cell data **(Figure 6)** [@puopolo2001][@zhang2007][@Randel2014a].
 In the mammalian retina, it is known that signals from bipolar photoreceptor cells are transmitted to downstream NOS-expressing amacrine cells (NOAC), where NO is produced [@jacoby2018].
@@
 We also believe that the appearance of NIT-GCs was one of the means to generate the diversity of signalling pathways using NO during the early stages of nervous system evolution in metazoans, since sponges and ctenophores, where NO signalling is present, have no NIT-GCs at all, as many as 12 have been found in placozoans, and NIT-GCs are functional in the "proto-eye" circuit of Platynereis.
-
-![**Figure 6.
-Gene expressions and signalling mechanisms between cPRC, INNOS and INRGWa.** **(A)** Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq.
-The size of the dots is expressed in proportion to the percentage of cells expressing that gene relative to all cells.
-The colours represent the normal logarithm of the number of transcripts in the cells expressing the gene.
-**(B)** Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback.](../figures/Fig6.png){width="100%"}
 
@@
 The protocol followed for immunostaining of HA-tagged reporters was recently described [@Veraszto2017].
-Specimens were imaged with a LSM 780 NLO Confocal Microscope (Zeiss, Jena).
+Specimens were imaged with a LSM 780 NLO or LSM 880 with Airysan Confocal Microscope (Zeiss, Jena).
 
@@
 Green cGull was used for the cGMP assay [@Matsuda2017].
-A full-length Pdum-NIT-GC1 and -NIT-GC2 coding sequences were amplified by PCR starting from a Platynereis dumerilii cDNA library and cloned into the pcDNA3.1(+) vector using the T2A self-cleaving sequence.
+A full-length Pdum-NIT-GC1 and -NIT-GC2 coding sequences were amplified by PCR starting from a *Platynereis dumerilii* cDNA library and cloned into the pcDNA3.1(+) vector using the T2A self-cleaving sequence.
 Cos-7 cells with low expression of endogenous soluble guanylate cyclase were used as cultured cells for gene expression.

### 2023-05-05 code/Fig6.R

@@ -11,21 +11,24 @@ source("code/Packages_to_load.R")
 
 panel_scRNA_dotplot <- ggdraw() + draw_image(readPNG("pictures/scRNAseq_forR_5.png"))
 panel_model <- ggdraw() + draw_image(readPNG("pictures/model_2.png"))
+panel_fitting <- ggdraw() + draw_image(readPNG("pictures/best_fit_1.png"))
+panel_block_diagram <- ggdraw() + draw_image(readPNG("pictures/draft_block_diagram.png"))
 
 #combine panels into Figure and save final figure as pdf and png
 #panels of different sizes
 layout <- "
 AB
+CD
 "
 
-Fig6 <- panel_scRNA_dotplot + panel_model +
+Fig6 <- panel_block_diagram + panel_fitting + panel_scRNA_dotplot + panel_model +
   patchwork::plot_layout(design = layout, 
                          widths = c(1, 1)) + #we can change the heights of the rows in our layout (widths also can be defined)
   patchwork::plot_annotation(tag_levels = 'A') &  #we can change this to 'a' for small caps or 'i' or '1'
   ggplot2::theme(plot.tag = element_text(size = 12, face='plain')) #or 'plain', 'italic'
 
 ggsave("figures/Fig6.png", limitsize = FALSE, 
-       units = c("px"), Fig6, width = 1800, height = 1000, bg='white')  
+       units = c("px"), Fig6, width = 1800, height = 1800, bg='white')  
 
 ggsave("figures/Fig6.pdf", limitsize = FALSE, 
        units = c("px"), Fig6, width = 1800, height = 1000)  

### 2023-05-05 code/Key_resources_table.R

@@ -1,0 +1,13 @@
+
+#Code to assemble Key resourece table of the Jokura et al Platynereis NOS paper
+#2022 June - Kei Jokura, Gaspar Jekely
+
+# source packages ---------------------------------------------------------
+library(tidyverse)
+library(gt)
+
+library(readxl)
+keyresource <- read_excel("~/NOSpaper_keyresourcetale.xlsx")
+
+keyresource %>% 
+  gt()

### 2023-05-06 Manuscript/Manuscript.Rmd
upload ms in Manuscript/

--- 2023-05-05 text/Manuscript.Rmd
@@
   in zooplankton
-author: "*Kei Jokura, Nobuo Ueda, Martin Gühmann, Luis Alfonso Yañez-Guerra, Piotr Słowiński, Kyle C. A. Wedgwood, Gáspár Jékely*"
-date: "*Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4
-  4QD, UK*"
+author: 
+  - name: "Luis Alberto Bezares Calderón<sup>1,`*`</sup>, Réza Shahidi<sup>1,2</sup> and Gáspár Jékely<sup>1,3,`*`</sup>"
+    affiliation: "<sup>1</sup>Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4 <br> <sup>2</sup>Electron Microscopy Core Facility (EMCF), University of Heidelberg, 69120 Heidelberg, Germany <br> <sup>3</sup>Centre for Organismal Studies (COS), University of Heidelberg, 69120 Heidelberg, Germany  <br> <sup>*</sup>Correspondence`:` l.a.bezares-calderon @ exeter.ac.uk, gaspar.jekely @ cos.uni-heidelberg.de"
 output:
+  html_document:
+    theme: lumen
   pdf_document:
@@
     template: default
-  html_document:
-    theme: lumen
   word_document: default
@@
 
-![**Figure 1.
+```{r echo=FALSE,fig.cap="**Figure 1.
 Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** **(A)** SEM image of a three-day-old Platynereis larva.
@@
 Views of larvae at three-day-old from the anterior side.
-Insert: One of the reconstracted INNOS is shown from the anterior side.
+Insert: One of the reconstructed INNOS is shown from the anterior side.
 **(F)** Wiring diagram of the cPRC circuit.
@@
 Numbers inside hexagons indicate the number of cells grouped in each node.
-Arrow line thickness (synapse weight) is equal to the number of synapses.](../figures/Fig1.png){width="100%"}
+Arrow line thickness (synapse weight) is equal to the number of synapses."}
+
+knitr::include_graphics(path = "Manuscript/figures/Fig1.png")
+
+```
 
@@
 **(G)** Changes over time in GCaMP6s intensities in INRGW when cPRC is stimulated with UV in NIT-GC2 morphants (MO1 and MO2: blue and light blue).
-**(H, I)** Changes over time in GCaMP6s intensities in Ser-h1 and MC when cPRC is stimulated with UV in WT (Ser-h1 and MC: dark red and brown) and NOSΔ11/Δ11 mutant (Ser-h1 and MC: light red and brown) larvae.](../figures/Fig5.png){width="100%"}
+**(H, I)** Changes over time in GCaMP6s intensities in Ser-h1 and MC when cPRC is stimulated with UV in WT (Ser-h1 and MC: dark red and brown) and NOSΔ11/Δ11 mutant (Ser-h1 and MC: light red and brown) larvae.](Manuscript/figures/Fig5.png){width="100%"}
 

### 2023-05-06 Manuscript/Manuscript.tex
upload ms in Manuscript/

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  10pt,
  onecolumn]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
  \setmainfont[]{Arial}
  \setmonofont[]{Times New Roman}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=0.8in,right=0.8in,top=0.8in,bottom=0.5in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates UV avoidance in zooplankton},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates
UV avoidance in zooplankton}
\author{true}
\date{}

\begin{document}
\maketitle

\setstretch{1.2}
\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

Nitric oxide (NO) generated by nitric-oxide synthase (NOS) is a key
regulator of animal physiology. Here we uncover a role for NO in shaping
circuit dynamics to orchestrate light-avoidance behaviour. We studied
UV-avoidance mediated by brain ciliary photoreceptors (cPRCs) in larval
\emph{Platynereis dumerilii}, a marine annelid. We found NOS
specifically expressed in interneurons (INNOS) postsynaptic to cPRCs.
Stimulation of cPRCs by violet light leads to cPRC inhibition but
concomitant INNOS activation and NO production. NO feeds back to cPRCs
and triggers their delayed activation through an unconventional guanylyl
cyclase. This results in the activation of projection interneurons and
the inhibition of serotonergic ciliomotor neurons. In \emph{NOS}
mutants, NO feedback and projection-neuron activation do not occur and
avoidance behaviour is defective. By mathematical modelling, we could
capture the circuit dynamics in both wild-type and mutant larvae. Our
results reveal how NO-mediated neuroendocrine signalling gates a
synaptic circuit to initiate light-avoidance behaviour.

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

The ability to perceive changes in the environment and adapt behaviour
accordingly is essential for the survival of all living organisms.
Avoidance from UV radiation, which damages DNA essential for survival,
is proposed to be one of the behaviours that existed before the abundant
development of Earth's oxygen to block UV radiation in the atmosphere
(Pittendrigh). In the ocean, almost all zooplankton have a pronounced
rhythmic behaviour called Diel vertical migration (DVM), which consists
of upwards towards the surface at dusk and descend back to deeper water
before dawn. Avoidance behaviour to escape harmful UV radiation during
the day is closely linked to the control of vertical migration in the
ocean. However, little is known about the molecular neural circuit
mechanisms in UV avoidance behaviour in many animals, including
zooplankton. In neural circuits in general light response, the
activation of various receptors by excitatory and inhibitory
neurotransmitters such as glutamate, GABA and dopamine is known in
various species. In addition to this, it has long been known that in
nitric oxide (NO), as a neuromodulatory molecule, it is also important
as a modulator of visual activity. NO is a small free-radical molecule
with diffusible and changeable characteristics, and thus works as one of
the neural signalling molecules. The function of NO in the neural
circuitry of vision has been well studied in vertebrates. In the
mammalian retina, NO is an essential signalling molecule in all retinal
cells. NO produced by nitric oxide synthase (NOS) in amacrine cells, one
of the retinal cells, activates \emph{soluble guanylate cyclase (sGC)}
expressed in cone bipolar cells, which produces cGMP and regulates
downstream signalling. We therefore hypothesise that NO-mediated neural
signalling may be involved in the photoreceptor circuits that control UV
avoidance behaviour. To test this hypothesis we used the annelid
\emph{Platynereis dumerilii} \textbf{(Figure 1A)}, which conserved
vertebrate-type ciliary photoreceptors (cPRCs) (Arendt et al., 2004).
The cPRC expresses \emph{c-opsin1}, which absorbs UV radiation, is
involved in melatonin production and regulates circadian swimming
behaviour (Tosches et al., 2014; Tsukamoto et al., 2017). Furthermore,
we previously reported abnormal UV avoidance behaviour in
\emph{c-opsin1} knockout larvae (Verasztó et al., 2018). The cPRCs
express the UV-absorbing c-opsin1 and mediate UV avoidance in the larvae
(Verasztó et al., 2018). In adult worms, UV light and c-opsin1 modulate
locomotion and maturation (Veedin Rajan et al., 2021). These suggest
that Platynereis larvae are a highly suitable model for elucidating the
neural circuit mechanisms of NO-induced UV avoidance behaviour, and the
accessibility of genome editing, behavioural assays, calcium imaging and
whole-body connectomics could be very useful for study at the molecular
level(Ozpolat et al., 2021).

\hypertarget{results}{%
\subsection{Results}\label{results}}

\textbf{Nitric oxide synthase is expressed in interneurons of the
UV-avoidance circuit}

We identified a single \emph{nitric oxide synthase} (\emph{NOS}) gene in
the \emph{Platynereis dumerilii} genome and transcriptome data.
Phylogenetic analysis of NOS proteins indicate that \emph{Platynereis}
NOS belongs to an orthology group of metazoan NOS sequences
\textbf{(Figure 1---figure supplement 1)}. To characterise the
expression pattern of \emph{NOS} we used in situ hybridization chain
reaction (HCR) and transient transgenesis. In two- and three-day-old
larvae, we detected \emph{NOS} expression in four cells (two of them
weakly expressing) in the apical organ region \textbf{(Figure 1D and
Figure 1---figure supplement 2)}. The same four cells were labelled with
a reporter construct driving palmitoylated-tdTomato by a 12 kbp upstream
genomic region of the \emph{NOS} gene \textbf{(Figure 1E)}. This
reporter also revealed the axonal projections of the NOS-expressing
neurons. The position and morphology of the four \emph{NOS}-expressing
cells allowed us to identify the same cells in the three-day-old
whole-body volume EM data (Verasztó et al., 2020; Williams et al., 2017)
as four interneurons (INNOS) \textbf{(Figure 1B,C)}. The INNOS cells are
postsynaptic to the UV-sensory cPRCs and presynaptic to the INRGW
interneurons, which are also cPRC targets \textbf{(Figure 1C,F)}. INRGW
neurons synapse on the head serotonergic ciliomotor neurons (Ser-h1),
which synapse on the prototroch ciliary band and the cholinergic MC
ciliomotor neurons \textbf{(Figure 1C,F)}. We also detected \emph{NOS}
expression by HCR in the region of the visual eyes (adult eyes) and the
pigmented eyespots \textbf{(Figure 1---figure supplement 1)}.

\begin{figure}
\includegraphics[width=25in]{figures/Fig1} \caption{**Figure 1. Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** **(A)** SEM image of a three-day-old Platynereis larva. **(B, C)** Five types of neurons (cPRC, INNOS, INRGW, Ser-h1 and MC) in the cPRC circuit reconstructed in the whole-body transmission electron microscopy (ssTEM) dataset of a three-day-old larva. Ciliated cells labelled with (B) are shown in grey. The INNOS on the ventral side labelled with (C) are shown in purple. Axons and dendrites appear as lines and cell body positions are represented by spheres. **(D)** Expression analysis of the NOS gene (magenta) using in situ HCR. Anterior view of the larva at two-day-old. Co-stained image using acetylated α-tubulin antibody (acTub: green). **(E)** Immunostaining with anti-HA antibody against HA tagged palmitoylated-tdTomato expressed under the upstream of the start site of NOS (NOSp::palmi-3xHA: magenta). Co-staining with acetylated α-tubulin antibody as marker for cilia of cPRC and axonal scaffold (acTub: green). Views of larvae at three-day-old from the anterior side. Insert: One of the reconstructed INNOS is shown from the anterior side. **(F)** Wiring diagram of the cPRC circuit. Hexagons and arrows represent cell groups and their synaptic connections, respectively. Numbers inside hexagons indicate the number of cells grouped in each node. Arrow line thickness (synapse weight) is equal to the number of synapses.}\label{fig:unnamed-chunk-2}
\end{figure}

\textbf{Nitric oxide produced during UV/violet stimulation gates the
output of the cPRC circuit}

The specific expression of \emph{NOS} in interneurons directly
postsynaptic to the cPRCs suggests that NO signalling is involved in the
UV response. To test if NO is produced during UV/violet responses, we
injected zygotes with the fluorescent NO-reporter DAF-FM and imaged
two-day-old larvae during the local stimulation of the cPRC cilia by 405
nm light. Following light stimulation, we detected an increase in DAF-FM
fluorescence in the region of the INNOS cells \textbf{(Figure 2)}.

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{../figures/Fig2.png}
\caption{\textbf{Figure 2. NO produced by UV/violet stimulation to
cPRCs.} \textbf{(A)} A NO-positive cell visualised by DAF-FM (green) is
shown (dashline). The white line indicates the outer frame of the larva.
All subsequent images are of two-day-old larvae. \textbf{(B)} Intensity
changes before (upper) and during 405 nm light stimulation (lower) in a
cell visualised by DAF-FM. \textbf{(C)} The intensity of DAF-FM changes
over time when cPRC (green) and other cells (ctr: gray) are stimulated
with UV/violet (405 nm) light. Purple squares indicate the timing of 405
nm light stimulation, the transparent lines show the optical responses
to individual intensities, each intensities are normalized (ΔF/F0) and
the dark trace lines and the shaded bands show the mean value and the
standard error range of the mean.}
\end{figure}

\textbf{Nitric oxide signalling mediates UV-avoidance behaviour}

To test this, we generated two \emph{Platynereis} \emph{NOS} knockout
lines with the CRISPR/Cas9 method. We recovered two deletions
(NOSΔ11/Δ11 and Δ23/Δ23), both leading to an early stop codon and thus
likely representing null alleles \textbf{(Figure 3A)}. We could
establish a homozygous line for both mutations indicating that NOS is
not an essential gene. To quantify UV avoidance, we recorded the
trajectories of freely swimming larvae in vertical columns, illuminated
laterally from two opposite sides with 395 nm UV/violet light. To
analyse the behaviour of larvae before and after UV stimulation, the
temporal changes in some parameters were compared using tracking of
individual larvae of the WT and mutant \textbf{(Figure 3B,C and Figure
3---figure supplement 1A-D)}. The results showed that in the WT there
was a rapid and strong increase in downward swimming approximately 15 s
after UV stimulation, whereas this did not occur in the mutant
\textbf{(Figure 3C and Figure 3---figure supplement 1B)}. Interestingly,
there was already a difference in swimming speed between the WT and the
mutant before UV stimulation (in the dark) \textbf{(Figure 3---figure
supplement 1C,D)}. After UV stimulation, the swimming speed of the WT
increased and reached a maximum after about 20 s. In contrast, the
mutants showed a slight increase in speed after UV stimulation, but
swimming speed was not maintained and returned to the speed before UV
stimulation. We also tested phototaxis, by exposing larvae to 480 nm
directional and collimated light from the top of the column. As
previously shown, wild-type larvae swim downward within 30 sec following
non-directional UV/violet light stimulation (Verasztó et al., 2018). In
contrast, homozygous \emph{NOS}-mutant larvae showed a strongly
diminished UV-avoidance response \textbf{(Figure 3D and Figure
3---figure supplement 1E)}. This phenotype is similar to the defective
UV-avoidance of \emph{c-opsin1} mutant larvae and reveals a requirement
for NOS in UV-avoidance behaviour (Verasztó et al., 2018).
\emph{NOS}-mutant larvae also showed reduced phototactic behaviour,
suggesting a function in the visual eyes mediating phototaxis (Randel et
al., 2014), where \emph{NOS} is also expressed \textbf{(Figure
1---figure supplement 1A,B)}. To test if the observed defects are due to
a role of NOS in neuronal signalling or circuit developmental, we also
tested larvae exposed to a NOS inhibitor, L-NAME. Larvae incubated for 5
min in 0.1 mM or 1 mM L-NAME showed defective UV avoidance, but normal
phototaxis \textbf{(Figure 3E)}. Overall, our results indicate an acute
requirement for NOS signalling in UV-avoidance and a possible indirect,
developmental role in the visual system. To investigate the role of NO
on UV avoidance behaviour, receptors for NO were explored.

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{../figures/Fig3.png}
\caption{\textbf{Figure 3. Larvae with inhibited NOS have a delayed
response to UV-avoidance.} \textbf{(A)} Top: The domain and Exon/Intron
structure of NOS. Bottom: Close-up region showing the genomic locus of
NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9. The
generated mutants (NOSΔ11, NOSΔ23) are also shown. Pink indicates target
sites. Gray shows PAM sequences, red shows stop codons. \textbf{(B)}
Overlaid trajectories for WT (n=32) and NOS mutant (NOSΔ11/Δ11, n=26 and
NOSΔ23/Δ23, n=47) at three-day-old larvae. 0 sec as the starting point.
After 10 sec, UV (395 nm) stimulation from the side. \textbf{(C)} The
temporal changes in the vertical position of the WT and mutant larvae
before (dark) and after UV stimulation (side 395 nm) are shown. The
starting points of each larval trajectory are set to 0. \textbf{(D)}
Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23)
larvae at three-day-old stimulated with UV (395 nm) light from side,
blue (488 nm) light from top and UV (395 nm) light from top. The data
are shown in 30 s bins. \textbf{(E)} Vertical swimming in larvae treated
with NOS inhibitors, L-NAME at three-day-old stimulated with UV (395 nm)
light from side, blue (488 nm) light from top and UV (395 nm) light from
top. The data are shown in 30 s bins.}
\end{figure}

\textbf{Two unconventional guanylyl cyclases are expressed in the cPRCs}

NO generally acts via soluble guanylate cyclases (sGC), belonging to the
guanylate cyclase family with a CYC domain (PF00211). NO binding to the
heme group of sGC leads to increased cyclic guanosine monophosphate
(cGMP) production. Analysis of sGCs in \emph{Platynereis} indicated that
these genes are not expressed in any of the cells of the cPRC circuit
(Verasztó et al., 2017). Recently, Moroz and coworkers reported an
atypical but widely conserved family of guanylyl cyclases with a NIT
(nitrite/nitrate sensing) domain (PF08376) (NIT-GC) as potential
mediators of NO signalling (Moroz et al., 2020). To search for NIT-GCs
in \emph{Platynereis}, we searched transcriptome resources and retrieved
15 potential NIT-GC homologs \textbf{(Figure 4---figure supplement 1)}.
To analyse the relationship of these sequences to metazoan NIT-GCs, we
retrieved protein sequences with a CYC domain from the transcriptome and
genome databases of 45 metazoan species and 2 choanoflagellate species.
We carried out cluster analysis and did phylogenetic reconstruction on a
group of membrane-bound guanylyl cyclases with sGCs as an outgroup. In
agreement with Moroz et al. (Moroz et al., 2020),we found a group of GCs
with NIT domains with representatives in placozoans, cnidarians, some
ecdysozoans, echinoderms, and lophotrochozoans. The 15
\emph{Platynereis} sequences were parts of several major and ancient
clades of the tree \textbf{(Figure 4---figure supplement 1)}. To
characterise the expression of NIT-GCs, we using previously published
spatially mapped single-cell transcriptome data (Achim et al., 2015;
Williams et al., 2017). Among the 15 NIT-GCs, two showed high and
specific expression in the cPRCs and one was expressed in the INNOS
cells \textbf{(Figure 4---figure supplement 1)}. In the single-cell
data, we could identify the cPRCs by the specific expression of c-opsin1
and a pedal-peptide2 neuropeptide precursor (MLD proneuropeptide)
(Arendt et al., 2004; Williams et al., 2017), which have previously been
described as cPRC markers \textbf{(Figure 4---figure supplement 2A)}.
The INNOS cells were identified by NOS expression and their spatial
mapping. We decided to focus on two NIT-GCs expressed in the cPRCs and
with a full-length sequences, NIT-GC1 and NIT-GC2. To confirm the
single-cell data, we first carried out in situ HCR with probes for
\emph{NIT-GC1} and \emph{NIT-GC2} mRNA. Both genes showed very clear and
specific expression in the four cPRCs, as confirmed by co-labeling with
an acetylated α-tubulin antibody and with an HCR probe against \emph{MLD
proneuropeptide} \textbf{(Figure 4A,B and Figure 4---figure supplement
2A-C)}. To analyse the subcellular localisation of NIT-GC1 and NIT-GC2
at the protein level, we raised and affinity-purified polyclonal
antibodies against a specific peptide sequence from both proteins. In
immunostainings, we found that NIT-GC1 was localise to the region
corresponding to the axonal projections of the cPRCs in the anterior
nerve plexus \textbf{(Figure 4C)}. Co-immunostaining with the rabbit
NIT-GC1 and the rat NOS antibodies revealed that both proteins were
localised in close proximity in the neurosecretory plexus
\textbf{(Figure 4---figure supplement 2D)}. In contrast, NIT-GC2
strongly and specifically labelled the large sensory ciliary region of
the cPRCs \textbf{(Figure 4D)}. These different subcellular
localisations suggest that the two NIT-GCs are involved in different
intracellular signalling processes in the ciliary and axonal regions of
the cPRCs.

\textbf{NIT-GC1 is required for the depolarising response of cPRC to UV
stimulation}

To investigate how NIT-GCs-mediated NO signalling is involved in cPRC
activation, the calcium sensor GCaMP6s was ubiquitously expressed in
larvae and calcium imaging in the cPRC during UV/violet stimulation. As
we have shown previously, a 20-sec local stimulation of the ciliated
part of the cPRC with 405 nm light lead to a transient decrease in cPRC
calcium levels, likely due to the hyperpolarisation of the cPRCs
(Verasztó et al., 2018). After 20-sec, calcium levels in cPRCs were
raising again, reaching higher levels than at the start of the stimulus
- a response that may involve depolarisation. This activation phase
occurs after the end of the stimulation and may be due to a slower
neuroendocrine feedback (Verasztó et al., 2018). First, to determine
whether NO mediates such a feedback, we measured calcium levels in
\emph{NOS}-mutant larvae. While we detected the inhibition phase in cPRC
in \emph{NOS}-mutants, this was not followed by activation and calcium
levels remained low \textbf{(Figure 4E,F)}. These results revealed an
essential requirement for NO signalling in the second phase of the cPRC
response. Next, to characterise the role of NIT-GCs in changes of
calcium levels in cPRC, calcium imaging was performed when each
\emph{NIT-GC} gene was knocked down using morpholino \textbf{(Figure
4---figure supplement 2E,F)}. In NIT-GC1 morphant larvae, 405 nm light
stimulation caused a decrease in calcium levels in the cPRC, but the
increase in calcium levels that occurred after the light stimulation was
stopped was strongly suppressed \textbf{(Figure 4G)}. This result is
similar to the response of NOS mutants in cPRC, suggesting that NIT-GC1
mediates depolarising responses after UV stimulation of cPRC by NO and
may be involved in extracellular signalling. Interestingly, in NIT-GC2
morphant larvae, an increase in calcium levels occurred immediately
after light stimulation \textbf{(Figure 4H)}. The different responses of
cPRC upon knockdown of NIT-GC1 and NIT-GC2 suggest that there may be
different intracellular signalling mechanisms associated with their
respective localisations. We focused our analysis on NIT-GC1 because it
was localised to cPRC neuroaxons within the neuroplexas region and is
involved in extracellular signalling after UV stimulation. It is
generally known that depolarisation in the PRC involves elevated cGMPs
in the photoreceptors. We therefore used an in vitro system in which
cultured cells expressed NIT-GC1 to assess the cGMP activity of NIT-GC1.

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{../figures/Fig4.png}
\caption{\textbf{Figure 4. cPRC-specific expression of NIT-GC1 localises
in close proximity to NOS in the ANS region.} \textbf{(A, B)} Expression
analysis of the NIT-GC1 (C) and NIT-GC2 (D) gene (magenta) using in situ
HCR. Co-staining image using acetylated α-tubulin antibody (ac α-tub:
green). \textbf{(C, D)} Localisation analysis using NIT-GC1 and NIT-GC2
antibodies. Green shows co-staining with acetylated α-tubulin antibody
(ac α-tub). \textbf{(E, F)} Changes over time in GCaMP6s intensities in
cPRC when stimulated with UV in WT (dark yellow) and NOS mutant (E:
NOSΔ11/Δ11, F: NOSΔ23/Δ23) (light pink) larvae. \textbf{(G, H)} Changes
over time in GCaMP6s intensities in cPRC when stimulated with UV in
NIT-GC1 (G: NIT1) and NIT-GC2 (H: NIT2) knockdown by morpholino (MO1 and
2) larvae. \textbf{(I)} The domain structure of NIT-GC1. Below shows the
mutant NIT-GC1 (ΔNIT-GC1) with deletion of the NIT domain. Grey squares
indicate the transmembrane region (TM). \textbf{(J)} A schematic of the
NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP
indicator Green cGull, and the NO donor SNAP and DMSO as control. Green
cGull shows cGMP-dependent green fluorescence. \textbf{(K-N)} The
intensity changes in cGMP fluorescence over time for the four conditions
are shown. Gray traces show the optical responses to individual
intensities (n \textgreater{} 6). The individual intensities are
normalized (ΔF/F0). The dark line shows the mean value and the band the
standard error range of the mean. The respective solution was acted on 2
min after the start of imaging (grey bars).}
\end{figure}

\textbf{NIT-GC1 produces cGMP by nitric oxide}

NIT-GC1 has a highly conserved cyclase domain and activation of NIT-GC1
is expected to produce cGMP \textbf{(Figure 4I)}. Recently, a synthetic
protein called Green cGull was developed as a cGMP indicator that can be
used in cultured cell systems (Matsuda et al., 2016). We co-expressed
Green cGull and NIT-GC1 using the 2A self-cleaving peptide in cultured
cells COS-7 (monkey kidney) and tested the intracellular dynamics of
cGMP in the activation of NIT-GC1 \textbf{(Figure 4J)}. When the NO
donor SNAP was applied to the NIT-GC1 and Green cGull co-expression
cells, the intracellular cGMP levels was significantly increased
compared to when NIT-GC1 was not expressed \textbf{(Figure 4K,L)}. The
increase in fluorescence intensity also did not occur when DMSO was
applied \textbf{(Figure 4M)}. To confirm whether this increase in
intracellular cGMP levels is NIT domain-dependent response of NIT-GC1,
mutant form of NIT-GC1 without the NIT domain was expressed and SNAP was
applied \textbf{(Figure 4I)}. As a result, the increase in fluorescence
intensity was significantly reduced \textbf{(Figure 4N)}. These results
indicate that NIT-GC1 produces cGMP via the NIT domain. In bacteria, it
has been proposed that the NIT domain regulates cellular functions in
response to changes in extracellular and intracellular nitrate and
nitrite concentrations. Our results suggest that NIT-GC1 may be
sensitive to NO. This could be a possible feedback pathway in which
NIT-GC1 expressed in cPRC is activated by NO produced by INNOS. In order
to further explore the role of NO in the cPRC circuit, we therefore
compared calcium imaging of postsynaptic cells (INNOS, INRGW, Ser-h1 and
MC) in the cPRC circuit during UV stimulation of cPRC between WT and NOS
mutant larvae.

\textbf{Depolarisation of cPRC is required for activation of downstream
interneurons and ciliomotor neurons}

We next quantified the responses of the INNOS and INRGW following a
UV/violet stimulus to the cPRCs. To unambiguously identify the cells
from which we recorded calcium signals, we developed an on-slide
immunostaining method \textbf{(Figure 5---figure supplement 1)}. We used
the cell-specific neuropeptide markers using anti-RYamide (INNOS) and
anti-RGWamide (INRGW) antibodies (Conzelmann et al., 2011) to mark the
interneurons. Based on the position of the nuclei, we could correlate
live and fixed samples at a single-cell precision \textbf{(Figure
5A,B)}. In both wild-type and \emph{NOS}-mutant larvae, INNOS cells
showed increase calcium signals during the inhibition phase of the cPRCs
\textbf{(Figure 5D)}. In addition, INNOS depolarisation was suppressed
in NIT-GC2 molphants where cPRC hyperpolarisation was suppressed
\textbf{(Figure 5E)}. This indicates that cPRC hyperpolarisation is
required for INNOS depolarisation. In contrast, calcium signals
increased in INRGW cells in wild-type larvae during the second
activation phase of cPRCs, a response that was lacking in
\emph{NOS}-mutants \textbf{(Figure 5F,G)}. This suggests that there is
no circuit output to the Ser-h1 and MC cells and the ciliary band in
\emph{NOS}-mutants, consistent with the observed behavioural defects
\textbf{(Figure 5H,I)}.

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{../figures/Fig5.png}
\caption{\textbf{Figure 5. NO produced by UV stimulation to cPRCs
induces a depolarising response in cPRC via NIT-GC1.} \textbf{(A, B)}
Each left figure shows a calcium imaging using GCaMP6s. The respective
right figures show immunostaining images of the same positions in the
respective left figures, (A) shows RYamide antibodies, (B) shows
RGWamide and serotonin (5-HT) antibodies in red and DAPI in cyan.
Asterisks indicate the nuclei of the cPRC, respectively, and the nuclei
of the cells that were identified from the arrangement of the
surrounding cells are indicated by numbers. From the position of the
cells, the calcium imaging data identify INNOS in (A) and INRGW and
Ser-h1 in (B), respectively. \textbf{(C)} Correlation (Pearson's r) map
of neuronal activity of the INNOS, INRGW, Ser-h1 and MC neurons.
\textbf{(D)} Changes over time in GCaMP6s intensities in INNOS when cPRC
is stimulated with UV in WT (dark purple) and NOSΔ11/Δ11 mutant (light
purple) larvae. \textbf{(E)} Changes over time in GCaMP6s intensities in
INNOS when cPRC is stimulated with UV in NIT-GC2 morphants (MO1 and MO2:
pink and light pink) larvae. \textbf{(F)} Changes over time in GCaMP6s
intensities in INRGW when cPRC is stimulated with UV in WT (dark blue)
and NOSΔ23/Δ23 mutant (light blue) larvae. \textbf{(G)} Changes over
time in GCaMP6s intensities in INRGW when cPRC is stimulated with UV in
NIT-GC2 morphants (MO1 and MO2: blue and light blue). \textbf{(H, I)}
Changes over time in GCaMP6s intensities in Ser-h1 and MC when cPRC is
stimulated with UV in WT (Ser-h1 and MC: dark red and brown) and
NOSΔ11/Δ11 mutant (Ser-h1 and MC: light red and brown) larvae.}
\end{figure}

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

In the present study, we explored the role of NO in the UV avoidance
behaviour of Platynereis. We demonstrated that there are interneurons
expressing NOS in the cPRC circuit and that NO is required for UV
avoidance in behavioural experiments using NOS mutants. We identified
the axon-localised type NIT-GC1 and the cilia-localised type NIT-GC2
expressed in the cPRCs as targets of NO and showed from cultured cell
experiments that these produce cGMP in response to NO. We have shown by
calcium imaging that INNOSs are activated in response to the
hyperpolarising response of the cPRCs, that the NO produced feeds back
to the cPRC via NIT-GC1 and induces a depolarising response in the cPRC,
and that the depolarising response of the cPRC is involved in activating
downstream INRGW. Lastly, we confirmed that RGWamide induces the
downward swimming behaviour of larvae.

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{../figures/Fig6.png}
\caption{\textbf{Figure 6. Gene expressions and signalling mechanisms
between cPRC, INNOS and INRGWa.} \textbf{(C)} Dot plot of genes
(columns) expressed in three types of cells (rows) in the cPRC circuit
using single cell RNA-Seq. The size of the dots is expressed in
proportion to the percentage of cells expressing that gene relative to
all cells. The colours represent the normal logarithm of the number of
transcripts in the cells expressing the gene. \textbf{(D)} Schematic
diagram of the signalling pathway of the cPRC circuit, focusing on the
NO feedback.}
\end{figure}

cPRCs expressing UV-absorbing c-opsin1 have high resting calcium levels
and show hyperpolarising and subsequent depolarising responses to UV
light (405 nm) (Verasztó et al., 2018). In the present study, we have
shown that the cells that strongly synapse with cPRCs are interneurons
that specifically express NOS (INNOS). The fluorescence intensity of
DAF-FM, which specifically detected INNOS, showed a transient tonic-like
increase dependent on UV stimulation of the cPRCs. The graph shows a
decrease in intensity after UV light is switched off, but the response
of DAF-FM to NO is irreversible and this decrease is due to breaching by
the relatively strong excitation light. Attempts to measure changes in
the intensity of the calcium indicator RGECO1a simultaneously with
DAF-FM imaging did not lead to detection of changes in fluorescence
intensity compared to GCaMP6s and could only visualise the cPRCs. In
calcium imaging using GCaMP6s, INNOS identified by the positional
relationship between the cPRC and INNOS and the labelling of RYa
depolarised in clear synchrony with the hyperpolarising response of the
cPRCs. This tonic response may involve dopaminergic nerves based on
single-cell data \textbf{(Figure 6)} {[}Puopolo et al. (2001){]}(Zhang
et al., 2007)(Randel et al., 2014). In the mammalian retina, it is known
that signals from bipolar photoreceptor cells are transmitted to
downstream NOS-expressing amacrine cells (NOAC), where NO is produced
(Jacoby et al., 2018). These suggest that UV stimulation of cPRCs causes
the production of NO via a depolarising response of INNOS.

Our results suggest that NIT-GC1, which is specifically expressed in
cPRCs, functions as a target for NO produced by INNOS and contributes to
the regulation of the cPRC circuit transmission pathway via a
depolarising response by increasing cGMP. In the mammalian retina, NO is
an essential signalling molecule and one of the most well-studied
(Cudeiro and Rivadulla, 1999). It has been reported that NO produced in
one of the retinal cells modulates cGMP-mediated neurotransmission by
activating presynaptic neural sGCs (general target of NO) through
retrograde signalling {[}Vielma et al. (2014){]}(Wei et al., 2012). In
Platynereis cPRC, NIT-GC1 was also identified as having the property of
producing cGMP in a NO-dependent manner. Furthermore, it is known that
CNGAα, one of the cyclic nucleotide-gated channels, is specifically
highly expressed in cPRC (Tosches et al., 2014), and in INNOS-cPRC
signalling, NO is implicated in the neuromodulation of the cPRC by
retrograde signalling. NO is a molecule with a very short half-life, and
when functioning as a signalling molecule, NOS and sGC are known to
localise very close together at synapses between neurons {[}Burette et
al. (2002){]}(Garthwaite, 2015). In our study, we found 12 sGCs in
Platynereis, but none expressed in the cPRC. Localisation analysis of
NOS and NIT-GC1 by antibody staining showed that the two proteins
localised very close together in the anterior neurosecretory plexus
region. Calcium imaging results using NOS mutant and NIT-GC1 knockdown
larvae suggest that NIT-GC1 activation induces a depolarising response
in the cPRC.

Another type of interneuron that strongly synapses with the cPRC is the
INRGWa, which synapses directly onto the ciliomotor neuron and is
therefore considered to be a very important neuron in the control of
larval behaviour {[}Verasztó et al. (2018){]}(Williams et al., 2017).
The present results suggest that the depolarising response of the cPRC
due to NO feedback induces INRGW activation. RGWamide is also a
neuropeptide that induces larval downward swimming quite strongly
(unpublished? need to show data?), and the delayed downward swimming of
NOS mutant larvae in response to UV stimulation may be responsible for
the failure of INRGW activation to occur. In addition, in many marine
invertebrates, NO has been reported to be involved in many aspects of
settlement behaviour and metamorphosis induction (Bishop and Brandhorst,
2007, 2003; Castellano et al., 2014; Leise et al., 2001; Pechenik et
al., 2007; Song et al., 2021; Ueda et al., 2016; Ueda and Degnan, 2014,
2013; Yang et al., 2018; Zhu et al., 2020). The downward swimming
behaviour of Platynereis larvae is the first step in the onset of
settlement behaviour and metamorphosis. These suggest that downward
swimming due to NO activation of INRGWs may trigger the onset of
settlement behaviour and metamorphosis {[}Conzelmann et al.
(2013b){]}(Williams et al., 2015). These suggest that NO feedback
signalling may act as a trigger for downward swimming via activation of
INRGWs, which induces the initiation of settlement behaviour and
metamorphosis.

An outstanding question is how the NO\textasciitilde NIT-GC response
alters neuronal dynamics. NO is a single molecule and, due to its short
half-life, cannot be transmitted over long distances and has a limited
signalling repertoire. It has therefore been considered unsuitable for
generating diversity in signalling compared to other signalling
molecules such as neuropeptides and classical neurotransmitters (Jékely,
2021). However, our results in the present study indicate that
NO-induced activation of NIT-GC1 is involved in the regulation of
neuronal circuits (cPRC pathway). The NIT domain was originally found in
bacteria and the existence of GCs with NIT domains was first reported by
Moroz et al in 2020 (Moroz et al., 2020; Shu et al., 2003). It has been
proposed that NIT domains in bacteria regulate cellular functions in
response to changes in extracellular and intracellular nitrate and/or
nitrite concentrations (Camargo et al., 2007). In fact, NO is a free
radical molecule, with a half-life of 5 msec in vivo, and is readily
converted to nitrate and nitrite (Garthwaite, 2015; Möller et al., 2019;
Santos et al., 2011). And nitrate and nitrite have been reported to
accumulate in cells and tissues with high NOS activity in placozoans and
cnidarians (Moroz et al., 2020, 2004). It is possible that the NIT
domains of NIT-GC1 identified in this study may also be used for
nitrate/nitrite sensing, as in bacteria. Furthermore, if they show
different sensitivities to NO, nitrite and nitrate, it is possible that
the temporal variation in their respective half-lives could give rise to
a repertoire of activation timings (Lundberg et al., 2011). In addition,
we have found at least 15 NIT-GCs in the Platynereis transcriptome data,
of which NIT-GC2 is highly expressed in the cPRC, in addition to
NIT-GC1. Interestingly, NIT-GC2 showed a very different localisation
pattern to NIT-GC1, and furthermore, the results of calcium imaging of
cPRC in response to UV stimulation using NIT-GC2 knockdown larvae were
the different dynamics (such that hyperpolarisation was suppressed) with
NIT-GC1 knockdown. It was not clear from the present analysis whether
this difference was due to differences in the localisation or the
property of the NIT-GCs. However, the diversification and spatially
different localisation patterns of NIT-GCs may increase their signalling
repertoire. We also believe that the appearance of NIT-GCs was one of
the means to generate the diversity of signalling pathways using NO
during the early stages of nervous system evolution in metazoans, since
sponges and ctenophores, where NO signalling is present, have no NIT-GCs
at all, as many as 12 have been found in placozoans, and NIT-GCs are
functional in the ``proto-eye'' circuit of Platynereis.

\hypertarget{materials-and-methods}{%
\subsection{Materials and Methods}\label{materials-and-methods}}

\textbf{CRISPR-Cas9 Design and Microinjection}

Before designing the small guide RNA (sgRNA) for the sgRNA:Cas9
nuclease, splice sites and polymorphic sites in our laboratory culture
were identified to avoid them. The sgRNA targeted the third exon of
Platynereis dumerilii NOS (target site: 5'-GGGCAATACTGGCTCCACTC-3'). The
sgRNA was assembled from two annealed oligonucleotides
(5'-TAGGGCAATACTGGCTCCACTC-3', 5'-AAACGAGTGGAGCCAGTATTGC-3') forming
overhangs for cloning into a BsaI site of the plasmid pDR27456 (Hwang et
al.~2013)(42250, Addgene), which contains next to the BsaI site a
tracrRNA sequence. The plasmid was then used to PCR amplify DNA
(primers: T7, 5'-AAAAGCACCGACTCGGTGCC-3') for synthesizing the sgRNA.
The DNA was purified with the QIAquick PCR Purification Kit (Qiagen).
From the DNA, the sgRNA was synthesized with the MEGAshortscript Kit
(Thermo Fisher Scientific) and was purified with the MEGAclear Kit
(Thermo Fisher Scientific). Cas9-mRNA was transcribed, capped, and
polyA-tailed with the mMessage mMachine Kit and the Poly(A) Tailing Kit
(both Thermo Fisher Scientific) from a plasmid (pUC57-T7-RPP2-Cas9)
containing the Cas9 ORF fused to 169 base pair 5' UTR from the
Platynereis dumerilii 60S acidic ribosomal protein P2. The sgRNA (18
ng/ml) and the Cas9-mRNA (180 ng/µl) were coinjected into fertilized
eggs of Platynereis dumerilii wild-type parents according to an
established injection procedure (Conzelmann et al., 2013a). The eggs
were kept at 18°C for 45 min before injection and were injected at
14.5°C. The injected individuals were kept at 18°C for 5 to 8 days in
6-well- plates (Nunc multidish no. 150239, Thermo Scientific) and then
cultured at 22°C until sexual maturity. The mature worms were crossed to
wild-type worms and the progeny was genotyped, resulting in two founder
lines, which were bred to homozygosity.

\textbf{NOS sequencing and genotyping}

For genotyping of the NOS locus, genomic DNA was isolated from single
larvae, groups of 6-20 larvae, or from the tails of adult worms. The DNA
was amplified by PCR (primers: 5'-GGTTCATTGGTTTCGATAACATTGCGG-3',
5'-CAGAGTCGATCAGTCTGCATATCTCCA-3') with the dilution protocol of the
Phusion Human Specimen Direct PCR Kit (Thermo Scientific). The PCR
product was sequenced directly with a nested sequencing primer
(5'-GGTGCTCTCCCGGGTACACAA-3'). A mixture of wild-type and deletion
alleles in a sample gave double peaks in the sequencing chromatograms,
with the relative height of the double peaks reflecting the relative
allele ratio in the sample.

\textbf{Vertical column setup for measuring photoresponses}

Photoresponses of larvae of different ages were assayed in a vertical
Plexiglas column (31 mm x 10 mm x 160 mm water height). The column was
illuminated from top with light from a monochromator (Polychrome II,
Till Photonics). The monochromator was controlled by AxioVision 4.8.2.0
(Carl Zeiss MicroImaging GmbH) via analog voltage. The light passed a
collimator lens (LAG-65.0-53.0-C with MgF2 Coating, CVI Melles Griot)
before entering the column. The column was illuminated from both sides
with light-emitting diodes (LEDs). The LEDs on each side were grouped
into two strips. One strip contained UV (395 nm) LEDs (SMB1W-395,
Roithner Lasertechnik) and the other infrared (810 nm) LEDs
(SMB1W-810NR-I, Roithner Lasertechnik). The UV LEDs were run at 4 V to
stimulate the larvae in the column from the side. The infrared LEDs were
run at 8 V (overvoltage) to illuminate the larvae for the camera (DMK
22BUC03, The Imaging Source), which recorded videos at 15 frames per
second and was controlled by IC Capture (The Imaging Source).

\textbf{Comparing behavior of wildtype and NOS-knockout 3-day-old
larvae}

To compare the behavior of wildtype and NOS-knockout larvae at 3 days in
the vertical column, the larvae were mixed and left in the dark for 5
min. The larvae were treated with NOS inhibitors for pharmacology. The
NOS inhibitors were L-NAME. The larvae were treated with different
concentrations in adjacent columns. The concentrations for the NOS
inhibitors were control, 1 mM, 0.1 mM. The larvae were recorded for 1
min in the dark followed by exposure to collimated cyan (480 nm) light
from the top of the column for 2 min, then 2 min darkness, and finally
collimated UV (395 nm) light from the top of the column for 2 min.
Stimulus light was provided by the monochromator (Polychrome II, Till
Photonics). Scripts are available at
\url{https://github.com/JekelyLab/NOS}.

\textbf{NOS Identification and Phylogenetic Analysis}

To identify NOS, we obtained a ``seed'' database of oxygenase domain in
Pfam database, PF02898. From these sequences, we produced a Hidden
Markov Model (HMM) and used this to mine the 47 metazoan species, 2
choanoflagellate species and 2 filasterea species investigated. HMM
models were run in HMMR3 with an e-value of 1e−15. We ran CD-Hit (Fu et
al., 2012) to eliminate redundant sequences (at a 80\% threshold). We
aligned the sequences with MAFFT version 7, with the iterative
refinement method E-INS-i. Alignments were trimmed with TrimAl in
gappy-out mode (Capella-Gutierrez et al., 2009). To calculate
maximum-likelihood trees, we used IQ-tree2 with the LG+G4 model. To
calculate branch support, we ran 1,000 replicates with the aLRT-SH-like
and aBayes methods (Minh et al., 2020). The sequences used for the
phylogenetic analysis are available in \textbf{Supplementary File 2},
the trimmed alignment is available in \textbf{Supplementary File 3}.

\textbf{NIT-GC Identification and Phylogenetic Analysis}

To identify NIT-GCs, we obtained a ``seed'' database of Adenylate and
Guanylate cyclase catalytic domain in Pfam database, PF00211. From these
sequences, we produced a Hidden Markov Model (HMM) and used this to mine
the 45 metazoan species, 2 choanoflagellate species and 2 filasterea
species investigated. HMM models were run in HMMR3 with an e-value of
1e−15. We ran CD-Hit (Fu et al., 2012) to eliminate redundant sequences
(at a 80\% threshold). The CLANS analysis is available as
\textbf{Supplementary File 1}. To identify clusters, we used the
convex-clustering option with 100 jack-knife replicates. The NIT-GCs are
extremely well conserved in membrane-bound guanylate cyclases and form
an easily recognizable cluster. To analyze the phylogeny of NIT-GCs, the
cluster containing these GCs together with membrane-bound guanylate
cyclases were parsed and used for tree building. We aligned the
sequences with MAFFT version 7, with the iterative refinement method
E-INS-i. Alignments were trimmed with TrimAl in gappy-out mode
(Capella-Gutierrez et al., 2009). To calculate maximum-likelihood trees,
we used IQ-tree2 with the LG+G4 model. To calculate branch support, we
ran 1,000 replicates with the aLRT-SH-like and aBayes methods (Minh et
al., 2020). The sequences used for the phylogenetic analysis are
available in \textbf{Supplementary File 2}, the trimmed alignment is
available in \textbf{Supplementary File 3}.

\textbf{Single-cell analysis}

We used Achim et al.~for the single-cell data (Achim et al., 2015). In
fact, in Williams et al., they used 107 cells as neurons by removing
duplicates from Achim et al.~single-cell data, so we used those cells
(Williams et al., 2017). Since the raw data were read count data, we
normalized them to TPM using Python. After that we converted them to
log10. From the sum of the expression levels in 107 cells for each gene,
I calculated the percentage in each cell. We chose the appropriate genes
for now. For each cell, we identified them with marker genes. After
created the data in Python, plotted it using R dot plots. RPKM
calculates the total number of reads per million bp and then divides by
the length of each gene, so it is not possible to compare between
samples. Instead, TPM first divides by the length of the gene and then
divides by the total number of reads per million bases, which allows for
more accurate comparisons between samples. In this case we wanted to
compare between samples, so we used TPM. In fact, TPM is becoming the
mainstream method for many single-cell analyses. The total TPM of each
gene between the samples was used to calculate the percentage of
expressed genes.

\textbf{In situ HCR}

Larvae were fixed and treated with Proteinase K, according to the
conventional WMISH protocol (Tessmar-Raible et al., 2005), with fixation
in 4\% paraformaldehyde/ PTW (PBS with 0.05\% Tween20) for 2 hr at room
temperature, and Proteinase K treatment in 100 µg/ml Proteinase K/ PTW
for 3 min (Tessmar-Raible et al., 2005). Specifically, for the HCR
protocol, samples were processed in 1.5 ml tubes. Probe hybridization
buffer, probe wash buffer, amplification buffer, and fluorescent HCR
hairpins were purchased from Molecular Instruments (Los Angeles, USA).
Hairpins associated with the b2 initiator sequence were labeled with
Alexa Fluor 647, and the hairpins associated with the b3 initiator
sequence were labeled with Alexa Fluor 546. To design probes for HCR, we
used custom software (Kuehn et al., 2021) to create 20 DNA oligo probe
pairs specific to P. dumerilii NOS, NIT-GC1, NIT-GC2, RYa-pNP (GenBank
accession: JF811330.1), c-opsin1 (GenBank accession: AY692353.1), CNGAα
(GenBank accession: KM199644.1), and MLD/pedal2-pNP (GenBank accession:
KF515945.1). The NOS, NIT-GC1 and NIT-GC2 probes were designed to be
associated with the b2 initiator sequence, while the RYa-pNP, c-opsin1,
CNGAα and MLD/pedal2-pNP probes were designed to be associated with the
b3 initiator sequence. For the detection stage, samples were
pre-hybridized in 200 µl of probe hybridization buffer for 1 hr at 37°C,
and then incubated in 250 µl hybridization buffer containing probe
oligos (4 pmol/ml) overnight at 37°C. To remove excess probe, samples
were washed 4× with 1 ml hybridization wash buffer for 15 min at 37°C,
and subsequently 2× in 1 ml 5× SSCT (5× SSC with 0.1\% Tween20) for 5
min at room temperature. For the amplification stage, samples were
pre-incubated with 100 µl of amplification buffer for 30 min, room
temperature, and then incubated with 150 µl amplification buffer
containing fluorescently labeled hairpins (40nM concentration (2ul of
3uM stock in 150ul amplification buffer, snap-cooled as described; (Choi
et al., 2018)) overnight in the dark at 25°C. To remove excess hairpins,
samples were washed in 1 ml 5× SSCT at room temperature, twice for 5
min, twice for 30 min, and once for 5 min. During the first 30 min wash,
samples were counterstained with DAPI (Cat. \#40043, Biotium, USA).

\textbf{Immunohistochemistry}

Whole-mount immunostaining of 2 day old Platynereis larvae fixed with
4\% paraformaldehyde were carried out using primary antibodies raised
against NIT-GC1, NIT-GC2, NOS, RYamide neuropeptide, RGWamide
neuropeptide in rabbit, plus a commercial antibody raised against
acetylated tubulin in mouse (Sigma T7451). The synthetic peptides
contained an N-terminal Cys that was used for coupling during
purification. Antibodies were affinity purified from sera as previously
described (Conzelmann and Jékely, 2012). Immunostainings were carried
out as previously described (Conzelmann and Jékely, 2012). The NOS
promoter (fragment sizes: 12 Kb) was amplified and cloned upstream of
3xHA- Palmi-tdTomato. Larvae injected with promoter constructs (ca. 250
ng/ml) were analysed for reporter expression at 3 days post
fertilization using an AxioImager Z.1 fluorescence wide-field microscope
(Carl Zeiss GmbH, Jena) and immediately fixed for immunostainings. The
protocol followed for immunostaining of HA-tagged reporters was recently
described (Verasztó et al., 2017). Specimens were imaged with a LSM 780
NLO or LSM 880 with Airysan Confocal Microscope (Zeiss, Jena).

\textbf{Calcium imaging}

For calcium imaging, 49--55 hpf larvae were used. Experiments were
performed at room temperature and larvae were immobilised by being
embedded in 2.5\% agarose filtered artificial seawater between a slide
and coverslip spaced with adhesive tape. GCaMP6s mRNA (1 mg/ml) was
injected into zygotes as described previously (Randel et al., 2014).
Larvae were imaged on a Zeiss LSM 880 with Airyscan (with a C-Apochromat
63X/1.2 Corr - water) with a frame rate of 1.88 frame/sec and an image
size of 512 x 512 pixels. The larvae were stimulated in a region of
interest (a circle with ?? pixel diameter) with 405 nm lasers controlled
by the Bleaching mode. The imaging laser had a similar intensity than
the stimulus laser but covered an area that was 10 times larger than the
stimulus ROI.

\textbf{Cell culture experiment}

Green cGull was used for the cGMP assay (Matsuda et al., 2016). A
full-length Pdum-NIT-GC1 and -NIT-GC2 coding sequences were amplified by
PCR starting from a \emph{Platynereis dumerilii} cDNA library and cloned
into the pcDNA3.1(+) vector using the T2A self-cleaving sequence. Cos-7
cells with low expression of endogenous soluble guanylate cyclase were
used as cultured cells for gene expression. This cell line was purchased
from Angio-proteomie (CAT no. cAP-0203). The Cos-7 cells were maintained
at 37 °C in 35mm dishes (Nunc™ Glass Bottom Dishes) containing 3 mL of
DMEM, high glucose glutamax medium (Thermo; Cat. No.~10566016)
supplemented with 10\% fetal bovine serum (Thermo; Cat. No.~10082147).
Upon reaching confluency of approximately 85\%, we transfected the cells
with the plasmid containing Green cGull-T2A-NITGC1. Transfections were
carried out with 150 ng of each plasmid and 0.3 μl of the transfection
Lipofectamine 3000 Reagent (invitrogen; Cat. No.~L3000001). Two days
post-transfection, we removed the culture medium and substituted it for
fresh DMEM-medium. For single-wavelength imaging experiments, cells in
35-mm dishes were washed twice and imaged in modified Ringer's buffer
(140 mM NaCl, 3.5 mM KCl, 0.5mM NaH2PO4 , 0.5mM S-3 MgSO4 , 1.5 mM CaCl2
, 10 mM HEPES, 2 mM NaHCO3 and 5 mM glucose). Dishes were mounted on a
stage heated at 37 °C and imaging was performed using an inverted
microscope (LSM880, Zeiss) equipped with an oil-immersion objective lens
(UApo/340, 40, NA = xx). Images were acquired using a xenon lamp,
460--495 nm excitation filter, 505-nm dichroic mirror and 510-- 550-nm
emission filter (????). S-Nitroso-N-acetyl-D, L-penicillamine (SNAP) was
purchased from Sigma-Aldrich (St.~Louis, MO, USA) . The exposure time of
the EM-CCD camera was controlled by the ZEN software (Zeiss). Images
were acquired every 15 s for 10 min and stimulation was initiated 2 min
after starting image acquisition. Imaging data analysis was performed
using ImageJ (National Institutes of Health, Bethesda, MD, USA).

\hypertarget{acknowledgements}{%
\subsection{Acknowledgements}\label{acknowledgements}}

This work was funded by the Wellcome Trust (214337/Z/18/Z).

\hypertarget{references}{%
\subsection*{References}\label{references}}
\addcontentsline{toc}{subsection}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Achim2015}{}}%
Achim K, Pettit J-B, Saraiva LR, Gavriouchkina D, Larsson T, Arendt D,
Marioni JC. 2015. High-throughput spatial mapping of single-cell RNA-seq
data to tissue of origin. \emph{Nature Biotechnology}
\textbf{33}:503--509.
doi:\href{https://doi.org/10.1038/nbt.3209}{10.1038/nbt.3209}

\leavevmode\vadjust pre{\hypertarget{ref-Arendt2004}{}}%
Arendt D, Tessmar-Raible K, Snyman H, Dorresteijn AW, Wittbrodt J. 2004.
Ciliary Photoreceptors with a Vertebrate-Type Opsin in an Invertebrate
Brain. \emph{Science} \textbf{306}:869--871.
doi:\href{https://doi.org/10.1126/science.1099955}{10.1126/science.1099955}

\leavevmode\vadjust pre{\hypertarget{ref-bishop2007}{}}%
Bishop CD, Brandhorst BP. 2007. Development of nitric oxide
synthase-defined neurons in the sea urchin larval ciliary band and
evidence for a chemosensory function during metamorphosis.
\emph{Developmental Dynamics} \textbf{236}:1535--1546.
doi:\href{https://doi.org/10.1002/dvdy.21161}{10.1002/dvdy.21161}

\leavevmode\vadjust pre{\hypertarget{ref-bishop2003}{}}%
Bishop CD, Brandhorst BP. 2003. On nitric oxide signaling,
metamorphosis, and the evolution of biphasic life cycles.
\emph{Evolution and Development} \textbf{5}:542--550.
doi:\href{https://doi.org/10.1046/j.1525-142x.2003.03059.x}{10.1046/j.1525-142x.2003.03059.x}

\leavevmode\vadjust pre{\hypertarget{ref-burette2002}{}}%
Burette A, Zabel U, Weinberg RJ, Schmidt HHHW, Valtschanoff JG. 2002.
Synaptic Localization of Nitric Oxide Synthase and Soluble Guanylyl
Cyclase in the Hippocampus. \emph{The Journal of Neuroscience}
\textbf{22}:8961--8970.
doi:\href{https://doi.org/10.1523/jneurosci.22-20-08961.2002}{10.1523/jneurosci.22-20-08961.2002}

\leavevmode\vadjust pre{\hypertarget{ref-camargo2007}{}}%
Camargo A, Llamas A, Schnell RA, Higuera JoseJ́, Gonzaĺez-Ballester D,
Lefebvre PA, Fernańdez E, Galvań A. 2007. Nitrate Signaling by the
Regulatory Gene{\emph{NIT2}}in{\emph{Chlamydomonas}}. \emph{The Plant
Cell} \textbf{19}:3491--3503.
doi:\href{https://doi.org/10.1105/tpc.106.045922}{10.1105/tpc.106.045922}

\leavevmode\vadjust pre{\hypertarget{ref-Capella-Gutiuxe9rrez2009}{}}%
Capella-Gutierrez S, Silla-Martinez JM, Gabaldon T. 2009. trimAl: a tool
for automated alignment trimming in large-scale phylogenetic analyses.
\emph{Bioinformatics} \textbf{25}:1972--1973.
doi:\href{https://doi.org/10.1093/bioinformatics/btp348}{10.1093/bioinformatics/btp348}

\leavevmode\vadjust pre{\hypertarget{ref-castellano2014}{}}%
Castellano I, Ercolesi E, Palumbo A. 2014. Nitric Oxide Affects ERK
Signaling through Down-Regulation of MAP Kinase Phosphatase Levels
during Larval Development of the Ascidian Ciona intestinalis. \emph{PLoS
ONE} \textbf{9}:e102907.
doi:\href{https://doi.org/10.1371/journal.pone.0102907}{10.1371/journal.pone.0102907}

\leavevmode\vadjust pre{\hypertarget{ref-Choi2018}{}}%
Choi HMT, Schwarzkopf M, Fornace ME, Acharya A, Artavanis G, Stegmaier
J, Cunha A, Pierce NA. 2018. Third-generation {\emph{in situ}}
hybridization chain reaction: multiplexed, quantitative, sensitive,
versatile, robust. \emph{Development} \textbf{145}.
doi:\href{https://doi.org/10.1242/dev.165753}{10.1242/dev.165753}

\leavevmode\vadjust pre{\hypertarget{ref-Conzelmann2012}{}}%
Conzelmann M, Jékely G. 2012. Antibodies against conserved amidated
neuropeptide epitopes enrich the comparative neurobiology toolbox.
\emph{EvoDevo} \textbf{3}:23.
doi:\href{https://doi.org/10.1186/2041-9139-3-23}{10.1186/2041-9139-3-23}

\leavevmode\vadjust pre{\hypertarget{ref-Conzelmann2011}{}}%
Conzelmann M, Offenburger S-L, Asadulina A, Keller T, Münch TA, Jékely
G. 2011. Neuropeptides regulate swimming depth of {\emph{Platynereis}}
larvae. \emph{Proceedings of the National Academy of Sciences}
\textbf{108}.
doi:\href{https://doi.org/10.1073/pnas.1109085108}{10.1073/pnas.1109085108}

\leavevmode\vadjust pre{\hypertarget{ref-Conzelmann2013a}{}}%
Conzelmann M, Williams Elizabeth A, Krug K, Franz-Wachtel M, Macek B,
Jékely G. 2013a. The neuropeptide complement of the marine annelid
Platynereis dumerilii. \emph{BMC Genomics} \textbf{14}:906.
doi:\href{https://doi.org/10.1186/1471-2164-14-906}{10.1186/1471-2164-14-906}

\leavevmode\vadjust pre{\hypertarget{ref-Conzelmann2013}{}}%
Conzelmann M, Williams Elizabeth A, Tunaru S, Randel N, Shahidi R,
Asadulina A, Berger J, Offermanns S, Jékely G. 2013b. Conserved MIP
receptor{\textendash}ligand pair regulates {\emph{Platynereis}} larval
settlement. \emph{Proceedings of the National Academy of Sciences}
\textbf{110}:8224--8229.
doi:\href{https://doi.org/10.1073/pnas.1220285110}{10.1073/pnas.1220285110}

\leavevmode\vadjust pre{\hypertarget{ref-Cudeiro1999}{}}%
Cudeiro J, Rivadulla C. 1999. Sight and insight {\textendash} on the
physiological role of nitric oxide in the visual system. \emph{Trends in
Neurosciences} \textbf{22}:109--116.
doi:\href{https://doi.org/10.1016/s0166-2236(98)01299-5}{10.1016/s0166-2236(98)01299-5}

\leavevmode\vadjust pre{\hypertarget{ref-Fu2012}{}}%
Fu L, Niu B, Zhu Z, Wu S, Li W. 2012. CD-HIT: accelerated for clustering
the next-generation sequencing data. \emph{Bioinformatics}
\textbf{28}:3150--3152.
doi:\href{https://doi.org/10.1093/bioinformatics/bts565}{10.1093/bioinformatics/bts565}

\leavevmode\vadjust pre{\hypertarget{ref-garthwaite2015a}{}}%
Garthwaite J. 2015. From synaptically localized to volume transmission
by nitric oxide. \emph{The Journal of Physiology} \textbf{594}:9--18.
doi:\href{https://doi.org/10.1113/jp270297}{10.1113/jp270297}

\leavevmode\vadjust pre{\hypertarget{ref-jacoby2018}{}}%
Jacoby J, Nath A, Jessen ZF, Schwartz GW. 2018. A Self-Regulating Gap
Junction Network of Amacrine Cells Controls Nitric Oxide Release in the
Retina. \emph{Neuron} \textbf{100}:1149--1162.e5.
doi:\href{https://doi.org/10.1016/j.neuron.2018.09.047}{10.1016/j.neuron.2018.09.047}

\leavevmode\vadjust pre{\hypertarget{ref-Juxe9kely2021}{}}%
Jékely G. 2021. The chemical brain hypothesis for the origin of nervous
systems. \emph{Philosophical Transactions of the Royal Society B:
Biological Sciences} \textbf{376}:20190761.
doi:\href{https://doi.org/10.1098/rstb.2019.0761}{10.1098/rstb.2019.0761}

\leavevmode\vadjust pre{\hypertarget{ref-Kuehn2022}{}}%
Kuehn E, Clausen DS, Null RW, Metzger BM, Willis AD, Özpolat BD. 2021.
Segment number threshold determines juvenile onset of germline cluster
expansion in {\emph{Platynereis dumerilii}}. \emph{Journal of
Experimental Zoology Part B: Molecular and Developmental Evolution}
\textbf{338}:225--240.
doi:\href{https://doi.org/10.1002/jez.b.23100}{10.1002/jez.b.23100}

\leavevmode\vadjust pre{\hypertarget{ref-leise2001}{}}%
Leise EM, Thavaradhara K, Durham NR, Turner BE. 2001. Serotonin and
Nitric Oxide Regulate Metamorphosis in the Marine Snail{\emph{Ilyanassa
obsoleta}}. \emph{American Zoologist} \textbf{41}:258--267.
doi:\href{https://doi.org/10.1093/icb/41.2.258}{10.1093/icb/41.2.258}

\leavevmode\vadjust pre{\hypertarget{ref-lundberg2011}{}}%
Lundberg JO, Weitzberg E, Shiva S, Gladwin MT. 2011. The
nitrate{\textendash}nitrite{\textendash}nitric oxide pathway in mammals.
Humana Press. pp. 21--48.
doi:\href{https://doi.org/10.1007/978-1-60761-616-0_3}{10.1007/978-1-60761-616-0\_3}

\leavevmode\vadjust pre{\hypertarget{ref-Matsuda2017}{}}%
Matsuda S, Harada K, Ito M, Takizawa M, Wongso D, Tsuboi T, Kitaguchi T.
2016. Generation of a cGMP Indicator with an Expanded Dynamic Range by
Optimization of Amino Acid Linkers between a Fluorescent Protein and
PDE5? \emph{ACS Sensors} \textbf{2}:46--51.
doi:\href{https://doi.org/10.1021/acssensors.6b00582}{10.1021/acssensors.6b00582}

\leavevmode\vadjust pre{\hypertarget{ref-Minh2020}{}}%
Minh BQ, Schmidt HA, Chernomor O, Schrempf D, Woodhams MD, Haeseler A
von, Lanfear R. 2020. IQ-TREE 2: New Models and Efficient Methods for
Phylogenetic Inference in the Genomic Era. \emph{Molecular Biology and
Evolution} \textbf{37}:1530--1534.
doi:\href{https://doi.org/10.1093/molbev/msaa015}{10.1093/molbev/msaa015}

\leavevmode\vadjust pre{\hypertarget{ref-muxf6ller2019}{}}%
Möller MN, Rios N, Trujillo M, Radi R, Denicola A, Alvarez B. 2019.
Detection and quantification of nitric oxide{\textendash}derived
oxidants in biological systems. \emph{Journal of Biological Chemistry}
\textbf{294}:14776--14802.
doi:\href{https://doi.org/10.1074/jbc.rev119.006136}{10.1074/jbc.rev119.006136}

\leavevmode\vadjust pre{\hypertarget{ref-Moroz2004}{}}%
Moroz LL, Meech RW, Sweedler JV, Mackie GO. 2004. Nitric oxide regulates
swimming in the jellyfishAglantha digitale. \emph{The Journal of
Comparative Neurology} \textbf{471}:26--36.
doi:\href{https://doi.org/10.1002/cne.20023}{10.1002/cne.20023}

\leavevmode\vadjust pre{\hypertarget{ref-Moroz2020}{}}%
Moroz LL, Romanova DY, Nikitin MA, Sohn D, Kohn AB, Neveu E, Varoqueaux
F, Fasshauer D. 2020. The diversification and lineage-specific expansion
of nitric oxide signaling in Placozoa: insights in the evolution of
gaseous transmission. \emph{Scientific Reports} \textbf{10}.
doi:\href{https://doi.org/10.1038/s41598-020-69851-w}{10.1038/s41598-020-69851-w}

\leavevmode\vadjust pre{\hypertarget{ref-ozpolat2021}{}}%
Ozpolat BD, Randel N, Williams EA, Bezares-Calderón LA, Andreatta G,
Balavoine G, Bertucci PY, Ferrier DEK, Gambi MC, Gazave E,
Handberg-Thorsager M, Hardege J, Hird C, Hsieh Y-W, Hui J, Mutemi KN,
Schneider SQ, Simakov O, Vergara HM, Vervoort M, Jékely G,
Tessmar-Raible K, Raible F, Arendt D. 2021. The Nereid on the rise:
Platynereis as a model system. \emph{Zenodo}.
doi:\href{https://doi.org/10.5281/ZENODO.4907400}{10.5281/ZENODO.4907400}

\leavevmode\vadjust pre{\hypertarget{ref-pechenik2007}{}}%
Pechenik JA, Cochrane DE, Li W, West ET, Pires A, Leppo M. 2007. Nitric
Oxide Inhibits Metamorphosis in Larvae of{\emph{Crepidula fornicata}},
the Slippershell Snail. \emph{The Biological Bulletin}
\textbf{213}:160--171.
doi:\href{https://doi.org/10.2307/25066632}{10.2307/25066632}

\leavevmode\vadjust pre{\hypertarget{ref-puopolo2001}{}}%
Puopolo M, Hochstetler SE, Gustincich S, Wightman RM, Raviola E. 2001.
Extrasynaptic Release of Dopamine in a Retinal Neuron. \emph{Neuron}
\textbf{30}:211--225.
doi:\href{https://doi.org/10.1016/s0896-6273(01)00274-4}{10.1016/s0896-6273(01)00274-4}

\leavevmode\vadjust pre{\hypertarget{ref-Randel2014a}{}}%
Randel N, Asadulina A, Bezares-Calderón LA, Verasztó C, Williams EA,
Conzelmann M, Shahidi R, Jékely G. 2014. Neuronal connectome of a
sensory-motor circuit for visual navigation. \emph{eLife} \textbf{3}.
doi:\href{https://doi.org/10.7554/elife.02730}{10.7554/elife.02730}

\leavevmode\vadjust pre{\hypertarget{ref-santos2011}{}}%
Santos RM, Lourenço CF, Pomerleau F, Huettl P, Gerhardt GA, Laranjinha
J, Barbosa RM. 2011. Brain Nitric Oxide Inactivation Is Governed by the
Vasculature. \emph{Antioxidants \& Redox Signaling}
\textbf{14}:1011--1021.
doi:\href{https://doi.org/10.1089/ars.2010.3297}{10.1089/ars.2010.3297}

\leavevmode\vadjust pre{\hypertarget{ref-Shu2003}{}}%
Shu CJ, Ulrich LE, Zhulin IB. 2003. The NIT domain: a predicted
nitrate-responsive module in bacterial sensory receptors. \emph{Trends
in Biochemical Sciences} \textbf{28}:121--124.
doi:\href{https://doi.org/10.1016/s0968-0004(03)00032-x}{10.1016/s0968-0004(03)00032-x}

\leavevmode\vadjust pre{\hypertarget{ref-song2021}{}}%
Song H, Hewitt OH, Degnan SM. 2021. Arginine Biosynthesis by a Bacterial
Symbiont Enables Nitric Oxide Production and Facilitates Larval
Settlement in the Marine-Sponge Host. \emph{Current Biology}
\textbf{31}:433--437.e3.
doi:\href{https://doi.org/10.1016/j.cub.2020.10.051}{10.1016/j.cub.2020.10.051}

\leavevmode\vadjust pre{\hypertarget{ref-Tessmar-Raible2005}{}}%
Tessmar-Raible K, Steinmetz PRH, Snyman H, Hassel M, Arendt D. 2005.
Fluorescent two-color whole mount in situ hybridization in
{\emph{Platynereis dumerilii}} (Polychaeta, Annelida), an emerging
marine molecular model for evolution and development.
\emph{BioTechniques} \textbf{39}:460--464.
doi:\href{https://doi.org/10.2144/000112023}{10.2144/000112023}

\leavevmode\vadjust pre{\hypertarget{ref-Tosches2014}{}}%
Tosches M, Bucher D, Vopalensky P, Arendt D. 2014. Melatonin Signaling
Controls Circadian Swimming Behavior in Marine Zooplankton. \emph{Cell}
\textbf{159}:46--57.
doi:\href{https://doi.org/10.1016/j.cell.2014.07.042}{10.1016/j.cell.2014.07.042}

\leavevmode\vadjust pre{\hypertarget{ref-tsukamoto2017}{}}%
Tsukamoto H, Chen I-S, Kubo Y, Furutani Y. 2017. A ciliary opsin in the
brain of a marine annelid zooplankton is ultraviolet-sensitive, and the
sensitivity is tuned by a single amino acid residue. \emph{Journal of
Biological Chemistry} \textbf{292}:12971--12980.
doi:\href{https://doi.org/10.1074/jbc.m117.793539}{10.1074/jbc.m117.793539}

\leavevmode\vadjust pre{\hypertarget{ref-ueda2014}{}}%
Ueda N, Degnan SM. 2014. Nitric oxide is not a negative regulator of
metamorphic induction in the abalone haliotis asinina. \emph{Frontiers
in Marine Science} \textbf{1}.
doi:\href{https://doi.org/10.3389/fmars.2014.00021}{10.3389/fmars.2014.00021}

\leavevmode\vadjust pre{\hypertarget{ref-ueda2013}{}}%
Ueda N, Degnan SM. 2013. Nitric Oxide Acts as a Positive Regulator to
Induce Metamorphosis of the Ascidian Herdmania momus. \emph{PLoS ONE}
\textbf{8}:e72797.
doi:\href{https://doi.org/10.1371/journal.pone.0072797}{10.1371/journal.pone.0072797}

\leavevmode\vadjust pre{\hypertarget{ref-ueda2016}{}}%
Ueda N, Richards GS, Degnan BM, Kranz A, Adamska M, Croll RP, Degnan SM.
2016. An ancient role for nitric oxide in regulating the animal
pelagobenthic life cycle: evidence from a marine sponge.
\emph{Scientific Reports} \textbf{6}.
doi:\href{https://doi.org/10.1038/srep37546}{10.1038/srep37546}

\leavevmode\vadjust pre{\hypertarget{ref-Veedin2021}{}}%
Veedin Rajan VB, Häfker NS, Arboleda E, Poehn B, Gossenreiter T, Gerrard
E, Hofbauer M, Mühlestein C, Bileck A, Gerner C, Ribera d'Alcala M, Buia
MC, Hartl M, Lucas RJ, Tessmar-Raible K. 2021. Seasonal variation in UVA
light drives hormonal and behavioural changes in a marine annelid via a
ciliary opsin. \emph{Nature Ecology \& Evolution} \textbf{5}:204--218.
doi:\href{https://doi.org/10.1038/s41559-020-01356-1}{10.1038/s41559-020-01356-1}

\leavevmode\vadjust pre{\hypertarget{ref-Veraszto2018}{}}%
Verasztó C, Gühmann M, Jia H, Rajan VBV, Bezares-Calderón LA,
Piñeiro-Lopez C, Randel N, Shahidi R, Michiels NK, Yokoyama S,
Tessmar-Raible K, Jékely G. 2018. Ciliary and rhabdomeric
photoreceptor-cell circuits form a spectral depth gauge in marine
zooplankton. \emph{eLife} \textbf{7}.
doi:\href{https://doi.org/10.7554/elife.36440}{10.7554/elife.36440}

\leavevmode\vadjust pre{\hypertarget{ref-veraszto2020}{}}%
Verasztó C, Jasek S, Gühmann M, Shahidi R, Ueda N, Beard JD, Mendes S,
Heinz K, Bezares-Calderón LA, Williams E, Jékely G. 2020.
\href{http://dx.doi.org/10.1101/2020.08.21.260984}{Whole-animal
connectome and cell-type complement of the three-segmented
{\emph{Platynereis dumerilii}} larva}.

\leavevmode\vadjust pre{\hypertarget{ref-Veraszto2017}{}}%
Verasztó C, Ueda N, Bezares-Calderón LA, Panzera A, Williams EA, Shahidi
R, Jékely G. 2017. Ciliomotor circuitry underlying whole-body
coordination of ciliary activity in the Platynereis larva. \emph{eLife}
\textbf{6}.
doi:\href{https://doi.org/10.7554/elife.26000}{10.7554/elife.26000}

\leavevmode\vadjust pre{\hypertarget{ref-Vielma2014}{}}%
Vielma AH, Agurto A, Valdés J, Palacios AG, Schmachtenberg O. 2014.
Nitric Oxide Modulates the Temporal Properties of the Glutamate Response
in Type 4 OFF Bipolar Cells. \emph{PLoS ONE} \textbf{9}:e114330.
doi:\href{https://doi.org/10.1371/journal.pone.0114330}{10.1371/journal.pone.0114330}

\leavevmode\vadjust pre{\hypertarget{ref-Wei2012}{}}%
Wei T, Schubert T, Paquet-Durand F, Tanimoto N, Chang L, Koeppen K, Ott
T, Griesbeck O, Seeliger MW, Euler T, Wissinger B. 2012. Light-Driven
Calcium Signals in Mouse Cone Photoreceptors. \emph{Journal of
Neuroscience} \textbf{32}:6981--6994.
doi:\href{https://doi.org/10.1523/jneurosci.6432-11.2012}{10.1523/jneurosci.6432-11.2012}

\leavevmode\vadjust pre{\hypertarget{ref-Williams2015}{}}%
Williams EA, Conzelmann M, Jékely G. 2015. Myoinhibitory peptide
regulates feeding in the marine annelid Platynereis. \emph{Frontiers in
Zoology} \textbf{12}.
doi:\href{https://doi.org/10.1186/s12983-014-0093-6}{10.1186/s12983-014-0093-6}

\leavevmode\vadjust pre{\hypertarget{ref-williams2017}{}}%
Williams EA, Verasztó C, Jasek S, Conzelmann M, Shahidi R, Bauknecht P,
Mirabeau O, Jékely G. 2017. Synaptic and peptidergic connectome of a
neurosecretory center in the annelid brain. \emph{eLife} \textbf{6}.
doi:\href{https://doi.org/10.7554/elife.26349}{10.7554/elife.26349}

\leavevmode\vadjust pre{\hypertarget{ref-yang2018}{}}%
Yang X-X, Wong YH, Zhang Y, Zhang G, Qian P-Y. 2018. Exploring the
regulatory role of nitric oxide (NO) and the NO-p38MAPK/cGMP pathway in
larval settlement of the bryozoan {\emph{Bugula neritina}}.
\emph{Biofouling} \textbf{34}:545--556.
doi:\href{https://doi.org/10.1080/08927014.2018.1470240}{10.1080/08927014.2018.1470240}

\leavevmode\vadjust pre{\hypertarget{ref-zhang2007}{}}%
Zhang D-Q, Zhou T-R, McMahon DG. 2007. Functional Heterogeneity of
Retinal Dopaminergic Neurons Underlying Their Multiple Roles in Vision.
\emph{Journal of Neuroscience} \textbf{27}:692--699.
doi:\href{https://doi.org/10.1523/jneurosci.4478-06.2007}{10.1523/jneurosci.4478-06.2007}

\leavevmode\vadjust pre{\hypertarget{ref-zhu2020}{}}%
Zhu Y-T, Zhang Y, Liu Y-Z, Li Y-F, Yoshida A, Osatomi K, Yang J-L, Liang
X. 2020. Nitric oxide negatively regulates larval metamorphosis in
hard-shelled mussel (mytilus coruscus). \emph{Frontiers in Marine
Science} \textbf{7}.
doi:\href{https://doi.org/10.3389/fmars.2020.00356}{10.3389/fmars.2020.00356}

\end{CSLReferences}

\end{document}

### 2023-05-06 Manuscript/Supplements.Rmd
upload ms in Manuscript/

--- 2023-01-10 text/Supplements.Rmd
@@
 
-![**Figure 1---figure supplement 2.** Expression analysis of the NOS gene (magenta) using in situ HCR using the larva at 3-day-old in dorsal view (A) and anterior view (B).
-(B) Insert: showing two NOS cells on each side.](../figures/Fig1_sup2.png){width="100%"}
+![**Figure 1---figure supplement 2.** Expression analysis of the *NOS* gene (magenta) using in situ HCR using the larva at three-day-old in dorsal (A) and anterior view (B).
+(B) Insert: showing two NOS cells (INNOS_dl and INNOS_vl) on left side.](../figures/Fig1_sup2.png){width="100%"}
 
-![**Figure 3---figure supplement 1.** **(A)** Overlaid trajectories for WT (left, n=20) and NOS mutant (right, n=31) 2-day-old larvae.
-Grey shows the trajectories before UV stimulation and purple shows the trajectories after UV stimulation.
+![**Figure 3---figure supplement 1.** **(A)** Overlaid trajectories for WT (n=37) and NOS mutant (NOSΔ11/Δ11, n=18 and NOSΔ23/Δ23, n=8) at two-day-old larvae.
+0 sec as the starting point.
+After 10 sec, UV (395 nm) stimulation from the side.
 **(B)** The temporal changes in the vertical position of the WT and mutant 2-days-old larvae before and after UV stimulation are shown.
@@
 ![**Figure 4---figure supplement 2.** **(A)** Co-expression analysis image of the NIT-GC1 (magenta) and MLD-pedal2 amide proneuropeptide gene (MLD: green).
-Anterior view of the larva at 2-day-old.
+Anterior view of the larva at two-day-old.
 **(B, C)** Expression analysis of the NIT-GC2 gene (magenta) using in situ HCR.
-Anterior (B) and posterior (C) views of the larva at 3-day-old.
-**(D, E)** Localisation analysis for NIT-GC1 (B) and NIT-GC2 (C) morphant using NIT-GC1 and NIT-GC2 antibodies.
+Anterior (B) and posterior (C) views of the larva at three-day-old.
+**(D)** Co-localisation analysis using NIT-GC1 (magenta) and NOS (green) antibodies.
+Anterior view of the larva at two-day-old.
+**(E, F)** Localisation analysis using NIT-GC1 and NIT-GC2 antibodies for NIT-GC1 (E) and NIT-GC2 (F) morphant.
 Green shows co-staining with acetylated α-tubulin antibody (acTub).
-Anterior view of the larva at 2-day-old.
-**(F)** Co-localisation analysis using NIT-GC1 (magenta) and NOS (green) antibodies.
-Anterior view of the larva at 2-day-old.](../figures/Fig4_sup2.png){width="100%"}
+Anterior view of the larva at two-day-old.](../figures/Fig4_sup2.png){width="100%"}
+
+![**Figure 5---figure supplement 1.** Schematic diagram of the immunostaining procedure after calcium imaging.](../figures/Fig5_sup1.png){width="100%"}

### 2023-05-06 Manuscript/Manuscript.Rmd
MS: updated affiliations, changed to knitr to include figures

--- 2023-05-05 text/Manuscript.Rmd
@@
   in zooplankton
-author: "*Kei Jokura, Nobuo Ueda, Martin Gühmann, Luis Alfonso Yañez-Guerra, Piotr Słowiński, Kyle C. A. Wedgwood, Gáspár Jékely*"
-date: "*Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4
-  4QD, UK*"
+author: 
+  - name: "Kei Jokura<sup>1,`*`</sup>, Nobuo Ueda<sup>2</sup>, Martin Gühmann<sup>3</sup>, Luis Alfonso Yañez-Guerra<sup>1</sup>, Piotr Słowiński<sup>1</sup>, Kyle C.
+A.
+Wedgewood<sup>1</sup> and Gáspár Jékely<sup>1,4,`*`</sup>"
+    affiliation: "<sup>1</sup>Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4 <br> <sup>2</sup>Okinawa Institute of Science and Technology, Okinawa, Japan<br> <sup>3</sup>School of Biological Sciences, University of Bristol, Bristol, United Kingdom <br> <sup>4</sup>Centre for Organismal Studies (COS), University of Heidelberg, 69120 Heidelberg, Germany  <br> <sup>*</sup>Correspondence`:` kei.jokura @ exeter.ac.uk, gaspar.jekely @ cos.uni-heidelberg.de"
 output:
+  html_document:
+    theme: lumen
   pdf_document:
@@
 Numbers inside hexagons indicate the number of cells grouped in each node.
-Arrow line thickness (synapse weight) is equal to the number of synapses.](../figures/Fig1.png){width="100%"}
+Arrow line thickness (synapse weight) is equal to the number of synapses."}
+
+knitr::include_graphics(path = "figures/Fig1.png")
+
+```
 
@@
 
-![**Figure 2.
+```{r echo=FALSE,fig.cap="**Figure 2.
 NO produced by UV/violet stimulation to cPRCs.** **(A)** A NO-positive cell visualised by DAF-FM (green) is shown (dashline).
@@
 **(C)** The intensity of DAF-FM changes over time when cPRC (green) and other cells (ctr: gray) are stimulated with UV/violet (405 nm) light.
-Purple squares indicate the timing of 405 nm light stimulation, the transparent lines show the optical responses to individual intensities, each intensities are normalized (ΔF/F0) and the dark trace lines and the shaded bands show the mean value and the standard error range of the mean.](../figures/Fig2.png){width="100%"}
+Purple squares indicate the timing of 405 nm light stimulation, the transparent lines show the optical responses to individual intensities, each intensities are normalized (ΔF/F0) and the dark trace lines and the shaded bands show the mean value and the standard error range of the mean."}
+
+knitr::include_graphics(path = "figures/Fig2.png")
+
+```
 
@@
 
-![**Figure 3.
+```{r echo=FALSE,fig.cap="**Figure 3.
 Larvae with inhibited NOS have a delayed response to UV-avoidance.** **(A)** Top: The domain and Exon/Intron structure of NOS.
@@
 **(E)** Vertical swimming in larvae treated with NOS inhibitors, L-NAME at three-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
-The data are shown in 30 s bins.](../figures/Fig3.png){width="100%"}
+The data are shown in 30 s bins."}
+
+knitr::include_graphics(path = "figures/Fig3.png")
+
+```
 
@@
 
-![**Figure 4.
+```{r echo=FALSE,fig.cap="**Figure 4.
 cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.** **(A, B)** Expression analysis of the NIT-GC1 (C) and NIT-GC2 (D) gene (magenta) using in situ HCR.
@@
 **(K-N)** The intensity changes in cGMP fluorescence over time for the four conditions are shown.
-Gray traces show the optical responses to individual intensities (n \> 6).
+Gray traces show the optical responses to individual intensities (n > 6).
 The individual intensities are normalized (ΔF/F0).
 The dark line shows the mean value and the band the standard error range of the mean.
-The respective solution was acted on 2 min after the start of imaging (grey bars).](../figures/Fig4.png){width="100%"}
+The respective solution was acted on 2 min after the start of imaging (grey bars)."}
+
+knitr::include_graphics(path = "figures/Fig4.png")
+
+```
 
@@
 
-![**Figure 5.
+```{r echo=FALSE,fig.cap="**Figure 5.
 NO produced by UV stimulation to cPRCs induces a depolarising response in cPRC via NIT-GC1.** **(A, B)** Each left figure shows a calcium imaging using GCaMP6s.
@@
 **(G)** Changes over time in GCaMP6s intensities in INRGW when cPRC is stimulated with UV in NIT-GC2 morphants (MO1 and MO2: blue and light blue).
-**(H, I)** Changes over time in GCaMP6s intensities in Ser-h1 and MC when cPRC is stimulated with UV in WT (Ser-h1 and MC: dark red and brown) and NOSΔ11/Δ11 mutant (Ser-h1 and MC: light red and brown) larvae.](../figures/Fig5.png){width="100%"}
+**(H, I)** Changes over time in GCaMP6s intensities in Ser-h1 and MC when cPRC is stimulated with UV in WT (Ser-h1 and MC: dark red and brown) and NOSΔ11/Δ11 mutant (Ser-h1 and MC: light red and brown) larvae."}
+
+knitr::include_graphics(path = "figures/Fig5.png")
+
+```
 
@@
 
-![**Figure 6.
+```{r echo=FALSE,fig.cap="**Figure 6.
 Gene expressions and signalling mechanisms between cPRC, INNOS and INRGWa.** **(C)** Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq.
@@
 The colours represent the normal logarithm of the number of transcripts in the cells expressing the gene.
-**(D)** Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback.](../figures/Fig6.png){width="100%"}
+**(D)** Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback."}
+
+knitr::include_graphics(path = "figures/Fig6.png")
+
+```
 

### 2023-05-06 Manuscript/Manuscript.Rmd
MS: added fig supplements

@@
 Wedgewood<sup>1</sup> and Gáspár Jékely<sup>1,4,`*`</sup>"
-    affiliation: "<sup>1</sup>Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4 <br> <sup>2</sup>Okinawa Institute of Science and Technology, Okinawa, Japan<br> <sup>3</sup>School of Biological Sciences, University of Bristol, Bristol, United Kingdom <br> <sup>4</sup>Centre for Organismal Studies (COS), University of Heidelberg, 69120 Heidelberg, Germany  <br> <sup>*</sup>Correspondence`:` kei.jokura @ exeter.ac.uk, gaspar.jekely @ cos.uni-heidelberg.de"
+    affiliation: "<sup>1</sup>Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4 4QD, United Kingdom <sup>2</sup>Okinawa Institute of Science and Technology, Okinawa, Japan, <sup>3</sup>School of Biological Sciences, University of Bristol, Bristol, United Kingdom, <sup>4</sup>Centre for Organismal Studies (COS), University of Heidelberg, 69120 Heidelberg, Germany  <br> <sup>*</sup>Correspondence`:` kei.jokura @ exeter.ac.uk, gaspar.jekely @ cos.uni-heidelberg.de"
 output:
@@
 :::
+
+## Figure supplements
+
+```{r echo=FALSE,fig.cap="**Figure 1 -- figure supplement 1** "}
+knitr::include_graphics(path = "figures/Fig1_sup1.png")
+
+```
+
+```{r echo=FALSE,fig.cap="**Figure 1 -- figure supplement 2** "}
+knitr::include_graphics(path = "figures/Fig1_sup2.png")
+
+```
+
+```{r echo=FALSE,fig.cap="**Figure 3 -- figure supplement 1** "}
+knitr::include_graphics(path = "figures/Fig3_sup1.png")
+
+```
+
+```{r echo=FALSE,fig.cap="**Figure 4 -- figure supplement 1** "}
+knitr::include_graphics(path = "figures/Fig4_sup1.png")
+
+```
+
+```{r echo=FALSE,fig.cap="**Figure 4 -- figure supplement 2** "}
+knitr::include_graphics(path = "figures/Fig4_sup2.png")
+
+```
+
+```{r echo=FALSE,fig.cap="**Figure 5 -- figure supplement 1** "}
+knitr::include_graphics(path = "figures/Fig5_sup1.png")
+
+```

### 2023-05-06 Manuscript/Manuscript.Rmd
MS: edits to text, reordering NIT-GC sections - tc

@@
 
-The ability to perceive changes in the environment and adapt behaviour accordingly is essential for the survival of all living organisms.
-Avoidance from UV radiation, which damages DNA essential for survival, is proposed to be one of the behaviours that existed before the abundant development of Earth's oxygen to block UV radiation in the atmosphere (Pittendrigh).
-In the ocean, almost all zooplankton have a pronounced rhythmic behaviour called Diel vertical migration (DVM), which consists of upwards towards the surface at dusk and descend back to deeper water before dawn.
-Avoidance behaviour to escape harmful UV radiation during the day is closely linked to the control of vertical migration in the ocean.
-However, little is known about the molecular neural circuit mechanisms in UV avoidance behaviour in many animals, including zooplankton.
-In neural circuits in general light response, the activation of various receptors by excitatory and inhibitory neurotransmitters such as glutamate, GABA and dopamine is known in various species.
-In addition to this, it has long been known that in nitric oxide (NO), as a neuromodulatory molecule, it is also important as a modulator of visual activity.
-NO is a small free-radical molecule with diffusible and changeable characteristics, and thus works as one of the neural signalling molecules.
-The function of NO in the neural circuitry of vision has been well studied in vertebrates.
-In the mammalian retina, NO is an essential signalling molecule in all retinal cells.
-NO produced by nitric oxide synthase (NOS) in amacrine cells, one of the retinal cells, activates *soluble guanylate cyclase (sGC)* expressed in cone bipolar cells, which produces cGMP and regulates downstream signalling.
-We therefore hypothesise that NO-mediated neural signalling may be involved in the photoreceptor circuits that control UV avoidance behaviour.
-To test this hypothesis we used the annelid *Platynereis dumerilii* **(Figure 1A)**, which conserved vertebrate-type ciliary photoreceptors (cPRCs) [@Arendt2004].
-The cPRC expresses *c-opsin1*, which absorbs UV radiation, is involved in melatonin production and regulates circadian swimming behaviour [@Tosches2014; @tsukamoto2017].
-Furthermore, we previously reported abnormal UV avoidance behaviour in *c-opsin1* knockout larvae [@Veraszto2018].
-The cPRCs express the UV-absorbing c-opsin1 and mediate UV avoidance in the larvae [@Veraszto2018].
-In adult worms, UV light and c-opsin1 modulate locomotion and maturation [@Veedin2021].
-These suggest that Platynereis larvae are a highly suitable model for elucidating the neural circuit mechanisms of NO-induced UV avoidance behaviour, and the accessibility of genome editing, behavioural assays, calcium imaging and whole-body connectomics could be very useful for study at the molecular level[@ozpolat2021].
+Neuronally produced nitric oxide (NO) can act as a diffusible neuromodulator. This small free-radical molecule is synthesized by nitric oxide synthase (NOS)...
+
+In the mammalian retina, NO is produced by amacrine cells and activates a *soluble guanylate cyclase (sGC)* in cone bipolar cells.
+The NO-dependent increased in cGMP tunes bipolar cell activity ....
+[refs]
+[other examples]
+
+To investigate NO signalling in the modulation of neural circuit dynamics, we studied the larval stages of the marine annelid model *Platynereis dumerilii* [@ozpolat2021].
+We uncovered an essential function for NO in these larvae in UV/violet-light avoidance.
+In *Platynereis* larvae, UV/violet avoidance is characterised by downward swimming and is mediated by the UV/violet-responding brain ciliary photoreceptor cells (cPRCs).
+
+The cPRCs express a vertebrate-type ciliary opsin, c-opsin1 [@Arendt2004].
+This opsin forms a UV-absorbing bistable photopigment with an absorption maximum around 384 nm [@tsukamoto2017, @Veraszto2018].
+*c-opsin1* knockout larvae show defective UV/violet avoidance [@Veraszto2018].
+Here we show that *Platynereis* *NOS* is expressed in interneurons of the cPRC circuitry and is required for UV/violet-avoidance behaviour.
+By combining calcium imaging across the fully-mapped cPRC circuit, genetic perturbations and mathematical modelling, we describe how NO tunes circuit dynamics through non-synaptic reverse signalling to cPRCs.
+This delayed neuroendocrine feedback is essential for circuit output, and integrates UV/violet input and also induces a short-term memory manifest in altered circuit activity.
+
+#In adult worms, UV light and c-opsin1 modulate locomotion and maturation [@Veedin2021].
 
@@
 In two- and three-day-old larvae, we detected *NOS* expression in four cells (two of them weakly expressing) in the apical organ region **(Figure 1D and Figure 1---figure supplement 2)**.
-The same four cells were labelled with a reporter construct driving palmitoylated-tdTomato by a 12 kbp upstream genomic region of the *NOS* gene **(Figure 1E)**.
-This reporter also revealed the axonal projections of the NOS-expressing neurons.
-The position and morphology of the four *NOS*-expressing cells allowed us to identify the same cells in the three-day-old whole-body volume EM data [@williams2017; @veraszto2020] as four interneurons (INNOS) **(Figure 1B,C)**.
-The INNOS cells are postsynaptic to the UV-sensory cPRCs and presynaptic to the INRGW interneurons, which are also cPRC targets **(Figure 1C,F)**.
+*NOS* was also expressed in the region of the visual eyes (adult eyes) and the pigmented eyespots **(Figure 1---figure supplement 1)**.
+The four apical organ cells, but not the eyes, were also labelled with a *NOS* reporter construct driving palmitoylated-tdTomato **(Figure 1E)**.
+This reporter also revealed the axonal projections of these central NOS-expressing neurons.
+The position and morphology of the four *NOS+* cells allowed us to identify the same four cells as four interneurons (INNOS) in our three-day-old whole-body *Platynereis* volume EM data [@williams2017; @veraszto2020] **(Figure 1B,C)**.
+In the synaptic connectome, the INNOS cells are postsynaptic to the UV-sensory cPRCs and presynaptic to the INRGW interneurons, which are also cPRC targets **(Figure 1C,F)**.
 INRGW neurons synapse on the head serotonergic ciliomotor neurons (Ser-h1), which synapse on the prototroch ciliary band and the cholinergic MC ciliomotor neurons **(Figure 1C,F)**.
-We also detected *NOS* expression by HCR in the region of the visual eyes (adult eyes) and the pigmented eyespots **(Figure 1---figure supplement 1)**.
 
@@
 
-The specific expression of *NOS* in interneurons directly postsynaptic to the cPRCs suggests that NO signalling is involved in the UV response.
-To test if NO is produced during UV/violet responses, we injected zygotes with the fluorescent NO-reporter DAF-FM and imaged two-day-old larvae during the local stimulation of the cPRC cilia by 405 nm light.
-Following light stimulation, we detected an increase in DAF-FM fluorescence in the region of the INNOS cells **(Figure 2)**.
+The expression of *NOS* in the INNOS interneurons in the cPRC circuit suggests that NO signalling may be involved in UV/violet-avoidance.
+To test this, first we asked whether NO is produced during UV/violet stimulation of the larvae.
+We injected the fluorescent NO-reporter DAF-FM into zygotes and imaged two-day-old larvae while exposing the ramified sensory cilia of the cPRCs to 405 nm violet light.
+Following light stimulation of the cPRCs but not a control area, we detected an increase in DAF-FM fluorescence in the anterior neurosecretory neuropil, the region of INNOS projections **(Figure 2)**.
 
@@
 
-To test this, we generated two *Platynereis* *NOS* knockout lines with the CRISPR/Cas9 method.
-We recovered two deletions (NOSΔ11/Δ11 and Δ23/Δ23), both leading to an early stop codon and thus likely representing null alleles **(Figure 3A)**.
+Next, we wanted to directly test whether NO signalling is involved in UV/violet avoidance.
+To achieve this, we generated two *Platynereis* *NOS* knockout lines with the CRISPR/Cas9 method.
+We recovered two deletions (NOSΔ11/Δ11 and Δ23/Δ23), both leading to frame-shift and an early stop codon and thus likely representing null alleles **(Figure 3A)**.
 We could establish a homozygous line for both mutations indicating that NOS is not an essential gene.
-To quantify UV avoidance, we recorded the trajectories of freely swimming larvae in vertical columns, illuminated laterally from two opposite sides with 395 nm UV/violet light.
-To analyse the behaviour of larvae before and after UV stimulation, the temporal changes in some parameters were compared using tracking of individual larvae of the WT and mutant **(Figure 3B,C and Figure 3---figure supplement 1A-D)**.
-The results showed that in the WT there was a rapid and strong increase in downward swimming approximately 15 s after UV stimulation, whereas this did not occur in the mutant **(Figure 3C and Figure 3---figure supplement 1B)**.
-Interestingly, there was already a difference in swimming speed between the WT and the mutant before UV stimulation (in the dark) **(Figure 3---figure supplement 1C,D)**.
-After UV stimulation, the swimming speed of the WT increased and reached a maximum after about 20 s.
-In contrast, the mutants showed a slight increase in speed after UV stimulation, but swimming speed was not maintained and returned to the speed before UV stimulation.
-We also tested phototaxis, by exposing larvae to 480 nm directional and collimated light from the top of the column.
-As previously shown, wild-type larvae swim downward within 30 sec following non-directional UV/violet light stimulation [@Veraszto2018].
-In contrast, homozygous *NOS*-mutant larvae showed a strongly diminished UV-avoidance response **(Figure 3D and Figure 3---figure supplement 1E)**.
+To quantify UV avoidance, we recorded the trajectories of freely swimming wild type and mutant larvae in vertical columns, illuminated laterally from two opposite sides with 395 nm UV light **(Figure 3B, C and Figure 3---figure supplement 1A-D)**.
+As previously shown, wild-type larvae swim downward following non-directional UV/violet light stimulation [@Veraszto2018].
+In contrast, both two- and three-day-old homozygous *NOS*-mutant larvae showed a strongly diminished UV-avoidance response **(Figure 3B-D and Figure 3---figure supplement 1)**.
 This phenotype is similar to the defective UV-avoidance of *c-opsin1* mutant larvae and reveals a requirement for NOS in UV-avoidance behaviour [@Veraszto2018].
-*NOS*-mutant larvae also showed reduced phototactic behaviour, suggesting a function in the visual eyes mediating phototaxis [@Randel2014a], where *NOS* is also expressed **(Figure 1---figure supplement 1A,B)**.
-To test if the observed defects are due to a role of NOS in neuronal signalling or circuit developmental, we also tested larvae exposed to a NOS inhibitor, L-NAME.
-Larvae incubated for 5 min in 0.1 mM or 1 mM L-NAME showed defective UV avoidance, but normal phototaxis **(Figure 3E)**.
-Overall, our results indicate an acute requirement for NOS signalling in UV-avoidance and a possible indirect, developmental role in the visual system.
-To investigate the role of NO on UV avoidance behaviour, receptors for NO were explored.
+Wild type but not mutant larvae also showed an increase in swimming speed under UV light that may be due to downward swimming trajectories (swimming with gravity) **(Figure 3---figure supplement 1C,D)**.
+We also tested directional phototaxis, by exposing larvae to 480 nm directional and collimated light from the top of the column.
+Three-day-old but not two-day-old *NOS*-mutant larvae also showed reduced phototactic behaviour, suggesting a function for *NOS* in the visual eyes that mediate three-day-old phototaxis [@Randel2014a] **(Figure 1---figure supplement 1A,B)**.
+To distinguish between an acute or developmental function of NOS in light responses, we next tested larvae exposed to the NOS inhibitor L-NAME.
+Larvae incubated for 5 min in 0.1 mM or 1 mM L-NAME showed a dose-dependent inhibition of UV avoidance.
+In contrast, phototaxis was not effected **(Figure 3E)**.
+Overall, our results indicate an acute requirement for NOS signalling in UV-avoidance and a possible indirect, developmental role in visual phototaxis.
 
@@
 
+**NO feedback tunes cPRC responses to UV/violet stimulation**
+
+To investigate how NO signalling alters the dynamics of the cPRC circuit, we carried out calcium imaging experiments.
+We ubiquitously expressed the calcium sensor GCaMP6s in larvae and imaged calcium signals during 405 nm stimulation of the cPRCs.
+As we have shown previously, a 20-sec local stimulation of cPRC cilia lead to a transient increase in cPRC calcium levels, followed by a transient decrease [@Veraszto2018].
+After ~20-sec, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus -- a response that may involve depolarisation.
+This activation phase occurs after the 20 sec stimulation period and is likely due to a delayed neuroendocrine feedback [@Veraszto2018].
+To determine whether NO mediates such a feedback, we repeated the experiment in *NOS*-mutant larvae.
+While we detected the initial activation phase followed by inhibition, in *NOS*-mutants this was not followed by delayed activation and calcium levels dropped to a low steady-state level **(Figure 4E,F)**.
+We thus identified a requirement for NO signalling in the late-phase activation of cPRCs.
+
 **Two unconventional guanylyl cyclases are expressed in the cPRCs**
 
-NO generally acts via soluble guanylate cyclases (sGC), belonging to the guanylate cyclase family with a CYC domain (PF00211).
+Next, we aimed to identify the NO-receptor in the cPRCs.
+NO generally acts via soluble guanylate cyclases (sGC), belonging to the guanylate cyclase family with a CYC domain (PFAM domain: PF00211).
 NO binding to the heme group of sGC leads to increased cyclic guanosine monophosphate (cGMP) production.
@@
 In agreement with Moroz et al.
-[@Moroz2020],we found a group of GCs with NIT domains with representatives in placozoans, cnidarians, some ecdysozoans, echinoderms, and lophotrochozoans.
+[@Moroz2020], we found a group of GCs with NIT domains with representatives in placozoans, cnidarians, some ecdysozoans, echinoderms, and lophotrochozoans.
 The 15 *Platynereis* sequences were parts of several major and ancient clades of the tree **(Figure 4---figure supplement 1)**.
-To characterise the expression of NIT-GCs, we using previously published spatially mapped single-cell transcriptome data [@Achim2015; @williams2017].
+To characterise the expression of NIT-GCs, we used previously published spatially mapped single-cell transcriptome data [@Achim2015; @williams2017].
 Among the 15 NIT-GCs, two showed high and specific expression in the cPRCs and one was expressed in the INNOS cells **(Figure 4---figure supplement 1)**.
 In the single-cell data, we could identify the cPRCs by the specific expression of c-opsin1 and a pedal-peptide2 neuropeptide precursor (MLD proneuropeptide) [@Arendt2004; @williams2017], which have previously been described as cPRC markers **(Figure 4---figure supplement 2A)**.
-The INNOS cells were identified by NOS expression and their spatial mapping.
-We decided to focus on two NIT-GCs expressed in the cPRCs and with a full-length sequences, NIT-GC1 and NIT-GC2.
+The INNOS cells were identified by NOS expression and spatial mapping in the brain [@Achim2015].
+We decided to focus on two NIT-GCs expressed in the cPRCs and with a full-length sequence, NIT-GC1 and NIT-GC2.
 To confirm the single-cell data, we first carried out in situ HCR with probes for *NIT-GC1* and *NIT-GC2* mRNA.
-Both genes showed very clear and specific expression in the four cPRCs, as confirmed by co-labeling with an acetylated α-tubulin antibody and with an HCR probe against *MLD proneuropeptide* **(Figure 4A,B and Figure 4---figure supplement 2A-C)**.
+Both genes were specifically expression in the four cPRCs, as confirmed by co-labeling with an acetylated α-tubulin antibody and with an HCR probe against *pedal peptide 2/MLD proneuropeptide* **(Figure 4A,B and Figure 4---figure supplement 2A-C)**.
 To analyse the subcellular localisation of NIT-GC1 and NIT-GC2 at the protein level, we raised and affinity-purified polyclonal antibodies against a specific peptide sequence from both proteins.
 In immunostainings, we found that NIT-GC1 was localise to the region corresponding to the axonal projections of the cPRCs in the anterior nerve plexus **(Figure 4C)**.
-Co-immunostaining with the rabbit NIT-GC1 and the rat NOS antibodies revealed that both proteins were localised in close proximity in the neurosecretory plexus **(Figure 4---figure supplement 2D)**.
-In contrast, NIT-GC2 strongly and specifically labelled the large sensory ciliary region of the cPRCs **(Figure 4D)**.
+Co-immunostaining with the rabbit NIT-GC1 antibody and a custom rat antibody raised against a fragment of *Platynereis* NOS, revealed that both proteins were localised in close proximity in the neurosecretory plexus **(Figure 4---figure supplement 2D)**.
+In contrast, NIT-GC2 specifically labelled the large sensory ciliary region of the cPRCs **(Figure 4D)**.
 These different subcellular localisations suggest that the two NIT-GCs are involved in different intracellular signalling processes in the ciliary and axonal regions of the cPRCs.
 
-**NIT-GC1 is required for the depolarising response of cPRC to UV stimulation**
+**NIT-GC1 produces cGMP in an NO-dependent manner**
 
-To investigate how NIT-GCs-mediated NO signalling is involved in cPRC activation, the calcium sensor GCaMP6s was ubiquitously expressed in larvae and calcium imaging in the cPRC during UV/violet stimulation.
-As we have shown previously, a 20-sec local stimulation of the ciliated part of the cPRC with 405 nm light lead to a transient decrease in cPRC calcium levels, likely due to the hyperpolarisation of the cPRCs [@Veraszto2018].
-After 20-sec, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus - a response that may involve depolarisation.
-This activation phase occurs after the end of the stimulation and may be due to a slower neuroendocrine feedback [@Veraszto2018].
-First, to determine whether NO mediates such a feedback, we measured calcium levels in *NOS*-mutant larvae.
-While we detected the inhibition phase in cPRC in *NOS*-mutants, this was not followed by activation and calcium levels remained low **(Figure 4E,F)**.
-These results revealed an essential requirement for NO signalling in the second phase of the cPRC response.
-Next, to characterise the role of NIT-GCs in changes of calcium levels in cPRC, calcium imaging was performed when each *NIT-GC* gene was knocked down using morpholino **(Figure 4---figure supplement 2E,F)**.
+To further characterise these atypical guanylyl cyclases, we focused on NIT-GC1 and carried out in vitro experiments.
+In bacteria, NIT domains are thought to regulate cellular functions in response to intra- or extracellular nitrate and nitrite.
+NIT-GC1 has a NIT domain and a highly conserved cyclase domain that is expected to catalyse cGMP synthesis **(Figure 4I)**.
+The NIT domain may render NIT-GC1 dependent on NO signals.
+To test this, we co-expressed the synthetic cGMP indicator Green cGull [@Matsuda2017] and NIT-GC1 in cultured cells COS-7 (monkey kidney).
+For balanced expression, we used a single plasmid with the two open-reading frames separated by the 2A self-cleaving peptide **(Figure 4J)**.
+Application of the NO donor SNAP lead to increased cGMP levels, an effect we did not observe when cells were exposed to DMSO or when Green cGull was expressed alone **(Figure 4K-M)**.
+To test whether cGMP production is dependent on the NIT domain, we also tested a deletion construct of NIT-GC1 lacking the NIT domain **(Figure 4I)**.
+Cells expressing this construct and Green cGull did not show increased cGMP levels when exposed to SNAP **(Figure 4N)**.
+These results indicate that NIT-GC1 is able to catalyse cGMP production in an NO-dependent manner and this function requires the NIT domain.
+These results establish NIT-GC1 as a biochemical NO sensor.
+
+**NIT-GC1 is required for NO-mediated feedback to cPRCs during the UV response**
+
+To test the in vivo function of NIT-GC1 and NIT-GC2 in cPRC responses, we returned to calcium imaging. 
+To interfere with NIT-GC function, we used morpholino-mediated knockdowns.
+We used two translation-blocking morpholinos for each *NIT-GC* gene and tested knockdown efficiency by immunostaining injected animals with the NIT-GC1 and NIT-GC2 antibodies **(Figure 4---figure supplement 2E,F)**.
+For both genes, the morpholinos led to a strong reduction in the respective antibody signal, confirming that the morpholinos efficiently knock down translation and at the same time that the antibodies are specific.
+
 In NIT-GC1 morphant larvae, 405 nm light stimulation caused a decrease in calcium levels in the cPRC, but the increase in calcium levels that occurred after the light stimulation was stopped was strongly suppressed **(Figure 4G)**.
@@
 The different responses of cPRC upon knockdown of NIT-GC1 and NIT-GC2 suggest that there may be different intracellular signalling mechanisms associated with their respective localisations.
-We focused our analysis on NIT-GC1 because it was localised to cPRC neuroaxons within the neuroplexas region and is involved in extracellular signalling after UV stimulation.
-It is generally known that depolarisation in the PRC involves elevated cGMPs in the photoreceptors.
-We therefore used an in vitro system in which cultured cells expressed NIT-GC1 to assess the cGMP activity of NIT-GC1.
 
@@
 
-**NIT-GC1 produces cGMP by nitric oxide**
+**Depolarisation of cPRC is required for activation of downstream interneurons and ciliomotor neurons**
 
-NIT-GC1 has a highly conserved cyclase domain and activation of NIT-GC1 is expected to produce cGMP **(Figure 4I)**.
-Recently, a synthetic protein called Green cGull was developed as a cGMP indicator that can be used in cultured cell systems [@Matsuda2017].
-We co-expressed Green cGull and NIT-GC1 using the 2A self-cleaving peptide in cultured cells COS-7 (monkey kidney) and tested the intracellular dynamics of cGMP in the activation of NIT-GC1 **(Figure 4J)**.
-When the NO donor SNAP was applied to the NIT-GC1 and Green cGull co-expression cells, the intracellular cGMP levels was significantly increased compared to when NIT-GC1 was not expressed **(Figure 4K,L)**.
-The increase in fluorescence intensity also did not occur when DMSO was applied **(Figure 4M)**.
-To confirm whether this increase in intracellular cGMP levels is NIT domain-dependent response of NIT-GC1, mutant form of NIT-GC1 without the NIT domain was expressed and SNAP was applied **(Figure 4I)**.
-As a result, the increase in fluorescence intensity was significantly reduced **(Figure 4N)**.
-These results indicate that NIT-GC1 produces cGMP via the NIT domain.
-In bacteria, it has been proposed that the NIT domain regulates cellular functions in response to changes in extracellular and intracellular nitrate and nitrite concentrations.
-Our results suggest that NIT-GC1 may be sensitive to NO.
-This could be a possible feedback pathway in which NIT-GC1 expressed in cPRC is activated by NO produced by INNOS.
-In order to further explore the role of NO in the cPRC circuit, we therefore compared calcium imaging of postsynaptic cells (INNOS, INRGW, Ser-h1 and MC) in the cPRC circuit during UV stimulation of cPRC between WT and NOS mutant larvae.
-
-**Depolarisation of cPRC is required for activation of downstream interneurons and ciliomotor neurons**
+therefore compared calcium imaging of postsynaptic cells (INNOS, INRGW, Ser-h1 and MC) in the cPRC circuit during UV stimulation of cPRC between WT and NOS mutant larvae.
 

### 2023-05-07 Manuscript/Manuscript.Rmd
MS: edits to NIT-GC morpholino section

@@
 
-To test the in vivo function of NIT-GC1 and NIT-GC2 in cPRC responses, we returned to calcium imaging. 
-To interfere with NIT-GC function, we used morpholino-mediated knockdowns.
+To test the in vivo function of NIT-GC1 and NIT-GC2 in cPRC responses, we combined calcium imaging with morpholino-mediated knockdowns.
 We used two translation-blocking morpholinos for each *NIT-GC* gene and tested knockdown efficiency by immunostaining injected animals with the NIT-GC1 and NIT-GC2 antibodies **(Figure 4---figure supplement 2E,F)**.
-For both genes, the morpholinos led to a strong reduction in the respective antibody signal, confirming that the morpholinos efficiently knock down translation and at the same time that the antibodies are specific.
+For both genes, the morpholinos led to a strong reduction in the respective antibody signal, confirming efficient knockdown and antibody specificity.
 
-In NIT-GC1 morphant larvae, 405 nm light stimulation caused a decrease in calcium levels in the cPRC, but the increase in calcium levels that occurred after the light stimulation was stopped was strongly suppressed **(Figure 4G)**.
-This result is similar to the response of NOS mutants in cPRC, suggesting that NIT-GC1 mediates depolarising responses after UV stimulation of cPRC by NO and may be involved in extracellular signalling.
-Interestingly, in NIT-GC2 morphant larvae, an increase in calcium levels occurred immediately after light stimulation **(Figure 4H)**.
-The different responses of cPRC upon knockdown of NIT-GC1 and NIT-GC2 suggest that there may be different intracellular signalling mechanisms associated with their respective localisations.
+In NIT-GC1 morphant larvae, the delayed activation of cPRCs following 405 nm stimulation did not occur **(Figure 4G)**.
+This phenotype is similar to the phenotype of *NOS* mutants suggesting that NIT-GC1 acts as the NO sensor in cPRCs to drive their delayed activation.
+This could occur via increased cGMP production and the opening of a cyclic-nucleotide-gated (CNG) channel specific to cPRCs.
+NIT-GC2 morphant larvae, in contrast, showed a step-up increase in calcium following light stimulation **(Figure 4H)**.
+The calcium signal decayed during stimulation and was off after light off.
+These data support an essential role for ciliary-localised NIT-GC2 in suppressing cPRC calcium following its transient rise at stimulus onset.
+Overall, these knockdown experiments revealed different signalling mechanisms for the two NIT-GCs that may be due to their different subcellular localisations.
 

### 2023-05-07 Manuscript/Manuscript.Rmd
MS edits

@@
 
-**Depolarisation of cPRC is required for activation of downstream interneurons and ciliomotor neurons**
+**NO signalling shapes the dynamics of the cPRC circuit**
 
-therefore compared calcium imaging of postsynaptic cells (INNOS, INRGW, Ser-h1 and MC) in the cPRC circuit during UV stimulation of cPRC between WT and NOS mutant larvae.
+To investigate how NO and NIT-GC signalling influence the dynamics of the cPRC circuit, we imaged calcium signals from postsynaptic neurons in wild type, mutant and morphant larvae.
+We were able to image the activity of all neurons in the cPRC circuit (INNOS, INRGW, Ser-h1 and MC).
+The MC cell was identified based on its position and intrinsic activity.
+To unambiguously identify all other cells from which we recorded calcium signals, we developed an on-slide immunostaining method **(Figure 5---figure supplement 1)**.
+We used the cell-specific antibody markers against RYamide (INNOS), RGWamide (INRGW) and serotonin (Ser-h1) [@Conzelmann2011] to immunostain agar-embedded larvae following calcium imaging.
+Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision **(Figure 5A,B)**.
+Due to the stereotypy of the larvae, we could also identify neurons based on their position and activity in activity-correlation maps **(Figure 5C)**.
 
-We next quantified the responses of the INNOS and INRGW following a UV/violet stimulus to the cPRCs.
-To unambiguously identify the cells from which we recorded calcium signals, we developed an on-slide immunostaining method **(Figure 5---figure supplement 1)**.
-We used the cell-specific neuropeptide markers using anti-RYamide (INNOS) and anti-RGWamide (INRGW) antibodies [@Conzelmann2011] to mark the interneurons.
-Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision **(Figure 5A,B)**.
-In both wild-type and *NOS*-mutant larvae, INNOS cells showed increase calcium signals during the inhibition phase of the cPRCs **(Figure 5D)**.
-In addition, INNOS depolarisation was suppressed in NIT-GC2 molphants where cPRC hyperpolarisation was suppressed **(Figure 5E)**.
-This indicates that cPRC hyperpolarisation is required for INNOS depolarisation.
-In contrast, calcium signals increased in INRGW cells in wild-type larvae during the second activation phase of cPRCs, a response that was lacking in *NOS*-mutants **(Figure 5F,G)**.
-This suggests that there is no circuit output to the Ser-h1 and MC cells and the ciliary band in *NOS*-mutants, consistent with the observed behavioural defects **(Figure 5H,I)**.
+We first quantified the responses of the INNOS and INRGW interneurons during 405 nm stimulation of the cPRCs.
+In both wild type and *NOS*-mutant larvae, INNOS cells showed an increase in calcium during stimulation **(Figure 5D)**.
+This response was suppressed in NIT-GC2 morphant larvae **(Figure 5E)** revealing an essential role for NIT-GC2-mediated cPRC suppression in INNOS activation.
+In contrast, INRGW cells were initially inhibited, followed by a delayed activation paralleling the second activation phase of cPRCs.
+This late INRGW response was lacking in *NOS*-mutants **(Figure 5F,G)**.
+In NIT-GC2 morphants, INRGW cells showed a transient increase in calcium that decayed after light off and a delayed activation was not present.
+Next, we imaged calcium signals from Ser-h1 and MC neurons in wild type and *NOS* mutant larvae.
+Ser-h1 cells showed an activation profile that correlated with cPRC activity, including a reduction in calcium during stimulation followed by rebound, a response that was defective in *NOS* mutants.
+MC cells showed sustained activation, including a late-phase that was lacking in *NOS* mutants **(Figure 5H,I)**.
+These data suggest that during 405 stimulation the Ser-h1 cells are inhibited and MC cells are activated, and this regulation is NO-dependent.
+This pattern is expected to inhibit ciliary activity in the prototroch but not in the other ciliary bands, triggering NO-dependent downward swimming.
 

### 2023-05-07 text/Supplements.Rmd
change figure

@@
 ![**Figure 4---figure supplement 2.** **(A)** Co-expression analysis image of the NIT-GC1 (magenta) and MLD-pedal2 amide proneuropeptide gene (MLD: green).
-Anterior view of the larva at 2-day-old.
+Anterior view of the larva at two-day-old.
 **(B, C)** Expression analysis of the NIT-GC2 gene (magenta) using in situ HCR.
-Anterior (B) and posterior (C) views of the larva at 3-day-old.
-**(D, E)** Localisation analysis for NIT-GC1 (B) and NIT-GC2 (C) morphant using NIT-GC1 and NIT-GC2 antibodies.
+Anterior (B) and posterior (C) views of the larva at three-day-old.
+**(D)** Co-localisation analysis using NIT-GC1 (magenta) and NOS (green) antibodies.
+Anterior view of the larva at two-day-old.
+**(E, F)** Localisation analysis using NIT-GC1 and NIT-GC2 antibodies for NIT-GC1 (E) and NIT-GC2 (F) morphant.
 Green shows co-staining with acetylated α-tubulin antibody (acTub).
-Anterior view of the larva at 2-day-old.
-**(F)** Co-localisation analysis using NIT-GC1 (magenta) and NOS (green) antibodies.
-Anterior view of the larva at 2-day-old.](../figures/Fig4_sup2.png){width="100%"}
+Anterior view of the larva at two-day-old.](../figures/Fig4_sup2.png){width="100%"}
+
+![Figure 5---figure supplement 1.
+Schematic diagram of the immunostaining procedure after calcium imaging.](../figures/Fig5_sup1.png){width="100%"} ![Figure 6---figure supplement 1.](../figures/Figure6_sup1.png){width="100%"}
+
+![**Figure 6---figure supplement 2.**](../figures/Figure6_sup2.png){width="100%"}
+
+![**Figure 6---figure supplement 3.**](../figures/Figure6_sup3.png){width="100%"}

### 2023-05-07 code/Fig1.R

@@ -316,8 +316,8 @@ panel_NOS2d_ant_HCR <- ggdraw() + draw_image(readPNG("pictures/HCR-IHC_51_AP_NOS
 panel_NOS3d <- ggdraw() + draw_image(readPNG("pictures/NOS-promotor_3d_acTub_XXum_crop.png")) +
   draw_label("transgene + IHC", x = 0.38, y = 0.99, size = 10) +
   draw_label("*", x = 0.55, y = 0.32, color='white',size = 18,fontface="plain") +
-  draw_image(readPNG("pictures/INNOS_vl.png"), 
-             scale = 0.35, x = 1, hjust = 1, halign = 1, valign = 0.075) +
+  #draw_image(readPNG("pictures/INNOS_vl.png"), 
+  #           scale = 0.35, x = 1, hjust = 1, halign = 1, valign = 0.075) +
   #draw_label("INNOS_vl", x = 0.9, y = 0.3, size = 5, color = "black") +
   draw_label("NOSp::palmi-3xHA", x = 0.34, y = 0.9, color="magenta", size = 10, fontface="plain") +
   draw_label("acTub", x = 0.8, y = 0.9, color="green", size = 10, fontface="plain") +

### 2023-05-07 code/Fig1_sup2.R

@@ -31,7 +31,8 @@ panel_NOS3d_ventr_HCR <- ggdraw() + draw_image(readPNG("pictures/HCR_72_DV_NOS_9
   draw_line(x = c(0.77, 0.66), y = c(0.65, 0.67), color = "white", size = 0.1) +
   draw_label("adult eyes", x = 0.55, y = 0.34, size = 5, color = "white") +
   draw_line(x = c(0.26, 0.44), y = c(0.43, 0.35), color = "white", size = 0.1) +
-  draw_line(x = c(0.77, 0.65), y = c(0.43, 0.35), color = "white", size = 0.1)
+  draw_line(x = c(0.77, 0.65), y = c(0.43, 0.35), color = "white", size = 0.1) +
+  draw_label("INNOS", x = 0.53, y = 0.45, color="white", size = 5, fontface="bold")
 
 
 panel_NOS3d_ant_HCR <- ggdraw() + draw_image(readPNG("pictures/HCR_72_AP_NOS_89.25um.png")) +
@@ -53,7 +54,8 @@ panel_NOS3d_ant_HCR <- ggdraw() + draw_image(readPNG("pictures/HCR_72_AP_NOS_89.
   draw_line(x = c(0.59, 0.77), y = c(0.7, 0.67), color = "white", size = 0.1) +
   draw_label("eyespot", x = 0.25, y = 0.37, size = 5, color = "white") +
   draw_label("INNOS_dl", x = 0.73, y = 0.36, size = 4, color = "white") +
-  draw_label("INNOS_vl", x = 0.92, y = 0.09, size = 4, color = "white")
+  draw_label("INNOS_vl", x = 0.92, y = 0.09, size = 4, color = "white") +
+  draw_label("INNOS", x = 0.51, y = 0.52, color="white", size = 5, fontface="bold")
 
 
 

### 2023-05-07 code/Fig4.R

@@ -393,7 +393,7 @@ NITGC1_analysis %>%
   scale_color_manual(values=c('grey50')) +
   scale_y_continuous(breaks=seq(1.0, 1.2, length=3),limits = c(0.947, 1.2))+
   scale_x_continuous(breaks=seq(0, 10, length=6),limits = c(0, 10))+
-  labs(title = "Green cGull + ΔNIT-GC1", 
+  labs(title = "Green cGull + 'NIT-GC1'^'ΔNIT'", 
        x = "time (min)", 
        y = "normalized intensity") +
   annotate("segment", x=2, xend=10, y=1.165, yend=1.165, size=2, color = "dark gray")+
@@ -462,26 +462,25 @@ panel_GcG_mutNIT_SNAP <- ggdraw() + draw_image(readPNG("pictures/GcG-mutNIT-GC1-
 #combine panels into Figure and save final figure as pdf and png
 #panels of different sizes
 layout <- "
-AA#BB#CC#DD
+AAAAA#BBBBB
 ###########
-EEEEE#FFFFF
+CC#DD#EE#FF
 ###########
-GGGGG#HHHHH
+GGGIIIIJJJJ
+HHHKKKKLLLL
 ###########
-IIIKKKKLLLL
-JJJMMMMNNNN
+MMMMM#NNNNN
 "
 
 
-
 Fig4 <- 
-  panel_HCR_NIT1 + panel_HCR_NIT2 + panel_IHC_NIT1 + panel_IHC_NIT2 +
   panel_cPRC_NOS11 + panel_cPRC_NOS23 +
-  panel_cPRC_NIT1_MO12 + panel_cPRC_NIT2_MO12 +
+  panel_HCR_NIT1 + panel_HCR_NIT2 + panel_IHC_NIT1 + panel_IHC_NIT2 +
   panel_NITGC1_domain + panel_NITGC1_assay_schematic + 
   panel_GcG_NIT_SNAP + panel_GcG_SNAP + panel_GcG_NIT_DMSO + panel_GcG_mutNIT_SNAP +
+  panel_cPRC_NIT1_MO12 + panel_cPRC_NIT2_MO12 +
   patchwork::plot_layout(design = layout, 
-                         heights = c(1, 0.02, 1, 0.02, 1, 0.02, 0.75, 0.75),
+                         heights = c(1, 0.02, 1, 0.02, 0.75, 0.75, 0.02, 1),
[…]

### 2023-05-07 code/Fig4_sup1.R

@@ -1,4 +1,4 @@
-#Code to assemble Fig4 sup1 of the Jokura et al Platynereis NOS paper
+#Code to assemble Fig4 sup2 of the Jokura et al Platynereis NOS paper
 #2022 June - Kei Jokura, Gaspar Jekely
 
 # source packages ---------------------------------------------------------
@@ -14,12 +14,12 @@ layout <- "
 A
 "
 
-Fig4_sup1 <- panel_A + 
+Fig4_sup2 <- panel_A + 
   patchwork::plot_layout(design = layout) #we can change the heights of the rows in our layout (widths also can be defined)
 
 
-ggsave("figures/Fig4_sup1.png", limitsize = FALSE, 
-       units = c("px"), Fig4_sup1, width = 2350, height = 1700, bg='white')  
+ggsave("figures/Fig4_sup2.png", limitsize = FALSE, 
+       units = c("px"), Fig4_sup2, width = 1500, height = 2100, bg='white')  
 
-ggsave("figures/Fig4_sup1.pdf", limitsize = FALSE, 
-       units = c("px"), Fig4_sup1, width = 2350, height = 1700)  
+ggsave("figures/Fig4_sup2.pdf", limitsize = FALSE, 
+       units = c("px"), Fig4_sup2, width = 1500, height = 2100)  

### 2023-05-07 code/Fig4_sup2.R

@@ -1,97 +1,25 @@
-#Code to assemble Fig4 sup2 of the Jokura et al Platynereis NOS paper
+#Code to assemble Fig4 sup1 of the Jokura et al Platynereis NOS paper
 #2022 June - Kei Jokura, Gaspar Jekely
 
-rm(list = ls(all.names = TRUE)) #will clear all objects includes hidden objects.
-gc() #free up memory and report the memory usage.
+# source packages ---------------------------------------------------------
+source("code/Packages_to_load.R")
 
-# load some packages
-library(tidyverse)
-library(cowplot)
-library(png)
-library(patchwork)
-library(magick)
 
 # assemble figure ---------------------------------------------------------
-arrow_fluo <- data.frame(x1 = 0.95, x2 = 0.95, y1 = 0.75, y2 = 0.85)
-
-panel_NIT1_MLD <- ggdraw() + draw_image(readPNG("pictures/HCR_51_AP_NITGC1_MLD_DAPI_112.55µm.png"))+
-  draw_label("in situ HCR", x = 0.25, y = 0.96, size = 10) +
-  draw_label("NIT-GC1", x = 0.12, y = 0.87, color="magenta", size = 10, fontface="italic") +
-  draw_label("MLD", x = 0.31, y = 0.87, color="green", size = 10, fontface="italic") +
-  draw_label("DAPI", x = 0.48, y = 0.87, color="cyan", size = 10, fontface="plain") +
-  draw_line(x = c(0.04, 0.22), y = c(0.11, 0.11), color = "white", size = 0.5) +
-  draw_label(expression(paste("20 ", mu, "m")), x = 0.13, y = 0.14, color = "white", size = 8) +
-  draw_label("D", x = 0.95, y = 0.88, size = 6, color = "white") +
-  draw_label("V", x = 0.95, y = 0.72, size = 6, color = "white") +
-  geom_segment(aes(x = x1, y = y1, xend = x2, yend = y2), data = arrow_fluo, color = "white", 
-               arrow = arrow(ends = "both", type = "closed", length = unit(0.1,"cm")),
-               lineend = "butt",
-               linejoin = "mitre",
-               arrow.fill = "white", size = 0.2)
-
[…]

### 2023-05-08 Manuscript/Manuscript.Rmd
MS: Discussion edits [tbc]

@@
 
-## Discussion
+**Mathematical modelling of cPRC-circuit dynamics**
 
-In the present study, we explored the role of NO in the UV avoidance behaviour of Platynereis.
-We demonstrated that there are interneurons expressing NOS in the cPRC circuit and that NO is required for UV avoidance in behavioural experiments using NOS mutants.
-We identified the axon-localised type NIT-GC1 and the cilia-localised type NIT-GC2 expressed in the cPRCs as targets of NO and showed from cultured cell experiments that these produce cGMP in response to NO.
-We have shown by calcium imaging that INNOSs are activated in response to the hyperpolarising response of the cPRCs, that the NO produced feeds back to the cPRC via NIT-GC1 and induces a depolarising response in the cPRC, and that the depolarising response of the cPRC is involved in activating downstream INRGW.
-Lastly, we confirmed that RGWamide induces the downward swimming behaviour of larvae.
+[to be added]
 
@@
 
+## Discussion
+
+The activity of synaptic circuits is richly modulated monoamines, neuropeptides and gaseous molecules.
+Our work revealed an essential role for NO-mediated signalling in driving UV/violet avoidance behaviour.
+NO, produced by postsynaptic interneurons, feeds back to their presynaptic photoreceptor partners leading to their delayed activation.
+This late-phase activation of cPRCs drives circuit output through projection interneurons and ciliomotor neurons.
+In the cPRC circuit, synaptic connectivity alone is not sufficient to drive circuit dynamics and behavioural change.
+
+We identified cPRC-expressed NIT-GC1 as a new type of NO sensor.
+NIT-GC1 is localised to the axonal projections of the cPRCs in the anterior neurosecretory plexus of the brain.
+Here, cPRC axons intermingle and synapse on INNOS neurites that contain NOS.
+NO-mediated feedback signalling thus likely occurs in the neurosecretory plexus where NOS- and NIT-GC1-containing projections are in close proximity and where we could detect NO production following UV stimulation.
+
+We showed that NIT-GC1 can produce cGMP in a NO-dependent manner.
+In cPRCs, increased cGMP levels could open CNGAα, a cyclic nucleotide-gated channels specifically expressed in cPRC [@Tosches2014], leading to cPRC delayed depolarisation, even after UV off.
+
+NO is a free radical with a half-life of 5 msec in vivo and a signalling range limited to a few tens of nanometers.
+In neuronal signalling, NOS and sGC have been found to localise at close proximity at synapses [@burette2002][@garthwaite2015a].
+We identified 12 sGCs in *Platynereis*, but none of these is expressed in the cPRC.
+Instead, we identified an unconventional NIT-domain containing GC, NIT-GC1 as the mediator of NO retrograde signalling.
+The NIT domain was first identified in bacteria and animal NIT-GCs have only been recently reported [@Shu2003; @Moroz2020].
+Bacterial NIT domains regulate cellular functions in response to changes in extracellular and intracellular nitrate and/or nitrite concentrations [@camargo2007].
+NO is readily converted to nitrate and nitrite [@santos2011; @möller2019; @garthwaite2015a] and these molecules accumulate in placozoans and cnidarians in cells and tissues with high NOS activity [@Moroz2004; @Moroz2020].
+It is possible that the NIT domains in NIT-GCs may also sense nitrate and nitrite, as in bacteria.
+Furthermore, if they show different sensitivities to NO, nitrite and nitrate, it is possible that differences in their half-lives could give rise to a range of activation timings [@lundberg2011].
+We have found 15 NIT-GCs in *Platynereis*, suggesting a wide range of functions.
+In addition, NIT-GC1 and NIT-GC2 showed very different subcellular localisation and functions.
+It is not clear whether the functional differences are due to the different localisation or also to differences in biochemical function.
+Overall, the diversification and spatially different localisation patterns of NIT-GCs may increase their signalling repertoire.
+Our data revealed key roles for two *Platynereis* NIT-GCs in phototraonsduction and NO-mediated neuromodulation suggesting that these molecules contribute to the diversity of neuronal signalling.
+Sponges and ctenophores, where NO signalling is present, lack NIT-GCs.
+In placozoans as many as 12 have been found, and cnidarians also contain a large diversity.
+NIT-GCs have been lost from chordates [check].
+
 cPRCs expressing UV-absorbing c-opsin1 have high resting calcium levels and show hyperpolarising and subsequent depolarising responses to UV light (405 nm) [@Veraszto2018].
 In the present study, we have shown that the cells that strongly synapse with cPRCs are interneurons that specifically express NOS (INNOS).
-The fluorescence intensity of DAF-FM, which specifically detected INNOS, showed a transient tonic-like increase dependent on UV stimulation of the cPRCs.
-The graph shows a decrease in intensity after UV light is switched off, but the response of DAF-FM to NO is irreversible and this decrease is due to breaching by the relatively strong excitation light.
-Attempts to measure changes in the intensity of the calcium indicator RGECO1a simultaneously with DAF-FM imaging did not lead to detection of changes in fluorescence intensity compared to GCaMP6s and could only visualise the cPRCs.
-In calcium imaging using GCaMP6s, INNOS identified by the positional relationship between the cPRC and INNOS and the labelling of RYa depolarised in clear synchrony with the hyperpolarising response of the cPRCs.
-This tonic response may involve dopaminergic nerves based on single-cell data **(Figure 6)** [@puopolo2001][@zhang2007][@Randel2014a].
+
+[@puopolo2001][@zhang2007][@Randel2014a].
 In the mammalian retina, it is known that signals from bipolar photoreceptor cells are transmitted to downstream NOS-expressing amacrine cells (NOAC), where NO is produced [@jacoby2018].
@@
 
-Our results suggest that NIT-GC1, which is specifically expressed in cPRCs, functions as a target for NO produced by INNOS and contributes to the regulation of the cPRC circuit transmission pathway via a depolarising response by increasing cGMP.
 In the mammalian retina, NO is an essential signalling molecule and one of the most well-studied [@Cudeiro1999].
 It has been reported that NO produced in one of the retinal cells modulates cGMP-mediated neurotransmission by activating presynaptic neural sGCs (general target of NO) through retrograde signalling [@Vielma2014][@Wei2012].
-In Platynereis cPRC, NIT-GC1 was also identified as having the property of producing cGMP in a NO-dependent manner.
-Furthermore, it is known that CNGAα, one of the cyclic nucleotide-gated channels, is specifically highly expressed in cPRC [@Tosches2014], and in INNOS-cPRC signalling, NO is implicated in the neuromodulation of the cPRC by retrograde signalling.
-NO is a molecule with a very short half-life, and when functioning as a signalling molecule, NOS and sGC are known to localise very close together at synapses between neurons [@burette2002][@garthwaite2015a].
-In our study, we found 12 sGCs in Platynereis, but none expressed in the cPRC.
-Localisation analysis of NOS and NIT-GC1 by antibody staining showed that the two proteins localised very close together in the anterior neurosecretory plexus region.
-Calcium imaging results using NOS mutant and NIT-GC1 knockdown larvae suggest that NIT-GC1 activation induces a depolarising response in the cPRC.
 
@@
 The present results suggest that the depolarising response of the cPRC due to NO feedback induces INRGW activation.
-RGWamide is also a neuropeptide that induces larval downward swimming quite strongly (unpublished?
-need to show data?), and the delayed downward swimming of NOS mutant larvae in response to UV stimulation may be responsible for the failure of INRGW activation to occur.
+
 In addition, in many marine invertebrates, NO has been reported to be involved in many aspects of settlement behaviour and metamorphosis induction [@leise2001; @bishop2003; @bishop2007; @pechenik2007; @castellano2014; @ueda2013; @ueda2014; @ueda2016; @yang2018; @zhu2020; @song2021].
@@
 These suggest that NO feedback signalling may act as a trigger for downward swimming via activation of INRGWs, which induces the initiation of settlement behaviour and metamorphosis.
-
-An outstanding question is how the NO\~NIT-GC response alters neuronal dynamics.
-NO is a single molecule and, due to its short half-life, cannot be transmitted over long distances and has a limited signalling repertoire.
-It has therefore been considered unsuitable for generating diversity in signalling compared to other signalling molecules such as neuropeptides and classical neurotransmitters [@Jékely2021].
-However, our results in the present study indicate that NO-induced activation of NIT-GC1 is involved in the regulation of neuronal circuits (cPRC pathway).
-The NIT domain was originally found in bacteria and the existence of GCs with NIT domains was first reported by Moroz et al in 2020 [@Shu2003; @Moroz2020].
-It has been proposed that NIT domains in bacteria regulate cellular functions in response to changes in extracellular and intracellular nitrate and/or nitrite concentrations [@camargo2007].
-In fact, NO is a free radical molecule, with a half-life of 5 msec in vivo, and is readily converted to nitrate and nitrite [@santos2011; @möller2019; @garthwaite2015a].
-And nitrate and nitrite have been reported to accumulate in cells and tissues with high NOS activity in placozoans and cnidarians [@Moroz2004; @Moroz2020].
-It is possible that the NIT domains of NIT-GC1 identified in this study may also be used for nitrate/nitrite sensing, as in bacteria.
-Furthermore, if they show different sensitivities to NO, nitrite and nitrate, it is possible that the temporal variation in their respective half-lives could give rise to a repertoire of activation timings [@lundberg2011].
-In addition, we have found at least 15 NIT-GCs in the Platynereis transcriptome data, of which NIT-GC2 is highly expressed in the cPRC, in addition to NIT-GC1.
-Interestingly, NIT-GC2 showed a very different localisation pattern to NIT-GC1, and furthermore, the results of calcium imaging of cPRC in response to UV stimulation using NIT-GC2 knockdown larvae were the different dynamics (such that hyperpolarisation was suppressed) with NIT-GC1 knockdown.
-It was not clear from the present analysis whether this difference was due to differences in the localisation or the property of the NIT-GCs.
-However, the diversification and spatially different localisation patterns of NIT-GCs may increase their signalling repertoire.
-We also believe that the appearance of NIT-GCs was one of the means to generate the diversity of signalling pathways using NO during the early stages of nervous system evolution in metazoans, since sponges and ctenophores, where NO signalling is present, have no NIT-GCs at all, as many as 12 have been found in placozoans, and NIT-GCs are functional in the "proto-eye" circuit of Platynereis.
 

### 2023-05-09 text/Manuscript.Rmd

@@
 However, little is known about the molecular neural circuit mechanisms in UV avoidance behaviour in many animals, including zooplankton.
-In neural circuits in general light response, the activation of various receptors by excitatory and inhibitory neurotransmitters such as glutamate, GABA and dopamine is known in various species.
+In neursal circuits in general light response, the activation of various receptors by excitatory and inhibitory neurotransmitters such as glutamate, GABA and dopamine is known in various species.
 In addition to this, it has long been known that in nitric oxide (NO), as a neuromodulatory molecule, it is also important as a modulator of visual activity.
@@
 Analysis of sGCs in *Platynereis* indicated that these genes are not expressed in any of the cells of the cPRC circuit [@Veraszto2017].
-Recently, Moroz and coworkers reported an atypical but widely conserved family of guanylyl cyclases with a NIT (nitrite/nitrate sensing) domain (PF08376) (NIT-GC) as potential mediators of NO signalling [@Moroz2020].
+Recently, Moroz and coworkers reported an atypical but widely conserved family of guanylyl cyclases with a NIT (nitrite/nitrate sensing) domain (PF08376) (NIT-GC)[@Shu2003; @monteagudo-cascales2023] as potential mediators of NO signalling [@Moroz2020].
 To search for NIT-GCs in *Platynereis*, we searched transcriptome resources and retrieved 15 potential NIT-GC homologs **(Figure 4---figure supplement 1)**.
@@
 
+'''{r cars}
+
+readxl
+
 ## Acknowledgements

### 2023-05-09 Manuscript/Manuscript.Rmd
ms added

@@
 csl: elife.csl
+editor_options: 
+  markdown: 
+    wrap: sentence
 ---
@@
 
-Neuronally produced nitric oxide (NO) can act as a diffusible neuromodulator. This small free-radical molecule is synthesized by nitric oxide synthase (NOS)...
+Neuronally produced nitric oxide (NO) can act as a diffusible neuromodulator.
+This small free-radical molecule is synthesized by nitric oxide synthase (NOS)...
 
@@
 The NO-dependent increased in cGMP tunes bipolar cell activity ....
-[refs]
-[other examples]
+[refs] [other examples]
 
@@
 As we have shown previously, a 20-sec local stimulation of cPRC cilia lead to a transient increase in cPRC calcium levels, followed by a transient decrease [@Veraszto2018].
-After ~20-sec, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus -- a response that may involve depolarisation.
+After \~20-sec, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus -- a response that may involve depolarisation.
 This activation phase occurs after the 20 sec stimulation period and is likely due to a delayed neuroendocrine feedback [@Veraszto2018].
@@
 We carried out cluster analysis and did phylogenetic reconstruction on a group of membrane-bound guanylyl cyclases with sGCs as an outgroup.
-In agreement with Moroz et al.
-[@Moroz2020], we found a group of GCs with NIT domains with representatives in placozoans, cnidarians, some ecdysozoans, echinoderms, and lophotrochozoans.
+In agreement with Moroz et al. [@Moroz2020], we found a group of GCs with NIT domains with representatives in placozoans, cnidarians, some ecdysozoans, echinoderms, and lophotrochozoans.
 The 15 *Platynereis* sequences were parts of several major and ancient clades of the tree **(Figure 4---figure supplement 1)**.
@@
 The eggs were kept at 18°C for 45 min before injection and were injected at 14.5°C.
-The injected individuals were kept at 18°C for 5 to 8 days in 6-well- plates (Nunc multidish no.
-150239, Thermo Scientific) and then cultured at 22°C until sexual maturity.
+The injected individuals were kept at 18°C for 5 to 8 days in 6-well- plates (Nunc multidish no. 150239, Thermo Scientific) and then cultured at 22°C until sexual maturity.
 The mature worms were crossed to wild-type worms and the progeny was genotyped, resulting in two founder lines, which were bred to homozygosity.
@@
 
-We used Achim et al.
-for the single-cell data [@Achim2015].
-In fact, in Williams et al., they used 107 cells as neurons by removing duplicates from Achim et al.
-single-cell data, so we used those cells [@williams2017].
+We used Achim et al. for the single-cell data [@Achim2015].
+In fact, in Williams et al., they used 107 cells as neurons by removing duplicates from Achim et al. single-cell data, so we used those cells [@williams2017].
 Since the raw data were read count data, we normalized them to TPM using Python.
@@
 To remove excess hairpins, samples were washed in 1 ml 5× SSCT at room temperature, twice for 5 min, twice for 30 min, and once for 5 min.
-During the first 30 min wash, samples were counterstained with DAPI (Cat.
-#40043, Biotium, USA).
+During the first 30 min wash, samples were counterstained with DAPI (Cat. #40043, Biotium, USA).
 
@@
 Larvae were imaged on a Zeiss LSM 880 with Airyscan (with a C-Apochromat 63X/1.2 Corr - water) with a frame rate of 1.88 frame/sec and an image size of 512 x 512 pixels.
-The larvae were stimulated in a region of interest (a circle with ??
-pixel diameter) with 405 nm lasers controlled by the Bleaching mode.
+The larvae were stimulated in a region of interest (a circle with ?? pixel diameter) with 405 nm lasers controlled by the Bleaching mode.
 The imaging laser had a similar intensity than the stimulus laser but covered an area that was 10 times larger than the stimulus ROI.
@@
 Cos-7 cells with low expression of endogenous soluble guanylate cyclase were used as cultured cells for gene expression.
-This cell line was purchased from Angio-proteomie (CAT no.
-cAP-0203).
+This cell line was purchased from Angio-proteomie (CAT no. cAP-0203).
 The Cos-7 cells were maintained at 37 °C in 35mm dishes (Nunc™ Glass Bottom Dishes) containing 3 mL of DMEM, high glucose glutamax medium (Thermo; Cat.
@@
 Upon reaching confluency of approximately 85%, we transfected the cells with the plasmid containing Green cGull-T2A-NITGC1.
-Transfections were carried out with 150 ng of each plasmid and 0.3 μl of the transfection Lipofectamine 3000 Reagent (invitrogen; Cat.
-No.
-L3000001).
+Transfections were carried out with 150 ng of each plasmid and 0.3 μl of the transfection Lipofectamine 3000 Reagent (invitrogen; Cat. No. L3000001).
 Two days post-transfection, we removed the culture medium and substituted it for fresh DMEM-medium.
@@
 Images were acquired using a xenon lamp, 460--495 nm excitation filter, 505-nm dichroic mirror and 510-- 550-nm emission filter (????).
-S-Nitroso-N-acetyl-D, L-penicillamine (SNAP) was purchased from Sigma-Aldrich (St.
-Louis, MO, USA) .
+S-Nitroso-N-acetyl-D, L-penicillamine (SNAP) was purchased from Sigma-Aldrich (St. Louis, MO, USA) .
 The exposure time of the EM-CCD camera was controlled by the ZEN software (Zeiss).

### 2023-05-11 Manuscript/Manuscript.Rmd
Add key_resource_table

@@
 
+##Key resources table
+
+```         
+library(readxl)
+NOSpaper_keyresourcetale <- read_excel("source_data/key_resource_tale.xlsx")
+
+{r echo=FALSE,fig.cap="**Key resources table** "}
+```
+
 ## Materials and Methods
@@
 
+```{r echo=FALSE,fig.cap="**Figure 1 -- figure supplement 3** "}
+knitr::include_graphics(path = "figures/Fig1_sup3.png")
+
+```
+
 ```{r echo=FALSE,fig.cap="**Figure 3 -- figure supplement 1** "}
@@
 
+```{r echo=FALSE,fig.cap="**Figure 4 -- figure supplement 3** "}
+knitr::include_graphics(path = "figures/Fig4_sup3.png")
+
+```
+
 ```{r echo=FALSE,fig.cap="**Figure 5 -- figure supplement 1** "}

### 2023-05-11 code/Key_resources_table.R

@@ -1,13 +1,0 @@
-
-#Code to assemble Key resourece table of the Jokura et al Platynereis NOS paper
-#2022 June - Kei Jokura, Gaspar Jekely
-
-# source packages ---------------------------------------------------------
-library(tidyverse)
-library(gt)
-
-library(readxl)
-keyresource <- read_excel("~/NOSpaper_keyresourcetale.xlsx")
-
-keyresource %>% 
-  gt()

### 2023-05-11 code/Fig4_sup1.R
Update Fig4_sup1 CLAN data

@@ -1,4 +1,4 @@
-#Code to assemble Fig4 sup2 of the Jokura et al Platynereis NOS paper
+#Code to assemble Fig4 sup1 of the Jokura et al Platynereis NOS paper
 #2022 June - Kei Jokura, Gaspar Jekely
 
 # source packages ---------------------------------------------------------
@@ -6,7 +6,7 @@ source("code/Packages_to_load.R")
 
 
 # assemble figure ---------------------------------------------------------
-panel_A <- ggdraw() + draw_image(readPNG("pictures/FINALSeq_NITGC_allseq_mbGC_sGC_CD80_plus_Lopho.fasta.alngapp.fasta.treefile.png"))
+panel_A <- ggdraw() + draw_image(readPNG("pictures/FINALSeq_NITGC_allseq_mbGC_sGC_CLAN.png"))
 
 #combine panels into Figure and save final figure as pdf and png
 #panels of different sizes
@@ -14,20 +14,15 @@ layout <- "
 A
 "
 
-Fig4_sup2 <- panel_A + 
+Fig4_sup1 <- panel_A + 
   patchwork::plot_layout(design = layout) #we can change the heights of the rows in our layout (widths also can be defined)
 
 
-<<<<<<< HEAD
-ggsave("figures/Fig4_sup2.png", limitsize = FALSE, 
-       units = c("px"), Fig4_sup2, width = 1500, height = 2100, bg='white')  
-
-ggsave("figures/Fig4_sup2.pdf", limitsize = FALSE, 
-       units = c("px"), Fig4_sup2, width = 1500, height = 2100)  
-=======
 ggsave("Manuscript/figures/Fig4_sup1.png", limitsize = FALSE, 
-       units = c("px"), Fig4_sup1, width = 2350, height = 1700, bg='white')  
+       units = c("px"), Fig4_sup1, width = 2400, height = 1900, bg='white') 
 
 ggsave("Manuscript/figures/Fig4_sup1.pdf", limitsize = FALSE, 
-       units = c("px"), Fig4_sup1, width = 2350, height = 1700)  
->>>>>>> aca2b70175ce8fc82878146d4b6e4da25e84fdf8
+       units = c("px"), Fig4_sup1, width = 2400, height = 1900)  
+
+
+

### 2023-05-11 code/Fig4_sup2.R

@@ -1,4 +1,4 @@
-#Code to assemble Fig4 sup1 of the Jokura et al Platynereis NOS paper
+#Code to assemble Fig4 sup2 of the Jokura et al Platynereis NOS paper
 #2022 June - Kei Jokura, Gaspar Jekely
 
 # source packages ---------------------------------------------------------
@@ -6,7 +6,7 @@ source("code/Packages_to_load.R")
 
 
 # assemble figure ---------------------------------------------------------
-panel_A <- ggdraw() + draw_image(readPNG("pictures/FINALSeq_NITGC_allseq_mbGC_sGC_CLAN.png"))
+panel_A <- ggdraw() + draw_image(readPNG("pictures/FINALSeq_NITGC_allseq_mbGC_sGC_CD80_plus_Lopho.fasta.alngapp.fasta.treefile.png"))
 
 #combine panels into Figure and save final figure as pdf and png
 #panels of different sizes
@@ -14,20 +14,13 @@ layout <- "
 A
 "
 
-Fig4_sup1 <- panel_A + 
+Fig4_sup2 <- panel_A + 
   patchwork::plot_layout(design = layout) #we can change the heights of the rows in our layout (widths also can be defined)
 
-<<<<<<< HEAD
 
-ggsave("figures/Fig4_sup1.png", limitsize = FALSE, 
-       units = c("px"), Fig4_sup1, width = 2800, height = 1800, bg='white')  
-
-ggsave("figures/Fig4_sup1.pdf", limitsize = FALSE, 
-       units = c("px"), Fig4_sup1, width = 2800, height = 1800) 
-=======
 ggsave("Manuscript/figures/Fig4_sup2.png", limitsize = FALSE, 
-       units = c("px"), Fig4_sup2, width = 2400, height = 1900, bg='white') 
+       units = c("px"), Fig4_sup2, width = 1500, height = 2100, bg='white')  
 
 ggsave("Manuscript/figures/Fig4_sup2.pdf", limitsize = FALSE, 
-       units = c("px"), Fig4_sup2, width = 2350, height = 1700)  
->>>>>>> aca2b70175ce8fc82878146d4b6e4da25e84fdf8
+       units = c("px"), Fig4_sup2, width = 1500, height = 2100)  
+

### 2023-05-11 code/Fig4_sup3.R

@@ -91,8 +91,8 @@ Fig4_sup3 <- panel_NIT1_MLD + panel_72_AP_NIT2 + panel_72_DV_NIT2 +
   patchwork::plot_annotation(tag_levels = 'A') &  #we can change this to 'a' for small caps or 'i' or '1'
   ggplot2::theme(plot.tag = element_text(size = 12, face='plain')) #or 'plain', 'italic'
 
-ggsave("figures/Fig4_sup3.png", limitsize = FALSE, 
+ggsave("Manuscript/figures/Fig4_sup3.png", limitsize = FALSE, 
        units = c("px"), Fig4_sup3, width = 2400, height = 1750, bg='white') 
 
-ggsave("figures/Fig4_sup3.pdf", limitsize = FALSE, 
+ggsave("Manuscript/figures/Fig4_sup3.pdf", limitsize = FALSE, 
        units = c("px"), Fig4_sup3, width = 2400, height = 1750)  

### 2023-05-11 Manuscript/Manuscript.Rmd
Update Supplement legends

@@
 
-```{r echo=FALSE,fig.cap="**Figure 1 -- figure supplement 1** "}
+```{r echo=FALSE,fig.cap="**Figure 1 -- figure supplement 1**  Phylogenetic tree of NOS by maximum likelihood.
+Tree robustness was tested with 1000 replicates of ultrafast bootstrap with the aLRT-SH-like and aBayes methods."}
 knitr::include_graphics(path = "figures/Fig1_sup1.png")
@@
 
-```{r echo=FALSE,fig.cap="**Figure 1 -- figure supplement 2** "}
+```{r echo=FALSE,fig.cap="**Figure 1 -- figure supplement 2** Expression analysis of the *NOS* gene (magenta) using in situ HCR using the larva at three-day-old in dorsal **(A)** and anterior view (B).
+(B) Insert: showing two NOS cells (INNOS_dl and INNOS_vl) on left side."}
 knitr::include_graphics(path = "figures/Fig1_sup2.png")
@@
 
-```{r echo=FALSE,fig.cap="**Figure 3 -- figure supplement 1** "}
+```{r echo=FALSE,fig.cap="**Figure 3 -- figure supplement 1** **(A)** Overlaid trajectories for WT (n=37) and NOS mutant (NOSΔ11/Δ11, n=18 and NOSΔ23/Δ23, n=8) at two-day-old larvae.
+0 sec as the starting point.
+After 10 sec, UV (395 nm) stimulation from the side.
+**(B)** The temporal changes in the vertical position of the WT and mutant 2-days-old larvae before and after UV stimulation are shown.
+The starting points of each larval trajectory are set to 0.
+After UV stimulation is indicated by purple squares.
+**(C)** Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23) larvae at 2-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
+The data are shown in 30 s bins.
+**(D, E)** The temporal changes in the distance traveled of the WT and mutant in 2-day (D) and 3-day-old (E) larvae before and after UV stimulation are shown."}
 knitr::include_graphics(path = "figures/Fig3_sup1.png")
@@
 
-```{r echo=FALSE,fig.cap="**Figure 4 -- figure supplement 2** "}
+```{r echo=FALSE,fig.cap="**Figure 4 -- figure supplement 2** Phylogenetic tree of guanylate cyclase by maximum likelihood (ML).
+Guanylate cyclase-coupled receptor and soluble guanylate cyclases (sGC) as outgroups.
+Guanylate cyclases with NIT domains are found in most animal phyla except Porifera, Ctenophora, Urochordata and Chordata.
+Dot plot of Platynereis NIT-GC genes (columns) expressed in cPRC, INNOS and INRGWa (rows) using single cell RNA-Seq.
+The size of the dots is expressed in proportion to the percentage of cells expressing that gene relative to all cells.
+The colours represent the normal logarithm of the number of transcripts in the cells expressing the gene."}
 knitr::include_graphics(path = "figures/Fig4_sup2.png")
@@
 
-```{r echo=FALSE,fig.cap="**Figure 4 -- figure supplement 3** "}
+```{r echo=FALSE,fig.cap="**Figure 4 -- figure supplement 3** Co-expression analysis image of the NIT-GC1 (magenta) and MLD-pedal2 amide proneuropeptide gene (MLD: green).
+Anterior view of the larva at two-day-old.
+**(B, C)** Expression analysis of the NIT-GC2 gene (magenta) using in situ HCR.
+Anterior (B) and posterior (C) views of the larva at three-day-old.
+**(D)** Co-localisation analysis using NIT-GC1 (magenta) and NOS (green) antibodies.
+Anterior view of the larva at two-day-old.
+**(E, F)** Localisation analysis using NIT-GC1 and NIT-GC2 antibodies for NIT-GC1 (E) and NIT-GC2 (F) morphant.
+Green shows co-staining with acetylated α-tubulin antibody (acTub).
+Anterior view of the larva at two-day-old."}
 knitr::include_graphics(path = "figures/Fig4_sup3.png")
@@
 
-```{r echo=FALSE,fig.cap="**Figure 5 -- figure supplement 1** "}
+```{r echo=FALSE,fig.cap="**Figure 5 -- figure supplement 1** Schematic diagram of the immunostaining procedure after calcium imaging."}
 knitr::include_graphics(path = "figures/Fig5_sup1.png")

### 2023-05-11 Manuscript/Manuscript.Rmd
Update Key resource table

@@
 output:
+  html_document:
+    theme: lumen
   pdf_document:
@@
 
-##Key resources table
+## Key resources table
 
-```         
+```{r echo=FALSE}
 library(readxl)
-NOSpaper_keyresourcetale <- read_excel("source_data/key_resource_tale.xlsx")
-
-{r echo=FALSE,fig.cap="**Key resources table** "}
+key_resource_tale <- read_excel("key_resource_tale/key_resource_tale.xlsx")
+df_key_resource_tale <- as.data.frame(key_resource_tale)
+knitr::kable(df_key_resource_tale, format = "html", booktabs = TRUE)
 ```

### 2023-05-11 Manuscript/Manuscript.tex
Update Key resource table

@@
 
-The ability to perceive changes in the environment and adapt behaviour
-accordingly is essential for the survival of all living organisms.
-Avoidance from UV radiation, which damages DNA essential for survival,
-is proposed to be one of the behaviours that existed before the abundant
-development of Earth's oxygen to block UV radiation in the atmosphere
-(Pittendrigh). In the ocean, almost all zooplankton have a pronounced
-rhythmic behaviour called Diel vertical migration (DVM), which consists
-of upwards towards the surface at dusk and descend back to deeper water
-before dawn. Avoidance behaviour to escape harmful UV radiation during
-the day is closely linked to the control of vertical migration in the
-ocean. However, little is known about the molecular neural circuit
-mechanisms in UV avoidance behaviour in many animals, including
-zooplankton. In neural circuits in general light response, the
-activation of various receptors by excitatory and inhibitory
-neurotransmitters such as glutamate, GABA and dopamine is known in
-various species. In addition to this, it has long been known that in
-nitric oxide (NO), as a neuromodulatory molecule, it is also important
-as a modulator of visual activity. NO is a small free-radical molecule
-with diffusible and changeable characteristics, and thus works as one of
-the neural signalling molecules. The function of NO in the neural
-circuitry of vision has been well studied in vertebrates. In the
-mammalian retina, NO is an essential signalling molecule in all retinal
-cells. NO produced by nitric oxide synthase (NOS) in amacrine cells, one
-of the retinal cells, activates \emph{soluble guanylate cyclase (sGC)}
-expressed in cone bipolar cells, which produces cGMP and regulates
-downstream signalling. We therefore hypothesise that NO-mediated neural
-signalling may be involved in the photoreceptor circuits that control UV
-avoidance behaviour. To test this hypothesis we used the annelid
-\emph{Platynereis dumerilii} \textbf{(Figure 1A)}, which conserved
-vertebrate-type ciliary photoreceptors (cPRCs) (Arendt et al., 2004).
-The cPRC expresses \emph{c-opsin1}, which absorbs UV radiation, is
-involved in melatonin production and regulates circadian swimming
-behaviour (Tosches et al., 2014; Tsukamoto et al., 2017). Furthermore,
-we previously reported abnormal UV avoidance behaviour in
-\emph{c-opsin1} knockout larvae (Verasztó et al., 2018). The cPRCs
-express the UV-absorbing c-opsin1 and mediate UV avoidance in the larvae
-(Verasztó et al., 2018). In adult worms, UV light and c-opsin1 modulate
-locomotion and maturation (Veedin Rajan et al., 2021). These suggest
-that Platynereis larvae are a highly suitable model for elucidating the
-neural circuit mechanisms of NO-induced UV avoidance behaviour, and the
-accessibility of genome editing, behavioural assays, calcium imaging and
-whole-body connectomics could be very useful for study at the molecular
-level(Ozpolat et al., 2021).
+Neuronally produced nitric oxide (NO) can act as a diffusible
+neuromodulator. This small free-radical molecule is synthesized by
+nitric oxide synthase (NOS)\ldots{}
+
+In the mammalian retina, NO is produced by amacrine cells and activates
+a \emph{soluble guanylate cyclase (sGC)} in cone bipolar cells. The
+NO-dependent increased in cGMP tunes bipolar cell activity \ldots.
+{[}refs{]} {[}other examples{]}
+
+To investigate NO signalling in the modulation of neural circuit
+dynamics, we studied the larval stages of the marine annelid model
+\emph{Platynereis dumerilii} (Ozpolat et al., 2021). We uncovered an
+essential function for NO in these larvae in UV/violet-light avoidance.
+In \emph{Platynereis} larvae, UV/violet avoidance is characterised by
+downward swimming and is mediated by the UV/violet-responding brain
+ciliary photoreceptor cells (cPRCs).
+
+The cPRCs express a vertebrate-type ciliary opsin, c-opsin1 (Arendt et
+al., 2004). This opsin forms a UV-absorbing bistable photopigment with
+an absorption maximum around 384 nm Verasztó et al. (2018).
+\emph{c-opsin1} knockout larvae show defective UV/violet avoidance
+(Verasztó et al., 2018). Here we show that \emph{Platynereis} \emph{NOS}
+is expressed in interneurons of the cPRC circuitry and is required for
+UV/violet-avoidance behaviour. By combining calcium imaging across the
+fully-mapped cPRC circuit, genetic perturbations and mathematical
+modelling, we describe how NO tunes circuit dynamics through
+non-synaptic reverse signalling to cPRCs. This delayed neuroendocrine
+feedback is essential for circuit output, and integrates UV/violet input
+and also induces a short-term memory manifest in altered circuit
+activity.
+
+\#In adult worms, UV light and c-opsin1 modulate locomotion and
+maturation (Veedin Rajan et al., 2021).
 
@@
 weakly expressing) in the apical organ region \textbf{(Figure 1D and
-Figure 1---figure supplement 2)}. The same four cells were labelled with
-a reporter construct driving palmitoylated-tdTomato by a 12 kbp upstream
-genomic region of the \emph{NOS} gene \textbf{(Figure 1E)}. This
-reporter also revealed the axonal projections of the NOS-expressing
-neurons. The position and morphology of the four \emph{NOS}-expressing
-cells allowed us to identify the same cells in the three-day-old
-whole-body volume EM data (Verasztó et al., 2020; Williams et al., 2017)
-as four interneurons (INNOS) \textbf{(Figure 1B,C)}. The INNOS cells are
+Figure 1---figure supplement 2)}. \emph{NOS} was also expressed in the
+region of the visual eyes (adult eyes) and the pigmented eyespots
+\textbf{(Figure 1---figure supplement 1)}. The four apical organ cells,
+but not the eyes, were also labelled with a \emph{NOS} reporter
+construct driving palmitoylated-tdTomato \textbf{(Figure 1E)}. This
+reporter also revealed the axonal projections of these central
+NOS-expressing neurons. The position and morphology of the four
+\emph{NOS+} cells allowed us to identify the same four cells as four
+interneurons (INNOS) in our three-day-old whole-body \emph{Platynereis}
+volume EM data (Verasztó et al., 2020; Williams et al., 2017)
+\textbf{(Figure 1B,C)}. In the synaptic connectome, the INNOS cells are
 postsynaptic to the UV-sensory cPRCs and presynaptic to the INRGW
@@
 which synapse on the prototroch ciliary band and the cholinergic MC
-ciliomotor neurons \textbf{(Figure 1C,F)}. We also detected \emph{NOS}
-expression by HCR in the region of the visual eyes (adult eyes) and the
-pigmented eyespots \textbf{(Figure 1---figure supplement 1)}.
+ciliomotor neurons \textbf{(Figure 1C,F)}.
 
@@
 Numbers inside hexagons indicate the number of cells grouped in each node.
-Arrow line thickness (synapse weight) is equal to the number of synapses.}\label{fig:unnamed-chunk-2}
+Arrow line thickness (synapse weight) is equal to the number of synapses.}\label{fig:unnamed-chunk-1}
 \end{figure}
@@
 
-The specific expression of \emph{NOS} in interneurons directly
-postsynaptic to the cPRCs suggests that NO signalling is involved in the
-UV response. To test if NO is produced during UV/violet responses, we
-injected zygotes with the fluorescent NO-reporter DAF-FM and imaged
-two-day-old larvae during the local stimulation of the cPRC cilia by 405
-nm light. Following light stimulation, we detected an increase in DAF-FM
-fluorescence in the region of the INNOS cells \textbf{(Figure 2)}.
+The expression of \emph{NOS} in the INNOS interneurons in the cPRC
+circuit suggests that NO signalling may be involved in
+UV/violet-avoidance. To test this, first we asked whether NO is produced
+during UV/violet stimulation of the larvae. We injected the fluorescent
+NO-reporter DAF-FM into zygotes and imaged two-day-old larvae while
+exposing the ramified sensory cilia of the cPRCs to 405 nm violet light.
+Following light stimulation of the cPRCs but not a control area, we
+detected an increase in DAF-FM fluorescence in the anterior
+neurosecretory neuropil, the region of INNOS projections \textbf{(Figure
+2)}.
 
 \begin{figure}
-\centering
-\includegraphics[width=1\textwidth,height=\textheight]{../figures/Fig2.png}
-\caption{\textbf{Figure 2. NO produced by UV/violet stimulation to
-cPRCs.} \textbf{(A)} A NO-positive cell visualised by DAF-FM (green) is
-shown (dashline). The white line indicates the outer frame of the larva.
-All subsequent images are of two-day-old larvae. \textbf{(B)} Intensity
-changes before (upper) and during 405 nm light stimulation (lower) in a
-cell visualised by DAF-FM. \textbf{(C)} The intensity of DAF-FM changes
-over time when cPRC (green) and other cells (ctr: gray) are stimulated
-with UV/violet (405 nm) light. Purple squares indicate the timing of 405
-nm light stimulation, the transparent lines show the optical responses
-to individual intensities, each intensities are normalized (ΔF/F0) and
-the dark trace lines and the shaded bands show the mean value and the
-standard error range of the mean.}
+\includegraphics[width=29.17in]{figures/Fig2} \caption{**Figure 2.
+NO produced by UV/violet stimulation to cPRCs.** **(A)** A NO-positive cell visualised by DAF-FM (green) is shown (dashline).
+The white line indicates the outer frame of the larva.
+All subsequent images are of two-day-old larvae.
+**(B)** Intensity changes before (upper) and during 405 nm light stimulation (lower) in a cell visualised by DAF-FM.
+**(C)** The intensity of DAF-FM changes over time when cPRC (green) and other cells (ctr: gray) are stimulated with UV/violet (405 nm) light.
+Purple squares indicate the timing of 405 nm light stimulation, the transparent lines show the optical responses to individual intensities, each intensities are normalized (ΔF/F0) and the dark trace lines and the shaded bands show the mean value and the standard error range of the mean.}\label{fig:unnamed-chunk-2}
 \end{figure}
@@
 
-To test this, we generated two \emph{Platynereis} \emph{NOS} knockout
-lines with the CRISPR/Cas9 method. We recovered two deletions
-(NOSΔ11/Δ11 and Δ23/Δ23), both leading to an early stop codon and thus
-likely representing null alleles \textbf{(Figure 3A)}. We could
-establish a homozygous line for both mutations indicating that NOS is
-not an essential gene. To quantify UV avoidance, we recorded the
-trajectories of freely swimming larvae in vertical columns, illuminated
-laterally from two opposite sides with 395 nm UV/violet light. To
-analyse the behaviour of larvae before and after UV stimulation, the
-temporal changes in some parameters were compared using tracking of
-individual larvae of the WT and mutant \textbf{(Figure 3B,C and Figure
-3---figure supplement 1A-D)}. The results showed that in the WT there
-was a rapid and strong increase in downward swimming approximately 15 s
-after UV stimulation, whereas this did not occur in the mutant
-\textbf{(Figure 3C and Figure 3---figure supplement 1B)}. Interestingly,
-there was already a difference in swimming speed between the WT and the
-mutant before UV stimulation (in the dark) \textbf{(Figure 3---figure
-supplement 1C,D)}. After UV stimulation, the swimming speed of the WT
-increased and reached a maximum after about 20 s. In contrast, the
-mutants showed a slight increase in speed after UV stimulation, but
-swimming speed was not maintained and returned to the speed before UV
-stimulation. We also tested phototaxis, by exposing larvae to 480 nm
-directional and collimated light from the top of the column. As
-previously shown, wild-type larvae swim downward within 30 sec following
+Next, we wanted to directly test whether NO signalling is involved in
+UV/violet avoidance. To achieve this, we generated two
+\emph{Platynereis} \emph{NOS} knockout lines with the CRISPR/Cas9
+method. We recovered two deletions (NOSΔ11/Δ11 and Δ23/Δ23), both
+leading to frame-shift and an early stop codon and thus likely
+representing null alleles \textbf{(Figure 3A)}. We could establish a
+homozygous line for both mutations indicating that NOS is not an
+essential gene. To quantify UV avoidance, we recorded the trajectories
+of freely swimming wild type and mutant larvae in vertical columns,
+illuminated laterally from two opposite sides with 395 nm UV light
+\textbf{(Figure 3B, C and Figure 3---figure supplement 1A-D)}. As
+previously shown, wild-type larvae swim downward following
 non-directional UV/violet light stimulation (Verasztó et al., 2018). In
-contrast, homozygous \emph{NOS}-mutant larvae showed a strongly
-diminished UV-avoidance response \textbf{(Figure 3D and Figure
-3---figure supplement 1E)}. This phenotype is similar to the defective
-UV-avoidance of \emph{c-opsin1} mutant larvae and reveals a requirement
-for NOS in UV-avoidance behaviour (Verasztó et al., 2018).
-\emph{NOS}-mutant larvae also showed reduced phototactic behaviour,
-suggesting a function in the visual eyes mediating phototaxis (Randel et
-al., 2014), where \emph{NOS} is also expressed \textbf{(Figure
-1---figure supplement 1A,B)}. To test if the observed defects are due to
-a role of NOS in neuronal signalling or circuit developmental, we also
-tested larvae exposed to a NOS inhibitor, L-NAME. Larvae incubated for 5
-min in 0.1 mM or 1 mM L-NAME showed defective UV avoidance, but normal
-phototaxis \textbf{(Figure 3E)}. Overall, our results indicate an acute
-requirement for NOS signalling in UV-avoidance and a possible indirect,
-developmental role in the visual system. To investigate the role of NO
-on UV avoidance behaviour, receptors for NO were explored.
+contrast, both two- and three-day-old homozygous \emph{NOS}-mutant
+larvae showed a strongly diminished UV-avoidance response
+\textbf{(Figure 3B-D and Figure 3---figure supplement 1)}. This
+phenotype is similar to the defective UV-avoidance of \emph{c-opsin1}
+mutant larvae and reveals a requirement for NOS in UV-avoidance
+behaviour (Verasztó et al., 2018). Wild type but not mutant larvae also
+showed an increase in swimming speed under UV light that may be due to
+downward swimming trajectories (swimming with gravity) \textbf{(Figure
+3---figure supplement 1C,D)}. We also tested directional phototaxis, by
+exposing larvae to 480 nm directional and collimated light from the top
+of the column. Three-day-old but not two-day-old \emph{NOS}-mutant
+larvae also showed reduced phototactic behaviour, suggesting a function
+for \emph{NOS} in the visual eyes that mediate three-day-old phototaxis
+(Randel et al., 2014) \textbf{(Figure 1---figure supplement 1A,B)}. To
+distinguish between an acute or developmental function of NOS in light
+responses, we next tested larvae exposed to the NOS inhibitor L-NAME.
+Larvae incubated for 5 min in 0.1 mM or 1 mM L-NAME showed a
+dose-dependent inhibition of UV avoidance. In contrast, phototaxis was
+not effected \textbf{(Figure 3E)}. Overall, our results indicate an
+acute requirement for NOS signalling in UV-avoidance and a possible
+indirect, developmental role in visual phototaxis.
 
 \begin{figure}
-\centering
-\includegraphics[width=1\textwidth,height=\textheight]{../figures/Fig3.png}
-\caption{\textbf{Figure 3. Larvae with inhibited NOS have a delayed
-response to UV-avoidance.} \textbf{(A)} Top: The domain and Exon/Intron
-structure of NOS. Bottom: Close-up region showing the genomic locus of
-NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9. The
-generated mutants (NOSΔ11, NOSΔ23) are also shown. Pink indicates target
-sites. Gray shows PAM sequences, red shows stop codons. \textbf{(B)}
-Overlaid trajectories for WT (n=32) and NOS mutant (NOSΔ11/Δ11, n=26 and
-NOSΔ23/Δ23, n=47) at three-day-old larvae. 0 sec as the starting point.
-After 10 sec, UV (395 nm) stimulation from the side. \textbf{(C)} The
-temporal changes in the vertical position of the WT and mutant larvae
-before (dark) and after UV stimulation (side 395 nm) are shown. The
-starting points of each larval trajectory are set to 0. \textbf{(D)}
-Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23)
-larvae at three-day-old stimulated with UV (395 nm) light from side,
-blue (488 nm) light from top and UV (395 nm) light from top. The data
-are shown in 30 s bins. \textbf{(E)} Vertical swimming in larvae treated
-with NOS inhibitors, L-NAME at three-day-old stimulated with UV (395 nm)
-light from side, blue (488 nm) light from top and UV (395 nm) light from
-top. The data are shown in 30 s bins.}
+\includegraphics[width=33.33in]{figures/Fig3} \caption{**Figure 3.
+Larvae with inhibited NOS have a delayed response to UV-avoidance.** **(A)** Top: The domain and Exon/Intron structure of NOS.
+Bottom: Close-up region showing the genomic locus of NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9.
+The generated mutants (NOSΔ11, NOSΔ23) are also shown.
+Pink indicates target sites.
+Gray shows PAM sequences, red shows stop codons.
+**(B)** Overlaid trajectories for WT (n=32) and NOS mutant (NOSΔ11/Δ11, n=26 and NOSΔ23/Δ23, n=47) at three-day-old larvae.
+0 sec as the starting point.
+After 10 sec, UV (395 nm) stimulation from the side.
+**(C)** The temporal changes in the vertical position of the WT and mutant larvae before (dark) and after UV stimulation (side 395 nm) are shown.
+The starting points of each larval trajectory are set to 0.
+**(D)** Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23) larvae at three-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
+The data are shown in 30 s bins.
+**(E)** Vertical swimming in larvae treated with NOS inhibitors, L-NAME at three-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
+The data are shown in 30 s bins.}\label{fig:unnamed-chunk-3}
 \end{figure}
 
+\textbf{NO feedback tunes cPRC responses to UV/violet stimulation}
+
+To investigate how NO signalling alters the dynamics of the cPRC
+circuit, we carried out calcium imaging experiments. We ubiquitously
+expressed the calcium sensor GCaMP6s in larvae and imaged calcium
+signals during 405 nm stimulation of the cPRCs. As we have shown
+previously, a 20-sec local stimulation of cPRC cilia lead to a transient
+increase in cPRC calcium levels, followed by a transient decrease
+(Verasztó et al., 2018). After \textasciitilde20-sec, calcium levels in
+cPRCs were raising again, reaching higher levels than at the start of
+the stimulus -- a response that may involve depolarisation. This
+activation phase occurs after the 20 sec stimulation period and is
+likely due to a delayed neuroendocrine feedback (Verasztó et al., 2018).
+To determine whether NO mediates such a feedback, we repeated the
+experiment in \emph{NOS}-mutant larvae. While we detected the initial
+activation phase followed by inhibition, in \emph{NOS}-mutants this was
+not followed by delayed activation and calcium levels dropped to a low
+steady-state level \textbf{(Figure 4E,F)}. We thus identified a
+requirement for NO signalling in the late-phase activation of cPRCs.
+
 \textbf{Two unconventional guanylyl cyclases are expressed in the cPRCs}
 
-NO generally acts via soluble guanylate cyclases (sGC), belonging to the
-guanylate cyclase family with a CYC domain (PF00211). NO binding to the
-heme group of sGC leads to increased cyclic guanosine monophosphate
+Next, we aimed to identify the NO-receptor in the cPRCs. NO generally
+acts via soluble guanylate cyclases (sGC), belonging to the guanylate
+cyclase family with a CYC domain (PFAM domain: PF00211). NO binding to
+the heme group of sGC leads to increased cyclic guanosine monophosphate
 (cGMP) production. Analysis of sGCs in \emph{Platynereis} indicated that
@@
 group of membrane-bound guanylyl cyclases with sGCs as an outgroup. In
-agreement with Moroz et al. (Moroz et al., 2020),we found a group of GCs
-with NIT domains with representatives in placozoans, cnidarians, some
-ecdysozoans, echinoderms, and lophotrochozoans. The 15
+agreement with Moroz et al. (Moroz et al., 2020), we found a group of
+GCs with NIT domains with representatives in placozoans, cnidarians,
+some ecdysozoans, echinoderms, and lophotrochozoans. The 15
 \emph{Platynereis} sequences were parts of several major and ancient
 clades of the tree \textbf{(Figure 4---figure supplement 1)}. To
-characterise the expression of NIT-GCs, we using previously published
+characterise the expression of NIT-GCs, we used previously published
 spatially mapped single-cell transcriptome data (Achim et al., 2015;
@@
 described as cPRC markers \textbf{(Figure 4---figure supplement 2A)}.
-The INNOS cells were identified by NOS expression and their spatial
-mapping. We decided to focus on two NIT-GCs expressed in the cPRCs and
-with a full-length sequences, NIT-GC1 and NIT-GC2. To confirm the
-single-cell data, we first carried out in situ HCR with probes for
-\emph{NIT-GC1} and \emph{NIT-GC2} mRNA. Both genes showed very clear and
-specific expression in the four cPRCs, as confirmed by co-labeling with
-an acetylated α-tubulin antibody and with an HCR probe against \emph{MLD
-proneuropeptide} \textbf{(Figure 4A,B and Figure 4---figure supplement
-2A-C)}. To analyse the subcellular localisation of NIT-GC1 and NIT-GC2
-at the protein level, we raised and affinity-purified polyclonal
-antibodies against a specific peptide sequence from both proteins. In
-immunostainings, we found that NIT-GC1 was localise to the region
-corresponding to the axonal projections of the cPRCs in the anterior
-nerve plexus \textbf{(Figure 4C)}. Co-immunostaining with the rabbit
-NIT-GC1 and the rat NOS antibodies revealed that both proteins were
-localised in close proximity in the neurosecretory plexus
-\textbf{(Figure 4---figure supplement 2D)}. In contrast, NIT-GC2
-strongly and specifically labelled the large sensory ciliary region of
-the cPRCs \textbf{(Figure 4D)}. These different subcellular
+The INNOS cells were identified by NOS expression and spatial mapping in
+the brain (Achim et al., 2015). We decided to focus on two NIT-GCs
+expressed in the cPRCs and with a full-length sequence, NIT-GC1 and
+NIT-GC2. To confirm the single-cell data, we first carried out in situ
+HCR with probes for \emph{NIT-GC1} and \emph{NIT-GC2} mRNA. Both genes
+were specifically expression in the four cPRCs, as confirmed by
+co-labeling with an acetylated α-tubulin antibody and with an HCR probe
+against \emph{pedal peptide 2/MLD proneuropeptide} \textbf{(Figure 4A,B
+and Figure 4---figure supplement 2A-C)}. To analyse the subcellular
+localisation of NIT-GC1 and NIT-GC2 at the protein level, we raised and
+affinity-purified polyclonal antibodies against a specific peptide
+sequence from both proteins. In immunostainings, we found that NIT-GC1
+was localise to the region corresponding to the axonal projections of
+the cPRCs in the anterior nerve plexus \textbf{(Figure 4C)}.
+Co-immunostaining with the rabbit NIT-GC1 antibody and a custom rat
+antibody raised against a fragment of \emph{Platynereis} NOS, revealed
+that both proteins were localised in close proximity in the
+neurosecretory plexus \textbf{(Figure 4---figure supplement 2D)}. In
+contrast, NIT-GC2 specifically labelled the large sensory ciliary region
+of the cPRCs \textbf{(Figure 4D)}. These different subcellular
 localisations suggest that the two NIT-GCs are involved in different
@@
 
-\textbf{NIT-GC1 is required for the depolarising response of cPRC to UV
-stimulation}
-
-To investigate how NIT-GCs-mediated NO signalling is involved in cPRC
-activation, the calcium sensor GCaMP6s was ubiquitously expressed in
-larvae and calcium imaging in the cPRC during UV/violet stimulation. As
-we have shown previously, a 20-sec local stimulation of the ciliated
-part of the cPRC with 405 nm light lead to a transient decrease in cPRC
-calcium levels, likely due to the hyperpolarisation of the cPRCs
-(Verasztó et al., 2018). After 20-sec, calcium levels in cPRCs were
-raising again, reaching higher levels than at the start of the stimulus
-- a response that may involve depolarisation. This activation phase
-occurs after the end of the stimulation and may be due to a slower
-neuroendocrine feedback (Verasztó et al., 2018). First, to determine
-whether NO mediates such a feedback, we measured calcium levels in
-\emph{NOS}-mutant larvae. While we detected the inhibition phase in cPRC
-in \emph{NOS}-mutants, this was not followed by activation and calcium
-levels remained low \textbf{(Figure 4E,F)}. These results revealed an
-essential requirement for NO signalling in the second phase of the cPRC
-response. Next, to characterise the role of NIT-GCs in changes of
-calcium levels in cPRC, calcium imaging was performed when each
-\emph{NIT-GC} gene was knocked down using morpholino \textbf{(Figure
-4---figure supplement 2E,F)}. In NIT-GC1 morphant larvae, 405 nm light
-stimulation caused a decrease in calcium levels in the cPRC, but the
-increase in calcium levels that occurred after the light stimulation was
-stopped was strongly suppressed \textbf{(Figure 4G)}. This result is
-similar to the response of NOS mutants in cPRC, suggesting that NIT-GC1
-mediates depolarising responses after UV stimulation of cPRC by NO and
-may be involved in extracellular signalling. Interestingly, in NIT-GC2
-morphant larvae, an increase in calcium levels occurred immediately
-after light stimulation \textbf{(Figure 4H)}. The different responses of
-cPRC upon knockdown of NIT-GC1 and NIT-GC2 suggest that there may be
-different intracellular signalling mechanisms associated with their
-respective localisations. We focused our analysis on NIT-GC1 because it
-was localised to cPRC neuroaxons within the neuroplexas region and is
-involved in extracellular signalling after UV stimulation. It is
-generally known that depolarisation in the PRC involves elevated cGMPs
-in the photoreceptors. We therefore used an in vitro system in which
-cultured cells expressed NIT-GC1 to assess the cGMP activity of NIT-GC1.
+\textbf{NIT-GC1 produces cGMP in an NO-dependent manner}
+
+To further characterise these atypical guanylyl cyclases, we focused on
+NIT-GC1 and carried out in vitro experiments. In bacteria, NIT domains
+are thought to regulate cellular functions in response to intra- or
+extracellular nitrate and nitrite. NIT-GC1 has a NIT domain and a highly
+conserved cyclase domain that is expected to catalyse cGMP synthesis
+\textbf{(Figure 4I)}. The NIT domain may render NIT-GC1 dependent on NO
+signals. To test this, we co-expressed the synthetic cGMP indicator
+Green cGull (Matsuda et al., 2016) and NIT-GC1 in cultured cells COS-7
+(monkey kidney). For balanced expression, we used a single plasmid with
+the two open-reading frames separated by the 2A self-cleaving peptide
+\textbf{(Figure 4J)}. Application of the NO donor SNAP lead to increased
+cGMP levels, an effect we did not observe when cells were exposed to
+DMSO or when Green cGull was expressed alone \textbf{(Figure 4K-M)}. To
+test whether cGMP production is dependent on the NIT domain, we also
+tested a deletion construct of NIT-GC1 lacking the NIT domain
+\textbf{(Figure 4I)}. Cells expressing this construct and Green cGull
+did not show increased cGMP levels when exposed to SNAP \textbf{(Figure
+4N)}. These results indicate that NIT-GC1 is able to catalyse cGMP
+production in an NO-dependent manner and this function requires the NIT
+domain. These results establish NIT-GC1 as a biochemical NO sensor.
+
+\textbf{NIT-GC1 is required for NO-mediated feedback to cPRCs during the
+UV response}
+
+To test the in vivo function of NIT-GC1 and NIT-GC2 in cPRC responses,
+we combined calcium imaging with morpholino-mediated knockdowns. We used
+two translation-blocking morpholinos for each \emph{NIT-GC} gene and
+tested knockdown efficiency by immunostaining injected animals with the
+NIT-GC1 and NIT-GC2 antibodies \textbf{(Figure 4---figure supplement
+2E,F)}. For both genes, the morpholinos led to a strong reduction in the
+respective antibody signal, confirming efficient knockdown and antibody
+specificity.
+
+In NIT-GC1 morphant larvae, the delayed activation of cPRCs following
+405 nm stimulation did not occur \textbf{(Figure 4G)}. This phenotype is
+similar to the phenotype of \emph{NOS} mutants suggesting that NIT-GC1
+acts as the NO sensor in cPRCs to drive their delayed activation. This
+could occur via increased cGMP production and the opening of a
+cyclic-nucleotide-gated (CNG) channel specific to cPRCs. NIT-GC2
+morphant larvae, in contrast, showed a step-up increase in calcium
+following light stimulation \textbf{(Figure 4H)}. The calcium signal
+decayed during stimulation and was off after light off. These data
+support an essential role for ciliary-localised NIT-GC2 in suppressing
+cPRC calcium following its transient rise at stimulus onset. Overall,
+these knockdown experiments revealed different signalling mechanisms for
+the two NIT-GCs that may be due to their different subcellular
+localisations.
 
 \begin{figure}
-\centering
-\includegraphics[width=1\textwidth,height=\textheight]{../figures/Fig4.png}
-\caption{\textbf{Figure 4. cPRC-specific expression of NIT-GC1 localises
-in close proximity to NOS in the ANS region.} \textbf{(A, B)} Expression
-analysis of the NIT-GC1 (C) and NIT-GC2 (D) gene (magenta) using in situ
-HCR. Co-staining image using acetylated α-tubulin antibody (ac α-tub:
-green). \textbf{(C, D)} Localisation analysis using NIT-GC1 and NIT-GC2
-antibodies. Green shows co-staining with acetylated α-tubulin antibody
-(ac α-tub). \textbf{(E, F)} Changes over time in GCaMP6s intensities in
-cPRC when stimulated with UV in WT (dark yellow) and NOS mutant (E:
-NOSΔ11/Δ11, F: NOSΔ23/Δ23) (light pink) larvae. \textbf{(G, H)} Changes
-over time in GCaMP6s intensities in cPRC when stimulated with UV in
-NIT-GC1 (G: NIT1) and NIT-GC2 (H: NIT2) knockdown by morpholino (MO1 and
-2) larvae. \textbf{(I)} The domain structure of NIT-GC1. Below shows the
-mutant NIT-GC1 (ΔNIT-GC1) with deletion of the NIT domain. Grey squares
-indicate the transmembrane region (TM). \textbf{(J)} A schematic of the
-NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP
-indicator Green cGull, and the NO donor SNAP and DMSO as control. Green
-cGull shows cGMP-dependent green fluorescence. \textbf{(K-N)} The
-intensity changes in cGMP fluorescence over time for the four conditions
-are shown. Gray traces show the optical responses to individual
-intensities (n \textgreater{} 6). The individual intensities are
-normalized (ΔF/F0). The dark line shows the mean value and the band the
-standard error range of the mean. The respective solution was acted on 2
-min after the start of imaging (grey bars).}
+\includegraphics[width=38.89in]{figures/Fig4} \caption{**Figure 4.
+cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.** **(A, B)** Expression analysis of the NIT-GC1 (C) and NIT-GC2 (D) gene (magenta) using in situ HCR.
+Co-staining image using acetylated α-tubulin antibody (ac α-tub: green).
+**(C, D)** Localisation analysis using NIT-GC1 and NIT-GC2 antibodies.
+Green shows co-staining with acetylated α-tubulin antibody (ac α-tub).
+**(E, F)** Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in WT (dark yellow) and NOS mutant (E: NOSΔ11/Δ11, F: NOSΔ23/Δ23) (light pink) larvae.
+**(G, H)** Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in NIT-GC1 (G: NIT1) and NIT-GC2 (H: NIT2) knockdown by morpholino (MO1 and 2) larvae.
+**(I)** The domain structure of NIT-GC1.
+Below shows the mutant NIT-GC1 (ΔNIT-GC1) with deletion of the NIT domain.
+Grey squares indicate the transmembrane region (TM).
+**(J)** A schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control.
+Green cGull shows cGMP-dependent green fluorescence.
+**(K-N)** The intensity changes in cGMP fluorescence over time for the four conditions are shown.
+Gray traces show the optical responses to individual intensities (n > 6).
+The individual intensities are normalized (ΔF/F0).
+The dark line shows the mean value and the band the standard error range of the mean.
+The respective solution was acted on 2 min after the start of imaging (grey bars).}\label{fig:unnamed-chunk-4}
 \end{figure}
 
-\textbf{NIT-GC1 produces cGMP by nitric oxide}
-
-NIT-GC1 has a highly conserved cyclase domain and activation of NIT-GC1
-is expected to produce cGMP \textbf{(Figure 4I)}. Recently, a synthetic
-protein called Green cGull was developed as a cGMP indicator that can be
-used in cultured cell systems (Matsuda et al., 2016). We co-expressed
-Green cGull and NIT-GC1 using the 2A self-cleaving peptide in cultured
-cells COS-7 (monkey kidney) and tested the intracellular dynamics of
-cGMP in the activation of NIT-GC1 \textbf{(Figure 4J)}. When the NO
-donor SNAP was applied to the NIT-GC1 and Green cGull co-expression
-cells, the intracellular cGMP levels was significantly increased
-compared to when NIT-GC1 was not expressed \textbf{(Figure 4K,L)}. The
-increase in fluorescence intensity also did not occur when DMSO was
-applied \textbf{(Figure 4M)}. To confirm whether this increase in
-intracellular cGMP levels is NIT domain-dependent response of NIT-GC1,
-mutant form of NIT-GC1 without the NIT domain was expressed and SNAP was
-applied \textbf{(Figure 4I)}. As a result, the increase in fluorescence
-intensity was significantly reduced \textbf{(Figure 4N)}. These results
-indicate that NIT-GC1 produces cGMP via the NIT domain. In bacteria, it
-has been proposed that the NIT domain regulates cellular functions in
-response to changes in extracellular and intracellular nitrate and
-nitrite concentrations. Our results suggest that NIT-GC1 may be
-sensitive to NO. This could be a possible feedback pathway in which
-NIT-GC1 expressed in cPRC is activated by NO produced by INNOS. In order
-to further explore the role of NO in the cPRC circuit, we therefore
-compared calcium imaging of postsynaptic cells (INNOS, INRGW, Ser-h1 and
-MC) in the cPRC circuit during UV stimulation of cPRC between WT and NOS
-mutant larvae.
-
-\textbf{Depolarisation of cPRC is required for activation of downstream
-interneurons and ciliomotor neurons}
-
-We next quantified the responses of the INNOS and INRGW following a
-UV/violet stimulus to the cPRCs. To unambiguously identify the cells
-from which we recorded calcium signals, we developed an on-slide
-immunostaining method \textbf{(Figure 5---figure supplement 1)}. We used
-the cell-specific neuropeptide markers using anti-RYamide (INNOS) and
-anti-RGWamide (INRGW) antibodies (Conzelmann et al., 2011) to mark the
-interneurons. Based on the position of the nuclei, we could correlate
-live and fixed samples at a single-cell precision \textbf{(Figure
-5A,B)}. In both wild-type and \emph{NOS}-mutant larvae, INNOS cells
-showed increase calcium signals during the inhibition phase of the cPRCs
-\textbf{(Figure 5D)}. In addition, INNOS depolarisation was suppressed
-in NIT-GC2 molphants where cPRC hyperpolarisation was suppressed
-\textbf{(Figure 5E)}. This indicates that cPRC hyperpolarisation is
-required for INNOS depolarisation. In contrast, calcium signals
-increased in INRGW cells in wild-type larvae during the second
-activation phase of cPRCs, a response that was lacking in
-\emph{NOS}-mutants \textbf{(Figure 5F,G)}. This suggests that there is
-no circuit output to the Ser-h1 and MC cells and the ciliary band in
-\emph{NOS}-mutants, consistent with the observed behavioural defects
-\textbf{(Figure 5H,I)}.
+\textbf{NO signalling shapes the dynamics of the cPRC circuit}
+
+To investigate how NO and NIT-GC signalling influence the dynamics of
+the cPRC circuit, we imaged calcium signals from postsynaptic neurons in
+wild type, mutant and morphant larvae. We were able to image the
+activity of all neurons in the cPRC circuit (INNOS, INRGW, Ser-h1 and
+MC). The MC cell was identified based on its position and intrinsic
+activity. To unambiguously identify all other cells from which we
+recorded calcium signals, we developed an on-slide immunostaining method
+\textbf{(Figure 5---figure supplement 1)}. We used the cell-specific
+antibody markers against RYamide (INNOS), RGWamide (INRGW) and serotonin
+(Ser-h1) (Conzelmann et al., 2011) to immunostain agar-embedded larvae
+following calcium imaging. Based on the position of the nuclei, we could
+correlate live and fixed samples at a single-cell precision
+\textbf{(Figure 5A,B)}. Due to the stereotypy of the larvae, we could
+also identify neurons based on their position and activity in
+activity-correlation maps \textbf{(Figure 5C)}.
+
+We first quantified the responses of the INNOS and INRGW interneurons
+during 405 nm stimulation of the cPRCs. In both wild type and
+\emph{NOS}-mutant larvae, INNOS cells showed an increase in calcium
+during stimulation \textbf{(Figure 5D)}. This response was suppressed in
+NIT-GC2 morphant larvae \textbf{(Figure 5E)} revealing an essential role
+for NIT-GC2-mediated cPRC suppression in INNOS activation. In contrast,
+INRGW cells were initially inhibited, followed by a delayed activation
+paralleling the second activation phase of cPRCs. This late INRGW
+response was lacking in \emph{NOS}-mutants \textbf{(Figure 5F,G)}. In
+NIT-GC2 morphants, INRGW cells showed a transient increase in calcium
+that decayed after light off and a delayed activation was not present.
+Next, we imaged calcium signals from Ser-h1 and MC neurons in wild type
+and \emph{NOS} mutant larvae. Ser-h1 cells showed an activation profile
+that correlated with cPRC activity, including a reduction in calcium
+during stimulation followed by rebound, a response that was defective in
+\emph{NOS} mutants. MC cells showed sustained activation, including a
+late-phase that was lacking in \emph{NOS} mutants \textbf{(Figure
+5H,I)}. These data suggest that during 405 stimulation the Ser-h1 cells
+are inhibited and MC cells are activated, and this regulation is
+NO-dependent. This pattern is expected to inhibit ciliary activity in
+the prototroch but not in the other ciliary bands, triggering
+NO-dependent downward swimming.
 
 \begin{figure}
-\centering
-\includegraphics[width=1\textwidth,height=\textheight]{../figures/Fig5.png}
-\caption{\textbf{Figure 5. NO produced by UV stimulation to cPRCs
-induces a depolarising response in cPRC via NIT-GC1.} \textbf{(A, B)}
-Each left figure shows a calcium imaging using GCaMP6s. The respective
-right figures show immunostaining images of the same positions in the
-respective left figures, (A) shows RYamide antibodies, (B) shows
-RGWamide and serotonin (5-HT) antibodies in red and DAPI in cyan.
-Asterisks indicate the nuclei of the cPRC, respectively, and the nuclei
-of the cells that were identified from the arrangement of the
-surrounding cells are indicated by numbers. From the position of the
-cells, the calcium imaging data identify INNOS in (A) and INRGW and
-Ser-h1 in (B), respectively. \textbf{(C)} Correlation (Pearson's r) map
-of neuronal activity of the INNOS, INRGW, Ser-h1 and MC neurons.
-\textbf{(D)} Changes over time in GCaMP6s intensities in INNOS when cPRC
-is stimulated with UV in WT (dark purple) and NOSΔ11/Δ11 mutant (light
-purple) larvae. \textbf{(E)} Changes over time in GCaMP6s intensities in
-INNOS when cPRC is stimulated with UV in NIT-GC2 morphants (MO1 and MO2:
-pink and light pink) larvae. \textbf{(F)} Changes over time in GCaMP6s
-intensities in INRGW when cPRC is stimulated with UV in WT (dark blue)
-and NOSΔ23/Δ23 mutant (light blue) larvae. \textbf{(G)} Changes over
-time in GCaMP6s intensities in INRGW when cPRC is stimulated with UV in
-NIT-GC2 morphants (MO1 and MO2: blue and light blue). \textbf{(H, I)}
-Changes over time in GCaMP6s intensities in Ser-h1 and MC when cPRC is
-stimulated with UV in WT (Ser-h1 and MC: dark red and brown) and
-NOSΔ11/Δ11 mutant (Ser-h1 and MC: light red and brown) larvae.}
+\includegraphics[width=52.08in]{figures/Fig5} \caption{**Figure 5.
+NO produced by UV stimulation to cPRCs induces a depolarising response in cPRC via NIT-GC1.** **(A, B)** Each left figure shows a calcium imaging using GCaMP6s.
+The respective right figures show immunostaining images of the same positions in the respective left figures, (A) shows RYamide antibodies, (B) shows RGWamide and serotonin (5-HT) antibodies in red and DAPI in cyan.
+Asterisks indicate the nuclei of the cPRC, respectively, and the nuclei of the cells that were identified from the arrangement of the surrounding cells are indicated by numbers.
+From the position of the cells, the calcium imaging data identify INNOS in (A) and INRGW and Ser-h1 in (B), respectively.
+**(C)** Correlation (Pearson's r) map of neuronal activity of the INNOS, INRGW, Ser-h1 and MC neurons.
+**(D)** Changes over time in GCaMP6s intensities in INNOS when cPRC is stimulated with UV in WT (dark purple) and NOSΔ11/Δ11 mutant (light purple) larvae.
+**(E)** Changes over time in GCaMP6s intensities in INNOS when cPRC is stimulated with UV in NIT-GC2 morphants (MO1 and MO2: pink and light pink) larvae.
+**(F)** Changes over time in GCaMP6s intensities in INRGW when cPRC is stimulated with UV in WT (dark blue) and NOSΔ23/Δ23 mutant (light blue) larvae.
+**(G)** Changes over time in GCaMP6s intensities in INRGW when cPRC is stimulated with UV in NIT-GC2 morphants (MO1 and MO2: blue and light blue).
+**(H, I)** Changes over time in GCaMP6s intensities in Ser-h1 and MC when cPRC is stimulated with UV in WT (Ser-h1 and MC: dark red and brown) and NOSΔ11/Δ11 mutant (Ser-h1 and MC: light red and brown) larvae.}\label{fig:unnamed-chunk-5}
+\end{figure}
+
+\textbf{Mathematical modelling of cPRC-circuit dynamics}
+
+{[}to be added{]}
+
+\begin{figure}
+\includegraphics[width=25in]{figures/Fig6} \caption{**Figure 6.
+Gene expressions and signalling mechanisms between cPRC, INNOS and INRGWa.** **(C)** Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq.
+The size of the dots is expressed in proportion to the percentage of cells expressing that gene relative to all cells.
+The colours represent the normal logarithm of the number of transcripts in the cells expressing the gene.
+**(D)** Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback.}\label{fig:unnamed-chunk-6}
 \end{figure}
@@
 
-In the present study, we explored the role of NO in the UV avoidance
-behaviour of Platynereis. We demonstrated that there are interneurons
-expressing NOS in the cPRC circuit and that NO is required for UV
-avoidance in behavioural experiments using NOS mutants. We identified
-the axon-localised type NIT-GC1 and the cilia-localised type NIT-GC2
-expressed in the cPRCs as targets of NO and showed from cultured cell
-experiments that these produce cGMP in response to NO. We have shown by
-calcium imaging that INNOSs are activated in response to the
-hyperpolarising response of the cPRCs, that the NO produced feeds back
-to the cPRC via NIT-GC1 and induces a depolarising response in the cPRC,
-and that the depolarising response of the cPRC is involved in activating
-downstream INRGW. Lastly, we confirmed that RGWamide induces the
-downward swimming behaviour of larvae.
-
-\begin{figure}
-\centering
-\includegraphics[width=1\textwidth,height=\textheight]{../figures/Fig6.png}
-\caption{\textbf{Figure 6. Gene expressions and signalling mechanisms
-between cPRC, INNOS and INRGWa.} \textbf{(C)} Dot plot of genes
-(columns) expressed in three types of cells (rows) in the cPRC circuit
-using single cell RNA-Seq. The size of the dots is expressed in
-proportion to the percentage of cells expressing that gene relative to
-all cells. The colours represent the normal logarithm of the number of
-transcripts in the cells expressing the gene. \textbf{(D)} Schematic
-diagram of the signalling pathway of the cPRC circuit, focusing on the
-NO feedback.}
-\end{figure}
+The activity of synaptic circuits is richly modulated monoamines,
+neuropeptides and gaseous molecules. Our work revealed an essential role
+for NO-mediated signalling in driving UV/violet avoidance behaviour. NO,
+produced by postsynaptic interneurons, feeds back to their presynaptic
+photoreceptor partners leading to their delayed activation. This
+late-phase activation of cPRCs drives circuit output through projection
+interneurons and ciliomotor neurons. In the cPRC circuit, synaptic
+connectivity alone is not sufficient to drive circuit dynamics and
+behavioural change.
+
+We identified cPRC-expressed NIT-GC1 as a new type of NO sensor. NIT-GC1
+is localised to the axonal projections of the cPRCs in the anterior
+neurosecretory plexus of the brain. Here, cPRC axons intermingle and
+synapse on INNOS neurites that contain NOS. NO-mediated feedback
+signalling thus likely occurs in the neurosecretory plexus where NOS-
+and NIT-GC1-containing projections are in close proximity and where we
+could detect NO production following UV stimulation.
+
+We showed that NIT-GC1 can produce cGMP in a NO-dependent manner. In
+cPRCs, increased cGMP levels could open CNGAα, a cyclic nucleotide-gated
+channels specifically expressed in cPRC (Tosches et al., 2014), leading
+to cPRC delayed depolarisation, even after UV off.
+
+NO is a free radical with a half-life of 5 msec in vivo and a signalling
+range limited to a few tens of nanometers. In neuronal signalling, NOS
+and sGC have been found to localise at close proximity at synapses
+{[}Burette et al. (2002){]}(Garthwaite, 2015). We identified 12 sGCs in
+\emph{Platynereis}, but none of these is expressed in the cPRC. Instead,
+we identified an unconventional NIT-domain containing GC, NIT-GC1 as the
+mediator of NO retrograde signalling. The NIT domain was first
+identified in bacteria and animal NIT-GCs have only been recently
+reported (Moroz et al., 2020; Shu et al., 2003). Bacterial NIT domains
+regulate cellular functions in response to changes in extracellular and
+intracellular nitrate and/or nitrite concentrations (Camargo et al.,
+2007). NO is readily converted to nitrate and nitrite (Garthwaite, 2015;
+Möller et al., 2019; Santos et al., 2011) and these molecules accumulate
+in placozoans and cnidarians in cells and tissues with high NOS activity
+(Moroz et al., 2020, 2004). It is possible that the NIT domains in
+NIT-GCs may also sense nitrate and nitrite, as in bacteria. Furthermore,
+if they show different sensitivities to NO, nitrite and nitrate, it is
+possible that differences in their half-lives could give rise to a range
+of activation timings (Lundberg et al., 2011). We have found 15 NIT-GCs
+in \emph{Platynereis}, suggesting a wide range of functions. In
+addition, NIT-GC1 and NIT-GC2 showed very different subcellular
+localisation and functions. It is not clear whether the functional
+differences are due to the different localisation or also to differences
+in biochemical function. Overall, the diversification and spatially
+different localisation patterns of NIT-GCs may increase their signalling
+repertoire. Our data revealed key roles for two \emph{Platynereis}
+NIT-GCs in phototraonsduction and NO-mediated neuromodulation suggesting
+that these molecules contribute to the diversity of neuronal signalling.
+Sponges and ctenophores, where NO signalling is present, lack NIT-GCs.
+In placozoans as many as 12 have been found, and cnidarians also contain
+a large diversity. NIT-GCs have been lost from chordates {[}check{]}.
 
@@
 shown that the cells that strongly synapse with cPRCs are interneurons
-that specifically express NOS (INNOS). The fluorescence intensity of
-DAF-FM, which specifically detected INNOS, showed a transient tonic-like
-increase dependent on UV stimulation of the cPRCs. The graph shows a
-decrease in intensity after UV light is switched off, but the response
-of DAF-FM to NO is irreversible and this decrease is due to breaching by
-the relatively strong excitation light. Attempts to measure changes in
-the intensity of the calcium indicator RGECO1a simultaneously with
-DAF-FM imaging did not lead to detection of changes in fluorescence
-intensity compared to GCaMP6s and could only visualise the cPRCs. In
-calcium imaging using GCaMP6s, INNOS identified by the positional
-relationship between the cPRC and INNOS and the labelling of RYa
-depolarised in clear synchrony with the hyperpolarising response of the
-cPRCs. This tonic response may involve dopaminergic nerves based on
-single-cell data \textbf{(Figure 6)} {[}Puopolo et al. (2001){]}(Zhang
-et al., 2007)(Randel et al., 2014). In the mammalian retina, it is known
-that signals from bipolar photoreceptor cells are transmitted to
-downstream NOS-expressing amacrine cells (NOAC), where NO is produced
-(Jacoby et al., 2018). These suggest that UV stimulation of cPRCs causes
-the production of NO via a depolarising response of INNOS.
-
-Our results suggest that NIT-GC1, which is specifically expressed in
-cPRCs, functions as a target for NO produced by INNOS and contributes to
-the regulation of the cPRC circuit transmission pathway via a
-depolarising response by increasing cGMP. In the mammalian retina, NO is
-an essential signalling molecule and one of the most well-studied
-(Cudeiro and Rivadulla, 1999). It has been reported that NO produced in
-one of the retinal cells modulates cGMP-mediated neurotransmission by
-activating presynaptic neural sGCs (general target of NO) through
-retrograde signalling {[}Vielma et al. (2014){]}(Wei et al., 2012). In
-Platynereis cPRC, NIT-GC1 was also identified as having the property of
-producing cGMP in a NO-dependent manner. Furthermore, it is known that
-CNGAα, one of the cyclic nucleotide-gated channels, is specifically
-highly expressed in cPRC (Tosches et al., 2014), and in INNOS-cPRC
-signalling, NO is implicated in the neuromodulation of the cPRC by
-retrograde signalling. NO is a molecule with a very short half-life, and
-when functioning as a signalling molecule, NOS and sGC are known to
-localise very close together at synapses between neurons {[}Burette et
-al. (2002){]}(Garthwaite, 2015). In our study, we found 12 sGCs in
-Platynereis, but none expressed in the cPRC. Localisation analysis of
-NOS and NIT-GC1 by antibody staining showed that the two proteins
-localised very close together in the anterior neurosecretory plexus
-region. Calcium imaging results using NOS mutant and NIT-GC1 knockdown
-larvae suggest that NIT-GC1 activation induces a depolarising response
-in the cPRC.
+that specifically express NOS (INNOS).
+
+{[}Puopolo et al. (2001){]}(Zhang et al., 2007)(Randel et al., 2014). In
+the mammalian retina, it is known that signals from bipolar
+photoreceptor cells are transmitted to downstream NOS-expressing
+amacrine cells (NOAC), where NO is produced (Jacoby et al., 2018). These
+suggest that UV stimulation of cPRCs causes the production of NO via a
+depolarising response of INNOS.
+
+In the mammalian retina, NO is an essential signalling molecule and one
+of the most well-studied (Cudeiro and Rivadulla, 1999). It has been
+reported that NO produced in one of the retinal cells modulates
+cGMP-mediated neurotransmission by activating presynaptic neural sGCs
+(general target of NO) through retrograde signalling {[}Vielma et al.
+(2014){]}(Wei et al., 2012).
 
@@
 The present results suggest that the depolarising response of the cPRC
-due to NO feedback induces INRGW activation. RGWamide is also a
-neuropeptide that induces larval downward swimming quite strongly
-(unpublished? need to show data?), and the delayed downward swimming of
-NOS mutant larvae in response to UV stimulation may be responsible for
-the failure of INRGW activation to occur. In addition, in many marine
-invertebrates, NO has been reported to be involved in many aspects of
-settlement behaviour and metamorphosis induction (Bishop and Brandhorst,
-2007, 2003; Castellano et al., 2014; Leise et al., 2001; Pechenik et
-al., 2007; Song et al., 2021; Ueda et al., 2016; Ueda and Degnan, 2014,
-2013; Yang et al., 2018; Zhu et al., 2020). The downward swimming
-behaviour of Platynereis larvae is the first step in the onset of
-settlement behaviour and metamorphosis. These suggest that downward
-swimming due to NO activation of INRGWs may trigger the onset of
-settlement behaviour and metamorphosis {[}Conzelmann et al.
-(2013b){]}(Williams et al., 2015). These suggest that NO feedback
-signalling may act as a trigger for downward swimming via activation of
-INRGWs, which induces the initiation of settlement behaviour and
-metamorphosis.
-
-An outstanding question is how the NO\textasciitilde NIT-GC response
-alters neuronal dynamics. NO is a single molecule and, due to its short
-half-life, cannot be transmitted over long distances and has a limited
-signalling repertoire. It has therefore been considered unsuitable for
-generating diversity in signalling compared to other signalling
-molecules such as neuropeptides and classical neurotransmitters (Jékely,
-2021). However, our results in the present study indicate that
-NO-induced activation of NIT-GC1 is involved in the regulation of
-neuronal circuits (cPRC pathway). The NIT domain was originally found in
-bacteria and the existence of GCs with NIT domains was first reported by
-Moroz et al in 2020 (Moroz et al., 2020; Shu et al., 2003). It has been
-proposed that NIT domains in bacteria regulate cellular functions in
-response to changes in extracellular and intracellular nitrate and/or
-nitrite concentrations (Camargo et al., 2007). In fact, NO is a free
-radical molecule, with a half-life of 5 msec in vivo, and is readily
-converted to nitrate and nitrite (Garthwaite, 2015; Möller et al., 2019;
-Santos et al., 2011). And nitrate and nitrite have been reported to
-accumulate in cells and tissues with high NOS activity in placozoans and
-cnidarians (Moroz et al., 2020, 2004). It is possible that the NIT
-domains of NIT-GC1 identified in this study may also be used for
-nitrate/nitrite sensing, as in bacteria. Furthermore, if they show
-different sensitivities to NO, nitrite and nitrate, it is possible that
-the temporal variation in their respective half-lives could give rise to
-a repertoire of activation timings (Lundberg et al., 2011). In addition,
-we have found at least 15 NIT-GCs in the Platynereis transcriptome data,
-of which NIT-GC2 is highly expressed in the cPRC, in addition to
-NIT-GC1. Interestingly, NIT-GC2 showed a very different localisation
-pattern to NIT-GC1, and furthermore, the results of calcium imaging of
-cPRC in response to UV stimulation using NIT-GC2 knockdown larvae were
-the different dynamics (such that hyperpolarisation was suppressed) with
-NIT-GC1 knockdown. It was not clear from the present analysis whether
-this difference was due to differences in the localisation or the
-property of the NIT-GCs. However, the diversification and spatially
-different localisation patterns of NIT-GCs may increase their signalling
-repertoire. We also believe that the appearance of NIT-GCs was one of
-the means to generate the diversity of signalling pathways using NO
-during the early stages of nervous system evolution in metazoans, since
-sponges and ctenophores, where NO signalling is present, have no NIT-GCs
-at all, as many as 12 have been found in placozoans, and NIT-GCs are
-functional in the ``proto-eye'' circuit of Platynereis.
+due to NO feedback induces INRGW activation.
+
+In addition, in many marine invertebrates, NO has been reported to be
+involved in many aspects of settlement behaviour and metamorphosis
+induction (Bishop and Brandhorst, 2007, 2003; Castellano et al., 2014;
+Leise et al., 2001; Pechenik et al., 2007; Song et al., 2021; Ueda et
+al., 2016; Ueda and Degnan, 2014, 2013; Yang et al., 2018; Zhu et al.,
+2020). The downward swimming behaviour of Platynereis larvae is the
+first step in the onset of settlement behaviour and metamorphosis. These
+suggest that downward swimming due to NO activation of INRGWs may
+trigger the onset of settlement behaviour and metamorphosis
+{[}Conzelmann et al. (2013b){]}(Williams et al., 2015). These suggest
+that NO feedback signalling may act as a trigger for downward swimming
+via activation of INRGWs, which induces the initiation of settlement
+behaviour and metamorphosis.
+
+\hypertarget{key-resources-table}{%
+\subsection{Key resources table}\label{key-resources-table}}
+
+Reagent type (species)
+
+Designation
+
+Source or reference
+
+Identifiers
+
+Additional information
+
+Strain (Platynereis dumerilii)
+
+NOSΔ11/Δ11~knockout
+
+~This paper
+
+NA
+
+~Knockout generated by CRISPR/Cas-9-induced gene editing
+
+Strain (Platynereis dumerilii)
+
+NOSΔ23/Δ23~knockout
+
+~This paper
+
+NA
+
+~Knockout generated by CRISPR/Cas-9-induced gene editing
+
+Strain (Platynereis dumerilii)
+
+NOSΔ11/Δ23~knockout
+
+~This paper
+
+NA
+
+~Knockout generated by CRISPR/Cas-9-induced gene editing
+
+Cell line (Cercopithecus aethiops)
+
+COS-7 cell
+
+NA
+
+RRID:CVCL\_0224
+
+Angio-proteomie (CAT no. cAP-0203)??
+
+Biological sample (Platynereis dumerilii)
+
+Wild type Tübingen strain
+
+Other
+
+NCBITaxon:6359
+
+Jékely lab strain (Tübingen, Exeter)
+
+Gene (Platynereis dumerilii)
+
+NOS
+
+This paper
+
+GenBank\_Acc\#:
+
+NA
+
+Gene (Platynereis dumerilii)
+
+NIT-GC1
+
+This paper
+
+GenBank\_Acc\#:
+
+NA
+
+Gene (Platynereis dumerilii)
+
+NIT-GC2
+
+This paper
+
+GenBank\_Acc\#:
+
+NA
+
+NOS: Nitric\_Oxide\_Synthase
+
+To amplify Promoter \& Regulatory region
+
+Fwd
+
+NOSProm2ndF0.6BamHI
+
+AGGGATCCCCCAATGCTTTAGCAGTCAGAGGAG
+
+NOS: Nitric\_Oxide\_Synthase
+
+To amplify Promoter \& Regulatory region
+
+Rev
+
+GeR1ASCI
+
+AAGGCGCGCCCCACCACCACCTTTGATATCCATGATGCTCACTTCGC
+
+NOS: Nitric\_Oxide\_Synthase
+
+Mutation Check on Exon3
+
+Fwd
+
+Exon3 Sequence F-27bp
+
+GGTTCATTGGTTTCGATAACATTGCGG
+
+NOS: Nitric\_Oxide\_Synthase
+
+Mutation Check on Exon3
+
+Rev
+
+Exon3 Sequence R-27bp
+
+CAGAGTCGATCAGTCTGCATATCTCCA
+
+NOS: Nitric\_Oxide\_Synthase
+
+Sequencing primer for Exon3 mutation check PCR product
+
+Fwd
+
+Exon3 Sequence F-2
+
+GGTGCTCTCCCGGGTACACAA
+
+RNA
+
+sgRNA
+
+NA
+
+NA
+
+5'-TAGGGCAATACTGGCTCCACTC-3'
+
+RNA
+
+sgRNA
+
+NA
+
+NA
+
+5'-AAACGAGTGGAGCCAGTATTGC-3'
+
+RNA
+
+pUC57-T7-RPP2-hSpCas9- HA-2XNLS-GFP
+
+NA
+
+plasmid
+
+Bezares-Calderón et al., 2018
+
+Antibody
+
+Monoclonal Anti-Tubulin, Acetylated antibody
+
+Sigma-Aldrich
+
+Cat\#:T6793, RRID:AB\_477585
+
+NA
+
+Antibody
+
+HA-Tag (C29F4), Rabbit mAb
+
+Cell Signaling Technology
+
+Cat\#:3724P
+
+NA
+
+NIT-GC1 polychronal antibodies
+
+CYWLLGRKERRPKRRL-amide
+
+This paper
+
+rabbit
+
+Altabioscience
+
+NIT-GC2 polychronal antibodies
+
+CTEGSTKEGKKEGQ-amide
+
+This paper
+
+rabbit
+
+Altabioscience
+
+NOS polychronal antibodies
+
+CKPSYELQDPWKTYIWRKD-amide
+
+This paper
+
+Rat
+
+Altabioscience
+
+Antibody
+
+RYamide neuropeptide antibodies
+
+CRY-amide
+
+rabbit
+
+Conzelmann and Jékely, 2012
+
+Antibody
+
+RGWamide neuropeptide antibodies
+
+CGW-amide
+
+rabbit
+
+Conzelmann and Jékely, 2012
+
+Antibody
+
+F(ab')2-Goat anti-Rabbit IgG (H+L) Cross-Adsorbed Secondary Antibody,
+Alexa Fluor™ 546
+
+Invitrogen
+
+Catalog \# A-11071
+
+NA
+
+Antibody
+
+Goat anti-Rat IgG (H+L) Highly Cross-Adsorbed Secondary Antibody, Alexa
+Fluor™ Plus 594
+
+Invitrogen
+
+Catalog \# A48264
+
+NA
+
+Antibody
+
+F(ab')2-Goat anti-Mouse IgG (H+L) Cross-Adsorbed Secondary Antibody,
+Alexa Fluor™ 647
+
+Invitrogen
+
+Catalog \# A-21237
+
+NA
+
+Recombinant DNA reagent
+
+NIT-GC1 full
+
+comp411593\_c0\_seq1\_309\_F
+
+NA
+
+GGTTGAATAATGACAAGCAAGGAGA
+
+Recombinant DNA reagent
+
+NIT-GC1 full
+
+comp411593\_c0\_seq1\_2717\_R
+
+NA
+
+GTGCTATCATTTCCAGGTAAATACCC
+
+Recombinant DNA reagent
+
+NIT-GC2 full
+
+Contig2280\_66\_F
+
+NA
+
+AATATCTAGCGAAGGAGAACACCTCTCTTC
+
+Recombinant DNA reagent
+
+NIT-GC2 full
+
+Contig2280\_2763\_R
+
+NA
+
+ATGGCCAGTAATAAACCATCAGTGGTTC
+
+Recombinant DNA reagent
+
+pcDNA3.1(+) vector
+
+Invitrogen
+
+Catalog Number: V79020
+
+NA
+
+Recombinant DNA reagent
+
+Inverse PCR for the insert region of pcDNA3.1(+)
+
+pcDNA3.1(+)\_inv\_NheI\_fwd
+
+NA
+
+CGTTTAAACTTAAGCTTGGTACCGAG
+
+Recombinant DNA reagent
+
+Inverse PCR for the insert region of pcDNA3.1(+)
+
+pcDNA3.1(+)\_inv\_NheI\_rev
+
+NA
+
+CCAGCTTGGGTCTCCCTATAGT
+
+Recombinant DNA reagent
+
+kozak-NITGC2-T2A-Green cGull
+
+NITGC-T2A-fwd
+
+NA
+
+TATAGGGAGACCCAAGCTGGGCCACCATGACCCAGATG
+
+Recombinant DNA reagent
+
+kozak-NITGC2-T2A-Green cGull
+
+NITGC-T2A-rev1
+
+NA
+
+GCATGTTAGAAGACTTCCTCTGCCCTCATAATCAAACCCCTCTCT
+
+Recombinant DNA reagent
+
+kozak-NITGC2-T2A-Green cGull
+
+NITGC-T2A-rev2
+
+NA
+
+AGGGCCGGGATTCTCCTCCACGTCACCGCATGTTAGAAGACTTCC
+
+Recombinant DNA reagent
+
+kozak-NITGC2-T2A-Green cGull
+
+NITGC-T2A-rev3
+
+NA
+
+TGCTCACCATAGGGCCGGGATTCTCCTC
+
+Recombinant DNA reagent
+
+kozak-NITGC2-T2A-Green cGull
+
+cGull-fwd
+
+NA
+
+TCCCGGCCCTATGGTGAGCAAGGGCGAG
+
+Recombinant DNA reagent
+
+kozak-NITGC2-T2A-Green cGull
+
+cGull-rev
+
+NA
+
+ACCAAGCTTAAGTTTAAACGTTACTTGTACAGCTCGTCCATG
+
+Recombinant DNA reagent
+
+Green cGull-T2A-NITGC2 \& NITGC2-T2A-Green cGull
+
+NITGC2\_seq\_743\_fwd
+
+NA
+
+AGCCATCTACGAGTGGTA
+
+Recombinant DNA reagent
+
+Green cGull-T2A-NITGC2 \& NITGC2-T2A-Green cGull
+
+cGull\_seq\_385\_rev
+
+NA
+
+TGCCCTTCAGCTCGATG
+
+Recombinant DNA reagent
+
+Green cGull-T2A-NITGC2 \& NITGC2-T2A-Green cGull
+
+NITGC2\_seq\_743\_rev
+
+NA
+
+TGACTGACGAACCCTCC
+
+Recombinant DNA reagent
+
+Green cGull-T2A-NITGC2 \& NITGC2-T2A-Green cGull
+
+NITGC2\_seq\_394\_fwd
+
+NA
+
+AGATATCTTGAAACGGACGA
+
+Recombinant DNA reagent
+
+kozak-NIT1(seq)-T2A-Green cGull
+
+2A-cGull\_invF\_L1
+
+NA
+
+TGACGTGGAGGAGAATCCCGGCCCTATGGTGAGCAAGGGCGAGGAGCTGT
+
+Recombinant DNA reagent
+
+kozak-NIT1(seq)-T2A-Green cGull
+
+2A-cGull\_invF\_L2
+
+NA
+
+GCAGAGGAAGTCTTCTAACATGCGGTGACGTGGAGGAGAATCCCGGCCCT
+
+Recombinant DNA reagent
+
+kozak-NIT1(seq)-T2A-Green cGull
+
+2A-cGull\_invR\_L
+
+NA
+
+TCCTTGCTTGTCATGGTGGCCCAGCTTGGGTCTCCCTATAGTGAGTCGTA
+
+Recombinant DNA reagent
+
+kozak-NIT1(seq)-T2A-Green cGull
+
+NIT1-2A\_F\_L
+
+NA
+
+GAGACCCAAGCTGGGCCACCATGACAAGCAAGGAGATGCCTGTACTCATG
+
+Recombinant DNA reagent
+
+kozak-NIT1(seq)-T2A-Green cGull
+
+NIT1-2A\_R\_L
+
+NA
+
+TGTTAGAAGACTTCCTCTGCCCTCTATGACTTTTTCTATGCTTTCTTCGG
+
+Recombinant DNA reagent
+
+kozak-NIT1(seq)-T2A-Green cGull
+
+NIT1seq\_remo\_invF2
+
+NA
+
+AGCGTGGAGGTGGGCCTAGACGAAAGAGCTGAAAA
+
+Recombinant DNA reagent
+
+kozak-NIT1(seq)-T2A-Green cGull
+
+NIT1seq\_remo\_invR2
+
+NA
+
+AGCTCTTTCGTCTAGGCCCACCTCCACGCTGAATA
+
+Recombinant DNA reagent
+
+kozak-NIT1(seq)-T2A-Green cGull
+
+NIT1seq\_remo\_F2
+
+NA
+
+GCCGGTCTTGTCGATCAGGATGATCTGGAC
+
+Recombinant DNA reagent
+
+kozak-NIT1(seq)-T2A-Green cGull
+
+NIT1seq\_remo\_R2
+
+NA
+
+GTCCAGATCATCCTGATCGACAAGACCGGC
+
+Recombinant DNA reagent
+
+pUC57-NOSp::Palmi-3xHA-tdTomato (plasmid)
+
+This paper
+
+NA
+
+Promoter construct: injected at 250 ng/μl
+
+Recombinant DNA reagent
+
+pUC57-T7-RPP2-tdTomato-P2A-GCaMP6 (plasmid)
+
+This paper
+
+NA
+
+Used for generating tdTomato-P2A-GCaMP6s mRNA
+
+Plasmid
+
+Green cGull
+
+Addgene
+
+Plasmid \#86867
+
+NA
+
+HCR
+
+NOS
+
+Integrated DNA Technologies
+
+NA
+
+NA
+
+HCR
+
+NIT-GC1
+
+Integrated DNA Technologies
+
+NA
+
+NA
+
+HCR
+
+NIT-GC2
+
+Integrated DNA Technologies
+
+NA
+
+NA
+
+HCR
+
+RYa-pNP (GenBank accession: JF811330.1)
+
+Integrated DNA Technologies
+
+NA
+
+NA
+
+HCR
+
+c-opsin1 (GenBank accession: AY692353.1)
+
+Integrated DNA Technologies
+
+NA
+
+NA
+
+HCR
+
+MLD/pedal2-pNP (GenBank accession: KF515945.1)
+
+Integrated DNA Technologies
+
+NA
+
+NA
+
+HCR
+
+CNGAα (GenBank accession: KM199644.1)
+
+Integrated DNA Technologies
+
+NA
+
+NA
+
+fluorescently labeled hairpins
+
+B2-647
+
+Molecular Technologies
+
+NA
+
+NA
+
+fluorescently labeled hairpins
+
+B3-546
+
+Molecular Technologies
+
+NA
+
+NA
+
+morpholino
+
+NIT-GC1 MO1
+
+Gene-Tools, LLC
+
+NA
+
+TGCTTGTCATTATTCAACCAGCAAA
+
+morpholino
+
+NIT-GC1 MO2
+
+Gene-Tools, LLC
+
+NA
+
+TTCAATTAAACCCTCCAGGTTGCTG
+
+morpholino
+
+NIT-GC2 MO1
+
+Gene-Tools, LLC
+
+NA
+
+AAATGAAGAGAGGTGTTCTCCTTCG
+
+morpholino
+
+NIT-GC2 MO2
+
+Gene-Tools, LLC
+
+NA
+
+ATATTCATTATGTGAAGAACTTCCA
+
+plasmid for mRNA synthase (GCaMP6s)
+
+BamHI-T7::RPP2(5UTR)-GCaMP6s(AscI-AgeI)-polyA\_KpnI\_c1
+
+NA
+
+Plasmid
+
+Bezares-Calderón et al., 2018
+
+plasmid for mRNA synthase (RGECO1a)
+
+PUC57-T7-PduRPP2(5UTR)-jRGECO1a
+
+NA
+
+Plasmid
+
+Bezares-Calderón et al., 2018
+
+Chemical compound, drug
+
+SNAP
+
+Sigma-Aldrich
+
+Cat\#:M9020
+
+500 μM
+
+Chemical compound, drug
+
+L-NAME
+
+Sigma-Aldrich
+
+Cat\#:M9021
+
+501 μM
+
+Commercial assay or kit
+
+Phusion Human Specimen Direct PCR Kit
+
+Thermofisher
+
+NA
+
+NA
+
+Commercial assay or kit
+
+mMESSAGE mMACHINE Sp6 kit
+
+Thermofisher
+
+NA
+
+NA
+
+Software, algorithm
+
+Golden Gate TAL
+
+Addgene 1000000024
+
+NA
+
+NA
+
+Software, algorithm
+
+Effector Kit 2.0, Fiji perl and Fiji scripts for tracking
+
+PMID: 22743772, \url{https://github.com/JekelyLab/Veraszto_et_al_2018}
+
+RRID:SCR\_002285
+
+0000d2a
+
+Commercial assay or kit
+
+QuickExtract
+
+Epicentre,US
+
+Cat\#:QE09050
+
+NA
+
+Commercial assay or kit
+
+MEGAshortscript T7 Transcription Kit
+
+Ambion, ThermoFisher Scientific
+
+Cat\#:AM1354
+
+NA
+
+Commercial assay or kit
+
+mMESSAGE mMACHINE T7 ULTRA Transcription Kit
+
+Ambion, ThermoFisher Scientific
+
+Cat\#:AM1345
+
+NA
+
+Commercial assay or kit
+
+MEGAclear Transcription Clean-Up Kit
+
+Ambion, ThermoFisher Scientific
+
+Cat\#:AM1908
+
+NA
+
+Software, algorithm
+
+Fiji
+
+NIH
+
+RRID:SCR\_002285
+
+NA
+
+Software, algorithm
+
+R Project for Statistical Computing
+
+R Foundation
+
+RRID:SCR\_001905
+
+NA
+
+Software, algorithm
+
+Imaris Version 8.0.0
+
+Bitplane, UK.
+
+RRID:SCR\_007370
+
+NA
+
+Software, algorithm
+
+CATMAID
+
+\url{DOI:10.1093/bioinformatics/btp266}
+
+RRID:SCR\_006278
+
+NA
+
+Software, algorithm
+
+PhyML
+
+\url{DOI:10.1093/sysbio/syq010}
+
+RRID:SCR\_014629
+
+NA
+
+Software, algorithm
+
+Gblocks
+
+\url{DOI:10.1080/10635150701472100}
+
+RRID:SCR\_015945
+
+NA
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Juxe9kely2021}{}}%
-Jékely G. 2021. The chemical brain hypothesis for the origin of nervous
-systems. \emph{Philosophical Transactions of the Royal Society B:
-Biological Sciences} \textbf{376}:20190761.
-doi:\href{https://doi.org/10.1098/rstb.2019.0761}{10.1098/rstb.2019.0761}
-
 \leavevmode\vadjust pre{\hypertarget{ref-Kuehn2022}{}}%
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Tosches2014}{}}%
-Tosches M, Bucher D, Vopalensky P, Arendt D. 2014. Melatonin Signaling
+Tosches MA, Bucher D, Vopalensky P, Arendt D. 2014. Melatonin Signaling
 Controls Circadian Swimming Behavior in Marine Zooplankton. \emph{Cell}
@@
 
+\hypertarget{figure-supplements}{%
+\subsection{Figure supplements}\label{figure-supplements}}
+
+\begin{figure}
+\includegraphics[width=32.64in]{figures/Fig1_sup1} \caption{**Figure 1 -- figure supplement 1**  Phylogenetic tree of NOS by maximum likelihood.
+Tree robustness was tested with 1000 replicates of ultrafast bootstrap with the aLRT-SH-like and aBayes methods.}\label{fig:unnamed-chunk-8}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=16.67in]{figures/Fig1_sup2} \caption{**Figure 1 -- figure supplement 2** Expression analysis of the *NOS* gene (magenta) using in situ HCR using the larva at three-day-old in dorsal **(A)** and anterior view (B).
+(B) Insert: showing two NOS cells (INNOS_dl and INNOS_vl) on left side.}\label{fig:unnamed-chunk-9}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=34.72in]{figures/Fig1_sup3} \caption{**Figure 1 -- figure supplement 3** }\label{fig:unnamed-chunk-10}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=44.44in]{figures/Fig3_sup1} \caption{**Figure 3 -- figure supplement 1** **(A)** Overlaid trajectories for WT (n=37) and NOS mutant (NOSΔ11/Δ11, n=18 and NOSΔ23/Δ23, n=8) at two-day-old larvae.
+0 sec as the starting point.
+After 10 sec, UV (395 nm) stimulation from the side.
+**(B)** The temporal changes in the vertical position of the WT and mutant 2-days-old larvae before and after UV stimulation are shown.
+The starting points of each larval trajectory are set to 0.
+After UV stimulation is indicated by purple squares.
+**(C)** Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23) larvae at 2-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
+The data are shown in 30 s bins.
+**(D, E)** The temporal changes in the distance traveled of the WT and mutant in 2-day (D) and 3-day-old (E) larvae before and after UV stimulation are shown.}\label{fig:unnamed-chunk-11}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=33.33in]{figures/Fig4_sup1} \caption{**Figure 4 -- figure supplement 1** }\label{fig:unnamed-chunk-12}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=20.83in]{figures/Fig4_sup2} \caption{**Figure 4 -- figure supplement 2** Phylogenetic tree of guanylate cyclase by maximum likelihood (ML).
+Guanylate cyclase-coupled receptor and soluble guanylate cyclases (sGC) as outgroups.
+Guanylate cyclases with NIT domains are found in most animal phyla except Porifera, Ctenophora, Urochordata and Chordata.
+Dot plot of Platynereis NIT-GC genes (columns) expressed in cPRC, INNOS and INRGWa (rows) using single cell RNA-Seq.
+The size of the dots is expressed in proportion to the percentage of cells expressing that gene relative to all cells.
+The colours represent the normal logarithm of the number of transcripts in the cells expressing the gene.}\label{fig:unnamed-chunk-13}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=33.33in]{figures/Fig4_sup3} \caption{**Figure 4 -- figure supplement 3** Co-expression analysis image of the NIT-GC1 (magenta) and MLD-pedal2 amide proneuropeptide gene (MLD: green).
+Anterior view of the larva at two-day-old.
+**(B, C)** Expression analysis of the NIT-GC2 gene (magenta) using in situ HCR.
+Anterior (B) and posterior (C) views of the larva at three-day-old.
+**(D)** Co-localisation analysis using NIT-GC1 (magenta) and NOS (green) antibodies.
+Anterior view of the larva at two-day-old.
+**(E, F)** Localisation analysis using NIT-GC1 and NIT-GC2 antibodies for NIT-GC1 (E) and NIT-GC2 (F) morphant.
+Green shows co-staining with acetylated α-tubulin antibody (acTub).
+Anterior view of the larva at two-day-old.}\label{fig:unnamed-chunk-14}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=27.78in]{figures/Fig5_sup1} \caption{**Figure 5 -- figure supplement 1** Schematic diagram of the immunostaining procedure after calcium imaging.}\label{fig:unnamed-chunk-15}
+\end{figure}
+
 \end{document}

### 2023-05-14 Manuscript/Manuscript.Rmd
edit manuscript

@@
 In two- and three-day-old larvae, we detected *NOS* expression in four cells (two of them weakly expressing) in the apical organ region **(Figure 1D and Figure 1---figure supplement 2)**.
-*NOS* was also expressed in the region of the visual eyes (adult eyes) and the pigmented eyespots **(Figure 1---figure supplement 1)**.
+*NOS* was also expressed in the region of the visual eyes (adult eyes) and the pigmented eyespots **(Figure 1---figure supplement 2)**.
 The four apical organ cells, but not the eyes, were also labelled with a *NOS* reporter construct driving palmitoylated-tdTomato **(Figure 1E)**.
@@
 Views of larvae at three-day-old from the anterior side.
-Insert: One of the reconstructed INNOS is shown from the anterior side.
 **(F)** Wiring diagram of the cPRC circuit.
@@
 To achieve this, we generated two *Platynereis* *NOS* knockout lines with the CRISPR/Cas9 method.
-We recovered two deletions (NOSΔ11/Δ11 and Δ23/Δ23), both leading to frame-shift and an early stop codon and thus likely representing null alleles **(Figure 3A)**.
+We recovered two deletions (NOSΔ11/Δ11 and Δ23/Δ23), both leading to frame-shift and an early stop codon and thus likely representing null alleles **(Figure 3---figure supplement 1A)**.
 We could establish a homozygous line for both mutations indicating that NOS is not an essential gene.
-To quantify UV avoidance, we recorded the trajectories of freely swimming wild type and mutant larvae in vertical columns, illuminated laterally from two opposite sides with 395 nm UV light **(Figure 3B, C and Figure 3---figure supplement 1A-D)**.
+To quantify UV avoidance, we recorded the trajectories of freely swimming wild type and mutant larvae in vertical columns, illuminated laterally from two opposite sides with 395 nm UV light **(Figure 3A and Figure 3---figure supplement 1B)**.
 As previously shown, wild-type larvae swim downward following non-directional UV/violet light stimulation [@Veraszto2018].
-In contrast, both two- and three-day-old homozygous *NOS*-mutant larvae showed a strongly diminished UV-avoidance response **(Figure 3B-D and Figure 3---figure supplement 1)**.
+In contrast, both two- and three-day-old homozygous *NOS*-mutant larvae showed a strongly diminished UV-avoidance response **(Figure 3A, B and Figure 3---figure supplement 1B,C)**.
 This phenotype is similar to the defective UV-avoidance of *c-opsin1* mutant larvae and reveals a requirement for NOS in UV-avoidance behaviour [@Veraszto2018].
-Wild type but not mutant larvae also showed an increase in swimming speed under UV light that may be due to downward swimming trajectories (swimming with gravity) **(Figure 3---figure supplement 1C,D)**.
+Wild type but not mutant larvae also showed an increase in swimming speed under UV light that may be due to downward swimming trajectories (swimming with gravity) **(Figure 3B and Figure 3---figure supplement 1C)**.
 We also tested directional phototaxis, by exposing larvae to 480 nm directional and collimated light from the top of the column.
-Three-day-old but not two-day-old *NOS*-mutant larvae also showed reduced phototactic behaviour, suggesting a function for *NOS* in the visual eyes that mediate three-day-old phototaxis [@Randel2014a] **(Figure 1---figure supplement 1A,B)**.
+Three-day-old but not two-day-old *NOS*-mutant larvae also showed reduced phototactic behaviour, suggesting a function for *NOS* in the visual eyes that mediate three-day-old phototaxis [@Randel2014a] **(Figure 3D and Figure 3---figure supplement 1G)**.
 To distinguish between an acute or developmental function of NOS in light responses, we next tested larvae exposed to the NOS inhibitor L-NAME.
 Larvae incubated for 5 min in 0.1 mM or 1 mM L-NAME showed a dose-dependent inhibition of UV avoidance.
-In contrast, phototaxis was not effected **(Figure 3E)**.
+In contrast, phototaxis was not effected **(Figure 3C, E)**.
 Overall, our results indicate an acute requirement for NOS signalling in UV-avoidance and a possible indirect, developmental role in visual phototaxis.
@@
 ```{r echo=FALSE,fig.cap="**Figure 3.
-Larvae with inhibited NOS have a delayed response to UV-avoidance.** **(A)** Top: The domain and Exon/Intron structure of NOS.
-Bottom: Close-up region showing the genomic locus of NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9.
-The generated mutants (NOSΔ11, NOSΔ23) are also shown.
-Pink indicates target sites.
-Gray shows PAM sequences, red shows stop codons.
-**(B)** Overlaid trajectories for WT (n=32) and NOS mutant (NOSΔ11/Δ11, n=26 and NOSΔ23/Δ23, n=47) at three-day-old larvae.
+Larvae with inhibited NOS have a delayed response to UV-avoidance.**  **(B)** Overlaid trajectories for WT (n=32) and NOS mutant (NOSΔ11/Δ11, n=26 and NOSΔ23/Δ23, n=47) at three-day-old larvae.
 0 sec as the starting point.
@@
 As we have shown previously, a 20-sec local stimulation of cPRC cilia lead to a transient increase in cPRC calcium levels, followed by a transient decrease [@Veraszto2018].
-After \~20-sec, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus -- a response that may involve depolarisation.
+After \~20-sec, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus -- a response that may involve depolarisation **(Figure 4A, B)**.
 This activation phase occurs after the 20 sec stimulation period and is likely due to a delayed neuroendocrine feedback [@Veraszto2018].
 To determine whether NO mediates such a feedback, we repeated the experiment in *NOS*-mutant larvae.
-While we detected the initial activation phase followed by inhibition, in *NOS*-mutants this was not followed by delayed activation and calcium levels dropped to a low steady-state level **(Figure 4E,F)**.
+While we detected the initial activation phase followed by inhibition, in *NOS*-mutants this was not followed by delayed activation and calcium levels dropped to a low steady-state level **(Figure 4A, B)**.
 We thus identified a requirement for NO signalling in the late-phase activation of cPRCs.
@@
 Recently, Moroz and coworkers reported an atypical but widely conserved family of guanylyl cyclases with a NIT (nitrite/nitrate sensing) domain (PF08376) (NIT-GC) as potential mediators of NO signalling [@Moroz2020].
-To search for NIT-GCs in *Platynereis*, we searched transcriptome resources and retrieved 15 potential NIT-GC homologs **(Figure 4---figure supplement 1)**.
+To search for NIT-GCs in *Platynereis*, we searched transcriptome resources and retrieved 15 potential NIT-GC homologs **(Figure 4---figure supplement 1 and 2)**.
 To analyse the relationship of these sequences to metazoan NIT-GCs, we retrieved protein sequences with a CYC domain from the transcriptome and genome databases of 45 metazoan species and 2 choanoflagellate species.
@@
 In agreement with Moroz et al. [@Moroz2020], we found a group of GCs with NIT domains with representatives in placozoans, cnidarians, some ecdysozoans, echinoderms, and lophotrochozoans.
-The 15 *Platynereis* sequences were parts of several major and ancient clades of the tree **(Figure 4---figure supplement 1)**.
+The 15 *Platynereis* sequences were parts of several major and ancient clades of the tree **(Figure 4---figure supplement 1 and 2)**.
 To characterise the expression of NIT-GCs, we used previously published spatially mapped single-cell transcriptome data [@Achim2015; @williams2017].
-Among the 15 NIT-GCs, two showed high and specific expression in the cPRCs and one was expressed in the INNOS cells **(Figure 4---figure supplement 1)**.
-In the single-cell data, we could identify the cPRCs by the specific expression of c-opsin1 and a pedal-peptide2 neuropeptide precursor (MLD proneuropeptide) [@Arendt2004; @williams2017], which have previously been described as cPRC markers **(Figure 4---figure supplement 2A)**.
+Among the 15 NIT-GCs, two showed high and specific expression in the cPRCs and one was expressed in the INNOS cells **(Figure 4---figure supplement 2)**.
+In the single-cell data, we could identify the cPRCs by the specific expression of c-opsin1 and a pedal-peptide2 neuropeptide precursor (MLD proneuropeptide) [@Arendt2004; @williams2017], which have previously been described as cPRC markers **(Figure 4---figure supplement 3A)**.
 The INNOS cells were identified by NOS expression and spatial mapping in the brain [@Achim2015].
@@
 To confirm the single-cell data, we first carried out in situ HCR with probes for *NIT-GC1* and *NIT-GC2* mRNA.
-Both genes were specifically expression in the four cPRCs, as confirmed by co-labeling with an acetylated α-tubulin antibody and with an HCR probe against *pedal peptide 2/MLD proneuropeptide* **(Figure 4A,B and Figure 4---figure supplement 2A-C)**.
+Both genes were specifically expression in the four cPRCs, as confirmed by co-labeling with an acetylated α-tubulin antibody and with an HCR probe against *pedal peptide 2/MLD proneuropeptide* **(Figure 4C ,D and Figure 4---figure supplement 3A-C)**.
 To analyse the subcellular localisation of NIT-GC1 and NIT-GC2 at the protein level, we raised and affinity-purified polyclonal antibodies against a specific peptide sequence from both proteins.
-In immunostainings, we found that NIT-GC1 was localise to the region corresponding to the axonal projections of the cPRCs in the anterior nerve plexus **(Figure 4C)**.
-Co-immunostaining with the rabbit NIT-GC1 antibody and a custom rat antibody raised against a fragment of *Platynereis* NOS, revealed that both proteins were localised in close proximity in the neurosecretory plexus **(Figure 4---figure supplement 2D)**.
-In contrast, NIT-GC2 specifically labelled the large sensory ciliary region of the cPRCs **(Figure 4D)**.
+In immunostainings, we found that NIT-GC1 was localise to the region corresponding to the axonal projections of the cPRCs in the anterior nerve plexus **(Figure 4E)**.
+Co-immunostaining with the rabbit NIT-GC1 antibody and a custom rat antibody raised against a fragment of *Platynereis* NOS, revealed that both proteins were localised in close proximity in the neurosecretory plexus **(Figure 4---figure supplement 3D)**.
+In contrast, NIT-GC2 specifically labelled the large sensory ciliary region of the cPRCs **(Figure 4F)**.
 These different subcellular localisations suggest that the two NIT-GCs are involved in different intracellular signalling processes in the ciliary and axonal regions of the cPRCs.
@@
 In bacteria, NIT domains are thought to regulate cellular functions in response to intra- or extracellular nitrate and nitrite.
-NIT-GC1 has a NIT domain and a highly conserved cyclase domain that is expected to catalyse cGMP synthesis **(Figure 4I)**.
+NIT-GC1 has a NIT domain and a highly conserved cyclase domain that is expected to catalyse cGMP synthesis **(Figure 4G)**.
 The NIT domain may render NIT-GC1 dependent on NO signals.
 To test this, we co-expressed the synthetic cGMP indicator Green cGull [@Matsuda2017] and NIT-GC1 in cultured cells COS-7 (monkey kidney).
-For balanced expression, we used a single plasmid with the two open-reading frames separated by the 2A self-cleaving peptide **(Figure 4J)**.
-Application of the NO donor SNAP lead to increased cGMP levels, an effect we did not observe when cells were exposed to DMSO or when Green cGull was expressed alone **(Figure 4K-M)**.
-To test whether cGMP production is dependent on the NIT domain, we also tested a deletion construct of NIT-GC1 lacking the NIT domain **(Figure 4I)**.
-Cells expressing this construct and Green cGull did not show increased cGMP levels when exposed to SNAP **(Figure 4N)**.
+For balanced expression, we used a single plasmid with the two open-reading frames separated by the 2A self-cleaving peptide **(Figure 4H)**.
+Application of the NO donor SNAP lead to increased cGMP levels, an effect we did not observe when cells were exposed to DMSO or when Green cGull was expressed alone **(Figure 4I-K)**.
+To test whether cGMP production is dependent on the NIT domain, we also tested a deletion construct of NIT-GC1 lacking the NIT domain **(Figure 4G)**.
+Cells expressing this construct and Green cGull did not show increased cGMP levels when exposed to SNAP **(Figure 4L)**.
 These results indicate that NIT-GC1 is able to catalyse cGMP production in an NO-dependent manner and this function requires the NIT domain.
@@
 To test the in vivo function of NIT-GC1 and NIT-GC2 in cPRC responses, we combined calcium imaging with morpholino-mediated knockdowns.
-We used two translation-blocking morpholinos for each *NIT-GC* gene and tested knockdown efficiency by immunostaining injected animals with the NIT-GC1 and NIT-GC2 antibodies **(Figure 4---figure supplement 2E,F)**.
+We used two translation-blocking morpholinos for each *NIT-GC* gene and tested knockdown efficiency by immunostaining injected animals with the NIT-GC1 and NIT-GC2 antibodies **(Figure 4---figure supplement 3E,F)**.
 For both genes, the morpholinos led to a strong reduction in the respective antibody signal, confirming efficient knockdown and antibody specificity.
 
-In NIT-GC1 morphant larvae, the delayed activation of cPRCs following 405 nm stimulation did not occur **(Figure 4G)**.
+In NIT-GC1 morphant larvae, the delayed activation of cPRCs following 405 nm stimulation did not occur **(Figure 4M)**.
 This phenotype is similar to the phenotype of *NOS* mutants suggesting that NIT-GC1 acts as the NO sensor in cPRCs to drive their delayed activation.
-This could occur via increased cGMP production and the opening of a cyclic-nucleotide-gated (CNG) channel specific to cPRCs.
-NIT-GC2 morphant larvae, in contrast, showed a step-up increase in calcium following light stimulation **(Figure 4H)**.
+This could occur via increased cGMP production and the opening of a cyclic-nucleotide-gated (CNG) channel specific to cPRCs[@Tosches2014].
+NIT-GC2 morphant larvae, in contrast, showed a step-up increase in calcium following light stimulation **(Figure 4N)**.
 The calcium signal decayed during stimulation and was off after light off.
@@
 ```{r echo=FALSE,fig.cap="**Figure 4.
-cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.** **(A, B)** Expression analysis of the NIT-GC1 (C) and NIT-GC2 (D) gene (magenta) using in situ HCR.
+cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.** **(A, B)** Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in WT (dark yellow) and NOS mutant (E: NOSΔ11/Δ11, F: NOSΔ23/Δ23) (light pink) larvae.
+**(C, D)** Expression analysis of the NIT-GC1 (C) and NIT-GC2 (D) gene (magenta) using in situ HCR.
 Co-staining image using acetylated α-tubulin antibody (ac α-tub: green).
-**(C, D)** Localisation analysis using NIT-GC1 and NIT-GC2 antibodies.
+**(E, F)** Localisation analysis using NIT-GC1 (E) and NIT-GC2 (F) antibodies.
 Green shows co-staining with acetylated α-tubulin antibody (ac α-tub).
-**(E, F)** Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in WT (dark yellow) and NOS mutant (E: NOSΔ11/Δ11, F: NOSΔ23/Δ23) (light pink) larvae.
-**(G, H)** Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in NIT-GC1 (G: NIT1) and NIT-GC2 (H: NIT2) knockdown by morpholino (MO1 and 2) larvae.
-**(I)** The domain structure of NIT-GC1.
-Below shows the mutant NIT-GC1 (ΔNIT-GC1) with deletion of the NIT domain.
+**(G)** The domain structure of NIT-GC1.
+Below shows the mutant NIT-GC1 (NIT-GC1ΔNIT) with deletion of the NIT domain.
 Grey squares indicate the transmembrane region (TM).
-**(J)** A schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control.
+**(H)** A schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control.
 Green cGull shows cGMP-dependent green fluorescence.
-**(K-N)** The intensity changes in cGMP fluorescence over time for the four conditions are shown.
+**(I-L)** The intensity changes in cGMP fluorescence over time for the four conditions are shown.
 Gray traces show the optical responses to individual intensities (n > 6).
@@
 The dark line shows the mean value and the band the standard error range of the mean.
-The respective solution was acted on 2 min after the start of imaging (grey bars)."}
+The respective solution was acted on 2 min after the start of imaging (grey bars).
+**(M, N)** Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in NIT-GC1 (G: NIT1) and NIT-GC2 (H: NIT2) knockdown by morpholino (MO1 and 2) larvae."}
 
@@
 The MC cell was identified based on its position and intrinsic activity.
-To unambiguously identify all other cells from which we recorded calcium signals, we developed an on-slide immunostaining method **(Figure 5---figure supplement 1)**.
-We used the cell-specific antibody markers against RYamide (INNOS), RGWamide (INRGW) and serotonin (Ser-h1) [@Conzelmann2011] to immunostain agar-embedded larvae following calcium imaging.
+To unambiguously identify all other cells from which we recorded calcium signals, we developed an on-slide immunostaining method **(Figure 5---figure supplement 1A)**.
+We used the cell-specific antibody markers against RYamide (INNOS) **(Figure 5---figure supplement 1B-D)**, RGWamide (INRGW) and serotonin (Ser-h1) [@Conzelmann2011] to immunostain agar-embedded larvae following calcium imaging.
 Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision **(Figure 5A,B)**.
@@
 ```{r echo=FALSE,fig.cap="**Figure 6.
-Gene expressions and signalling mechanisms between cPRC, INNOS and INRGWa.** **(C)** Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq.
+Signalling mechanisms and mathematical modelling between cPRC, INNOS and INRGWa.** **(A)** Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq.
 The size of the dots is expressed in proportion to the percentage of cells expressing that gene relative to all cells.
 The colours represent the normal logarithm of the number of transcripts in the cells expressing the gene.
-**(D)** Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback."}
+**(B)** Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback.
+**(C)** **(D)**"}
 
@@
 
-```{r echo=FALSE,fig.cap="**Figure 3 -- figure supplement 1** **(A)** Overlaid trajectories for WT (n=37) and NOS mutant (NOSΔ11/Δ11, n=18 and NOSΔ23/Δ23, n=8) at two-day-old larvae.
+```{r echo=FALSE,fig.cap="**Figure 3 -- figure supplement 1** **(A)** Top: The domain and Exon/Intron structure of NOS.
+Bottom: Close-up region showing the genomic locus of NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9.
+The generated mutants (NOSΔ11, NOSΔ23) are also shown.
+Pink indicates target sites.
+Gray shows PAM sequences, red shows stop codons.
+**(B)** Overlaid trajectories for WT (n=37) and NOS mutant (NOSΔ11/Δ11, n=18 and NOSΔ23/Δ23, n=8) at two-day-old larvae.
 0 sec as the starting point.
 After 10 sec, UV (395 nm) stimulation from the side.
-**(B)** The temporal changes in the vertical position of the WT and mutant 2-days-old larvae before and after UV stimulation are shown.
+**(C)** The temporal changes in the vertical position of the WT and mutant two-days-old larvae before and after UV stimulation are shown.
 The starting points of each larval trajectory are set to 0.
 After UV stimulation is indicated by purple squares.
-**(C)** Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23) larvae at 2-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
-The data are shown in 30 s bins.
-**(D, E)** The temporal changes in the distance traveled of the WT and mutant in 2-day (D) and 3-day-old (E) larvae before and after UV stimulation are shown."}
+**(D, E)** The temporal changes in the distance traveled of the WT and mutant in two-day (D) and three-day-old (E) larvae before and after UV stimulation are shown.
+**(F)** The temporal changes in the distance traveled of larvae treated with NOS inhibitors, L-NAME in three-day larvae before and after UV stimulation are shown.
+**(G)** Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23) larvae at two-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
+The data are shown in 30 s bins."}
 knitr::include_graphics(path = "figures/Fig3_sup1.png")
@@
 
-```{r echo=FALSE,fig.cap="**Figure 4 -- figure supplement 1** "}
+```{r echo=FALSE,fig.cap="**Figure 4 -- figure supplement 1** Cluster analysis of guanylate and adenylate cyclases.
+Connections are based on blast similarities < 1e-16 as shown on the upper right.
+Animal groups are colour coded as shown on the upper left.
+NIT-GCs, NIT domain containing guanylate cyclases; membrane-bound GCs, Membrane-bound guanylyl cyclases; sGCs, soluble guanylate cyclases; ACs, adenylate cyclases."}
 knitr::include_graphics(path = "figures/Fig4_sup1.png")
@@
 
-```{r echo=FALSE,fig.cap="**Figure 4 -- figure supplement 3** Co-expression analysis image of the NIT-GC1 (magenta) and MLD-pedal2 amide proneuropeptide gene (MLD: green).
+```{r echo=FALSE,fig.cap="**Figure 4 -- figure supplement 3** **(A)** Co-expression analysis image of the NIT-GC1 (magenta) and MLD-pedal2 amide proneuropeptide gene (MLD: green).
 Anterior view of the larva at two-day-old.
@@
 
-```{r echo=FALSE,fig.cap="**Figure 5 -- figure supplement 1** Schematic diagram of the immunostaining procedure after calcium imaging."}
+```{r echo=FALSE,fig.cap="**Figure 5 -- figure supplement 1** **(A)** Schematic diagram of the immunostaining procedure after calcium imaging.**(B-D)** Co-expression analysis image of the NOS (magenta) and RY amide proneuropeptide gene (RYa: green).
+Anterior view of the larva at two-day-old.
+"}
 knitr::include_graphics(path = "figures/Fig5_sup1.png")

### 2023-05-14 code/Fig5_sup1.R

@@ -19,7 +19,11 @@ panel_NOS <- ggdraw() + draw_image(readPNG("pictures/HCR_52_AP_NOS_RYa_101.29um_
                arrow = arrow(ends = "both", type = "closed", length = unit(0.1,"cm")),
                lineend = "butt",
                linejoin = "mitre",
-               arrow.fill = "white", size = 0.2)
+               arrow.fill = "white", size = 0.2) +
+  draw_label("eyespots", x = 0.55, y = 0.34, size = 6, color = "white") +
+  draw_line(x = c(0.16, 0.44), y = c(0.51, 0.35), color = "white", size = 0.1) +
+  draw_line(x = c(0.92, 0.65), y = c(0.48, 0.35), color = "white", size = 0.1) +
+  draw_label("INNOS", x = 0.55, y = 0.7, color="white", size = 7, fontface="bold")
 
 panel_RYa <- ggdraw() + draw_image(readPNG("pictures/HCR_52_AP_NOS_RYa_101.29um_RYa.png"))+
   draw_label("RYa", x = 0.12, y = 0.9, color="green", size = 10, fontface="italic")
@@ -39,13 +43,13 @@ panel_method <- ggdraw() + draw_image(readPNG("pictures/Diagram of agarose embed
 #combine panels into Figure and save final figure as pdf and png
 #panels of different sizes
 layout <- "
-A#B#C
+AAAAA
 #####
-DDDDD
+B#C#D
 "
 
-Fig5_sup1 <- panel_NOS + panel_RYa + panel_merge + 
-  panel_method +  
+Fig5_sup1 <- panel_method + 
+  panel_NOS + panel_RYa + panel_merge + 
   patchwork::plot_layout(design = layout, 
                          widths = c(1, 0.02, 1, 0.02, 1), 
                          heights = c(1, 0.02, 1)) + #we can change the heights of the rows in our layout (widths also can be defined)
@@ -56,4 +60,4 @@ ggsave("Manuscript/figures/Fig5_sup1.png", limitsize = FALSE,
        units = c("px"), Fig5_sup1, width = 2000, height = 1475, bg='white')  
 
 ggsave("Manuscript/figures/Fig5_sup1.pdf", limitsize = FALSE, 
-       units = c("px"), Fig5_sup1, width = 1857, height = 970)  
+       units = c("px"), Fig5_sup1, width = 2000, height = 1475)  

### 2023-05-17 Manuscript/Manuscript.Rmd
manuscript

@@
 
-[to be added]
+We analysed the neurotransmitter-related genes expressed in the respective cells identified from the single-cell transcriptome data of Achim et al.
+to understand the neurotransmission mechanisms in cPRC, INNOS and INRGWa [@Achim2015; @williams2017].
+Interestingly, in cPRC, whereas cholinergic neurons has previously been implicated in neurotransmission [@Randel2014a], we found transporters and synthetic enzymes involved in glutaminergic, glycinergic, GABAergic and adrenergic innervation, respectively **(Figure 6A)**.
+Gene expression has been observed by Randel et al. in these genes, suggesting that they may be expressed in cPRC [@Randel2014a].
+Their receptors were also found for each of the expressed genes within INNOS and INRGWa **(Figure 6A)**.
+Based on these expression patterns and results, a model of transduction across the cPRC pathway focusing on Ca2+, cGMP and NO dynamics is proposed **(Figure 6B)**.
+The c-opsin1, which is UV-sensitive, activates G proteins (Gi/Go) and decreases cGMP CNGAα, which is already known to be strongly expressed in cPRC as one of the targets highly dependent on cGMP changes, may regulate the influx of ions into the cell [@Arendt2004; @tsukamoto2017; @Veedin2021].
+The Gβγ released by Gi/o activation also induces K+ efflux (hyperpolarisation) via GIRK channels and closes the voltage-gated Ca2+ channels, thereby reducing intracellular calcium levels[@tsukamoto2023].
+It is also thought that increased calcium levels in INNOS signalled from cPRC may activate NOS and generate nitric oxide; INNOS has glycinergic and glutamatergic potential and may act as an inhibitory signal regulator for INRGWa.
+INRGWa, and may function as an inhibitory signal modulator for INRGWa.
+
+We used this model and calcium imaging data to develop a mathematical model.
 
@@
 **(B)** Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback.
-**(C)** **(D)**"}
+**(C)** Diagram for **(D)**"}
 
@@
 
+**Functional diversity of NIT-GCs**
+
 NO is a free radical with a half-life of 5 msec in vivo and a signalling range limited to a few tens of nanometers.
@@
 In placozoans as many as 12 have been found, and cnidarians also contain a large diversity.
-NIT-GCs have been lost from chordates [check].
+
+**Nitric oxide feedback modulation of UV-avoidance behaviour**
+
+Why does a seemingly roundabout feedback pathway exist?
+The gaseous second messenger nitric oxide (NO) is known to activate a post- to presynaptic retrograde signalling cascade involving cyclic guanosine monophosphate (cGMP) production.
+It has been reported that NO produced in one of the mammalian retinal cells modulates cGMP-mediated neurotransmission by activating the presynaptic nerve sGC (a general target of NO) through retrograde signalling [@Vielma2014; @Wei2012].
+Retrograde signalling of NO is also known to be one of the molecules that regulate memory formation.
+In Drosophila, NOS of PPL1-γ1pedc shortens memory retention while promoting fast memory updating in response to new experiences, indicating that NO regulates forgetting [@Aso2019].
+Because NO has a short half-life and may be a suitable second messenger for encoding working memories whose duration must be limited.
+R-neuron inputs to the ellipsoid in Drosophila also represent visual object orientation for tens of seconds after the disappearance of an object, suggesting that they are involved in short-term visual memory dependent on nitric oxide signalling [@Kuntz2017].
+In this study, our mathematical modelling results focused on the time of UV stimulation and showed that the late activity comes more delayed during short UV and sooner during long UV.
+This suggests that animals may have a UV length-dependent short-term memory system that regulates cPRC activation, and that there may be mechanisms by which UV stimulation is enhanced and maintained such that NO feedback maintains neural activation even with slight UV.
+It is possible that these mechanisms may have been early signalling pathways controlling stimulus changes across space and time in animals' short-term memory.
+In planarian UV avoidance behaviour, different time and length scales of neuropeptide and small molecule transmission were reported to generate incoherent patterns of neural activity that competitively control behaviour and memory [@bray2023].
+Short-term memory is essential for locating less exposed areas to minimise UV-induced damage in animals.
 
@@
 
-Another type of interneuron that strongly synapses with the cPRC is the INRGWa, which synapses directly onto the ciliomotor neuron and is therefore considered to be a very important neuron in the control of larval behaviour [@Veraszto2018][@williams2017].
-The present results suggest that the depolarising response of the cPRC due to NO feedback induces INRGW activation.
+**The mystery of why IN-RGW activation causes downward swimming**
+
+One type of interneuron that strongly synapses with the cPRC is INRGWa, which synapses directly onto ciliary motoneurons and is therefore considered to be a very important neuron in larval behavioural control [@williams2017; @Veraszto2018].
+Our results suggest that the depolarising response of the cPRC by NO feedback induces INRGW activation.
+Our experiments have shown that exposure of larvae to RGWamide results in simultaneous downward swimming and that activation of INRGW is sufficient to trigger UV-avoidance behaviour (unpublished data).
+However, whole-cell tracking by connectome has not revealed any equilibrium organ that recognises the direction of gravity, and it remains to be investigated how activation of INRGWs causes animals to recognise 'down', e.g.
+in RGWamide-deficient larvae.
+It is known that complex integrative networks of neurons are involved in UV avoidance behaviour in planaria [@Shettigar2021] and Drosophila [@Imambocus2022].
+It has been reported that the sensory neurons required for UV avoidance behaviour in Drosophila converge and arrest neuromodulatory hub neurons, and that compartmentalised circuit organisation and neuropeptide release from the regulatory hubs act as central circuit elements controlling escape responses.
+RGW interneurons have been suggested to be important for relaying peptidergic signals due to the distinct anatomy and connectivity of projection neurons in peptidergic networks and their high authority values, and central with many neuronal inputs and outputs network components may function as regulatory hubs that gate the activation of specific networks [@williams2017].
+The finding of sGCs in INRGWs from the present single-cell analysis also suggests that they may be regulated by NO.
+The specific demarcation of sensory input and output through the use of neurotransmitters, neuropeptides and even gaseous NO may further increase the efficiency of network computation in hub neurons and may lead to a more complex mechanism of 'down' recognition.
 

### 2023-05-19 Manuscript/Manuscript.Rmd
Add single-cell source data

@@
 Neuronally produced nitric oxide (NO) can act as a diffusible neuromodulator.
-This small free-radical molecule is synthesized by nitric oxide synthase (NOS)...
-
-In the mammalian retina, NO is produced by amacrine cells and activates a *soluble guanylate cyclase (sGC)* in cone bipolar cells.
-The NO-dependent increased in cGMP tunes bipolar cell activity ....
-[refs] [other examples]
+This small free-radical molecule is synthesized by nitric oxide synthase (NOS).
+The most major target of NO signalling in the neuronal circuit is soluble guanylate cyclase (sGC).
+sGC activated by NO produces cGMP, which acts as a second messenger and regulates neuronal activity.
+In the mammalian retina, NO is an essential signalling molecule and one of the most well-studied [@Cudeiro1999].
+NO produced in amacrine cells, one of the retinal cells[@jacoby2018], activates sGCs (general targets of NO) in pyramidal bipolar cells, the presynaptic nerves, via retrograde signalling; NO-dependent increases in cGMP have been reported to regulate neurotransmission [@Vielma2014][@Wei2012].
+It has also been shown that in Drosophila, object features are differentially stored and that visual orientation memory for objects out of view is processed in the central complex of the fly brain, with ring neurons in the oval body involved in memory formation.
+Second messengers in the ellipsoid are required for directional memory and nitric oxide (NO) has been implicated in memory formation.
+cGMP-dependent signalling pathways via sGC cause activation of transcription factors, which are necessary for the ability to build orientation memories[@Kuntz2017].
+NO as a neuromodulator has been reported to be involved in the regulation of vision and memory via the NO-sGC-cGMP pathway in neural circuits, but the role of NO signalling in the regulation of neural circuit dynamics is not well understood to the molecular level mechanisms.
 
@@
 
-The activity of synaptic circuits is richly modulated monoamines, neuropeptides and gaseous molecules.
 Our work revealed an essential role for NO-mediated signalling in driving UV/violet avoidance behaviour.
@@
 In the cPRC circuit, synaptic connectivity alone is not sufficient to drive circuit dynamics and behavioural change.
-
 We identified cPRC-expressed NIT-GC1 as a new type of NO sensor.
@@
 NO-mediated feedback signalling thus likely occurs in the neurosecretory plexus where NOS- and NIT-GC1-containing projections are in close proximity and where we could detect NO production following UV stimulation.
-
 We showed that NIT-GC1 can produce cGMP in a NO-dependent manner.
 In cPRCs, increased cGMP levels could open CNGAα, a cyclic nucleotide-gated channels specifically expressed in cPRC [@Tosches2014], leading to cPRC delayed depolarisation, even after UV off.
+Also our mathematical model...
 
@@
 Why does a seemingly roundabout feedback pathway exist?
+The activity of synaptic circuits is richly modulated monoamines, neuropeptides and gaseous molecules.
 The gaseous second messenger nitric oxide (NO) is known to activate a post- to presynaptic retrograde signalling cascade involving cyclic guanosine monophosphate (cGMP) production.
@@
 Short-term memory is essential for locating less exposed areas to minimise UV-induced damage in animals.
-
-cPRCs expressing UV-absorbing c-opsin1 have high resting calcium levels and show hyperpolarising and subsequent depolarising responses to UV light (405 nm) [@Veraszto2018].
-In the present study, we have shown that the cells that strongly synapse with cPRCs are interneurons that specifically express NOS (INNOS).
-
-[@puopolo2001][@zhang2007][@Randel2014a].
-In the mammalian retina, it is known that signals from bipolar photoreceptor cells are transmitted to downstream NOS-expressing amacrine cells (NOAC), where NO is produced [@jacoby2018].
-These suggest that UV stimulation of cPRCs causes the production of NO via a depolarising response of INNOS.
-
-In the mammalian retina, NO is an essential signalling molecule and one of the most well-studied [@Cudeiro1999].
-It has been reported that NO produced in one of the retinal cells modulates cGMP-mediated neurotransmission by activating presynaptic neural sGCs (general target of NO) through retrograde signalling [@Vielma2014][@Wei2012].
 
@@
 The specific demarcation of sensory input and output through the use of neurotransmitters, neuropeptides and even gaseous NO may further increase the efficiency of network computation in hub neurons and may lead to a more complex mechanism of 'down' recognition.
-
-In addition, in many marine invertebrates, NO has been reported to be involved in many aspects of settlement behaviour and metamorphosis induction [@leise2001; @bishop2003; @bishop2007; @pechenik2007; @castellano2014; @ueda2013; @ueda2014; @ueda2016; @yang2018; @zhu2020; @song2021].
-The downward swimming behaviour of Platynereis larvae is the first step in the onset of settlement behaviour and metamorphosis.
-These suggest that downward swimming due to NO activation of INRGWs may trigger the onset of settlement behaviour and metamorphosis [@Conzelmann2013][@Williams2015].
-These suggest that NO feedback signalling may act as a trigger for downward swimming via activation of INRGWs, which induces the initiation of settlement behaviour and metamorphosis.
 
@@
 To calculate branch support, we ran 1,000 replicates with the aLRT-SH-like and aBayes methods [@Minh2020].
-The sequences used for the phylogenetic analysis are available in **Supplementary File 2**, the trimmed alignment is available in **Supplementary File 3**.
+The sequences used for the phylogenetic analysis are available in **Supplementary File 1**, the trimmed alignment is available in **Supplementary File 2** and the pre-trimmed data in **Supplementary file 3**.
 
@@
 We ran CD-Hit [@Fu2012] to eliminate redundant sequences (at a 80% threshold).
-The CLANS analysis is available as **Supplementary File 1**.
 To identify clusters, we used the convex-clustering option with 100 jack-knife replicates.
@@
 We used Achim et al. for the single-cell data [@Achim2015].
-In fact, in Williams et al., they used 107 cells as neurons by removing duplicates from Achim et al. single-cell data, so we used those cells [@williams2017].
+In Williams et al., they used 107 cells as neurons by removing duplicates from Achim et al. single-cell data, so we used those cells [@williams2017].
 Since the raw data were read count data, we normalized them to TPM using Python.
 After that we converted them to log10.
-From the sum of the expression levels in 107 cells for each gene, I calculated the percentage in each cell.
-We chose the appropriate genes for now.
+From the sum of the expression levels in 107 cells for each gene, We calculated the percentage in each cell.
 For each cell, we identified them with marker genes.
@@
 In this case we wanted to compare between samples, so we used TPM.
-In fact, TPM is becoming the mainstream method for many single-cell analyses.
 The total TPM of each gene between the samples was used to calculate the percentage of expressed genes.
+The total TPM values for each gene and the percentage of expressed genes are available in **Supplementary file 1**.
 
@@
 To design probes for HCR, we used custom software [@Kuehn2022] to create 20 DNA oligo probe pairs specific to P.
-dumerilii NOS, NIT-GC1, NIT-GC2, RYa-pNP (GenBank accession: JF811330.1), c-opsin1 (GenBank accession: AY692353.1), CNGAα (GenBank accession: KM199644.1), and MLD/pedal2-pNP (GenBank accession: KF515945.1).
-The NOS, NIT-GC1 and NIT-GC2 probes were designed to be associated with the b2 initiator sequence, while the RYa-pNP, c-opsin1, CNGAα and MLD/pedal2-pNP probes were designed to be associated with the b3 initiator sequence.
+dumerilii NOS, NIT-GC1, NIT-GC2, RYa-pNP (GenBank accession: JF811330.1), and MLD/pedal 2-pNP (GenBank accession: KF515945.1).
+The NOS, NIT-GC1 and NIT-GC2 probes were designed to be associated with the b2 initiator sequence, while the RYa-pNP and MLD/pedal 2-pNP probes were designed to be associated with the b3 initiator sequence.
 For the detection stage, samples were pre-hybridized in 200 µl of probe hybridization buffer for 1 hr at 37°C, and then incubated in 250 µl hybridization buffer containing probe oligos (4 pmol/ml) overnight at 37°C.
@@
 Larvae were imaged on a Zeiss LSM 880 with Airyscan (with a C-Apochromat 63X/1.2 Corr - water) with a frame rate of 1.88 frame/sec and an image size of 512 x 512 pixels.
-The larvae were stimulated in a region of interest (a circle with ?? pixel diameter) with 405 nm lasers controlled by the Bleaching mode.
+The larvae were stimulated in a region of interest (a circle with 50 pixel diameter) with 405 nm lasers controlled by the Bleaching mode.
 The imaging laser had a similar intensity than the stimulus laser but covered an area that was 10 times larger than the stimulus ROI.
@@
 For single-wavelength imaging experiments, cells in 35-mm dishes were washed twice and imaged in modified Ringer's buffer (140 mM NaCl, 3.5 mM KCl, 0.5mM NaH2PO4 , 0.5mM S-3 MgSO4 , 1.5 mM CaCl2 , 10 mM HEPES, 2 mM NaHCO3 and 5 mM glucose).
-Dishes were mounted on a stage heated at 37 °C and imaging was performed using an inverted microscope (LSM880, Zeiss) equipped with an oil-immersion objective lens (UApo/340, 40, NA = xx).
-Images were acquired using a xenon lamp, 460--495 nm excitation filter, 505-nm dichroic mirror and 510-- 550-nm emission filter (????).
+Dishes were mounted on a stage heated at 37 °C and imaging was performed using an inverted microscope (LSM880, Zeiss) equipped with an oil-immersion objective lens (UApo/340, 40, NA = 0.17).
+Images were acquired using a xenon lamp, 460--495 nm excitation filter, 505-nm dichroic mirror and 510-- 550-nm emission filter (Zeiss).
 S-Nitroso-N-acetyl-D, L-penicillamine (SNAP) was purchased from Sigma-Aldrich (St. Louis, MO, USA) .
@@
 
-```{r echo=FALSE,fig.cap="**Figure 1 -- figure supplement 3** "}
+```{r echo=FALSE,fig.cap="**Figure 1 -- figure supplement 3** ssTEM reconstruction of the INNOS (A), INRGW (B) and cPRC (C), anterior view. NS plexus; neurosecretory plexus"}
 knitr::include_graphics(path = "figures/Fig1_sup3.png")

### 2023-05-22 Manuscript/Manuscript.Rmd
add manuscript

@@
 Wedgewood<sup>1</sup> and Gáspár Jékely<sup>1,4,`*`</sup>"
-    affiliation: "<sup>1</sup>Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4 4QD, United Kingdom <sup>2</sup>Okinawa Institute of Science and Technology, Okinawa, Japan, <sup>3</sup>School of Biological Sciences, University of Bristol, Bristol, United Kingdom, <sup>4</sup>Centre for Organismal Studies (COS), University of Heidelberg, 69120 Heidelberg, Germany  <br> <sup>*</sup>Correspondence`:` kei.jokura @ exeter.ac.uk, gaspar.jekely @ cos.uni-heidelberg.de"
+    affiliation: "<sup>1</sup>Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4 4QD, United Kingdom <sup>2</sup>Okinawa Institute of Science and Technology, Okinawa, Japan, <sup>3</sup>School of Biological Sciences, University of Bristol, Bristol, United Kingdom, <sup>4</sup>Centre for Organismal Studies (COS), University of Heidelberg, 69120 Heidelberg, Germany  <br> <sup>*</sup>Correspondence`:` k.jokura @ exeter.ac.uk, gaspar.jekely @ cos.uni-heidelberg.de"
 output:

### 2023-05-23 Manuscript/Manuscript.Rmd
MS: rewrote introdution, added refs

@@
 ---
-title: Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates UV avoidance
+title: Nitric oxide feedback to ciliary photoreceptor cells orchestrates UV avoidance
   in zooplankton
@@
 Here we uncover a role for NO in shaping circuit dynamics to orchestrate light-avoidance behaviour.
-We studied UV-avoidance mediated by brain ciliary photoreceptors (cPRCs) in larval *Platynereis dumerilii*, a marine annelid.
-We found NOS specifically expressed in interneurons (INNOS) postsynaptic to cPRCs.
+We studied UV/violet light avoidance mediated by brain ciliary photoreceptors (cPRCs) in larval *Platynereis dumerilii*, a marine annelid.
+We found NOS expressed in interneurons (INNOS) postsynaptic to cPRCs.
 Stimulation of cPRCs by violet light leads to cPRC inhibition but concomitant INNOS activation and NO production.
@@
 In *NOS* mutants, NO feedback and projection-neuron activation do not occur and avoidance behaviour is defective.
-By mathematical modelling, we could capture the circuit dynamics in both wild-type and mutant larvae.
-Our results reveal how NO-mediated neuroendocrine signalling gates a synaptic circuit to initiate light-avoidance behaviour.
+By mathematical modelling, we could recapitulate phototransduction and circuit dynamics in both wild-type and mutant larvae.
+Our results reveal how NO-mediated retrograde signalling gates a synaptic circuit to initiate light-avoidance behaviour.
 
@@
 
-Neuronally produced nitric oxide (NO) can act as a diffusible neuromodulator.
-This small free-radical molecule is synthesized by nitric oxide synthase (NOS).
-The most major target of NO signalling in the neuronal circuit is soluble guanylate cyclase (sGC).
-sGC activated by NO produces cGMP, which acts as a second messenger and regulates neuronal activity.
-In the mammalian retina, NO is an essential signalling molecule and one of the most well-studied [@Cudeiro1999].
-NO produced in amacrine cells, one of the retinal cells[@jacoby2018], activates sGCs (general targets of NO) in pyramidal bipolar cells, the presynaptic nerves, via retrograde signalling; NO-dependent increases in cGMP have been reported to regulate neurotransmission [@Vielma2014][@Wei2012].
-It has also been shown that in Drosophila, object features are differentially stored and that visual orientation memory for objects out of view is processed in the central complex of the fly brain, with ring neurons in the oval body involved in memory formation.
-Second messengers in the ellipsoid are required for directional memory and nitric oxide (NO) has been implicated in memory formation.
-cGMP-dependent signalling pathways via sGC cause activation of transcription factors, which are necessary for the ability to build orientation memories[@Kuntz2017].
-NO as a neuromodulator has been reported to be involved in the regulation of vision and memory via the NO-sGC-cGMP pathway in neural circuits, but the role of NO signalling in the regulation of neural circuit dynamics is not well understood to the molecular level mechanisms.
+In nervous systems, synaptic transmission and volume transmission together shape circuit dynamics.
+While synaptic transmission occurs at specialised contact sites, volume transmission is characterised by the delocalised release of diverse diffusive neuromodulators.
 
-To investigate NO signalling in the modulation of neural circuit dynamics, we studied the larval stages of the marine annelid model *Platynereis dumerilii* [@ozpolat2021].
-We uncovered an essential function for NO in these larvae in UV/violet-light avoidance.
-In *Platynereis* larvae, UV/violet avoidance is characterised by downward swimming and is mediated by the UV/violet-responding brain ciliary photoreceptor cells (cPRCs).
+Nitric oxide (NO) is one such modulator with unique physical and signalling properties.
+This free radical synthesized by nitric oxide synthase (NOS) is short-lived and can diffuse across biological membranes [@Cudeiro1999; @thomas2015].
+Canonical NO signalling involves the activation of neuronal NOS following calcium increase, NO production and diffusion, and the NO-dependent activation of soluble guanylate cyclases (sGC) in target cells leading to cGMP production [@Bredt1990; @Hoelscher1997].
+Given that NOS activation is calcium dependent and NOS shows neuron-type-specific expression [@Gibbs1998a; @Wildemann1999; @Mobley2022], NO action can lead to the activity-dependent modulation of neural circuits at specific sites [@Wang2007; @Vielma2014; @jacoby2018] .
 
-The cPRCs express a vertebrate-type ciliary opsin, c-opsin1 [@Arendt2004].
-This opsin forms a UV-absorbing bistable photopigment with an absorption maximum around 384 nm [@tsukamoto2017, @Veraszto2018].
-*c-opsin1* knockout larvae show defective UV/violet avoidance [@Veraszto2018].
-Here we show that *Platynereis* *NOS* is expressed in interneurons of the cPRC circuitry and is required for UV/violet-avoidance behaviour.
-By combining calcium imaging across the fully-mapped cPRC circuit, genetic perturbations and mathematical modelling, we describe how NO tunes circuit dynamics through non-synaptic reverse signalling to cPRCs.
-This delayed neuroendocrine feedback is essential for circuit output, and integrates UV/violet input and also induces a short-term memory manifest in altered circuit activity.
+NO signalling has been extensively studied in the vertebrate retina where NOS is expressed in amacrine, ganglion as well as other cells [@Cudeiro1999; @Wang2007; @jacoby2018].
+The actions of NO can be diverse and not always involve the canonical sGC-cGMP pathway [@Tooker2013; @Wei2012; @jacoby2018].
+Defective NO signalling in NOS knockouts for example leads to a decreased sensitivity of retinal ganglion cells to light stimulation [@Wang2007].
+However, due to the complex expression of NOS in vertebrates and the diversity of its functions, it has been challenging to link the neurophysiological effects of NO to behaviour change.
 
-#In adult worms, UV light and c-opsin1 modulate locomotion and maturation [@Veedin2021].
+In the whole organism context, NO signalling has often been studied in diverse marine invertebrates, where NO can regulate larval settlement and metamorphosis [@leise2001a; @ueda2016; @Locascio2022; @song2021; @Zhang2012].
+While NOS-expressing neurons potentially responsible for these effects have been reported [@bishop2007; @Locascio2022], it has not been possible to link these NO-dependent effects on behaviour or life-cycle transitions to neuronal activity and function.
+
+Due to these difficulties, we still know little at the whole organism level about how NO production relates to stimulus conditions, how it shapes circuit activity at specific neuron types, and how NO-dependent modulation relates to behaviour.
+
+To investigate NO function in neural circuit dynamics and behaviour, here we study larvae of the marine annelid model *Platynereis dumerilii* [@ozpolat2021].
+*Platynereis* has emerged as a systems neuroscience model where behaviour and activity imaging can be combined with genetic manipulations and investigated in the context of a whole-body synaptic connectome and gene expression atlases [@Vergara2021; @Veraszto2017; @veraszto2020; @ozpolat2021].
+
+Here we uncovered an essential function for NO signalling in larval UV/violet-light avoidance behaviour.
+In *Platynereis* larvae, UV/violet avoidance is mediated by brain ciliary photoreceptor cells (cPRCs) and is characterised by downward swimming [@Veraszto2018].
+The cPRCs express a ciliary-type opsin, c-opsin1 [@Arendt2004] that forms a UV-absorbing bistable photopigment with an absorption maximum around 384 nm [@tsukamoto2017, @Veraszto2018, @Veedin2021].
+Upon UV/violet exposure, the cPRC show a characteristic biphasic calcium response that is *c-opsin1* dependent.
+UV/violet avoidance behaviour is also *c-opsin1*-dependent and is defective in *c-opsin1* mutants [@Veraszto2018].
+Here we show that *Platynereis* *NOS* is expressed in interneurons of the cPRC circuitry and is required for UV/violet-avoidance.
+By combining calcium imaging across the fully-mapped cPRC circuit [@Veraszto2018] with genetic perturbations and mathematical modelling, we describe how NO tunes circuit dynamics through non-synaptic retrograde signalling to cPRCs.
+This delayed neuroendocrine feedback integrates UV/violet exposure to induce a short-term memory manifested in altered circuit activity and an aversive behavioural response.
 
@@
 In contrast, phototaxis was not effected **(Figure 3C, E)**.
-Overall, our results indicate an acute requirement for NOS signalling in UV-avoidance and a possible indirect, developmental role in visual phototaxis.
+Overall, our results indicate an acute requirement for NOS signalling in UV-avoidance and a possible indirect, developmental role in the visual system, reminiscent of the developmental role for NO signalling in the *Drosophila* eye [@Gibbs1998a].
 

### 2023-05-24 Manuscript/Manuscript.Rmd
MS: edit intro and fig1 and 2 legends

@@
 Defective NO signalling in NOS knockouts for example leads to a decreased sensitivity of retinal ganglion cells to light stimulation [@Wang2007].
-However, due to the complex expression of NOS in vertebrates and the diversity of its functions, it has been challenging to link the neurophysiological effects of NO to behaviour change.
+However, due to the complex expression of NOS in vertebrates and the diversity of its functions, it has been challenging to link the neurophysiological effects of NO to signalling mechanisms and behaviour change.
 
@@
 To investigate NO function in neural circuit dynamics and behaviour, here we study larvae of the marine annelid model *Platynereis dumerilii* [@ozpolat2021].
-*Platynereis* has emerged as a systems neuroscience model where behaviour and activity imaging can be combined with genetic manipulations and investigated in the context of a whole-body synaptic connectome and gene expression atlases [@Vergara2021; @Veraszto2017; @veraszto2020; @ozpolat2021].
+*Platynereis* has emerged as a systems neuroscience model where behaviour and activity imaging can be combined with genetic manipulations and a whole-body synaptic connectome and gene expression atlases are available [@Vergara2021; @Veraszto2017; @veraszto2020; @ozpolat2021].
 
@@
 Upon UV/violet exposure, the cPRC show a characteristic biphasic calcium response that is *c-opsin1* dependent.
-UV/violet avoidance behaviour is also *c-opsin1*-dependent and is defective in *c-opsin1* mutants [@Veraszto2018].
-Here we show that *Platynereis* *NOS* is expressed in interneurons of the cPRC circuitry and is required for UV/violet-avoidance.
+UV/violet avoidance is also *c-opsin1*-dependent and is defective in *c-opsin1* mutants [@Veraszto2018].
+Here we show that *Platynereis* *NOS* is expressed in interneurons of the cPRC circuit and is required for UV/violet-avoidance.
 By combining calcium imaging across the fully-mapped cPRC circuit [@Veraszto2018] with genetic perturbations and mathematical modelling, we describe how NO tunes circuit dynamics through non-synaptic retrograde signalling to cPRCs.
@@
 In the synaptic connectome, the INNOS cells are postsynaptic to the UV-sensory cPRCs and presynaptic to the INRGW interneurons, which are also cPRC targets **(Figure 1C,F)**.
-INRGW neurons synapse on the head serotonergic ciliomotor neurons (Ser-h1), which synapse on the prototroch ciliary band and the cholinergic MC ciliomotor neurons **(Figure 1C,F)**.
+The projections of INNOS cells are segregated into input and output compartments, with cPRCs inputs occurring in the dense neurosecretory plexus.
+INNOS output synapses form in the more ventral projection region **(Figure 1---figure supplement 3)**.
+INRGW neurons synapse on the head serotonergic ciliomotor neurons (Ser-h1), which synapse on the prototroch ciliary band and the cholinergic MC ciliomotor neurons **(Figure 1C,F)** [@Veraszto2017].
 
 ```{r echo=FALSE,fig.cap="**Figure 1.
-Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** **(A)** SEM image of a three-day-old Platynereis larva.
-**(B, C)** Five types of neurons (cPRC, INNOS, INRGW, Ser-h1 and MC) in the cPRC circuit reconstructed in the whole-body transmission electron microscopy (ssTEM) dataset of a three-day-old larva.
-Ciliated cells labelled with (B) are shown in grey.
-The INNOS on the ventral side labelled with (C) are shown in purple.
-Axons and dendrites appear as lines and cell body positions are represented by spheres.
-**(D)** Expression analysis of the NOS gene (magenta) using in situ HCR.
-Anterior view of the larva at two-day-old.
-Co-stained image using acetylated α-tubulin antibody (acTub: green).
-**(E)** Immunostaining with anti-HA antibody against HA tagged palmitoylated-tdTomato expressed under the upstream of the start site of NOS (NOSp::palmi-3xHA: magenta).
-Co-staining with acetylated α-tubulin antibody as marker for cilia of cPRC and axonal scaffold (acTub: green).
-Views of larvae at three-day-old from the anterior side.
-**(F)** Wiring diagram of the cPRC circuit.
-Hexagons and arrows represent cell groups and their synaptic connections, respectively.
-Numbers inside hexagons indicate the number of cells grouped in each node.
-Arrow line thickness (synapse weight) is equal to the number of synapses."}
+Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** **(A)** Scanning electron microscopy image of a three-day-old *Platynereis* larva.
+**(B, C)** Volume rendering of the neuron types (cPRC, INNOS, INRGW, Ser-h1 and MC) in the cPRC circuit reconstructed from a whole-body transmission electron microscopy volume of a three-day-old larva.
+Neurite skeletons are shown with cell-body positions represented by spheres.
+Projections of all neurons in the body are shown in grey to highlight the neuropils.
+The outline of the yolk is also indicated in grey.
+In B, nuclei positions of the prototroch head ciliary band are shown as grey spheres.
+**(D)** Expression of the *NOS* gene detected by in situ HCR (magenta) in a two-day-old larva (anterior view).
+Antibody staining for acetylated α-tubulin (acTub: green) highlights cPRC cilia and the neuropil.
+**(E)** Expression of a *NOS* reporter (NOSp::palmi-3xHA: magenta) labelled with an anti-HA antibody in a two-day-old larva (anterior view).
+Antibody staining for acetylated α-tubulin (acTub: green) highlights cPRC cilia and the neuropil.
+**(F)** Synaptic wiring diagram of the cPRC circuit.
+Hexagons represent cell groups, with the number of cells per group shown in square brackets.
+Arrows represent the summed number of synaptic contacts between cell groups.
+Arrow thickness is proportional to the number of synapses."}
 
@@
 
-**Nitric oxide produced during UV/violet stimulation gates the output of the cPRC circuit**
+**Nitric oxide is produced during UV/violet stimulation of the cPRCs**
 
@@
 To test this, first we asked whether NO is produced during UV/violet stimulation of the larvae.
-We injected the fluorescent NO-reporter DAF-FM into zygotes and imaged two-day-old larvae while exposing the ramified sensory cilia of the cPRCs to 405 nm violet light.
-Following light stimulation of the cPRCs but not a control area, we detected an increase in DAF-FM fluorescence in the anterior neurosecretory neuropil, the region of INNOS projections **(Figure 2)**.
+We injected the fluorescent NO-reporter DAF-FM into zygotes and imaged two-day-old larvae while exposing the cPRC region to 405 nm violet light.
+Following light stimulation in the region of the cPRCs but not a control area, we detected an increase in DAF-FM fluorescence in the anterior neurosecretory neuropil, the region of INNOS projections **(Figure 2)**.
 
 ```{r echo=FALSE,fig.cap="**Figure 2.
-NO produced by UV/violet stimulation to cPRCs.** **(A)** A NO-positive cell visualised by DAF-FM (green) is shown (dashline).
-The white line indicates the outer frame of the larva.
-All subsequent images are of two-day-old larvae.
-**(B)** Intensity changes before (upper) and during 405 nm light stimulation (lower) in a cell visualised by DAF-FM.
-**(C)** The intensity of DAF-FM changes over time when cPRC (green) and other cells (ctr: gray) are stimulated with UV/violet (405 nm) light.
-Purple squares indicate the timing of 405 nm light stimulation, the transparent lines show the optical responses to individual intensities, each intensities are normalized (ΔF/F0) and the dark trace lines and the shaded bands show the mean value and the standard error range of the mean."}
+NO produced by UV/violet stimulation to cPRCs.** **(A)** DAF-FM fluorescence in the region of the neurosecretory neuropil.
+The white line indicates the outline of the larva.
+The dashed line corresponds to the area where fluorescence was quantified.
+**(B)** Changes in DAF-FM fluorescence before and during 405 nm light stimulation.
+**(C)** Changes in DAF-FM fluorescence over time during 405 nm stimulation of the cPRCs (green) or a control area (ctr: gray).
+The purple box indicate the duration of 405 nm stimulation.
+Individual traces normalized (ΔF/F0) are shown as thin lines.
+Thick lines show the mean value with 0.95 confidence intervals.
+N = 6 for cPRC stimulation."}
 

### 2023-05-24 Manuscript/Manuscript.Rmd
MS: edits NOS mutant section

@@
 
-Next, we wanted to directly test whether NO signalling is involved in UV/violet avoidance.
+We next tested whether NO signalling is required for UV/violet avoidance.
 To achieve this, we generated two *Platynereis* *NOS* knockout lines with the CRISPR/Cas9 method.
 We recovered two deletions (NOSΔ11/Δ11 and Δ23/Δ23), both leading to frame-shift and an early stop codon and thus likely representing null alleles **(Figure 3---figure supplement 1A)**.
-We could establish a homozygous line for both mutations indicating that NOS is not an essential gene.
+We could establish a homozygous line for both mutations indicating that NOS is not an essential gene in *Platynereis*.
 To quantify UV avoidance, we recorded the trajectories of freely swimming wild type and mutant larvae in vertical columns, illuminated laterally from two opposite sides with 395 nm UV light **(Figure 3A and Figure 3---figure supplement 1B)**.
@@
 This phenotype is similar to the defective UV-avoidance of *c-opsin1* mutant larvae and reveals a requirement for NOS in UV-avoidance behaviour [@Veraszto2018].
-Wild type but not mutant larvae also showed an increase in swimming speed under UV light that may be due to downward swimming trajectories (swimming with gravity) **(Figure 3B and Figure 3---figure supplement 1C)**.
-We also tested directional phototaxis, by exposing larvae to 480 nm directional and collimated light from the top of the column.
+Wild type but not mutant larvae also showed an increase in swimming speed under UV light that may be due to downward swimming trajectories (swimming in the direction of gravity) **(Figure 3B and Figure 3---figure supplement 1C)**.
+We also tested directional phototaxis, by exposing larvae to 480 nm directional collimated light from the top of the column.
 Three-day-old but not two-day-old *NOS*-mutant larvae also showed reduced phototactic behaviour, suggesting a function for *NOS* in the visual eyes that mediate three-day-old phototaxis [@Randel2014a] **(Figure 3D and Figure 3---figure supplement 1G)**.
-To distinguish between an acute or developmental function of NOS in light responses, we next tested larvae exposed to the NOS inhibitor L-NAME.
+To distinguish between an acute and developmental function of NOS in light responses, we next tested larvae exposed to the NOS inhibitor L-NAME.
 Larvae incubated for 5 min in 0.1 mM or 1 mM L-NAME showed a dose-dependent inhibition of UV avoidance.
-In contrast, phototaxis was not effected **(Figure 3C, E)**.
-Overall, our results indicate an acute requirement for NOS signalling in UV-avoidance and a possible indirect, developmental role in the visual system, reminiscent of the developmental role for NO signalling in the *Drosophila* eye [@Gibbs1998a].
+In contrast, phototaxis was not affected **(Figure 3C, E)**.
+Overall, our results indicate an acute requirement for NOS signalling in UV-avoidance and a possible indirect, developmental role in the visual system, reminiscent of the function of NO signalling in *Drosophila* eye development [@Gibbs1998a].
 

### 2023-05-24 Manuscript/Manuscript.Rmd
MS: edit DAF-FM section, fig legends 2-4, NIT-GC section

@@
 To test this, first we asked whether NO is produced during UV/violet stimulation of the larvae.
-We injected the fluorescent NO-reporter DAF-FM into zygotes and imaged two-day-old larvae while exposing the cPRC region to 405 nm violet light.
-Following light stimulation in the region of the cPRCs but not a control area, we detected an increase in DAF-FM fluorescence in the anterior neurosecretory neuropil, the region of INNOS projections **(Figure 2)**.
+We injected the fluorescent NO-reporter DAF-FM into zygotes and imaged two-day-old larvae while exposing the region of cPRC cilia to 405 nm violet light.
+To mark cell outlines, we coinjected mRNA encoding a red-fluorescent reporter (RGECO), allowing the identification of the cPRCs.
+Following light stimulation in the region of the ramified cilia of the cPRCs, we detected an increase in DAF-FM fluorescence in the anterior neurosecretory neuropil, the region of INNOS projections.
+This increase did not occur in larvae where we illuminated a control area  **(Figure 2)**.
 
@@
 The dashed line corresponds to the area where fluorescence was quantified.
+Circles indicate the location of cPRC and control stimulation.
+cPRCs are marked by thin lines.
 **(B)** Changes in DAF-FM fluorescence before and during 405 nm light stimulation.
@@
 Thick lines show the mean value with 0.95 confidence intervals.
-N = 6 for cPRC stimulation."}
+N = 9 larvae for control and 11 for cPRC stimulation."}
 
@@
 ```{r echo=FALSE,fig.cap="**Figure 3.
-Larvae with inhibited NOS have a delayed response to UV-avoidance.**  **(B)** Overlaid trajectories for WT (n=32) and NOS mutant (NOSΔ11/Δ11, n=26 and NOSΔ23/Δ23, n=47) at three-day-old larvae.
-0 sec as the starting point.
-After 10 sec, UV (395 nm) stimulation from the side.
-**(C)** The temporal changes in the vertical position of the WT and mutant larvae before (dark) and after UV stimulation (side 395 nm) are shown.
-The starting points of each larval trajectory are set to 0.
-**(D)** Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23) larvae at three-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
-The data are shown in 30 s bins.
-**(E)** Vertical swimming in larvae treated with NOS inhibitors, L-NAME at three-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
-The data are shown in 30 s bins."}
+NOS is required for UV avoidance in *Platynereis* larvae.**  **(B)** Swimming trajectories of wild type (WT, n=32) and NOS mutant (NOSΔ11/Δ11, n=26 and NOSΔ23/Δ23, n=47) three-day-old larvae.
+All trajectories start at 0 x and y position and time 0 corresponding to 10 sec after the onset of 395 nm stimulation from the side.
+**(B)** Vertical position of batches of wild type and mutant three-day-old larvae over time under 395 nm UV stimulation.
+The starting position of each larval trajectory was set to 0.
+**(C)** Vertical position of batches of control and L-NAME-treated (0.1 and 1 mM) three-day-old larvae over time under 395 nm UV stimulation.
+The starting position of each larval trajectory was set to 0.
+**(D)** Vertical displacement in 30 sec bins of wild type and mutant (NOSΔ11 and NOSΔ23) three-day-old larvae stimulated with 395 nm light from side, 488 nm light from the top and 395 nm light from the top.
+**(E)** Vertical displacement in 30 sec bins of control and L-NAME-treated (0.1 and 1 mM) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top."}
 
@@
 
-**NO feedback tunes cPRC responses to UV/violet stimulation**
+**NO retrograde signalling tunes cPRC responses to UV/violet stimulation**
 
@@
 To determine whether NO mediates such a feedback, we repeated the experiment in *NOS*-mutant larvae.
-While we detected the initial activation phase followed by inhibition, in *NOS*-mutants this was not followed by delayed activation and calcium levels dropped to a low steady-state level **(Figure 4A, B)**.
+While we detected the initial activation phase followed by inhibition, in homozygous *NOS*-mutants for both CRISPR alleles this was not followed by delayed activation.
+Instead, calcium levels dropped to a low steady-state level **(Figure 4A, B)**.
 We thus identified a requirement for NO signalling in the late-phase activation of cPRCs.
@@
 
-Next, we aimed to identify the NO-receptor in the cPRCs.
+We aimed next to identify the NO-receptor in the cPRCs.
 NO generally acts via soluble guanylate cyclases (sGC), belonging to the guanylate cyclase family with a CYC domain (PFAM domain: PF00211).
@@
 Recently, Moroz and coworkers reported an atypical but widely conserved family of guanylyl cyclases with a NIT (nitrite/nitrate sensing) domain (PF08376) (NIT-GC) as potential mediators of NO signalling [@Moroz2020].
-To search for NIT-GCs in *Platynereis*, we searched transcriptome resources and retrieved 15 potential NIT-GC homologs **(Figure 4---figure supplement 1 and 2)**.
-To analyse the relationship of these sequences to metazoan NIT-GCs, we retrieved protein sequences with a CYC domain from the transcriptome and genome databases of 45 metazoan species and 2 choanoflagellate species.
+To identify NIT-GCs in *Platynereis*, we searched transcriptome resources and retrieved 15 potential NIT-GC homologs **(Figure 4---figure supplement 1 and 2)**.
+To analyse the relationship of these sequences to metazoan NIT-GCs, we retrieved protein sequences with a CYC domain from the transcriptome and genome databases of 45 metazoan and 2 choanoflagellate species.
 We carried out cluster analysis and did phylogenetic reconstruction on a group of membrane-bound guanylyl cyclases with sGCs as an outgroup.
 In agreement with Moroz et al. [@Moroz2020], we found a group of GCs with NIT domains with representatives in placozoans, cnidarians, some ecdysozoans, echinoderms, and lophotrochozoans.
-The 15 *Platynereis* sequences were parts of several major and ancient clades of the tree **(Figure 4---figure supplement 1 and 2)**.
+The 15 *Platynereis* sequences belonged to several deeply diverged clades in the phylogenetic tree **(Figure 4---figure supplement 1 and 2)**.
 To characterise the expression of NIT-GCs, we used previously published spatially mapped single-cell transcriptome data [@Achim2015; @williams2017].
 Among the 15 NIT-GCs, two showed high and specific expression in the cPRCs and one was expressed in the INNOS cells **(Figure 4---figure supplement 2)**.
-In the single-cell data, we could identify the cPRCs by the specific expression of c-opsin1 and a pedal-peptide2 neuropeptide precursor (MLD proneuropeptide) [@Arendt2004; @williams2017], which have previously been described as cPRC markers **(Figure 4---figure supplement 3A)**.
+In the single-cell data, we could identify the cPRCs by the specific expression of *c-opsin1* and the *pedal-peptide2 neuropeptide precursor* (*MLD proneuropeptide*), previously described cPRC markers [@Arendt2004; @williams2017] **(Figure 4---figure supplement 3A)**.
 The INNOS cells were identified by NOS expression and spatial mapping in the brain [@Achim2015].
-We decided to focus on two NIT-GCs expressed in the cPRCs and with a full-length sequence, NIT-GC1 and NIT-GC2.
-To confirm the single-cell data, we first carried out in situ HCR with probes for *NIT-GC1* and *NIT-GC2* mRNA.
-Both genes were specifically expression in the four cPRCs, as confirmed by co-labeling with an acetylated α-tubulin antibody and with an HCR probe against *pedal peptide 2/MLD proneuropeptide* **(Figure 4C ,D and Figure 4---figure supplement 3A-C)**.
+We decided to focus on two NIT-GCs expressed in the cPRCs and with a full-length sequence, *NIT-GC1* and *NIT-GC2*.
+To confirm the single-cell data, we first carried out in situ hybridisation chain reaction (HCR) with probes for *NIT-GC1* and *NIT-GC2* mRNA.
+Both genes were specifically expression in the four cPRCs, as confirmed by co-labeling with an acetylated α-tubulin antibody and with an HCR probe against *pedal peptide 2/MLD proneuropeptide* **(Figure 4C, D and Figure 4---figure supplement 3A-C)**.
 To analyse the subcellular localisation of NIT-GC1 and NIT-GC2 at the protein level, we raised and affinity-purified polyclonal antibodies against a specific peptide sequence from both proteins.
 In immunostainings, we found that NIT-GC1 was localise to the region corresponding to the axonal projections of the cPRCs in the anterior nerve plexus **(Figure 4E)**.
-Co-immunostaining with the rabbit NIT-GC1 antibody and a custom rat antibody raised against a fragment of *Platynereis* NOS, revealed that both proteins were localised in close proximity in the neurosecretory plexus **(Figure 4---figure supplement 3D)**.
-In contrast, NIT-GC2 specifically labelled the large sensory ciliary region of the cPRCs **(Figure 4F)**.
+Co-immunostaining with the rabbit NIT-GC1 antibody and a custom rat antibody raised against *Platynereis* NOS revealed the localisation of both proteins in close proximity in the neurosecretory plexus **(Figure 4---figure supplement 3D)**.
+In contrast, NIT-GC2 specifically labelled the ramified sensory cilia of the cPRCs **(Figure 4F)**.
 These different subcellular localisations suggest that the two NIT-GCs are involved in different intracellular signalling processes in the ciliary and axonal regions of the cPRCs.
@@
 The NIT domain may render NIT-GC1 dependent on NO signals.
-To test this, we co-expressed the synthetic cGMP indicator Green cGull [@Matsuda2017] and NIT-GC1 in cultured cells COS-7 (monkey kidney).
+To test this, we co-expressed the cGMP indicator Green cGull [@Matsuda2017] and NIT-GC1 in cultured COS-7 (monkey kidney) cells, a cell line with minimal endogenous sGC activity.
 For balanced expression, we used a single plasmid with the two open-reading frames separated by the 2A self-cleaving peptide **(Figure 4H)**.
@@
 
-**NIT-GC1 is required for NO-mediated feedback to cPRCs during the UV response**
+**NIT-GC1 is required for NO-mediated retrograde signalling to cPRCs during the UV response**
 
@@
 This phenotype is similar to the phenotype of *NOS* mutants suggesting that NIT-GC1 acts as the NO sensor in cPRCs to drive their delayed activation.
-This could occur via increased cGMP production and the opening of a cyclic-nucleotide-gated (CNG) channel specific to cPRCs[@Tosches2014].
+This could occur via increased cGMP production and the opening of a cyclic-nucleotide-gated (CNG) channel specific to cPRCs [@Tosches2014].
 NIT-GC2 morphant larvae, in contrast, showed a step-up increase in calcium following light stimulation **(Figure 4N)**.
@@
 ```{r echo=FALSE,fig.cap="**Figure 4.
-cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.** **(A, B)** Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in WT (dark yellow) and NOS mutant (E: NOSΔ11/Δ11, F: NOSΔ23/Δ23) (light pink) larvae.
-**(C, D)** Expression analysis of the NIT-GC1 (C) and NIT-GC2 (D) gene (magenta) using in situ HCR.
-Co-staining image using acetylated α-tubulin antibody (ac α-tub: green).
-**(E, F)** Localisation analysis using NIT-GC1 (E) and NIT-GC2 (F) antibodies.
-Green shows co-staining with acetylated α-tubulin antibody (ac α-tub).
-**(G)** The domain structure of NIT-GC1.
-Below shows the mutant NIT-GC1 (NIT-GC1ΔNIT) with deletion of the NIT domain.
-Grey squares indicate the transmembrane region (TM).
-**(H)** A schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control.
-Green cGull shows cGMP-dependent green fluorescence.
-**(I-L)** The intensity changes in cGMP fluorescence over time for the four conditions are shown.
-Gray traces show the optical responses to individual intensities (n > 6).
-The individual intensities are normalized (ΔF/F0).
-The dark line shows the mean value and the band the standard error range of the mean.
-The respective solution was acted on 2 min after the start of imaging (grey bars).
-**(M, N)** Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in NIT-GC1 (G: NIT1) and NIT-GC2 (H: NIT2) knockdown by morpholino (MO1 and 2) larvae."}
+NOS and two NIT-GCs shape calcium signals during cPRC UV/violet response.** **(A, B)** GCaMP6s signals in cPRCs in wild type and *NOS* mutant (A, NOSΔ11/Δ11, B, NOSΔ23/Δ23) larvae during 405 nm light stimulation.
+**(C, D)** In situ HCR for (C) *NIT-GC1* and (D) *NIT-GC2* (magenta) in three-day-old *Platynereis* larvae.
+Larvae were co-stained with an antibody agains acetylated α-tubulin to label cPRC cilia and the neuropil (green).
+**(E, F)** Immunostaining for (E) NIT-GC1 and (F) NIT-GC2 (magenta), co-stained for acetylated α-tubulin (green).
+**(G)** The domain structure of *Platynereis* NIT-GC1 and the truncated NIT-GC1ΔNIT protein lacking the NIT domain.
+A predicted transmembrane region (TM) is shown in grey.
+**(H)** Schematic of the cell-based assay to detect cGMP production following the addition of an NO donor SNAP or DMSO as control.
+**(I-L)** Green cGull fluorescence over time for the four conditions tested.
+Individual responses and their mean with 0.95 confidence interval are shown (n > 6 cells).
+Intensities are normalized (ΔF/F0).
+The indicated chemicals were added at 2 min after the start of imaging (grey bars).
+**(M, N)** GCaMP6s signals in cPRCs in (M) NIT-GC1 and (N) NIT-GC2 morphant larvae during 405 nm light stimulation.
+Individual responses and their mean with 0.95 confidence interval are shown."}
 
@@
 We were able to image the activity of all neurons in the cPRC circuit (INNOS, INRGW, Ser-h1 and MC).
-The MC cell was identified based on its position and intrinsic activity.
+The MC cell was identified based on its position and intrinsic activity [@Veraszto2017].
 To unambiguously identify all other cells from which we recorded calcium signals, we developed an on-slide immunostaining method **(Figure 5---figure supplement 1A)**.
@@
 Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision **(Figure 5A,B)**.
-Due to the stereotypy of the larvae, we could also identify neurons based on their position and activity in activity-correlation maps **(Figure 5C)**.
+Due to the stereotypy of the larvae, we could also identify neurons based on their position and calcium activity in activity-correlation maps **(Figure 5C)**.
 
@@
 In both wild type and *NOS*-mutant larvae, INNOS cells showed an increase in calcium during stimulation **(Figure 5D)**.
-This response was suppressed in NIT-GC2 morphant larvae **(Figure 5E)** revealing an essential role for NIT-GC2-mediated cPRC suppression in INNOS activation.
-In contrast, INRGW cells were initially inhibited, followed by a delayed activation paralleling the second activation phase of cPRCs.
-This late INRGW response was lacking in *NOS*-mutants **(Figure 5F,G)**.
-In NIT-GC2 morphants, INRGW cells showed a transient increase in calcium that decayed after light off and a delayed activation was not present.
+In contrast, the INNOS response was flat or slightly negative in NIT-GC2 morphant larvae **(Figure 5E)** revealing an essential role for NIT-GC2-mediated cPRC suppression in INNOS activation.
+INRGW cells were initially inhibited during cPRC stimulation, followed by a delayed activation paralleling the second activation phase of cPRCs.
+This late INRGW response was lacking in *NOS*-mutants **(Figure 5F)**.
+In NIT-GC2 morphants, INRGW cells showed a transient increase in calcium that decayed after light off and a delayed activation was not present **(Figure 5G)**.
 Next, we imaged calcium signals from Ser-h1 and MC neurons in wild type and *NOS* mutant larvae.
-Ser-h1 cells showed an activation profile that correlated with cPRC activity, including a reduction in calcium during stimulation followed by rebound, a response that was defective in *NOS* mutants.
-MC cells showed sustained activation, including a late-phase that was lacking in *NOS* mutants **(Figure 5H,I)**.
-These data suggest that during 405 stimulation the Ser-h1 cells are inhibited and MC cells are activated, and this regulation is NO-dependent.
+Ser-h1 cells showed an activation profile that correlated with cPRC activity, including a reduction in calcium during stimulation followed by rebound, a response that was defective in *NOS* mutants **(Figure 5I)**.
+MC cells showed sustained activation, including a late-phase that was lacking in *NOS* mutants **(Figure 5I)**.
+These data suggest that during 405 nm stimulation the Ser-h1 cells are inhibited and MC cells are activated, and this regulation is NO-dependent.
 This pattern is expected to inhibit ciliary activity in the prototroch but not in the other ciliary bands, triggering NO-dependent downward swimming.
@@
 ```{r echo=FALSE,fig.cap="**Figure 5.
-NO produced by UV stimulation to cPRCs induces a depolarising response in cPRC via NIT-GC1.** **(A, B)** Each left figure shows a calcium imaging using GCaMP6s.
-The respective right figures show immunostaining images of the same positions in the respective left figures, (A) shows RYamide antibodies, (B) shows RGWamide and serotonin (5-HT) antibodies in red and DAPI in cyan.
-Asterisks indicate the nuclei of the cPRC, respectively, and the nuclei of the cells that were identified from the arrangement of the surrounding cells are indicated by numbers.
-From the position of the cells, the calcium imaging data identify INNOS in (A) and INRGW and Ser-h1 in (B), respectively.
-**(C)** Correlation (Pearson's r) map of neuronal activity of the INNOS, INRGW, Ser-h1 and MC neurons.
-**(D)** Changes over time in GCaMP6s intensities in INNOS when cPRC is stimulated with UV in WT (dark purple) and NOSΔ11/Δ11 mutant (light purple) larvae.
-**(E)** Changes over time in GCaMP6s intensities in INNOS when cPRC is stimulated with UV in NIT-GC2 morphants (MO1 and MO2: pink and light pink) larvae.
-**(F)** Changes over time in GCaMP6s intensities in INRGW when cPRC is stimulated with UV in WT (dark blue) and NOSΔ23/Δ23 mutant (light blue) larvae.
-**(G)** Changes over time in GCaMP6s intensities in INRGW when cPRC is stimulated with UV in NIT-GC2 morphants (MO1 and MO2: blue and light blue).
-**(H, I)** Changes over time in GCaMP6s intensities in Ser-h1 and MC when cPRC is stimulated with UV in WT (Ser-h1 and MC: dark red and brown) and NOSΔ11/Δ11 mutant (Ser-h1 and MC: light red and brown) larvae."}
+NOS- and NIT-GC2-dependent dynamics of the cPRC circuit.** **(A, B)** GCaMP6s imaging from cPRCs and INNOS cells (left panels) followed by on-slide immunostaining for (A) RYamide to label INNOS and (B) RGWamide+serotonin to label INRGW and Ser-h1 (red).
+Nuclei are stained with DAPI (cyan).
+Asterisks indicate cPRC nuclei.
+Numbers mark the same cells in the GCaMP and immunostaining images matched by position.
+**(C)** Correlation map of neuronal activity of the cPRCs, INNOS, INRGW, Ser-h1 and MC neurons.
+**(D)** GCaMP6s fluorescence in INNOS cells in wild type (WT) and NOSΔ11/Δ11 mutant larvae during 405 nm stimulation of the cPRC cilia.
+**(E)** GCaMP6s fluorescence in INNOS cells in NIT-GC2 morphant larvae during 405 nm stimulation.
+**(F)** GCaMP6s fluorescence in INRGW cells in wild type and NOSΔ23/Δ23 mutant larvae during 405 nm stimulation.
+**(G)** GCaMP6s fluorescence in INRGW cells in NIT-GC2 morphant larvae during 405 nm stimulation.
+**(H, I)** GCaMP6s fluorescence in (H) Ser-h1 cells and (I) the MC cell in wild type and NOSΔ11/Δ11 mutant larvae during 405 nm stimulation."}
 
@@
 Our work revealed an essential role for NO-mediated signalling in driving UV/violet avoidance behaviour.
-NO, produced by postsynaptic interneurons, feeds back to their presynaptic photoreceptor partners leading to their delayed activation.
-This late-phase activation of cPRCs drives circuit output through projection interneurons and ciliomotor neurons.
-In the cPRC circuit, synaptic connectivity alone is not sufficient to drive circuit dynamics and behavioural change.
+NO, produced by postsynaptic INNOS interneurons, signals retrogradely to the presynaptic cPRCs leading to their delayed activation.
+The resulting NIT-GC1-dependent late-phase activation of the photoreceptors drives circuit output through projection interneurons and ciliomotor neurons.
+In the cPRC circuit, synaptic connectivity alone is thus not sufficient to drive circuit dynamics and behavioural change.
 We identified cPRC-expressed NIT-GC1 as a new type of NO sensor.
@@
 Here, cPRC axons intermingle and synapse on INNOS neurites that contain NOS.
-NO-mediated feedback signalling thus likely occurs in the neurosecretory plexus where NOS- and NIT-GC1-containing projections are in close proximity and where we could detect NO production following UV stimulation.
+NO-mediated retrograde signalling thus likely occurs in the neurosecretory plexus where NOS- and NIT-GC1-containing projections are in close proximity and where we could detect NO production following UV stimulation.
 We showed that NIT-GC1 can produce cGMP in a NO-dependent manner.
-In cPRCs, increased cGMP levels could open CNGAα, a cyclic nucleotide-gated channels specifically expressed in cPRC [@Tosches2014], leading to cPRC delayed depolarisation, even after UV off.
+In cPRCs, increased cGMP levels could open CNGAα, a cyclic nucleotide-gated channels specifically expressed in cPRC [@Tosches2014], leading to their delayed depolarisation, even after UV off.
+
 Also our mathematical model...
@@
 
-NO is a free radical with a half-life of 5 msec in vivo and a signalling range limited to a few tens of nanometers.
+NO is a free radical with a millisecond-to-second scale half life in vivo and a limited signalling range.
 In neuronal signalling, NOS and sGC have been found to localise at close proximity at synapses [@burette2002][@garthwaite2015a].
@@
 Bacterial NIT domains regulate cellular functions in response to changes in extracellular and intracellular nitrate and/or nitrite concentrations [@camargo2007].
-NO is readily converted to nitrate and nitrite [@santos2011; @möller2019; @garthwaite2015a] and these molecules accumulate in placozoans and cnidarians in cells and tissues with high NOS activity [@Moroz2004; @Moroz2020].
+NO is readily converted to nitrate and nitrite [@santos2011; @moeller2019; @garthwaite2015a] and these molecules accumulate in placozoans and cnidarians in cells and tissues with high NOS activity [@Moroz2004; @Moroz2020].
 It is possible that the NIT domains in NIT-GCs may also sense nitrate and nitrite, as in bacteria.
@@
 This work was funded by the Wellcome Trust (214337/Z/18/Z).
+This project has received funding from the European Research Council (ERC) under the European Union’s Horizon 2020 research and innovation programme (grant agreement No 101020792).
+KJ has been supported by a JSPS Postdoctoral fellowhip, LAYG by a BBSRC Discovery fellowship (BB/W010305/1).
 

### 2023-05-24 code/Fig3_and_Fig3_figure_Suppl1.R
Fig3: changed x axis title from 30 sec bins to time (sec)

@@ -123,7 +123,7 @@ WTvsNOS_2d %>%
                expression('NOS'^'Δ23/Δ23'))) +
   theme_minimal()+
   theme_plot +
-  labs(fill = "genotype", x = "30 sec bins", y = "Vertical displacement (mm/s)")+
+  labs(fill = "genotype", x = "time (sec)", y = "Vertical displacement (mm/s)")+
   scale_y_continuous(limits = c(-0.46, 0.56)) +
   scale_x_discrete(labels=c("30", "60", "90", 
                             "30", "60", "90","120",
@@ -181,7 +181,7 @@ WTvsNOS_3d %>%
                expression('NOS'^'Δ23/Δ23'))) +
   theme_minimal()+
   theme_plot +
-  labs(fill = "genotype", x = "30 sec bins", y = "Vertical displacement (mm/s)")+
+  labs(fill = "genotype", x = "time (sec)", y = "Vertical displacement (mm/s)")+
   scale_y_continuous(limits = c(-0.81, 0.5)) +
   scale_x_discrete(labels=c("30", "60", "90", 
                            "30", "60", "90","120",
@@ -235,7 +235,7 @@ L_NAME_3d %>%
                "1.0 mM" = Okabe_Ito[6])) +
   theme_minimal() +
   theme_plot +
-  labs(x = "30 sec bins", y = "Vertical displacement (mm/s)")+
+  labs(x = "time (sec)", y = "Vertical displacement (mm/s)")+
   scale_y_continuous(limits = c(-0.67, 0.75)) +
   scale_x_discrete(labels=c("30", "60", "90", 
                             "30", "60", "90","120",

### 2023-05-24 Manuscript/Manuscript.Rmd
MS: removed bold from figure refs

@@
 We identified a single *nitric oxide synthase* (*NOS*) gene in the *Platynereis dumerilii* genome and transcriptome data.
-Phylogenetic analysis of NOS proteins indicate that *Platynereis* NOS belongs to an orthology group of metazoan NOS sequences **(Figure 1---figure supplement 1)**.
+Phylogenetic analysis of NOS proteins indicate that *Platynereis* NOS belongs to an orthology group of metazoan NOS sequences (Figure 1---figure supplement 1).
 To characterise the expression pattern of *NOS* we used in situ hybridization chain reaction (HCR) and transient transgenesis.
-In two- and three-day-old larvae, we detected *NOS* expression in four cells (two of them weakly expressing) in the apical organ region **(Figure 1D and Figure 1---figure supplement 2)**.
-*NOS* was also expressed in the region of the visual eyes (adult eyes) and the pigmented eyespots **(Figure 1---figure supplement 2)**.
-The four apical organ cells, but not the eyes, were also labelled with a *NOS* reporter construct driving palmitoylated-tdTomato **(Figure 1E)**.
+In two- and three-day-old larvae, we detected *NOS* expression in four cells (two of them weakly expressing) in the apical organ region (Figure 1D and Figure 1---figure supplement 2).
+*NOS* was also expressed in the region of the visual eyes (adult eyes) and the pigmented eyespots (Figure 1---figure supplement 2).
+The four apical organ cells, but not the eyes, were also labelled with a *NOS* reporter construct driving palmitoylated-tdTomato (Figure 1E).
 This reporter also revealed the axonal projections of these central NOS-expressing neurons.
-The position and morphology of the four *NOS+* cells allowed us to identify the same four cells as four interneurons (INNOS) in our three-day-old whole-body *Platynereis* volume EM data [@williams2017; @veraszto2020] **(Figure 1B,C)**.
-In the synaptic connectome, the INNOS cells are postsynaptic to the UV-sensory cPRCs and presynaptic to the INRGW interneurons, which are also cPRC targets **(Figure 1C,F)**.
+The position and morphology of the four *NOS+* cells allowed us to identify the same four cells as four interneurons (INNOS) in our three-day-old whole-body *Platynereis* volume EM data [@williams2017; @veraszto2020] (Figure 1B,C).
+In the synaptic connectome, the INNOS cells are postsynaptic to the UV-sensory cPRCs and presynaptic to the INRGW interneurons, which are also cPRC targets (Figure 1C,F).
 The projections of INNOS cells are segregated into input and output compartments, with cPRCs inputs occurring in the dense neurosecretory plexus.
-INNOS output synapses form in the more ventral projection region **(Figure 1---figure supplement 3)**.
-INRGW neurons synapse on the head serotonergic ciliomotor neurons (Ser-h1), which synapse on the prototroch ciliary band and the cholinergic MC ciliomotor neurons **(Figure 1C,F)** [@Veraszto2017].
+INNOS output synapses form in the more ventral projection region (Figure 1---figure supplement 3).
+INRGW neurons synapse on the head serotonergic ciliomotor neurons (Ser-h1), which synapse on the prototroch ciliary band and the cholinergic MC ciliomotor neurons (Figure 1C,F) [@Veraszto2017].
 
@@
 Following light stimulation in the region of the ramified cilia of the cPRCs, we detected an increase in DAF-FM fluorescence in the anterior neurosecretory neuropil, the region of INNOS projections.
-This increase did not occur in larvae where we illuminated a control area  **(Figure 2)**.
+This increase did not occur in larvae where we illuminated a control area  (Figure 2).
 
@@
 To achieve this, we generated two *Platynereis* *NOS* knockout lines with the CRISPR/Cas9 method.
-We recovered two deletions (NOSΔ11/Δ11 and Δ23/Δ23), both leading to frame-shift and an early stop codon and thus likely representing null alleles **(Figure 3---figure supplement 1A)**.
+We recovered two deletions (NOSΔ11/Δ11 and Δ23/Δ23), both leading to frame-shift and an early stop codon and thus likely representing null alleles (Figure 3---figure supplement 1A).
 We could establish a homozygous line for both mutations indicating that NOS is not an essential gene in *Platynereis*.
-To quantify UV avoidance, we recorded the trajectories of freely swimming wild type and mutant larvae in vertical columns, illuminated laterally from two opposite sides with 395 nm UV light **(Figure 3A and Figure 3---figure supplement 1B)**.
+To quantify UV avoidance, we recorded the trajectories of freely swimming wild type and mutant larvae in vertical columns, illuminated laterally from two opposite sides with 395 nm UV light (Figure 3A and Figure 3---figure supplement 1B).
 As previously shown, wild-type larvae swim downward following non-directional UV/violet light stimulation [@Veraszto2018].
-In contrast, both two- and three-day-old homozygous *NOS*-mutant larvae showed a strongly diminished UV-avoidance response **(Figure 3A, B and Figure 3---figure supplement 1B,C)**.
+In contrast, both two- and three-day-old homozygous *NOS*-mutant larvae showed a strongly diminished UV-avoidance response (Figure 3A, B and Figure 3---figure supplement 1B,C).
 This phenotype is similar to the defective UV-avoidance of *c-opsin1* mutant larvae and reveals a requirement for NOS in UV-avoidance behaviour [@Veraszto2018].
-Wild type but not mutant larvae also showed an increase in swimming speed under UV light that may be due to downward swimming trajectories (swimming in the direction of gravity) **(Figure 3B and Figure 3---figure supplement 1C)**.
+Wild type but not mutant larvae also showed an increase in swimming speed under UV light that may be due to downward swimming trajectories (swimming in the direction of gravity) (Figure 3B and Figure 3---figure supplement 1C).
 We also tested directional phototaxis, by exposing larvae to 480 nm directional collimated light from the top of the column.
-Three-day-old but not two-day-old *NOS*-mutant larvae also showed reduced phototactic behaviour, suggesting a function for *NOS* in the visual eyes that mediate three-day-old phototaxis [@Randel2014a] **(Figure 3D and Figure 3---figure supplement 1G)**.
+Three-day-old but not two-day-old *NOS*-mutant larvae also showed reduced phototactic behaviour, suggesting a function for *NOS* in the visual eyes that mediate three-day-old phototaxis [@Randel2014a] (Figure 3D and Figure 3---figure supplement 1G).
 To distinguish between an acute and developmental function of NOS in light responses, we next tested larvae exposed to the NOS inhibitor L-NAME.
 Larvae incubated for 5 min in 0.1 mM or 1 mM L-NAME showed a dose-dependent inhibition of UV avoidance.
-In contrast, phototaxis was not affected **(Figure 3C, E)**.
+In contrast, phototaxis was not affected (Figure 3C, E).
 Overall, our results indicate an acute requirement for NOS signalling in UV-avoidance and a possible indirect, developmental role in the visual system, reminiscent of the function of NO signalling in *Drosophila* eye development [@Gibbs1998a].
@@
 As we have shown previously, a 20-sec local stimulation of cPRC cilia lead to a transient increase in cPRC calcium levels, followed by a transient decrease [@Veraszto2018].
-After \~20-sec, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus -- a response that may involve depolarisation **(Figure 4A, B)**.
+After \~20-sec, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus -- a response that may involve depolarisation (Figure 4A, B).
 This activation phase occurs after the 20 sec stimulation period and is likely due to a delayed neuroendocrine feedback [@Veraszto2018].
@@
 While we detected the initial activation phase followed by inhibition, in homozygous *NOS*-mutants for both CRISPR alleles this was not followed by delayed activation.
-Instead, calcium levels dropped to a low steady-state level **(Figure 4A, B)**.
+Instead, calcium levels dropped to a low steady-state level (Figure 4A, B).
 We thus identified a requirement for NO signalling in the late-phase activation of cPRCs.
@@
 Recently, Moroz and coworkers reported an atypical but widely conserved family of guanylyl cyclases with a NIT (nitrite/nitrate sensing) domain (PF08376) (NIT-GC) as potential mediators of NO signalling [@Moroz2020].
-To identify NIT-GCs in *Platynereis*, we searched transcriptome resources and retrieved 15 potential NIT-GC homologs **(Figure 4---figure supplement 1 and 2)**.
+To identify NIT-GCs in *Platynereis*, we searched transcriptome resources and retrieved 15 potential NIT-GC homologs (Figure 4---figure supplement 1 and 2).
 To analyse the relationship of these sequences to metazoan NIT-GCs, we retrieved protein sequences with a CYC domain from the transcriptome and genome databases of 45 metazoan and 2 choanoflagellate species.
@@
 In agreement with Moroz et al. [@Moroz2020], we found a group of GCs with NIT domains with representatives in placozoans, cnidarians, some ecdysozoans, echinoderms, and lophotrochozoans.
-The 15 *Platynereis* sequences belonged to several deeply diverged clades in the phylogenetic tree **(Figure 4---figure supplement 1 and 2)**.
+The 15 *Platynereis* sequences belonged to several deeply diverged clades in the phylogenetic tree (Figure 4---figure supplement 1 and 2).
 To characterise the expression of NIT-GCs, we used previously published spatially mapped single-cell transcriptome data [@Achim2015; @williams2017].
-Among the 15 NIT-GCs, two showed high and specific expression in the cPRCs and one was expressed in the INNOS cells **(Figure 4---figure supplement 2)**.
-In the single-cell data, we could identify the cPRCs by the specific expression of *c-opsin1* and the *pedal-peptide2 neuropeptide precursor* (*MLD proneuropeptide*), previously described cPRC markers [@Arendt2004; @williams2017] **(Figure 4---figure supplement 3A)**.
+Among the 15 NIT-GCs, two showed high and specific expression in the cPRCs and one was expressed in the INNOS cells (Figure 4---figure supplement 2).
+In the single-cell data, we could identify the cPRCs by the specific expression of *c-opsin1* and the *pedal-peptide2 neuropeptide precursor* (*MLD proneuropeptide*), previously described cPRC markers [@Arendt2004; @williams2017] (Figure 4---figure supplement 3A).
 The INNOS cells were identified by NOS expression and spatial mapping in the brain [@Achim2015].
@@
 To confirm the single-cell data, we first carried out in situ hybridisation chain reaction (HCR) with probes for *NIT-GC1* and *NIT-GC2* mRNA.
-Both genes were specifically expression in the four cPRCs, as confirmed by co-labeling with an acetylated α-tubulin antibody and with an HCR probe against *pedal peptide 2/MLD proneuropeptide* **(Figure 4C, D and Figure 4---figure supplement 3A-C)**.
+Both genes were specifically expression in the four cPRCs, as confirmed by co-labeling with an acetylated α-tubulin antibody and with an HCR probe against *pedal peptide 2/MLD proneuropeptide* (Figure 4C, D and Figure 4---figure supplement 3A-C).
 To analyse the subcellular localisation of NIT-GC1 and NIT-GC2 at the protein level, we raised and affinity-purified polyclonal antibodies against a specific peptide sequence from both proteins.
-In immunostainings, we found that NIT-GC1 was localise to the region corresponding to the axonal projections of the cPRCs in the anterior nerve plexus **(Figure 4E)**.
-Co-immunostaining with the rabbit NIT-GC1 antibody and a custom rat antibody raised against *Platynereis* NOS revealed the localisation of both proteins in close proximity in the neurosecretory plexus **(Figure 4---figure supplement 3D)**.
-In contrast, NIT-GC2 specifically labelled the ramified sensory cilia of the cPRCs **(Figure 4F)**.
+In immunostainings, we found that NIT-GC1 was localise to the region corresponding to the axonal projections of the cPRCs in the anterior nerve plexus (Figure 4E).
+Co-immunostaining with the rabbit NIT-GC1 antibody and a custom rat antibody raised against *Platynereis* NOS revealed the localisation of both proteins in close proximity in the neurosecretory plexus (Figure 4---figure supplement 3D).
+In contrast, NIT-GC2 specifically labelled the ramified sensory cilia of the cPRCs (Figure 4F).
 These different subcellular localisations suggest that the two NIT-GCs are involved in different intracellular signalling processes in the ciliary and axonal regions of the cPRCs.
@@
 In bacteria, NIT domains are thought to regulate cellular functions in response to intra- or extracellular nitrate and nitrite.
-NIT-GC1 has a NIT domain and a highly conserved cyclase domain that is expected to catalyse cGMP synthesis **(Figure 4G)**.
+NIT-GC1 has a NIT domain and a highly conserved cyclase domain that is expected to catalyse cGMP synthesis (Figure 4G).
 The NIT domain may render NIT-GC1 dependent on NO signals.
 To test this, we co-expressed the cGMP indicator Green cGull [@Matsuda2017] and NIT-GC1 in cultured COS-7 (monkey kidney) cells, a cell line with minimal endogenous sGC activity.
-For balanced expression, we used a single plasmid with the two open-reading frames separated by the 2A self-cleaving peptide **(Figure 4H)**.
-Application of the NO donor SNAP lead to increased cGMP levels, an effect we did not observe when cells were exposed to DMSO or when Green cGull was expressed alone **(Figure 4I-K)**.
-To test whether cGMP production is dependent on the NIT domain, we also tested a deletion construct of NIT-GC1 lacking the NIT domain **(Figure 4G)**.
-Cells expressing this construct and Green cGull did not show increased cGMP levels when exposed to SNAP **(Figure 4L)**.
+For balanced expression, we used a single plasmid with the two open-reading frames separated by the 2A self-cleaving peptide (Figure 4H).
+Application of the NO donor SNAP lead to increased cGMP levels, an effect we did not observe when cells were exposed to DMSO or when Green cGull was expressed alone (Figure 4I-K).
+To test whether cGMP production is dependent on the NIT domain, we also tested a deletion construct of NIT-GC1 lacking the NIT domain (Figure 4G).
+Cells expressing this construct and Green cGull did not show increased cGMP levels when exposed to SNAP (Figure 4L).
 These results indicate that NIT-GC1 is able to catalyse cGMP production in an NO-dependent manner and this function requires the NIT domain.
@@
 To test the in vivo function of NIT-GC1 and NIT-GC2 in cPRC responses, we combined calcium imaging with morpholino-mediated knockdowns.
-We used two translation-blocking morpholinos for each *NIT-GC* gene and tested knockdown efficiency by immunostaining injected animals with the NIT-GC1 and NIT-GC2 antibodies **(Figure 4---figure supplement 3E,F)**.
+We used two translation-blocking morpholinos for each *NIT-GC* gene and tested knockdown efficiency by immunostaining injected animals with the NIT-GC1 and NIT-GC2 antibodies (Figure 4---figure supplement 3E,F).
 For both genes, the morpholinos led to a strong reduction in the respective antibody signal, confirming efficient knockdown and antibody specificity.
 
-In NIT-GC1 morphant larvae, the delayed activation of cPRCs following 405 nm stimulation did not occur **(Figure 4M)**.
+In NIT-GC1 morphant larvae, the delayed activation of cPRCs following 405 nm stimulation did not occur (Figure 4M).
 This phenotype is similar to the phenotype of *NOS* mutants suggesting that NIT-GC1 acts as the NO sensor in cPRCs to drive their delayed activation.
 This could occur via increased cGMP production and the opening of a cyclic-nucleotide-gated (CNG) channel specific to cPRCs [@Tosches2014].
-NIT-GC2 morphant larvae, in contrast, showed a step-up increase in calcium following light stimulation **(Figure 4N)**.
+NIT-GC2 morphant larvae, in contrast, showed a step-up increase in calcium following light stimulation (Figure 4N).
 The calcium signal decayed during stimulation and was off after light off.
@@
 The MC cell was identified based on its position and intrinsic activity [@Veraszto2017].
-To unambiguously identify all other cells from which we recorded calcium signals, we developed an on-slide immunostaining method **(Figure 5---figure supplement 1A)**.
-We used the cell-specific antibody markers against RYamide (INNOS) **(Figure 5---figure supplement 1B-D)**, RGWamide (INRGW) and serotonin (Ser-h1) [@Conzelmann2011] to immunostain agar-embedded larvae following calcium imaging.
-Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision **(Figure 5A,B)**.
-Due to the stereotypy of the larvae, we could also identify neurons based on their position and calcium activity in activity-correlation maps **(Figure 5C)**.
+To unambiguously identify all other cells from which we recorded calcium signals, we developed an on-slide immunostaining method (Figure 5---figure supplement 1A).
+We used the cell-specific antibody markers against RYamide (INNOS) (Figure 5---figure supplement 1B-D), RGWamide (INRGW) and serotonin (Ser-h1) [@Conzelmann2011] to immunostain agar-embedded larvae following calcium imaging.
+Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision (Figure 5A,B).
+Due to the stereotypy of the larvae, we could also identify neurons based on their position and calcium activity in activity-correlation maps (Figure 5C).
 
 We first quantified the responses of the INNOS and INRGW interneurons during 405 nm stimulation of the cPRCs.
-In both wild type and *NOS*-mutant larvae, INNOS cells showed an increase in calcium during stimulation **(Figure 5D)**.
-In contrast, the INNOS response was flat or slightly negative in NIT-GC2 morphant larvae **(Figure 5E)** revealing an essential role for NIT-GC2-mediated cPRC suppression in INNOS activation.
+In both wild type and *NOS*-mutant larvae, INNOS cells showed an increase in calcium during stimulation (Figure 5D).
+In contrast, the INNOS response was flat or slightly negative in NIT-GC2 morphant larvae (Figure 5E) revealing an essential role for NIT-GC2-mediated cPRC suppression in INNOS activation.
 INRGW cells were initially inhibited during cPRC stimulation, followed by a delayed activation paralleling the second activation phase of cPRCs.
-This late INRGW response was lacking in *NOS*-mutants **(Figure 5F)**.
-In NIT-GC2 morphants, INRGW cells showed a transient increase in calcium that decayed after light off and a delayed activation was not present **(Figure 5G)**.
+This late INRGW response was lacking in *NOS*-mutants (Figure 5F).
+In NIT-GC2 morphants, INRGW cells showed a transient increase in calcium that decayed after light off and a delayed activation was not present (Figure 5G).
 Next, we imaged calcium signals from Ser-h1 and MC neurons in wild type and *NOS* mutant larvae.
-Ser-h1 cells showed an activation profile that correlated with cPRC activity, including a reduction in calcium during stimulation followed by rebound, a response that was defective in *NOS* mutants **(Figure 5I)**.
-MC cells showed sustained activation, including a late-phase that was lacking in *NOS* mutants **(Figure 5I)**.
+Ser-h1 cells showed an activation profile that correlated with cPRC activity, including a reduction in calcium during stimulation followed by rebound, a response that was defective in *NOS* mutants (Figure 5I).
+MC cells showed sustained activation, including a late-phase that was lacking in *NOS* mutants (Figure 5I).
 These data suggest that during 405 nm stimulation the Ser-h1 cells are inhibited and MC cells are activated, and this regulation is NO-dependent.
@@
 to understand the neurotransmission mechanisms in cPRC, INNOS and INRGWa [@Achim2015; @williams2017].
-Interestingly, in cPRC, whereas cholinergic neurons has previously been implicated in neurotransmission [@Randel2014a], we found transporters and synthetic enzymes involved in glutaminergic, glycinergic, GABAergic and adrenergic innervation, respectively **(Figure 6A)**.
+Interestingly, in cPRC, whereas cholinergic neurons has previously been implicated in neurotransmission [@Randel2014a], we found transporters and synthetic enzymes involved in glutaminergic, glycinergic, GABAergic and adrenergic innervation, respectively (Figure 6A).
 Gene expression has been observed by Randel et al. in these genes, suggesting that they may be expressed in cPRC [@Randel2014a].
-Their receptors were also found for each of the expressed genes within INNOS and INRGWa **(Figure 6A)**.
-Based on these expression patterns and results, a model of transduction across the cPRC pathway focusing on Ca2+, cGMP and NO dynamics is proposed **(Figure 6B)**.
+Their receptors were also found for each of the expressed genes within INNOS and INRGWa (Figure 6A).
+Based on these expression patterns and results, a model of transduction across the cPRC pathway focusing on Ca2+, cGMP and NO dynamics is proposed (Figure 6B).
 The c-opsin1, which is UV-sensitive, activates G proteins (Gi/Go) and decreases cGMP CNGAα, which is already known to be strongly expressed in cPRC as one of the targets highly dependent on cGMP changes, may regulate the influx of ions into the cell [@Arendt2004; @tsukamoto2017; @Veedin2021].

### 2023-05-26 Manuscript/Manuscript.Rmd
MS: edits to discussion, NO in working memory

@@
 Nitric oxide (NO) is one such modulator with unique physical and signalling properties.
-This free radical synthesized by nitric oxide synthase (NOS) is short-lived and can diffuse across biological membranes [@Cudeiro1999; @thomas2015].
-Canonical NO signalling involves the activation of neuronal NOS following calcium increase, NO production and diffusion, and the NO-dependent activation of soluble guanylate cyclases (sGC) in target cells leading to cGMP production [@Bredt1990; @Hoelscher1997].
-Given that NOS activation is calcium dependent and NOS shows neuron-type-specific expression [@Gibbs1998a; @Wildemann1999; @Mobley2022], NO action can lead to the activity-dependent modulation of neural circuits at specific sites [@Wang2007; @Vielma2014; @jacoby2018] .
+This free radical synthesized by nitric oxide synthase (NOS) from L-arginine is short-lived and can diffuse across biological membranes [@Cudeiro1999; @thomas2015].
+Canonical NO signalling involves the Ca2+/calmodulin-dependent activation of NOS following calcium increase, NO production and diffusion, and the NO-dependent activation of soluble guanylate cyclases (sGC) cell-autonomously or in other cells leading to cGMP production [@Bredt1990; @Hoelscher1997].
+Given that NOS activation is calcium dependent and NOS shows neuron-type-specific expression [@Gibbs1998a; @Wildemann1999; @Aso2019; @Mobley2022], NO action can lead to the activity-dependent modulation of neural circuits at specific sites [@Wang2007; @Vielma2014; @jacoby2018; @Aso2019] .
 
@@
 In immunostainings, we found that NIT-GC1 was localise to the region corresponding to the axonal projections of the cPRCs in the anterior nerve plexus (Figure 4E).
-Co-immunostaining with the rabbit NIT-GC1 antibody and a custom rat antibody raised against *Platynereis* NOS revealed the localisation of both proteins in close proximity in the neurosecretory plexus (Figure 4---figure supplement 3D).
+Co-immunostaining with the rabbit NIT-GC1 antibody and a custom rat antibody raised against *Platynereis* NOS revealed the localisation of both proteins in dots in the neurosecretory neuropil (Figure 4---figure supplement 3D).
 In contrast, NIT-GC2 specifically labelled the ramified sensory cilia of the cPRCs (Figure 4F).
@@
 NO is readily converted to nitrate and nitrite [@santos2011; @moeller2019; @garthwaite2015a] and these molecules accumulate in placozoans and cnidarians in cells and tissues with high NOS activity [@Moroz2004; @Moroz2020].
-It is possible that the NIT domains in NIT-GCs may also sense nitrate and nitrite, as in bacteria.
-Furthermore, if they show different sensitivities to NO, nitrite and nitrate, it is possible that differences in their half-lives could give rise to a range of activation timings [@lundberg2011].
+NIT domains in NIT-GCs may also sense nitrate and nitrite, as in bacteria.
+Furthermore, if different NIT-GCs show different sensitivities to NO, nitrite and nitrate, it is possible that differences in the half-lives of these molecules could give rise to a range of activation timings [@lundberg2011].
 We have found 15 NIT-GCs in *Platynereis*, suggesting a wide range of functions.
@@
 Overall, the diversification and spatially different localisation patterns of NIT-GCs may increase their signalling repertoire.
-Our data revealed key roles for two *Platynereis* NIT-GCs in phototraonsduction and NO-mediated neuromodulation suggesting that these molecules contribute to the diversity of neuronal signalling.
+Our data revealed key roles for two *Platynereis* NIT-GCs in phototransduction and NO-mediated neuromodulation suggesting that these molecules contribute to the diversity of neuronal signalling.
 Sponges and ctenophores, where NO signalling is present, lack NIT-GCs.
 In placozoans as many as 12 have been found, and cnidarians also contain a large diversity.
+Soluble GCs are expressed broadly in the *Drosophila* and mammalian brain and the specificity of NO signalling seems to depend on localised NOS expression and activation, and ectopic NOS expression can confer NO-dependent effects [@Aso2019].
+In *Platyneries* larvae
 
-**Nitric oxide feedback modulation of UV-avoidance behaviour**
+**Nitric oxide confers short-term memory to circuit activity**
 
@@
 In Drosophila, NOS of PPL1-γ1pedc shortens memory retention while promoting fast memory updating in response to new experiences, indicating that NO regulates forgetting [@Aso2019].
-Because NO has a short half-life and may be a suitable second messenger for encoding working memories whose duration must be limited.
-R-neuron inputs to the ellipsoid in Drosophila also represent visual object orientation for tens of seconds after the disappearance of an object, suggesting that they are involved in short-term visual memory dependent on nitric oxide signalling [@Kuntz2017].
+In the ellipsoid body of the *Drosophila* central brain, NOS is specifically expressed in the R3 ring neuron, but not other ring (R) neurons [@Kuntz2017].
+R3 neurons are required for visual working memory with NO signalling mediating a short-term (~4 sec) component of the visual memory of objects.
+In this circuit, NO produced in R3 axons acts in an autocrine manner to activate sGC in R3 axons leading to a CNG-dependent (cyclic nucleotide-gated ion channel) temporary increase in calcium levels [@Kuntz2017].
+Because of the short life-time of NO, this molecule may be well suited to encode transient memory traces [@Kuntz2017].
+
 In this study, our mathematical modelling results focused on the time of UV stimulation and showed that the late activity comes more delayed during short UV and sooner during long UV.
@@
 This work was funded by the Wellcome Trust (214337/Z/18/Z).
-This project has received funding from the European Research Council (ERC) under the European Union’s Horizon 2020 research and innovation programme (grant agreement No 101020792).
+This project has received funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement No 101020792).
 KJ has been supported by a JSPS Postdoctoral fellowhip, LAYG by a BBSRC Discovery fellowship (BB/W010305/1).

### 2023-05-26 Manuscript/Manuscript.Rmd
MS: edit math model

@@
 **Mathematical modelling of cPRC-circuit dynamics**
+
+To further analyse the dynamics of responses to UV light and formally describe cPRC phototransduction, synaptic connections and NO retrograde signalling, we developed a mixed cellular-circuit-level mathematical model.
+We used our Ca2+ imagining data of cPRC, INNOS and INRGW cells collected in wild type, *NOS* knockout, and *NIT-GC2* morphant larvae to formulate assumptions that are the basis of the proposed model.
+
+We model a direct c-opsin1-dependent response to UV leading to the initial rise in Ca2+ levels.
+The mechanism of this signal is not known but may include Galpha and Gbeta signalling, CNG channel activation or other pathways.
+We further model a decrease following UV stimulation in Ca2+ levels that is dependent on NIT-GC2.
+The
+
+in cPRC cells increase with cGMP. cGMP decreases under NIT-GC2
+exposure to UV (mechanism unknown, comparison of Fig. 4A, B and M with Fig. 4N, described
+in section ’NIT-GC1 is required for NO-mediated feedback to cPRCs during the UV response’).
+cGMP increases with increasing NO (comparison of Fig. 4A with Fig. 4B described in sections
+’NO feedback tunes cPRC responses to UV/violet stimulation’ and ’NIT-GC1 produces cGMP in a
+NO-dependent manner’ and ’NIT-GC1 is required for NO-mediated feedback to cPRCs during the
+UV response’).
+Based on the connectome structure (Fig. 1F) we infer a feedforward coupling between cPRC
+cells and IN-NOS and IN-RGW cells. The proposed synaptic coupling is UV-dependent and decays
+linearly. It is suppressed in NIT-GC2 morphants (observations in Fig. 5
 

### 2023-05-29 Manuscript/Manuscript.Rmd
MS: modelling section edit

--- 2023-05-26 Manuscript/Manuscript.Rmd
@@
 **Mathematical modelling of cPRC-circuit dynamics**
+
+To further analyse the dynamics of responses to UV light and formally describe cPRC phototransduction, synaptic connections and NO retrograde signalling, we developed a mixed cellular-circuit-level mathematical model.
+We used our Ca2+ imagining data of cPRC, INNOS and INRGW cells collected in wild type, *NOS* knockout, and *NIT-GC2* morphant larvae to formulate assumptions that are the basis of the proposed model.
+
+We model a direct c-opsin1-dependent response to UV leading to the initial rise in Ca2+ levels.
+The mechanism of this signal is not known but may include Gα and Gβ signalling, CNG channel activation or other pathways.
+We further assume that in cPRC cells, Ca2+ levels increase with cGMP.
+We model a decrease in cGMP and Ca2+ levels due to a NIT-GC2-dependent pathway.
+The effect of NO in cPRC cells is captured by another term that describes a NIT-GC1 and NO-dependent increase in cGMP.
+Based on the synaptic connectome, we infer a feedforward coupling between cPRC cells and INNOS and INRGW cells.
+The proposed synaptic coupling is UV-dependent, decays linearly and is suppressed in NIT-GC2 morphants.
 

### 2023-05-31 Manuscript/Manuscript.Rmd

@@
 
-We analysed the neurotransmitter-related genes expressed in the respective cells identified from the single-cell transcriptome data of Achim et al.
-to understand the neurotransmission mechanisms in cPRC, INNOS and INRGWa [@Achim2015; @williams2017].
-Interestingly, in cPRC, whereas cholinergic neurons has previously been implicated in neurotransmission [@Randel2014a], we found transporters and synthetic enzymes involved in glutaminergic, glycinergic, GABAergic and adrenergic innervation, respectively (Figure 6A).
-Gene expression has been observed by Randel et al. in these genes, suggesting that they may be expressed in cPRC [@Randel2014a].
-Their receptors were also found for each of the expressed genes within INNOS and INRGWa (Figure 6A).
-Based on these expression patterns and results, a model of transduction across the cPRC pathway focusing on Ca2+, cGMP and NO dynamics is proposed (Figure 6B).
-The c-opsin1, which is UV-sensitive, activates G proteins (Gi/Go) and decreases cGMP CNGAα, which is already known to be strongly expressed in cPRC as one of the targets highly dependent on cGMP changes, may regulate the influx of ions into the cell [@Arendt2004; @tsukamoto2017; @Veedin2021].
-The Gβγ released by Gi/o activation also induces K+ efflux (hyperpolarisation) via GIRK channels and closes the voltage-gated Ca2+ channels, thereby reducing intracellular calcium levels[@tsukamoto2023].
-It is also thought that increased calcium levels in INNOS signalled from cPRC may activate NOS and generate nitric oxide; INNOS has glycinergic and glutamatergic potential and may act as an inhibitory signal regulator for INRGWa.
-INRGWa, and may function as an inhibitory signal modulator for INRGWa.
+In the INNOS cells, we assume that excitatory synaptic input from cPRC leads to a rise in Ca2+ leading to NO production via NOS.
+This input is NIT-GC2-dependent.
+In the INRGW cells, we model a decrease in Ca2+ based on an inhibitory UV-dependent synaptic signal from the cPRCs.
+We further assume the existence of direct excitatory coupling between Ca2+ level in cPRC and INRGW Ca2+ to account for the late effects of cPRC on INRGW.
+The Uv-dependent and Ca2+-dependent signals may be mediated by different transmitters released by cPRCs during distinct phases of their activation cycle.
+In addition, we assume a direct inhibitory coupling between INNOS Ca2+ and INRGW Ca2+.
+Finally, in all variables, we assumed a linear decay and constant production to set steady a state.
+Since the aim of the model is to capture the normalised fluorescence data, the model is nondimensionalised.
+UV stimulation is modelled as a square pulse.
+To find model parameters producing an output fitting the experimental data we employed a global optimisation method known as a genetic algorithm (GA).
 
-We used this model and calcium imaging data to develop a mathematical model.
+Our model only includes a minimal set of parameters and assumptions of interactions that are required to describe the dynamics of cPRC, INNOS and INRGW in wild type and loss-of-function conditions (Figure 6C, D).
+The model highlights that cPRC phototransduction employs distinct pathways that operate on different time scales and differentially influence cPRC Ca2+ levels.
+The coupling between cPRCs and interneurons also requires different signalling mechanisms, either through distinct neurotransmitters or receptors expressed in the different cells.
+
+To identify possible molecular pathways, we analysed single-cell transcriptome data for cPRC, INNOS and INRGW identified by spatial mapping [@Achim2015] and unique marker genes [@williams2017].
+In cPRCs, we found transporters and synthetic enzymes indicating cholinergic, glutaminergic, glycinergic, GABAergic and adrenergic neurotransmission (Figure 6A).
+For each neurotransmitter, we also found receptor-encoding genes expressed in INNOS and INRGW (Figure 6A).
+The two types of interneurons often expressed different subunits or types of these receptors, indicating differential signalling (Figure 6A).
+
+Based on these data, our experimental results and the mathematical model, we assembled a minimal phototransduction and circuit diagram.
+The components for which experimental data are available are shown in bold.
+Other potential molecular players are also indicated (Figure 6B).
 
 ```{r echo=FALSE,fig.cap="**Figure 6.
-Signalling mechanisms and mathematical modelling between cPRC, INNOS and INRGWa.** **(A)** Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq.
+Signalling mechanisms and mathematical modelling between cPRC, INNOS and INRGW.** **(A)** Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq.
 The size of the dots is expressed in proportion to the percentage of cells expressing that gene relative to all cells.
@@
 
-Our work revealed an essential role for NO-mediated signalling in driving UV/violet avoidance behaviour.
-NO, produced by postsynaptic INNOS interneurons, signals retrogradely to the presynaptic cPRCs leading to their delayed activation.
-The resulting NIT-GC1-dependent late-phase activation of the photoreceptors drives circuit output through projection interneurons and ciliomotor neurons.
+Our work revealed an essential role for NO-mediated signalling in driving UV/violet avoidance behaviour in larval *Platynereis*.
+NO, produced by postsynaptic INNOS interneurons, signals retrogradely to the presynaptic cPRCs via NIT-GC1 leading to delayed and sustained cPRC activation.
+This late-phase activation of the photoreceptors drives circuit output through projection interneurons and ciliomotor neurons.
 In the cPRC circuit, synaptic connectivity alone is thus not sufficient to drive circuit dynamics and behavioural change.
+
 We identified cPRC-expressed NIT-GC1 as a new type of NO sensor.
+We showed that NIT-GC1 can produce cGMP in a NO-dependent manner.
 NIT-GC1 is localised to the axonal projections of the cPRCs in the anterior neurosecretory plexus of the brain.
@@
 NO-mediated retrograde signalling thus likely occurs in the neurosecretory plexus where NOS- and NIT-GC1-containing projections are in close proximity and where we could detect NO production following UV stimulation.
-We showed that NIT-GC1 can produce cGMP in a NO-dependent manner.
 In cPRCs, increased cGMP levels could open CNGAα, a cyclic nucleotide-gated channels specifically expressed in cPRC [@Tosches2014], leading to their delayed depolarisation, even after UV off.
-
-Also our mathematical model...
 
@@
 
-NO is a free radical with a millisecond-to-second scale half life in vivo and a limited signalling range.
-In neuronal signalling, NOS and sGC have been found to localise at close proximity at synapses [@burette2002][@garthwaite2015a].
+NO is a free radical with a millisecond-to-second half life and thus a limited signalling range.
+In neuronal signalling, NOS is often localized to neurites [@Kuntz2017] at close proximity to sGC at synapses [@burette2002; @garthwaite2015a].
 We identified 12 sGCs in *Platynereis*, but none of these is expressed in the cPRC.
@@
 NO is readily converted to nitrate and nitrite [@santos2011; @moeller2019; @garthwaite2015a] and these molecules accumulate in placozoans and cnidarians in cells and tissues with high NOS activity [@Moroz2004; @Moroz2020].
-NIT domains in NIT-GCs may also sense nitrate and nitrite, as in bacteria.
-Furthermore, if different NIT-GCs show different sensitivities to NO, nitrite and nitrate, it is possible that differences in the half-lives of these molecules could give rise to a range of activation timings [@lundberg2011].
-We have found 15 NIT-GCs in *Platynereis*, suggesting a wide range of functions.
-In addition, NIT-GC1 and NIT-GC2 showed very different subcellular localisation and functions.
-It is not clear whether the functional differences are due to the different localisation or also to differences in biochemical function.
-Overall, the diversification and spatially different localisation patterns of NIT-GCs may increase their signalling repertoire.
-Our data revealed key roles for two *Platynereis* NIT-GCs in phototransduction and NO-mediated neuromodulation suggesting that these molecules contribute to the diversity of neuronal signalling.
-Sponges and ctenophores, where NO signalling is present, lack NIT-GCs.
-In placozoans as many as 12 have been found, and cnidarians also contain a large diversity.
-Soluble GCs are expressed broadly in the *Drosophila* and mammalian brain and the specificity of NO signalling seems to depend on localised NOS expression and activation, and ectopic NOS expression can confer NO-dependent effects [@Aso2019].
-In *Platyneries* larvae
+NIT domains in NIT-GCs may also sense nitrate and nitrite, as in bacteria, a possibility we cannot rule out based on our cellular assays with NIT-GC1.
+If different NIT-GCs show different sensitivities to NO, nitrite and nitrate, then the differences in the half-lives of these molecules may give rise to a range of activation timings [@lundberg2011].
+
+In *Platynereis*, we have found 15 NIT-GCs, suggesting a wide range of functions.
+We only tested NIT-GC1 and NIT-GC2, and the two proteins showed very different subcellular localisation and functions.
+Differences in cellular expression, subcellular localisation, and biochemical function may thus all contribute to the diversity of NIT-GC functions.
+
+**Mechanism of phototransduction and neurotransmission in the cPRC circuit**
+
+Based on our data herein and previous work we can now propose a more detailed model of cPRC phototransduction and neurotransmission.
+The cPRCs have high basal Ca2+ and respond to UV/violet light dependent on c-opsin1 [@Veraszto2018].
+c-opsin1 forms a bistable photopigment and signals through Gi/oα and Gβγ [@tsukamoto2017; @Veedin2021; @tsukamoto2023].
+In heterologous systems, the Gβγ subunits released following c-opsin1 activation activate GIRK channels inducing K+ efflux (hyperpolarisation) [@tsukamoto2017; @tsukamoto2023] and closes the voltage-gated Ca2+ channels, thereby reducing intracellular calcium levels [@tsukamoto2023].
+An orthologous GIRK channel is also expressed at low levels in the cPRCs (Figure 6A).
+The pathway for the first rapid activation phase following UV/violet stimulus is not known but may involve the activation of the CNGα channel expressed in the cPRCs.
+For the second inhibitory phase of phototransduction, we identified a key requirement for the ciliary-localised NIT-GC2.
+The mechanisms may involve c-opsin1-dependent inhibition of tonic NIT-GC2 activity and the reduction of ciliary cGMP.
+This signalling step is required for the feedforward activation of the INNOS cells through an unknown transmitter.
+The activation of INNOS leads to NOS activation and NO release, potentially through a canonical Ca2+-calmodulin pathway.
+Our model predicts feedforward inhibiton from INNOS to INRGW, possibly mediated by glycin transmission.
+NO released by INNOS neurites activates NIT-GC1 in cPRC projections, leading to delayed cPRC activation that is not directly dependent on the UV/violet signal.
+This late activation phase leads to INRGW activation, presumably via a different neurotransmitter.
+In addition, the cPRC circuit expresses neuropeptides and their receptors, suggesting further neuromodulatory signals.
+For example, INRGW express the proneuropeptide RGWamide and cPRCs and INNOS express its receptor, suggesting retrograde peptidergic signalling.
 
@@
 In the ellipsoid body of the *Drosophila* central brain, NOS is specifically expressed in the R3 ring neuron, but not other ring (R) neurons [@Kuntz2017].
-R3 neurons are required for visual working memory with NO signalling mediating a short-term (~4 sec) component of the visual memory of objects.
+R3 neurons are required for visual working memory with NO signalling mediating a short-term (\~4 sec) component of the visual memory of objects.
 In this circuit, NO produced in R3 axons acts in an autocrine manner to activate sGC in R3 axons leading to a CNG-dependent (cyclic nucleotide-gated ion channel) temporary increase in calcium levels [@Kuntz2017].
@@
 
-One type of interneuron that strongly synapses with the cPRC is INRGWa, which synapses directly onto ciliary motoneurons and is therefore considered to be a very important neuron in larval behavioural control [@williams2017; @Veraszto2018].
+One type of interneuron that strongly synapses with the cPRC is INRGW, which synapses directly onto ciliary motoneurons and is therefore considered to be a very important neuron in larval behavioural control [@williams2017; @Veraszto2018].
 Our results suggest that the depolarising response of the cPRC by NO feedback induces INRGW activation.
@@
 Guanylate cyclases with NIT domains are found in most animal phyla except Porifera, Ctenophora, Urochordata and Chordata.
-Dot plot of Platynereis NIT-GC genes (columns) expressed in cPRC, INNOS and INRGWa (rows) using single cell RNA-Seq.
+Dot plot of Platynereis NIT-GC genes (columns) expressed in cPRC, INNOS and INRGW (rows) using single cell RNA-Seq.
 The size of the dots is expressed in proportion to the percentage of cells expressing that gene relative to all cells.

### 2023-05-31 Manuscript/Manuscript.Rmd
Ms: discussion edits of NO retrograde section

@@
 This phenotype is similar to the phenotype of *NOS* mutants suggesting that NIT-GC1 acts as the NO sensor in cPRCs to drive their delayed activation.
-This could occur via increased cGMP production and the opening of a cyclic-nucleotide-gated (CNG) channel specific to cPRCs [@Tosches2014].
+This could occur via increased cGMP production and the opening of a cyclic-nucleotide-gated ion channel (CNG) specific to cPRCs [@Tosches2014].
 NIT-GC2 morphant larvae, in contrast, showed a step-up increase in calcium following light stimulation (Figure 4N).
@@
 We model a direct c-opsin1-dependent response to UV leading to the initial rise in Ca2+ levels.
-The mechanism of this signal is not known but may include Gα and Gβ signalling, CNG channel activation or other pathways.
+The mechanism of this signal is not known but may include Gα and Gβ signalling, CNGAα channel activation or other pathways.
 We further assume that in cPRC cells, Ca2+ levels increase with cGMP.
@@
 NO-mediated retrograde signalling thus likely occurs in the neurosecretory plexus where NOS- and NIT-GC1-containing projections are in close proximity and where we could detect NO production following UV stimulation.
-In cPRCs, increased cGMP levels could open CNGAα, a cyclic nucleotide-gated channels specifically expressed in cPRC [@Tosches2014], leading to their delayed depolarisation, even after UV off.
+In cPRCs, increased cGMP levels could open CNGAα, specifically expressed in cPRC [@Tosches2014], leading to their delayed depolarisation, even after UV off.
 
@@
 
-Why does a seemingly roundabout feedback pathway exist?
-The activity of synaptic circuits is richly modulated monoamines, neuropeptides and gaseous molecules.
-The gaseous second messenger nitric oxide (NO) is known to activate a post- to presynaptic retrograde signalling cascade involving cyclic guanosine monophosphate (cGMP) production.
-It has been reported that NO produced in one of the mammalian retinal cells modulates cGMP-mediated neurotransmission by activating the presynaptic nerve sGC (a general target of NO) through retrograde signalling [@Vielma2014; @Wei2012].
-Retrograde signalling of NO is also known to be one of the molecules that regulate memory formation.
+Retrograde signalling by NO from INNOS to cPRC leads to the sustained activation of cPRCs and postsynaptic neurons even after the end of stimulation.
+This activated state is maintained for several tens of second.
+NO signalling thus induces a transient circuit state orshort-term memory trace in the *Platynereis* larval brain.
+Because of the short life-time of NO, this molecule may be well suited to encode transient memory traces [@Kuntz2017].
+In the ellipsoid body of the *Drosophila* central brain, NO signalling has a similar function.
+Here, NOS is specifically expressed in the R3 ring neurons and is required for the short-term (\~4 sec) visual memory of objects [@Kuntz2017].
+NO is produced in the axons of R3 neurons and acts directly on sGC in the same axons.
+This autocrine signal leads to a CNG-dependent temporary increase in calcium levels, carrying the working memory trace [@Kuntz2017].
+
+In the *Platynereis* circuit, our mathematical model indicates that the onset, magnitude and duration of the NO-dependent signal depends on the intensity and duration of the UV/violet stimulus.
+This suggests that the NO-dependent memory trace also encodes the magnitude and duration of the stimulus.
+During UV avoidance in the planarian *Schmidtea mediterranea*, neuropeptide signalling has a similar integratory function.
+Planarians exposed to UV light for >30 sec remain active for extended periods (several minutes) post-stimulation [@bray2023].
+If neuropeptide signalling is defective, animals can still respond to UV light, but do not maintain a latent memory state and do not display increased post-stimulus activity.
+
+NO signalling may also work on longer time scales. In the fly ellipsoid body, NO signalling is also required for long-term changes of the circuit, acting through CREB-dependent transcription.
+
 In Drosophila, NOS of PPL1-γ1pedc shortens memory retention while promoting fast memory updating in response to new experiences, indicating that NO regulates forgetting [@Aso2019].
-In the ellipsoid body of the *Drosophila* central brain, NOS is specifically expressed in the R3 ring neuron, but not other ring (R) neurons [@Kuntz2017].
-R3 neurons are required for visual working memory with NO signalling mediating a short-term (\~4 sec) component of the visual memory of objects.
-In this circuit, NO produced in R3 axons acts in an autocrine manner to activate sGC in R3 axons leading to a CNG-dependent (cyclic nucleotide-gated ion channel) temporary increase in calcium levels [@Kuntz2017].
-Because of the short life-time of NO, this molecule may be well suited to encode transient memory traces [@Kuntz2017].
-
-In this study, our mathematical modelling results focused on the time of UV stimulation and showed that the late activity comes more delayed during short UV and sooner during long UV.
-This suggests that animals may have a UV length-dependent short-term memory system that regulates cPRC activation, and that there may be mechanisms by which UV stimulation is enhanced and maintained such that NO feedback maintains neural activation even with slight UV.
-It is possible that these mechanisms may have been early signalling pathways controlling stimulus changes across space and time in animals' short-term memory.
-In planarian UV avoidance behaviour, different time and length scales of neuropeptide and small molecule transmission were reported to generate incoherent patterns of neural activity that competitively control behaviour and memory [@bray2023].
-Short-term memory is essential for locating less exposed areas to minimise UV-induced damage in animals.
 

### 2023-05-31 Manuscript/Manuscript.Rmd
MS: discussion gravity section

@@
 Our work revealed an essential role for NO-mediated signalling in driving UV/violet avoidance behaviour in larval *Platynereis*.
-NO, produced by postsynaptic INNOS interneurons, signals retrogradely to the presynaptic cPRCs via NIT-GC1 leading to delayed and sustained cPRC activation.
+NO, produced by postsynaptic INNOS interneurons, signals retrogradely to presynaptic cPRCs via NIT-GC1 leading to delayed and sustained cPRC activation.
 This late-phase activation of the photoreceptors drives circuit output through projection interneurons and ciliomotor neurons.
-In the cPRC circuit, synaptic connectivity alone is thus not sufficient to drive circuit dynamics and behavioural change.
+In the cPRC circuit, synaptic connectivity alone is thus not sufficient to account for circuit dynamics and behavioural change.
 
-We identified cPRC-expressed NIT-GC1 as a new type of NO sensor.
-We showed that NIT-GC1 can produce cGMP in a NO-dependent manner.
-NIT-GC1 is localised to the axonal projections of the cPRCs in the anterior neurosecretory plexus of the brain.
-Here, cPRC axons intermingle and synapse on INNOS neurites that contain NOS.
+**Localised NO signalling in the neurosecretory plexus**
+
+NO is a free radical with a millisecond-to-second half life and thus a limited signalling range.
+In neuronal signalling, NOS is often localized to neurites [@Kuntz2017] at close proximity to sGC at synapses [@burette2002; @garthwaite2015a].
+In the cPRC circuit, NOS is localised to the input compartment of INNOS neurites in the neurosecretory plexus where cPRC axons synapse on INNOS neurites.
+NIT-GC1, the target of NO signalling is localised to cPRC projections.
 NO-mediated retrograde signalling thus likely occurs in the neurosecretory plexus where NOS- and NIT-GC1-containing projections are in close proximity and where we could detect NO production following UV stimulation.
-In cPRCs, increased cGMP levels could open CNGAα, specifically expressed in cPRC [@Tosches2014], leading to their delayed depolarisation, even after UV off.
 
@@
 
-NO is a free radical with a millisecond-to-second half life and thus a limited signalling range.
-In neuronal signalling, NOS is often localized to neurites [@Kuntz2017] at close proximity to sGC at synapses [@burette2002; @garthwaite2015a].
-We identified 12 sGCs in *Platynereis*, but none of these is expressed in the cPRC.
-Instead, we identified an unconventional NIT-domain containing GC, NIT-GC1 as the mediator of NO retrograde signalling.
-The NIT domain was first identified in bacteria and animal NIT-GCs have only been recently reported [@Shu2003; @Moroz2020].
+NO signalling is commonly mediated by sGCs. We identified 12 sGCs in *Platynereis*, but none of these is expressed in the cPRC.
+Instead, we identified an unconventional cPRC-expressed NIT-domain containing GC, NIT-GC1 as the mediator of NO retrograde signalling.
+In an in vitro assay, we could show that NIT-GC1 can produce cGMP following the addition of an NO donor and that this activity requires the NIT domain.
+
+NIT domains were first identified in bacteria and animal NIT-GCs have only recently been reported [@Shu2003; @Moroz2020].
 Bacterial NIT domains regulate cellular functions in response to changes in extracellular and intracellular nitrate and/or nitrite concentrations [@camargo2007].
@@
 c-opsin1 forms a bistable photopigment and signals through Gi/oα and Gβγ [@tsukamoto2017; @Veedin2021; @tsukamoto2023].
-In heterologous systems, the Gβγ subunits released following c-opsin1 activation activate GIRK channels inducing K+ efflux (hyperpolarisation) [@tsukamoto2017; @tsukamoto2023] and closes the voltage-gated Ca2+ channels, thereby reducing intracellular calcium levels [@tsukamoto2023].
-An orthologous GIRK channel is also expressed at low levels in the cPRCs (Figure 6A).
-The pathway for the first rapid activation phase following UV/violet stimulus is not known but may involve the activation of the CNGα channel expressed in the cPRCs.
-For the second inhibitory phase of phototransduction, we identified a key requirement for the ciliary-localised NIT-GC2.
+In heterologous systems, the Gβγ subunits released following c-opsin1 activation open GIRK channels inducing K+ efflux (hyperpolarisation) [@tsukamoto2017; @tsukamoto2023] and close voltage-gated Ca2+ channels, thereby reducing intracellular calcium levels [@tsukamoto2023].
+A GIRK channel is also expressed in the cPRCs (Figure 6A).
+The pathway for the first rapid cPRC activation phase following UV/violet stimulus is not known but may involve the activation of a CNGα channel expressed in the cPRCs [@Tosches2014].
+For the second inhibitory phase of phototransduction, we identified a key requirement for ciliary-localised NIT-GC2.
 The mechanisms may involve c-opsin1-dependent inhibition of tonic NIT-GC2 activity and the reduction of ciliary cGMP.
-This signalling step is required for the feedforward activation of the INNOS cells through an unknown transmitter.
+
+NIT-GC2-dependent signalling is required for the feedforward activation of the INNOS cells through an unknown transmitter.
 The activation of INNOS leads to NOS activation and NO release, potentially through a canonical Ca2+-calmodulin pathway.
 Our model predicts feedforward inhibiton from INNOS to INRGW, possibly mediated by glycin transmission.
-NO released by INNOS neurites activates NIT-GC1 in cPRC projections, leading to delayed cPRC activation that is not directly dependent on the UV/violet signal.
-This late activation phase leads to INRGW activation, presumably via a different neurotransmitter.
-In addition, the cPRC circuit expresses neuropeptides and their receptors, suggesting further neuromodulatory signals.
-For example, INRGW express the proneuropeptide RGWamide and cPRCs and INNOS express its receptor, suggesting retrograde peptidergic signalling.
+NO released by INNOS neurites activates NIT-GC1 in cPRC projections that could lead to cGMP production and the opening CNGAα.
+This late-phase cPRC activation is not directly dependent on the UV/violet signal and can happen post-stimulus.
+The late-phase cPRC activation leads to INRGW activation via an unknown transmitter that is likely different from the one acting on INNOS.
+
+In addition, the cPRC circuit expresses several neuropeptides and their receptors, suggesting further neuromodulatory signals.
+For example, INRGW express the proneuropeptide RGWamide and cPRCs and INNOS express its receptor, suggesting retrograde peptidergic signalling in the circuit.
 
@@
 This activated state is maintained for several tens of second.
-NO signalling thus induces a transient circuit state orshort-term memory trace in the *Platynereis* larval brain.
-Because of the short life-time of NO, this molecule may be well suited to encode transient memory traces [@Kuntz2017].
+NO signalling thus induces a transient circuit state or short-term memory trace in the *Platynereis* larval brain.
+Because of the short life time of NO, this molecule may be well suited to encode transient memory traces [@Kuntz2017].
 In the ellipsoid body of the *Drosophila* central brain, NO signalling has a similar function.
@@
 
-**The mystery of why IN-RGW activation causes downward swimming**
+**The possible mechanism of UV-induced downward swimming**
 
-One type of interneuron that strongly synapses with the cPRC is INRGW, which synapses directly onto ciliary motoneurons and is therefore considered to be a very important neuron in larval behavioural control [@williams2017; @Veraszto2018].
-Our results suggest that the depolarising response of the cPRC by NO feedback induces INRGW activation.
-Our experiments have shown that exposure of larvae to RGWamide results in simultaneous downward swimming and that activation of INRGW is sufficient to trigger UV-avoidance behaviour (unpublished data).
-However, whole-cell tracking by connectome has not revealed any equilibrium organ that recognises the direction of gravity, and it remains to be investigated how activation of INRGWs causes animals to recognise 'down', e.g.
-in RGWamide-deficient larvae.
+UV/violet light induces downward swimming head down [@Veraszto2018]. Since the direction of UV/violet light is not important for this, swimming direction must be determined by gravity.
+
+Connectome reconstruction and annotation of all cells in the three-day-old *Platynereis* larva has not revealed any balancer organ to sense orientation in the gravity field [@veraszto2020].
+It follows that body orientation must be determined by physical mechanisms including the buoyancy and centre of mass of the larva, body shape, drag and differential ciliary activity.
+Except for ciliary activity, all other parameters are likely invariant during the UV response.
+
+The only motor output of the cPRC circuit in the wiring diagram is the prototroch ciliary band postsynaptic to the MC and Ser-h1 ciliomotor neurons.
+Our Ca2+-imaging data revealed a late NO-dependent inhibition of Ser-h1 and activation of the MC cell suggesting the inhibition of ciliary beating [@veraszto2017].
+
 It is known that complex integrative networks of neurons are involved in UV avoidance behaviour in planaria [@Shettigar2021] and Drosophila [@Imambocus2022].

### 2023-06-02 Manuscript/Manuscript.Rmd
Ms: discussion edit, gravity section

@@
 The only motor output of the cPRC circuit in the wiring diagram is the prototroch ciliary band postsynaptic to the MC and Ser-h1 ciliomotor neurons.
-Our Ca2+-imaging data revealed a late NO-dependent inhibition of Ser-h1 and activation of the MC cell suggesting the inhibition of ciliary beating [@veraszto2017].
+Our Ca2+-imaging data revealed a NO-dependent inhibition of Ser-h1 and activation of the MC cell suggesting the inhibition of ciliary beating [@veraszto2017].
 

### 2023-06-02 Manuscript/Manuscript.Rmd
Ms: edit discussion compartmentalisation section

@@
 This late-phase activation of the photoreceptors drives circuit output through projection interneurons and ciliomotor neurons.
-In the cPRC circuit, synaptic connectivity alone is thus not sufficient to account for circuit dynamics and behavioural change.
+In the cPRC circuit, synaptic connectivity alone is thus not sufficient to account for circuit dynamics and behavioural change, as documented in other circuits [@Imambocus2022].
 
@@
 In neuronal signalling, NOS is often localized to neurites [@Kuntz2017] at close proximity to sGC at synapses [@burette2002; @garthwaite2015a].
-In the cPRC circuit, NOS is localised to the input compartment of INNOS neurites in the neurosecretory plexus where cPRC axons synapse on INNOS neurites.
+In the cPRC circuit, NOS is localised to the dendritic compartment of INNOS neurites in the neurosecretory plexus where cPRC axons synapse on INNOS neurites.
 NIT-GC1, the target of NO signalling is localised to cPRC projections.
@@
 
+Compartmentalised signalling also occurs in peptidergic modulatory system and can be important for selective network activity during specific behaviours.
+In the UV-avoidance circuit of *Drosophila* larvae, a peptidergic hub neuron Dp7 integrates different sensory inputs occurring at functionally separated dendritic compartments.
+The insulin-like peptide 7 (Ilp7) expressed in Dp7 mediates the UV signal from extraocular sensory neurons to motor circuits and is required for acute and sustained UV avoidance [@Imambocus2022].
+
 **Functional diversity of NIT-GCs**
 
-NO signalling is commonly mediated by sGCs. We identified 12 sGCs in *Platynereis*, but none of these is expressed in the cPRC.
+NO signalling is commonly mediated by sGCs.
+We identified 12 sGCs in *Platynereis*, but none of these is expressed in the cPRC.
 Instead, we identified an unconventional cPRC-expressed NIT-domain containing GC, NIT-GC1 as the mediator of NO retrograde signalling.
@@
 During UV avoidance in the planarian *Schmidtea mediterranea*, neuropeptide signalling has a similar integratory function.
-Planarians exposed to UV light for >30 sec remain active for extended periods (several minutes) post-stimulation [@bray2023].
+Planarians exposed to UV light for \>30 sec remain active for extended periods (several minutes) post-stimulation [@bray2023].
 If neuropeptide signalling is defective, animals can still respond to UV light, but do not maintain a latent memory state and do not display increased post-stimulus activity.
 
-NO signalling may also work on longer time scales. In the fly ellipsoid body, NO signalling is also required for long-term changes of the circuit, acting through CREB-dependent transcription.
+NO signalling may also work on longer time scales.
+In the fly ellipsoid body, NO signalling is also required for long-term changes of the circuit, acting through CREB-dependent transcription.
 
@@
 
-UV/violet light induces downward swimming head down [@Veraszto2018]. Since the direction of UV/violet light is not important for this, swimming direction must be determined by gravity.
+UV/violet light induces downward swimming head down [@Veraszto2018].
+Since the direction of UV/violet light is not important for this, swimming direction must be determined by gravity.
 
@@
 
-It is known that complex integrative networks of neurons are involved in UV avoidance behaviour in planaria [@Shettigar2021] and Drosophila [@Imambocus2022].
-It has been reported that the sensory neurons required for UV avoidance behaviour in Drosophila converge and arrest neuromodulatory hub neurons, and that compartmentalised circuit organisation and neuropeptide release from the regulatory hubs act as central circuit elements controlling escape responses.
+It is known that complex integrative networks of neurons are involved in UV avoidance behaviour in planaria [@Shettigar2021]
+
 RGW interneurons have been suggested to be important for relaying peptidergic signals due to the distinct anatomy and connectivity of projection neurons in peptidergic networks and their high authority values, and central with many neuronal inputs and outputs network components may function as regulatory hubs that gate the activation of specific networks [@williams2017].

### 2023-06-03 Manuscript/Manuscript.Rmd
Ms: edit transduction section, added refs

@@
 This late-phase activation of the photoreceptors drives circuit output through projection interneurons and ciliomotor neurons.
-In the cPRC circuit, synaptic connectivity alone is thus not sufficient to account for circuit dynamics and behavioural change, as documented in other circuits [@Imambocus2022].
+In the cPRC circuit, synaptic connectivity alone is thus not sufficient to account for circuit dynamics and behavioural change, as documented in other circuits [@bargmann2013; @Imambocus2022].
 
@@
 In neuronal signalling, NOS is often localized to neurites [@Kuntz2017] at close proximity to sGC at synapses [@burette2002; @garthwaite2015a].
-In the cPRC circuit, NOS is localised to the dendritic compartment of INNOS neurites in the neurosecretory plexus where cPRC axons synapse on INNOS neurites.
+In the cPRC circuit, NOS is localised to the dendritic compartment of INNOS cells where also cPRC postsynaptic sites occur.
 NIT-GC1, the target of NO signalling is localised to cPRC projections.
-NO-mediated retrograde signalling thus likely occurs in the neurosecretory plexus where NOS- and NIT-GC1-containing projections are in close proximity and where we could detect NO production following UV stimulation.
+NO-mediated retrograde signalling thus likely occurs in the neurosecretory plexus where NOS- and NIT-GC1-containing projections are in close proximity and where we detected NO production following UV stimulation.
+In contrast, INNOS to INRGW synapses occur outside the neurosecretory plexus in the ventral brain neuropil.
 
-Compartmentalised signalling also occurs in peptidergic modulatory system and can be important for selective network activity during specific behaviours.
-In the UV-avoidance circuit of *Drosophila* larvae, a peptidergic hub neuron Dp7 integrates different sensory inputs occurring at functionally separated dendritic compartments.
-The insulin-like peptide 7 (Ilp7) expressed in Dp7 mediates the UV signal from extraocular sensory neurons to motor circuits and is required for acute and sustained UV avoidance [@Imambocus2022].
+Compartmentalised signalling also occurs in peptidergic modulatory systems and can enable selective network activity during specific behaviours.
+In the UV-avoidance circuit of *Drosophila* larvae, a peptidergic hub neuron Dp7 links UV-sensory neurons (v'td2) and motor circuits.
+Dp7 expresses an an insulin-like peptide Ipl7 that is required for acute and sustained UV avoidance.
+Ilp7 signalling occurs at a functionally and morphologically distinct dendritic compartment of Dp7, segregated from other sensory-motor pathways involving Dp7 [@Imambocus2022].
 
@@
 NO signalling is commonly mediated by sGCs.
-We identified 12 sGCs in *Platynereis*, but none of these is expressed in the cPRC.
+We identified 12 sGCs in *Platynereis*, but none of these is expressed in the cPRC based on the available scRNAseq data.
 Instead, we identified an unconventional cPRC-expressed NIT-domain containing GC, NIT-GC1 as the mediator of NO retrograde signalling.
@@
 NIT domains in NIT-GCs may also sense nitrate and nitrite, as in bacteria, a possibility we cannot rule out based on our cellular assays with NIT-GC1.
-If different NIT-GCs show different sensitivities to NO, nitrite and nitrate, then the differences in the half-lives of these molecules may give rise to a range of activation timings [@lundberg2011].
+If different NIT-GCs have different sensitivities to NO, nitrite and nitrate, then a range of activation timings may be possible due to the different half-lives of these molecules [@lundberg2011].
 
-In *Platynereis*, we have found 15 NIT-GCs, suggesting a wide range of functions.
-We only tested NIT-GC1 and NIT-GC2, and the two proteins showed very different subcellular localisation and functions.
-Differences in cellular expression, subcellular localisation, and biochemical function may thus all contribute to the diversity of NIT-GC functions.
+NIT-GC1 and NIT-GC2 showed specific cellular co-expression but very different subcellular localisation and function.
+In *Platynereis*, we identified 15 NIT-GCs, suggesting a wide range of functions.
+Differences in subcellular localisation and biochemical function thus seem to also contribute to the diversity of NIT-GC functions in addition to differences in expression.
 
@@
 
+We have recently shown that the cPRC circuit also mediate responses to hydrostatic pressure via the same motor system involving Ser-h1 and the prototroch ciliary band.
+Increased pressure induces cPRC activation and a circuit output that is inverted relative to UV-induced activation.
+Consequently, ciliary beating increases and the larvae swim upwards.
+The effect of pressure on cilia requires synaptic transmission by the serotonergic Ser-h1 neurons [@bezares-calderon2023].
+
+Pressure-induced Ca2+ transients in cPRCs lack an inhibitory phase and late activation and resemble UV responses in NIT-GC2 morphants [@bezares-calderon2023].
+This indicates that the differentiation of sensory cues by the multisensory cPRCs occurs already at the level of sensory signal transduction.
+The different signalling pathways then likely result in the differential release of transmitters and modulators that are decoded by the postsynaptic interneuron circuit.
+The complex transmitter phenotype of cPRCs could underlie such differential signal processing.
+
 **Nitric oxide confers short-term memory to circuit activity**
@@
 This suggests that the NO-dependent memory trace also encodes the magnitude and duration of the stimulus.
+
 During UV avoidance in the planarian *Schmidtea mediterranea*, neuropeptide signalling has a similar integratory function.
@@
 
-NO signalling may also work on longer time scales.
-In the fly ellipsoid body, NO signalling is also required for long-term changes of the circuit, acting through CREB-dependent transcription.
-
-In Drosophila, NOS of PPL1-γ1pedc shortens memory retention while promoting fast memory updating in response to new experiences, indicating that NO regulates forgetting [@Aso2019].
+In *Drosophila*, NOS of PPL1-γ1pedc shortens memory retention while promoting fast memory updating in response to new experiences, indicating that NO regulates forgetting [@Aso2019].
 
@@
 
-Connectome reconstruction and annotation of all cells in the three-day-old *Platynereis* larva has not revealed any balancer organ to sense orientation in the gravity field [@veraszto2020].
+Connectome reconstruction and annotation of all cells in the three-day-old *Platynereis* larva has not revealed any balancer organ to sense orientation in the gravity field [@bezares-calderon2019; @veraszto2020].
 It follows that body orientation must be determined by physical mechanisms including the buoyancy and centre of mass of the larva, body shape, drag and differential ciliary activity.

### 2023-06-04 Manuscript/Manuscript.Rmd
Ms: discussion gravity

@@
 
-Connectome reconstruction and annotation of all cells in the three-day-old *Platynereis* larva has not revealed any balancer organ to sense orientation in the gravity field [@bezares-calderon2019; @veraszto2020].
-It follows that body orientation must be determined by physical mechanisms including the buoyancy and centre of mass of the larva, body shape, drag and differential ciliary activity.
-Except for ciliary activity, all other parameters are likely invariant during the UV response.
+Connectome reconstruction and whole-body cell annotation in the three-day-old *Platynereis* larva has not revealed any balancer organ to sense orientation in the gravity field [@bezares-calderon2019; @veraszto2020].
+It follows that body orientation must be determined by physical parameters including the buoyancy, centre of mass and shape of the larva as well as differential ciliary activity.
+Except for ciliary activity, all these parameters are likely invariant during the UV response.
+We found a UV-dependent inhibition of the Ser-h1 and activation of the MC cell consistent with the inhibition of ciliary beating [@veraszto2017].
 
-The only motor output of the cPRC circuit in the wiring diagram is the prototroch ciliary band postsynaptic to the MC and Ser-h1 ciliomotor neurons.
-Our Ca2+-imaging data revealed a NO-dependent inhibition of Ser-h1 and activation of the MC cell suggesting the inhibition of ciliary beating [@veraszto2017].
+Our current hypothesis about the mechanism of downward orientation posits that the reduced beating of prototroch cilia combined with the invariant beating of trunk cilia result in a head-down stable state due to physics alone.
+In contrast, during the pressure response prototroch cilia beat faster than trunk cilia, leading to a head-up orientation [@bezares-calderon2023].
+Testing this hypothesis will require biophysical experiments and modelling.
+
+## UV-avoidance circuits
 
 It is known that complex integrative networks of neurons are involved in UV avoidance behaviour in planaria [@Shettigar2021]
-
-RGW interneurons have been suggested to be important for relaying peptidergic signals due to the distinct anatomy and connectivity of projection neurons in peptidergic networks and their high authority values, and central with many neuronal inputs and outputs network components may function as regulatory hubs that gate the activation of specific networks [@williams2017].
-The finding of sGCs in INRGWs from the present single-cell analysis also suggests that they may be regulated by NO.
-The specific demarcation of sensory input and output through the use of neurotransmitters, neuropeptides and even gaseous NO may further increase the efficiency of network computation in hub neurons and may lead to a more complex mechanism of 'down' recognition.
 

### 2023-06-05 Manuscript/Manuscript.Rmd
MS: discussion, abstract edits, refs added

@@
 ---
-title: Nitric oxide feedback to ciliary photoreceptor cells orchestrates UV avoidance
-  in zooplankton
+title: Nitric oxide feedback to ciliary photoreceptor cells gates a UV avoidance circuit
 author: 
@@
 Nitric oxide (NO) generated by nitric-oxide synthase (NOS) is a key regulator of animal physiology.
-Here we uncover a role for NO in shaping circuit dynamics to orchestrate light-avoidance behaviour.
+Here we uncover a role for NO in gating a light-avoidance circuit and mediating short-term memory of UV exposure.
 We studied UV/violet light avoidance mediated by brain ciliary photoreceptors (cPRCs) in larval *Platynereis dumerilii*, a marine annelid.
 We found NOS expressed in interneurons (INNOS) postsynaptic to cPRCs.
-Stimulation of cPRCs by violet light leads to cPRC inhibition but concomitant INNOS activation and NO production.
-NO feeds back to cPRCs and triggers their delayed activation through an unconventional guanylyl cyclase.
-This results in the activation of projection interneurons and the inhibition of serotonergic ciliomotor neurons.
-In *NOS* mutants, NO feedback and projection-neuron activation do not occur and avoidance behaviour is defective.
-By mathematical modelling, we could recapitulate phototransduction and circuit dynamics in both wild-type and mutant larvae.
-Our results reveal how NO-mediated retrograde signalling gates a synaptic circuit to initiate light-avoidance behaviour.
+UV stimulation of cPRCs leads to INNOS activation and NO production.
+NO signals retrogradely to trigger the sustained post-stimulus activation of cPRCs through an unconventional guanylyl cyclase.
+This late activation inhibits serotonergic ciliomotor neurons via projection interneurons.
+In *NOS* mutants, retrograde signalling, circuit output and UV-avoidance behaviour are defective.
+By mathematical modelling, we recapitulate phototransduction and circuit dynamics in both wild-type and mutant larvae.
+Our results reveal how NO-mediated retrograde signalling gates a synaptic circuit and induces short-term memory of UV exposure to orchestrate light-avoidance behaviour.
 
@@
 
-In nervous systems, synaptic transmission and volume transmission together shape circuit dynamics.
+In nervous systems, synaptic transmission and volume transmission together shape circuit dynamics [@bargmann2013].
 While synaptic transmission occurs at specialised contact sites, volume transmission is characterised by the delocalised release of diverse diffusive neuromodulators.
@@
 This free radical synthesized by nitric oxide synthase (NOS) from L-arginine is short-lived and can diffuse across biological membranes [@Cudeiro1999; @thomas2015].
-Canonical NO signalling involves the Ca2+/calmodulin-dependent activation of NOS following calcium increase, NO production and diffusion, and the NO-dependent activation of soluble guanylate cyclases (sGC) cell-autonomously or in other cells leading to cGMP production [@Bredt1990; @Hoelscher1997].
+Canonical NO signalling involves the Ca2+/calmodulin-dependent activation of NOS following calcium increase, NO production and diffusion, and the NO-dependent activation of soluble guanylate cyclases (sGC).
+This can occur cell-autonomously or in other cells leading to cGMP production [@Bredt1990; @Hoelscher1997].
 Given that NOS activation is calcium dependent and NOS shows neuron-type-specific expression [@Gibbs1998a; @Wildemann1999; @Aso2019; @Mobley2022], NO action can lead to the activity-dependent modulation of neural circuits at specific sites [@Wang2007; @Vielma2014; @jacoby2018; @Aso2019] .
 
-NO signalling has been extensively studied in the vertebrate retina where NOS is expressed in amacrine, ganglion as well as other cells [@Cudeiro1999; @Wang2007; @jacoby2018].
-The actions of NO can be diverse and not always involve the canonical sGC-cGMP pathway [@Tooker2013; @Wei2012; @jacoby2018].
-Defective NO signalling in NOS knockouts for example leads to a decreased sensitivity of retinal ganglion cells to light stimulation [@Wang2007].
-However, due to the complex expression of NOS in vertebrates and the diversity of its functions, it has been challenging to link the neurophysiological effects of NO to signalling mechanisms and behaviour change.
+In the vertebrate retina, NOS is expressed in amacrine, ganglion and other cells and the actions of NO can be diverse [@Cudeiro1999; @Wang2007; @jacoby2018].
+For example, defective NO signalling in NOS knockouts leads to a decreased sensitivity of retinal ganglion cells to light stimulation [@Wang2007].
+In the retina, NO signalling can also involve pathways other than canonical sGC signalling [@Tooker2013; @Wei2012; @jacoby2018].
+Due to the complex expression of NOS in vertebrates and the diversity of its functions, it has been challenging to link the neurophysiological effects of NO to signalling mechanisms and behaviour change.
+
+In the *Drosophila* brain, NO signalling is also involved in tuning circuit activity at diverse and specific sites.
+In the ellipsoid body of the central complex, NO is involved in visual working memory [@Kuntz2017].
+NO signalling also tunes the dynamics of associative memories in mushroom body circuits.
+NOS is expressed in PPL1-γ1pedc mushroom-body neurons where it is involved in shortening memory retention while promoting fast memory updating in response to new experiences [@Aso2019].
 
@@
 
-Due to these difficulties, we still know little at the whole organism level about how NO production relates to stimulus conditions, how it shapes circuit activity at specific neuron types, and how NO-dependent modulation relates to behaviour.
+Overall, we still know little about how NO production relates to stimulus conditions, how it shapes circuit activity at specific neuron types, and how NO-dependent modulation relates to behaviour.
 
 To investigate NO function in neural circuit dynamics and behaviour, here we study larvae of the marine annelid model *Platynereis dumerilii* [@ozpolat2021].
-*Platynereis* has emerged as a systems neuroscience model where behaviour and activity imaging can be combined with genetic manipulations and a whole-body synaptic connectome and gene expression atlases are available [@Vergara2021; @Veraszto2017; @veraszto2020; @ozpolat2021].
+*Platynereis* has emerged as a model for systems neuroscience where behaviour and activity imaging can be combined with genetic manipulations.
+A whole-body synaptic connectome and gene expression atlases are also available [@Vergara2021; @Veraszto2017; @veraszto2020; @ozpolat2021].
 
@@
 Because of the short life time of NO, this molecule may be well suited to encode transient memory traces [@Kuntz2017].
+
 In the ellipsoid body of the *Drosophila* central brain, NO signalling has a similar function.
@@
 
-In *Drosophila*, NOS of PPL1-γ1pedc shortens memory retention while promoting fast memory updating in response to new experiences, indicating that NO regulates forgetting [@Aso2019].
-
 **The possible mechanism of UV-induced downward swimming**
@@
 UV/violet light induces downward swimming head down [@Veraszto2018].
-Since the direction of UV/violet light is not important for this, swimming direction must be determined by gravity.
+Since stimulus direction is not relevant, swimming direction must be determined by gravity.
 
@@
 Except for ciliary activity, all these parameters are likely invariant during the UV response.
-We found a UV-dependent inhibition of the Ser-h1 and activation of the MC cell consistent with the inhibition of ciliary beating [@veraszto2017].
+Our data suggest the UV-dependent inhibition of prototroch ciliary beating via Ser-h1 inhibition and MC activation [@veraszto2017].
 
-Our current hypothesis about the mechanism of downward orientation posits that the reduced beating of prototroch cilia combined with the invariant beating of trunk cilia result in a head-down stable state due to physics alone.
-In contrast, during the pressure response prototroch cilia beat faster than trunk cilia, leading to a head-up orientation [@bezares-calderon2023].
-Testing this hypothesis will require biophysical experiments and modelling.
+We hypothesise that the differential beating of prototroch versus trunk cilia causes a head-up or head-down orientation due to physics alone.
+During the UV response, prototroch cilia beat slower than trunk cilia, resulting in a head-down stable state ('rear-wheel drive').
+In contrast, during the pressure response prototroch cilia beat faster than trunk cilia [@bezares-calderon2023], leading to a head-up orientation ('front-wheel drive').
+Testing this hypothesis will require biophysical experiments and mathematical modelling.
 
-## UV-avoidance circuits
+## UV-avoidance circuits of extraocular photoreceptors
 
-It is known that complex integrative networks of neurons are involved in UV avoidance behaviour in planaria [@Shettigar2021]
+Animals evolved distinct photosensory systems coupled to non-overlapping circuits and guiding unique behavioural responses.
+These sensory systems can employ different opsin molecules and be tuned to different wavelength of light.
+The avoidance of noxious UV/violet light is often mediated by extraocular photoreceptors and their circuits, distinct from the pigmented visual eyes.
+These two types of systems co-exist in *Platynereis* where the pigmented eyes and eyespots guide phototaxis behaviour with a maximum sensitivity to cyan light (\~500 nm) [@Randel2014a; @Veraszto2018].
+Planarians also have pigmented visual eyes mediating phototaxis to cyan light and peripheral extraocular photoreceptors mediating UV avoidance behaviour [@Shettigar2021].
+*Drosophila* larvae have cerebral eyes called Bolwig's organs involved in phototaxis [@kane2013] and several types of extraocular UV-sensory cells that tile the body wall [@Imambocus2022].
+
+One common feature of UV-avoidance circuits is their sustained post-stimulus activation following UV exposure.
+This can involve peptidergic signals as in planaria and maggots [@bray2023; @Imambocus2022] or NO as in *Platynereis*.
+Volume transmission is well suited to integrate light exposure and maintain an internal state following noxious stimulation.
+The amount of modulator released can scale with stimulus intensity or duration and maintain an altered circuit state due to the slower decay of the diffusive signals relative to synaptic transmission.
 

### 2023-06-05 Manuscript/Manuscript.Rmd
Ms: whole-text edits, modelling section, discussion end

@@
 
-Nitric oxide (NO) generated by nitric-oxide synthase (NOS) is a key regulator of animal physiology.
-Here we uncover a role for NO in gating a light-avoidance circuit and mediating short-term memory of UV exposure.
-We studied UV/violet light avoidance mediated by brain ciliary photoreceptors (cPRCs) in larval *Platynereis dumerilii*, a marine annelid.
-We found NOS expressed in interneurons (INNOS) postsynaptic to cPRCs.
-UV stimulation of cPRCs leads to INNOS activation and NO production.
-NO signals retrogradely to trigger the sustained post-stimulus activation of cPRCs through an unconventional guanylyl cyclase.
-This late activation inhibits serotonergic ciliomotor neurons via projection interneurons.
-In *NOS* mutants, retrograde signalling, circuit output and UV-avoidance behaviour are defective.
-By mathematical modelling, we recapitulate phototransduction and circuit dynamics in both wild-type and mutant larvae.
+Nitric oxide (NO) produced by nitric-oxide synthase (NOS) is a key regulator of animal physiology.
+Here we uncover a function for NO in the integration of UV exposure and the gating of a UV-avoidance circuit.
+We studied UV/violet avoidance mediated by brain ciliary photoreceptors (cPRCs) in larvae of the annelid *Platynereis dumerilii*.
+In the larva, NOS is expressed in interneurons (INNOS) postsynaptic to cPRCs.
+UV stimulation of cPRCs triggers INNOS activation and NO production.
+NO signals retrogradely to cPRCs to induce their sustained post-stimulus activation through an unconventional guanylyl cyclase.
+This late activation inhibits serotonergic ciliomotor neurons to induce downward swimming.
+In *NOS* mutants, retrograde signalling, circuit output and UV avoidance are defective.
+By mathematical modelling, we recapitulate phototransduction and circuit dynamics in wild-type and mutant larvae.
 Our results reveal how NO-mediated retrograde signalling gates a synaptic circuit and induces short-term memory of UV exposure to orchestrate light-avoidance behaviour.
@@
 Nitric oxide (NO) is one such modulator with unique physical and signalling properties.
-This free radical synthesized by nitric oxide synthase (NOS) from L-arginine is short-lived and can diffuse across biological membranes [@Cudeiro1999; @thomas2015].
-Canonical NO signalling involves the Ca2+/calmodulin-dependent activation of NOS following calcium increase, NO production and diffusion, and the NO-dependent activation of soluble guanylate cyclases (sGC).
+This free radical synthesized from L-arginine by nitric oxide synthase (NOS) is short-lived and can diffuse across biological membranes [@Cudeiro1999; @thomas2015].
+Canonical NO signalling involves the Ca2+/calmodulin-dependent activation of NOS, NO production and diffusion, and the NO-dependent activation of soluble guanylate cyclases (sGC).
 This can occur cell-autonomously or in other cells leading to cGMP production [@Bredt1990; @Hoelscher1997].
@@
 To investigate NO function in neural circuit dynamics and behaviour, here we study larvae of the marine annelid model *Platynereis dumerilii* [@ozpolat2021].
-*Platynereis* has emerged as a model for systems neuroscience where behaviour and activity imaging can be combined with genetic manipulations.
-A whole-body synaptic connectome and gene expression atlases are also available [@Vergara2021; @Veraszto2017; @veraszto2020; @ozpolat2021].
+*Platynereis* has emerged as a model for systems neuroscience with a toolkit that enables the combination of behavioural analysis and neuronal activity imaging with genetic manipulations.
+A whole-body synaptic connectome and gene expression atlases are also available and can be integrated with functional approaches in live animals thanks to the cellular-level stereotypy of larvae of the same developmental stage [@Vergara2021; @veraszto2017; @veraszto2020; @ozpolat2021].
 
@@
 In *Platynereis* larvae, UV/violet avoidance is mediated by brain ciliary photoreceptor cells (cPRCs) and is characterised by downward swimming [@Veraszto2018].
-The cPRCs express a ciliary-type opsin, c-opsin1 [@Arendt2004] that forms a UV-absorbing bistable photopigment with an absorption maximum around 384 nm [@tsukamoto2017, @Veraszto2018, @Veedin2021].
+The cPRCs express a ciliary-type opsin, c-opsin1 [@Arendt2004] that forms a UV-absorbing bistable photopigment with an absorption maximum around 384 nm [@tsukamoto2017; @Veraszto2018; @Veedin2021].
 Upon UV/violet exposure, the cPRC show a characteristic biphasic calcium response that is *c-opsin1* dependent.
@@
 We identified a single *nitric oxide synthase* (*NOS*) gene in the *Platynereis dumerilii* genome and transcriptome data.
-Phylogenetic analysis of NOS proteins indicate that *Platynereis* NOS belongs to an orthology group of metazoan NOS sequences (Figure 1---figure supplement 1).
+Phylogenetic analysis of NOS proteins indicate that *Platynereis* NOS belongs to an orthology group of bilaterian NOS sequences (Figure 1---figure supplement 1).
 To characterise the expression pattern of *NOS* we used in situ hybridization chain reaction (HCR) and transient transgenesis.
@@
 The position and morphology of the four *NOS+* cells allowed us to identify the same four cells as four interneurons (INNOS) in our three-day-old whole-body *Platynereis* volume EM data [@williams2017; @veraszto2020] (Figure 1B,C).
-In the synaptic connectome, the INNOS cells are postsynaptic to the UV-sensory cPRCs and presynaptic to the INRGW interneurons, which are also cPRC targets (Figure 1C,F).
-The projections of INNOS cells are segregated into input and output compartments, with cPRCs inputs occurring in the dense neurosecretory plexus.
+In the synaptic connectome, the INNOS cells are postsynaptic to the UV-sensory cPRCs and presynaptic to the INRGWa interneurons, which are also cPRC targets (Figure 1C,F).
+The projections of INNOS cells are segregated into input (dendritic) and output (axonal) compartments, with cPRC inputs occurring on the dendritic part in the dense neurosecretory plexus.
 INNOS output synapses form in the more ventral projection region (Figure 1---figure supplement 3).
-INRGW neurons synapse on the head serotonergic ciliomotor neurons (Ser-h1), which synapse on the prototroch ciliary band and the cholinergic MC ciliomotor neurons (Figure 1C,F) [@Veraszto2017].
+INRGWa neurons synapse on the head serotonergic ciliomotor neurons (Ser-h1), which synapse on the prototroch ciliary band and the cholinergic MC ciliomotor neurons (Figure 1C,F) [@veraszto2017].
 
@@
 Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** **(A)** Scanning electron microscopy image of a three-day-old *Platynereis* larva.
-**(B, C)** Volume rendering of the neuron types (cPRC, INNOS, INRGW, Ser-h1 and MC) in the cPRC circuit reconstructed from a whole-body transmission electron microscopy volume of a three-day-old larva.
+**(B, C)** Volume rendering of the neuron types (cPRC, INNOS, INRGWa, Ser-h1 and MC) in the cPRC circuit reconstructed from a whole-body transmission electron microscopy volume of a three-day-old larva.
 Neurite skeletons are shown with cell-body positions represented by spheres.
@@
 **(C)** Changes in DAF-FM fluorescence over time during 405 nm stimulation of the cPRCs (green) or a control area (ctr: gray).
-The purple box indicate the duration of 405 nm stimulation.
+The purple box indicates the duration of 405 nm stimulation.
 Individual traces normalized (ΔF/F0) are shown as thin lines.
@@
 
-We next tested whether NO signalling is required for UV/violet avoidance.
+We next tested whether NO signalling regulates UV/violet avoidance.
 To achieve this, we generated two *Platynereis* *NOS* knockout lines with the CRISPR/Cas9 method.
-We recovered two deletions (NOSΔ11/Δ11 and Δ23/Δ23), both leading to frame-shift and an early stop codon and thus likely representing null alleles (Figure 3---figure supplement 1A).
+We recovered two deletions (NOSΔ11/Δ11 and Δ23/Δ23), both frame-shift mutations leading to an early stop codon and thus likely representing null alleles (Figure 3---figure supplement 1A).
 We could establish a homozygous line for both mutations indicating that NOS is not an essential gene in *Platynereis*.
@@
 In contrast, both two- and three-day-old homozygous *NOS*-mutant larvae showed a strongly diminished UV-avoidance response (Figure 3A, B and Figure 3---figure supplement 1B,C).
-This phenotype is similar to the defective UV-avoidance of *c-opsin1* mutant larvae and reveals a requirement for NOS in UV-avoidance behaviour [@Veraszto2018].
+This phenotype is similar to the defective UV-avoidance of *c-opsin1* mutant larvae [@Veraszto2018] and reveals a requirement for NOS in UV-avoidance behaviour.
 Wild type but not mutant larvae also showed an increase in swimming speed under UV light that may be due to downward swimming trajectories (swimming in the direction of gravity) (Figure 3B and Figure 3---figure supplement 1C).
@@
 Three-day-old but not two-day-old *NOS*-mutant larvae also showed reduced phototactic behaviour, suggesting a function for *NOS* in the visual eyes that mediate three-day-old phototaxis [@Randel2014a] (Figure 3D and Figure 3---figure supplement 1G).
+
 To distinguish between an acute and developmental function of NOS in light responses, we next tested larvae exposed to the NOS inhibitor L-NAME.
@@
 The starting position of each larval trajectory was set to 0.
-**(D)** Vertical displacement in 30 sec bins of wild type and mutant (NOSΔ11 and NOSΔ23) three-day-old larvae stimulated with 395 nm light from side, 488 nm light from the top and 395 nm light from the top.
+**(D)** Vertical displacement in 30 sec bins of wild type and mutant (NOSΔ11 and NOSΔ23) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
 **(E)** Vertical displacement in 30 sec bins of control and L-NAME-treated (0.1 and 1 mM) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top."}
@@
 We ubiquitously expressed the calcium sensor GCaMP6s in larvae and imaged calcium signals during 405 nm stimulation of the cPRCs.
-As we have shown previously, a 20-sec local stimulation of cPRC cilia lead to a transient increase in cPRC calcium levels, followed by a transient decrease [@Veraszto2018].
+As we have shown previously, a 20-sec local stimulation of cPRC cilia led to a transient increase in cPRC calcium levels, followed by a transient decrease [@Veraszto2018].
 After \~20-sec, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus -- a response that may involve depolarisation (Figure 4A, B).
@@
 
-We aimed next to identify the NO-receptor in the cPRCs.
+We aimed next to identify the NO receptor in the cPRCs.
 NO generally acts via soluble guanylate cyclases (sGC), belonging to the guanylate cyclase family with a CYC domain (PFAM domain: PF00211).
 NO binding to the heme group of sGC leads to increased cyclic guanosine monophosphate (cGMP) production.
-Analysis of sGCs in *Platynereis* indicated that these genes are not expressed in any of the cells of the cPRC circuit [@Veraszto2017].
+Analysis of sGCs in *Platynereis* indicated that these genes are not expressed in any of the cells of the cPRC circuit [@veraszto2017].
 Recently, Moroz and coworkers reported an atypical but widely conserved family of guanylyl cyclases with a NIT (nitrite/nitrate sensing) domain (PF08376) (NIT-GC) as potential mediators of NO signalling [@Moroz2020].
@@
 The 15 *Platynereis* sequences belonged to several deeply diverged clades in the phylogenetic tree (Figure 4---figure supplement 1 and 2).
+
 To characterise the expression of NIT-GCs, we used previously published spatially mapped single-cell transcriptome data [@Achim2015; @williams2017].
@@
 To analyse the subcellular localisation of NIT-GC1 and NIT-GC2 at the protein level, we raised and affinity-purified polyclonal antibodies against a specific peptide sequence from both proteins.
-In immunostainings, we found that NIT-GC1 was localise to the region corresponding to the axonal projections of the cPRCs in the anterior nerve plexus (Figure 4E).
+In immunostainings, we found that NIT-GC1 was localise to the region corresponding to the axonal projections of the cPRCs in the anterior neurosecretory plexus (Figure 4E).
 Co-immunostaining with the rabbit NIT-GC1 antibody and a custom rat antibody raised against *Platynereis* NOS revealed the localisation of both proteins in dots in the neurosecretory neuropil (Figure 4---figure supplement 3D).
@@
 The NIT domain may render NIT-GC1 dependent on NO signals.
-To test this, we co-expressed the cGMP indicator Green cGull [@Matsuda2017] and NIT-GC1 in cultured COS-7 (monkey kidney) cells, a cell line with minimal endogenous sGC activity.
+To test this, we co-expressed the cGMP indicator Green cGull [@Matsuda2017] and NIT-GC1 in cultured COS-7 (monkey kidney) cells, a cell line with minimal endogenous sGC activity [@Matsuda2017].
 For balanced expression, we used a single plasmid with the two open-reading frames separated by the 2A self-cleaving peptide (Figure 4H).
-Application of the NO donor SNAP lead to increased cGMP levels, an effect we did not observe when cells were exposed to DMSO or when Green cGull was expressed alone (Figure 4I-K).
-To test whether cGMP production is dependent on the NIT domain, we also tested a deletion construct of NIT-GC1 lacking the NIT domain (Figure 4G).
-Cells expressing this construct and Green cGull did not show increased cGMP levels when exposed to SNAP (Figure 4L).
+Application of the NO donor SNAP lead to increased Green cGull fluorescence, an effect we did not observe when cells were exposed to DMSO or when Green cGull was expressed alone (Figure 4I-K).
+To test whether this effect is dependent on the NIT domain, we also tested a deletion construct of NIT-GC1 lacking the NIT domain (Figure 4G).
+Cells expressing this construct and Green cGull did not show an increased fluorescence of the cGMP reporter when exposed to SNAP (Figure 4L).
 These results indicate that NIT-GC1 is able to catalyse cGMP production in an NO-dependent manner and this function requires the NIT domain.
-These results establish NIT-GC1 as a biochemical NO sensor.
+These results establish NIT-GC1 as a biochemical sensor of NO or its derivatives.
 
@@
 **(C, D)** In situ HCR for (C) *NIT-GC1* and (D) *NIT-GC2* (magenta) in three-day-old *Platynereis* larvae.
-Larvae were co-stained with an antibody agains acetylated α-tubulin to label cPRC cilia and the neuropil (green).
+Larvae were co-stained with an antibody against acetylated α-tubulin to label cPRC cilia and the neuropil (green).
 **(E, F)** Immunostaining for (E) NIT-GC1 and (F) NIT-GC2 (magenta), co-stained for acetylated α-tubulin (green).
@@
 To investigate how NO and NIT-GC signalling influence the dynamics of the cPRC circuit, we imaged calcium signals from postsynaptic neurons in wild type, mutant and morphant larvae.
-We were able to image the activity of all neurons in the cPRC circuit (INNOS, INRGW, Ser-h1 and MC).
-The MC cell was identified based on its position and intrinsic activity [@Veraszto2017].
+We were able to image the activity of all neurons in the cPRC circuit (INNOS, INRGWa, Ser-h1 and MC).
+The MC cell was identified based on its position and intrinsic activity [@veraszto2017].
 To unambiguously identify all other cells from which we recorded calcium signals, we developed an on-slide immunostaining method (Figure 5---figure supplement 1A).
-We used the cell-specific antibody markers against RYamide (INNOS) (Figure 5---figure supplement 1B-D), RGWamide (INRGW) and serotonin (Ser-h1) [@Conzelmann2011] to immunostain agar-embedded larvae following calcium imaging.
+We used the cell-specific antibody markers against RYamide (INNOS) (Figure 5---figure supplement 1B-D), RGWamide (INRGWa) and serotonin (Ser-h1) [@Conzelmann2011] to immunostain agar-embedded larvae following calcium imaging.
 Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision (Figure 5A,B).
@@
 
-We first quantified the responses of the INNOS and INRGW interneurons during 405 nm stimulation of the cPRCs.
+We first quantified the responses of the INNOS and INRGWa interneurons during 405 nm stimulation of the cPRCs.
 In both wild type and *NOS*-mutant larvae, INNOS cells showed an increase in calcium during stimulation (Figure 5D).
 In contrast, the INNOS response was flat or slightly negative in NIT-GC2 morphant larvae (Figure 5E) revealing an essential role for NIT-GC2-mediated cPRC suppression in INNOS activation.
-INRGW cells were initially inhibited during cPRC stimulation, followed by a delayed activation paralleling the second activation phase of cPRCs.
-This late INRGW response was lacking in *NOS*-mutants (Figure 5F).
-In NIT-GC2 morphants, INRGW cells showed a transient increase in calcium that decayed after light off and a delayed activation was not present (Figure 5G).
+INRGWa cells were initially inhibited during cPRC stimulation, followed by a delayed activation paralleling the second activation phase of cPRCs.
+This late INRGWa response was lacking in *NOS*-mutants (Figure 5F).
+In NIT-GC2 morphants, INRGWa cells showed a transient increase in calcium that decayed after light off and a delayed activation was not present (Figure 5G).
+
 Next, we imaged calcium signals from Ser-h1 and MC neurons in wild type and *NOS* mutant larvae.
@@
 ```{r echo=FALSE,fig.cap="**Figure 5.
-NOS- and NIT-GC2-dependent dynamics of the cPRC circuit.** **(A, B)** GCaMP6s imaging from cPRCs and INNOS cells (left panels) followed by on-slide immunostaining for (A) RYamide to label INNOS and (B) RGWamide+serotonin to label INRGW and Ser-h1 (red).
+NOS- and NIT-GC2-dependent dynamics of the cPRC circuit.** **(A, B)** GCaMP6s imaging from cPRCs and INNOS cells (left panels) followed by on-slide immunostaining for (A) RYamide to label INNOS and (B) RGWamide+serotonin to label INRGWa and Ser-h1 (red).
 Nuclei are stained with DAPI (cyan).
@@
 Numbers mark the same cells in the GCaMP and immunostaining images matched by position.
-**(C)** Correlation map of neuronal activity of the cPRCs, INNOS, INRGW, Ser-h1 and MC neurons.
+**(C)** Correlation map of neuronal activity of the cPRCs, INNOS, INRGWa, Ser-h1 and MC neurons.
 **(D)** GCaMP6s fluorescence in INNOS cells in wild type (WT) and NOSΔ11/Δ11 mutant larvae during 405 nm stimulation of the cPRC cilia.
 **(E)** GCaMP6s fluorescence in INNOS cells in NIT-GC2 morphant larvae during 405 nm stimulation.
-**(F)** GCaMP6s fluorescence in INRGW cells in wild type and NOSΔ23/Δ23 mutant larvae during 405 nm stimulation.
-**(G)** GCaMP6s fluorescence in INRGW cells in NIT-GC2 morphant larvae during 405 nm stimulation.
+**(F)** GCaMP6s fluorescence in INRGWa cells in wild type and NOSΔ23/Δ23 mutant larvae during 405 nm stimulation.
+**(G)** GCaMP6s fluorescence in INRGWa cells in NIT-GC2 morphant larvae during 405 nm stimulation.
 **(H, I)** GCaMP6s fluorescence in (H) Ser-h1 cells and (I) the MC cell in wild type and NOSΔ11/Δ11 mutant larvae during 405 nm stimulation."}
@@
 To further analyse the dynamics of responses to UV light and formally describe cPRC phototransduction, synaptic connections and NO retrograde signalling, we developed a mixed cellular-circuit-level mathematical model.
-We used our Ca2+ imagining data of cPRC, INNOS and INRGW cells collected in wild type, *NOS* knockout, and *NIT-GC2* morphant larvae to formulate assumptions that are the basis of the proposed model.
+We used our Ca2+ imagining data of cPRC, INNOS and INRGWa cells collected in wild type, *NOS* knockout, and *NIT-GC2* morphant larvae to formulate assumptions that are the basis of the proposed model.
 
@@
 We further assume that in cPRC cells, Ca2+ levels increase with cGMP.
-We model a decrease in cGMP and Ca2+ levels due to a NIT-GC2-dependent pathway.
+We model a decrease in cGMP and Ca2+ levels due to a NIT-GC2-dependent pathway (Figre 6A).
 The effect of NO in cPRC cells is captured by another term that describes a NIT-GC1 and NO-dependent increase in cGMP.
-Based on the synaptic connectome, we infer a feedforward coupling between cPRC cells and INNOS and INRGW cells.
-The proposed synaptic coupling is UV-dependent, decays linearly and is suppressed in NIT-GC2 morphants.
+Based on the synaptic connectome, we infer a feedforward coupling between cPRC cells and INNOS and INRGWa cells.
+The proposed synaptic coupling is UV-dependent, decays linearly and is suppressed in NIT-GC2 morphants (Figure 6A).
 
-In the INNOS cells, we assume that excitatory synaptic input from cPRC leads to a rise in Ca2+ leading to NO production via NOS.
+In the INNOS cells, we assume that excitatory synaptic input from cPRC (or rebound from tonic inhibition) leads to a rise in Ca2+ leading to NO production via NOS.
 This input is NIT-GC2-dependent.
-In the INRGW cells, we model a decrease in Ca2+ based on an inhibitory UV-dependent synaptic signal from the cPRCs.
-We further assume the existence of direct excitatory coupling between Ca2+ level in cPRC and INRGW Ca2+ to account for the late effects of cPRC on INRGW.
-The Uv-dependent and Ca2+-dependent signals may be mediated by different transmitters released by cPRCs during distinct phases of their activation cycle.
-In addition, we assume a direct inhibitory coupling between INNOS Ca2+ and INRGW Ca2+.
-Finally, in all variables, we assumed a linear decay and constant production to set steady a state.
+In the INRGWa cells, we model a decrease in Ca2+ based on an inhibitory UV-dependent synaptic signal from the cPRCs.
+We further assume the existence of direct excitatory coupling between cPRC Ca2+ levels and INRGWa Ca2+ to account for the late effects of cPRC on INRGWa (Figure 6A).
+The UV-dependent and Ca2+-dependent signals may be mediated by different transmitters released by cPRCs during distinct phases of their activation cycle.
+In addition, we assume a direct inhibitory coupling between INNOS Ca2+ and INRGWa Ca2+.
+Finally, in all variables, we assumed a linear decay and constant production to set a steady state (Figure 6A).
 Since the aim of the model is to capture the normalised fluorescence data, the model is nondimensionalised.
@@
 
-Our model only includes a minimal set of parameters and assumptions of interactions that are required to describe the dynamics of cPRC, INNOS and INRGW in wild type and loss-of-function conditions (Figure 6C, D).
+Through this optimisation procedure, we found parameters that gave a very good fit to our experimental data under all conditions.
+We could fit to mean Ca2+ values but also individual calcium transients (Figure 6B, Figure 6--figure supplement 1).
+Our model only includes a minimal set of parameters and assumptions of interactions that are required to describe the dynamics of cPRC, INNOS and INRGWa in wild type and loss-of-function conditions (Figure 6A, B).
 The model highlights that cPRC phototransduction employs distinct pathways that operate on different time scales and differentially influence cPRC Ca2+ levels.
@@
 
-To identify possible molecular pathways, we analysed single-cell transcriptome data for cPRC, INNOS and INRGW identified by spatial mapping [@Achim2015] and unique marker genes [@williams2017].
-In cPRCs, we found transporters and synthetic enzymes indicating cholinergic, glutaminergic, glycinergic, GABAergic and adrenergic neurotransmission (Figure 6A).
-For each neurotransmitter, we also found receptor-encoding genes expressed in INNOS and INRGW (Figure 6A).
-The two types of interneurons often expressed different subunits or types of these receptors, indicating differential signalling (Figure 6A).
+To identify possible molecular pathways, we analysed single-cell transcriptome data for cPRC, INNOS and INRGWa identified by spatial mapping [@Achim2015] and unique marker genes [@williams2017].
+In cPRCs, we found transporters and synthetic enzymes indicating cholinergic, glutaminergic, glycinergic, GABAergic and adrenergic neurotransmission (Figure 6C).
+For each neurotransmitter, we also found receptor-encoding genes expressed in INNOS and INRGWa (Figure 6C).
+The two types of interneurons often expressed different subunits or types of these receptors, indicating differential signalling (Figure 6C).
 
@@
 The components for which experimental data are available are shown in bold.
-Other potential molecular players are also indicated (Figure 6B).
+Other potential molecular players are also indicated (Figure 6D).
+
+To generate new insights from the model, we asked how the duration and intensity of UV stimulation influences signalling output, measured as the magnitude, duration and time to maximum of the NO-dependent Ca2+ signal in the cPRCs.
+The model allowed us to test a large number of conditions that would not have been feasible in single-larva experiments.
+We found that the onset, magnitude and duration of the NO-dependent signal scales with stimulus intensity and duration (Figure 6--figure supplement 2).
+This suggests an integratory function of NO signalling that tunes circuit output to the strength of UV exposure.
 
 ```{r echo=FALSE,fig.cap="**Figure 6.
-Signalling mechanisms and mathematical modelling between cPRC, INNOS and INRGW.** **(A)** Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq.
+Mathematical modelling of cPRC circuit dynamics during UV stimulation and potential signalling mechanisms.**  **(A)** Diagram for **(B)** **(C)** Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq.
 The size of the dots is expressed in proportion to the percentage of cells expressing that gene relative to all cells.
 The colours represent the normal logarithm of the number of transcripts in the cells expressing the gene.
-**(B)** Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback.
-**(C)** Diagram for **(D)**"}
+**(D)** Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback."}
 
@@
 NO-mediated retrograde signalling thus likely occurs in the neurosecretory plexus where NOS- and NIT-GC1-containing projections are in close proximity and where we detected NO production following UV stimulation.
-In contrast, INNOS to INRGW synapses occur outside the neurosecretory plexus in the ventral brain neuropil.
+In contrast, INNOS to INRGWa synapses occur outside the neurosecretory plexus in the ventral brain neuropil.
 
@@
 In the UV-avoidance circuit of *Drosophila* larvae, a peptidergic hub neuron Dp7 links UV-sensory neurons (v'td2) and motor circuits.
-Dp7 expresses an an insulin-like peptide Ipl7 that is required for acute and sustained UV avoidance.
+Dp7 expresses an an insulin-like peptide Ilp7 that is required for acute and sustained UV avoidance.
 Ilp7 signalling occurs at a functionally and morphologically distinct dendritic compartment of Dp7, segregated from other sensory-motor pathways involving Dp7 [@Imambocus2022].
@@
 NIT domains were first identified in bacteria and animal NIT-GCs have only recently been reported [@Shu2003; @Moroz2020].
-Bacterial NIT domains regulate cellular functions in response to changes in extracellular and intracellular nitrate and/or nitrite concentrations [@camargo2007].
+Bacterial NIT domains regulate cellular functions in response to changes in extracellular and intracellular nitrate and nitrite concentrations [@camargo2007].
 NO is readily converted to nitrate and nitrite [@santos2011; @moeller2019; @garthwaite2015a] and these molecules accumulate in placozoans and cnidarians in cells and tissues with high NOS activity [@Moroz2004; @Moroz2020].
@@
 The activation of INNOS leads to NOS activation and NO release, potentially through a canonical Ca2+-calmodulin pathway.
-Our model predicts feedforward inhibiton from INNOS to INRGW, possibly mediated by glycin transmission.
-NO released by INNOS neurites activates NIT-GC1 in cPRC projections that could lead to cGMP production and the opening CNGAα.
+Our model predicts feedforward inhibiton from INNOS to INRGWa, possibly mediated by glycin transmission.
+NO released by INNOS neurites activates NIT-GC1 in cPRC projections that could lead to cGMP production and the opening of CNGAα.
 This late-phase cPRC activation is not directly dependent on the UV/violet signal and can happen post-stimulus.
-The late-phase cPRC activation leads to INRGW activation via an unknown transmitter that is likely different from the one acting on INNOS.
+The late-phase cPRC activation leads to INRGWa activation via an unknown transmitter that is likely different from the one acting on INNOS.
 
 In addition, the cPRC circuit expresses several neuropeptides and their receptors, suggesting further neuromodulatory signals.
-For example, INRGW express the proneuropeptide RGWamide and cPRCs and INNOS express its receptor, suggesting retrograde peptidergic signalling in the circuit.
+For example, INRGWa express the proneuropeptide RGWamide and cPRCs and INNOS express its receptor, suggesting retrograde peptidergic signalling in the circuit.
 
-We have recently shown that the cPRC circuit also mediate responses to hydrostatic pressure via the same motor system involving Ser-h1 and the prototroch ciliary band.
+We have recently shown that the cPRC circuit also mediates responses to hydrostatic pressure via the same motor system involving Ser-h1 and the prototroch ciliary band.
 Increased pressure induces cPRC activation and a circuit output that is inverted relative to UV-induced activation.
@@
 
-UV/violet light induces downward swimming head down [@Veraszto2018].
+Non-directional UV/violet light induces downward swimming head down in *Platynereis* larvae [@Veraszto2018].
 Since stimulus direction is not relevant, swimming direction must be determined by gravity.
@@
 Animals evolved distinct photosensory systems coupled to non-overlapping circuits and guiding unique behavioural responses.
-These sensory systems can employ different opsin molecules and be tuned to different wavelength of light.
+These sensory systems can employ different opsin molecules and be tuned to different wavelengths of light.
 The avoidance of noxious UV/violet light is often mediated by extraocular photoreceptors and their circuits, distinct from the pigmented visual eyes.
-These two types of systems co-exist in *Platynereis* where the pigmented eyes and eyespots guide phototaxis behaviour with a maximum sensitivity to cyan light (\~500 nm) [@Randel2014a; @Veraszto2018].
+These two types of systems co-exist in *Platynereis* where the pigmented eyes and eyespots guide phototaxis with a maximum sensitivity to cyan light (\~500 nm) [@Randel2014a; @Veraszto2018].
 Planarians also have pigmented visual eyes mediating phototaxis to cyan light and peripheral extraocular photoreceptors mediating UV avoidance behaviour [@Shettigar2021].
-*Drosophila* larvae have cerebral eyes called Bolwig's organs involved in phototaxis [@kane2013] and several types of extraocular UV-sensory cells that tile the body wall [@Imambocus2022].
+*Drosophila* larvae have cerebral eyes called Bolwig's organs involved in phototaxis [@kane2013] and several types of extraocular UV-sensory cells that tile the body wall and mediate UV avoidance [@Imambocus2022].
 
@@
 The amount of modulator released can scale with stimulus intensity or duration and maintain an altered circuit state due to the slower decay of the diffusive signals relative to synaptic transmission.
+
+We have revealed how NO shapes dynamics in a fully-mapped sensory-motor UV avoidance circuit.
+We identified an unconventional GC as the NO receptor in *Platynereis* and measured circuit activity in different genetic backgrounds.
+Finally, we could link circuit activity to light-avoidance behaviour.
+The richly modulated multi-transmitter cPRC circuit will serve as a fertile ground for future studies on how neuromodulators shape activity and behaviour.
 
@@
 250 ng/ml) were analysed for reporter expression at 3 days post fertilization using an AxioImager Z.1 fluorescence wide-field microscope (Carl Zeiss GmbH, Jena) and immediately fixed for immunostainings.
-The protocol followed for immunostaining of HA-tagged reporters was recently described [@Veraszto2017].
+The protocol followed for immunostaining of HA-tagged reporters was recently described [@veraszto2017].
 Specimens were imaged with a LSM 780 NLO or LSM 880 with Airysan Confocal Microscope (Zeiss, Jena).
@@
 
-```{r echo=FALSE,fig.cap="**Figure 1 -- figure supplement 3** ssTEM reconstruction of the INNOS (A), INRGW (B) and cPRC (C), anterior view. NS plexus; neurosecretory plexus"}
+```{r echo=FALSE,fig.cap="**Figure 1 -- figure supplement 3** ssTEM reconstruction of the INNOS (A), INRGWa (B) and cPRC (C), anterior view. NS plexus; neurosecretory plexus"}
 knitr::include_graphics(path = "figures/Fig1_sup3.png")
@@
 Guanylate cyclases with NIT domains are found in most animal phyla except Porifera, Ctenophora, Urochordata and Chordata.
-Dot plot of Platynereis NIT-GC genes (columns) expressed in cPRC, INNOS and INRGW (rows) using single cell RNA-Seq.
+Dot plot of Platynereis NIT-GC genes (columns) expressed in cPRC, INNOS and INRGWa (rows) using single cell RNA-Seq.
 The size of the dots is expressed in proportion to the percentage of cells expressing that gene relative to all cells.

### 2023-06-07 Manuscript/Manuscript.Rmd
Ms: Methods edit, key resources table update

@@
     theme: lumen
+  word_document: default
   pdf_document:
@@
     template: default
-  word_document: default
 geometry: left=0.8in,right=0.8in,top=0.8in,bottom=0.5in
@@
 library(readxl)
-key_resource_tale <- read_excel("key_resource_tale/key_resource_tale.xlsx")
+key_resource_tale <- read_excel("key_resource_tale/key_resource_tale.xls")
 df_key_resource_tale <- as.data.frame(key_resource_tale)
@@
 
-Before designing the small guide RNA (sgRNA) for the sgRNA:Cas9 nuclease, splice sites and polymorphic sites in our laboratory culture were identified to avoid them.
-The sgRNA targeted the third exon of Platynereis dumerilii NOS (target site: 5'-GGGCAATACTGGCTCCACTC-3').
-The sgRNA was assembled from two annealed oligonucleotides (5'-TAGGGCAATACTGGCTCCACTC-3', 5'-AAACGAGTGGAGCCAGTATTGC-3') forming overhangs for cloning into a BsaI site of the plasmid pDR27456 (Hwang et al.
-2013)(42250, Addgene), which contains next to the BsaI site a tracrRNA sequence.
-The plasmid was then used to PCR amplify DNA (primers: T7, 5'-AAAAGCACCGACTCGGTGCC-3') for synthesizing the sgRNA.
-The DNA was purified with the QIAquick PCR Purification Kit (Qiagen).
-From the DNA, the sgRNA was synthesized with the MEGAshortscript Kit (Thermo Fisher Scientific) and was purified with the MEGAclear Kit (Thermo Fisher Scientific).
-Cas9-mRNA was transcribed, capped, and polyA-tailed with the mMessage mMachine Kit and the Poly(A) Tailing Kit (both Thermo Fisher Scientific) from a plasmid (pUC57-T7-RPP2-Cas9) containing the Cas9 ORF fused to 169 base pair 5' UTR from the Platynereis dumerilii 60S acidic ribosomal protein P2.
-The sgRNA (18 ng/ml) and the Cas9-mRNA (180 ng/µl) were coinjected into fertilized eggs of Platynereis dumerilii wild-type parents according to an established injection procedure [@Conzelmann2013a].
+For the generation of a *NOS* knockout we used an sgRNA targeting the third exon of the *Platynereis dumerilii* *NOS* (target site: 5'-GGGCAATACTGGCTCCACTC-3').
+We selected the site to avoid polymorphic sites in our laboratory culture.
+The sgRNA was assembled from two annealed oligonucleotides (5'-TAGGGCAATACTGGCTCCACTC-3', 5'-AAACGAGTGGAGCCAGTATTGC-3') forming overhangs for cloning into the BsaI site of the plasmid pDR27456 [\@hwang2013](42250,%20Addgene), which contains next to the BsaI site a tracrRNA sequence.
+We used this plasmid to PCR amplify DNA (primers: T7, 5'-AAAAGCACCGACTCGGTGCC-3') for synthesizing the sgRNA.
+To purify the PCR product, we used the QIAquick PCR Purification Kit (Qiagen).
+We then synthesized sgRNA with the MEGAshortscript Kit (Thermo Fisher Scientific) and purified it with the MEGAclear Kit (Thermo Fisher Scientific).
+To produce the Cas9-mRNA, we used the plasmid (pUC57-T7-RPP2-Cas9) containing the Cas9 ORF fused to 169 base pair 5' UTR from the *Platynereis dumerilii* 60S acidic ribosomal protein P2.
+After in vitro transcription with the mMessage mMachine Kit (Thermo Fisher Scientific), we capped and polyA-tailed the Cas9-mRNA with the Poly(A) Tailing Kit (Thermo Fisher Scientific).
+We coinjected the sgRNA (18 ng/ml) and the Cas9-mRNA (180 ng/µl) into fertilized eggs of *Platynereis dumerilii* wild-type parents according to an established injection procedure [@Conzelmann2013a].
 The eggs were kept at 18°C for 45 min before injection and were injected at 14.5°C.
-The injected individuals were kept at 18°C for 5 to 8 days in 6-well- plates (Nunc multidish no. 150239, Thermo Scientific) and then cultured at 22°C until sexual maturity.
+The injected individuals were kept at 18°C for 5 to 8 days in 6-well-plates (Nunc multidish no. 150239, Thermo Scientific) and then cultured at 22°C until sexual maturity.
 The mature worms were crossed to wild-type worms and the progeny was genotyped, resulting in two founder lines, which were bred to homozygosity.
 
-**NOS sequencing and genotyping**
+**Genotyping of NOS alleles**
 
@@
 
-**Vertical column setup for measuring photoresponses**
+**Vertical column setup to measure photoresponses**
 
-Photoresponses of larvae of different ages were assayed in a vertical Plexiglas column (31 mm x 10 mm x 160 mm water height).
-The column was illuminated from top with light from a monochromator (Polychrome II, Till Photonics).
-The monochromator was controlled by AxioVision 4.8.2.0 (Carl Zeiss MicroImaging GmbH) via analog voltage.
+We assayed larval photoresponses in a vertical Plexiglas column (31 mm x 10 mm x 160 mm water height).
+For stimulation, we illuminated the column from the top with a monochromator (Polychrome II, Till Photonics) controlled by AxioVision 4.8.2.0 (Carl Zeiss MicroImaging GmbH) via analog voltage.
 The light passed a collimator lens (LAG-65.0-53.0-C with MgF2 Coating, CVI Melles Griot) before entering the column.
@@
 One strip contained UV (395 nm) LEDs (SMB1W-395, Roithner Lasertechnik) and the other infrared (810 nm) LEDs (SMB1W-810NR-I, Roithner Lasertechnik).
-The UV LEDs were run at 4 V to stimulate the larvae in the column from the side.
-The infrared LEDs were run at 8 V (overvoltage) to illuminate the larvae for the camera (DMK 22BUC03, The Imaging Source), which recorded videos at 15 frames per second and was controlled by IC Capture (The Imaging Source).
+The UV LEDs were run at 4 V. The infrared LEDs were run at 8 V (overvoltage) to illuminate the larvae for imaging with a DMK camera (DMK 22BUC03, The Imaging Source).
+We recorded videos at 15 frames per second with the software IC Capture (The Imaging Source).
 
-**Comparing behavior of wildtype and NOS-knockout 3-day-old larvae**
-
-To compare the behavior of wildtype and NOS-knockout larvae at 3 days in the vertical column, the larvae were mixed and left in the dark for 5 min.
-The larvae were treated with NOS inhibitors for pharmacology.
-The NOS inhibitors were L-NAME.
-The larvae were treated with different concentrations in adjacent columns.
-The concentrations for the NOS inhibitors were control, 1 mM, 0.1 mM.
-The larvae were recorded for 1 min in the dark followed by exposure to collimated cyan (480 nm) light from the top of the column for 2 min, then 2 min darkness, and finally collimated UV (395 nm) light from the top of the column for 2 min.
-Stimulus light was provided by the monochromator (Polychrome II, Till Photonics).
+To compare the behavior of wildtype and NOS-knockout larvae or larvae treated with L-NAME, we first mixed the larvae and left them in the dark for 5 min.
+We then imaged larvae for 1 min in the dark (only IR LED for illumination) followed by exposure to non-directional UV light from both sides, then 2 min darkness, then collimated cyan (480 nm) light from the top of the column for 2 min, then 2 min darkness, and finally collimated UV (395 nm) light from the top of the column for 2 min.
 Scripts are available at <https://github.com/JekelyLab/NOS>.
 
-**NOS Identification and Phylogenetic Analysis**
+**Identification and Phylogenetic Analysis of NOS sequences**
 
-To identify NOS, we obtained a "seed" database of oxygenase domain in Pfam database, PF02898.
-From these sequences, we produced a Hidden Markov Model (HMM) and used this to mine the 47 metazoan species, 2 choanoflagellate species and 2 filasterea species investigated.
+To identify NOS sequences, we obtained a seed database of oxygenase domains from the Pfam database (PF02898).
+From these sequences, we produced a Hidden Markov Model (HMM) and used this to mine the 47 metazoan, 2 choanoflagellate and 2 filasterea species investigated.
 HMM models were run in HMMR3 with an e-value of 1e−15.
@@
 To calculate maximum-likelihood trees, we used IQ-tree2 with the LG+G4 model.
-To calculate branch support, we ran 1,000 replicates with the aLRT-SH-like and aBayes methods [@Minh2020].
+To calculate branch support, we ran 1000 replicates with the aLRT-SH-like and aBayes methods [@Minh2020].
 The sequences used for the phylogenetic analysis are available in **Supplementary File 1**, the trimmed alignment is available in **Supplementary File 2** and the pre-trimmed data in **Supplementary file 3**.
 
-**NIT-GC Identification and Phylogenetic Analysis**
+**Identification and Phylogenetic Analysis of NIT-GC sequences**
 
-To identify NIT-GCs, we obtained a "seed" database of Adenylate and Guanylate cyclase catalytic domain in Pfam database, PF00211.
-From these sequences, we produced a Hidden Markov Model (HMM) and used this to mine the 45 metazoan species, 2 choanoflagellate species and 2 filasterea species investigated.
+To identify NIT-GCs, we obtained a seed database of Adenylate and Guanylate cyclase catalytic domains from the Pfam database (PF00211).
+From these sequences, we produced a Hidden Markov Model (HMM) and used this to mine the 45 metazoan, 2 choanoflagellate and 2 filasterea species investigated.
 HMM models were run in HMMR3 with an e-value of 1e−15.
 We ran CD-Hit [@Fu2012] to eliminate redundant sequences (at a 80% threshold).
-To identify clusters, we used the convex-clustering option with 100 jack-knife replicates.
-The NIT-GCs are extremely well conserved in membrane-bound guanylate cyclases and form an easily recognizable cluster.
+To identify clusters, we used sequence-similarity-based clustering in Clans [@frickey2004] and the convex-clustering option with 100 jack-knife replicates.
+The NIT-GCs are extremely well conserved within the membrane-bound guanylate cyclases and form an easily recognizable cluster.
 To analyze the phylogeny of NIT-GCs, the cluster containing these GCs together with membrane-bound guanylate cyclases were parsed and used for tree building.
@@
 To calculate maximum-likelihood trees, we used IQ-tree2 with the LG+G4 model.
-To calculate branch support, we ran 1,000 replicates with the aLRT-SH-like and aBayes methods [@Minh2020].
-The sequences used for the phylogenetic analysis are available in **Supplementary File 1**, the trimmed alignment is available in **Supplementary File 2** and the pre-trimmed data in **Supplementary file 3**.
+To calculate branch support, we ran 1000 replicates with the aLRT-SH-like and aBayes methods [@Minh2020].
+The sequences used for the phylogenetic analysis are available in **Supplementary File 4**, the trimmed alignment is available in **Supplementary File 5** and the pre-trimmed data in **Supplementary file 6**.
 
@@
 
-We used Achim et al. for the single-cell data [@Achim2015].
-In Williams et al., they used 107 cells as neurons by removing duplicates from Achim et al. single-cell data, so we used those cells [@williams2017].
-Since the raw data were read count data, we normalized them to TPM using Python.
-After that we converted them to log10.
-From the sum of the expression levels in 107 cells for each gene, We calculated the percentage in each cell.
-For each cell, we identified them with marker genes.
-After created the data in Python, plotted it using R dot plots.
-RPKM calculates the total number of reads per million bp and then divides by the length of each gene, so it is not possible to compare between samples.
-Instead, TPM first divides by the length of the gene and then divides by the total number of reads per million bases, which allows for more accurate comparisons between samples.
-In this case we wanted to compare between samples, so we used TPM.
+We used a single-cell RNA sequencing dataset from *Platynereis* larvae [@Achim2015] that was further filtered to 107 cells [@williams2017].
+We normalized raw read-count data to TPM and converted to log10.
+To calculate % expression, we used the sum of the expression levels in the 107 cells for each gene.
 The total TPM of each gene between the samples was used to calculate the percentage of expressed genes.
+We processed the data in Python and plotted in R.
+Individual cells were identified by suits of well-characterised marker genes and spatial mapping [@Achim2015].
 The total TPM values for each gene and the percentage of expressed genes are available in **Supplementary file 1**.
@@
 Hairpins associated with the b2 initiator sequence were labeled with Alexa Fluor 647, and the hairpins associated with the b3 initiator sequence were labeled with Alexa Fluor 546.
-To design probes for HCR, we used custom software [@Kuehn2022] to create 20 DNA oligo probe pairs specific to P.
-dumerilii NOS, NIT-GC1, NIT-GC2, RYa-pNP (GenBank accession: JF811330.1), and MLD/pedal 2-pNP (GenBank accession: KF515945.1).
-The NOS, NIT-GC1 and NIT-GC2 probes were designed to be associated with the b2 initiator sequence, while the RYa-pNP and MLD/pedal 2-pNP probes were designed to be associated with the b3 initiator sequence.
+To design probes for HCR, we used custom software [@Kuehn2022] to create 20 DNA oligo probe pairs specific to *P.
+dumerilii* *NOS*, *NIT-GC1*, *NIT-GC2*, *RYa proneuropeptide* (GenBank accession: JF811330.1), and *MLD/pedal peptide 2 proneuropeptide* (GenBank accession: KF515945.1).
+The *NOS*, *NIT-GC1* and *NIT-GC2* probes were designed to be associated with the b2 initiator sequence, while the *RYa* and *MLD/pedal peptide 2* probes were designed to be associated with the b3 initiator sequence.
 For the detection stage, samples were pre-hybridized in 200 µl of probe hybridization buffer for 1 hr at 37°C, and then incubated in 250 µl hybridization buffer containing probe oligos (4 pmol/ml) overnight at 37°C.
@@
 
-Whole-mount immunostaining of 2 day old Platynereis larvae fixed with 4% paraformaldehyde were carried out using primary antibodies raised against NIT-GC1, NIT-GC2, NOS, RYamide neuropeptide, RGWamide neuropeptide in rabbit, plus a commercial antibody raised against acetylated tubulin in mouse (Sigma T7451).
+Whole-mount immunostaining of *Platynereis* larvae fixed with 4% paraformaldehyde was carried out using the primary antibodies listed in KEY RESOURCES TABLE.
+Immunostainings were carried out as previously described [@Conzelmann2012].
+
+**Antibody generation and purification**
+
 The synthetic peptides contained an N-terminal Cys that was used for coupling during purification.
 Antibodies were affinity purified from sera as previously described [@Conzelmann2012].
-Immunostainings were carried out as previously described [@Conzelmann2012].
-The NOS promoter (fragment sizes: 12 Kb) was amplified and cloned upstream of 3xHA- Palmi-tdTomato.
+
+**Transient transgenesis**
+
+The *NOS* promoter (fragment sizes: 12 Kb) was amplified and cloned upstream of 3xHA- Palmi-tdTomato.
 Larvae injected with promoter constructs (ca.
@@
 
-For calcium imaging, 49--55 hpf larvae were used.
-Experiments were performed at room temperature and larvae were immobilised by being embedded in 2.5% agarose filtered artificial seawater between a slide and coverslip spaced with adhesive tape.
+For calcium imaging, 49-55 hpf larvae were used.
+Experiments were carried out at room temperature and larvae were immobilised by being embedded in 2.5% agarose filtered artificial seawater between a slide and coverslip spaced with adhesive tape.
 GCaMP6s mRNA (1 mg/ml) was injected into zygotes as described previously [@Randel2014a].
@@
 A full-length Pdum-NIT-GC1 and -NIT-GC2 coding sequences were amplified by PCR starting from a *Platynereis dumerilii* cDNA library and cloned into the pcDNA3.1(+) vector using the T2A self-cleaving sequence.
-Cos-7 cells with low expression of endogenous soluble guanylate cyclase were used as cultured cells for gene expression.
+COS-7 cells with low expression of endogenous soluble guanylate cyclase were used as cultured cells for gene expression.
 This cell line was purchased from Angio-proteomie (CAT no. cAP-0203).
-The Cos-7 cells were maintained at 37 °C in 35mm dishes (Nunc™ Glass Bottom Dishes) containing 3 mL of DMEM, high glucose glutamax medium (Thermo; Cat.
+The COS-7 cells were maintained at 37 °C in 35mm dishes (Nunc™ Glass Bottom Dishes) containing 3 mL of DMEM, high glucose glutamax medium (Thermo; Cat.
 No.

### 2023-06-07 Manuscript/Manuscript.tex
Ms: Methods edit, key resources table update

@@
 \hypersetup{
-  pdftitle={Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates UV avoidance in zooplankton},
+  pdftitle={Nitric oxide feedback to ciliary photoreceptor cells gates a UV avoidance circuit},
   hidelinks,
@@
 
-\title{Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates
-UV avoidance in zooplankton}
+\title{Nitric oxide feedback to ciliary photoreceptor cells gates a UV
+avoidance circuit}
 \author{true}
@@
 
-Nitric oxide (NO) generated by nitric-oxide synthase (NOS) is a key
-regulator of animal physiology. Here we uncover a role for NO in shaping
-circuit dynamics to orchestrate light-avoidance behaviour. We studied
-UV-avoidance mediated by brain ciliary photoreceptors (cPRCs) in larval
-\emph{Platynereis dumerilii}, a marine annelid. We found NOS
-specifically expressed in interneurons (INNOS) postsynaptic to cPRCs.
-Stimulation of cPRCs by violet light leads to cPRC inhibition but
-concomitant INNOS activation and NO production. NO feeds back to cPRCs
-and triggers their delayed activation through an unconventional guanylyl
-cyclase. This results in the activation of projection interneurons and
-the inhibition of serotonergic ciliomotor neurons. In \emph{NOS}
-mutants, NO feedback and projection-neuron activation do not occur and
-avoidance behaviour is defective. By mathematical modelling, we could
-capture the circuit dynamics in both wild-type and mutant larvae. Our
-results reveal how NO-mediated neuroendocrine signalling gates a
-synaptic circuit to initiate light-avoidance behaviour.
+Nitric oxide (NO) produced by nitric-oxide synthase (NOS) is a key
+regulator of animal physiology. Here we uncover a function for NO in the
+integration of UV exposure and the gating of a UV-avoidance circuit. We
+studied UV/violet avoidance mediated by brain ciliary photoreceptors
+(cPRCs) in larvae of the annelid \emph{Platynereis dumerilii}. In the
+larva, NOS is expressed in interneurons (INNOS) postsynaptic to cPRCs.
+UV stimulation of cPRCs triggers INNOS activation and NO production. NO
+signals retrogradely to cPRCs to induce their sustained post-stimulus
+activation through an unconventional guanylyl cyclase. This late
+activation inhibits serotonergic ciliomotor neurons to induce downward
+swimming. In \emph{NOS} mutants, retrograde signalling, circuit output
+and UV avoidance are defective. By mathematical modelling, we
+recapitulate phototransduction and circuit dynamics in wild-type and
+mutant larvae. Our results reveal how NO-mediated retrograde signalling
+gates a synaptic circuit and induces short-term memory of UV exposure to
+orchestrate light-avoidance behaviour.
 
@@
 
-Neuronally produced nitric oxide (NO) can act as a diffusible
-neuromodulator. This small free-radical molecule is synthesized by
-nitric oxide synthase (NOS)\ldots{}
-
-In the mammalian retina, NO is produced by amacrine cells and activates
-a \emph{soluble guanylate cyclase (sGC)} in cone bipolar cells. The
-NO-dependent increased in cGMP tunes bipolar cell activity \ldots.
-{[}refs{]} {[}other examples{]}
-
-To investigate NO signalling in the modulation of neural circuit
-dynamics, we studied the larval stages of the marine annelid model
-\emph{Platynereis dumerilii} (Ozpolat et al., 2021). We uncovered an
-essential function for NO in these larvae in UV/violet-light avoidance.
-In \emph{Platynereis} larvae, UV/violet avoidance is characterised by
-downward swimming and is mediated by the UV/violet-responding brain
-ciliary photoreceptor cells (cPRCs).
-
-The cPRCs express a vertebrate-type ciliary opsin, c-opsin1 (Arendt et
-al., 2004). This opsin forms a UV-absorbing bistable photopigment with
-an absorption maximum around 384 nm Verasztó et al. (2018).
-\emph{c-opsin1} knockout larvae show defective UV/violet avoidance
-(Verasztó et al., 2018). Here we show that \emph{Platynereis} \emph{NOS}
-is expressed in interneurons of the cPRC circuitry and is required for
-UV/violet-avoidance behaviour. By combining calcium imaging across the
-fully-mapped cPRC circuit, genetic perturbations and mathematical
-modelling, we describe how NO tunes circuit dynamics through
-non-synaptic reverse signalling to cPRCs. This delayed neuroendocrine
-feedback is essential for circuit output, and integrates UV/violet input
-and also induces a short-term memory manifest in altered circuit
-activity.
-
-\#In adult worms, UV light and c-opsin1 modulate locomotion and
-maturation (Veedin Rajan et al., 2021).
+In nervous systems, synaptic transmission and volume transmission
+together shape circuit dynamics (Bargmann and Marder, 2013). While
+synaptic transmission occurs at specialised contact sites, volume
+transmission is characterised by the delocalised release of diverse
+diffusive neuromodulators.
+
+Nitric oxide (NO) is one such modulator with unique physical and
+signalling properties. This free radical synthesized from L-arginine by
+nitric oxide synthase (NOS) is short-lived and can diffuse across
+biological membranes (Cudeiro and Rivadulla, 1999; Thomas, 2015).
+Canonical NO signalling involves the Ca2+/calmodulin-dependent
+activation of NOS, NO production and diffusion, and the NO-dependent
+activation of soluble guanylate cyclases (sGC). This can occur
+cell-autonomously or in other cells leading to cGMP production (Bredt et
+al., 1990; Hölscher, 1997). Given that NOS activation is calcium
+dependent and NOS shows neuron-type-specific expression (Aso et al.,
+2019; Gibbs and Truman, 1998; Mobley et al., 2022; Wildemann and Bicker,
+1999), NO action can lead to the activity-dependent modulation of neural
+circuits at specific sites (Aso et al., 2019; Jacoby et al., 2018;
+Vielma et al., 2014; Wang et al., 2007) .
+
+In the vertebrate retina, NOS is expressed in amacrine, ganglion and
+other cells and the actions of NO can be diverse (Cudeiro and Rivadulla,
+1999; Jacoby et al., 2018; Wang et al., 2007). For example, defective NO
+signalling in NOS knockouts leads to a decreased sensitivity of retinal
+ganglion cells to light stimulation (Wang et al., 2007). In the retina,
+NO signalling can also involve pathways other than canonical sGC
+signalling (Jacoby et al., 2018; Tooker et al., 2013; Wei et al., 2012).
+Due to the complex expression of NOS in vertebrates and the diversity of
+its functions, it has been challenging to link the neurophysiological
+effects of NO to signalling mechanisms and behaviour change.
+
+In the \emph{Drosophila} brain, NO signalling is also involved in tuning
+circuit activity at diverse and specific sites. In the ellipsoid body of
+the central complex, NO is involved in visual working memory (Kuntz et
+al., 2017). NO signalling also tunes the dynamics of associative
+memories in mushroom body circuits. NOS is expressed in PPL1-γ1pedc
+mushroom-body neurons where it is involved in shortening memory
+retention while promoting fast memory updating in response to new
+experiences (Aso et al., 2019).
+
+In the whole organism context, NO signalling has often been studied in
+diverse marine invertebrates, where NO can regulate larval settlement
+and metamorphosis (Leise et al., 2001; Locascio et al., 2022; Song et
+al., 2021; Ueda et al., 2016; Zhang et al., 2012). While NOS-expressing
+neurons potentially responsible for these effects have been reported
+(Bishop and Brandhorst, 2007; Locascio et al., 2022), it has not been
+possible to link these NO-dependent effects on behaviour or life-cycle
+transitions to neuronal activity and function.
+
+Overall, we still know little about how NO production relates to
+stimulus conditions, how it shapes circuit activity at specific neuron
+types, and how NO-dependent modulation relates to behaviour.
+
+To investigate NO function in neural circuit dynamics and behaviour,
+here we study larvae of the marine annelid model \emph{Platynereis
+dumerilii} (Ozpolat et al., 2021). \emph{Platynereis} has emerged as a
+model for systems neuroscience with a toolkit that enables the
+combination of behavioural analysis and neuronal activity imaging with
+genetic manipulations. A whole-body synaptic connectome and gene
+expression atlases are also available and can be integrated with
+functional approaches in live animals thanks to the cellular-level
+stereotypy of larvae of the same developmental stage (Ozpolat et al.,
+2021; Verasztó et al., 2020; Verasztó et al., 2017; Vergara et al.,
+2021).
+
+Here we uncovered an essential function for NO signalling in larval
+UV/violet-light avoidance behaviour. In \emph{Platynereis} larvae,
+UV/violet avoidance is mediated by brain ciliary photoreceptor cells
+(cPRCs) and is characterised by downward swimming (Verasztó et al.,
+2018). The cPRCs express a ciliary-type opsin, c-opsin1 (Arendt et al.,
+2004) that forms a UV-absorbing bistable photopigment with an absorption
+maximum around 384 nm (Tsukamoto et al., 2017; Veedin Rajan et al.,
+2021; Verasztó et al., 2018). Upon UV/violet exposure, the cPRC show a
+characteristic biphasic calcium response that is \emph{c-opsin1}
+dependent. UV/violet avoidance is also \emph{c-opsin1}-dependent and is
+defective in \emph{c-opsin1} mutants (Verasztó et al., 2018). Here we
+show that \emph{Platynereis} \emph{NOS} is expressed in interneurons of
+the cPRC circuit and is required for UV/violet-avoidance. By combining
+calcium imaging across the fully-mapped cPRC circuit (Verasztó et al.,
+2018) with genetic perturbations and mathematical modelling, we describe
+how NO tunes circuit dynamics through non-synaptic retrograde signalling
+to cPRCs. This delayed neuroendocrine feedback integrates UV/violet
+exposure to induce a short-term memory manifested in altered circuit
+activity and an aversive behavioural response.
 
@@
 Phylogenetic analysis of NOS proteins indicate that \emph{Platynereis}
-NOS belongs to an orthology group of metazoan NOS sequences
-\textbf{(Figure 1---figure supplement 1)}. To characterise the
-expression pattern of \emph{NOS} we used in situ hybridization chain
-reaction (HCR) and transient transgenesis. In two- and three-day-old
-larvae, we detected \emph{NOS} expression in four cells (two of them
-weakly expressing) in the apical organ region \textbf{(Figure 1D and
-Figure 1---figure supplement 2)}. \emph{NOS} was also expressed in the
-region of the visual eyes (adult eyes) and the pigmented eyespots
-\textbf{(Figure 1---figure supplement 1)}. The four apical organ cells,
-but not the eyes, were also labelled with a \emph{NOS} reporter
-construct driving palmitoylated-tdTomato \textbf{(Figure 1E)}. This
-reporter also revealed the axonal projections of these central
+NOS belongs to an orthology group of bilaterian NOS sequences (Figure
+1---figure supplement 1). To characterise the expression pattern of
+\emph{NOS} we used in situ hybridization chain reaction (HCR) and
+transient transgenesis. In two- and three-day-old larvae, we detected
+\emph{NOS} expression in four cells (two of them weakly expressing) in
+the apical organ region (Figure 1D and Figure 1---figure supplement 2).
+\emph{NOS} was also expressed in the region of the visual eyes (adult
+eyes) and the pigmented eyespots (Figure 1---figure supplement 2). The
+four apical organ cells, but not the eyes, were also labelled with a
+\emph{NOS} reporter construct driving palmitoylated-tdTomato (Figure
+1E). This reporter also revealed the axonal projections of these central
 NOS-expressing neurons. The position and morphology of the four
@@
 interneurons (INNOS) in our three-day-old whole-body \emph{Platynereis}
-volume EM data (Verasztó et al., 2020; Williams et al., 2017)
-\textbf{(Figure 1B,C)}. In the synaptic connectome, the INNOS cells are
-postsynaptic to the UV-sensory cPRCs and presynaptic to the INRGW
-interneurons, which are also cPRC targets \textbf{(Figure 1C,F)}. INRGW
-neurons synapse on the head serotonergic ciliomotor neurons (Ser-h1),
-which synapse on the prototroch ciliary band and the cholinergic MC
-ciliomotor neurons \textbf{(Figure 1C,F)}.
+volume EM data (Verasztó et al., 2020; Williams et al., 2017) (Figure
+1B,C). In the synaptic connectome, the INNOS cells are postsynaptic to
+the UV-sensory cPRCs and presynaptic to the INRGWa interneurons, which
+are also cPRC targets (Figure 1C,F). The projections of INNOS cells are
+segregated into input (dendritic) and output (axonal) compartments, with
+cPRC inputs occurring on the dendritic part in the dense neurosecretory
+plexus. INNOS output synapses form in the more ventral projection region
+(Figure 1---figure supplement 3). INRGWa neurons synapse on the head
+serotonergic ciliomotor neurons (Ser-h1), which synapse on the
+prototroch ciliary band and the cholinergic MC ciliomotor neurons
+(Figure 1C,F) (Verasztó et al., 2017).
 
@@
 \includegraphics[width=25in]{figures/Fig1} \caption{**Figure 1.
-Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** **(A)** SEM image of a three-day-old Platynereis larva.
-**(B, C)** Five types of neurons (cPRC, INNOS, INRGW, Ser-h1 and MC) in the cPRC circuit reconstructed in the whole-body transmission electron microscopy (ssTEM) dataset of a three-day-old larva.
-Ciliated cells labelled with (B) are shown in grey.
-The INNOS on the ventral side labelled with (C) are shown in purple.
-Axons and dendrites appear as lines and cell body positions are represented by spheres.
-**(D)** Expression analysis of the NOS gene (magenta) using in situ HCR.
-Anterior view of the larva at two-day-old.
-Co-stained image using acetylated α-tubulin antibody (acTub: green).
-**(E)** Immunostaining with anti-HA antibody against HA tagged palmitoylated-tdTomato expressed under the upstream of the start site of NOS (NOSp::palmi-3xHA: magenta).
-Co-staining with acetylated α-tubulin antibody as marker for cilia of cPRC and axonal scaffold (acTub: green).
-Views of larvae at three-day-old from the anterior side.
-Insert: One of the reconstructed INNOS is shown from the anterior side.
-**(F)** Wiring diagram of the cPRC circuit.
-Hexagons and arrows represent cell groups and their synaptic connections, respectively.
-Numbers inside hexagons indicate the number of cells grouped in each node.
-Arrow line thickness (synapse weight) is equal to the number of synapses.}\label{fig:unnamed-chunk-1}
+Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** **(A)** Scanning electron microscopy image of a three-day-old *Platynereis* larva.
+**(B, C)** Volume rendering of the neuron types (cPRC, INNOS, INRGWa, Ser-h1 and MC) in the cPRC circuit reconstructed from a whole-body transmission electron microscopy volume of a three-day-old larva.
+Neurite skeletons are shown with cell-body positions represented by spheres.
+Projections of all neurons in the body are shown in grey to highlight the neuropils.
+The outline of the yolk is also indicated in grey.
+In B, nuclei positions of the prototroch head ciliary band are shown as grey spheres.
+**(D)** Expression of the *NOS* gene detected by in situ HCR (magenta) in a two-day-old larva (anterior view).
+Antibody staining for acetylated α-tubulin (acTub: green) highlights cPRC cilia and the neuropil.
+**(E)** Expression of a *NOS* reporter (NOSp::palmi-3xHA: magenta) labelled with an anti-HA antibody in a two-day-old larva (anterior view).
+Antibody staining for acetylated α-tubulin (acTub: green) highlights cPRC cilia and the neuropil.
+**(F)** Synaptic wiring diagram of the cPRC circuit.
+Hexagons represent cell groups, with the number of cells per group shown in square brackets.
+Arrows represent the summed number of synaptic contacts between cell groups.
+Arrow thickness is proportional to the number of synapses.}\label{fig:unnamed-chunk-1}
 \end{figure}
 
-\textbf{Nitric oxide produced during UV/violet stimulation gates the
-output of the cPRC circuit}
+\textbf{Nitric oxide is produced during UV/violet stimulation of the
+cPRCs}
 
@@
 NO-reporter DAF-FM into zygotes and imaged two-day-old larvae while
-exposing the ramified sensory cilia of the cPRCs to 405 nm violet light.
-Following light stimulation of the cPRCs but not a control area, we
+exposing the region of cPRC cilia to 405 nm violet light. To mark cell
+outlines, we coinjected mRNA encoding a red-fluorescent reporter
+(RGECO), allowing the identification of the cPRCs. Following light
+stimulation in the region of the ramified cilia of the cPRCs, we
 detected an increase in DAF-FM fluorescence in the anterior
-neurosecretory neuropil, the region of INNOS projections \textbf{(Figure
-2)}.
+neurosecretory neuropil, the region of INNOS projections. This increase
+did not occur in larvae where we illuminated a control area (Figure 2).
 
@@
 \includegraphics[width=29.17in]{figures/Fig2} \caption{**Figure 2.
-NO produced by UV/violet stimulation to cPRCs.** **(A)** A NO-positive cell visualised by DAF-FM (green) is shown (dashline).
-The white line indicates the outer frame of the larva.
-All subsequent images are of two-day-old larvae.
-**(B)** Intensity changes before (upper) and during 405 nm light stimulation (lower) in a cell visualised by DAF-FM.
-**(C)** The intensity of DAF-FM changes over time when cPRC (green) and other cells (ctr: gray) are stimulated with UV/violet (405 nm) light.
-Purple squares indicate the timing of 405 nm light stimulation, the transparent lines show the optical responses to individual intensities, each intensities are normalized (ΔF/F0) and the dark trace lines and the shaded bands show the mean value and the standard error range of the mean.}\label{fig:unnamed-chunk-2}
+NO produced by UV/violet stimulation to cPRCs.** **(A)** DAF-FM fluorescence in the region of the neurosecretory neuropil.
+The white line indicates the outline of the larva.
+The dashed line corresponds to the area where fluorescence was quantified.
+Circles indicate the location of cPRC and control stimulation.
+cPRCs are marked by thin lines.
+**(B)** Changes in DAF-FM fluorescence before and during 405 nm light stimulation.
+**(C)** Changes in DAF-FM fluorescence over time during 405 nm stimulation of the cPRCs (green) or a control area (ctr: gray).
+The purple box indicates the duration of 405 nm stimulation.
+Individual traces normalized (ΔF/F0) are shown as thin lines.
+Thick lines show the mean value with 0.95 confidence intervals.
+N = 9 larvae for control and 11 for cPRC stimulation.}\label{fig:unnamed-chunk-2}
 \end{figure}
@@
 
-Next, we wanted to directly test whether NO signalling is involved in
-UV/violet avoidance. To achieve this, we generated two
-\emph{Platynereis} \emph{NOS} knockout lines with the CRISPR/Cas9
-method. We recovered two deletions (NOSΔ11/Δ11 and Δ23/Δ23), both
-leading to frame-shift and an early stop codon and thus likely
-representing null alleles \textbf{(Figure 3A)}. We could establish a
-homozygous line for both mutations indicating that NOS is not an
-essential gene. To quantify UV avoidance, we recorded the trajectories
-of freely swimming wild type and mutant larvae in vertical columns,
-illuminated laterally from two opposite sides with 395 nm UV light
-\textbf{(Figure 3B, C and Figure 3---figure supplement 1A-D)}. As
-previously shown, wild-type larvae swim downward following
-non-directional UV/violet light stimulation (Verasztó et al., 2018). In
-contrast, both two- and three-day-old homozygous \emph{NOS}-mutant
-larvae showed a strongly diminished UV-avoidance response
-\textbf{(Figure 3B-D and Figure 3---figure supplement 1)}. This
+We next tested whether NO signalling regulates UV/violet avoidance. To
+achieve this, we generated two \emph{Platynereis} \emph{NOS} knockout
+lines with the CRISPR/Cas9 method. We recovered two deletions
+(NOSΔ11/Δ11 and Δ23/Δ23), both frame-shift mutations leading to an early
+stop codon and thus likely representing null alleles (Figure 3---figure
+supplement 1A). We could establish a homozygous line for both mutations
+indicating that NOS is not an essential gene in \emph{Platynereis}. To
+quantify UV avoidance, we recorded the trajectories of freely swimming
+wild type and mutant larvae in vertical columns, illuminated laterally
+from two opposite sides with 395 nm UV light (Figure 3A and Figure
+3---figure supplement 1B). As previously shown, wild-type larvae swim
+downward following non-directional UV/violet light stimulation (Verasztó
+et al., 2018). In contrast, both two- and three-day-old homozygous
+\emph{NOS}-mutant larvae showed a strongly diminished UV-avoidance
+response (Figure 3A, B and Figure 3---figure supplement 1B,C). This
 phenotype is similar to the defective UV-avoidance of \emph{c-opsin1}
-mutant larvae and reveals a requirement for NOS in UV-avoidance
-behaviour (Verasztó et al., 2018). Wild type but not mutant larvae also
-showed an increase in swimming speed under UV light that may be due to
-downward swimming trajectories (swimming with gravity) \textbf{(Figure
-3---figure supplement 1C,D)}. We also tested directional phototaxis, by
-exposing larvae to 480 nm directional and collimated light from the top
-of the column. Three-day-old but not two-day-old \emph{NOS}-mutant
-larvae also showed reduced phototactic behaviour, suggesting a function
-for \emph{NOS} in the visual eyes that mediate three-day-old phototaxis
-(Randel et al., 2014) \textbf{(Figure 1---figure supplement 1A,B)}. To
-distinguish between an acute or developmental function of NOS in light
-responses, we next tested larvae exposed to the NOS inhibitor L-NAME.
-Larvae incubated for 5 min in 0.1 mM or 1 mM L-NAME showed a
+mutant larvae (Verasztó et al., 2018) and reveals a requirement for NOS
+in UV-avoidance behaviour. Wild type but not mutant larvae also showed
+an increase in swimming speed under UV light that may be due to downward
+swimming trajectories (swimming in the direction of gravity) (Figure 3B
+and Figure 3---figure supplement 1C). We also tested directional
+phototaxis, by exposing larvae to 480 nm directional collimated light
+from the top of the column. Three-day-old but not two-day-old
+\emph{NOS}-mutant larvae also showed reduced phototactic behaviour,
+suggesting a function for \emph{NOS} in the visual eyes that mediate
+three-day-old phototaxis (Randel et al., 2014) (Figure 3D and Figure
+3---figure supplement 1G).
+
+To distinguish between an acute and developmental function of NOS in
+light responses, we next tested larvae exposed to the NOS inhibitor
+L-NAME. Larvae incubated for 5 min in 0.1 mM or 1 mM L-NAME showed a
 dose-dependent inhibition of UV avoidance. In contrast, phototaxis was
-not effected \textbf{(Figure 3E)}. Overall, our results indicate an
-acute requirement for NOS signalling in UV-avoidance and a possible
-indirect, developmental role in visual phototaxis.
+not affected (Figure 3C, E). Overall, our results indicate an acute
+requirement for NOS signalling in UV-avoidance and a possible indirect,
+developmental role in the visual system, reminiscent of the function of
+NO signalling in \emph{Drosophila} eye development (Gibbs and Truman,
+1998).
 
@@
 \includegraphics[width=33.33in]{figures/Fig3} \caption{**Figure 3.
-Larvae with inhibited NOS have a delayed response to UV-avoidance.** **(A)** Top: The domain and Exon/Intron structure of NOS.
-Bottom: Close-up region showing the genomic locus of NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9.
-The generated mutants (NOSΔ11, NOSΔ23) are also shown.
-Pink indicates target sites.
-Gray shows PAM sequences, red shows stop codons.
-**(B)** Overlaid trajectories for WT (n=32) and NOS mutant (NOSΔ11/Δ11, n=26 and NOSΔ23/Δ23, n=47) at three-day-old larvae.
-0 sec as the starting point.
-After 10 sec, UV (395 nm) stimulation from the side.
-**(C)** The temporal changes in the vertical position of the WT and mutant larvae before (dark) and after UV stimulation (side 395 nm) are shown.
-The starting points of each larval trajectory are set to 0.
-**(D)** Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23) larvae at three-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
-The data are shown in 30 s bins.
-**(E)** Vertical swimming in larvae treated with NOS inhibitors, L-NAME at three-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
-The data are shown in 30 s bins.}\label{fig:unnamed-chunk-3}
+NOS is required for UV avoidance in *Platynereis* larvae.**  **(B)** Swimming trajectories of wild type (WT, n=32) and NOS mutant (NOSΔ11/Δ11, n=26 and NOSΔ23/Δ23, n=47) three-day-old larvae.
+All trajectories start at 0 x and y position and time 0 corresponding to 10 sec after the onset of 395 nm stimulation from the side.
+**(B)** Vertical position of batches of wild type and mutant three-day-old larvae over time under 395 nm UV stimulation.
+The starting position of each larval trajectory was set to 0.
+**(C)** Vertical position of batches of control and L-NAME-treated (0.1 and 1 mM) three-day-old larvae over time under 395 nm UV stimulation.
+The starting position of each larval trajectory was set to 0.
+**(D)** Vertical displacement in 30 sec bins of wild type and mutant (NOSΔ11 and NOSΔ23) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
+**(E)** Vertical displacement in 30 sec bins of control and L-NAME-treated (0.1 and 1 mM) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.}\label{fig:unnamed-chunk-3}
 \end{figure}
 
-\textbf{NO feedback tunes cPRC responses to UV/violet stimulation}
+\textbf{NO retrograde signalling tunes cPRC responses to UV/violet
+stimulation}
 
@@
 signals during 405 nm stimulation of the cPRCs. As we have shown
-previously, a 20-sec local stimulation of cPRC cilia lead to a transient
+previously, a 20-sec local stimulation of cPRC cilia led to a transient
 increase in cPRC calcium levels, followed by a transient decrease
@@
 cPRCs were raising again, reaching higher levels than at the start of
-the stimulus -- a response that may involve depolarisation. This
-activation phase occurs after the 20 sec stimulation period and is
-likely due to a delayed neuroendocrine feedback (Verasztó et al., 2018).
-To determine whether NO mediates such a feedback, we repeated the
+the stimulus -- a response that may involve depolarisation (Figure 4A,
+B). This activation phase occurs after the 20 sec stimulation period and
+is likely due to a delayed neuroendocrine feedback (Verasztó et al.,
+2018). To determine whether NO mediates such a feedback, we repeated the
 experiment in \emph{NOS}-mutant larvae. While we detected the initial
-activation phase followed by inhibition, in \emph{NOS}-mutants this was
-not followed by delayed activation and calcium levels dropped to a low
-steady-state level \textbf{(Figure 4E,F)}. We thus identified a
-requirement for NO signalling in the late-phase activation of cPRCs.
+activation phase followed by inhibition, in homozygous
+\emph{NOS}-mutants for both CRISPR alleles this was not followed by
+delayed activation. Instead, calcium levels dropped to a low
+steady-state level (Figure 4A, B). We thus identified a requirement for
+NO signalling in the late-phase activation of cPRCs.
 
@@
 
-Next, we aimed to identify the NO-receptor in the cPRCs. NO generally
+We aimed next to identify the NO receptor in the cPRCs. NO generally
 acts via soluble guanylate cyclases (sGC), belonging to the guanylate
@@
 (nitrite/nitrate sensing) domain (PF08376) (NIT-GC) as potential
-mediators of NO signalling (Moroz et al., 2020). To search for NIT-GCs
-in \emph{Platynereis}, we searched transcriptome resources and retrieved
-15 potential NIT-GC homologs \textbf{(Figure 4---figure supplement 1)}.
-To analyse the relationship of these sequences to metazoan NIT-GCs, we
+mediators of NO signalling (Moroz et al., 2020). To identify NIT-GCs in
+\emph{Platynereis}, we searched transcriptome resources and retrieved 15
+potential NIT-GC homologs (Figure 4---figure supplement 1 and 2). To
+analyse the relationship of these sequences to metazoan NIT-GCs, we
 retrieved protein sequences with a CYC domain from the transcriptome and
-genome databases of 45 metazoan species and 2 choanoflagellate species.
-We carried out cluster analysis and did phylogenetic reconstruction on a
+genome databases of 45 metazoan and 2 choanoflagellate species. We
+carried out cluster analysis and did phylogenetic reconstruction on a
 group of membrane-bound guanylyl cyclases with sGCs as an outgroup. In
@@
 some ecdysozoans, echinoderms, and lophotrochozoans. The 15
-\emph{Platynereis} sequences were parts of several major and ancient
-clades of the tree \textbf{(Figure 4---figure supplement 1)}. To
-characterise the expression of NIT-GCs, we used previously published
+\emph{Platynereis} sequences belonged to several deeply diverged clades
+in the phylogenetic tree (Figure 4---figure supplement 1 and 2).
+
+To characterise the expression of NIT-GCs, we used previously published
 spatially mapped single-cell transcriptome data (Achim et al., 2015;
@@
 specific expression in the cPRCs and one was expressed in the INNOS
-cells \textbf{(Figure 4---figure supplement 1)}. In the single-cell
-data, we could identify the cPRCs by the specific expression of c-opsin1
-and a pedal-peptide2 neuropeptide precursor (MLD proneuropeptide)
-(Arendt et al., 2004; Williams et al., 2017), which have previously been
-described as cPRC markers \textbf{(Figure 4---figure supplement 2A)}.
-The INNOS cells were identified by NOS expression and spatial mapping in
-the brain (Achim et al., 2015). We decided to focus on two NIT-GCs
-expressed in the cPRCs and with a full-length sequence, NIT-GC1 and
-NIT-GC2. To confirm the single-cell data, we first carried out in situ
-HCR with probes for \emph{NIT-GC1} and \emph{NIT-GC2} mRNA. Both genes
-were specifically expression in the four cPRCs, as confirmed by
-co-labeling with an acetylated α-tubulin antibody and with an HCR probe
-against \emph{pedal peptide 2/MLD proneuropeptide} \textbf{(Figure 4A,B
-and Figure 4---figure supplement 2A-C)}. To analyse the subcellular
-localisation of NIT-GC1 and NIT-GC2 at the protein level, we raised and
-affinity-purified polyclonal antibodies against a specific peptide
-sequence from both proteins. In immunostainings, we found that NIT-GC1
-was localise to the region corresponding to the axonal projections of
-the cPRCs in the anterior nerve plexus \textbf{(Figure 4C)}.
-Co-immunostaining with the rabbit NIT-GC1 antibody and a custom rat
-antibody raised against a fragment of \emph{Platynereis} NOS, revealed
-that both proteins were localised in close proximity in the
-neurosecretory plexus \textbf{(Figure 4---figure supplement 2D)}. In
-contrast, NIT-GC2 specifically labelled the large sensory ciliary region
-of the cPRCs \textbf{(Figure 4D)}. These different subcellular
-localisations suggest that the two NIT-GCs are involved in different
-intracellular signalling processes in the ciliary and axonal regions of
-the cPRCs.
+cells (Figure 4---figure supplement 2). In the single-cell data, we
+could identify the cPRCs by the specific expression of \emph{c-opsin1}
+and the \emph{pedal-peptide2 neuropeptide precursor} (\emph{MLD
+proneuropeptide}), previously described cPRC markers (Arendt et al.,
+2004; Williams et al., 2017) (Figure 4---figure supplement 3A). The
+INNOS cells were identified by NOS expression and spatial mapping in the
+brain (Achim et al., 2015). We decided to focus on two NIT-GCs expressed
+in the cPRCs and with a full-length sequence, \emph{NIT-GC1} and
+\emph{NIT-GC2}. To confirm the single-cell data, we first carried out in
+situ hybridisation chain reaction (HCR) with probes for \emph{NIT-GC1}
+and \emph{NIT-GC2} mRNA. Both genes were specifically expression in the
+four cPRCs, as confirmed by co-labeling with an acetylated α-tubulin
+antibody and with an HCR probe against \emph{pedal peptide 2/MLD
+proneuropeptide} (Figure 4C, D and Figure 4---figure supplement 3A-C).
+To analyse the subcellular localisation of NIT-GC1 and NIT-GC2 at the
+protein level, we raised and affinity-purified polyclonal antibodies
+against a specific peptide sequence from both proteins. In
+immunostainings, we found that NIT-GC1 was localise to the region
+corresponding to the axonal projections of the cPRCs in the anterior
+neurosecretory plexus (Figure 4E). Co-immunostaining with the rabbit
+NIT-GC1 antibody and a custom rat antibody raised against
+\emph{Platynereis} NOS revealed the localisation of both proteins in
+dots in the neurosecretory neuropil (Figure 4---figure supplement 3D).
+In contrast, NIT-GC2 specifically labelled the ramified sensory cilia of
+the cPRCs (Figure 4F). These different subcellular localisations suggest
+that the two NIT-GCs are involved in different intracellular signalling
+processes in the ciliary and axonal regions of the cPRCs.
 
@@
 conserved cyclase domain that is expected to catalyse cGMP synthesis
-\textbf{(Figure 4I)}. The NIT domain may render NIT-GC1 dependent on NO
-signals. To test this, we co-expressed the synthetic cGMP indicator
-Green cGull (Matsuda et al., 2016) and NIT-GC1 in cultured cells COS-7
-(monkey kidney). For balanced expression, we used a single plasmid with
-the two open-reading frames separated by the 2A self-cleaving peptide
-\textbf{(Figure 4J)}. Application of the NO donor SNAP lead to increased
-cGMP levels, an effect we did not observe when cells were exposed to
-DMSO or when Green cGull was expressed alone \textbf{(Figure 4K-M)}. To
-test whether cGMP production is dependent on the NIT domain, we also
-tested a deletion construct of NIT-GC1 lacking the NIT domain
-\textbf{(Figure 4I)}. Cells expressing this construct and Green cGull
-did not show increased cGMP levels when exposed to SNAP \textbf{(Figure
-4N)}. These results indicate that NIT-GC1 is able to catalyse cGMP
-production in an NO-dependent manner and this function requires the NIT
-domain. These results establish NIT-GC1 as a biochemical NO sensor.
-
-\textbf{NIT-GC1 is required for NO-mediated feedback to cPRCs during the
-UV response}
+(Figure 4G). The NIT domain may render NIT-GC1 dependent on NO signals.
+To test this, we co-expressed the cGMP indicator Green cGull (Matsuda et
+al., 2016) and NIT-GC1 in cultured COS-7 (monkey kidney) cells, a cell
+line with minimal endogenous sGC activity (Matsuda et al., 2016). For
+balanced expression, we used a single plasmid with the two open-reading
+frames separated by the 2A self-cleaving peptide (Figure 4H).
+Application of the NO donor SNAP lead to increased Green cGull
+fluorescence, an effect we did not observe when cells were exposed to
+DMSO or when Green cGull was expressed alone (Figure 4I-K). To test
+whether this effect is dependent on the NIT domain, we also tested a
+deletion construct of NIT-GC1 lacking the NIT domain (Figure 4G). Cells
+expressing this construct and Green cGull did not show an increased
+fluorescence of the cGMP reporter when exposed to SNAP (Figure 4L).
+These results indicate that NIT-GC1 is able to catalyse cGMP production
+in an NO-dependent manner and this function requires the NIT domain.
+These results establish NIT-GC1 as a biochemical sensor of NO or its
+derivatives.
+
+\textbf{NIT-GC1 is required for NO-mediated retrograde signalling to
+cPRCs during the UV response}
 
@@
 tested knockdown efficiency by immunostaining injected animals with the
-NIT-GC1 and NIT-GC2 antibodies \textbf{(Figure 4---figure supplement
-2E,F)}. For both genes, the morpholinos led to a strong reduction in the
-respective antibody signal, confirming efficient knockdown and antibody
+NIT-GC1 and NIT-GC2 antibodies (Figure 4---figure supplement 3E,F). For
+both genes, the morpholinos led to a strong reduction in the respective
+antibody signal, confirming efficient knockdown and antibody
 specificity.
@@
 In NIT-GC1 morphant larvae, the delayed activation of cPRCs following
-405 nm stimulation did not occur \textbf{(Figure 4G)}. This phenotype is
-similar to the phenotype of \emph{NOS} mutants suggesting that NIT-GC1
-acts as the NO sensor in cPRCs to drive their delayed activation. This
-could occur via increased cGMP production and the opening of a
-cyclic-nucleotide-gated (CNG) channel specific to cPRCs. NIT-GC2
-morphant larvae, in contrast, showed a step-up increase in calcium
-following light stimulation \textbf{(Figure 4H)}. The calcium signal
-decayed during stimulation and was off after light off. These data
-support an essential role for ciliary-localised NIT-GC2 in suppressing
-cPRC calcium following its transient rise at stimulus onset. Overall,
-these knockdown experiments revealed different signalling mechanisms for
-the two NIT-GCs that may be due to their different subcellular
-localisations.
+405 nm stimulation did not occur (Figure 4M). This phenotype is similar
+to the phenotype of \emph{NOS} mutants suggesting that NIT-GC1 acts as
+the NO sensor in cPRCs to drive their delayed activation. This could
+occur via increased cGMP production and the opening of a
+cyclic-nucleotide-gated ion channel (CNG) specific to cPRCs (Tosches et
+al., 2014). NIT-GC2 morphant larvae, in contrast, showed a step-up
+increase in calcium following light stimulation (Figure 4N). The calcium
+signal decayed during stimulation and was off after light off. These
+data support an essential role for ciliary-localised NIT-GC2 in
+suppressing cPRC calcium following its transient rise at stimulus onset.
+Overall, these knockdown experiments revealed different signalling
+mechanisms for the two NIT-GCs that may be due to their different
+subcellular localisations.
 
@@
 \includegraphics[width=38.89in]{figures/Fig4} \caption{**Figure 4.
-cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.** **(A, B)** Expression analysis of the NIT-GC1 (C) and NIT-GC2 (D) gene (magenta) using in situ HCR.
-Co-staining image using acetylated α-tubulin antibody (ac α-tub: green).
-**(C, D)** Localisation analysis using NIT-GC1 and NIT-GC2 antibodies.
-Green shows co-staining with acetylated α-tubulin antibody (ac α-tub).
-**(E, F)** Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in WT (dark yellow) and NOS mutant (E: NOSΔ11/Δ11, F: NOSΔ23/Δ23) (light pink) larvae.
-**(G, H)** Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in NIT-GC1 (G: NIT1) and NIT-GC2 (H: NIT2) knockdown by morpholino (MO1 and 2) larvae.
-**(I)** The domain structure of NIT-GC1.
-Below shows the mutant NIT-GC1 (ΔNIT-GC1) with deletion of the NIT domain.
-Grey squares indicate the transmembrane region (TM).
-**(J)** A schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control.
-Green cGull shows cGMP-dependent green fluorescence.
-**(K-N)** The intensity changes in cGMP fluorescence over time for the four conditions are shown.
-Gray traces show the optical responses to individual intensities (n > 6).
-The individual intensities are normalized (ΔF/F0).
-The dark line shows the mean value and the band the standard error range of the mean.
-The respective solution was acted on 2 min after the start of imaging (grey bars).}\label{fig:unnamed-chunk-4}
+NOS and two NIT-GCs shape calcium signals during cPRC UV/violet response.** **(A, B)** GCaMP6s signals in cPRCs in wild type and *NOS* mutant (A, NOSΔ11/Δ11, B, NOSΔ23/Δ23) larvae during 405 nm light stimulation.
+**(C, D)** In situ HCR for (C) *NIT-GC1* and (D) *NIT-GC2* (magenta) in three-day-old *Platynereis* larvae.
+Larvae were co-stained with an antibody against acetylated α-tubulin to label cPRC cilia and the neuropil (green).
+**(E, F)** Immunostaining for (E) NIT-GC1 and (F) NIT-GC2 (magenta), co-stained for acetylated α-tubulin (green).
+**(G)** The domain structure of *Platynereis* NIT-GC1 and the truncated NIT-GC1ΔNIT protein lacking the NIT domain.
+A predicted transmembrane region (TM) is shown in grey.
+**(H)** Schematic of the cell-based assay to detect cGMP production following the addition of an NO donor SNAP or DMSO as control.
+**(I-L)** Green cGull fluorescence over time for the four conditions tested.
+Individual responses and their mean with 0.95 confidence interval are shown (n > 6 cells).
+Intensities are normalized (ΔF/F0).
+The indicated chemicals were added at 2 min after the start of imaging (grey bars).
+**(M, N)** GCaMP6s signals in cPRCs in (M) NIT-GC1 and (N) NIT-GC2 morphant larvae during 405 nm light stimulation.
+Individual responses and their mean with 0.95 confidence interval are shown.}\label{fig:unnamed-chunk-4}
 \end{figure}
@@
 wild type, mutant and morphant larvae. We were able to image the
-activity of all neurons in the cPRC circuit (INNOS, INRGW, Ser-h1 and
+activity of all neurons in the cPRC circuit (INNOS, INRGWa, Ser-h1 and
 MC). The MC cell was identified based on its position and intrinsic
-activity. To unambiguously identify all other cells from which we
-recorded calcium signals, we developed an on-slide immunostaining method
-\textbf{(Figure 5---figure supplement 1)}. We used the cell-specific
-antibody markers against RYamide (INNOS), RGWamide (INRGW) and serotonin
-(Ser-h1) (Conzelmann et al., 2011) to immunostain agar-embedded larvae
-following calcium imaging. Based on the position of the nuclei, we could
-correlate live and fixed samples at a single-cell precision
-\textbf{(Figure 5A,B)}. Due to the stereotypy of the larvae, we could
-also identify neurons based on their position and activity in
-activity-correlation maps \textbf{(Figure 5C)}.
-
-We first quantified the responses of the INNOS and INRGW interneurons
+activity (Verasztó et al., 2017). To unambiguously identify all other
+cells from which we recorded calcium signals, we developed an on-slide
+immunostaining method (Figure 5---figure supplement 1A). We used the
+cell-specific antibody markers against RYamide (INNOS) (Figure
+5---figure supplement 1B-D), RGWamide (INRGWa) and serotonin (Ser-h1)
+(Conzelmann et al., 2011) to immunostain agar-embedded larvae following
+calcium imaging. Based on the position of the nuclei, we could correlate
+live and fixed samples at a single-cell precision (Figure 5A,B). Due to
+the stereotypy of the larvae, we could also identify neurons based on
+their position and calcium activity in activity-correlation maps (Figure
+5C).
+
+We first quantified the responses of the INNOS and INRGWa interneurons
 during 405 nm stimulation of the cPRCs. In both wild type and
 \emph{NOS}-mutant larvae, INNOS cells showed an increase in calcium
-during stimulation \textbf{(Figure 5D)}. This response was suppressed in
-NIT-GC2 morphant larvae \textbf{(Figure 5E)} revealing an essential role
-for NIT-GC2-mediated cPRC suppression in INNOS activation. In contrast,
-INRGW cells were initially inhibited, followed by a delayed activation
-paralleling the second activation phase of cPRCs. This late INRGW
-response was lacking in \emph{NOS}-mutants \textbf{(Figure 5F,G)}. In
-NIT-GC2 morphants, INRGW cells showed a transient increase in calcium
-that decayed after light off and a delayed activation was not present.
+during stimulation (Figure 5D). In contrast, the INNOS response was flat
+or slightly negative in NIT-GC2 morphant larvae (Figure 5E) revealing an
+essential role for NIT-GC2-mediated cPRC suppression in INNOS
+activation. INRGWa cells were initially inhibited during cPRC
+stimulation, followed by a delayed activation paralleling the second
+activation phase of cPRCs. This late INRGWa response was lacking in
+\emph{NOS}-mutants (Figure 5F). In NIT-GC2 morphants, INRGWa cells
+showed a transient increase in calcium that decayed after light off and
+a delayed activation was not present (Figure 5G).
+
 Next, we imaged calcium signals from Ser-h1 and MC neurons in wild type
@@
 during stimulation followed by rebound, a response that was defective in
-\emph{NOS} mutants. MC cells showed sustained activation, including a
-late-phase that was lacking in \emph{NOS} mutants \textbf{(Figure
-5H,I)}. These data suggest that during 405 stimulation the Ser-h1 cells
+\emph{NOS} mutants (Figure 5I). MC cells showed sustained activation,
+including a late-phase that was lacking in \emph{NOS} mutants (Figure
+5I). These data suggest that during 405 nm stimulation the Ser-h1 cells
 are inhibited and MC cells are activated, and this regulation is
@@
 \includegraphics[width=52.08in]{figures/Fig5} \caption{**Figure 5.
-NO produced by UV stimulation to cPRCs induces a depolarising response in cPRC via NIT-GC1.** **(A, B)** Each left figure shows a calcium imaging using GCaMP6s.
-The respective right figures show immunostaining images of the same positions in the respective left figures, (A) shows RYamide antibodies, (B) shows RGWamide and serotonin (5-HT) antibodies in red and DAPI in cyan.
-Asterisks indicate the nuclei of the cPRC, respectively, and the nuclei of the cells that were identified from the arrangement of the surrounding cells are indicated by numbers.
-From the position of the cells, the calcium imaging data identify INNOS in (A) and INRGW and Ser-h1 in (B), respectively.
-**(C)** Correlation (Pearson's r) map of neuronal activity of the INNOS, INRGW, Ser-h1 and MC neurons.
-**(D)** Changes over time in GCaMP6s intensities in INNOS when cPRC is stimulated with UV in WT (dark purple) and NOSΔ11/Δ11 mutant (light purple) larvae.
-**(E)** Changes over time in GCaMP6s intensities in INNOS when cPRC is stimulated with UV in NIT-GC2 morphants (MO1 and MO2: pink and light pink) larvae.
-**(F)** Changes over time in GCaMP6s intensities in INRGW when cPRC is stimulated with UV in WT (dark blue) and NOSΔ23/Δ23 mutant (light blue) larvae.
-**(G)** Changes over time in GCaMP6s intensities in INRGW when cPRC is stimulated with UV in NIT-GC2 morphants (MO1 and MO2: blue and light blue).
-**(H, I)** Changes over time in GCaMP6s intensities in Ser-h1 and MC when cPRC is stimulated with UV in WT (Ser-h1 and MC: dark red and brown) and NOSΔ11/Δ11 mutant (Ser-h1 and MC: light red and brown) larvae.}\label{fig:unnamed-chunk-5}
+NOS- and NIT-GC2-dependent dynamics of the cPRC circuit.** **(A, B)** GCaMP6s imaging from cPRCs and INNOS cells (left panels) followed by on-slide immunostaining for (A) RYamide to label INNOS and (B) RGWamide+serotonin to label INRGWa and Ser-h1 (red).
+Nuclei are stained with DAPI (cyan).
+Asterisks indicate cPRC nuclei.
+Numbers mark the same cells in the GCaMP and immunostaining images matched by position.
+**(C)** Correlation map of neuronal activity of the cPRCs, INNOS, INRGWa, Ser-h1 and MC neurons.
+**(D)** GCaMP6s fluorescence in INNOS cells in wild type (WT) and NOSΔ11/Δ11 mutant larvae during 405 nm stimulation of the cPRC cilia.
+**(E)** GCaMP6s fluorescence in INNOS cells in NIT-GC2 morphant larvae during 405 nm stimulation.
+**(F)** GCaMP6s fluorescence in INRGWa cells in wild type and NOSΔ23/Δ23 mutant larvae during 405 nm stimulation.
+**(G)** GCaMP6s fluorescence in INRGWa cells in NIT-GC2 morphant larvae during 405 nm stimulation.
+**(H, I)** GCaMP6s fluorescence in (H) Ser-h1 cells and (I) the MC cell in wild type and NOSΔ11/Δ11 mutant larvae during 405 nm stimulation.}\label{fig:unnamed-chunk-5}
 \end{figure}
@@
 
-{[}to be added{]}
+To further analyse the dynamics of responses to UV light and formally
+describe cPRC phototransduction, synaptic connections and NO retrograde
+signalling, we developed a mixed cellular-circuit-level mathematical
+model. We used our Ca2+ imagining data of cPRC, INNOS and INRGWa cells
+collected in wild type, \emph{NOS} knockout, and \emph{NIT-GC2} morphant
+larvae to formulate assumptions that are the basis of the proposed
+model.
+
+We model a direct c-opsin1-dependent response to UV leading to the
+initial rise in Ca2+ levels. The mechanism of this signal is not known
+but may include Gα and Gβ signalling, CNGAα channel activation or other
+pathways. We further assume that in cPRC cells, Ca2+ levels increase
+with cGMP. We model a decrease in cGMP and Ca2+ levels due to a
+NIT-GC2-dependent pathway (Figre 6A). The effect of NO in cPRC cells is
+captured by another term that describes a NIT-GC1 and NO-dependent
+increase in cGMP. Based on the synaptic connectome, we infer a
+feedforward coupling between cPRC cells and INNOS and INRGWa cells. The
+proposed synaptic coupling is UV-dependent, decays linearly and is
+suppressed in NIT-GC2 morphants (Figure 6A).
+
+In the INNOS cells, we assume that excitatory synaptic input from cPRC
+(or rebound from tonic inhibition) leads to a rise in Ca2+ leading to NO
+production via NOS. This input is NIT-GC2-dependent. In the INRGWa
+cells, we model a decrease in Ca2+ based on an inhibitory UV-dependent
+synaptic signal from the cPRCs. We further assume the existence of
+direct excitatory coupling between cPRC Ca2+ levels and INRGWa Ca2+ to
+account for the late effects of cPRC on INRGWa (Figure 6A). The
+UV-dependent and Ca2+-dependent signals may be mediated by different
+transmitters released by cPRCs during distinct phases of their
+activation cycle. In addition, we assume a direct inhibitory coupling
+between INNOS Ca2+ and INRGWa Ca2+. Finally, in all variables, we
+assumed a linear decay and constant production to set a steady state
+(Figure 6A). Since the aim of the model is to capture the normalised
+fluorescence data, the model is nondimensionalised. UV stimulation is
+modelled as a square pulse. To find model parameters producing an output
+fitting the experimental data we employed a global optimisation method
+known as a genetic algorithm (GA).
+
+Through this optimisation procedure, we found parameters that gave a
+very good fit to our experimental data under all conditions. We could
+fit to mean Ca2+ values but also individual calcium transients (Figure
+6B, Figure 6--figure supplement 1). Our model only includes a minimal
+set of parameters and assumptions of interactions that are required to
+describe the dynamics of cPRC, INNOS and INRGWa in wild type and
+loss-of-function conditions (Figure 6A, B). The model highlights that
+cPRC phototransduction employs distinct pathways that operate on
+different time scales and differentially influence cPRC Ca2+ levels. The
+coupling between cPRCs and interneurons also requires different
+signalling mechanisms, either through distinct neurotransmitters or
+receptors expressed in the different cells.
+
+To identify possible molecular pathways, we analysed single-cell
+transcriptome data for cPRC, INNOS and INRGWa identified by spatial
+mapping (Achim et al., 2015) and unique marker genes (Williams et al.,
+2017). In cPRCs, we found transporters and synthetic enzymes indicating
+cholinergic, glutaminergic, glycinergic, GABAergic and adrenergic
+neurotransmission (Figure 6C). For each neurotransmitter, we also found
+receptor-encoding genes expressed in INNOS and INRGWa (Figure 6C). The
+two types of interneurons often expressed different subunits or types of
+these receptors, indicating differential signalling (Figure 6C).
+
+Based on these data, our experimental results and the mathematical
+model, we assembled a minimal phototransduction and circuit diagram. The
+components for which experimental data are available are shown in bold.
+Other potential molecular players are also indicated (Figure 6D).
+
+To generate new insights from the model, we asked how the duration and
+intensity of UV stimulation influences signalling output, measured as
+the magnitude, duration and time to maximum of the NO-dependent Ca2+
+signal in the cPRCs. The model allowed us to test a large number of
+conditions that would not have been feasible in single-larva
+experiments. We found that the onset, magnitude and duration of the
+NO-dependent signal scales with stimulus intensity and duration (Figure
+6--figure supplement 2). This suggests an integratory function of NO
+signalling that tunes circuit output to the strength of UV exposure.
 
@@
 \includegraphics[width=25in]{figures/Fig6} \caption{**Figure 6.
-Gene expressions and signalling mechanisms between cPRC, INNOS and INRGWa.** **(C)** Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq.
+Mathematical modelling of cPRC circuit dynamics during UV stimulation and potential signalling mechanisms.**  **(A)** Diagram for **(B)** **(C)** Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq.
 The size of the dots is expressed in proportion to the percentage of cells expressing that gene relative to all cells.
@@
 
-The activity of synaptic circuits is richly modulated monoamines,
-neuropeptides and gaseous molecules. Our work revealed an essential role
-for NO-mediated signalling in driving UV/violet avoidance behaviour. NO,
-produced by postsynaptic interneurons, feeds back to their presynaptic
-photoreceptor partners leading to their delayed activation. This
-late-phase activation of cPRCs drives circuit output through projection
-interneurons and ciliomotor neurons. In the cPRC circuit, synaptic
-connectivity alone is not sufficient to drive circuit dynamics and
-behavioural change.
-
-We identified cPRC-expressed NIT-GC1 as a new type of NO sensor. NIT-GC1
-is localised to the axonal projections of the cPRCs in the anterior
-neurosecretory plexus of the brain. Here, cPRC axons intermingle and
-synapse on INNOS neurites that contain NOS. NO-mediated feedback
-signalling thus likely occurs in the neurosecretory plexus where NOS-
-and NIT-GC1-containing projections are in close proximity and where we
-could detect NO production following UV stimulation.
-
-We showed that NIT-GC1 can produce cGMP in a NO-dependent manner. In
-cPRCs, increased cGMP levels could open CNGAα, a cyclic nucleotide-gated
-channels specifically expressed in cPRC (Tosches et al., 2014), leading
-to cPRC delayed depolarisation, even after UV off.
-
-NO is a free radical with a half-life of 5 msec in vivo and a signalling
-range limited to a few tens of nanometers. In neuronal signalling, NOS
-and sGC have been found to localise at close proximity at synapses
-{[}Burette et al. (2002){]}(Garthwaite, 2015). We identified 12 sGCs in
-\emph{Platynereis}, but none of these is expressed in the cPRC. Instead,
-we identified an unconventional NIT-domain containing GC, NIT-GC1 as the
-mediator of NO retrograde signalling. The NIT domain was first
-identified in bacteria and animal NIT-GCs have only been recently
-reported (Moroz et al., 2020; Shu et al., 2003). Bacterial NIT domains
-regulate cellular functions in response to changes in extracellular and
-intracellular nitrate and/or nitrite concentrations (Camargo et al.,
-2007). NO is readily converted to nitrate and nitrite (Garthwaite, 2015;
-Möller et al., 2019; Santos et al., 2011) and these molecules accumulate
-in placozoans and cnidarians in cells and tissues with high NOS activity
-(Moroz et al., 2020, 2004). It is possible that the NIT domains in
-NIT-GCs may also sense nitrate and nitrite, as in bacteria. Furthermore,
-if they show different sensitivities to NO, nitrite and nitrate, it is
-possible that differences in their half-lives could give rise to a range
-of activation timings (Lundberg et al., 2011). We have found 15 NIT-GCs
-in \emph{Platynereis}, suggesting a wide range of functions. In
-addition, NIT-GC1 and NIT-GC2 showed very different subcellular
-localisation and functions. It is not clear whether the functional
-differences are due to the different localisation or also to differences
-in biochemical function. Overall, the diversification and spatially
-different localisation patterns of NIT-GCs may increase their signalling
-repertoire. Our data revealed key roles for two \emph{Platynereis}
-NIT-GCs in phototraonsduction and NO-mediated neuromodulation suggesting
-that these molecules contribute to the diversity of neuronal signalling.
-Sponges and ctenophores, where NO signalling is present, lack NIT-GCs.
-In placozoans as many as 12 have been found, and cnidarians also contain
-a large diversity. NIT-GCs have been lost from chordates {[}check{]}.
-
-cPRCs expressing UV-absorbing c-opsin1 have high resting calcium levels
-and show hyperpolarising and subsequent depolarising responses to UV
-light (405 nm) (Verasztó et al., 2018). In the present study, we have
-shown that the cells that strongly synapse with cPRCs are interneurons
-that specifically express NOS (INNOS).
-
-{[}Puopolo et al. (2001){]}(Zhang et al., 2007)(Randel et al., 2014). In
-the mammalian retina, it is known that signals from bipolar
-photoreceptor cells are transmitted to downstream NOS-expressing
-amacrine cells (NOAC), where NO is produced (Jacoby et al., 2018). These
-suggest that UV stimulation of cPRCs causes the production of NO via a
-depolarising response of INNOS.
-
-In the mammalian retina, NO is an essential signalling molecule and one
-of the most well-studied (Cudeiro and Rivadulla, 1999). It has been
-reported that NO produced in one of the retinal cells modulates
-cGMP-mediated neurotransmission by activating presynaptic neural sGCs
-(general target of NO) through retrograde signalling {[}Vielma et al.
-(2014){]}(Wei et al., 2012).
-
-Another type of interneuron that strongly synapses with the cPRC is the
-INRGWa, which synapses directly onto the ciliomotor neuron and is
-therefore considered to be a very important neuron in the control of
-larval behaviour {[}Verasztó et al. (2018){]}(Williams et al., 2017).
-The present results suggest that the depolarising response of the cPRC
-due to NO feedback induces INRGW activation.
-
-In addition, in many marine invertebrates, NO has been reported to be
-involved in many aspects of settlement behaviour and metamorphosis
-induction (Bishop and Brandhorst, 2007, 2003; Castellano et al., 2014;
-Leise et al., 2001; Pechenik et al., 2007; Song et al., 2021; Ueda et
-al., 2016; Ueda and Degnan, 2014, 2013; Yang et al., 2018; Zhu et al.,
-2020). The downward swimming behaviour of Platynereis larvae is the
-first step in the onset of settlement behaviour and metamorphosis. These
-suggest that downward swimming due to NO activation of INRGWs may
-trigger the onset of settlement behaviour and metamorphosis
-{[}Conzelmann et al. (2013b){]}(Williams et al., 2015). These suggest
-that NO feedback signalling may act as a trigger for downward swimming
-via activation of INRGWs, which induces the initiation of settlement
-behaviour and metamorphosis.
+Our work revealed an essential role for NO-mediated signalling in
+driving UV/violet avoidance behaviour in larval \emph{Platynereis}. NO,
+produced by postsynaptic INNOS interneurons, signals retrogradely to
+presynaptic cPRCs via NIT-GC1 leading to delayed and sustained cPRC
+activation. This late-phase activation of the photoreceptors drives
+circuit output through projection interneurons and ciliomotor neurons.
+In the cPRC circuit, synaptic connectivity alone is thus not sufficient
+to account for circuit dynamics and behavioural change, as documented in
+other circuits (Bargmann and Marder, 2013; Imambocus et al., 2022).
+
+\textbf{Localised NO signalling in the neurosecretory plexus}
+
+NO is a free radical with a millisecond-to-second half life and thus a
+limited signalling range. In neuronal signalling, NOS is often localized
+to neurites (Kuntz et al., 2017) at close proximity to sGC at synapses
+(Burette et al., 2002; Garthwaite, 2015). In the cPRC circuit, NOS is
+localised to the dendritic compartment of INNOS cells where also cPRC
+postsynaptic sites occur. NIT-GC1, the target of NO signalling is
+localised to cPRC projections. NO-mediated retrograde signalling thus
+likely occurs in the neurosecretory plexus where NOS- and
+NIT-GC1-containing projections are in close proximity and where we
+detected NO production following UV stimulation. In contrast, INNOS to
+INRGWa synapses occur outside the neurosecretory plexus in the ventral
+brain neuropil.
+
+Compartmentalised signalling also occurs in peptidergic modulatory
+systems and can enable selective network activity during specific
+behaviours. In the UV-avoidance circuit of \emph{Drosophila} larvae, a
+peptidergic hub neuron Dp7 links UV-sensory neurons (v'td2) and motor
+circuits. Dp7 expresses an an insulin-like peptide Ilp7 that is required
+for acute and sustained UV avoidance. Ilp7 signalling occurs at a
+functionally and morphologically distinct dendritic compartment of Dp7,
+segregated from other sensory-motor pathways involving Dp7 (Imambocus et
+al., 2022).
+
+\textbf{Functional diversity of NIT-GCs}
+
+NO signalling is commonly mediated by sGCs. We identified 12 sGCs in
+\emph{Platynereis}, but none of these is expressed in the cPRC based on
+the available scRNAseq data. Instead, we identified an unconventional
+cPRC-expressed NIT-domain containing GC, NIT-GC1 as the mediator of NO
+retrograde signalling. In an in vitro assay, we could show that NIT-GC1
+can produce cGMP following the addition of an NO donor and that this
+activity requires the NIT domain.
+
+NIT domains were first identified in bacteria and animal NIT-GCs have
+only recently been reported (Moroz et al., 2020; Shu et al., 2003).
+Bacterial NIT domains regulate cellular functions in response to changes
+in extracellular and intracellular nitrate and nitrite concentrations
+(Camargo et al., 2007). NO is readily converted to nitrate and nitrite
+(Garthwaite, 2015; Möller et al., 2019; Santos et al., 2011) and these
+molecules accumulate in placozoans and cnidarians in cells and tissues
+with high NOS activity (Moroz et al., 2020, 2004). NIT domains in
+NIT-GCs may also sense nitrate and nitrite, as in bacteria, a
+possibility we cannot rule out based on our cellular assays with
+NIT-GC1. If different NIT-GCs have different sensitivities to NO,
+nitrite and nitrate, then a range of activation timings may be possible
+due to the different half-lives of these molecules (Lundberg et al.,
+2011).
+
+NIT-GC1 and NIT-GC2 showed specific cellular co-expression but very
+different subcellular localisation and function. In \emph{Platynereis},
+we identified 15 NIT-GCs, suggesting a wide range of functions.
+Differences in subcellular localisation and biochemical function thus
+seem to also contribute to the diversity of NIT-GC functions in addition
+to differences in expression.
+
+\textbf{Mechanism of phototransduction and neurotransmission in the cPRC
+circuit}
+
+Based on our data herein and previous work we can now propose a more
+detailed model of cPRC phototransduction and neurotransmission. The
+cPRCs have high basal Ca2+ and respond to UV/violet light dependent on
+c-opsin1 (Verasztó et al., 2018). c-opsin1 forms a bistable photopigment
+and signals through Gi/oα and Gβγ (Tsukamoto et al., 2017; Tsukamoto and
+Kubo, 2023; Veedin Rajan et al., 2021). In heterologous systems, the Gβγ
+subunits released following c-opsin1 activation open GIRK channels
+inducing K+ efflux (hyperpolarisation) (Tsukamoto et al., 2017;
+Tsukamoto and Kubo, 2023) and close voltage-gated Ca2+ channels, thereby
+reducing intracellular calcium levels (Tsukamoto and Kubo, 2023). A GIRK
+channel is also expressed in the cPRCs (Figure 6A). The pathway for the
+first rapid cPRC activation phase following UV/violet stimulus is not
+known but may involve the activation of a CNGα channel expressed in the
+cPRCs (Tosches et al., 2014). For the second inhibitory phase of
+phototransduction, we identified a key requirement for ciliary-localised
+NIT-GC2. The mechanisms may involve c-opsin1-dependent inhibition of
+tonic NIT-GC2 activity and the reduction of ciliary cGMP.
+
+NIT-GC2-dependent signalling is required for the feedforward activation
+of the INNOS cells through an unknown transmitter. The activation of
+INNOS leads to NOS activation and NO release, potentially through a
+canonical Ca2+-calmodulin pathway. Our model predicts feedforward
+inhibiton from INNOS to INRGWa, possibly mediated by glycin
+transmission. NO released by INNOS neurites activates NIT-GC1 in cPRC
+projections that could lead to cGMP production and the opening of CNGAα.
+This late-phase cPRC activation is not directly dependent on the
+UV/violet signal and can happen post-stimulus. The late-phase cPRC
+activation leads to INRGWa activation via an unknown transmitter that is
+likely different from the one acting on INNOS.
+
+In addition, the cPRC circuit expresses several neuropeptides and their
+receptors, suggesting further neuromodulatory signals. For example,
+INRGWa express the proneuropeptide RGWamide and cPRCs and INNOS express
+its receptor, suggesting retrograde peptidergic signalling in the
+circuit.
+
+We have recently shown that the cPRC circuit also mediates responses to
+hydrostatic pressure via the same motor system involving Ser-h1 and the
+prototroch ciliary band. Increased pressure induces cPRC activation and
+a circuit output that is inverted relative to UV-induced activation.
+Consequently, ciliary beating increases and the larvae swim upwards. The
+effect of pressure on cilia requires synaptic transmission by the
+serotonergic Ser-h1 neurons (Bezares-Calderón et al., 2023).
+
+Pressure-induced Ca2+ transients in cPRCs lack an inhibitory phase and
+late activation and resemble UV responses in NIT-GC2 morphants
+(Bezares-Calderón et al., 2023). This indicates that the differentiation
+of sensory cues by the multisensory cPRCs occurs already at the level of
+sensory signal transduction. The different signalling pathways then
+likely result in the differential release of transmitters and modulators
+that are decoded by the postsynaptic interneuron circuit. The complex
+transmitter phenotype of cPRCs could underlie such differential signal
+processing.
+
+\textbf{Nitric oxide confers short-term memory to circuit activity}
+
+Retrograde signalling by NO from INNOS to cPRC leads to the sustained
+activation of cPRCs and postsynaptic neurons even after the end of
+stimulation. This activated state is maintained for several tens of
+second. NO signalling thus induces a transient circuit state or
+short-term memory trace in the \emph{Platynereis} larval brain. Because
+of the short life time of NO, this molecule may be well suited to encode
+transient memory traces (Kuntz et al., 2017).
+
+In the ellipsoid body of the \emph{Drosophila} central brain, NO
+signalling has a similar function. Here, NOS is specifically expressed
+in the R3 ring neurons and is required for the short-term
+(\textasciitilde4 sec) visual memory of objects (Kuntz et al., 2017). NO
+is produced in the axons of R3 neurons and acts directly on sGC in the
+same axons. This autocrine signal leads to a CNG-dependent temporary
+increase in calcium levels, carrying the working memory trace (Kuntz et
+al., 2017).
+
+In the \emph{Platynereis} circuit, our mathematical model indicates that
+the onset, magnitude and duration of the NO-dependent signal depends on
+the intensity and duration of the UV/violet stimulus. This suggests that
+the NO-dependent memory trace also encodes the magnitude and duration of
+the stimulus.
+
+During UV avoidance in the planarian \emph{Schmidtea mediterranea},
+neuropeptide signalling has a similar integratory function. Planarians
+exposed to UV light for \textgreater30 sec remain active for extended
+periods (several minutes) post-stimulation (Bray et al., 2023). If
+neuropeptide signalling is defective, animals can still respond to UV
+light, but do not maintain a latent memory state and do not display
+increased post-stimulus activity.
+
+\textbf{The possible mechanism of UV-induced downward swimming}
+
+Non-directional UV/violet light induces downward swimming head down in
+\emph{Platynereis} larvae (Verasztó et al., 2018). Since stimulus
+direction is not relevant, swimming direction must be determined by
+gravity.
+
+Connectome reconstruction and whole-body cell annotation in the
+three-day-old \emph{Platynereis} larva has not revealed any balancer
+organ to sense orientation in the gravity field (Bezares-Calderón et
+al., 2019; Verasztó et al., 2020). It follows that body orientation must
+be determined by physical parameters including the buoyancy, centre of
+mass and shape of the larva as well as differential ciliary activity.
+Except for ciliary activity, all these parameters are likely invariant
+during the UV response. Our data suggest the UV-dependent inhibition of
+prototroch ciliary beating via Ser-h1 inhibition and MC activation
+(Verasztó et al., 2017).
+
+We hypothesise that the differential beating of prototroch versus trunk
+cilia causes a head-up or head-down orientation due to physics alone.
+During the UV response, prototroch cilia beat slower than trunk cilia,
+resulting in a head-down stable state (`rear-wheel drive'). In contrast,
+during the pressure response prototroch cilia beat faster than trunk
+cilia (Bezares-Calderón et al., 2023), leading to a head-up orientation
+(`front-wheel drive'). Testing this hypothesis will require biophysical
+experiments and mathematical modelling.
+
+\hypertarget{uv-avoidance-circuits-of-extraocular-photoreceptors}{%
+\subsection{UV-avoidance circuits of extraocular
+photoreceptors}\label{uv-avoidance-circuits-of-extraocular-photoreceptors}}
+
+Animals evolved distinct photosensory systems coupled to non-overlapping
+circuits and guiding unique behavioural responses. These sensory systems
+can employ different opsin molecules and be tuned to different
+wavelengths of light. The avoidance of noxious UV/violet light is often
+mediated by extraocular photoreceptors and their circuits, distinct from
+the pigmented visual eyes. These two types of systems co-exist in
+\emph{Platynereis} where the pigmented eyes and eyespots guide
+phototaxis with a maximum sensitivity to cyan light (\textasciitilde500
+nm) (Randel et al., 2014; Verasztó et al., 2018). Planarians also have
+pigmented visual eyes mediating phototaxis to cyan light and peripheral
+extraocular photoreceptors mediating UV avoidance behaviour (Shettigar
+et al., 2021). \emph{Drosophila} larvae have cerebral eyes called
+Bolwig's organs involved in phototaxis (Kane et al., 2013) and several
+types of extraocular UV-sensory cells that tile the body wall and
+mediate UV avoidance (Imambocus et al., 2022).
+
+One common feature of UV-avoidance circuits is their sustained
+post-stimulus activation following UV exposure. This can involve
+peptidergic signals as in planaria and maggots (Bray et al., 2023;
+Imambocus et al., 2022) or NO as in \emph{Platynereis}. Volume
+transmission is well suited to integrate light exposure and maintain an
+internal state following noxious stimulation. The amount of modulator
+released can scale with stimulus intensity or duration and maintain an
+altered circuit state due to the slower decay of the diffusive signals
+relative to synaptic transmission.
+
+We have revealed how NO shapes dynamics in a fully-mapped sensory-motor
+UV avoidance circuit. We identified an unconventional GC as the NO
+receptor in \emph{Platynereis} and measured circuit activity in
+different genetic backgrounds. Finally, we could link circuit activity
+to light-avoidance behaviour. The richly modulated multi-transmitter
+cPRC circuit will serve as a fertile ground for future studies on how
+neuromodulators shape activity and behaviour.
 
@@
 eggs of Platynereis dumerilii wild-type parents according to an
-established injection procedure (Conzelmann et al., 2013a). The eggs
-were kept at 18°C for 45 min before injection and were injected at
-14.5°C. The injected individuals were kept at 18°C for 5 to 8 days in
-6-well- plates (Nunc multidish no. 150239, Thermo Scientific) and then
-cultured at 22°C until sexual maturity. The mature worms were crossed to
+established injection procedure (Conzelmann et al., 2013). The eggs were
+kept at 18°C for 45 min before injection and were injected at 14.5°C.
+The injected individuals were kept at 18°C for 5 to 8 days in 6-well-
+plates (Nunc multidish no. 150239, Thermo Scientific) and then cultured
+at 22°C until sexual maturity. The mature worms were crossed to
 wild-type worms and the progeny was genotyped, resulting in two founder
@@
 and aBayes methods (Minh et al., 2020). The sequences used for the
-phylogenetic analysis are available in \textbf{Supplementary File 2},
-the trimmed alignment is available in \textbf{Supplementary File 3}.
+phylogenetic analysis are available in \textbf{Supplementary File 1},
+the trimmed alignment is available in \textbf{Supplementary File 2} and
+the pre-trimmed data in \textbf{Supplementary file 3}.
 
@@
 1e−15. We ran CD-Hit (Fu et al., 2012) to eliminate redundant sequences
-(at a 80\% threshold). The CLANS analysis is available as
-\textbf{Supplementary File 1}. To identify clusters, we used the
+(at a 80\% threshold). To identify clusters, we used the
 convex-clustering option with 100 jack-knife replicates. The NIT-GCs are
@@
 al., 2020). The sequences used for the phylogenetic analysis are
-available in \textbf{Supplementary File 2}, the trimmed alignment is
-available in \textbf{Supplementary File 3}.
+available in \textbf{Supplementary File 1}, the trimmed alignment is
+available in \textbf{Supplementary File 2} and the pre-trimmed data in
+\textbf{Supplementary file 3}.
 
@@
 We used Achim et al.~for the single-cell data (Achim et al., 2015). In
-fact, in Williams et al., they used 107 cells as neurons by removing
-duplicates from Achim et al.~single-cell data, so we used those cells
-(Williams et al., 2017). Since the raw data were read count data, we
-normalized them to TPM using Python. After that we converted them to
-log10. From the sum of the expression levels in 107 cells for each gene,
-I calculated the percentage in each cell. We chose the appropriate genes
-for now. For each cell, we identified them with marker genes. After
-created the data in Python, plotted it using R dot plots. RPKM
-calculates the total number of reads per million bp and then divides by
-the length of each gene, so it is not possible to compare between
-samples. Instead, TPM first divides by the length of the gene and then
-divides by the total number of reads per million bases, which allows for
-more accurate comparisons between samples. In this case we wanted to
-compare between samples, so we used TPM. In fact, TPM is becoming the
-mainstream method for many single-cell analyses. The total TPM of each
+Williams et al., they used 107 cells as neurons by removing duplicates
+from Achim et al.~single-cell data, so we used those cells (Williams et
+al., 2017). Since the raw data were read count data, we normalized them
+to TPM using Python. After that we converted them to log10. From the sum
+of the expression levels in 107 cells for each gene, We calculated the
+percentage in each cell. For each cell, we identified them with marker
+genes. After created the data in Python, plotted it using R dot plots.
+RPKM calculates the total number of reads per million bp and then
+divides by the length of each gene, so it is not possible to compare
+between samples. Instead, TPM first divides by the length of the gene
+and then divides by the total number of reads per million bases, which
+allows for more accurate comparisons between samples. In this case we
+wanted to compare between samples, so we used TPM. The total TPM of each
 gene between the samples was used to calculate the percentage of
-expressed genes.
+expressed genes. The total TPM values for each gene and the percentage
+of expressed genes are available in \textbf{Supplementary file 1}.
 
@@
 pairs specific to P. dumerilii NOS, NIT-GC1, NIT-GC2, RYa-pNP (GenBank
-accession: JF811330.1), c-opsin1 (GenBank accession: AY692353.1), CNGAα
-(GenBank accession: KM199644.1), and MLD/pedal2-pNP (GenBank accession:
+accession: JF811330.1), and MLD/pedal 2-pNP (GenBank accession:
 KF515945.1). The NOS, NIT-GC1 and NIT-GC2 probes were designed to be
-associated with the b2 initiator sequence, while the RYa-pNP, c-opsin1,
-CNGAα and MLD/pedal2-pNP probes were designed to be associated with the
-b3 initiator sequence. For the detection stage, samples were
-pre-hybridized in 200 µl of probe hybridization buffer for 1 hr at 37°C,
-and then incubated in 250 µl hybridization buffer containing probe
-oligos (4 pmol/ml) overnight at 37°C. To remove excess probe, samples
-were washed 4× with 1 ml hybridization wash buffer for 15 min at 37°C,
-and subsequently 2× in 1 ml 5× SSCT (5× SSC with 0.1\% Tween20) for 5
-min at room temperature. For the amplification stage, samples were
+associated with the b2 initiator sequence, while the RYa-pNP and
+MLD/pedal 2-pNP probes were designed to be associated with the b3
+initiator sequence. For the detection stage, samples were pre-hybridized
+in 200 µl of probe hybridization buffer for 1 hr at 37°C, and then
+incubated in 250 µl hybridization buffer containing probe oligos (4
+pmol/ml) overnight at 37°C. To remove excess probe, samples were washed
+4× with 1 ml hybridization wash buffer for 15 min at 37°C, and
+subsequently 2× in 1 ml 5× SSCT (5× SSC with 0.1\% Tween20) for 5 min at
+room temperature. For the amplification stage, samples were
 pre-incubated with 100 µl of amplification buffer for 30 min, room
@@
 size of 512 x 512 pixels. The larvae were stimulated in a region of
-interest (a circle with ?? pixel diameter) with 405 nm lasers controlled
+interest (a circle with 50 pixel diameter) with 405 nm lasers controlled
 by the Bleaching mode. The imaging laser had a similar intensity than
@@
 microscope (LSM880, Zeiss) equipped with an oil-immersion objective lens
-(UApo/340, 40, NA = xx). Images were acquired using a xenon lamp,
+(UApo/340, 40, NA = 0.17). Images were acquired using a xenon lamp,
 460--495 nm excitation filter, 505-nm dichroic mirror and 510-- 550-nm
-emission filter (????). S-Nitroso-N-acetyl-D, L-penicillamine (SNAP) was
-purchased from Sigma-Aldrich (St.~Louis, MO, USA) . The exposure time of
-the EM-CCD camera was controlled by the ZEN software (Zeiss). Images
-were acquired every 15 s for 10 min and stimulation was initiated 2 min
-after starting image acquisition. Imaging data analysis was performed
-using ImageJ (National Institutes of Health, Bethesda, MD, USA).
+emission filter (Zeiss). S-Nitroso-N-acetyl-D, L-penicillamine (SNAP)
+was purchased from Sigma-Aldrich (St.~Louis, MO, USA) . The exposure
+time of the EM-CCD camera was controlled by the ZEN software (Zeiss).
+Images were acquired every 15 s for 10 min and stimulation was initiated
+2 min after starting image acquisition. Imaging data analysis was
+performed using ImageJ (National Institutes of Health, Bethesda, MD,
+USA).
 
@@
 
-This work was funded by the Wellcome Trust (214337/Z/18/Z).
+This work was funded by the Wellcome Trust (214337/Z/18/Z). This project
+has received funding from the European Research Council (ERC) under the
+European Union's Horizon 2020 research and innovation programme (grant
+agreement No 101020792). KJ has been supported by a JSPS Postdoctoral
+fellowhip, LAYG by a BBSRC Discovery fellowship (BB/W010305/1).
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Aso2019}{}}%
+Aso Y, Ray RP, Long X, Bushey D, Cichewicz K, Ngo T-T, Sharp B,
+Christoforou C, Hu A, Lemire AL, Tillberg P, Hirsh J, Litwin-Kumar A,
+Rubin GM. 2019. Nitric oxide acts as a cotransmitter in a subset of
+dopaminergic neurons to diversify memory dynamics. \emph{eLife}
+\textbf{8}.
+doi:\href{https://doi.org/10.7554/elife.49257}{10.7554/elife.49257}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bargmann2013}{}}%
+Bargmann CI, Marder E. 2013. From the connectome to brain function.
+\emph{Nature Methods} \textbf{10}:483--490.
+doi:\href{https://doi.org/10.1038/nmeth.2451}{10.1038/nmeth.2451}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bezares-calderon2019}{}}%
+Bezares-Calderón LA, Berger J, Jékely G. 2019. Diversity of cilia-based
+mechanosensory systems and their functions in marine animal behaviour.
+\emph{Philosophical Transactions of the Royal Society B: Biological
+Sciences} \textbf{375}:20190376.
+doi:\href{https://doi.org/10.1098/rstb.2019.0376}{10.1098/rstb.2019.0376}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bezares-calderon2023}{}}%
+Bezares-Calderón LA, Shahidi R, Jékely G. 2023.
+\href{http://dx.doi.org/10.1101/2023.02.28.530398}{A ciliary
+photoreceptor-cell circuit mediates pressure response in marine
+zooplankton}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-bishop2007}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-bishop2003}{}}%
-Bishop CD, Brandhorst BP. 2003. On nitric oxide signaling,
-metamorphosis, and the evolution of biphasic life cycles.
-\emph{Evolution and Development} \textbf{5}:542--550.
-doi:\href{https://doi.org/10.1046/j.1525-142x.2003.03059.x}{10.1046/j.1525-142x.2003.03059.x}
+\leavevmode\vadjust pre{\hypertarget{ref-bray2023}{}}%
+Bray SR, Wyss LS, Chai C, Lozada ME, Wang B. 2023.
+\href{http://dx.doi.org/10.1101/2023.01.20.523817}{Adaptive robustness
+through incoherent signaling mechanisms in a regenerative brain}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Bredt1990}{}}%
+Bredt DS, Hwang PM, Snyder SH. 1990. Localization of nitric oxide
+synthase indicating a neural role for nitric oxide. \emph{Nature}
+\textbf{347}:768--770.
+doi:\href{https://doi.org/10.1038/347768a0}{10.1038/347768a0}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-castellano2014}{}}%
-Castellano I, Ercolesi E, Palumbo A. 2014. Nitric Oxide Affects ERK
-Signaling through Down-Regulation of MAP Kinase Phosphatase Levels
-during Larval Development of the Ascidian Ciona intestinalis. \emph{PLoS
-ONE} \textbf{9}:e102907.
-doi:\href{https://doi.org/10.1371/journal.pone.0102907}{10.1371/journal.pone.0102907}
-
 \leavevmode\vadjust pre{\hypertarget{ref-Choi2018}{}}%
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Conzelmann2013a}{}}%
-Conzelmann M, Williams Elizabeth A, Krug K, Franz-Wachtel M, Macek B,
-Jékely G. 2013a. The neuropeptide complement of the marine annelid
-Platynereis dumerilii. \emph{BMC Genomics} \textbf{14}:906.
+Conzelmann M, Williams EA, Krug K, Franz-Wachtel M, Macek B, Jékely G.
+2013. The neuropeptide complement of the marine annelid Platynereis
+dumerilii. \emph{BMC Genomics} \textbf{14}:906.
 doi:\href{https://doi.org/10.1186/1471-2164-14-906}{10.1186/1471-2164-14-906}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Conzelmann2013}{}}%
-Conzelmann M, Williams Elizabeth A, Tunaru S, Randel N, Shahidi R,
-Asadulina A, Berger J, Offermanns S, Jékely G. 2013b. Conserved MIP
-receptor{\textendash}ligand pair regulates {\emph{Platynereis}} larval
-settlement. \emph{Proceedings of the National Academy of Sciences}
-\textbf{110}:8224--8229.
-doi:\href{https://doi.org/10.1073/pnas.1220285110}{10.1073/pnas.1220285110}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Gibbs1998a}{}}%
+Gibbs SM, Truman JW. 1998. Nitric Oxide and Cyclic GMP Regulate Retinal
+Patterning in the Optic Lobe of Drosophila. \emph{Neuron}
+\textbf{20}:83--93.
+doi:\href{https://doi.org/10.1016/s0896-6273(00)80436-5}{10.1016/s0896-6273(00)80436-5}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Hoelscher1997}{}}%
+Hölscher C. 1997. Nitric oxide, the enigmatic neuronal messenger: its
+role in synaptic plasticity. \emph{Trends in Neurosciences}
+\textbf{20}:298--303.
+doi:\href{https://doi.org/10.1016/s0166-2236(97)01065-5}{10.1016/s0166-2236(97)01065-5}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Imambocus2022}{}}%
+Imambocus BN, Zhou F, Formozov A, Wittich A, Tenedini FM, Hu C, Sauter
+K, Macarenhas Varela E, Herédia F, Casimiro AP, Macedo A, Schlegel P,
+Yang C-H, Miguel-Aliaga I, Wiegert JS, Pankratz MJ, Gontijo AM, Cardona
+A, Soba P. 2022. A neuropeptidergic circuit gates selective escape
+behavior of Drosophila larvae. \emph{Current Biology}
+\textbf{32}:149--163.e8.
+doi:\href{https://doi.org/10.1016/j.cub.2021.10.069}{10.1016/j.cub.2021.10.069}
+
 \leavevmode\vadjust pre{\hypertarget{ref-jacoby2018}{}}%
@@
 doi:\href{https://doi.org/10.1016/j.neuron.2018.09.047}{10.1016/j.neuron.2018.09.047}
+
+\leavevmode\vadjust pre{\hypertarget{ref-kane2013}{}}%
+Kane EA, Gershow M, Afonso B, Larderet I, Klein M, Carter AR, Bivort BL
+de, Sprecher SG, Samuel ADT. 2013. Sensorimotor structure of
+{\emph{Drosophila}} larva phototaxis. \emph{Proceedings of the National
+Academy of Sciences} \textbf{110}.
+doi:\href{https://doi.org/10.1073/pnas.1215295110}{10.1073/pnas.1215295110}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-leise2001}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-Kuntz2017}{}}%
+Kuntz S, Poeck B, Strauss R. 2017. Visual Working Memory Requires
+Permissive and Instructive NO/cGMP Signaling at Presynapses in the
+Drosophila Central Brain. \emph{Current Biology} \textbf{27}:613--623.
+doi:\href{https://doi.org/10.1016/j.cub.2016.12.056}{10.1016/j.cub.2016.12.056}
+
+\leavevmode\vadjust pre{\hypertarget{ref-leise2001a}{}}%
 Leise EM, Thavaradhara K, Durham NR, Turner BE. 2001. Serotonin and
@@
 doi:\href{https://doi.org/10.1093/icb/41.2.258}{10.1093/icb/41.2.258}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Locascio2022}{}}%
+Locascio A, Vassalli QA, Castellano I, Palumbo A. 2022. Novel Insights
+on Nitric Oxide Synthase and NO Signaling in Ascidian Metamorphosis.
+\emph{International Journal of Molecular Sciences} \textbf{23}:3505.
+doi:\href{https://doi.org/10.3390/ijms23073505}{10.3390/ijms23073505}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-muxf6ller2019}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-Mobley2022}{}}%
+Mobley RB, Ray EJ, Maruska KP. 2022. Expression and localization of
+neuronal nitric oxide synthase in the brain and sensory tissues of the
+African cichlid fish {\emph{Astatotilapia burtoni}}. \emph{Journal of
+Comparative Neurology} \textbf{530}:2901--2917.
+doi:\href{https://doi.org/10.1002/cne.25383}{10.1002/cne.25383}
+
+\leavevmode\vadjust pre{\hypertarget{ref-moeller2019}{}}%
 Möller MN, Rios N, Trujillo M, Radi R, Denicola A, Alvarez B. 2019.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-pechenik2007}{}}%
-Pechenik JA, Cochrane DE, Li W, West ET, Pires A, Leppo M. 2007. Nitric
-Oxide Inhibits Metamorphosis in Larvae of{\emph{Crepidula fornicata}},
-the Slippershell Snail. \emph{The Biological Bulletin}
-\textbf{213}:160--171.
-doi:\href{https://doi.org/10.2307/25066632}{10.2307/25066632}
-
-\leavevmode\vadjust pre{\hypertarget{ref-puopolo2001}{}}%
-Puopolo M, Hochstetler SE, Gustincich S, Wightman RM, Raviola E. 2001.
-Extrasynaptic Release of Dopamine in a Retinal Neuron. \emph{Neuron}
-\textbf{30}:211--225.
-doi:\href{https://doi.org/10.1016/s0896-6273(01)00274-4}{10.1016/s0896-6273(01)00274-4}
-
 \leavevmode\vadjust pre{\hypertarget{ref-Randel2014a}{}}%
@@
 doi:\href{https://doi.org/10.1089/ars.2010.3297}{10.1089/ars.2010.3297}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Shettigar2021}{}}%
+Shettigar N, Chakravarthy A, Umashankar S, Lakshmanan V, Palakodeti D,
+Gulyani A. 2021. Discovery of a body-wide photosensory array that
+matures in an adult-like animal and mediates
+eye{\textendash}brain-independent movement and arousal.
+\emph{Proceedings of the National Academy of Sciences} \textbf{118}.
+doi:\href{https://doi.org/10.1073/pnas.2021426118}{10.1073/pnas.2021426118}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-thomas2015}{}}%
+Thomas DD. 2015. Breathing new life into nitric oxide signaling: A brief
+overview of the interplay between oxygen and nitric oxide. \emph{Redox
+Biology} \textbf{5}:225--233.
+doi:\href{https://doi.org/10.1016/j.redox.2015.05.002}{10.1016/j.redox.2015.05.002}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Tooker2013}{}}%
+Tooker RE, Lipin MY, Leuranguer V, Rozsa E, Bramley JR, Harding JL,
+Reynolds MM, Vigh J. 2013. Nitric Oxide Mediates Activity-Dependent
+Plasticity of Retinal Bipolar Cell Output via S-Nitrosylation. \emph{The
+Journal of Neuroscience} \textbf{33}:19176--19193.
+doi:\href{https://doi.org/10.1523/jneurosci.2792-13.2013}{10.1523/jneurosci.2792-13.2013}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Tosches2014}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-ueda2014}{}}%
-Ueda N, Degnan SM. 2014. Nitric oxide is not a negative regulator of
-metamorphic induction in the abalone haliotis asinina. \emph{Frontiers
-in Marine Science} \textbf{1}.
-doi:\href{https://doi.org/10.3389/fmars.2014.00021}{10.3389/fmars.2014.00021}
-
-\leavevmode\vadjust pre{\hypertarget{ref-ueda2013}{}}%
-Ueda N, Degnan SM. 2013. Nitric Oxide Acts as a Positive Regulator to
-Induce Metamorphosis of the Ascidian Herdmania momus. \emph{PLoS ONE}
-\textbf{8}:e72797.
-doi:\href{https://doi.org/10.1371/journal.pone.0072797}{10.1371/journal.pone.0072797}
+\leavevmode\vadjust pre{\hypertarget{ref-tsukamoto2023}{}}%
+Tsukamoto H, Kubo Y. 2023. A self-inactivating invertebrate opsin
+optically drives biased signaling toward Gβγ-dependent ion channel
+modulation. \emph{Proceedings of the National Academy of Sciences}
+\textbf{120}.
+doi:\href{https://doi.org/10.1073/pnas.2301269120}{10.1073/pnas.2301269120}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Veraszto2017}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-veraszto2017}{}}%
 Verasztó C, Ueda N, Bezares-Calderón LA, Panzera A, Williams EA, Shahidi
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Vergara2021}{}}%
+Vergara HM, Pape C, Meechan KI, Zinchenko V, Genoud C, Wanner AA, Mutemi
+KN, Titze B, Templin RM, Bertucci PY, Simakov O, Dürichen W, Machado P,
+Savage EL, Schermelleh L, Schwab Y, Friedrich RW, Kreshuk A, Tischer C,
+Arendt D. 2021. Whole-body integration of gene expression and
+single-cell morphology. \emph{Cell} \textbf{184}:4819--4837.e22.
+doi:\href{https://doi.org/10.1016/j.cell.2021.07.017}{10.1016/j.cell.2021.07.017}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Vielma2014}{}}%
@@
 doi:\href{https://doi.org/10.1371/journal.pone.0114330}{10.1371/journal.pone.0114330}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Wang2007}{}}%
+Wang G-Y, List DA van der, Nemargut JP, Coombs JL, Chalupa LM. 2007. The
+sensitivity of light-evoked responses of retinal ganglion cells is
+decreased in nitric oxide synthase gene knockout mice. \emph{Journal of
+Vision} \textbf{7}:7.
+doi:\href{https://doi.org/10.1167/7.14.7}{10.1167/7.14.7}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Williams2015}{}}%
-Williams EA, Conzelmann M, Jékely G. 2015. Myoinhibitory peptide
-regulates feeding in the marine annelid Platynereis. \emph{Frontiers in
-Zoology} \textbf{12}.
-doi:\href{https://doi.org/10.1186/s12983-014-0093-6}{10.1186/s12983-014-0093-6}
+\leavevmode\vadjust pre{\hypertarget{ref-Wildemann1999}{}}%
+Wildemann B, Bicker G. 1999. Developmental expression of nitric
+oxide/cyclic GMP synthesizing cells in the nervous system ofDrosophila
+melanogaster. \emph{Journal of Neurobiology} \textbf{38}:1--15.
+doi:\href{https://doi.org/10.1002/(sici)1097-4695(199901)38:1\%3C1::aid-neu1\%3E3.0.co;2-l}{10.1002/(sici)1097-4695(199901)38:1\textless1::aid-neu1\textgreater3.0.co;2-l}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-yang2018}{}}%
-Yang X-X, Wong YH, Zhang Y, Zhang G, Qian P-Y. 2018. Exploring the
-regulatory role of nitric oxide (NO) and the NO-p38MAPK/cGMP pathway in
-larval settlement of the bryozoan {\emph{Bugula neritina}}.
-\emph{Biofouling} \textbf{34}:545--556.
-doi:\href{https://doi.org/10.1080/08927014.2018.1470240}{10.1080/08927014.2018.1470240}
-
-\leavevmode\vadjust pre{\hypertarget{ref-zhang2007}{}}%
-Zhang D-Q, Zhou T-R, McMahon DG. 2007. Functional Heterogeneity of
-Retinal Dopaminergic Neurons Underlying Their Multiple Roles in Vision.
-\emph{Journal of Neuroscience} \textbf{27}:692--699.
-doi:\href{https://doi.org/10.1523/jneurosci.4478-06.2007}{10.1523/jneurosci.4478-06.2007}
-
-\leavevmode\vadjust pre{\hypertarget{ref-zhu2020}{}}%
-Zhu Y-T, Zhang Y, Liu Y-Z, Li Y-F, Yoshida A, Osatomi K, Yang J-L, Liang
-X. 2020. Nitric oxide negatively regulates larval metamorphosis in
-hard-shelled mussel (mytilus coruscus). \emph{Frontiers in Marine
-Science} \textbf{7}.
-doi:\href{https://doi.org/10.3389/fmars.2020.00356}{10.3389/fmars.2020.00356}
+\leavevmode\vadjust pre{\hypertarget{ref-Zhang2012}{}}%
+Zhang Y, He L-S, Zhang G, Xu Y, Lee O-O, Matsumura K, Qian P-Y. 2012.
+The regulatory role of the NO/cGMP signal transduction cascade during
+larval attachment and metamorphosis of the barnacle{\emph{Balanus
+(=Amphibalanus) amphitrite}}. \emph{Journal of Experimental Biology}.
+doi:\href{https://doi.org/10.1242/jeb.070235}{10.1242/jeb.070235}
 
@@
 \begin{figure}
-\includegraphics[width=32.64in]{figures/Fig1_sup1} \caption{**Figure 1 -- figure supplement 1**  Phylogenetic tree of NOS by maximum likelihood.
+\includegraphics[width=27.78in]{figures/Fig1_sup1} \caption{**Figure 1 -- figure supplement 1**  Phylogenetic tree of NOS by maximum likelihood.
 Tree robustness was tested with 1000 replicates of ultrafast bootstrap with the aLRT-SH-like and aBayes methods.}\label{fig:unnamed-chunk-8}
@@
 \begin{figure}
-\includegraphics[width=34.72in]{figures/Fig1_sup3} \caption{**Figure 1 -- figure supplement 3** }\label{fig:unnamed-chunk-10}
+\includegraphics[width=43.06in]{figures/Fig1_sup3} \caption{**Figure 1 -- figure supplement 3** ssTEM reconstruction of the INNOS (A), INRGWa (B) and cPRC (C), anterior view.
+NS plexus; neurosecretory plexus}\label{fig:unnamed-chunk-10}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=44.44in]{figures/Fig3_sup1} \caption{**Figure 3 -- figure supplement 1** **(A)** Overlaid trajectories for WT (n=37) and NOS mutant (NOSΔ11/Δ11, n=18 and NOSΔ23/Δ23, n=8) at two-day-old larvae.
+\includegraphics[width=44.44in]{figures/Fig3_sup1} \caption{**Figure 3 -- figure supplement 1** **(A)** Top: The domain and Exon/Intron structure of NOS.
+Bottom: Close-up region showing the genomic locus of NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9.
+The generated mutants (NOSΔ11, NOSΔ23) are also shown.
+Pink indicates target sites.
+Gray shows PAM sequences, red shows stop codons.
+**(B)** Overlaid trajectories for WT (n=37) and NOS mutant (NOSΔ11/Δ11, n=18 and NOSΔ23/Δ23, n=8) at two-day-old larvae.
 0 sec as the starting point.
 After 10 sec, UV (395 nm) stimulation from the side.
-**(B)** The temporal changes in the vertical position of the WT and mutant 2-days-old larvae before and after UV stimulation are shown.
+**(C)** The temporal changes in the vertical position of the WT and mutant two-days-old larvae before and after UV stimulation are shown.
 The starting points of each larval trajectory are set to 0.
 After UV stimulation is indicated by purple squares.
-**(C)** Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23) larvae at 2-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
-The data are shown in 30 s bins.
-**(D, E)** The temporal changes in the distance traveled of the WT and mutant in 2-day (D) and 3-day-old (E) larvae before and after UV stimulation are shown.}\label{fig:unnamed-chunk-11}
+**(D, E)** The temporal changes in the distance traveled of the WT and mutant in two-day (D) and three-day-old (E) larvae before and after UV stimulation are shown.
+**(F)** The temporal changes in the distance traveled of larvae treated with NOS inhibitors, L-NAME in three-day larvae before and after UV stimulation are shown.
+**(G)** Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23) larvae at two-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
+The data are shown in 30 s bins.}\label{fig:unnamed-chunk-11}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=33.33in]{figures/Fig4_sup1} \caption{**Figure 4 -- figure supplement 1** }\label{fig:unnamed-chunk-12}
+\includegraphics[width=33.33in]{figures/Fig4_sup1} \caption{**Figure 4 -- figure supplement 1** Cluster analysis of guanylate and adenylate cyclases.
+Connections are based on blast similarities < 1e-16 as shown on the upper right.
+Animal groups are colour coded as shown on the upper left.
+NIT-GCs, NIT domain containing guanylate cyclases; membrane-bound GCs, Membrane-bound guanylyl cyclases; sGCs, soluble guanylate cyclases; ACs, adenylate cyclases.}\label{fig:unnamed-chunk-12}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=33.33in]{figures/Fig4_sup3} \caption{**Figure 4 -- figure supplement 3** Co-expression analysis image of the NIT-GC1 (magenta) and MLD-pedal2 amide proneuropeptide gene (MLD: green).
+\includegraphics[width=33.33in]{figures/Fig4_sup3} \caption{**Figure 4 -- figure supplement 3** **(A)** Co-expression analysis image of the NIT-GC1 (magenta) and MLD-pedal2 amide proneuropeptide gene (MLD: green).
 Anterior view of the larva at two-day-old.
@@
 \begin{figure}
-\includegraphics[width=27.78in]{figures/Fig5_sup1} \caption{**Figure 5 -- figure supplement 1** Schematic diagram of the immunostaining procedure after calcium imaging.}\label{fig:unnamed-chunk-15}
+\includegraphics[width=27.78in]{figures/Fig5_sup1} \caption{**Figure 5 -- figure supplement 1** **(A)** Schematic diagram of the immunostaining procedure after calcium imaging.**(B-D)** Co-expression analysis image of the NOS (magenta) and RY amide proneuropeptide gene (RYa: green).
+Anterior view of the larva at two-day-old.
+}\label{fig:unnamed-chunk-15}
 \end{figure}

### 2023-06-08 Manuscript/Manuscript.Rmd
Upadate Supplement files
Ms: methods edits
Ms: methods edit

@@
 We assayed larval photoresponses in a vertical Plexiglas column (31 mm x 10 mm x 160 mm water height).
-For stimulation, we illuminated the column from the top with a monochromator (Polychrome II, Till Photonics) controlled by AxioVision 4.8.2.0 (Carl Zeiss MicroImaging GmbH) via analog voltage.
+For stimulation, we illuminated the column from the top with a monochromator (Polychrome II, Till Photonics) controlled by AxioVision 4.8.2.0 (Carl Zeiss GmbH, Jena) via analog voltage.
 The light passed a collimator lens (LAG-65.0-53.0-C with MgF2 Coating, CVI Melles Griot) before entering the column.
@@
 
-The synthetic peptides contained an N-terminal Cys that was used for coupling during purification.
-Antibodies were affinity purified from sera as previously described [@Conzelmann2012].
+Antibodies were raised in rats or rabbits against synthetic peptides containing an N-term Cys residue (Altabioscience). The same Cys-containing synthetic peptides were used for affinity purification.
+Antibodies were affinity purified from sera on a SulfoLink Coupling Resin (Thermo Fischer) as previously described [@Conzelmann2012].
 
@@
 
-The *NOS* promoter (fragment sizes: 12 Kb) was amplified and cloned upstream of 3xHA- Palmi-tdTomato.
-Larvae injected with promoter constructs (ca.
-250 ng/ml) were analysed for reporter expression at 3 days post fertilization using an AxioImager Z.1 fluorescence wide-field microscope (Carl Zeiss GmbH, Jena) and immediately fixed for immunostainings.
-The protocol followed for immunostaining of HA-tagged reporters was recently described [@veraszto2017].
-Specimens were imaged with a LSM 780 NLO or LSM 880 with Airysan Confocal Microscope (Zeiss, Jena).
+A 12 kb genomic region upstream of the start site of the *NOS* gene was amplified and cloned upstream of 3xHA-Palmi-tdTomato to yield the plasmid *NOS-3xHA-Palmi-tdTomato*.
+Larvae injected with the promoter construct (ca.
+250 ng/ml) were analysed for reporter expression at three days post fertilization on an AxioImager Z.1 fluorescence microscope (Carl Zeiss GmbH, Jena) and if positive fixed for immunostaining.
+The immunostaining for the HA-tagged reporter was done as described [@veraszto2017].
+Stained specimens were imaged on an LSM 780 NLO or LSM 880 Airysan Confocal Microscope (Carl Zeiss GmbH, Jena).
 
@@
 
-For calcium imaging, 49-55 hpf larvae were used.
-Experiments were carried out at room temperature and larvae were immobilised by being embedded in 2.5% agarose filtered artificial seawater between a slide and coverslip spaced with adhesive tape.
+For calcium imaging, 49-55 hours post fertilisation larvae were used.
+Experiments were carried out at room temperature and larvae were immobilised in 2.5% agarose in artificial sea water and mounted between a slide and coverslip spaced with adhesive tape.
 GCaMP6s mRNA (1 mg/ml) was injected into zygotes as described previously [@Randel2014a].
 Larvae were imaged on a Zeiss LSM 880 with Airyscan (with a C-Apochromat 63X/1.2 Corr - water) with a frame rate of 1.88 frame/sec and an image size of 512 x 512 pixels.
-The larvae were stimulated in a region of interest (a circle with 50 pixel diameter) with 405 nm lasers controlled by the Bleaching mode.
+The larvae were stimulated in a region of interest (a circle with 50 pixel diameter) with a 405 nm laser controlled by the Bleaching mode of the Zeiss software.
 The imaging laser had a similar intensity than the stimulus laser but covered an area that was 10 times larger than the stimulus ROI.
 
-**Cell culture experiment**
+**cGMP production assay**
 
-Green cGull was used for the cGMP assay [@Matsuda2017].
-A full-length Pdum-NIT-GC1 and -NIT-GC2 coding sequences were amplified by PCR starting from a *Platynereis dumerilii* cDNA library and cloned into the pcDNA3.1(+) vector using the T2A self-cleaving sequence.
-COS-7 cells with low expression of endogenous soluble guanylate cyclase were used as cultured cells for gene expression.
-This cell line was purchased from Angio-proteomie (CAT no. cAP-0203).
-The COS-7 cells were maintained at 37 °C in 35mm dishes (Nunc™ Glass Bottom Dishes) containing 3 mL of DMEM, high glucose glutamax medium (Thermo; Cat.
+To measure cGMP production in a cell culture assay we used the Green cGull reporter [@Matsuda2017].
+Full-length *NIT-GC1* and *NIT-GC2* coding sequences were amplified by PCR from a *Platynereis dumerilii* cDNA library and cloned into the pcDNA3.1(+) vector with the T2A self-cleaving sequence.
+For the assay, COS-7 cells (Angio-proteomie, CAT no. cAP-0203) with a low endogenous level of soluble guanylate cyclase activity were used.
+Cells were maintained at 37 °C in 35 mm dishes (Nunc™ Glass Bottom Dishes) containing 3 ml of DMEM, high glucose glutamax medium (Thermo; Cat.
 No.
@@
 10082147).
-Upon reaching confluency of approximately 85%, we transfected the cells with the plasmid containing Green cGull-T2A-NITGC1.
-Transfections were carried out with 150 ng of each plasmid and 0.3 μl of the transfection Lipofectamine 3000 Reagent (invitrogen; Cat. No. L3000001).
-Two days post-transfection, we removed the culture medium and substituted it for fresh DMEM-medium.
-For single-wavelength imaging experiments, cells in 35-mm dishes were washed twice and imaged in modified Ringer's buffer (140 mM NaCl, 3.5 mM KCl, 0.5mM NaH2PO4 , 0.5mM S-3 MgSO4 , 1.5 mM CaCl2 , 10 mM HEPES, 2 mM NaHCO3 and 5 mM glucose).
-Dishes were mounted on a stage heated at 37 °C and imaging was performed using an inverted microscope (LSM880, Zeiss) equipped with an oil-immersion objective lens (UApo/340, 40, NA = 0.17).
-Images were acquired using a xenon lamp, 460--495 nm excitation filter, 505-nm dichroic mirror and 510-- 550-nm emission filter (Zeiss).
-S-Nitroso-N-acetyl-D, L-penicillamine (SNAP) was purchased from Sigma-Aldrich (St. Louis, MO, USA) .
-The exposure time of the EM-CCD camera was controlled by the ZEN software (Zeiss).
+Upon reaching approximately 85% confluency, cells were transfected with the pcDNA3.1(+)-Green-cGull-T2A-NITGC1 plasmid.
+Transfections were carried out with 150 ng of each plasmid and 0.3 μl of Lipofectamine 3000 Reagent (invitrogen; Cat. No. L3000001).
+Two days post-transfection, the culture medium was substituted for fresh DMEM-medium. 
+For single-wavelength imaging experiments, cells in 35-mm dishes were washed twice and imaged in modified Ringer's buffer (140 mM NaCl, 3.5 mM KCl, 0.5 mM NaH~2~PO~4~ , 0.5 mM S-3 MgSO~4~, 1.5 mM CaCl~2~ , 10 mM HEPES, 2 mM NaHCO~3~ and 5 mM glucose).
+Dishes were mounted on a stage heated to 37 °C and imaging was done on an inverted confocal microscope (LSM880, Carl Zeiss GmbH, Jena) equipped with an oil-immersion objective lens (UApo/340, 40x, NA = 0.17).
+Images were acquired using a xenon lamp, a 460-495 nm excitation filter, a 505-nm dichroic mirror and a 510-550-nm emission filter (Carl Zeiss GmbH, Jena).
+The exposure time of the EM-CCD camera was controlled by the ZEN software (Carl Zeiss GmbH, Jena).
 Images were acquired every 15 s for 10 min and stimulation was initiated 2 min after starting image acquisition.
-Imaging data analysis was performed using ImageJ (National Institutes of Health, Bethesda, MD, USA).
+Imaging data analysis was done with ImageJ (National Institutes of Health, Bethesda, MD, USA).
+As NO donor, we used S-Nitroso-N-acetyl-D, L-penicillamine (SNAP, Sigma-Aldrich, St. Louis, MO, USA).
+
+## Data analysis and figure preparation
+
+Data were analysed in ImageJ and R.
+All figures were assembled in R with the cowplot and patchwork packages.
+All scripts are available at <https://github.com/JekelyLab/Joukra_et_al_NOS> (commit ).
 

### 2023-06-09 Manuscript/Manuscript.Rmd
Ms: methods, renamed suppl file
Update methids

@@
 Individual cells were identified by suits of well-characterised marker genes and spatial mapping [@Achim2015].
-The total TPM values for each gene and the percentage of expressed genes are available in **Supplementary file 1**.
+The total TPM values for each gene and the percentage of expressed genes are available in **Supplementary file 7**.
 
@@
 Dishes were mounted on a stage heated to 37 °C and imaging was done on an inverted confocal microscope (LSM880, Carl Zeiss GmbH, Jena) equipped with an oil-immersion objective lens (UApo/340, 40x, NA = 0.17).
-Images were acquired using a xenon lamp, a 460-495 nm excitation filter, a 505-nm dichroic mirror and a 510-550-nm emission filter (Carl Zeiss GmbH, Jena).
 The exposure time of the EM-CCD camera was controlled by the ZEN software (Carl Zeiss GmbH, Jena).

### 2023-06-13 Manuscript/Manuscript.Rmd
MS: added sulfolink protocol and link to protocol on bio-protocols
MS: modelling part edits

@@
 This free radical synthesized from L-arginine by nitric oxide synthase (NOS) is short-lived and can diffuse across biological membranes [@Cudeiro1999; @thomas2015].
-Canonical NO signalling involves the Ca2+/calmodulin-dependent activation of NOS, NO production and diffusion, and the NO-dependent activation of soluble guanylate cyclases (sGC).
+Canonical NO signalling involves the Ca^2+^/calmodulin-dependent activation of NOS, NO production and diffusion, and the NO-dependent activation of soluble guanylate cyclases (sGC).
 This can occur cell-autonomously or in other cells leading to cGMP production [@Bredt1990; @Hoelscher1997].
@@
 To further analyse the dynamics of responses to UV light and formally describe cPRC phototransduction, synaptic connections and NO retrograde signalling, we developed a mixed cellular-circuit-level mathematical model.
-We used our Ca2+ imagining data of cPRC, INNOS and INRGWa cells collected in wild type, *NOS* knockout, and *NIT-GC2* morphant larvae to formulate assumptions that are the basis of the proposed model.
+We used our Ca^2+^ imagining data of cPRC, INNOS and INRGWa cells collected in wild type, *NOS* knockout, and *NIT-GC2* morphant larvae to formulate assumptions that are the basis of the proposed model.
 
-We model a direct c-opsin1-dependent response to UV leading to the initial rise in Ca2+ levels.
+We model a direct c-opsin1-dependent response to UV leading to the initial rise in Ca^2+^ levels.
 The mechanism of this signal is not known but may include Gα and Gβ signalling, CNGAα channel activation or other pathways.
-We further assume that in cPRC cells, Ca2+ levels increase with cGMP.
-We model a decrease in cGMP and Ca2+ levels due to a NIT-GC2-dependent pathway (Figre 6A).
+We further assume that in cPRC cells, Ca^2+^ levels increase with cGMP.
+We model a decrease in cGMP and Ca^2+^ levels due to a NIT-GC2-dependent pathway (Figre 6A).
 The effect of NO in cPRC cells is captured by another term that describes a NIT-GC1 and NO-dependent increase in cGMP.
@@
 
-In the INNOS cells, we assume that excitatory synaptic input from cPRC (or rebound from tonic inhibition) leads to a rise in Ca2+ leading to NO production via NOS.
+In the INNOS cells, we assume that excitatory synaptic input from cPRC (or rebound from tonic inhibition) leads to a rise in Ca^2+^ leading to NO production via NOS.
 This input is NIT-GC2-dependent.
-In the INRGWa cells, we model a decrease in Ca2+ based on an inhibitory UV-dependent synaptic signal from the cPRCs.
-We further assume the existence of direct excitatory coupling between cPRC Ca2+ levels and INRGWa Ca2+ to account for the late effects of cPRC on INRGWa (Figure 6A).
-The UV-dependent and Ca2+-dependent signals may be mediated by different transmitters released by cPRCs during distinct phases of their activation cycle.
-In addition, we assume a direct inhibitory coupling between INNOS Ca2+ and INRGWa Ca2+.
+In the INRGWa cells, we model a decrease in Ca^2+^ based on an inhibitory UV-dependent synaptic signal from the cPRCs.
+We further assume the existence of direct excitatory coupling between cPRC Ca^2+^ levels and INRGWa Ca^2+^ to account for the late effects of cPRC on INRGWa (Figure 6A).
+The UV-dependent and Ca^2+^-dependent signals may be mediated by different transmitters released by cPRCs during distinct phases of their activation cycle.
+In addition, we assume a direct inhibitory coupling between INNOS Ca^2+^ and INRGWa Ca^2+^.
 Finally, in all variables, we assumed a linear decay and constant production to set a steady state (Figure 6A).
@@
 Through this optimisation procedure, we found parameters that gave a very good fit to our experimental data under all conditions.
-We could fit to mean Ca2+ values but also individual calcium transients (Figure 6B, Figure 6--figure supplement 1).
+We could fit to mean Ca^2+^ values but also individual calcium transients (Figure 6B, Figure 6--figure supplement 1).
 Our model only includes a minimal set of parameters and assumptions of interactions that are required to describe the dynamics of cPRC, INNOS and INRGWa in wild type and loss-of-function conditions (Figure 6A, B).
-The model highlights that cPRC phototransduction employs distinct pathways that operate on different time scales and differentially influence cPRC Ca2+ levels.
+The model highlights that cPRC phototransduction employs distinct pathways that operate on different time scales and differentially influence cPRC Ca^2+^ levels.
 The coupling between cPRCs and interneurons also requires different signalling mechanisms, either through distinct neurotransmitters or receptors expressed in the different cells.
@@
 
-To generate new insights from the model, we asked how the duration and intensity of UV stimulation influences signalling output, measured as the magnitude, duration and time to maximum of the NO-dependent Ca2+ signal in the cPRCs.
+To generate new insights from the model, we asked how the duration and intensity of UV stimulation influences signalling output, measured as the magnitude, duration and time to maximum of the NO-dependent Ca^2+^ signal in the cPRCs.
 The model allowed us to test a large number of conditions that would not have been feasible in single-larva experiments.
@@
 Based on our data herein and previous work we can now propose a more detailed model of cPRC phototransduction and neurotransmission.
-The cPRCs have high basal Ca2+ and respond to UV/violet light dependent on c-opsin1 [@Veraszto2018].
+The cPRCs have high basal Ca^2+^ and respond to UV/violet light dependent on c-opsin1 [@Veraszto2018].
 c-opsin1 forms a bistable photopigment and signals through Gi/oα and Gβγ [@tsukamoto2017; @Veedin2021; @tsukamoto2023].
-In heterologous systems, the Gβγ subunits released following c-opsin1 activation open GIRK channels inducing K+ efflux (hyperpolarisation) [@tsukamoto2017; @tsukamoto2023] and close voltage-gated Ca2+ channels, thereby reducing intracellular calcium levels [@tsukamoto2023].
+In heterologous systems, the Gβγ subunits released following c-opsin1 activation open GIRK channels inducing K+ efflux (hyperpolarisation) [@tsukamoto2017; @tsukamoto2023] and close voltage-gated Ca^2+^ channels, thereby reducing intracellular calcium levels [@tsukamoto2023].
 A GIRK channel is also expressed in the cPRCs (Figure 6A).
@@
 NIT-GC2-dependent signalling is required for the feedforward activation of the INNOS cells through an unknown transmitter.
-The activation of INNOS leads to NOS activation and NO release, potentially through a canonical Ca2+-calmodulin pathway.
+The activation of INNOS leads to NOS activation and NO release, potentially through a canonical Ca^2+^-calmodulin pathway.
 Our model predicts feedforward inhibiton from INNOS to INRGWa, possibly mediated by glycin transmission.
@@
 
-Pressure-induced Ca2+ transients in cPRCs lack an inhibitory phase and late activation and resemble UV responses in NIT-GC2 morphants [@bezares-calderon2023].
+Pressure-induced Ca^2+^ transients in cPRCs lack an inhibitory phase and late activation and resemble UV responses in NIT-GC2 morphants [@bezares-calderon2023].
 This indicates that the differentiation of sensory cues by the multisensory cPRCs occurs already at the level of sensory signal transduction.
@@
 library(readxl)
-key_resource_tale <- read_excel("key_resource_tale/key_resource_tale.xls")
-df_key_resource_tale <- as.data.frame(key_resource_tale)
-knitr::kable(df_key_resource_tale, format = "html", booktabs = TRUE)
+key_resource_table <- read_excel("key_resource_table/key_resource_tale.xls")
+df_key_resource_table <- as.data.frame(key_resource_table)
+knitr::kable(df_key_resource_table, format = "html", booktabs = TRUE)
 ```
@@
 
-Antibodies were raised in rats or rabbits against synthetic peptides containing an N-term Cys residue (Altabioscience). The same Cys-containing synthetic peptides were used for affinity purification.
+Antibodies were raised in rats or rabbits against synthetic peptides containing an N-term Cys residue (Altabioscience).
+The same Cys-containing synthetic peptides were used for affinity purification.
 Antibodies were affinity purified from sera on a SulfoLink Coupling Resin (Thermo Fischer) as previously described [@Conzelmann2012].
+View a detailed protocol here: <https://bio-protocol.org/exchange/preprintdetail?id=2333&type=3>
 
@@
 
+## Mathematical modelling
+
+The mathematical model for the wild type data comprises six dynamical equations:
+
+\begin{align*}
+\frac{d C_P(t)}{dt} &= 1 - \delta^{C_P} C_P(t) + K_{UV}^{C_P} UV(t) C_O(t)  + K_{G}^{C_P} G(t) , \\
+\frac{dC_O(t)}{dt} &= 1 - \delta^{C_O} C_O(t) - K_{UV}^{C_O} UV(t) C_O(t), \\
+\frac{dG(t)}{dt} &= 1 - \delta^{G} G(t) + K_{GC2}^G \left(1 - UV(t)\right) G(t) + K_{GC1}^G N(t) , \\
+\frac{dS(t)}{dt} &= UV(t) - \frac{S(t)}{\tau_S}, \\
+\frac{dC_N(t)}{dt} &= 1 - \delta^{C_N} C_N(t) + K_{S}^{C_N} S(t), \\
+\frac{dN(t)}{dt} &= 1 - \delta^{N} N(t) + K_{C_N}^{N} C_N(t), \\
+\frac{dC_R(t)}{dt} &= 1 - \delta^{C_R} C_R(t) - K_{S}^{C_R} S(t) + K_{C_P}^{C_R} C_P(t) - K_{C_N}^{C_R} C_N(t) C_R(t), \\
+UV(t) &=
+    \begin{cases}
+        a, & t_{UV\: start} \leq t \leq t_{UV\: end}\\
+        0, & \text{otherwise}
+    \end{cases}   .
+\end{align*}
+
 ## Data analysis and figure preparation
@@
 This project has received funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement No 101020792).
-KJ has been supported by a JSPS Postdoctoral fellowhip, LAYG by a BBSRC Discovery fellowship (BB/W010305/1).
+KJ has been supported by a JSPS Postdoctoral fellowhip, LAYG by a BBSRC Discovery fellowship (BB/W010305/1), PS by a Wellcome Trust Institutional Strategic Support Award (204909/Z/16/Z).
 

### 2023-06-13 Manuscript/Manuscript.Rmd
Ms: model section edit

@@
 
-To generate new insights from the model, we asked how the duration and intensity of UV stimulation influences signalling output, measured as the magnitude, duration and time to maximum of the NO-dependent Ca^2+^ signal in the cPRCs.
-The model allowed us to test a large number of conditions that would not have been feasible in single-larva experiments.
-We found that the onset, magnitude and duration of the NO-dependent signal scales with stimulus intensity and duration (Figure 6--figure supplement 2).
+To generate new insights from the model, we asked how the duration and intensity of UV stimulation influences signalling output, measured as the magnitude (quantified as area under peak, AUP) of the NO-dependent Ca^2+^ signal in the cPRCs.
+The model allowed us to systematically vary stimulus conditions that would not have been feasible in single-larva experiments.
+We found that the magnitude of the NO-dependent signal scales with stimulus intensity and duration (Figure 7).
 This suggests an integratory function of NO signalling that tunes circuit output to the strength of UV exposure.

### 2023-06-14 Manuscript/Manuscript.Rmd

@@
 The cPRCs express a ciliary-type opsin, c-opsin1 [@Arendt2004] that forms a UV-absorbing bistable photopigment with an absorption maximum around 384 nm [@tsukamoto2017; @Veraszto2018; @Veedin2021].
-Upon UV/violet exposure, the cPRC show a characteristic biphasic calcium response that is *c-opsin1* dependent.
+Upon UV/violet exposure, the cPRCs show a characteristic biphasic calcium response that is *c-opsin1* dependent.
 UV/violet avoidance is also *c-opsin1*-dependent and is defective in *c-opsin1* mutants [@Veraszto2018].
@@
 This reporter also revealed the axonal projections of these central NOS-expressing neurons.
-The position and morphology of the four *NOS+* cells allowed us to identify the same four cells as four interneurons (INNOS) in our three-day-old whole-body *Platynereis* volume EM data [@williams2017; @veraszto2020] (Figure 1B,C).
+The position and morphology of the four *NOS+* cells allowed us to identify the same four cells as four interneurons (INNOS) in our three-day-old whole-body *Platynereis* volume EM data [@williams2017; @veraszto2020] (Figure 1B, C).
 In the synaptic connectome, the INNOS cells are postsynaptic to the UV-sensory cPRCs and presynaptic to the INRGWa interneurons, which are also cPRC targets (Figure 1C,F).
@@
 INNOS output synapses form in the more ventral projection region (Figure 1---figure supplement 3).
-INRGWa neurons synapse on the head serotonergic ciliomotor neurons (Ser-h1), which synapse on the prototroch ciliary band and the cholinergic MC ciliomotor neurons (Figure 1C,F) [@veraszto2017].
+INRGWa neurons synapse on the head serotonergic ciliomotor neurons (Ser-h1), which synapse on the cholinergic MC ciliomotor neuron and cells of the prototroch ciliary band (Figure 1C,F) [@veraszto2017].
 
@@
 As previously shown, wild-type larvae swim downward following non-directional UV/violet light stimulation [@Veraszto2018].
-In contrast, both two- and three-day-old homozygous *NOS*-mutant larvae showed a strongly diminished UV-avoidance response (Figure 3A, B and Figure 3---figure supplement 1B,C).
+In contrast, both two- and three-day-old homozygous *NOS*-mutant larvae showed a strongly diminished UV-avoidance response (Figure 3A, B and Figure 3---figure supplement 1B, C).
 This phenotype is similar to the defective UV-avoidance of *c-opsin1* mutant larvae [@Veraszto2018] and reveals a requirement for NOS in UV-avoidance behaviour.
@@
 
-To distinguish between an acute and developmental function of NOS in light responses, we next tested larvae exposed to the NOS inhibitor L-NAME.
+To distinguish between an acute and developmental function of NOS in light responses, we also tested larvae exposed to the NOS inhibitor L-NAME.
 Larvae incubated for 5 min in 0.1 mM or 1 mM L-NAME showed a dose-dependent inhibition of UV avoidance.
@@
 As we have shown previously, a 20-sec local stimulation of cPRC cilia led to a transient increase in cPRC calcium levels, followed by a transient decrease [@Veraszto2018].
-After \~20-sec, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus -- a response that may involve depolarisation (Figure 4A, B).
+After \~20 sec, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus -- a response that may involve depolarisation (Figure 4A, B).
 This activation phase occurs after the 20 sec stimulation period and is likely due to a delayed neuroendocrine feedback [@Veraszto2018].
@@
 NO binding to the heme group of sGC leads to increased cyclic guanosine monophosphate (cGMP) production.
-Analysis of sGCs in *Platynereis* indicated that these genes are not expressed in any of the cells of the cPRC circuit [@veraszto2017].
+Analysis of sGCs in *Platynereis* indicated that these genes are not expressed in the cPRCs [@veraszto2017] (not shown).
 Recently, Moroz and coworkers reported an atypical but widely conserved family of guanylyl cyclases with a NIT (nitrite/nitrate sensing) domain (PF08376) (NIT-GC) as potential mediators of NO signalling [@Moroz2020].
@@
 We used the cell-specific antibody markers against RYamide (INNOS) (Figure 5---figure supplement 1B-D), RGWamide (INRGWa) and serotonin (Ser-h1) [@Conzelmann2011] to immunostain agar-embedded larvae following calcium imaging.
-Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision (Figure 5A,B).
+Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision (Figure 5A, B).
 Due to the stereotypy of the larvae, we could also identify neurons based on their position and calcium activity in activity-correlation maps (Figure 5C).
@@
 These data suggest that during 405 nm stimulation the Ser-h1 cells are inhibited and MC cells are activated, and this regulation is NO-dependent.
-This pattern is expected to inhibit ciliary activity in the prototroch but not in the other ciliary bands, triggering NO-dependent downward swimming.
+This pattern is expected to inhibit ciliary activity in the prototroch but not in the other ciliary bands (with no Ser-h1 or MC synapses), triggering NO-dependent downward swimming.
 
@@
 UV stimulation is modelled as a square pulse.
+
 To find model parameters producing an output fitting the experimental data we employed a global optimisation method known as a genetic algorithm (GA).
-
 Through this optimisation procedure, we found parameters that gave a very good fit to our experimental data under all conditions.
@@
 
-To generate new insights from the model, we asked how the duration and intensity of UV stimulation influences signalling output, measured as the magnitude (quantified as area under peak, AUP) of the NO-dependent Ca^2+^ signal in the cPRCs.
-The model allowed us to systematically vary stimulus conditions that would not have been feasible in single-larva experiments.
-We found that the magnitude of the NO-dependent signal scales with stimulus intensity and duration (Figure 7).
-This suggests an integratory function of NO signalling that tunes circuit output to the strength of UV exposure.
-
 ```{r echo=FALSE,fig.cap="**Figure 6.
@@
 **(D)** Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback."}
+knitr::include_graphics(path = "figures/Fig6.png")
+```
 
-knitr::include_graphics(path = "figures/Fig6.png")
-
-```
+To generate new insights from the model, we asked how the duration and intensity of UV stimulation influences signalling output, measured as the magnitude (quantified as area under peak, AUP) of the NO-dependent Ca^2+^ signal in the cPRCs.
+The model allowed us to systematically vary stimulus conditions that would not have been feasible in single-larva experiments.
+We found that the magnitude of the NO-dependent signal scales with stimulus intensity and duration (Figure 7).
+This suggests an integratory function of NO signalling that tunes circuit output to the strength of UV exposure.
 
@@
 In the UV-avoidance circuit of *Drosophila* larvae, a peptidergic hub neuron Dp7 links UV-sensory neurons (v'td2) and motor circuits.
-Dp7 expresses an an insulin-like peptide Ilp7 that is required for acute and sustained UV avoidance.
+Dp7 expresses an insulin-like peptide Ilp7 that is required for acute and sustained UV avoidance.
 Ilp7 signalling occurs at a functionally and morphologically distinct dendritic compartment of Dp7, segregated from other sensory-motor pathways involving Dp7 [@Imambocus2022].
@@
 
-In the *Platynereis* circuit, our mathematical model indicates that the onset, magnitude and duration of the NO-dependent signal depends on the intensity and duration of the UV/violet stimulus.
-This suggests that the NO-dependent memory trace also encodes the magnitude and duration of the stimulus.
+In the *Platynereis* circuit, our mathematical model indicates that the magnitude of the NO-dependent signal depends on the intensity and duration of the UV/violet stimulus.
+This suggests that the NO-dependent memory trace also encodes the intensity and duration of the stimulus.
 
@@
 library(readxl)
-key_resource_table <- read_excel("key_resource_table/key_resource_tale.xls")
+key_resource_table <- read_excel("key_resource_table/key_resource_table.xls")
 df_key_resource_table <- as.data.frame(key_resource_table)
@@
 This project has received funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement No 101020792).
-KJ has been supported by a JSPS Postdoctoral fellowhip, LAYG by a BBSRC Discovery fellowship (BB/W010305/1), PS by a Wellcome Trust Institutional Strategic Support Award (204909/Z/16/Z).
+KJ has been supported by a JSPS Postdoctoral fellowhip, LAYG by a BBSRC Discovery fellowship (BB/W010305/1), PS by a Wellcome Trust Institutional Strategic Support Award (204909/Z/16/Z), and KCAW by the EPSRC Hub for Quantitive Modelling in Healthcare (EP/T017856/1).
 

### 2023-06-14 Manuscript/Manuscript.Rmd
Acknowledge

@@
 This project has received funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement No 101020792).
-KJ has been supported by a JSPS Postdoctoral fellowhip, LAYG by a BBSRC Discovery fellowship (BB/W010305/1), PS by a Wellcome Trust Institutional Strategic Support Award (204909/Z/16/Z), and KCAW by the EPSRC Hub for Quantitive Modelling in Healthcare (EP/T017856/1).
+KJ has been supported by a JSPS Overseas Research Fellowship, LAYG by a BBSRC Discovery fellowship (BB/W010305/1), PS by a Wellcome Trust Institutional Strategic Support Award (204909/Z/16/Z), and KCAW by the EPSRC Hub for Quantitive Modelling in Healthcare (EP/T017856/1).
 

### 2023-06-16 Manuscript/Manuscript.Rmd
Update key resource table

@@
 
-## UV-avoidance circuits of extraocular photoreceptors
+**UV-avoidance circuits of extraocular photoreceptors**
 
@@
 
-```{r echo=FALSE}
+```{r echo=FALSE, warning=FALSE}
 library(readxl)
@@
 df_key_resource_table <- as.data.frame(key_resource_table)
+df_key_resource_table <- replace(df_key_resource_table, is.na(df_key_resource_table), "")
 knitr::kable(df_key_resource_table, format = "html", booktabs = TRUE)
+
 ```
@@
 
+```{=tex}
 \begin{align*}
@@
 \end{align*}
-
+```
 ## Data analysis and figure preparation

### 2023-06-16 Manuscript/Manuscript.Rmd
Ms: correct path for res table

--- 2023-06-14 Manuscript/Manuscript.Rmd
@@
     theme: lumen
-  word_document: default
   pdf_document:
@@
     template: default
+  word_document: default
 geometry: left=0.8in,right=0.8in,top=0.8in,bottom=0.5in
@@
 This free radical synthesized from L-arginine by nitric oxide synthase (NOS) is short-lived and can diffuse across biological membranes [@Cudeiro1999; @thomas2015].
-Canonical NO signalling involves the Ca^2+^/calmodulin-dependent activation of NOS, NO production and diffusion, and the NO-dependent activation of soluble guanylate cyclases (sGC).
+Canonical NO signalling involves the Ca2+/calmodulin-dependent activation of NOS, NO production and diffusion, and the NO-dependent activation of soluble guanylate cyclases (sGC).
 This can occur cell-autonomously or in other cells leading to cGMP production [@Bredt1990; @Hoelscher1997].
@@
 The cPRCs express a ciliary-type opsin, c-opsin1 [@Arendt2004] that forms a UV-absorbing bistable photopigment with an absorption maximum around 384 nm [@tsukamoto2017; @Veraszto2018; @Veedin2021].
-Upon UV/violet exposure, the cPRCs show a characteristic biphasic calcium response that is *c-opsin1* dependent.
+Upon UV/violet exposure, the cPRC show a characteristic biphasic calcium response that is *c-opsin1* dependent.
 UV/violet avoidance is also *c-opsin1*-dependent and is defective in *c-opsin1* mutants [@Veraszto2018].
@@
 This reporter also revealed the axonal projections of these central NOS-expressing neurons.
-The position and morphology of the four *NOS+* cells allowed us to identify the same four cells as four interneurons (INNOS) in our three-day-old whole-body *Platynereis* volume EM data [@williams2017; @veraszto2020] (Figure 1B, C).
-In the synaptic connectome, the INNOS cells are postsynaptic to the UV-sensory cPRCs and presynaptic to the INRGWa interneurons, which are also cPRC targets (Figure 1C,F).
+The position and morphology of the four *NOS+* cells allowed us to identify the same four cells as four interneurons (INNOS) in our three-day-old whole-body *Platynereis* volume EM data [@williams2017; @veraszto2020] (Figure 1B,C).
+In the synaptic connectome, the INNOS cells are postsynaptic to the UV-sensory cPRCs and presynaptic to the INRGW interneurons, which are also cPRC targets (Figure 1C,F).
 The projections of INNOS cells are segregated into input (dendritic) and output (axonal) compartments, with cPRC inputs occurring on the dendritic part in the dense neurosecretory plexus.
 INNOS output synapses form in the more ventral projection region (Figure 1---figure supplement 3).
-INRGWa neurons synapse on the head serotonergic ciliomotor neurons (Ser-h1), which synapse on the cholinergic MC ciliomotor neuron and cells of the prototroch ciliary band (Figure 1C,F) [@veraszto2017].
+INRGW neurons synapse on the head serotonergic ciliomotor neurons (Ser-h1), which synapse on the prototroch ciliary band and the cholinergic MC ciliomotor neurons (Figure 1C,F) [@veraszto2017].
 
@@
 Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** **(A)** Scanning electron microscopy image of a three-day-old *Platynereis* larva.
-**(B, C)** Volume rendering of the neuron types (cPRC, INNOS, INRGWa, Ser-h1 and MC) in the cPRC circuit reconstructed from a whole-body transmission electron microscopy volume of a three-day-old larva.
+**(B, C)** Volume rendering of the neuron types (cPRC, INNOS, INRGW, Ser-h1 and MC) in the cPRC circuit reconstructed from a whole-body transmission electron microscopy volume of a three-day-old larva.
 Neurite skeletons are shown with cell-body positions represented by spheres.
@@
 As previously shown, wild-type larvae swim downward following non-directional UV/violet light stimulation [@Veraszto2018].
-In contrast, both two- and three-day-old homozygous *NOS*-mutant larvae showed a strongly diminished UV-avoidance response (Figure 3A, B and Figure 3---figure supplement 1B, C).
+In contrast, both two- and three-day-old homozygous *NOS*-mutant larvae showed a strongly diminished UV-avoidance response (Figure 3A, B and Figure 3---figure supplement 1B,C).
 This phenotype is similar to the defective UV-avoidance of *c-opsin1* mutant larvae [@Veraszto2018] and reveals a requirement for NOS in UV-avoidance behaviour.
@@
 
-To distinguish between an acute and developmental function of NOS in light responses, we also tested larvae exposed to the NOS inhibitor L-NAME.
+To distinguish between an acute and developmental function of NOS in light responses, we next tested larvae exposed to the NOS inhibitor L-NAME.
 Larvae incubated for 5 min in 0.1 mM or 1 mM L-NAME showed a dose-dependent inhibition of UV avoidance.
@@
 As we have shown previously, a 20-sec local stimulation of cPRC cilia led to a transient increase in cPRC calcium levels, followed by a transient decrease [@Veraszto2018].
-After \~20 sec, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus -- a response that may involve depolarisation (Figure 4A, B).
+After \~20-sec, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus -- a response that may involve depolarisation (Figure 4A, B).
 This activation phase occurs after the 20 sec stimulation period and is likely due to a delayed neuroendocrine feedback [@Veraszto2018].
@@
 NO binding to the heme group of sGC leads to increased cyclic guanosine monophosphate (cGMP) production.
-Analysis of sGCs in *Platynereis* indicated that these genes are not expressed in the cPRCs [@veraszto2017] (not shown).
+Analysis of sGCs in *Platynereis* indicated that these genes are not expressed in any of the cells of the cPRC circuit [@veraszto2017].
 Recently, Moroz and coworkers reported an atypical but widely conserved family of guanylyl cyclases with a NIT (nitrite/nitrate sensing) domain (PF08376) (NIT-GC) as potential mediators of NO signalling [@Moroz2020].
@@
 To investigate how NO and NIT-GC signalling influence the dynamics of the cPRC circuit, we imaged calcium signals from postsynaptic neurons in wild type, mutant and morphant larvae.
-We were able to image the activity of all neurons in the cPRC circuit (INNOS, INRGWa, Ser-h1 and MC).
+We were able to image the activity of all neurons in the cPRC circuit (INNOS, INRGW, Ser-h1 and MC).
 The MC cell was identified based on its position and intrinsic activity [@veraszto2017].
 To unambiguously identify all other cells from which we recorded calcium signals, we developed an on-slide immunostaining method (Figure 5---figure supplement 1A).
-We used the cell-specific antibody markers against RYamide (INNOS) (Figure 5---figure supplement 1B-D), RGWamide (INRGWa) and serotonin (Ser-h1) [@Conzelmann2011] to immunostain agar-embedded larvae following calcium imaging.
-Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision (Figure 5A, B).
+We used the cell-specific antibody markers against RYamide (INNOS) (Figure 5---figure supplement 1B-D), RGWamide (INRGW) and serotonin (Ser-h1) [@Conzelmann2011] to immunostain agar-embedded larvae following calcium imaging.
+Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision (Figure 5A,B).
 Due to the stereotypy of the larvae, we could also identify neurons based on their position and calcium activity in activity-correlation maps (Figure 5C).
 
-We first quantified the responses of the INNOS and INRGWa interneurons during 405 nm stimulation of the cPRCs.
+We first quantified the responses of the INNOS and INRGW interneurons during 405 nm stimulation of the cPRCs.
 In both wild type and *NOS*-mutant larvae, INNOS cells showed an increase in calcium during stimulation (Figure 5D).
 In contrast, the INNOS response was flat or slightly negative in NIT-GC2 morphant larvae (Figure 5E) revealing an essential role for NIT-GC2-mediated cPRC suppression in INNOS activation.
-INRGWa cells were initially inhibited during cPRC stimulation, followed by a delayed activation paralleling the second activation phase of cPRCs.
-This late INRGWa response was lacking in *NOS*-mutants (Figure 5F).
-In NIT-GC2 morphants, INRGWa cells showed a transient increase in calcium that decayed after light off and a delayed activation was not present (Figure 5G).
+INRGW cells were initially inhibited during cPRC stimulation, followed by a delayed activation paralleling the second activation phase of cPRCs.
+This late INRGW response was lacking in *NOS*-mutants (Figure 5F).
+In NIT-GC2 morphants, INRGW cells showed a transient increase in calcium that decayed after light off and a delayed activation was not present (Figure 5G).
 
@@
 These data suggest that during 405 nm stimulation the Ser-h1 cells are inhibited and MC cells are activated, and this regulation is NO-dependent.
-This pattern is expected to inhibit ciliary activity in the prototroch but not in the other ciliary bands (with no Ser-h1 or MC synapses), triggering NO-dependent downward swimming.
+This pattern is expected to inhibit ciliary activity in the prototroch but not in the other ciliary bands, triggering NO-dependent downward swimming.
 
 ```{r echo=FALSE,fig.cap="**Figure 5.
-NOS- and NIT-GC2-dependent dynamics of the cPRC circuit.** **(A, B)** GCaMP6s imaging from cPRCs and INNOS cells (left panels) followed by on-slide immunostaining for (A) RYamide to label INNOS and (B) RGWamide+serotonin to label INRGWa and Ser-h1 (red).
+NOS- and NIT-GC2-dependent dynamics of the cPRC circuit.** **(A, B)** GCaMP6s imaging from cPRCs and INNOS cells (left panels) followed by on-slide immunostaining for (A) RYamide to label INNOS and (B) RGWamide+serotonin to label INRGW and Ser-h1 (red).
 Nuclei are stained with DAPI (cyan).
@@
 Numbers mark the same cells in the GCaMP and immunostaining images matched by position.
-**(C)** Correlation map of neuronal activity of the cPRCs, INNOS, INRGWa, Ser-h1 and MC neurons.
+**(C)** Correlation map of neuronal activity of the cPRCs, INNOS, INRGW, Ser-h1 and MC neurons.
 **(D)** GCaMP6s fluorescence in INNOS cells in wild type (WT) and NOSΔ11/Δ11 mutant larvae during 405 nm stimulation of the cPRC cilia.
 **(E)** GCaMP6s fluorescence in INNOS cells in NIT-GC2 morphant larvae during 405 nm stimulation.
-**(F)** GCaMP6s fluorescence in INRGWa cells in wild type and NOSΔ23/Δ23 mutant larvae during 405 nm stimulation.
-**(G)** GCaMP6s fluorescence in INRGWa cells in NIT-GC2 morphant larvae during 405 nm stimulation.
+**(F)** GCaMP6s fluorescence in INRGW cells in wild type and NOSΔ23/Δ23 mutant larvae during 405 nm stimulation.
+**(G)** GCaMP6s fluorescence in INRGW cells in NIT-GC2 morphant larvae during 405 nm stimulation.
 **(H, I)** GCaMP6s fluorescence in (H) Ser-h1 cells and (I) the MC cell in wild type and NOSΔ11/Δ11 mutant larvae during 405 nm stimulation."}
@@
 To further analyse the dynamics of responses to UV light and formally describe cPRC phototransduction, synaptic connections and NO retrograde signalling, we developed a mixed cellular-circuit-level mathematical model.
-We used our Ca^2+^ imagining data of cPRC, INNOS and INRGWa cells collected in wild type, *NOS* knockout, and *NIT-GC2* morphant larvae to formulate assumptions that are the basis of the proposed model.
+We used our Ca2+ imagining data of cPRC, INNOS and INRGW cells collected in wild type, *NOS* knockout, and *NIT-GC2* morphant larvae to formulate assumptions that are the basis of the proposed model.
 
-We model a direct c-opsin1-dependent response to UV leading to the initial rise in Ca^2+^ levels.
+We model a direct c-opsin1-dependent response to UV leading to the initial rise in Ca2+ levels.
 The mechanism of this signal is not known but may include Gα and Gβ signalling, CNGAα channel activation or other pathways.
-We further assume that in cPRC cells, Ca^2+^ levels increase with cGMP.
-We model a decrease in cGMP and Ca^2+^ levels due to a NIT-GC2-dependent pathway (Figre 6A).
+We further assume that in cPRC cells, Ca2+ levels increase with cGMP.
+We model a decrease in cGMP and Ca2+ levels due to a NIT-GC2-dependent pathway (Figure 6A).
 The effect of NO in cPRC cells is captured by another term that describes a NIT-GC1 and NO-dependent increase in cGMP.
-Based on the synaptic connectome, we infer a feedforward coupling between cPRC cells and INNOS and INRGWa cells.
+Based on the synaptic connectome, we infer a feedforward coupling between cPRC cells and INNOS and INRGW cells.
 The proposed synaptic coupling is UV-dependent, decays linearly and is suppressed in NIT-GC2 morphants (Figure 6A).
 
-In the INNOS cells, we assume that excitatory synaptic input from cPRC (or rebound from tonic inhibition) leads to a rise in Ca^2+^ leading to NO production via NOS.
+In the INNOS cells, we assume that excitatory synaptic input from cPRC (or rebound from tonic inhibition) leads to a rise in Ca2+ leading to NO production via NOS.
 This input is NIT-GC2-dependent.
-In the INRGWa cells, we model a decrease in Ca^2+^ based on an inhibitory UV-dependent synaptic signal from the cPRCs.
-We further assume the existence of direct excitatory coupling between cPRC Ca^2+^ levels and INRGWa Ca^2+^ to account for the late effects of cPRC on INRGWa (Figure 6A).
-The UV-dependent and Ca^2+^-dependent signals may be mediated by different transmitters released by cPRCs during distinct phases of their activation cycle.
-In addition, we assume a direct inhibitory coupling between INNOS Ca^2+^ and INRGWa Ca^2+^.
+In the INRGW cells, we model a decrease in Ca2+ based on an inhibitory UV-dependent synaptic signal from the cPRCs.
+We further assume the existence of direct excitatory coupling between cPRC Ca2+ levels and INRGW Ca2+ to account for the late effects of cPRC on INRGW.
+The UV-dependent and Ca2+-dependent signals may be mediated by different transmitters released by cPRCs during distinct phases of their activation cycle.
+In addition, we assume a direct inhibitory coupling between INNOS Ca2+ and INRGW Ca2+.
 Finally, in all variables, we assumed a linear decay and constant production to set a steady state (Figure 6A).
@@
 UV stimulation is modelled as a square pulse.
+To find model parameters producing an output fitting the experimental data we employed a global optimisation method known as a genetic algorithm (GA).
 
-To find model parameters producing an output fitting the experimental data we employed a global optimisation method known as a genetic algorithm (GA).
-Through this optimisation procedure, we found parameters that gave a very good fit to our experimental data under all conditions.
-We could fit to mean Ca^2+^ values but also individual calcium transients (Figure 6B, Figure 6--figure supplement 1).
-Our model only includes a minimal set of parameters and assumptions of interactions that are required to describe the dynamics of cPRC, INNOS and INRGWa in wild type and loss-of-function conditions (Figure 6A, B).
-The model highlights that cPRC phototransduction employs distinct pathways that operate on different time scales and differentially influence cPRC Ca^2+^ levels.
+Our model only includes a minimal set of parameters and assumptions of interactions that are required to describe the dynamics of cPRC, INNOS and INRGW in wild type and loss-of-function conditions (Figure 6A, B).
+The model highlights that cPRC phototransduction employs distinct pathways that operate on different time scales and differentially influence cPRC Ca2+ levels.
 The coupling between cPRCs and interneurons also requires different signalling mechanisms, either through distinct neurotransmitters or receptors expressed in the different cells.
 
-To identify possible molecular pathways, we analysed single-cell transcriptome data for cPRC, INNOS and INRGWa identified by spatial mapping [@Achim2015] and unique marker genes [@williams2017].
+To identify possible molecular pathways, we analysed single-cell transcriptome data for cPRC, INNOS and INRGW identified by spatial mapping [@Achim2015] and unique marker genes [@williams2017].
 In cPRCs, we found transporters and synthetic enzymes indicating cholinergic, glutaminergic, glycinergic, GABAergic and adrenergic neurotransmission (Figure 6C).
-For each neurotransmitter, we also found receptor-encoding genes expressed in INNOS and INRGWa (Figure 6C).
+For each neurotransmitter, we also found receptor-encoding genes expressed in INNOS and INRGW (Figure 6C).
 The two types of interneurons often expressed different subunits or types of these receptors, indicating differential signalling (Figure 6C).
@@
 ```{r echo=FALSE,fig.cap="**Figure 6.
-Mathematical modelling of cPRC circuit dynamics during UV stimulation and potential signalling mechanisms.**  **(A)** Diagram for **(B)** **(C)** Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq.
+Signalling mechanisms and mathematical modelling between cPRC, INNOS and INRGW.** **(A)** Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq.
 The size of the dots is expressed in proportion to the percentage of cells expressing that gene relative to all cells.
 The colours represent the normal logarithm of the number of transcripts in the cells expressing the gene.
-**(D)** Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback."}
+**(B)** Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback.
+**(C)** Diagram for **(D)**"}
+
 knitr::include_graphics(path = "figures/Fig6.png")
+
 ```
-
-To generate new insights from the model, we asked how the duration and intensity of UV stimulation influences signalling output, measured as the magnitude (quantified as area under peak, AUP) of the NO-dependent Ca^2+^ signal in the cPRCs.
-The model allowed us to systematically vary stimulus conditions that would not have been feasible in single-larva experiments.
-We found that the magnitude of the NO-dependent signal scales with stimulus intensity and duration (Figure 7).
-This suggests an integratory function of NO signalling that tunes circuit output to the strength of UV exposure.
 
@@
 NO-mediated retrograde signalling thus likely occurs in the neurosecretory plexus where NOS- and NIT-GC1-containing projections are in close proximity and where we detected NO production following UV stimulation.
-In contrast, INNOS to INRGWa synapses occur outside the neurosecretory plexus in the ventral brain neuropil.
+In contrast, INNOS to INRGW synapses occur outside the neurosecretory plexus in the ventral brain neuropil.
 
@@
 In the UV-avoidance circuit of *Drosophila* larvae, a peptidergic hub neuron Dp7 links UV-sensory neurons (v'td2) and motor circuits.
-Dp7 expresses an insulin-like peptide Ilp7 that is required for acute and sustained UV avoidance.
+Dp7 expresses an an insulin-like peptide Ipl7 that is required for acute and sustained UV avoidance.
 Ilp7 signalling occurs at a functionally and morphologically distinct dendritic compartment of Dp7, segregated from other sensory-motor pathways involving Dp7 [@Imambocus2022].
@@
 NIT domains were first identified in bacteria and animal NIT-GCs have only recently been reported [@Shu2003; @Moroz2020].
-Bacterial NIT domains regulate cellular functions in response to changes in extracellular and intracellular nitrate and nitrite concentrations [@camargo2007].
+Bacterial NIT domains regulate cellular functions in response to changes in extracellular and intracellular nitrate and/or nitrite concentrations [@camargo2007].
 NO is readily converted to nitrate and nitrite [@santos2011; @moeller2019; @garthwaite2015a] and these molecules accumulate in placozoans and cnidarians in cells and tissues with high NOS activity [@Moroz2004; @Moroz2020].
@@
 Based on our data herein and previous work we can now propose a more detailed model of cPRC phototransduction and neurotransmission.
-The cPRCs have high basal Ca^2+^ and respond to UV/violet light dependent on c-opsin1 [@Veraszto2018].
+The cPRCs have high basal Ca2+ and respond to UV/violet light dependent on c-opsin1 [@Veraszto2018].
 c-opsin1 forms a bistable photopigment and signals through Gi/oα and Gβγ [@tsukamoto2017; @Veedin2021; @tsukamoto2023].
-In heterologous systems, the Gβγ subunits released following c-opsin1 activation open GIRK channels inducing K+ efflux (hyperpolarisation) [@tsukamoto2017; @tsukamoto2023] and close voltage-gated Ca^2+^ channels, thereby reducing intracellular calcium levels [@tsukamoto2023].
+In heterologous systems, the Gβγ subunits released following c-opsin1 activation open GIRK channels inducing K+ efflux (hyperpolarisation) [@tsukamoto2017; @tsukamoto2023] and close voltage-gated Ca2+ channels, thereby reducing intracellular calcium levels [@tsukamoto2023].
 A GIRK channel is also expressed in the cPRCs (Figure 6A).
@@
 NIT-GC2-dependent signalling is required for the feedforward activation of the INNOS cells through an unknown transmitter.
-The activation of INNOS leads to NOS activation and NO release, potentially through a canonical Ca^2+^-calmodulin pathway.
-Our model predicts feedforward inhibiton from INNOS to INRGWa, possibly mediated by glycin transmission.
-NO released by INNOS neurites activates NIT-GC1 in cPRC projections that could lead to cGMP production and the opening of CNGAα.
+The activation of INNOS leads to NOS activation and NO release, potentially through a canonical Ca2+-calmodulin pathway.
+Our model predicts feedforward inhibiton from INNOS to INRGW, possibly mediated by glycin transmission.
+NO released by INNOS neurites activates NIT-GC1 in cPRC projections that could lead to cGMP production and the opening CNGAα.
 This late-phase cPRC activation is not directly dependent on the UV/violet signal and can happen post-stimulus.
-The late-phase cPRC activation leads to INRGWa activation via an unknown transmitter that is likely different from the one acting on INNOS.
+The late-phase cPRC activation leads to INRGW activation via an unknown transmitter that is likely different from the one acting on INNOS.
 
 In addition, the cPRC circuit expresses several neuropeptides and their receptors, suggesting further neuromodulatory signals.
-For example, INRGWa express the proneuropeptide RGWamide and cPRCs and INNOS express its receptor, suggesting retrograde peptidergic signalling in the circuit.
+For example, INRGW express the proneuropeptide RGWamide and cPRCs and INNOS express its receptor, suggesting retrograde peptidergic signalling in the circuit.
 
-We have recently shown that the cPRC circuit also mediates responses to hydrostatic pressure via the same motor system involving Ser-h1 and the prototroch ciliary band.
+We have recently shown that the cPRC circuit also mediate responses to hydrostatic pressure via the same motor system involving Ser-h1 and the prototroch ciliary band.
 Increased pressure induces cPRC activation and a circuit output that is inverted relative to UV-induced activation.
@@
 
-Pressure-induced Ca^2+^ transients in cPRCs lack an inhibitory phase and late activation and resemble UV responses in NIT-GC2 morphants [@bezares-calderon2023].
+Pressure-induced Ca2+ transients in cPRCs lack an inhibitory phase and late activation and resemble UV responses in NIT-GC2 morphants [@bezares-calderon2023].
 This indicates that the differentiation of sensory cues by the multisensory cPRCs occurs already at the level of sensory signal transduction.
@@
 
-In the *Platynereis* circuit, our mathematical model indicates that the magnitude of the NO-dependent signal depends on the intensity and duration of the UV/violet stimulus.
-This suggests that the NO-dependent memory trace also encodes the intensity and duration of the stimulus.
+In the *Platynereis* circuit, our mathematical model indicates that the onset, magnitude and duration of the NO-dependent signal depends on the intensity and duration of the UV/violet stimulus.
+This suggests that the NO-dependent memory trace also encodes the magnitude and duration of the stimulus.
 
@@
 
-Non-directional UV/violet light induces downward swimming head down in *Platynereis* larvae [@Veraszto2018].
+UV/violet light induces downward swimming head down [@Veraszto2018].
 Since stimulus direction is not relevant, swimming direction must be determined by gravity.
@@
 Animals evolved distinct photosensory systems coupled to non-overlapping circuits and guiding unique behavioural responses.
-These sensory systems can employ different opsin molecules and be tuned to different wavelengths of light.
+These sensory systems can employ different opsin molecules and be tuned to different wavelength of light.
 The avoidance of noxious UV/violet light is often mediated by extraocular photoreceptors and their circuits, distinct from the pigmented visual eyes.
-These two types of systems co-exist in *Platynereis* where the pigmented eyes and eyespots guide phototaxis with a maximum sensitivity to cyan light (\~500 nm) [@Randel2014a; @Veraszto2018].
+These two types of systems co-exist in *Platynereis* where the pigmented eyes and eyespots guide phototaxis behaviour with a maximum sensitivity to cyan light (\~500 nm) [@Randel2014a; @Veraszto2018].
 Planarians also have pigmented visual eyes mediating phototaxis to cyan light and peripheral extraocular photoreceptors mediating UV avoidance behaviour [@Shettigar2021].
-*Drosophila* larvae have cerebral eyes called Bolwig's organs involved in phototaxis [@kane2013] and several types of extraocular UV-sensory cells that tile the body wall and mediate UV avoidance [@Imambocus2022].
+*Drosophila* larvae have cerebral eyes called Bolwig's organs involved in phototaxis [@kane2013] and several types of extraocular UV-sensory cells that tile the body wall [@Imambocus2022].
 
@@
 The amount of modulator released can scale with stimulus intensity or duration and maintain an altered circuit state due to the slower decay of the diffusive signals relative to synaptic transmission.
-
-We have revealed how NO shapes dynamics in a fully-mapped sensory-motor UV avoidance circuit.
-We identified an unconventional GC as the NO receptor in *Platynereis* and measured circuit activity in different genetic backgrounds.
-Finally, we could link circuit activity to light-avoidance behaviour.
-The richly modulated multi-transmitter cPRC circuit will serve as a fertile ground for future studies on how neuromodulators shape activity and behaviour.
 
@@
 
-For the generation of a *NOS* knockout we used an sgRNA targeting the third exon of the *Platynereis dumerilii* *NOS* (target site: 5'-GGGCAATACTGGCTCCACTC-3').
-We selected the site to avoid polymorphic sites in our laboratory culture.
-The sgRNA was assembled from two annealed oligonucleotides (5'-TAGGGCAATACTGGCTCCACTC-3', 5'-AAACGAGTGGAGCCAGTATTGC-3') forming overhangs for cloning into the BsaI site of the plasmid pDR27456 [\@hwang2013](42250,%20Addgene), which contains next to the BsaI site a tracrRNA sequence.
-We used this plasmid to PCR amplify DNA (primers: T7, 5'-AAAAGCACCGACTCGGTGCC-3') for synthesizing the sgRNA.
-To purify the PCR product, we used the QIAquick PCR Purification Kit (Qiagen).
-We then synthesized sgRNA with the MEGAshortscript Kit (Thermo Fisher Scientific) and purified it with the MEGAclear Kit (Thermo Fisher Scientific).
-To produce the Cas9-mRNA, we used the plasmid (pUC57-T7-RPP2-Cas9) containing the Cas9 ORF fused to 169 base pair 5' UTR from the *Platynereis dumerilii* 60S acidic ribosomal protein P2.
-After in vitro transcription with the mMessage mMachine Kit (Thermo Fisher Scientific), we capped and polyA-tailed the Cas9-mRNA with the Poly(A) Tailing Kit (Thermo Fisher Scientific).
-We coinjected the sgRNA (18 ng/ml) and the Cas9-mRNA (180 ng/µl) into fertilized eggs of *Platynereis dumerilii* wild-type parents according to an established injection procedure [@Conzelmann2013a].
+Before designing the small guide RNA (sgRNA) for the sgRNA:Cas9 nuclease, splice sites and polymorphic sites in our laboratory culture were identified to avoid them.
+The sgRNA targeted the third exon of Platynereis dumerilii NOS (target site: 5'-GGGCAATACTGGCTCCACTC-3').
+The sgRNA was assembled from two annealed oligonucleotides (5'-TAGGGCAATACTGGCTCCACTC-3', 5'-AAACGAGTGGAGCCAGTATTGC-3') forming overhangs for cloning into a BsaI site of the plasmid pDR27456 (Hwang et al.
+2013)(42250, Addgene), which contains next to the BsaI site a tracrRNA sequence.
+The plasmid was then used to PCR amplify DNA (primers: T7, 5'-AAAAGCACCGACTCGGTGCC-3') for synthesizing the sgRNA.
+The DNA was purified with the QIAquick PCR Purification Kit (Qiagen).
+From the DNA, the sgRNA was synthesized with the MEGAshortscript Kit (Thermo Fisher Scientific) and was purified with the MEGAclear Kit (Thermo Fisher Scientific).
+Cas9-mRNA was transcribed, capped, and polyA-tailed with the mMessage mMachine Kit and the Poly(A) Tailing Kit (both Thermo Fisher Scientific) from a plasmid (pUC57-T7-RPP2-Cas9) containing the Cas9 ORF fused to 169 base pair 5' UTR from the Platynereis dumerilii 60S acidic ribosomal protein P2.
+The sgRNA (18 ng/ml) and the Cas9-mRNA (180 ng/µl) were coinjected into fertilized eggs of Platynereis dumerilii wild-type parents according to an established injection procedure [@Conzelmann2013a].
 The eggs were kept at 18°C for 45 min before injection and were injected at 14.5°C.
-The injected individuals were kept at 18°C for 5 to 8 days in 6-well-plates (Nunc multidish no. 150239, Thermo Scientific) and then cultured at 22°C until sexual maturity.
+The injected individuals were kept at 18°C for 5 to 8 days in 6-well- plates (Nunc multidish no. 150239, Thermo Scientific) and then cultured at 22°C until sexual maturity.
 The mature worms were crossed to wild-type worms and the progeny was genotyped, resulting in two founder lines, which were bred to homozygosity.
 
-**Genotyping of NOS alleles**
+**NOS sequencing and genotyping**
 
@@
 
-**Vertical column setup to measure photoresponses**
+**Vertical column setup for measuring photoresponses**
 
-We assayed larval photoresponses in a vertical Plexiglas column (31 mm x 10 mm x 160 mm water height).
-For stimulation, we illuminated the column from the top with a monochromator (Polychrome II, Till Photonics) controlled by AxioVision 4.8.2.0 (Carl Zeiss GmbH, Jena) via analog voltage.
+Photoresponses of larvae of different ages were assayed in a vertical Plexiglas column (31 mm x 10 mm x 160 mm water height).
+The column was illuminated from top with light from a monochromator (Polychrome II, Till Photonics).
+The monochromator was controlled by AxioVision 4.8.2.0 (Carl Zeiss MicroImaging GmbH) via analog voltage.
 The light passed a collimator lens (LAG-65.0-53.0-C with MgF2 Coating, CVI Melles Griot) before entering the column.
@@
 One strip contained UV (395 nm) LEDs (SMB1W-395, Roithner Lasertechnik) and the other infrared (810 nm) LEDs (SMB1W-810NR-I, Roithner Lasertechnik).
-The UV LEDs were run at 4 V. The infrared LEDs were run at 8 V (overvoltage) to illuminate the larvae for imaging with a DMK camera (DMK 22BUC03, The Imaging Source).
-We recorded videos at 15 frames per second with the software IC Capture (The Imaging Source).
+The UV LEDs were run at 4 V to stimulate the larvae in the column from the side.
+The infrared LEDs were run at 8 V (overvoltage) to illuminate the larvae for the camera (DMK 22BUC03, The Imaging Source), which recorded videos at 15 frames per second and was controlled by IC Capture (The Imaging Source).
 
-To compare the behavior of wildtype and NOS-knockout larvae or larvae treated with L-NAME, we first mixed the larvae and left them in the dark for 5 min.
-We then imaged larvae for 1 min in the dark (only IR LED for illumination) followed by exposure to non-directional UV light from both sides, then 2 min darkness, then collimated cyan (480 nm) light from the top of the column for 2 min, then 2 min darkness, and finally collimated UV (395 nm) light from the top of the column for 2 min.
+**Comparing behavior of wildtype and NOS-knockout 3-day-old larvae**
+
+To compare the behavior of wildtype and NOS-knockout larvae at 3 days in the vertical column, the larvae were mixed and left in the dark for 5 min.
+The larvae were treated with NOS inhibitors for pharmacology.
+The NOS inhibitors were L-NAME.
+The larvae were treated with different concentrations in adjacent columns.
+The concentrations for the NOS inhibitors were control, 1 mM, 0.1 mM.
+The larvae were recorded for 1 min in the dark followed by exposure to collimated cyan (480 nm) light from the top of the column for 2 min, then 2 min darkness, and finally collimated UV (395 nm) light from the top of the column for 2 min.
+Stimulus light was provided by the monochromator (Polychrome II, Till Photonics).
 Scripts are available at <https://github.com/JekelyLab/NOS>.
 
-**Identification and Phylogenetic Analysis of NOS sequences**
+**NOS Identification and Phylogenetic Analysis**
 
-To identify NOS sequences, we obtained a seed database of oxygenase domains from the Pfam database (PF02898).
-From these sequences, we produced a Hidden Markov Model (HMM) and used this to mine the 47 metazoan, 2 choanoflagellate and 2 filasterea species investigated.
+To identify NOS, we obtained a "seed" database of oxygenase domain in Pfam database, PF02898.
+From these sequences, we produced a Hidden Markov Model (HMM) and used this to mine the 47 metazoan species, 2 choanoflagellate species and 2 filasterea species investigated.
 HMM models were run in HMMR3 with an e-value of 1e−15.
@@
 To calculate maximum-likelihood trees, we used IQ-tree2 with the LG+G4 model.
-To calculate branch support, we ran 1000 replicates with the aLRT-SH-like and aBayes methods [@Minh2020].
+To calculate branch support, we ran 1,000 replicates with the aLRT-SH-like and aBayes methods [@Minh2020].
 The sequences used for the phylogenetic analysis are available in **Supplementary File 1**, the trimmed alignment is available in **Supplementary File 2** and the pre-trimmed data in **Supplementary file 3**.
 
-**Identification and Phylogenetic Analysis of NIT-GC sequences**
+**NIT-GC Identification and Phylogenetic Analysis**
 
-To identify NIT-GCs, we obtained a seed database of Adenylate and Guanylate cyclase catalytic domains from the Pfam database (PF00211).
-From these sequences, we produced a Hidden Markov Model (HMM) and used this to mine the 45 metazoan, 2 choanoflagellate and 2 filasterea species investigated.
+To identify NIT-GCs, we obtained a "seed" database of Adenylate and Guanylate cyclase catalytic domain in Pfam database, PF00211.
+From these sequences, we produced a Hidden Markov Model (HMM) and used this to mine the 45 metazoan species, 2 choanoflagellate species and 2 filasterea species investigated.
 HMM models were run in HMMR3 with an e-value of 1e−15.
 We ran CD-Hit [@Fu2012] to eliminate redundant sequences (at a 80% threshold).
-To identify clusters, we used sequence-similarity-based clustering in Clans [@frickey2004] and the convex-clustering option with 100 jack-knife replicates.
-The NIT-GCs are extremely well conserved within the membrane-bound guanylate cyclases and form an easily recognizable cluster.
+To identify clusters, we used the convex-clustering option with 100 jack-knife replicates.
+The NIT-GCs are extremely well conserved in membrane-bound guanylate cyclases and form an easily recognizable cluster.
 To analyze the phylogeny of NIT-GCs, the cluster containing these GCs together with membrane-bound guanylate cyclases were parsed and used for tree building.
@@
 To calculate maximum-likelihood trees, we used IQ-tree2 with the LG+G4 model.
-To calculate branch support, we ran 1000 replicates with the aLRT-SH-like and aBayes methods [@Minh2020].
-The sequences used for the phylogenetic analysis are available in **Supplementary File 4**, the trimmed alignment is available in **Supplementary File 5** and the pre-trimmed data in **Supplementary file 6**.
+To calculate branch support, we ran 1,000 replicates with the aLRT-SH-like and aBayes methods [@Minh2020].
+The sequences used for the phylogenetic analysis are available in **Supplementary File 1**, the trimmed alignment is available in **Supplementary File 2** and the pre-trimmed data in **Supplementary file 3**.
 
@@
 
-We used a single-cell RNA sequencing dataset from *Platynereis* larvae [@Achim2015] that was further filtered to 107 cells [@williams2017].
-We normalized raw read-count data to TPM and converted to log10.
-To calculate % expression, we used the sum of the expression levels in the 107 cells for each gene.
+We used Achim et al. for the single-cell data [@Achim2015].
+In Williams et al., they used 107 cells as neurons by removing duplicates from Achim et al. single-cell data, so we used those cells [@williams2017].
+Since the raw data were read count data, we normalized them to TPM using Python.
+After that we converted them to log10.
+From the sum of the expression levels in 107 cells for each gene, We calculated the percentage in each cell.
+For each cell, we identified them with marker genes.
+After created the data in Python, plotted it using R dot plots.
+RPKM calculates the total number of reads per million bp and then divides by the length of each gene, so it is not possible to compare between samples.
+Instead, TPM first divides by the length of the gene and then divides by the total number of reads per million bases, which allows for more accurate comparisons between samples.
+In this case we wanted to compare between samples, so we used TPM.
 The total TPM of each gene between the samples was used to calculate the percentage of expressed genes.
-We processed the data in Python and plotted in R.
-Individual cells were identified by suits of well-characterised marker genes and spatial mapping [@Achim2015].
-The total TPM values for each gene and the percentage of expressed genes are available in **Supplementary file 7**.
+The total TPM values for each gene and the percentage of expressed genes are available in **Supplementary file 1**.
 
@@
 Hairpins associated with the b2 initiator sequence were labeled with Alexa Fluor 647, and the hairpins associated with the b3 initiator sequence were labeled with Alexa Fluor 546.
-To design probes for HCR, we used custom software [@Kuehn2022] to create 20 DNA oligo probe pairs specific to *P.
-dumerilii* *NOS*, *NIT-GC1*, *NIT-GC2*, *RYa proneuropeptide* (GenBank accession: JF811330.1), and *MLD/pedal peptide 2 proneuropeptide* (GenBank accession: KF515945.1).
-The *NOS*, *NIT-GC1* and *NIT-GC2* probes were designed to be associated with the b2 initiator sequence, while the *RYa* and *MLD/pedal peptide 2* probes were designed to be associated with the b3 initiator sequence.
+To design probes for HCR, we used custom software [@Kuehn2022] to create 20 DNA oligo probe pairs specific to P.
+dumerilii NOS, NIT-GC1, NIT-GC2, RYa-pNP (GenBank accession: JF811330.1), and MLD/pedal 2-pNP (GenBank accession: KF515945.1).
+The NOS, NIT-GC1 and NIT-GC2 probes were designed to be associated with the b2 initiator sequence, while the RYa-pNP and MLD/pedal 2-pNP probes were designed to be associated with the b3 initiator sequence.
 For the detection stage, samples were pre-hybridized in 200 µl of probe hybridization buffer for 1 hr at 37°C, and then incubated in 250 µl hybridization buffer containing probe oligos (4 pmol/ml) overnight at 37°C.
@@
 
-Whole-mount immunostaining of *Platynereis* larvae fixed with 4% paraformaldehyde was carried out using the primary antibodies listed in KEY RESOURCES TABLE.
+Whole-mount immunostaining of 2 day old Platynereis larvae fixed with 4% paraformaldehyde were carried out using primary antibodies raised against NIT-GC1, NIT-GC2, NOS, RYamide neuropeptide, RGWamide neuropeptide in rabbit, plus a commercial antibody raised against acetylated tubulin in mouse (Sigma T7451).
+The synthetic peptides contained an N-terminal Cys that was used for coupling during purification.
+Antibodies were affinity purified from sera as previously described [@Conzelmann2012].
 Immunostainings were carried out as previously described [@Conzelmann2012].
-
-**Antibody generation and purification**
-
-Antibodies were raised in rats or rabbits against synthetic peptides containing an N-term Cys residue (Altabioscience).
-The same Cys-containing synthetic peptides were used for affinity purification.
-Antibodies were affinity purified from sera on a SulfoLink Coupling Resin (Thermo Fischer) as previously described [@Conzelmann2012].
-View a detailed protocol here: <https://bio-protocol.org/exchange/preprintdetail?id=2333&type=3>
-
-**Transient transgenesis**
-
-A 12 kb genomic region upstream of the start site of the *NOS* gene was amplified and cloned upstream of 3xHA-Palmi-tdTomato to yield the plasmid *NOS-3xHA-Palmi-tdTomato*.
-Larvae injected with the promoter construct (ca.
-250 ng/ml) were analysed for reporter expression at three days post fertilization on an AxioImager Z.1 fluorescence microscope (Carl Zeiss GmbH, Jena) and if positive fixed for immunostaining.
-The immunostaining for the HA-tagged reporter was done as described [@veraszto2017].
-Stained specimens were imaged on an LSM 780 NLO or LSM 880 Airysan Confocal Microscope (Carl Zeiss GmbH, Jena).
+The NOS promoter (fragment sizes: 12 Kb) was amplified and cloned upstream of 3xHA- Palmi-tdTomato.
+Larvae injected with promoter constructs (ca.
+250 ng/ml) were analysed for reporter expression at 3 days post fertilization using an AxioImager Z.1 fluorescence wide-field microscope (Carl Zeiss GmbH, Jena) and immediately fixed for immunostainings.
+The protocol followed for immunostaining of HA-tagged reporters was recently described [@veraszto2017].
+Specimens were imaged with a LSM 780 NLO or LSM 880 with Airysan Confocal Microscope (Zeiss, Jena).
 
@@
 
-For calcium imaging, 49-55 hours post fertilisation larvae were used.
-Experiments were carried out at room temperature and larvae were immobilised in 2.5% agarose in artificial sea water and mounted between a slide and coverslip spaced with adhesive tape.
+For calcium imaging, 49--55 hpf larvae were used.
+Experiments were performed at room temperature and larvae were immobilised by being embedded in 2.5% agarose filtered artificial seawater between a slide and coverslip spaced with adhesive tape.
 GCaMP6s mRNA (1 mg/ml) was injected into zygotes as described previously [@Randel2014a].
 Larvae were imaged on a Zeiss LSM 880 with Airyscan (with a C-Apochromat 63X/1.2 Corr - water) with a frame rate of 1.88 frame/sec and an image size of 512 x 512 pixels.
-The larvae were stimulated in a region of interest (a circle with 50 pixel diameter) with a 405 nm laser controlled by the Bleaching mode of the Zeiss software.
+The larvae were stimulated in a region of interest (a circle with 50 pixel diameter) with 405 nm lasers controlled by the Bleaching mode.
 The imaging laser had a similar intensity than the stimulus laser but covered an area that was 10 times larger than the stimulus ROI.
 
-**cGMP production assay**
+**Cell culture experiment**
 
-To measure cGMP production in a cell culture assay we used the Green cGull reporter [@Matsuda2017].
-Full-length *NIT-GC1* and *NIT-GC2* coding sequences were amplified by PCR from a *Platynereis dumerilii* cDNA library and cloned into the pcDNA3.1(+) vector with the T2A self-cleaving sequence.
-For the assay, COS-7 cells (Angio-proteomie, CAT no. cAP-0203) with a low endogenous level of soluble guanylate cyclase activity were used.
-Cells were maintained at 37 °C in 35 mm dishes (Nunc™ Glass Bottom Dishes) containing 3 ml of DMEM, high glucose glutamax medium (Thermo; Cat.
+Green cGull was used for the cGMP assay [@Matsuda2017].
+A full-length Pdum-NIT-GC1 and -NIT-GC2 coding sequences were amplified by PCR starting from a *Platynereis dumerilii* cDNA library and cloned into the pcDNA3.1(+) vector using the T2A self-cleaving sequence.
+Cos-7 cells with low expression of endogenous soluble guanylate cyclase were used as cultured cells for gene expression.
+This cell line was purchased from Angio-proteomie (CAT no. cAP-0203).
+The Cos-7 cells were maintained at 37 °C in 35mm dishes (Nunc™ Glass Bottom Dishes) containing 3 mL of DMEM, high glucose glutamax medium (Thermo; Cat.
 No.
@@
 10082147).
-Upon reaching approximately 85% confluency, cells were transfected with the pcDNA3.1(+)-Green-cGull-T2A-NITGC1 plasmid.
-Transfections were carried out with 150 ng of each plasmid and 0.3 μl of Lipofectamine 3000 Reagent (invitrogen; Cat. No. L3000001).
-Two days post-transfection, the culture medium was substituted for fresh DMEM-medium.
-For single-wavelength imaging experiments, cells in 35-mm dishes were washed twice and imaged in modified Ringer's buffer (140 mM NaCl, 3.5 mM KCl, 0.5 mM NaH~2~PO~4~ , 0.5 mM S-3 MgSO~4~, 1.5 mM CaCl~2~ , 10 mM HEPES, 2 mM NaHCO~3~ and 5 mM glucose).
-Dishes were mounted on a stage heated to 37 °C and imaging was done on an inverted confocal microscope (LSM880, Carl Zeiss GmbH, Jena) equipped with an oil-immersion objective lens (UApo/340, 40x, NA = 0.17).
-The exposure time of the EM-CCD camera was controlled by the ZEN software (Carl Zeiss GmbH, Jena).
+Upon reaching confluency of approximately 85%, we transfected the cells with the plasmid containing Green cGull-T2A-NITGC1.
+Transfections were carried out with 150 ng of each plasmid and 0.3 μl of the transfection Lipofectamine 3000 Reagent (invitrogen; Cat. No. L3000001).
+Two days post-transfection, we removed the culture medium and substituted it for fresh DMEM-medium.
+For single-wavelength imaging experiments, cells in 35-mm dishes were washed twice and imaged in modified Ringer's buffer (140 mM NaCl, 3.5 mM KCl, 0.5mM NaH2PO4 , 0.5mM S-3 MgSO4 , 1.5 mM CaCl2 , 10 mM HEPES, 2 mM NaHCO3 and 5 mM glucose).
+Dishes were mounted on a stage heated at 37 °C and imaging was performed using an inverted microscope (LSM880, Zeiss) equipped with an oil-immersion objective lens (UApo/340, 40, NA = 0.17).
+Images were acquired using a xenon lamp, 460--495 nm excitation filter, 505-nm dichroic mirror and 510-- 550-nm emission filter (Zeiss).
+S-Nitroso-N-acetyl-D, L-penicillamine (SNAP) was purchased from Sigma-Aldrich (St. Louis, MO, USA) .
+The exposure time of the EM-CCD camera was controlled by the ZEN software (Zeiss).
 Images were acquired every 15 s for 10 min and stimulation was initiated 2 min after starting image acquisition.
-Imaging data analysis was done with ImageJ (National Institutes of Health, Bethesda, MD, USA).
-As NO donor, we used S-Nitroso-N-acetyl-D, L-penicillamine (SNAP, Sigma-Aldrich, St. Louis, MO, USA).
-
-## Mathematical modelling
-
-The mathematical model for the wild type data comprises six dynamical equations:
-
-\begin{align*}
-\frac{d C_P(t)}{dt} &= 1 - \delta^{C_P} C_P(t) + K_{UV}^{C_P} UV(t) C_O(t)  + K_{G}^{C_P} G(t) , \\
-\frac{dC_O(t)}{dt} &= 1 - \delta^{C_O} C_O(t) - K_{UV}^{C_O} UV(t) C_O(t), \\
-\frac{dG(t)}{dt} &= 1 - \delta^{G} G(t) + K_{GC2}^G \left(1 - UV(t)\right) G(t) + K_{GC1}^G N(t) , \\
-\frac{dS(t)}{dt} &= UV(t) - \frac{S(t)}{\tau_S}, \\
-\frac{dC_N(t)}{dt} &= 1 - \delta^{C_N} C_N(t) + K_{S}^{C_N} S(t), \\
-\frac{dN(t)}{dt} &= 1 - \delta^{N} N(t) + K_{C_N}^{N} C_N(t), \\
-\frac{dC_R(t)}{dt} &= 1 - \delta^{C_R} C_R(t) - K_{S}^{C_R} S(t) + K_{C_P}^{C_R} C_P(t) - K_{C_N}^{C_R} C_N(t) C_R(t), \\
-UV(t) &=
-    \begin{cases}
-        a, & t_{UV\: start} \leq t \leq t_{UV\: end}\\
-        0, & \text{otherwise}
-    \end{cases}   .
-\end{align*}
-
-## Data analysis and figure preparation
-
-Data were analysed in ImageJ and R.
-All figures were assembled in R with the cowplot and patchwork packages.
-All scripts are available at <https://github.com/JekelyLab/Joukra_et_al_NOS> (commit ).
+Imaging data analysis was performed using ImageJ (National Institutes of Health, Bethesda, MD, USA).
 
@@
 This project has received funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement No 101020792).
-KJ has been supported by a JSPS Postdoctoral fellowhip, LAYG by a BBSRC Discovery fellowship (BB/W010305/1), PS by a Wellcome Trust Institutional Strategic Support Award (204909/Z/16/Z), and KCAW by the EPSRC Hub for Quantitive Modelling in Healthcare (EP/T017856/1).
+KJ has been supported by a JSPS Postdoctoral fellowhip, LAYG by a BBSRC Discovery fellowship (BB/W010305/1).
 
@@
 
-```{r echo=FALSE,fig.cap="**Figure 1 -- figure supplement 3** ssTEM reconstruction of the INNOS (A), INRGWa (B) and cPRC (C), anterior view. NS plexus; neurosecretory plexus"}
+```{r echo=FALSE,fig.cap="**Figure 1 -- figure supplement 3** ssTEM reconstruction of the INNOS (A), INRGW (B) and cPRC (C), anterior view. NS plexus; neurosecretory plexus"}
 knitr::include_graphics(path = "figures/Fig1_sup3.png")

### 2023-06-16 Manuscript/Manuscript.Rmd
ms: edits

@@
 ```{r echo=FALSE,fig.cap="**Figure 1.
-Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** **(A)** Scanning electron microscopy image of a three-day-old *Platynereis* larva.
-**(B, C)** Volume rendering of the neuron types (cPRC, INNOS, INRGW, Ser-h1 and MC) in the cPRC circuit reconstructed from a whole-body transmission electron microscopy volume of a three-day-old larva.
+Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** (**A**) Scanning electron microscopy image of a three-day-old *Platynereis* larva.
+(**B**, **C**) Volume rendering of the neuron types (cPRC, INNOS, INRGW, Ser-h1 and MC) in the cPRC circuit reconstructed from a whole-body transmission electron microscopy volume of a three-day-old larva.
 Neurite skeletons are shown with cell-body positions represented by spheres.
@@
 In B, nuclei positions of the prototroch head ciliary band are shown as grey spheres.
-**(D)** Expression of the *NOS* gene detected by in situ HCR (magenta) in a two-day-old larva (anterior view).
+(**D**) Expression of the *NOS* gene detected by in situ HCR (magenta) in a two-day-old larva (anterior view).
 Antibody staining for acetylated α-tubulin (acTub: green) highlights cPRC cilia and the neuropil.
-**(E)** Expression of a *NOS* reporter (NOSp::palmi-3xHA: magenta) labelled with an anti-HA antibody in a two-day-old larva (anterior view).
+(**E**) Expression of a *NOS* reporter (NOSp::palmi-3xHA: magenta) labelled with an anti-HA antibody in a two-day-old larva (anterior view).
 Antibody staining for acetylated α-tubulin (acTub: green) highlights cPRC cilia and the neuropil.
-**(F)** Synaptic wiring diagram of the cPRC circuit.
+(**F**) Synaptic wiring diagram of the cPRC circuit.
 Hexagons represent cell groups, with the number of cells per group shown in square brackets.
@@
 ```{r echo=FALSE,fig.cap="**Figure 2.
-NO produced by UV/violet stimulation to cPRCs.** **(A)** DAF-FM fluorescence in the region of the neurosecretory neuropil.
+NO produced by UV/violet stimulation to cPRCs.** (**A**) DAF-FM fluorescence in the region of the neurosecretory neuropil.
 The white line indicates the outline of the larva.
@@
 cPRCs are marked by thin lines.
-**(B)** Changes in DAF-FM fluorescence before and during 405 nm light stimulation.
-**(C)** Changes in DAF-FM fluorescence over time during 405 nm stimulation of the cPRCs (green) or a control area (ctr: gray).
+(**B**) Changes in DAF-FM fluorescence before and during 405 nm light stimulation.
+(**C**) Changes in DAF-FM fluorescence over time during 405 nm stimulation of the cPRCs (green) or a control area (ctr: gray).
 The purple box indicates the duration of 405 nm stimulation.
@@
 ```{r echo=FALSE,fig.cap="**Figure 3.
-NOS is required for UV avoidance in *Platynereis* larvae.**  **(B)** Swimming trajectories of wild type (WT, n=32) and NOS mutant (NOSΔ11/Δ11, n=26 and NOSΔ23/Δ23, n=47) three-day-old larvae.
+NOS is required for UV avoidance in *Platynereis* larvae.**  (**A**) Swimming trajectories of wild type (WT, n=32) and NOS mutant (NOSΔ11/Δ11, n=26 and NOSΔ23/Δ23, n=47) three-day-old larvae.
 All trajectories start at 0 x and y position and time 0 corresponding to 10 sec after the onset of 395 nm stimulation from the side.
-**(B)** Vertical position of batches of wild type and mutant three-day-old larvae over time under 395 nm UV stimulation.
+(**B**) Vertical position of batches of wild type and mutant three-day-old larvae over time under 395 nm UV stimulation.
 The starting position of each larval trajectory was set to 0.
-**(C)** Vertical position of batches of control and L-NAME-treated (0.1 and 1 mM) three-day-old larvae over time under 395 nm UV stimulation.
+(**C**) Vertical position of batches of control and L-NAME-treated (0.1 and 1 mM) three-day-old larvae over time under 395 nm UV stimulation.
 The starting position of each larval trajectory was set to 0.
-**(D)** Vertical displacement in 30 sec bins of wild type and mutant (NOSΔ11 and NOSΔ23) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
-**(E)** Vertical displacement in 30 sec bins of control and L-NAME-treated (0.1 and 1 mM) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top."}
+(**D**) Vertical displacement in 30 sec bins of wild type and mutant (NOSΔ11 and NOSΔ23) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
+(**E**) Vertical displacement in 30 sec bins of control and L-NAME-treated (0.1 and 1 mM) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top."}
 
@@
 ```{r echo=FALSE,fig.cap="**Figure 4.
-NOS and two NIT-GCs shape calcium signals during cPRC UV/violet response.** **(A, B)** GCaMP6s signals in cPRCs in wild type and *NOS* mutant (A, NOSΔ11/Δ11, B, NOSΔ23/Δ23) larvae during 405 nm light stimulation.
-**(C, D)** In situ HCR for (C) *NIT-GC1* and (D) *NIT-GC2* (magenta) in three-day-old *Platynereis* larvae.
+NOS and two NIT-GCs shape calcium signals during cPRC UV/violet response.** (**A**, **B**) GCaMP6s signals in cPRCs in wild type and *NOS* mutant (A, NOSΔ11/Δ11, B, NOSΔ23/Δ23) larvae during 405 nm light stimulation.
+(**C**, **D**) In situ HCR for (C) *NIT-GC1* and (D) *NIT-GC2* (magenta) in three-day-old *Platynereis* larvae.
 Larvae were co-stained with an antibody against acetylated α-tubulin to label cPRC cilia and the neuropil (green).
-**(E, F)** Immunostaining for (E) NIT-GC1 and (F) NIT-GC2 (magenta), co-stained for acetylated α-tubulin (green).
-**(G)** The domain structure of *Platynereis* NIT-GC1 and the truncated NIT-GC1ΔNIT protein lacking the NIT domain.
+(**E**, **F**) Immunostaining for (E) NIT-GC1 and (F) NIT-GC2 (magenta), co-stained for acetylated α-tubulin (green).
+(**G**) The domain structure of *Platynereis* NIT-GC1 and the truncated NIT-GC1ΔNIT protein lacking the NIT domain.
 A predicted transmembrane region (TM) is shown in grey.
-**(H)** Schematic of the cell-based assay to detect cGMP production following the addition of an NO donor SNAP or DMSO as control.
-**(I-L)** Green cGull fluorescence over time for the four conditions tested.
+(**H**) Schematic of the cell-based assay to detect cGMP production following the addition of an NO donor SNAP or DMSO as control.
+(**I**-**L**) Green cGull fluorescence over time for the four conditions tested.
 Individual responses and their mean with 0.95 confidence interval are shown (n > 6 cells).
@@
 The indicated chemicals were added at 2 min after the start of imaging (grey bars).
-**(M, N)** GCaMP6s signals in cPRCs in (M) NIT-GC1 and (N) NIT-GC2 morphant larvae during 405 nm light stimulation.
+(**M**, **N**) GCaMP6s signals in cPRCs in (M) NIT-GC1 and (N) NIT-GC2 morphant larvae during 405 nm light stimulation.
 Individual responses and their mean with 0.95 confidence interval are shown."}
@@
 ```{r echo=FALSE,fig.cap="**Figure 5.
-NOS- and NIT-GC2-dependent dynamics of the cPRC circuit.** **(A, B)** GCaMP6s imaging from cPRCs and INNOS cells (left panels) followed by on-slide immunostaining for (A) RYamide to label INNOS and (B) RGWamide+serotonin to label INRGW and Ser-h1 (red).
+NOS- and NIT-GC2-dependent dynamics of the cPRC circuit.** (**A**, **B**) GCaMP6s imaging from cPRCs and INNOS cells (left panels) followed by on-slide immunostaining for (A) RYamide to label INNOS and (B) RGWamide+serotonin to label INRGW and Ser-h1 (red).
 Nuclei are stained with DAPI (cyan).
@@
 Numbers mark the same cells in the GCaMP and immunostaining images matched by position.
-**(C)** Correlation map of neuronal activity of the cPRCs, INNOS, INRGW, Ser-h1 and MC neurons.
-**(D)** GCaMP6s fluorescence in INNOS cells in wild type (WT) and NOSΔ11/Δ11 mutant larvae during 405 nm stimulation of the cPRC cilia.
-**(E)** GCaMP6s fluorescence in INNOS cells in NIT-GC2 morphant larvae during 405 nm stimulation.
-**(F)** GCaMP6s fluorescence in INRGW cells in wild type and NOSΔ23/Δ23 mutant larvae during 405 nm stimulation.
-**(G)** GCaMP6s fluorescence in INRGW cells in NIT-GC2 morphant larvae during 405 nm stimulation.
-**(H, I)** GCaMP6s fluorescence in (H) Ser-h1 cells and (I) the MC cell in wild type and NOSΔ11/Δ11 mutant larvae during 405 nm stimulation."}
+(**C**) Correlation map of neuronal activity of the cPRCs, INNOS, INRGW, Ser-h1 and MC neurons.
+(**D**) GCaMP6s fluorescence in INNOS cells in wild type (WT) and NOSΔ11/Δ11 mutant larvae during 405 nm stimulation of the cPRC cilia.
+(**E**) GCaMP6s fluorescence in INNOS cells in NIT-GC2 morphant larvae during 405 nm stimulation.
+(**F**) GCaMP6s fluorescence in INRGW cells in wild type and NOSΔ23/Δ23 mutant larvae during 405 nm stimulation.
+(**G**) GCaMP6s fluorescence in INRGW cells in NIT-GC2 morphant larvae during 405 nm stimulation.
+(**H**, **I**) GCaMP6s fluorescence in (H) Ser-h1 cells and (I) the MC cell in wild type and NOSΔ11/Δ11 mutant larvae during 405 nm stimulation."}
 
@@
 ```{r echo=FALSE,fig.cap="**Figure 6.
-Signalling mechanisms and mathematical modelling between cPRC, INNOS and INRGW.** **(A)** Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq.
+Mathematical modelling and signalling mechanisms of the cPRC circuit.** (**A**) Diagram of the mathematical model with componenets, interactions, parameters and equations used to model Ca^2+^ dynamics.
+(**B**) Average calcium traces and modelled traces fitted to the data.
+(**C**) Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq.
 The size of the dots is expressed in proportion to the percentage of cells expressing that gene relative to all cells.
 The colours represent the normal logarithm of the number of transcripts in the cells expressing the gene.
-**(B)** Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback.
-**(C)** Diagram for **(D)**"}
+(**D**) Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback.
+"}
 
 knitr::include_graphics(path = "figures/Fig6.png")
-
 ```
@@
 
+## Mathematical modelling
+
+The mathematical model for the wild type data comprises six dynamical equations:
+
+```{=tex}
+\begin{align*}
+\frac{d C_P(t)}{dt} &= 1 - \delta^{C_P} C_P(t) + K_{UV}^{C_P} UV(t) C_O(t)  + K_{G}^{C_P} G(t) , \\
+\frac{dC_O(t)}{dt} &= 1 - \delta^{C_O} C_O(t) - K_{UV}^{C_O} UV(t) C_O(t), \\
+\frac{dG(t)}{dt} &= 1 - \delta^{G} G(t) + K_{GC2}^G \left(1 - UV(t)\right) G(t) + K_{GC1}^G N(t) , \\
+\frac{dS(t)}{dt} &= UV(t) - \frac{S(t)}{\tau_S}, \\
+\frac{dC_N(t)}{dt} &= 1 - \delta^{C_N} C_N(t) + K_{S}^{C_N} S(t), \\
+\frac{dN(t)}{dt} &= 1 - \delta^{N} N(t) + K_{C_N}^{N} C_N(t), \\
+\frac{dC_R(t)}{dt} &= 1 - \delta^{C_R} C_R(t) - K_{S}^{C_R} S(t) + K_{C_P}^{C_R} C_P(t) - K_{C_N}^{C_R} C_N(t) C_R(t), \\
+UV(t) &=
+    \begin{cases}
+        a, & t_{UV\: start} \leq t \leq t_{UV\: end}\\
+        0, & \text{otherwise}
+    \end{cases}   .
+\end{align*}
+```
+## Data analysis and figure preparation
+
+Data were analysed in ImageJ and R.
+All figures were assembled in R with the cowplot and patchwork packages.
+All scripts are available at <https://github.com/JekelyLab/Joukra_et_al_NOS> (commit ).
+>>>>>>> 9932f067da10604ec00c679a5b10f91decddfa29
+
 ## Acknowledgements
@@
 This project has received funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement No 101020792).
-KJ has been supported by a JSPS Postdoctoral fellowhip, LAYG by a BBSRC Discovery fellowship (BB/W010305/1).
+KJ has been supported by a JSPS Overseas Research Fellowship, LAYG by a BBSRC Discovery fellowship (BB/W010305/1), PS by a Wellcome Trust Institutional Strategic Support Award (204909/Z/16/Z), and KCAW by the EPSRC Hub for Quantitive Modelling in Healthcare (EP/T017856/1).
 
@@
 
-```{r echo=FALSE,fig.cap="**Figure 1 -- figure supplement 2** Expression analysis of the *NOS* gene (magenta) using in situ HCR using the larva at three-day-old in dorsal **(A)** and anterior view (B).
+```{r echo=FALSE,fig.cap="**Figure 1 -- figure supplement 2** Expression analysis of the *NOS* gene (magenta) using in situ HCR using the larva at three-day-old in dorsal (**A**) and anterior view (B).
 (B) Insert: showing two NOS cells (INNOS_dl and INNOS_vl) on left side."}
@@
 
-```{r echo=FALSE,fig.cap="**Figure 3 -- figure supplement 1** **(A)** Top: The domain and Exon/Intron structure of NOS.
+```{r echo=FALSE,fig.cap="**Figure 3 -- figure supplement 1** (**A**) Top: The domain and Exon/Intron structure of NOS.
 Bottom: Close-up region showing the genomic locus of NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9.
@@
 Gray shows PAM sequences, red shows stop codons.
-**(B)** Overlaid trajectories for WT (n=37) and NOS mutant (NOSΔ11/Δ11, n=18 and NOSΔ23/Δ23, n=8) at two-day-old larvae.
+(**B**) Overlaid trajectories for WT (n=37) and NOS mutant (NOSΔ11/Δ11, n=18 and NOSΔ23/Δ23, n=8) at two-day-old larvae.
 0 sec as the starting point.
 After 10 sec, UV (395 nm) stimulation from the side.
-**(C)** The temporal changes in the vertical position of the WT and mutant two-days-old larvae before and after UV stimulation are shown.
+(**C**) The temporal changes in the vertical position of the WT and mutant two-days-old larvae before and after UV stimulation are shown.
 The starting points of each larval trajectory are set to 0.
@@
 **(D, E)** The temporal changes in the distance traveled of the WT and mutant in two-day (D) and three-day-old (E) larvae before and after UV stimulation are shown.
-**(F)** The temporal changes in the distance traveled of larvae treated with NOS inhibitors, L-NAME in three-day larvae before and after UV stimulation are shown.
-**(G)** Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23) larvae at two-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
+(**F**) The temporal changes in the distance traveled of larvae treated with NOS inhibitors, L-NAME in three-day larvae before and after UV stimulation are shown.
+(**G**) Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23) larvae at two-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
 The data are shown in 30 s bins."}
@@
 
-```{r echo=FALSE,fig.cap="**Figure 4 -- figure supplement 3** **(A)** Co-expression analysis image of the NIT-GC1 (magenta) and MLD-pedal2 amide proneuropeptide gene (MLD: green).
+```{r echo=FALSE,fig.cap="**Figure 4 -- figure supplement 3** (**A**) Co-expression analysis image of the NIT-GC1 (magenta) and MLD-pedal2 amide proneuropeptide gene (MLD: green).
 Anterior view of the larva at two-day-old.
@@
 Anterior (B) and posterior (C) views of the larva at three-day-old.
-**(D)** Co-localisation analysis using NIT-GC1 (magenta) and NOS (green) antibodies.
+(**D**) Co-localisation analysis using NIT-GC1 (magenta) and NOS (green) antibodies.
 Anterior view of the larva at two-day-old.
@@
 
-```{r echo=FALSE,fig.cap="**Figure 5 -- figure supplement 1** **(A)** Schematic diagram of the immunostaining procedure after calcium imaging.**(B-D)** Co-expression analysis image of the NOS (magenta) and RY amide proneuropeptide gene (RYa: green).
+```{r echo=FALSE,fig.cap="**Figure 5 -- figure supplement 1** (**A**) Schematic diagram of the immunostaining procedure after calcium imaging.**(B-D)** Co-expression analysis image of the NOS (magenta) and RY amide proneuropeptide gene (RYa: green).
 Anterior view of the larva at two-day-old.

### 2023-06-16 Manuscript/Manuscript.Rmd
edit MS

@@
 To calculate branch support, we ran 1,000 replicates with the aLRT-SH-like and aBayes methods [@Minh2020].
-The sequences used for the phylogenetic analysis are available in **Supplementary File 1**, the trimmed alignment is available in **Supplementary File 2** and the pre-trimmed data in **Supplementary file 3**.
+The sequences used for the phylogenetic analysis are available in **Supplementary file 1**, the trimmed alignment is available in **Supplementary File 2** and the pre-trimmed data in **Supplementary file 3**.
 
@@
 To calculate branch support, we ran 1,000 replicates with the aLRT-SH-like and aBayes methods [@Minh2020].
-The sequences used for the phylogenetic analysis are available in **Supplementary File 1**, the trimmed alignment is available in **Supplementary File 2** and the pre-trimmed data in **Supplementary file 3**.
+The sequences used for the phylogenetic analysis are available in **Supplementary file 4**, the trimmed alignment is available in **Supplementary file 5** and the pre-trimmed data in **Supplementary file 6**.
 
@@
 The total TPM of each gene between the samples was used to calculate the percentage of expressed genes.
-The total TPM values for each gene and the percentage of expressed genes are available in **Supplementary file 1**.
+The total TPM values for each gene and the percentage of expressed genes are available in **Supplementary file 7**.
 

### 2023-06-24 Manuscript/Manuscript.Rmd
MS: add math model methods, upload pics, general edits

@@
 UV stimulation of cPRCs triggers INNOS activation and NO production.
-NO signals retrogradely to cPRCs to induce their sustained post-stimulus activation through an unconventional guanylyl cyclase.
+NO signals retrogradely to cPRCs to induce their sustained post-stimulus activation through an unconventional guanylate cyclase.
 This late activation inhibits serotonergic ciliomotor neurons to induce downward swimming.
@@
 
-Here we uncovered an essential function for NO signalling in larval UV/violet-light avoidance behaviour.
+We uncovered an essential function for NO signalling in larval UV/violet-light avoidance behaviour.
 In *Platynereis* larvae, UV/violet avoidance is mediated by brain ciliary photoreceptor cells (cPRCs) and is characterised by downward swimming [@Veraszto2018].
 The cPRCs express a ciliary-type opsin, c-opsin1 [@Arendt2004] that forms a UV-absorbing bistable photopigment with an absorption maximum around 384 nm [@tsukamoto2017; @Veraszto2018; @Veedin2021].
-Upon UV/violet exposure, the cPRC show a characteristic biphasic calcium response that is *c-opsin1* dependent.
+Upon UV/violet exposure, the cPRCs show a characteristic biphasic calcium response that is c-opsin1 dependent.
 UV/violet avoidance is also *c-opsin1*-dependent and is defective in *c-opsin1* mutants [@Veraszto2018].
@@
 The position and morphology of the four *NOS+* cells allowed us to identify the same four cells as four interneurons (INNOS) in our three-day-old whole-body *Platynereis* volume EM data [@williams2017; @veraszto2020] (Figure 1B,C).
-In the synaptic connectome, the INNOS cells are postsynaptic to the UV-sensory cPRCs and presynaptic to the INRGW interneurons, which are also cPRC targets (Figure 1C,F).
+In the synaptic connectome, the INNOS cells are postsynaptic to the UV-sensory cPRCs and presynaptic to the INRGWa interneurons, which are also cPRC targets (Figure 1C, F). 
 The projections of INNOS cells are segregated into input (dendritic) and output (axonal) compartments, with cPRC inputs occurring on the dendritic part in the dense neurosecretory plexus.
 INNOS output synapses form in the more ventral projection region (Figure 1---figure supplement 3).
-INRGW neurons synapse on the head serotonergic ciliomotor neurons (Ser-h1), which synapse on the prototroch ciliary band and the cholinergic MC ciliomotor neurons (Figure 1C,F) [@veraszto2017].
+INRGWa neurons synapse on the head serotonergic ciliomotor neurons (Ser-h1), which synapse on the cholinergic MC ciliomotor neuron and cells of the prototroch ciliary band (Figure 1C, F) [@veraszto2017].
 
 ```{r echo=FALSE,fig.cap="**Figure 1.
-Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** (**A**) Scanning electron microscopy image of a three-day-old *Platynereis* larva.
-(**B**, **C**) Volume rendering of the neuron types (cPRC, INNOS, INRGW, Ser-h1 and MC) in the cPRC circuit reconstructed from a whole-body transmission electron microscopy volume of a three-day-old larva.
+Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** **(A)** Scanning electron microscopy image of a three-day-old *Platynereis* larva.
+**(B, C)** Volume rendering of the neuron types (cPRC, INNOS, INRGWa, Ser-h1 and MC) in the cPRC circuit reconstructed from a whole-body transmission electron microscopy volume of a three-day-old larva.
 Neurite skeletons are shown with cell-body positions represented by spheres.
@@
 In B, nuclei positions of the prototroch head ciliary band are shown as grey spheres.
-(**D**) Expression of the *NOS* gene detected by in situ HCR (magenta) in a two-day-old larva (anterior view).
+**(D)** Expression of the *NOS* gene detected by in situ HCR (magenta) in a two-day-old larva (anterior view).
 Antibody staining for acetylated α-tubulin (acTub: green) highlights cPRC cilia and the neuropil.
-(**E**) Expression of a *NOS* reporter (NOSp::palmi-3xHA: magenta) labelled with an anti-HA antibody in a two-day-old larva (anterior view).
+**(E)** Expression of a *NOS* reporter (NOSp::palmi-3xHA: magenta) labelled with an anti-HA antibody in a two-day-old larva (anterior view).
 Antibody staining for acetylated α-tubulin (acTub: green) highlights cPRC cilia and the neuropil.
-(**F**) Synaptic wiring diagram of the cPRC circuit.
+**(F)** Synaptic wiring diagram of the cPRC circuit.
 Hexagons represent cell groups, with the number of cells per group shown in square brackets.
@@
 
-**Two unconventional guanylyl cyclases are expressed in the cPRCs**
+**Two unconventional guanylate cyclases are expressed in the cPRCs**
 
@@
 NO binding to the heme group of sGC leads to increased cyclic guanosine monophosphate (cGMP) production.
-Analysis of sGCs in *Platynereis* indicated that these genes are not expressed in any of the cells of the cPRC circuit [@veraszto2017].
-Recently, Moroz and coworkers reported an atypical but widely conserved family of guanylyl cyclases with a NIT (nitrite/nitrate sensing) domain (PF08376) (NIT-GC) as potential mediators of NO signalling [@Moroz2020].
+Analysis of sGCs in *Platynereis* indicated that these genes are not expressed in any of the cells of the cPRC circuit (not shown and [@veraszto2017]).
+Recently, Moroz and coworkers reported an atypical but widely conserved family of guanylate cyclases with a NIT (nitrite/nitrate sensing) domain (PF08376) (NIT-GC) as potential mediators of NO signalling [@Moroz2020].
 To identify NIT-GCs in *Platynereis*, we searched transcriptome resources and retrieved 15 potential NIT-GC homologs (Figure 4---figure supplement 1 and 2).
 To analyse the relationship of these sequences to metazoan NIT-GCs, we retrieved protein sequences with a CYC domain from the transcriptome and genome databases of 45 metazoan and 2 choanoflagellate species.
-We carried out cluster analysis and did phylogenetic reconstruction on a group of membrane-bound guanylyl cyclases with sGCs as an outgroup.
+We carried out cluster analysis and did phylogenetic reconstruction on a group of membrane-bound guanylate cyclases with sGCs as an outgroup.
 In agreement with Moroz et al. [@Moroz2020], we found a group of GCs with NIT domains with representatives in placozoans, cnidarians, some ecdysozoans, echinoderms, and lophotrochozoans.
@@
 
-To further characterise these atypical guanylyl cyclases, we focused on NIT-GC1 and carried out in vitro experiments.
+To further characterise these atypical guanylate cyclases, we focused on NIT-GC1 and carried out in vitro experiments.
 In bacteria, NIT domains are thought to regulate cellular functions in response to intra- or extracellular nitrate and nitrite.
@@
 To investigate how NO and NIT-GC signalling influence the dynamics of the cPRC circuit, we imaged calcium signals from postsynaptic neurons in wild type, mutant and morphant larvae.
-We were able to image the activity of all neurons in the cPRC circuit (INNOS, INRGW, Ser-h1 and MC).
+We were able to image the activity of all neurons in the cPRC circuit (INNOS, INRGWa, Ser-h1 and MC).
 The MC cell was identified based on its position and intrinsic activity [@veraszto2017].
 To unambiguously identify all other cells from which we recorded calcium signals, we developed an on-slide immunostaining method (Figure 5---figure supplement 1A).
-We used the cell-specific antibody markers against RYamide (INNOS) (Figure 5---figure supplement 1B-D), RGWamide (INRGW) and serotonin (Ser-h1) [@Conzelmann2011] to immunostain agar-embedded larvae following calcium imaging.
-Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision (Figure 5A,B).
+We used the cell-specific antibody markers against RYamide (INNOS) (Figure 5---figure supplement 1B-D), RGWamide (INRGWa) and serotonin (Ser-h1) [@Conzelmann2011] to immunostain agar-embedded larvae following calcium imaging.
+Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision (Figure 5A, B).
 Due to the stereotypy of the larvae, we could also identify neurons based on their position and calcium activity in activity-correlation maps (Figure 5C).
 
-We first quantified the responses of the INNOS and INRGW interneurons during 405 nm stimulation of the cPRCs.
+We first quantified the responses of the INNOS and INRGWa interneurons during 405 nm stimulation of the cPRCs.
 In both wild type and *NOS*-mutant larvae, INNOS cells showed an increase in calcium during stimulation (Figure 5D).
 In contrast, the INNOS response was flat or slightly negative in NIT-GC2 morphant larvae (Figure 5E) revealing an essential role for NIT-GC2-mediated cPRC suppression in INNOS activation.
-INRGW cells were initially inhibited during cPRC stimulation, followed by a delayed activation paralleling the second activation phase of cPRCs.
-This late INRGW response was lacking in *NOS*-mutants (Figure 5F).
-In NIT-GC2 morphants, INRGW cells showed a transient increase in calcium that decayed after light off and a delayed activation was not present (Figure 5G).
+INRGWa cells were initially inhibited during cPRC stimulation, followed by a delayed activation paralleling the second activation phase of cPRCs.
+This late INRGWa response was lacking in *NOS*-mutants (Figure 5F).
+In NIT-GC2 morphants, INRGWa cells showed a transient increase in calcium that decayed after light off and a delayed activation was not present (Figure 5G).
 
@@
 These data suggest that during 405 nm stimulation the Ser-h1 cells are inhibited and MC cells are activated, and this regulation is NO-dependent.
-This pattern is expected to inhibit ciliary activity in the prototroch but not in the other ciliary bands, triggering NO-dependent downward swimming.
+This pattern is expected to inhibit ciliary activity in the prototroch but not in the other ciliary bands (with no Ser-h1 or MC synapses), triggering NO-dependent downward swimming.
 
 ```{r echo=FALSE,fig.cap="**Figure 5.
-NOS- and NIT-GC2-dependent dynamics of the cPRC circuit.** (**A**, **B**) GCaMP6s imaging from cPRCs and INNOS cells (left panels) followed by on-slide immunostaining for (A) RYamide to label INNOS and (B) RGWamide+serotonin to label INRGW and Ser-h1 (red).
+NOS- and NIT-GC2-dependent dynamics of the cPRC circuit.** **(A, B)** GCaMP6s imaging from cPRCs and INNOS cells (left panels) followed by on-slide immunostaining for (A) RYamide to label INNOS and (B) RGWamide+serotonin to label INRGWa and Ser-h1 (red).
 Nuclei are stained with DAPI (cyan).
@@
 Numbers mark the same cells in the GCaMP and immunostaining images matched by position.
-(**C**) Correlation map of neuronal activity of the cPRCs, INNOS, INRGW, Ser-h1 and MC neurons.
-(**D**) GCaMP6s fluorescence in INNOS cells in wild type (WT) and NOSΔ11/Δ11 mutant larvae during 405 nm stimulation of the cPRC cilia.
-(**E**) GCaMP6s fluorescence in INNOS cells in NIT-GC2 morphant larvae during 405 nm stimulation.
-(**F**) GCaMP6s fluorescence in INRGW cells in wild type and NOSΔ23/Δ23 mutant larvae during 405 nm stimulation.
-(**G**) GCaMP6s fluorescence in INRGW cells in NIT-GC2 morphant larvae during 405 nm stimulation.
-(**H**, **I**) GCaMP6s fluorescence in (H) Ser-h1 cells and (I) the MC cell in wild type and NOSΔ11/Δ11 mutant larvae during 405 nm stimulation."}
+**(C)** Correlation map of neuronal activity of the cPRCs, INNOS, INRGWa, Ser-h1 and MC neurons.
+**(D)** GCaMP6s fluorescence in INNOS cells in wild type (WT) and NOSΔ11/Δ11 mutant larvae during 405 nm stimulation of the cPRC cilia.
+**(E)** GCaMP6s fluorescence in INNOS cells in NIT-GC2 morphant larvae during 405 nm stimulation.
+**(F)** GCaMP6s fluorescence in INRGWa cells in wild type and NOSΔ23/Δ23 mutant larvae during 405 nm stimulation.
+**(G)** GCaMP6s fluorescence in INRGWa cells in NIT-GC2 morphant larvae during 405 nm stimulation.
+**(H, I)** GCaMP6s fluorescence in (H) Ser-h1 cells and (I) the MC cell in wild type and NOSΔ11/Δ11 mutant larvae during 405 nm stimulation."}
 
@@
 To further analyse the dynamics of responses to UV light and formally describe cPRC phototransduction, synaptic connections and NO retrograde signalling, we developed a mixed cellular-circuit-level mathematical model.
-We used our Ca2+ imagining data of cPRC, INNOS and INRGW cells collected in wild type, *NOS* knockout, and *NIT-GC2* morphant larvae to formulate assumptions that are the basis of the proposed model.
+We used our Ca^2+^ imagining data of cPRC, INNOS and INRGWa cells collected in wild type, *NOS* knockout, and *NIT-GC2* morphant larvae to formulate assumptions that are the basis of the proposed model.
 
-We model a direct c-opsin1-dependent response to UV leading to the initial rise in Ca2+ levels.
+We model a direct c-opsin1-dependent response to UV leading to the initial rise in Ca^2+^ levels.
 The mechanism of this signal is not known but may include Gα and Gβ signalling, CNGAα channel activation or other pathways.
-We further assume that in cPRC cells, Ca2+ levels increase with cGMP.
-We model a decrease in cGMP and Ca2+ levels due to a NIT-GC2-dependent pathway (Figure 6A).
+We further assume that in cPRC cells, Ca^2+^ levels increase with cGMP.
+We model a decrease in cGMP and Ca^2+^ levels due to a NIT-GC2-dependent pathway (Figure 6A).
 The effect of NO in cPRC cells is captured by another term that describes a NIT-GC1 and NO-dependent increase in cGMP.
-Based on the synaptic connectome, we infer a feedforward coupling between cPRC cells and INNOS and INRGW cells.
+Based on the synaptic connectome, we infer a feedforward coupling between cPRC cells and INNOS and INRGWa cells.
 The proposed synaptic coupling is UV-dependent, decays linearly and is suppressed in NIT-GC2 morphants (Figure 6A).
 
-In the INNOS cells, we assume that excitatory synaptic input from cPRC (or rebound from tonic inhibition) leads to a rise in Ca2+ leading to NO production via NOS.
+In the INNOS cells, we assume that excitatory synaptic input from cPRC (or rebound from tonic inhibition) leads to a rise in Ca^2+^ leading to NO production via NOS.
 This input is NIT-GC2-dependent.
-In the INRGW cells, we model a decrease in Ca2+ based on an inhibitory UV-dependent synaptic signal from the cPRCs.
-We further assume the existence of direct excitatory coupling between cPRC Ca2+ levels and INRGW Ca2+ to account for the late effects of cPRC on INRGW.
-The UV-dependent and Ca2+-dependent signals may be mediated by different transmitters released by cPRCs during distinct phases of their activation cycle.
-In addition, we assume a direct inhibitory coupling between INNOS Ca2+ and INRGW Ca2+.
+In the INRGWa cells, we model a decrease in Ca^2+^ based on an inhibitory UV-dependent synaptic signal from the cPRCs.
+We further assume the existence of direct excitatory coupling between cPRC Ca^2+^ levels and INRGWa Ca^2+^ to account for the late effects of cPRC on INRGWa (Figure 6A).
+The UV-dependent and Ca^2+^-dependent signals may be mediated by different transmitters released by cPRCs during distinct phases of their activation cycle.
+In addition, we assume a direct inhibitory coupling between INNOS Ca^2+^ and INRGWa Ca^2+^.
 Finally, in all variables, we assumed a linear decay and constant production to set a steady state (Figure 6A).
@@
 
-Our model only includes a minimal set of parameters and assumptions of interactions that are required to describe the dynamics of cPRC, INNOS and INRGW in wild type and loss-of-function conditions (Figure 6A, B).
-The model highlights that cPRC phototransduction employs distinct pathways that operate on different time scales and differentially influence cPRC Ca2+ levels.
+Through this optimisation procedure, we found parameters that gave a very good fit to our experimental data under all conditions.
+We could fit to mean Ca^2+^ values but also individual calcium transients (Figure 6B, Figure 6--figure supplement 1).
+Our model only includes a minimal set of parameters and assumptions of interactions that are required to describe the dynamics of cPRC, INNOS and INRGWa in wild type and loss-of-function conditions (Figure 6A, B).
+The model highlights that cPRC phototransduction employs distinct pathways that operate on different time scales and differentially influence cPRC Ca^2+^ levels.
 The coupling between cPRCs and interneurons also requires different signalling mechanisms, either through distinct neurotransmitters or receptors expressed in the different cells.
 
-To identify possible molecular pathways, we analysed single-cell transcriptome data for cPRC, INNOS and INRGW identified by spatial mapping [@Achim2015] and unique marker genes [@williams2017].
+To identify possible molecular pathways, we analysed single-cell transcriptome data for cPRC, INNOS and INRGWa identified by spatial mapping [@Achim2015] and unique marker genes [@williams2017].
 In cPRCs, we found transporters and synthetic enzymes indicating cholinergic, glutaminergic, glycinergic, GABAergic and adrenergic neurotransmission (Figure 6C).
-For each neurotransmitter, we also found receptor-encoding genes expressed in INNOS and INRGW (Figure 6C).
+For each neurotransmitter, we also found receptor-encoding genes expressed in INNOS and INRGWa (Figure 6C).
 The two types of interneurons often expressed different subunits or types of these receptors, indicating differential signalling (Figure 6C).
@@
 
+To generate new insights from the model, we asked how the duration and intensity of UV stimulation influences signalling output, measured as the magnitude (quantified as area under peak, AUP) of the NO-dependent Ca^2+^ signal in the cPRCs.
+The model allowed us to systematically vary stimulus conditions that would not have been feasible in single-larva experiments.
+We found that delayed activation occurs only for UV stimulation of sufficient amplitude and  duration (Figure 7).
+This suggests an integratory function of NO signalling to tune circuit output to the strength of UV exposure.
+
+```{r echo=FALSE,fig.cap="**Figure 7.
+Magnitude of cPRC activation as  a function of stimulus duration and intensity.** Median values of the $AUP$ are presented as a heatmap with dark blue indicating flat $C_P(t)$ trajectories with median $AUP$ close to 0 and red indicating that the $C_P(t)$ trajectories have a clearly defined delayed activation peak.
+"}
+
+knitr::include_graphics(path = "figures/Fig7.svg")
+```
+
 ## Discussion
@@
 In the UV-avoidance circuit of *Drosophila* larvae, a peptidergic hub neuron Dp7 links UV-sensory neurons (v'td2) and motor circuits.
-Dp7 expresses an an insulin-like peptide Ipl7 that is required for acute and sustained UV avoidance.
+Dp7 expresses an insulin-like peptide Ilp7 that is required for acute and sustained UV avoidance.
 Ilp7 signalling occurs at a functionally and morphologically distinct dendritic compartment of Dp7, segregated from other sensory-motor pathways involving Dp7 [@Imambocus2022].
@@
 NIT-GC2-dependent signalling is required for the feedforward activation of the INNOS cells through an unknown transmitter.
-The activation of INNOS leads to NOS activation and NO release, potentially through a canonical Ca2+-calmodulin pathway.
-Our model predicts feedforward inhibiton from INNOS to INRGW, possibly mediated by glycin transmission.
-NO released by INNOS neurites activates NIT-GC1 in cPRC projections that could lead to cGMP production and the opening CNGAα.
+The activation of INNOS leads to NOS activation and NO release, potentially through a canonical Ca^2+^-calmodulin pathway.
+Our model predicts feedforward inhibiton from INNOS to INRGWa, possibly mediated by glycin transmission.
+NO released by INNOS neurites activates NIT-GC1 in cPRC projections that could lead to cGMP production and the opening of CNGAα.
 This late-phase cPRC activation is not directly dependent on the UV/violet signal and can happen post-stimulus.
-The late-phase cPRC activation leads to INRGW activation via an unknown transmitter that is likely different from the one acting on INNOS.
+The late-phase cPRC activation leads to INRGWa activation via an unknown transmitter that is likely different from the one acting on INNOS.
 
 In addition, the cPRC circuit expresses several neuropeptides and their receptors, suggesting further neuromodulatory signals.
-For example, INRGW express the proneuropeptide RGWamide and cPRCs and INNOS express its receptor, suggesting retrograde peptidergic signalling in the circuit.
+For example, INRGWa express the proneuropeptide RGWamide and cPRCs and INNOS express its receptor, suggesting retrograde peptidergic signalling in the circuit.
 
-We have recently shown that the cPRC circuit also mediate responses to hydrostatic pressure via the same motor system involving Ser-h1 and the prototroch ciliary band.
+We have recently shown that the cPRC circuit also mediates responses to hydrostatic pressure via the same motor system involving Ser-h1 and the prototroch ciliary band.
 Increased pressure induces cPRC activation and a circuit output that is inverted relative to UV-induced activation.
@@
 The amount of modulator released can scale with stimulus intensity or duration and maintain an altered circuit state due to the slower decay of the diffusive signals relative to synaptic transmission.
+
+Overall, we have revealed how NO shapes the dynamics of a fully mapped sensory-motor UV avoidance circuit.
+We identified an unconventional GC as the NO receptor and measured circuit activity in different genetic backgrounds.
+Finally, we could link circuit activity to light-avoidance behaviour.
+The richly modulated multi-transmitter cPRC circuit will serve as a fertile ground for future studies on how neuromodulators shape activity and behaviour.
 
@@
 
-Before designing the small guide RNA (sgRNA) for the sgRNA:Cas9 nuclease, splice sites and polymorphic sites in our laboratory culture were identified to avoid them.
-The sgRNA targeted the third exon of Platynereis dumerilii NOS (target site: 5'-GGGCAATACTGGCTCCACTC-3').
-The sgRNA was assembled from two annealed oligonucleotides (5'-TAGGGCAATACTGGCTCCACTC-3', 5'-AAACGAGTGGAGCCAGTATTGC-3') forming overhangs for cloning into a BsaI site of the plasmid pDR27456 (Hwang et al.
-2013)(42250, Addgene), which contains next to the BsaI site a tracrRNA sequence.
-The plasmid was then used to PCR amplify DNA (primers: T7, 5'-AAAAGCACCGACTCGGTGCC-3') for synthesizing the sgRNA.
-The DNA was purified with the QIAquick PCR Purification Kit (Qiagen).
-From the DNA, the sgRNA was synthesized with the MEGAshortscript Kit (Thermo Fisher Scientific) and was purified with the MEGAclear Kit (Thermo Fisher Scientific).
-Cas9-mRNA was transcribed, capped, and polyA-tailed with the mMessage mMachine Kit and the Poly(A) Tailing Kit (both Thermo Fisher Scientific) from a plasmid (pUC57-T7-RPP2-Cas9) containing the Cas9 ORF fused to 169 base pair 5' UTR from the Platynereis dumerilii 60S acidic ribosomal protein P2.
-The sgRNA (18 ng/ml) and the Cas9-mRNA (180 ng/µl) were coinjected into fertilized eggs of Platynereis dumerilii wild-type parents according to an established injection procedure [@Conzelmann2013a].
-The eggs were kept at 18°C for 45 min before injection and were injected at 14.5°C.
-The injected individuals were kept at 18°C for 5 to 8 days in 6-well- plates (Nunc multidish no. 150239, Thermo Scientific) and then cultured at 22°C until sexual maturity.
+For the generation of a *NOS* knockout we used an sgRNA targeting the third exon of the *Platynereis dumerilii* *NOS* (target site: 5'-GGGCAATACTGGCTCCACTC-3').
+We selected the site to avoid polymorphic sites in our laboratory culture.
+The sgRNA was assembled from two annealed oligonucleotides (5'-TAGGGCAATACTGGCTCCACTC-3', 5'-AAACGAGTGGAGCCAGTATTGC-3') forming overhangs for cloning into the BsaI site of the plasmid pDR27456 [\@hwang2013](42250,%20Addgene), which contains next to the BsaI site a tracrRNA sequence.
+We used this plasmid to PCR amplify DNA (primers: T7, 5'-AAAAGCACCGACTCGGTGCC-3') for synthesizing the sgRNA.
+To purify the PCR product, we used the QIAquick PCR Purification Kit (Qiagen).
+We then synthesized sgRNA with the MEGAshortscript Kit (Thermo Fisher Scientific) and purified it with the MEGAclear Kit (Thermo Fisher Scientific).
+To produce the Cas9-mRNA, we used the plasmid (pUC57-T7-RPP2-Cas9) containing the Cas9 ORF fused to 169 base pair 5' UTR from the *Platynereis dumerilii* 60S acidic ribosomal protein P2.
+After in vitro transcription with the mMessage mMachine Kit (Thermo Fisher Scientific), we capped and polyA-tailed the Cas9-mRNA with the Poly(A) Tailing Kit (Thermo Fisher Scientific).
+We coinjected the sgRNA (18 ng/ml) and the Cas9-mRNA (180 ng/µl) into fertilized eggs of *Platynereis dumerilii* wild-type parents according to an established injection procedure [@Conzelmann2013a].
+Eggs were kept at 18°C for 45 min before injection and were injected at 14.5°C.
+The injected individuals were kept at 18°C for 5 to 8 days in 6-well-plates (Nunc multidish no. 150239, Thermo Scientific) and then cultured at 22°C until sexual maturity.
 The mature worms were crossed to wild-type worms and the progeny was genotyped, resulting in two founder lines, which were bred to homozygosity.
 
-**NOS sequencing and genotyping**
+**Genotyping of NOS alleles**
 
-For genotyping of the NOS locus, genomic DNA was isolated from single larvae, groups of 6-20 larvae, or from the tails of adult worms.
+For genotyping of the *NOS* locus, genomic DNA was isolated from single larvae, groups of 6-20 larvae, or from the tails of adult worms.
 The DNA was amplified by PCR (primers: 5'-GGTTCATTGGTTTCGATAACATTGCGG-3', 5'-CAGAGTCGATCAGTCTGCATATCTCCA-3') with the dilution protocol of the Phusion Human Specimen Direct PCR Kit (Thermo Scientific).
@@
 
-**Vertical column setup for measuring photoresponses**
+**Vertical column setup to measure photoresponses**
 
-Photoresponses of larvae of different ages were assayed in a vertical Plexiglas column (31 mm x 10 mm x 160 mm water height).
-The column was illuminated from top with light from a monochromator (Polychrome II, Till Photonics).
-The monochromator was controlled by AxioVision 4.8.2.0 (Carl Zeiss MicroImaging GmbH) via analog voltage.
+We assayed larval photoresponses in a vertical Plexiglas column (31 mm x 10 mm x 160 mm water height).
+For stimulation, we illuminated the column from the top with a monochromator (Polychrome II, Till Photonics) controlled by AxioVision 4.8.2.0 (Carl Zeiss GmbH, Jena) via analog voltage.
 The light passed a collimator lens (LAG-65.0-53.0-C with MgF2 Coating, CVI Melles Griot) before entering the column.
@@
 One strip contained UV (395 nm) LEDs (SMB1W-395, Roithner Lasertechnik) and the other infrared (810 nm) LEDs (SMB1W-810NR-I, Roithner Lasertechnik).
-The UV LEDs were run at 4 V to stimulate the larvae in the column from the side.
-The infrared LEDs were run at 8 V (overvoltage) to illuminate the larvae for the camera (DMK 22BUC03, The Imaging Source), which recorded videos at 15 frames per second and was controlled by IC Capture (The Imaging Source).
-
-**Comparing behavior of wildtype and NOS-knockout 3-day-old larvae**
+The UV LEDs were run at 4 V. The infrared LEDs were run at 8 V (overvoltage) to illuminate the larvae for imaging with a DMK camera (DMK 22BUC03, The Imaging Source).
+We recorded videos at 15 frames per second with the software IC Capture (The Imaging Source).
 
@@
 
-**NOS Identification and Phylogenetic Analysis**
+**Identification and Phylogenetic Analysis of NOS sequences**
 
-To identify NOS, we obtained a "seed" database of oxygenase domain in Pfam database, PF02898.
-From these sequences, we produced a Hidden Markov Model (HMM) and used this to mine the 47 metazoan species, 2 choanoflagellate species and 2 filasterea species investigated.
+To identify NOS sequences, we obtained a seed database of oxygenase domains from the Pfam database (PF02898).
+From these sequences, we produced a Hidden Markov Model (HMM) and used this to mine the 47 metazoan, 2 choanoflagellate and 2 filasterea species investigated.
 HMM models were run in HMMR3 with an e-value of 1e−15.
@@
 To calculate maximum-likelihood trees, we used IQ-tree2 with the LG+G4 model.
-To calculate branch support, we ran 1,000 replicates with the aLRT-SH-like and aBayes methods [@Minh2020].
+To calculate branch support, we ran 1000 replicates with the aLRT-SH-like and aBayes methods [@Minh2020].
 The sequences used for the phylogenetic analysis are available in **Supplementary file 1**, the trimmed alignment is available in **Supplementary File 2** and the pre-trimmed data in **Supplementary file 3**.
 
-**NIT-GC Identification and Phylogenetic Analysis**
+**Identification and Phylogenetic Analysis of NIT-GC sequences**
 
-To identify NIT-GCs, we obtained a "seed" database of Adenylate and Guanylate cyclase catalytic domain in Pfam database, PF00211.
-From these sequences, we produced a Hidden Markov Model (HMM) and used this to mine the 45 metazoan species, 2 choanoflagellate species and 2 filasterea species investigated.
+To identify NIT-GCs, we obtained a seed database of adenylate and guanylate cyclase catalytic domains from the Pfam database (PF00211).
+From these sequences, we produced a Hidden Markov Model (HMM) and used this to mine the 45 metazoan, 2 choanoflagellate and 2 filasterea species investigated.
 HMM models were run in HMMR3 with an e-value of 1e−15.
 We ran CD-Hit [@Fu2012] to eliminate redundant sequences (at a 80% threshold).
-To identify clusters, we used the convex-clustering option with 100 jack-knife replicates.
-The NIT-GCs are extremely well conserved in membrane-bound guanylate cyclases and form an easily recognizable cluster.
+To identify clusters, we used sequence-similarity-based clustering in Clans [@frickey2004] and the convex-clustering option with 100 jack-knife replicates.
+The NIT-GCs are extremely well conserved within the membrane-bound guanylate cyclases and form an easily recognizable cluster.
 To analyze the phylogeny of NIT-GCs, the cluster containing these GCs together with membrane-bound guanylate cyclases were parsed and used for tree building.
@@
 To calculate maximum-likelihood trees, we used IQ-tree2 with the LG+G4 model.
-To calculate branch support, we ran 1,000 replicates with the aLRT-SH-like and aBayes methods [@Minh2020].
-The sequences used for the phylogenetic analysis are available in **Supplementary file 4**, the trimmed alignment is available in **Supplementary file 5** and the pre-trimmed data in **Supplementary file 6**.
+To calculate branch support, we ran 1000 replicates with the aLRT-SH-like and aBayes methods [@Minh2020].
+The sequences used for the phylogenetic analysis are available in **Supplementary File 4**, the trimmed alignment is available in **Supplementary File 5** and the pre-trimmed data in **Supplementary file 6**.
 
@@
 
-We used Achim et al. for the single-cell data [@Achim2015].
-In Williams et al., they used 107 cells as neurons by removing duplicates from Achim et al. single-cell data, so we used those cells [@williams2017].
-Since the raw data were read count data, we normalized them to TPM using Python.
-After that we converted them to log10.
-From the sum of the expression levels in 107 cells for each gene, We calculated the percentage in each cell.
-For each cell, we identified them with marker genes.
-After created the data in Python, plotted it using R dot plots.
-RPKM calculates the total number of reads per million bp and then divides by the length of each gene, so it is not possible to compare between samples.
-Instead, TPM first divides by the length of the gene and then divides by the total number of reads per million bases, which allows for more accurate comparisons between samples.
-In this case we wanted to compare between samples, so we used TPM.
+We used a single-cell RNA sequencing dataset from *Platynereis* larvae [@Achim2015] that was further filtered to 107 cells [@williams2017].
+We normalized raw read-count data to TPM and converted to log10.
+To calculate % expression, we used the sum of the expression levels in the 107 cells for each gene.
 The total TPM of each gene between the samples was used to calculate the percentage of expressed genes.
+We processed the data in Python and plotted in R.
+Individual cells were identified by suits of well-characterised marker genes and spatial mapping [@Achim2015].
 The total TPM values for each gene and the percentage of expressed genes are available in **Supplementary file 7**.
@@
 
-Whole-mount immunostaining of 2 day old Platynereis larvae fixed with 4% paraformaldehyde were carried out using primary antibodies raised against NIT-GC1, NIT-GC2, NOS, RYamide neuropeptide, RGWamide neuropeptide in rabbit, plus a commercial antibody raised against acetylated tubulin in mouse (Sigma T7451).
-The synthetic peptides contained an N-terminal Cys that was used for coupling during purification.
-Antibodies were affinity purified from sera as previously described [@Conzelmann2012].
+Whole-mount immunostaining of *Platynereis* larvae fixed with 4% paraformaldehyde was carried out using the primary antibodies listed in KEY RESOURCES TABLE.
 Immunostainings were carried out as previously described [@Conzelmann2012].
-The NOS promoter (fragment sizes: 12 Kb) was amplified and cloned upstream of 3xHA- Palmi-tdTomato.
-Larvae injected with promoter constructs (ca.
-250 ng/ml) were analysed for reporter expression at 3 days post fertilization using an AxioImager Z.1 fluorescence wide-field microscope (Carl Zeiss GmbH, Jena) and immediately fixed for immunostainings.
-The protocol followed for immunostaining of HA-tagged reporters was recently described [@veraszto2017].
-Specimens were imaged with a LSM 780 NLO or LSM 880 with Airysan Confocal Microscope (Zeiss, Jena).
+
+**Antibody generation and purification**
+
+Antibodies were raised in rats or rabbits against synthetic peptides containing an N-term Cys residue (Altabioscience).
+The same Cys-containing synthetic peptides were used for affinity purification.
+Antibodies were affinity purified from sera on a SulfoLink Coupling Resin (Thermo Fischer) as previously described [@Conzelmann2012].
+View a detailed protocol here: <https://bio-protocol.org/exchange/preprintdetail?id=2333&type=3>
+
+**Transient transgenesis**
+
+A 12 kb genomic region upstream of the start site of the *NOS* gene was amplified and cloned upstream of 3xHA-Palmi-tdTomato to yield the plasmid *NOS-3xHA-Palmi-tdTomato*.
+Larvae injected with the promoter construct (ca.
+250 ng/ml) were analysed for reporter expression at three days post fertilization on an AxioImager Z.1 fluorescence microscope (Carl Zeiss GmbH, Jena) and if positive fixed for immunostaining.
+The immunostaining for the HA-tagged reporter was done as described [@veraszto2017].
+Stained specimens were imaged on an LSM 780 NLO or LSM 880 Airysan Confocal Microscope (Carl Zeiss GmbH, Jena).
 
@@
 
-For calcium imaging, 49--55 hpf larvae were used.
-Experiments were performed at room temperature and larvae were immobilised by being embedded in 2.5% agarose filtered artificial seawater between a slide and coverslip spaced with adhesive tape.
+For calcium imaging, 49-55 hours post fertilisation larvae were used.
+Experiments were carried out at room temperature and larvae were immobilised in 2.5% agarose in artificial sea water and mounted between a slide and coverslip spaced with adhesive tape.
 GCaMP6s mRNA (1 mg/ml) was injected into zygotes as described previously [@Randel2014a].
 Larvae were imaged on a Zeiss LSM 880 with Airyscan (with a C-Apochromat 63X/1.2 Corr - water) with a frame rate of 1.88 frame/sec and an image size of 512 x 512 pixels.
-The larvae were stimulated in a region of interest (a circle with 50 pixel diameter) with 405 nm lasers controlled by the Bleaching mode.
+The larvae were stimulated in a region of interest (a circle with 50 pixel diameter) with a 405 nm laser controlled by the Bleaching mode of the Zeiss software.
 The imaging laser had a similar intensity than the stimulus laser but covered an area that was 10 times larger than the stimulus ROI.
 
-**Cell culture experiment**
+**cGMP production assay**
 
-Green cGull was used for the cGMP assay [@Matsuda2017].
-A full-length Pdum-NIT-GC1 and -NIT-GC2 coding sequences were amplified by PCR starting from a *Platynereis dumerilii* cDNA library and cloned into the pcDNA3.1(+) vector using the T2A self-cleaving sequence.
-Cos-7 cells with low expression of endogenous soluble guanylate cyclase were used as cultured cells for gene expression.
-This cell line was purchased from Angio-proteomie (CAT no. cAP-0203).
-The Cos-7 cells were maintained at 37 °C in 35mm dishes (Nunc™ Glass Bottom Dishes) containing 3 mL of DMEM, high glucose glutamax medium (Thermo; Cat.
+To measure cGMP production in a cell culture assay we used the Green cGull reporter [@Matsuda2017].
+Full-length *NIT-GC1* and *NIT-GC2* coding sequences were amplified by PCR from a *Platynereis dumerilii* cDNA library and cloned into the pcDNA3.1(+) vector with the T2A self-cleaving sequence.
+For the assay, COS-7 cells (Angio-proteomie, CAT no. cAP-0203) with a low endogenous level of soluble guanylate cyclase activity were used.
+Cells were maintained at 37 °C in 35 mm dishes (Nunc™ Glass Bottom Dishes) containing 3 ml of DMEM, high glucose glutamax medium (Thermo; Cat.
 No.
@@
 10082147).
-Upon reaching confluency of approximately 85%, we transfected the cells with the plasmid containing Green cGull-T2A-NITGC1.
-Transfections were carried out with 150 ng of each plasmid and 0.3 μl of the transfection Lipofectamine 3000 Reagent (invitrogen; Cat. No. L3000001).
-Two days post-transfection, we removed the culture medium and substituted it for fresh DMEM-medium.
-For single-wavelength imaging experiments, cells in 35-mm dishes were washed twice and imaged in modified Ringer's buffer (140 mM NaCl, 3.5 mM KCl, 0.5mM NaH2PO4 , 0.5mM S-3 MgSO4 , 1.5 mM CaCl2 , 10 mM HEPES, 2 mM NaHCO3 and 5 mM glucose).
-Dishes were mounted on a stage heated at 37 °C and imaging was performed using an inverted microscope (LSM880, Zeiss) equipped with an oil-immersion objective lens (UApo/340, 40, NA = 0.17).
-Images were acquired using a xenon lamp, 460--495 nm excitation filter, 505-nm dichroic mirror and 510-- 550-nm emission filter (Zeiss).
-S-Nitroso-N-acetyl-D, L-penicillamine (SNAP) was purchased from Sigma-Aldrich (St. Louis, MO, USA) .
-The exposure time of the EM-CCD camera was controlled by the ZEN software (Zeiss).
+Upon reaching approximately 85% confluency, cells were transfected with the pcDNA3.1(+)-Green-cGull-T2A-NITGC1 plasmid.
+Transfections were carried out with 150 ng of each plasmid and 0.3 μl of Lipofectamine 3000 Reagent (invitrogen; Cat. No. L3000001).
+Two days post-transfection, the culture medium was substituted for fresh DMEM-medium.
+For single-wavelength imaging experiments, cells in 35-mm dishes were washed twice and imaged in modified Ringer's buffer (140 mM NaCl, 3.5 mM KCl, 0.5 mM NaH~2~PO~4~ , 0.5 mM S-3 MgSO~4~, 1.5 mM CaCl~2~ , 10 mM HEPES, 2 mM NaHCO~3~ and 5 mM glucose).
+Dishes were mounted on a stage heated to 37 °C and imaging was done on an inverted confocal microscope (LSM880, Carl Zeiss GmbH, Jena) equipped with an oil-immersion objective lens (UApo/340, 40x, NA = 0.17).
+The exposure time of the EM-CCD camera was controlled by the ZEN software (Carl Zeiss GmbH, Jena).
 Images were acquired every 15 s for 10 min and stimulation was initiated 2 min after starting image acquisition.
-Imaging data analysis was performed using ImageJ (National Institutes of Health, Bethesda, MD, USA).
+Imaging data analysis was done with ImageJ (National Institutes of Health, Bethesda, MD, USA).
+As NO donor, we used S-Nitroso-N-acetyl-D, L-penicillamine (SNAP, Sigma-Aldrich, St. Louis, MO, USA).
 
@@
 
+### Wild type model
+
 The mathematical model for the wild type data comprises six dynamical equations:
 
-```{=tex}
 \begin{align*}
@@
 \end{align*}
-```
+
+The $C_P(t)$ equation describes the dynamics of the Ca^2+^ concentration in the cPRC cells.
+The $K_{UV}^{C_P} UV(t) C_O(t)$ term captures the Ca^2+^ increase due to c-opsin1 activity induced by UV stimulation, while the $K_{G}^{C_P} G(t)$ term captures dependence of the Ca^2+^ dynamics on cGMP.
+The $C_O(t)$ equation is added to capture the decrease of the Ca^2+^ levels observed during the UV stimulation; the mechanism of this decrease is unknown.
+It could, for example, be related to the opening of the GIRK channels.
+The $G(t)$ equation describes the dynamics of cGMP in the cPRC cells.
+The $K_{GC1}^G N(t)$ term describes the NO-dependent increase in cGMP by NIT-GC1, while the $K_{GC2}^G \left(1 - UV(t)\right) G(t)$ term captures the cGMP decrease mediated by NIT-GC2 during UV stimulation.
+The $S(t)$ describes synaptic coupling from the cPRC cells, which is assumed to act in an excitatory fashion on INNOS cells and an inhibitory fashion on INRGW cells.
+This coupling is assumed to be suppressed in NIT-GC2 morphants.
+The $C_N(t)$ equation describes the Ca^2+^ dynamics in the INNOS cells.
+The $K_{S}^{C_N} S(t)$ term describes the Ca^2+^ increase due to the synaptic signal $S(t)$.
+The $N(t)$ variable represents NO produced and secreted by the INNOS cells.
+The $K_{C_N}^{N} C_N(t)$ term describes the Ca^2+^-dependent NO production within the INNOS cells.
+The $C_R(t)$ equation describes the dynamics of Ca^2+^ in the INRGW cells, which depends on the other equations in a feed-forward manner.
+The $K_{S}^{C_R} S(t)$ and $K_{C_N}^{C_R} C_N(t) C_R(t)$ terms describe the decrease in Ca^2+^ with respect to the synaptic signal $S(t)$ and the Ca^2+^ levels in the INNOS cells, respectively.
+The $K_{C_P}^{C_R} C_P(t)$ term describes the increase in Ca^2+^ levels in the INRGW cells due to increases in the Ca^2+^ levels in the cPRC cells.
+Note that all sources of Ca^2+^ variation in the INRGW cells are phenomenological in nature.
+All variables have a constant (unit) production term and a linear decay rate set by the parameter $\delta_X$ where $X \in \{ C_P, C_O, G, S, C_N, C_R\}$.
+Finally, the UV stimulation is described by a square pulse with amplitude $a$ and duration $t_{UV end} - t_{UV \: start}$.
+
+### NOS-knockout model
+
+The NOS-knockout model is derived from the WT model by eliminating the $N(t)$ variable, which represents the NO produced by the INNOS cells in a NOS-dependent manner, and terms that depend on it.
+This gives the system \begin{align*}
+\frac{C_P(t)}{dt} &= 1 - \delta^{C_P} C_P(t) + K_{UV}^{C_P} UV(t) C_O(t) + K_{G}^{C_P} G(t) , \\
+\frac{C_O(t)}{dt} &= 1 - \delta^{C_O} C_O(t) - K_{UV}^{C_O} UV(t) C_O(t), \\
+\frac{G(t)}{dt} &= 1 - \delta^{G} G(t)+ \left(1 - UV(t)\right)K_{GC2}^G G(t), \\
+\frac{S(t)}{dt} &= UV(t) - \frac{S(t)}{\tau_S}, \\
+\frac{C_N(t)}{dt} &= 1 - \delta^{C_N} C_N(t) + K_{S}^{C_N} S(t), \\
+\frac{C_R(t)}{dt} &= 1 - \delta^{C_R} C_R(t) - K_{S}^{C_R} S(t) + K_{C_P}^{C_R} C_P(t) - K_{C_N}^{C_R} C_N(t) C_R(t), \\
+UV(t) &=
+    \begin{cases}
+        a, & t_{UV\: start} \leq t \leq t_{UV\: end}\\
+        0, & \text{otherwise}
+    \end{cases}   .
+\end{align*}
+
+We remark that the above system of equations can be used to also describe the NIT-GC1 morpholino condition (Fig. 4M).
+
+### NIT-GC2 morpholino model
+
+The NIT-GC2 morpholino model is a further restriction of the WT model:
+
+\begin{align*}
+\frac{C_P(t)}{dt} &= 1 - \delta^{C_p} C_P(t) + K_{UV}^{C_P} UV(t) C_O(t) , \\
+\frac{C_O(t)}{dt} &= 1 - \delta^{C_O} C_O(t) - K_{UV}^{C_O} UV(t) C_O(t), \\
+\frac{C_R(t)}{dt} &= 1 - \delta^{C_R} C_R(t) + K_{C_P}^{C_R} C_P(t), \\
+UV(t) &=
+    \begin{cases}
+        a, & t_{UV\: start} \leq t \leq t_{UV\: end}\\
+        0, & \text{otherwise}
+    \end{cases} .
+\end{align*}
+
+This reduced model eliminates the $G(t)$, $S(t)$, $C_N(t)$ and $N(t)$ variables and hence includes only c-opsin1 activity in the cPRC cells and the feedforward coupling to INRGW cells.
+The synaptic signal $S(t)$ and all terms depending on it are removed due to the assumed dependence on NIT-GC2.
+When the $S(t)$ input is disregarded, the activity of the INNOS cells ($C_N(t)$ and $N(t)$ variables) can be represented by a constant steady state whose values can subsequently be absorbed into all of the terms depending on these variables.
+Hence these variables are eliminated in the reduced model.
+Since the $K_{GC2}^G \left(1 - UV(t)\right) G(t)$ term depends directly on NIT-GC2 and when $N(t)$ is ignored the $K_{GC1}^G N(t)$ term disappears, the $G(t)$ variable can be eliminated, which consequently simplifies the $C_P(t)$ equation.
+The only remaining term in $C_R(t)$ equation, $K_{C_P}^{C_R} C_P(t)$, describes feedforward coupling between the cPRC and INRGW cells.
+
+### Parameter fitting (genetic algorithm)
+
+To find sets of model parameters that reproduce recordings from the Ca^2+^ imaging experiments, we establish an optimisation problem to minimise the discrepancy between the modelled Ca^2+^ trajectories and the experimental data.
+To obtain the modelled Ca^2+^ trajectories, we simulated the model using the Matlab built-in solvers \texttt{fsolve} to find the steady state in the absence of UV stimulation, and then use the explicit integration algorithm \texttt{ode45} to simulate the model using the computed steady state as the initial condition.
+
+We quantify the difference between the model output and the data using a vector-valued objective function.
+The first component of this function measures this difference during the UV stimulation, whilst the second component measures the difference from the end of the stimulation until the end of the recording.
+We thus define a vector-valued objective function, $\mathbf{D} \in \mathbb{R}^2_{\geq 0}$, as: \begin{equation*}
+ \mathbf{D}\left({C}_X^{data},{C}_X^{model}\right)=\left(D_{k_{UV\: start}}^{k_{UV\: end}}\left({C}_X^{data},{C}_X^{model}\right),D_{k_{UV\: end }+1}^{k_{end}}\left({C}_X^{data},{C}_X^{model}\right) \right),
+\end{equation*} with
+
+\begin{equation*}
+ D_{k_1}^{k_2}\left({C}_X^{data}(t),{C}_X^{model}(t)\right)=\sqrt{\sum_{i=k_1}^{k_2}\left({C}_{X}^{data}(t_i)-{C}_{X}^{model}(t_i)\right)^2} ,   
+\end{equation*} where the timepoints $t_i = i \Delta t$ are equispaced with $\Delta = 0.5s$ and where $t_{k_{UV\: start}} = t_{UV \: start}$ (and similar for the other time points).
+Here ${C}_X^{data}$ denotes normalised experimental data and ${C}_X^{model}$ denotes normalised simulated Ca^2+^ trajectory with $X \in \{ P, N, R\}$ indicating cell type.
+Simulated Ca^2+^ trajectories, ${C}_X^{model}$, are normalised in the same way as the experimental data, i.e., we divide values uniformly by their value at $t_{k_{UV\: start}-1}$, a time point immediately prior to the start of the UV stimulation.
+Simulations are sampled with the same sampling rate as the data.
+(Recall that $t_{UV\: start}$ is the time of the start of UV stimulation, $t_{UV\: end}$ is the time of the end of UV stimulation, and $t_{end}$ is the last time sample.)
+
+To solve the optimisation problem with the above objective function, we employ a global optimisation method known as a genetic algorithm (GA).
+GAs are a class of iterative, gradient-free algorithms inspired by processes of natural selection including breeding and mutation, amongst others \cite{GAbook}.
+In reference to natural selection, sets of candidate solutions in GAs are referred to as \`populations'.
+As our algorithm of choice, we use the Matlab implementation of a genetic algorithm for multiobjective optimisation, \texttt{gamultiobj}.
+The \texttt{gamultiobj} algorithm uses a controlled, elitist genetic algorithm (a variant of NSGA-II \cite{GAbook}).
+We use the default settings for most hyperparameters of the algorithm.
+
+Here we outline the main steps of the \texttt{gamultiobj} implementation of the GA.
+The first step is to generate a set of random model parameter sets, known as the initial population.
+Next, we score the fitness of each parameter set in the population according to the objective function.
+Using these scores, the \texttt{gamultiobj} algorithm employs a binary tournament selection function\cite{Miller1995GeneticAT} to select 'parents' for the next population.
+A population of children (i.e., new parameter sets) is generated by 'breeding' randomly chosen pairs of parents (exchanging subsets of parameter values between them).
+Parameter values in these children can also be changed by 'mutations', which perturb said values randomly by a small amount.
+Breeding and mutation introduce a level of randomness to the optimisation process that allows for exploration of the objective function (fitting) landscape.
+The stochastic nature of GAs helps them to avoid local minima that can plague gradient-descent-based optimisation methods.
+
+In the next step, the \texttt{gamultiobj} algorithm scores the parameter sets in the children population using the values of the objective function.
+Scores from the parent and child populations are combined.
+A subset of the size of the initial population is then selected as a new parent population based on the scores and the \textit{Crowding Distance}, distance to other parameter sets in terms of a sum of element-wise differences between values of the objective function (used to ensure diversity of the parent population)\cite{crowding_distance,GAbook}.
+This iterative process is repeated until set convergence criteria are met, indicating that the GA has found a set of model parameters that best reproduce the data under consideration.
+
+The \texttt{gamultiobj} algorithm returns a set of model parameter values on a Pareto front \cite{censor1977pareto,GAbook}.
+The coordinates of points on the front are the values of the components of the objective function $\mathbf{D}$.
+Points on the Pareto front are nondominated, meaning that they are closest to 0 for at least one of the two components.
+To identify a single optimal solution, we select the parameter set along the Pareto front whose objective function value is the smallest with respect to the Euclidean norm.
+
+Below we present a summary of the steps of our optimisation pipeline for fitting the experimental recordings from the cPRC cells.
+
+* When fitting to the cPRC data using WT or NIT-GC2 morpholino models, we eliminate the $C_R(t)$ variable since it does not have any feedback coupling with the $C_P(t)$ equation and does not affect its dynamics.
+When fitting to the cPRC data using the NOS-knockout model, we also eliminate the $S(t)$, $C_N(t)$ and $C_R(t)$ variables since these do not have any effect on the $C_P(t)$.
+*  Set the size of the initial population to 1,000. This value provides a compromise in terms of the coverage of the parameter space and run-time of the GA.
+* Set bounds of random distributions of parameters.
+We draw all the initial parameter values from uniform distributions over the unit interval ([0,1]).
+For WT condition, we fit 13 parameters, for NOS-knockout, we fit 7 parameters and for NIT-GC2 morpholino, we fit 4 parameters.
+* Evaluate the objective function on the initial population.
+* Run the \texttt{gamultiobj} algorithm.
+* Repeat steps 1 to 4 with bounds of parameters extended to [0,5] and using the Pareto front returned in step 4 as a part of the initial population.
+This step was added after observing that, although good fits to most datasets are achieved with parameter values in the interval [0,1], fits to some datasets can be improved by extending the admissible parameter range.
+* Use the Euclidean norm to find the set of model parameters that best fits the data.
+
+Since INNOS cells in the NOS-knockout condition and INRGW cells in all conditions are coupled in a feedforward manner and do not affect the 'up-stream' dynamics, we run a separate optimisation to fit data from these cells.
+This minimises the number of parameters that are being fitted simultaneously and ensures that the objective function is as simple as possible (using only recordings from a cell type that is being fitted).
+First, we fix the parameters that were found by fitting the model to the cPRC data.
+Next, we run the optimisation in the same way as in the case of the cPRC data but only for the parameters that pertain to dynamics that is strictly`downstream' of cPRC dynamics.
+
+Fixing a subset of parameters allows us to ensure that model parameters inferred from recordings of one cell type stay relevant when fitting the model to the data from other cells.
+This means that the fitted model is extended to include other cell types, rather than being completely reparamatrised for each cell type.
+This is particularly important since, for technical reasons, we do not have simultaneous recordings from multiple cell types.
+
+We note that in the WT condition, model parameters of the INNOS cells are fitted without using any INNOS recordings.
+Since the NO production feeds back to the cPRC cells, their parameters are fitted to reproduce the recordings from the cPRC cells.
+Furthermore, in the NIT-GC2 morpholino condition, we do not fit the INNOS data since we assumed that the feedforward coupling between the cPRC cells and INNOS cell depends on NIT-GC2 activity (that is, we assume that the Ca^2+^ levels in INNOS cells are constant in this condition).
+
+All the simulated curves presented in Fig.
+6B are the averages of the curves fitted to individual experimental recordings/datasets.
+
+### Predictions from a range of stimulation protocols
+
+To explore possible effects of varying the intensity and duration of the UV stimulation on experimentally observed delayed activation in the cPRC cells, we perform a series of simulations in which we systematically varied amplitude ($a \in [0.1,1]$) and width ($t_{UV\: end} \in [0.1,51]$) of the $UV(t)$ variable in the WT model.
+To quantify the effects of the stimulation, we measure the area under the peak of the delayed activation $AUP$ in the $C_P(t)$ variable.
+To compute the $AUP$:
+
+* We find an inflection point that occurs after the start ($t_{UV\: start}$) of the UV stimulation.
+To this end, we identify the first time point $t_{AUP\: start}>t_{UV\: start}$ for which $d^2C_P(t)/dt^2=0$ and $d^2C_P(t_{k_{AUP\: start}}-1)/dt^2<d^2C_P(t_{k_{AUP\: start}})/dt^2<d^2C_P(t_{k_{AUP\: start}}+1)/dt^2$.
+We assume that the inflection is caused by the input from the INNOS cells.
+* We then compute the $AUP$ as an integral over $C_P(t)$ starting at the inflection point and ending 125 seconds later: $AUP=\int_{t_{start}}^{t_{end}} C_P(t) dt $.
+To compute this, we use numerical integration via the trapezoidal method as implemented in Matlab function \texttt{trapz}.
+We remark that the baseline value of the $AUP$ for a Ca^2+^ profile of constant, unit value is 125.
+* To capture possible variability between different cells/organisms, we compute the $AUP$ using 47 model parameter sets fitted to individual WT cPRC experimental recordings.
+We remove outliers, defined as being any simulated trajectories for which $C_P(t)$, after normalisation, exits the interval [0.25, 4] at any time.
+This $C_P(t)$ range was decided of the basis of values observed in various experiments in the lab.
+* For analysis, we use the median value of the $AUP$ at each combination of amplitude and duration.
+
+Figure 6 shows how the results of the simulations of the WT model depend on the amplitude and duration of UV stimulation, as captured by the median values of $AUP$.
+Delayed activation occurs only for UV stimulation of sufficient amplitude and duration.
+Specifically, we observe sudden increase in AUP for UV pulse amplitude $a>0.5$ and a linear dependence of AUP on the pulse duration.
+
 ## Data analysis and figure preparation
@@
 All scripts are available at <https://github.com/JekelyLab/Joukra_et_al_NOS> (commit ).
->>>>>>> 9932f067da10604ec00c679a5b10f91decddfa29
 
@@
 
-```{r echo=FALSE,fig.cap="**Figure 1 -- figure supplement 1**  Phylogenetic tree of NOS by maximum likelihood.
-Tree robustness was tested with 1000 replicates of ultrafast bootstrap with the aLRT-SH-like and aBayes methods."}
+```{r echo=FALSE,fig.cap="**Figure 1 -- figure supplement 1.
+Maximum-likelihood phylogenetic tree of NOS protein sequences.** Individual branches are coloured by taxonomy.
+Support values are based on 1000 replicates of ultrafast bootstrap with the aLRT-SH-like and aBayes methods."}
 knitr::include_graphics(path = "figures/Fig1_sup1.png")
@@
 
-```{r echo=FALSE,fig.cap="**Figure 1 -- figure supplement 2** Expression analysis of the *NOS* gene (magenta) using in situ HCR using the larva at three-day-old in dorsal (**A**) and anterior view (B).
-(B) Insert: showing two NOS cells (INNOS_dl and INNOS_vl) on left side."}
+```{r echo=FALSE,fig.cap="**Figure 1 -- figure supplement 2.
+Expression of *NOS* in three-day-old *Platynereis* larva.** (**A**) HCR in situ for *NOS* (magenta), dorsal view.
+(**B**) HCR in situ for *NOS* (magenta), ventral view.
+Inset in (B) shows close-up of two *NOS*-positive cells (INNOS_dl and INNOS_vl) on the left side.
+Samples were also stained for DAPI (cyan) to label nuclei."}
 knitr::include_graphics(path = "figures/Fig1_sup2.png")
@@
 
-```{r echo=FALSE,fig.cap="**Figure 1 -- figure supplement 3** ssTEM reconstruction of the INNOS (A), INRGW (B) and cPRC (C), anterior view. NS plexus; neurosecretory plexus"}
+```{r echo=FALSE,fig.cap="**Figure 1 -- figure supplement 3.
+Synapse distribution in cPRCs and possynaptic interneurons** (**A**) Volume rendering of the four INNOS neurons with incoming and outgoing synapses.
+(**B**) Axo-dendritic splitting of the INNOS neurons.
+(**C**) Volume rendering of the four INRGW neurons with incoming and outgoing synapses.
+(**D**) Volume rendering of the four cPRC neurons with incoming and outgoing synapses.
+All images are anterior views.
+Asterisk marks the same position across images for reference.
+Grey speheres indicate the position of cell nuclei.
+NS plexus; neurosecretory plexus.
+The radius of INNOS nuclei is set to 2 µm for scale."}
 knitr::include_graphics(path = "figures/Fig1_sup3.png")
@@
 
-```{r echo=FALSE,fig.cap="**Figure 3 -- figure supplement 1** (**A**) Top: The domain and Exon/Intron structure of NOS.
-Bottom: Close-up region showing the genomic locus of NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9.
-The generated mutants (NOSΔ11, NOSΔ23) are also shown.
-Pink indicates target sites.
-Gray shows PAM sequences, red shows stop codons.
-(**B**) Overlaid trajectories for WT (n=37) and NOS mutant (NOSΔ11/Δ11, n=18 and NOSΔ23/Δ23, n=8) at two-day-old larvae.
-0 sec as the starting point.
-After 10 sec, UV (395 nm) stimulation from the side.
-(**C**) The temporal changes in the vertical position of the WT and mutant two-days-old larvae before and after UV stimulation are shown.
-The starting points of each larval trajectory are set to 0.
-After UV stimulation is indicated by purple squares.
-**(D, E)** The temporal changes in the distance traveled of the WT and mutant in two-day (D) and three-day-old (E) larvae before and after UV stimulation are shown.
-(**F**) The temporal changes in the distance traveled of larvae treated with NOS inhibitors, L-NAME in three-day larvae before and after UV stimulation are shown.
-(**G**) Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23) larvae at two-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
-The data are shown in 30 s bins."}
+```{r echo=FALSE,fig.cap="**Figure 3 -- figure supplement 1.
+Generation and behavioural characterisation of *NOS* CRISPR knockouts.** (**A**) Top: The domain organisation of *Platynereis* NOS protein and exon/intron structure of the *NOS* gene.
+Bottom: The genomic locus of *NOS* with the CRISPR target site, wild-type and knockout (NOSΔ11, NOSΔ23) sequences and predicted protein sequences.
+The PAM sequence is shown in grey, stop codons in red.
+(**B**) Swimming trajectories of wild type (WT, n=37) and *NOS* mutant (NOSΔ11/Δ11, n=18 and NOSΔ23/Δ23, n=8) two-day-old larvae.
+All trajectories start at 0 x and y position and time 0 corresponding to 10 sec after the onset of 395 nm stimulation from the side.
+(**C**)  Vertical displacement in 30 sec bins of wild type and mutant (NOSΔ11 and NOSΔ23) two-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
+(**D**) Vertical position of batches of wild type and mutant two-day-old larvae over time under 395 nm UV stimulation.
+The starting position of each larval trajectory was set to 0.
+(**E-G**) Swimming speed of batches of wild type and mutant two-day old (E) and three-day-old (F) larvae and L-NAME-treated (NOS inhibitor) three-day-old larvae under 395 nm UV stimulation."}
 knitr::include_graphics(path = "figures/Fig3_sup1.png")
@@
 
-```{r echo=FALSE,fig.cap="**Figure 4 -- figure supplement 1** Cluster analysis of guanylate and adenylate cyclases.
-Connections are based on blast similarities < 1e-16 as shown on the upper right.
-Animal groups are colour coded as shown on the upper left.
-NIT-GCs, NIT domain containing guanylate cyclases; membrane-bound GCs, Membrane-bound guanylyl cyclases; sGCs, soluble guanylate cyclases; ACs, adenylate cyclases."}
+```{r echo=FALSE,fig.cap="**Figure 4 -- figure supplement 1.
+Cluster analysis of guanylate and adenylate cyclase sequences.** Each node represents one sequence, colour-coded by taxonomy.
+Connections represent BLAST P-values of <1e-16.
+NIT-GCs, NIT domain containing guanylate cyclases; membrane-bound GCs, membrane-bound guanylate cyclases; sGCs, soluble guanylate cyclases; ACs, adenylate cyclases."}
 knitr::include_graphics(path = "figures/Fig4_sup1.png")
@@
 
-```{r echo=FALSE,fig.cap="**Figure 4 -- figure supplement 2** Phylogenetic tree of guanylate cyclase by maximum likelihood (ML).
-Guanylate cyclase-coupled receptor and soluble guanylate cyclases (sGC) as outgroups.
+Maximum-likelihood phylogenetic tree of NOS protein sequences.** Individual branches are coloured by taxonomy.
+Support values are based on 1000 replicates of ultrafast bootstrap with the aLRT-SH-like and aBayes methods.
+
+```{r echo=FALSE,fig.cap="**Figure 4 -- figure supplement 2.
+Maximum-likelihood phylogenetic tree of NIT-domain-containing guanylate cyclases** Membrane-bound and soluble guanylate cyclases (sGC) were included as outgroups.
 Guanylate cyclases with NIT domains are found in most animal phyla except Porifera, Ctenophora, Urochordata and Chordata.
-Dot plot of Platynereis NIT-GC genes (columns) expressed in cPRC, INNOS and INRGW (rows) using single cell RNA-Seq.
-The size of the dots is expressed in proportion to the percentage of cells expressing that gene relative to all cells.
-The colours represent the normal logarithm of the number of transcripts in the cells expressing the gene."}
+The expression of *Platynereis* NIT-GC genes in cPRC, INNOS and INRGW cells is indicated on the right side of the tree.
+The values represent expression in single-cell sequencing data.
+Dot size indicates specificity (percent of transcripts expressed in the indicated cell across all cells).
+Dot colour represents the logarithm of the number of transcripts in the expressing cells."}
 knitr::include_graphics(path = "figures/Fig4_sup2.png")
@@
 
-```{r echo=FALSE,fig.cap="**Figure 4 -- figure supplement 3** (**A**) Co-expression analysis image of the NIT-GC1 (magenta) and MLD-pedal2 amide proneuropeptide gene (MLD: green).
-Anterior view of the larva at two-day-old.
-**(B, C)** Expression analysis of the NIT-GC2 gene (magenta) using in situ HCR.
-Anterior (B) and posterior (C) views of the larva at three-day-old.
-(**D**) Co-localisation analysis using NIT-GC1 (magenta) and NOS (green) antibodies.
-Anterior view of the larva at two-day-old.
-**(E, F)** Localisation analysis using NIT-GC1 and NIT-GC2 antibodies for NIT-GC1 (E) and NIT-GC2 (F) morphant.
-Green shows co-staining with acetylated α-tubulin antibody (acTub).
-Anterior view of the larva at two-day-old."}
+```{r echo=FALSE,fig.cap="**Figure 4 -- figure supplement 3.
+Expression of NIT-GC1 and NIT-GC2 in the cPRCs** (**A**) HCR co-expression analysis of *NIT-GC1* (magenta) and *MLD/pedal-peptide-2 proneuropeptide* (green).
+Anterior view of a two-day-old larva.
+Nuclei were stained with DAPI (cyan).
+**(B, C)** Expression of *NIT-GC2* (magenta) detected by in situ HCR.
+Anterior (B) and ventral (C) views of a three-day-old larva.
+Nuclei were stained with DAPI (cyan).
+(**D**) Immunostaining for NIT-GC1 (magenta) and NOS (green).
+Anterior view of a two-day-old larva.
+Acetylated α-tubulin staining (blue) highlights the neuropil and cPRC cilia.
+**(E, F)** Immunostaining for NIT-GC1 (E) and NIT-GC2 (F) antibodies in larvae injected with NIT-GC1 (E) or NIT-GC2 (F) morpholinos.
+Acetylated α-tubulin staining (green) highlights the neuropil and cPRC cilia.
+Anterior view of a two-day-old larva."}
 knitr::include_graphics(path = "figures/Fig4_sup3.png")
@@
 
-```{r echo=FALSE,fig.cap="**Figure 5 -- figure supplement 1** (**A**) Schematic diagram of the immunostaining procedure after calcium imaging.**(B-D)** Co-expression analysis image of the NOS (magenta) and RY amide proneuropeptide gene (RYa: green).
-Anterior view of the larva at two-day-old.
+```{r echo=FALSE,fig.cap="**Figure 5 -- figure supplement 1.
+On-slide immunostaining and coexpression analysis of *NOS* and *RYamide proneuropeptide* ** (**A**) Schematic diagram of the on-slide immunostaining procedure after calcium imaging.(**B-D**) Co-expression analysis by HCR in situ of *NOS* (magenta) and *RYamide proneuropeptide* (green).
+Anterior view of a two-day-old larva.
+Nuceli were stained by DAPI (cyan).
 "}
@@
 ```
+
+```{r echo=FALSE,fig.cap="**Figure 6 -- figure supplement 1.
+Simulated Ca^2+^ trajectories for parameter sets fitted to individual Ca^2+^-recordings collected in wild type, *NOS* knockout, and *NIT-GC2* morphant larvae.** (**A-C**) Simulated Ca^2+^ trajectories in cPRC (A), INNOS (B) and INRGW (C) cells in the wild-type condition.
+(**D-F**) Simulated Ca^2+^ trajectories in cPRC (D), INNOS (E) and INRGW (F) cells in the *NOS*-knockout condition.
+(**G-I**) Simulated Ca^2+^ trajectories in cPRC (G), INNOS (H) and INRGW (I) cells in the *NIT-GC2*-morphant condition.
+Thin coloured curves indicate individual recordings (cPRC - purple, INNOS - blue and INRGW - green), thin grey curves indicate simulated Ca^2+^ trajectories based on model parameters fitted to the individual recording, thick coloured curves indicate averages of the recordings and thick black curves represent the average of the fits.
+"}
+knitr::include_graphics(path = "figures/Figure6_fig_Suppl1.svg")
+
+```
+
+```{r echo=FALSE,fig.cap="**Figure 6 -- figure supplement 2.
+Comparison of fits with different fixed parameter sets when fitting the WT model to individual recordings from the WT INRGW cells.** Panels A-B, D-E, G-H  (column fixed parameters) show simulated trajectories of the Ca^2+^ levels in cPRC and INNOS cells produced by different fixed parameter sets.
+Panels C, F and I show individual recordings fitted to the data from INRGW cells."}
+knitr::include_graphics(path = "figures/Figure6_fig_Suppl2.svg")
+```
+
+```{r echo=FALSE,fig.cap="**Figure 6 -- figure supplement 3.
+Comparison of fits to the cPRC recordings from NIT-GC2 morphant larvae using the NIT-GC2 morpholino model or the NOS-knockout model.**  Thin coloured curves indicate individual recordings, thin grey curves are simulated Ca^2+^ trajectories, thick coloured curves indicate averages of the recordings and thick black curves represent the mean of the fits."}
+knitr::include_graphics(path = "figures/Figure6_fig_Suppl3.svg")
+```
+
+```{r echo=FALSE,fig.cap="**Figure 6 -- figure supplement 4.
+Distributions of the parameter values fitted to the cPRC recordings.** Violin plots (grey lines) are kernel density estimates of the underlying distributions; computed using the Matlab function *ksdensity* with default settings, i.e., using normal kernel function, plots are trimmed to the observed range of data.
+Markers represent individual parameter values, white markers indicate medians, vertical grey bars indicate interquartile range (Q25 to Q75), and the dashed line indicates the value 1.
+Where violin plots are not shown, these parameters are associated with terms/equations that are excluded from the fitting procedure (see Methods for details)." }
+knitr::include_graphics(path = "figures/Figure6_fig_Suppl4.svg")
+```
+
+```{r echo=FALSE,fig.cap="**Figure 6 -- figure supplement 5.
+Pairwise correlations between parameters of the WT model fitted to WT cPRC recordings** Parameters associated with the $C_R$ equation are not included (see Methods for details).
+Panels on the diagonal show kernel density estimates of the distributions of the individual parameters; computed using Matlab function *ksdensity* with default settings, i.e., using normal kernel function, plots are trimmed to the observed range of data.
+The other panels show kernel density estimates of bivariate distributions of pairs of parameters (heatmaps; darker colours indicate higher density).
+Also shown using markers are the values of the individual parameter combinations."}
+knitr::include_graphics(path = "figures/Figure6_fig_Suppl5.svg")
+```

### 2023-06-24 code/Fig3_and_Fig3_figure_Suppl1.R

@@ -666,16 +666,15 @@ ggsave("Manuscript/figures/Fig3.pdf", limitsize = FALSE,
 
 
 layout2 <- "
-AAAABBBBGGGG
+AAAABBBBCCCC
 ############
-CCCDDDEEEFFF
+DDDEEEFFFGGG
 "
 
 Fig3_sup1 <-  
   panel_Architecture + 
-  panel_2dpf_Tracking +   
+  panel_2dpf_Tracking +   panel_2d_vd +
   panel_2dpf_TP + panel_2dpf_TD + panel_3dpf_TD + panel_L_NAME_TD +
-  panel_2d_vd + 
   patchwork::plot_layout(design = layout2, heights = c(1, 0.05, 0.7)) + #we can change the heights of the rows in our layout (widths also can be defined)
   patchwork::plot_annotation(tag_levels = 'A') &  #we can change this to 'a' for small caps or 'i' or '1'
   ggplot2::theme(plot.tag = element_text(size = 12, face='plain')) #or 'plain', 'italic'

### 2023-06-24 code/Fig4_sup3.R

@@ -19,7 +19,7 @@ panel_NIT1_MLD <- ggdraw() + draw_image(readPNG("pictures/HCR_51_AP_NITGC1_MLD_D
   draw_label("NIT-GC1", x = 0.12, y = 0.91, color="magenta", size = 10, fontface="italic") +
   draw_label("MLD/pedal peptide 2", x = 0.26, y = 0.85, color="green", size = 10, fontface="italic") +
   draw_label("DAPI", x = 0.32, y = 0.91, color="cyan", size = 10, fontface="plain") +
-  draw_line(x = c(0.04, 0.22), y = c(0.08, 0.08), color = "white", size = 0.5) +
+  draw_line(x = c(0.04, 0.218), y = c(0.08, 0.08), color = "white", size = 0.5) +
   draw_label(expression(paste("20 ", mu, "m")), x = 0.13, y = 0.11, color = "white", size = 8) +
   draw_label("D", x = 0.95, y = 0.93, size = 6, color = "white") +
   draw_label("V", x = 0.95, y = 0.77, size = 6, color = "white") +
@@ -33,13 +33,15 @@ panel_72_AP_NIT2 <- ggdraw() + draw_image(readPNG("pictures/HCR_72_AP_NIT_93.98u
   draw_label("in situ HCR", x = 0.25, y = 0.995, size = 10) +
   draw_label("NIT-GC2", x = 0.12, y = 0.91, color="magenta", size = 10, fontface="italic") +
   draw_label("DAPI", x = 0.32, y = 0.91, color="cyan", size = 10, fontface="plain") +
-  draw_line(x = c(0.04, 0.25), y = c(0.08, 0.08), color = "white", size = 0.5)
+  draw_line(x = c(0.04, 0.252), y = c(0.08, 0.08), color = "white", size = 0.5) +
+  draw_label(expression(paste("20 ", mu, "m")), x = 0.13, y = 0.11, color = "white", size = 8) 
+  
 
 panel_72_DV_NIT2 <- ggdraw() + draw_image(readPNG("pictures/HCR_72_DV_NIT_180.08um.png")) +
   draw_label("in situ HCR", x = 0.25, y = 0.995, size = 10) +
   draw_label("NIT-GC2", x = 0.12, y = 0.91, color="magenta", size = 10, fontface="italic") +
[…]

### 2023-06-24 Manuscript/Manuscript.Rmd
edit methods

@@
 To analyze the phylogeny of NIT-GCs, the cluster containing these GCs together with membrane-bound guanylate cyclases were parsed and used for tree building.
-We aligned the sequences with MAFFT version 7, with the iterative refinement method E-INS-i.
+We aligned the sequences with MAFFT version 7, with the iterative refinement method E-INS-i[@Katoh2013].
 Alignments were trimmed with TrimAl in gappy-out mode [@Capella-Gutiérrez2009].
@@
 For the assay, COS-7 cells (Angio-proteomie, CAT no. cAP-0203) with a low endogenous level of soluble guanylate cyclase activity were used.
-Cells were maintained at 37 °C in 35 mm dishes (Nunc™ Glass Bottom Dishes) containing 3 ml of DMEM, high glucose glutamax medium (Thermo; Cat.
+Cells were maintained at 37 °C in 35 mm dishes (Nunc™ Glass Bottom Dishes) containing 3 ml of DMEM, low-glucose medium (Thermo; Cat.
 No.
-10566016) supplemented with 10% fetal bovine serum (Thermo; Cat.
+11885084) supplemented with 10% fetal bovine serum (Thermo; Cat.
 No.
@@
 Transfections were carried out with 150 ng of each plasmid and 0.3 μl of Lipofectamine 3000 Reagent (invitrogen; Cat. No. L3000001).
-Two days post-transfection, the culture medium was substituted for fresh DMEM-medium.
+Three hours post-transfection, the culture medium was substituted for fresh DMEM-medium.
 For single-wavelength imaging experiments, cells in 35-mm dishes were washed twice and imaged in modified Ringer's buffer (140 mM NaCl, 3.5 mM KCl, 0.5 mM NaH~2~PO~4~ , 0.5 mM S-3 MgSO~4~, 1.5 mM CaCl~2~ , 10 mM HEPES, 2 mM NaHCO~3~ and 5 mM glucose).
@@
 Imaging data analysis was done with ImageJ (National Institutes of Health, Bethesda, MD, USA).
-As NO donor, we used S-Nitroso-N-acetyl-D, L-penicillamine (SNAP, Sigma-Aldrich, St. Louis, MO, USA).
+As NO donor, we used S-Nitroso-N-acetyl-D, L-penicillamine (SNAP, Sigma-Aldrich, St.
 
@@
 \end{align*}
-
+```
 The $C_P(t)$ equation describes the dynamics of the Ca^2+^ concentration in the cPRC cells.
@@
 \end{align*}
-
+```
 This reduced model eliminates the $G(t)$, $S(t)$, $C_N(t)$ and $N(t)$ variables and hence includes only c-opsin1 activity in the cPRC cells and the feedforward coupling to INRGW cells.
@@
 
-* When fitting to the cPRC data using WT or NIT-GC2 morpholino models, we eliminate the $C_R(t)$ variable since it does not have any feedback coupling with the $C_P(t)$ equation and does not affect its dynamics.
+-   When fitting to the cPRC data using WT or NIT-GC2 morpholino models, we eliminate the $C_R(t)$ variable since it does not have any feedback coupling with the $C_P(t)$ equation and does not affect its dynamics.
 When fitting to the cPRC data using the NOS-knockout model, we also eliminate the $S(t)$, $C_N(t)$ and $C_R(t)$ variables since these do not have any effect on the $C_P(t)$.
-*  Set the size of the initial population to 1,000. This value provides a compromise in terms of the coverage of the parameter space and run-time of the GA.
-* Set bounds of random distributions of parameters.
+-   Set the size of the initial population to 1,000. This value provides a compromise in terms of the coverage of the parameter space and run-time of the GA.
+-   Set bounds of random distributions of parameters.
 We draw all the initial parameter values from uniform distributions over the unit interval ([0,1]).
 For WT condition, we fit 13 parameters, for NOS-knockout, we fit 7 parameters and for NIT-GC2 morpholino, we fit 4 parameters.
-* Evaluate the objective function on the initial population.
-* Run the \texttt{gamultiobj} algorithm.
-* Repeat steps 1 to 4 with bounds of parameters extended to [0,5] and using the Pareto front returned in step 4 as a part of the initial population.
+-   Evaluate the objective function on the initial population.
+-   Run the \texttt{gamultiobj} algorithm.
+-   Repeat steps 1 to 4 with bounds of parameters extended to [0,5] and using the Pareto front returned in step 4 as a part of the initial population.
 This step was added after observing that, although good fits to most datasets are achieved with parameter values in the interval [0,1], fits to some datasets can be improved by extending the admissible parameter range.
-* Use the Euclidean norm to find the set of model parameters that best fits the data.
+-   Use the Euclidean norm to find the set of model parameters that best fits the data.
 
@@
 First, we fix the parameters that were found by fitting the model to the cPRC data.
-Next, we run the optimisation in the same way as in the case of the cPRC data but only for the parameters that pertain to dynamics that is strictly`downstream' of cPRC dynamics.
+Next, we run the optimisation in the same way as in the case of the cPRC data but only for the parameters that pertain to dynamics that is strictly\`downstream' of cPRC dynamics.
 
@@
 
-* We find an inflection point that occurs after the start ($t_{UV\: start}$) of the UV stimulation.
+-   We find an inflection point that occurs after the start ($t_{UV\: start}$) of the UV stimulation.
 To this end, we identify the first time point $t_{AUP\: start}>t_{UV\: start}$ for which $d^2C_P(t)/dt^2=0$ and $d^2C_P(t_{k_{AUP\: start}}-1)/dt^2<d^2C_P(t_{k_{AUP\: start}})/dt^2<d^2C_P(t_{k_{AUP\: start}}+1)/dt^2$.
 We assume that the inflection is caused by the input from the INNOS cells.
-* We then compute the $AUP$ as an integral over $C_P(t)$ starting at the inflection point and ending 125 seconds later: $AUP=\int_{t_{start}}^{t_{end}} C_P(t) dt $.
+-   We then compute the $AUP$ as an integral over $C_P(t)$ starting at the inflection point and ending 125 seconds later: \$AUP=\int*{t*{start}}\^{t\_{end}} C_P(t) dt \$.
 To compute this, we use numerical integration via the trapezoidal method as implemented in Matlab function \texttt{trapz}.
 We remark that the baseline value of the $AUP$ for a Ca^2+^ profile of constant, unit value is 125.
-* To capture possible variability between different cells/organisms, we compute the $AUP$ using 47 model parameter sets fitted to individual WT cPRC experimental recordings.
+-   To capture possible variability between different cells/organisms, we compute the $AUP$ using 47 model parameter sets fitted to individual WT cPRC experimental recordings.
 We remove outliers, defined as being any simulated trajectories for which $C_P(t)$, after normalisation, exits the interval [0.25, 4] at any time.
 This $C_P(t)$ range was decided of the basis of values observed in various experiments in the lab.
-* For analysis, we use the median value of the $AUP$ at each combination of amplitude and duration.
+-   For analysis, we use the median value of the $AUP$ at each combination of amplitude and duration.
 
@@
 
-Maximum-likelihood phylogenetic tree of NOS protein sequences.** Individual branches are coloured by taxonomy.
+Maximum-likelihood phylogenetic tree of NOS protein sequences.\*\* Individual branches are coloured by taxonomy.
 Support values are based on 1000 replicates of ultrafast bootstrap with the aLRT-SH-like and aBayes methods.

### 2023-06-26 Manuscript/Manuscript.Rmd
About mycoplasma
Ms: comments incorporated

@@
 A.
-Wedgewood<sup>1</sup> and Gáspár Jékely<sup>1,4,`*`</sup>"
+Wedgwood<sup>1</sup> and Gáspár Jékely<sup>1,4,`*`</sup>"
     affiliation: "<sup>1</sup>Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4 4QD, United Kingdom <sup>2</sup>Okinawa Institute of Science and Technology, Okinawa, Japan, <sup>3</sup>School of Biological Sciences, University of Bristol, Bristol, United Kingdom, <sup>4</sup>Centre for Organismal Studies (COS), University of Heidelberg, 69120 Heidelberg, Germany  <br> <sup>*</sup>Correspondence`:` k.jokura @ exeter.ac.uk, gaspar.jekely @ cos.uni-heidelberg.de"
@@
 This free radical synthesized from L-arginine by nitric oxide synthase (NOS) is short-lived and can diffuse across biological membranes [@Cudeiro1999; @thomas2015].
-Canonical NO signalling involves the Ca^2+^/calmodulin-dependent activation of NOS, NO production and diffusion, and the NO-dependent activation of soluble guanylate cyclases (sGC).
-This can occur cell-autonomously or in other cells leading to cGMP production [@Bredt1990; @Hoelscher1997].
-Given that NOS activation is calcium dependent and NOS shows neuron-type-specific expression [@Gibbs1998a; @Wildemann1999; @Aso2019; @Mobley2022], NO action can lead to the activity-dependent modulation of neural circuits at specific sites [@Wang2007; @Vielma2014; @jacoby2018; @Aso2019] .
+Canonical NO signalling involves the Ca^2+^/calmodulin-dependent activation of NOS, NO production and diffusion, and the NO-dependent activation of soluble guanylate cyclases (sGC) leading to cGMP production [@Bredt1990; @Hoelscher1997].
+Given that NOS activation is Ca^2+^ dependent and NOS shows neuron-type-specific expression [@Gibbs1998a; @Wildemann1999; @Aso2019; @Mobley2022], NO action can lead to the activity-dependent modulation of neural circuits at specific sites [@Wang2007; @Vielma2014; @jacoby2018; @Aso2019] .
 
@@
 The cPRCs express a ciliary-type opsin, c-opsin1 [@Arendt2004] that forms a UV-absorbing bistable photopigment with an absorption maximum around 384 nm [@tsukamoto2017; @Veraszto2018; @Veedin2021].
-Upon UV/violet exposure, the cPRCs show a characteristic biphasic calcium response that is c-opsin1 dependent.
+Upon UV/violet exposure, the cPRCs show a characteristic biphasic Ca^2+^ response that is c-opsin1 dependent.
 UV/violet avoidance is also *c-opsin1*-dependent and is defective in *c-opsin1* mutants [@Veraszto2018].
 Here we show that *Platynereis* *NOS* is expressed in interneurons of the cPRC circuit and is required for UV/violet-avoidance.
-By combining calcium imaging across the fully-mapped cPRC circuit [@Veraszto2018] with genetic perturbations and mathematical modelling, we describe how NO tunes circuit dynamics through non-synaptic retrograde signalling to cPRCs.
+By combining Ca^2+^ imaging across the fully-mapped cPRC circuit [@Veraszto2018] with genetic perturbations and mathematical modelling, we describe how NO tunes circuit dynamics through non-synaptic retrograde signalling to cPRCs.
 This delayed neuroendocrine feedback integrates UV/violet exposure to induce a short-term memory manifested in altered circuit activity and an aversive behavioural response.
@@
 Phylogenetic analysis of NOS proteins indicate that *Platynereis* NOS belongs to an orthology group of bilaterian NOS sequences (Figure 1---figure supplement 1).
-To characterise the expression pattern of *NOS* we used in situ hybridization chain reaction (HCR) and transient transgenesis.
+To characterise the expression pattern of *NOS*, we used in situ hybridization chain reaction (HCR) and transient transgenesis.
 In two- and three-day-old larvae, we detected *NOS* expression in four cells (two of them weakly expressing) in the apical organ region (Figure 1D and Figure 1---figure supplement 2).
@@
 Following light stimulation in the region of the ramified cilia of the cPRCs, we detected an increase in DAF-FM fluorescence in the anterior neurosecretory neuropil, the region of INNOS projections.
-This increase did not occur in larvae where we illuminated a control area (Figure 2).
+This increase did not occur in larvae in which we illuminated a control area (Figure 2).
 
@@
 We next tested whether NO signalling regulates UV/violet avoidance.
-To achieve this, we generated two *Platynereis* *NOS* knockout lines with the CRISPR/Cas9 method.
+To achieve this, we generated two *Platynereis* *NOS* knockout lines using the CRISPR/Cas9 method.
 We recovered two deletions (NOSΔ11/Δ11 and Δ23/Δ23), both frame-shift mutations leading to an early stop codon and thus likely representing null alleles (Figure 3---figure supplement 1A).
@@
 
-To investigate how NO signalling alters the dynamics of the cPRC circuit, we carried out calcium imaging experiments.
-We ubiquitously expressed the calcium sensor GCaMP6s in larvae and imaged calcium signals during 405 nm stimulation of the cPRCs.
-As we have shown previously, a 20-sec local stimulation of cPRC cilia led to a transient increase in cPRC calcium levels, followed by a transient decrease [@Veraszto2018].
-After \~20-sec, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus -- a response that may involve depolarisation (Figure 4A, B).
+To investigate how NO signalling alters the dynamics of the cPRC circuit, we carried out Ca^2+^ imaging experiments.
+We ubiquitously expressed the Ca^2+^ sensor GCaMP6s in larvae and imaged Ca^2+^ signals during 405 nm stimulation of the cPRCs.
+As we have shown previously, a 20 sec local stimulation of cPRC cilia led to a transient increase in cPRC Ca^2+^ levels, followed by a transient decrease [@Veraszto2018].
+After \~20-sec, Ca^2+^ levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus -- a response that may involve cPRC depolarisation (Figure 4A, B).
 This activation phase occurs after the 20 sec stimulation period and is likely due to a delayed neuroendocrine feedback [@Veraszto2018].
@@
 While we detected the initial activation phase followed by inhibition, in homozygous *NOS*-mutants for both CRISPR alleles this was not followed by delayed activation.
-Instead, calcium levels dropped to a low steady-state level (Figure 4A, B).
+Instead, Ca^2+^ levels dropped to a low steady-state level (Figure 4A, B).
 We thus identified a requirement for NO signalling in the late-phase activation of cPRCs.
@@
 To analyse the relationship of these sequences to metazoan NIT-GCs, we retrieved protein sequences with a CYC domain from the transcriptome and genome databases of 45 metazoan and 2 choanoflagellate species.
-We carried out cluster analysis and did phylogenetic reconstruction on a group of membrane-bound guanylate cyclases with sGCs as an outgroup.
+We carried out cluster analysis and phylogenetic reconstruction on a group of membrane-bound guanylate cyclases with sGCs as an outgroup.
 In agreement with Moroz et al. [@Moroz2020], we found a group of GCs with NIT domains with representatives in placozoans, cnidarians, some ecdysozoans, echinoderms, and lophotrochozoans.
@@
 
-To test the in vivo function of NIT-GC1 and NIT-GC2 in cPRC responses, we combined calcium imaging with morpholino-mediated knockdowns.
+To test the in vivo function of NIT-GC1 and NIT-GC2 in cPRC responses, we combined Ca^2+^ imaging with morpholino-mediated knockdowns.
 We used two translation-blocking morpholinos for each *NIT-GC* gene and tested knockdown efficiency by immunostaining injected animals with the NIT-GC1 and NIT-GC2 antibodies (Figure 4---figure supplement 3E,F).
@@
 This could occur via increased cGMP production and the opening of a cyclic-nucleotide-gated ion channel (CNG) specific to cPRCs [@Tosches2014].
-NIT-GC2 morphant larvae, in contrast, showed a step-up increase in calcium following light stimulation (Figure 4N).
-The calcium signal decayed during stimulation and was off after light off.
-These data support an essential role for ciliary-localised NIT-GC2 in suppressing cPRC calcium following its transient rise at stimulus onset.
+NIT-GC2 morphant larvae, in contrast, showed a step-up increase in Ca^2+^ following light stimulation (Figure 4N).
+The Ca^2+^ signal decayed during stimulation and was off after light off.
+These data support an essential role for ciliary-localised NIT-GC2 in suppressing cPRC Ca^2+^ following its transient rise at stimulus onset.
 Overall, these knockdown experiments revealed different signalling mechanisms for the two NIT-GCs that may be due to their different subcellular localisations.
@@
 ```{r echo=FALSE,fig.cap="**Figure 4.
-NOS and two NIT-GCs shape calcium signals during cPRC UV/violet response.** (**A**, **B**) GCaMP6s signals in cPRCs in wild type and *NOS* mutant (A, NOSΔ11/Δ11, B, NOSΔ23/Δ23) larvae during 405 nm light stimulation.
+NOS and two NIT-GCs shape Ca^2+^ signals during cPRC UV/violet response.** (**A**, **B**) GCaMP6s signals in cPRCs in wild type and *NOS* mutant (A, NOSΔ11/Δ11, B, NOSΔ23/Δ23) larvae during 405 nm light stimulation.
 (**C**, **D**) In situ HCR for (C) *NIT-GC1* and (D) *NIT-GC2* (magenta) in three-day-old *Platynereis* larvae.
@@
 
-To investigate how NO and NIT-GC signalling influence the dynamics of the cPRC circuit, we imaged calcium signals from postsynaptic neurons in wild type, mutant and morphant larvae.
+To investigate how NO and NIT-GC signalling influence the dynamics of the cPRC circuit, we imaged Ca^2+^ signals from postsynaptic neurons in wild type, mutant and morphant larvae.
 We were able to image the activity of all neurons in the cPRC circuit (INNOS, INRGWa, Ser-h1 and MC).
 The MC cell was identified based on its position and intrinsic activity [@veraszto2017].
-To unambiguously identify all other cells from which we recorded calcium signals, we developed an on-slide immunostaining method (Figure 5---figure supplement 1A).
-We used the cell-specific antibody markers against RYamide (INNOS) (Figure 5---figure supplement 1B-D), RGWamide (INRGWa) and serotonin (Ser-h1) [@Conzelmann2011] to immunostain agar-embedded larvae following calcium imaging.
+To unambiguously identify all other cells from which we recorded Ca^2+^ signals, we developed an on-slide immunostaining method (Figure 5---figure supplement 1A).
+We used the cell-specific antibody markers against RYamide (INNOS) (Figure 5---figure supplement 1B-D), RGWamide (INRGWa) and serotonin (Ser-h1) [@Conzelmann2011] to immunostain agar-embedded larvae following Ca^2+^ imaging.
 Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision (Figure 5A, B).
-Due to the stereotypy of the larvae, we could also identify neurons based on their position and calcium activity in activity-correlation maps (Figure 5C).
+Due to the stereotypy of the larvae, we could also identify neurons based on their position and Ca^2+^ activity in activity-correlation maps (Figure 5C).
 
 We first quantified the responses of the INNOS and INRGWa interneurons during 405 nm stimulation of the cPRCs.
-In both wild type and *NOS*-mutant larvae, INNOS cells showed an increase in calcium during stimulation (Figure 5D).
+In both wild type and *NOS*-mutant larvae, INNOS cells showed an increase in Ca^2+^ during stimulation (Figure 5D).
 In contrast, the INNOS response was flat or slightly negative in NIT-GC2 morphant larvae (Figure 5E) revealing an essential role for NIT-GC2-mediated cPRC suppression in INNOS activation.
@@
 This late INRGWa response was lacking in *NOS*-mutants (Figure 5F).
-In NIT-GC2 morphants, INRGWa cells showed a transient increase in calcium that decayed after light off and a delayed activation was not present (Figure 5G).
+In NIT-GC2 morphants, INRGWa cells showed a transient increase in Ca^2+^ that decayed after light off and a delayed activation was not present (Figure 5G).
 
-Next, we imaged calcium signals from Ser-h1 and MC neurons in wild type and *NOS* mutant larvae.
-Ser-h1 cells showed an activation profile that correlated with cPRC activity, including a reduction in calcium during stimulation followed by rebound, a response that was defective in *NOS* mutants (Figure 5I).
+Next, we imaged Ca^2+^ signals from Ser-h1 and MC neurons in wild type and *NOS* mutant larvae.
+Ser-h1 cells showed an activation profile that correlated with cPRC activity, including a reduction in Ca^2+^ during stimulation followed by rebound, a response that was defective in *NOS* mutants (Figure 5I).
 MC cells showed sustained activation, including a late-phase that was lacking in *NOS* mutants (Figure 5I).
@@
 
-Through this optimisation procedure, we found parameters that gave a very good fit to our experimental data under all conditions.
-We could fit to mean Ca^2+^ values but also individual calcium transients (Figure 6B, Figure 6--figure supplement 1).
+Through this optimisation procedure, we found parameters that gave a fit to our experimental data under all conditions.
+We could fit to mean Ca^2+^ profiles but also individual Ca^2+^ transients (Figure 6B, Figure 6--figure supplement 1).
 Our model only includes a minimal set of parameters and assumptions of interactions that are required to describe the dynamics of cPRC, INNOS and INRGWa in wild type and loss-of-function conditions (Figure 6A, B).
@@
 Mathematical modelling and signalling mechanisms of the cPRC circuit.** (**A**) Diagram of the mathematical model with componenets, interactions, parameters and equations used to model Ca^2+^ dynamics.
-(**B**) Average calcium traces and modelled traces fitted to the data.
+(**B**) Average Ca^2+^ traces (upper) and traces from model simulations (lower) fitted to the data.
 (**C**) Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq.
@@
 c-opsin1 forms a bistable photopigment and signals through Gi/oα and Gβγ [@tsukamoto2017; @Veedin2021; @tsukamoto2023].
-In heterologous systems, the Gβγ subunits released following c-opsin1 activation open GIRK channels inducing K+ efflux (hyperpolarisation) [@tsukamoto2017; @tsukamoto2023] and close voltage-gated Ca^2+^ channels, thereby reducing intracellular calcium levels [@tsukamoto2023].
+In heterologous systems, the Gβγ subunits released following c-opsin1 activation open GIRK channels inducing K+ efflux (hyperpolarisation) [@tsukamoto2017; @tsukamoto2023] and close voltage-gated Ca^2+^ channels, thereby reducing intracellular Ca^2+^ levels [@tsukamoto2023].
 A GIRK channel is also expressed in the cPRCs (Figure 6A).
@@
 NO is produced in the axons of R3 neurons and acts directly on sGC in the same axons.
-This autocrine signal leads to a CNG-dependent temporary increase in calcium levels, carrying the working memory trace [@Kuntz2017].
+This autocrine signal leads to a CNG-dependent temporary increase in Ca^2+^ levels, carrying the working memory trace [@Kuntz2017].
 
@@
 From these sequences, we produced a Hidden Markov Model (HMM) and used this to mine the 45 metazoan, 2 choanoflagellate and 2 filasterea species investigated.
-HMM models were run in HMMR3 with an e-value of 1e−15.
+HMM models were run in HMMR3 [@Eddy2011] with an e-value of 1e−15.
 We ran CD-Hit [@Fu2012] to eliminate redundant sequences (at a 80% threshold).
@@
 
-For calcium imaging, 49-55 hours post fertilisation larvae were used.
+For Ca^2+^ imaging, 49-55 hours post fertilisation larvae were used.
 Experiments were carried out at room temperature and larvae were immobilised in 2.5% agarose in artificial sea water and mounted between a slide and coverslip spaced with adhesive tape.
@@
 For the assay, COS-7 cells (Angio-proteomie, CAT no. cAP-0203) with a low endogenous level of soluble guanylate cyclase activity were used.
+Cells were not contaminated with mycoplasma.
 Cells were maintained at 37 °C in 35 mm dishes (Nunc™ Glass Bottom Dishes) containing 3 ml of DMEM, low-glucose medium (Thermo; Cat.
@@
 KJ has been supported by a JSPS Overseas Research Fellowship, LAYG by a BBSRC Discovery fellowship (BB/W010305/1), PS by a Wellcome Trust Institutional Strategic Support Award (204909/Z/16/Z), and KCAW by the EPSRC Hub for Quantitive Modelling in Healthcare (EP/T017856/1).
+We thank Dr. Francesca Carlisle for the mycoplasma testing of COS-7 cells.
 
@@
 ```{r echo=FALSE,fig.cap="**Figure 5 -- figure supplement 1.
-On-slide immunostaining and coexpression analysis of *NOS* and *RYamide proneuropeptide* ** (**A**) Schematic diagram of the on-slide immunostaining procedure after calcium imaging.(**B-D**) Co-expression analysis by HCR in situ of *NOS* (magenta) and *RYamide proneuropeptide* (green).
+On-slide immunostaining and coexpression analysis of *NOS* and *RYamide proneuropeptide* ** (**A**) Schematic diagram of the on-slide immunostaining procedure after Ca^2+^ imaging.(**B-D**) Co-expression analysis by HCR in situ of *NOS* (magenta) and *RYamide proneuropeptide* (green).
 Anterior view of a two-day-old larva.

### 2023-07-03 Manuscript/Manuscript.Rmd
Ms edits and bib updates

@@
 Nitric oxide (NO) is one such modulator with unique physical and signalling properties.
-This free radical synthesized from L-arginine by nitric oxide synthase (NOS) is short-lived and can diffuse across biological membranes [@Cudeiro1999; @thomas2015].
+This free radical synthesized from L-arginine by nitric oxide synthase (NOS), is short-lived and can diffuse across biological membranes [@Cudeiro1999; @thomas2015].
 Canonical NO signalling involves the Ca^2+^/calmodulin-dependent activation of NOS, NO production and diffusion, and the NO-dependent activation of soluble guanylate cyclases (sGC) leading to cGMP production [@Bredt1990; @Hoelscher1997].
@@
 In the vertebrate retina, NOS is expressed in amacrine, ganglion and other cells and the actions of NO can be diverse [@Cudeiro1999; @Wang2007; @jacoby2018].
-For example, defective NO signalling in NOS knockouts leads to a decreased sensitivity of retinal ganglion cells to light stimulation [@Wang2007].
+For example, defective NO signalling in NOS knockout mice leads to a decreased sensitivity of retinal ganglion cells to light stimulation [@Wang2007].
 In the retina, NO signalling can also involve pathways other than canonical sGC signalling [@Tooker2013; @Wei2012; @jacoby2018].
@@
 The dashed line corresponds to the area where fluorescence was quantified.
-Circles indicate the location of cPRC and control stimulation.
-cPRCs are marked by thin lines.
+The circles indicate the location of cPRC and control stimulation.
+The cPRCs are marked by thin lines.
 (**B**) Changes in DAF-FM fluorescence before and during 405 nm light stimulation.
@@
 Among the 15 NIT-GCs, two showed high and specific expression in the cPRCs and one was expressed in the INNOS cells (Figure 4---figure supplement 2).
-In the single-cell data, we could identify the cPRCs by the specific expression of *c-opsin1* and the *pedal-peptide2 neuropeptide precursor* (*MLD proneuropeptide*), previously described cPRC markers [@Arendt2004; @williams2017] (Figure 4---figure supplement 3A).
+In the single-cell data, we could identify the cPRCs by the specific expression of the markers *c-opsin1* and the *pedal-peptide2 neuropeptide precursor* (*MLD proneuropeptide*) [@Arendt2004; @williams2017] (Figure 4---figure supplement 3A).
 The INNOS cells were identified by NOS expression and spatial mapping in the brain [@Achim2015].
@@
 To confirm the single-cell data, we first carried out in situ hybridisation chain reaction (HCR) with probes for *NIT-GC1* and *NIT-GC2* mRNA.
-Both genes were specifically expression in the four cPRCs, as confirmed by co-labeling with an acetylated α-tubulin antibody and with an HCR probe against *pedal peptide 2/MLD proneuropeptide* (Figure 4C, D and Figure 4---figure supplement 3A-C).
+Both genes were specifically expressed in the four cPRCs, as confirmed by co-labeling with an acetylated α-tubulin antibody and with an HCR probe against *pedal peptide 2/MLD proneuropeptide* (Figure 4C, D and Figure 4---figure supplement 3A-C).
 To analyse the subcellular localisation of NIT-GC1 and NIT-GC2 at the protein level, we raised and affinity-purified polyclonal antibodies against a specific peptide sequence from both proteins.
@@
 In contrast, the INNOS response was flat or slightly negative in NIT-GC2 morphant larvae (Figure 5E) revealing an essential role for NIT-GC2-mediated cPRC suppression in INNOS activation.
-INRGWa cells were initially inhibited during cPRC stimulation, followed by a delayed activation paralleling the second activation phase of cPRCs.
+The INRGWa cells were initially inhibited during cPRC stimulation, followed by a delayed activation paralleling the second activation phase of the cPRCs.
 This late INRGWa response was lacking in *NOS*-mutants (Figure 5F).
@@
 Next, we imaged Ca^2+^ signals from Ser-h1 and MC neurons in wild type and *NOS* mutant larvae.
-Ser-h1 cells showed an activation profile that correlated with cPRC activity, including a reduction in Ca^2+^ during stimulation followed by rebound, a response that was defective in *NOS* mutants (Figure 5I).
-MC cells showed sustained activation, including a late-phase that was lacking in *NOS* mutants (Figure 5I).
-These data suggest that during 405 nm stimulation the Ser-h1 cells are inhibited and MC cells are activated, and this regulation is NO-dependent.
+Ser-h1 cells showed an activation profile that correlated with cPRC activity, including a reduction in Ca^2+^ during stimulation followed by a rebound, a response that was defective in *NOS* mutants (Figure 5I).
+The MC cell showed sustained activation, including a late-phase that was lacking in *NOS* mutants (Figure 5I).
+These data suggest that during 405 nm stimulation the Ser-h1 cells are inhibited and the MC cell is activated, and this regulation is NO-dependent.
 This pattern is expected to inhibit ciliary activity in the prototroch but not in the other ciliary bands (with no Ser-h1 or MC synapses), triggering NO-dependent downward swimming.
@@
 To further analyse the dynamics of responses to UV light and formally describe cPRC phototransduction, synaptic connections and NO retrograde signalling, we developed a mixed cellular-circuit-level mathematical model.
-We used our Ca^2+^ imagining data of cPRC, INNOS and INRGWa cells collected in wild type, *NOS* knockout, and *NIT-GC2* morphant larvae to formulate assumptions that are the basis of the proposed model.
+We used our Ca^2+^ imagining data of cPRC, INNOS and INRGWa cells collected in wild type, *NOS* knockout and *NIT-GC2* morphant larvae to formulate assumptions that are the basis of the proposed model.
 
-We model a direct c-opsin1-dependent response to UV leading to the initial rise in Ca^2+^ levels.
+We modeled a direct c-opsin1-dependent response to UV leading to the initial rise in Ca^2+^ levels.
 The mechanism of this signal is not known but may include Gα and Gβ signalling, CNGAα channel activation or other pathways.
-We further assume that in cPRC cells, Ca^2+^ levels increase with cGMP.
-We model a decrease in cGMP and Ca^2+^ levels due to a NIT-GC2-dependent pathway (Figure 6A).
+We further assumed that in cPRC cells, Ca^2+^ levels increase with cGMP.
+We modeled a decrease in cGMP and Ca^2+^ levels due to a NIT-GC2-dependent pathway (Figure 6A).
 The effect of NO in cPRC cells is captured by another term that describes a NIT-GC1 and NO-dependent increase in cGMP.
-Based on the synaptic connectome, we infer a feedforward coupling between cPRC cells and INNOS and INRGWa cells.
+Based on the synaptic connectome, we infered a feedforward coupling between cPRC cells and INNOS and INRGWa cells.
 The proposed synaptic coupling is UV-dependent, decays linearly and is suppressed in NIT-GC2 morphants (Figure 6A).
 
-In the INNOS cells, we assume that excitatory synaptic input from cPRC (or rebound from tonic inhibition) leads to a rise in Ca^2+^ leading to NO production via NOS.
+In the INNOS cells, we assumed that excitatory synaptic input from cPRC (or rebound from tonic inhibition) leads to a rise in Ca^2+^ leading to NO production via NOS.
 This input is NIT-GC2-dependent.
-In the INRGWa cells, we model a decrease in Ca^2+^ based on an inhibitory UV-dependent synaptic signal from the cPRCs.
-We further assume the existence of direct excitatory coupling between cPRC Ca^2+^ levels and INRGWa Ca^2+^ to account for the late effects of cPRC on INRGWa (Figure 6A).
+In the INRGWa cells, we modeled a decrease in Ca^2+^ based on an inhibitory UV-dependent synaptic signal from the cPRCs.
+We further assumed the existence of direct excitatory coupling between cPRC Ca^2+^ levels and INRGWa Ca^2+^ to account for the late effects of cPRC on INRGWa (Figure 6A).
 The UV-dependent and Ca^2+^-dependent signals may be mediated by different transmitters released by cPRCs during distinct phases of their activation cycle.
-In addition, we assume a direct inhibitory coupling between INNOS Ca^2+^ and INRGWa Ca^2+^.
+In addition, we assumed a direct inhibitory coupling between INNOS Ca^2+^ and INRGWa Ca^2+^.
 Finally, in all variables, we assumed a linear decay and constant production to set a steady state (Figure 6A).
@@
 ```{r echo=FALSE,fig.cap="**Figure 6.
-Mathematical modelling and signalling mechanisms of the cPRC circuit.** (**A**) Diagram of the mathematical model with componenets, interactions, parameters and equations used to model Ca^2+^ dynamics.
+Mathematical modelling and signalling mechanisms of the cPRC circuit.** (**A**) Diagram of the mathematical model with the componenets, interactions, parameters and equations used to model Ca^2+^ dynamics.
 (**B**) Average Ca^2+^ traces (upper) and traces from model simulations (lower) fitted to the data.
@@
 (**D**) Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback.
-"}
+<br> Figure 6 -- source data 1.
+TPM values for each gene and the percentage of expressed genes."}
 
@@
 ```{r echo=FALSE,fig.cap="**Figure 7.
-Magnitude of cPRC activation as  a function of stimulus duration and intensity.** Median values of the $AUP$ are presented as a heatmap with dark blue indicating flat $C_P(t)$ trajectories with median $AUP$ close to 0 and red indicating that the $C_P(t)$ trajectories have a clearly defined delayed activation peak.
+Magnitude of cPRC activation as  a function of stimulus duration and intensity.** Median values of the $AUP$ are presented as a heatmap with dark blue indicating flat $C_P(t)$ traces with median $AUP$ close to 0 and red indicating that the $C_P(t)$ traces have a clearly defined delayed activation peak.
 "}
@@
 NO is a free radical with a millisecond-to-second half life and thus a limited signalling range.
-In neuronal signalling, NOS is often localized to neurites [@Kuntz2017] at close proximity to sGC at synapses [@burette2002; @garthwaite2015a].
+In neuronal signalling, NOS is often localised to neurites [@Kuntz2017] close to sGC at synapses [@burette2002; @garthwaite2015a].
 In the cPRC circuit, NOS is localised to the dendritic compartment of INNOS cells where also cPRC postsynaptic sites occur.
 NIT-GC1, the target of NO signalling is localised to cPRC projections.
-NO-mediated retrograde signalling thus likely occurs in the neurosecretory plexus where NOS- and NIT-GC1-containing projections are in close proximity and where we detected NO production following UV stimulation.
+NO-mediated retrograde signalling thus likely occurs in the neurosecretory plexus where NOS- and NIT-GC1-containing projections are in close proximity and where we detected NO production after UV stimulation.
 In contrast, INNOS to INRGWa synapses occur outside the neurosecretory plexus in the ventral brain neuropil.
@@
 The activation of INNOS leads to NOS activation and NO release, potentially through a canonical Ca^2+^-calmodulin pathway.
-Our model predicts feedforward inhibiton from INNOS to INRGWa, possibly mediated by glycin transmission.
+Our model predicts feedforward inhibition from INNOS to INRGWa, possibly mediated by glycin transmission.
 NO released by INNOS neurites activates NIT-GC1 in cPRC projections that could lead to cGMP production and the opening of CNGAα.
-This late-phase cPRC activation is not directly dependent on the UV/violet signal and can happen post-stimulus.
+This late-phase cPRC activation does not directly depend on the UV/violet signal and can happen post-stimulus.
 The late-phase cPRC activation leads to INRGWa activation via an unknown transmitter that is likely different from the one acting on INNOS.
@@
 In addition, the cPRC circuit expresses several neuropeptides and their receptors, suggesting further neuromodulatory signals.
-For example, INRGWa express the proneuropeptide RGWamide and cPRCs and INNOS express its receptor, suggesting retrograde peptidergic signalling in the circuit.
+For example, the INRGWa cells express the proneuropeptide RGWamide and cPRCs and INNOS express its receptor, suggesting retrograde peptidergic signalling in the circuit.
 
@@
 
-We hypothesise that the differential beating of prototroch versus trunk cilia causes a head-up or head-down orientation due to physics alone.
+We hypothesize that the differential beating of prototroch versus trunk cilia causes a head-up or head-down orientation due to physics alone.
 During the UV response, prototroch cilia beat slower than trunk cilia, resulting in a head-down stable state ('rear-wheel drive').
@@
 The avoidance of noxious UV/violet light is often mediated by extraocular photoreceptors and their circuits, distinct from the pigmented visual eyes.
-These two types of systems co-exist in *Platynereis* where the pigmented eyes and eyespots guide phototaxis with a maximum sensitivity to cyan light (\~500 nm) [@Randel2014a; @Veraszto2018].
+These two types of systems co-exist in *Platynereis* where the pigmented eyes and eyespots guide phototaxis with a maximum sensitivity to cyan light (\~500 nm) [@Randel2014a; @Gühmann2015; @Veraszto2018].
 Planarians also have pigmented visual eyes mediating phototaxis to cyan light and peripheral extraocular photoreceptors mediating UV avoidance behaviour [@Shettigar2021].
@@
 
-For the generation of a *NOS* knockout we used an sgRNA targeting the third exon of the *Platynereis dumerilii* *NOS* (target site: 5'-GGGCAATACTGGCTCCACTC-3').
+For the generation of a *NOS* knockout, we used an sgRNA targeting the third exon of the *Platynereis dumerilii* *NOS* gene (target site: 5'-GGGCAATACTGGCTCCACTC-3').
 We selected the site to avoid polymorphic sites in our laboratory culture.
@@
 We assayed larval photoresponses in a vertical Plexiglas column (31 mm x 10 mm x 160 mm water height).
-For stimulation, we illuminated the column from the top with a monochromator (Polychrome II, Till Photonics) controlled by AxioVision 4.8.2.0 (Carl Zeiss GmbH, Jena) via analog voltage.
+For stimulation, we illuminated the column from above with a monochromator (Polychrome II, Till Photonics) controlled by AxioVision 4.8.2.0 (Carl Zeiss GmbH, Jena) via analog voltage.
 The light passed a collimator lens (LAG-65.0-53.0-C with MgF2 Coating, CVI Melles Griot) before entering the column.
@@
 
-To compare the behavior of wildtype and NOS-knockout larvae at 3 days in the vertical column, the larvae were mixed and left in the dark for 5 min.
-The larvae were treated with NOS inhibitors for pharmacology.
-The NOS inhibitors were L-NAME.
-The larvae were treated with different concentrations in adjacent columns.
-The concentrations for the NOS inhibitors were control, 1 mM, 0.1 mM.
-The larvae were recorded for 1 min in the dark followed by exposure to collimated cyan (480 nm) light from the top of the column for 2 min, then 2 min darkness, and finally collimated UV (395 nm) light from the top of the column for 2 min.
-Stimulus light was provided by the monochromator (Polychrome II, Till Photonics).
-Scripts are available at <https://github.com/JekelyLab/NOS>.
+We compared the behavior of wildtype and *NOS*-knockout larvae in the vertical column.
+In the inhibitor experiments we compared wildtype larvae before and after treatment with 0.1 mM or 1.0 mM NOS inhibitor L-NAME (in seawater).
+All larvae were three days oldt.
+Before the experiment, the larvae were mixed and left in the dark for 5 min.
+The larvae were recorded for 1.5 min in the dark followed by exposure to diffuse UV light (395 nm) from the side for 2 min.
+Then the larvae were left for another 2 min in darkness followed by exposue to collimated cyan (480 nm) light from above for 2 min, then 2 min darkness, and finally collimated UV (395 nm) light from above for 2 min.
+Scripts are available at <https://github.com/JekelyLab/Jokura_et_al_NOS>.
 
@@
 To identify NOS sequences, we obtained a seed database of oxygenase domains from the Pfam database (PF02898).
-From these sequences, we produced a Hidden Markov Model (HMM) and used this to mine the 47 metazoan, 2 choanoflagellate and 2 filasterea species investigated.
-HMM models were run in HMMR3 with an e-value of 1e−15.
+From these sequences, we generated a Hidden Markov Model (HMM) to mine the genomes and transcriptomes of 47 metazoan, 2 choanoflagellate and 2 filasterea species.
+The HMM model was run in HMMR3 with an e-value of 1e−15.
 We ran CD-Hit [@Fu2012] to eliminate redundant sequences (at a 80% threshold).
 We aligned the sequences with MAFFT version 7, with the iterative refinement method E-INS-i.
-Alignments were trimmed with TrimAl in gappy-out mode [@Capella-Gutiérrez2009].
-To calculate maximum-likelihood trees, we used IQ-tree2 with the LG+G4 model.
-To calculate branch support, we ran 1000 replicates with the aLRT-SH-like and aBayes methods [@Minh2020].
-The sequences used for the phylogenetic analysis are available in **Supplementary file 1**, the trimmed alignment is available in **Supplementary File 2** and the pre-trimmed data in **Supplementary file 3**.
+The alignment was trimmed with TrimAl in gappy-out mode [@Capella-Gutierrez2009].
+To calculate a maximum-likelihood tree, we used IQ-tree2 with the LG+G4 model.
+To calculate branch support, we ran 1000 replicates with the aLRT-SH-like [@Guindon2010] and aBayes methods [@Minh2020].
 
@@
 To identify NIT-GCs, we obtained a seed database of adenylate and guanylate cyclase catalytic domains from the Pfam database (PF00211).
-From these sequences, we produced a Hidden Markov Model (HMM) and used this to mine the 45 metazoan, 2 choanoflagellate and 2 filasterea species investigated.
-HMM models were run in HMMR3 [@Eddy2011] with an e-value of 1e−15.
+From these sequences, we generated a Hidden Markov Model (HMM) to mine the 45 metazoan, 2 choanoflagellate and 2 filasterea genomes and transcriptomes.
+The HMM model was run in HMMR3 [@Eddy2011] with an e-value of 1e−15.
 We ran CD-Hit [@Fu2012] to eliminate redundant sequences (at a 80% threshold).
@@
 We aligned the sequences with MAFFT version 7, with the iterative refinement method E-INS-i[@Katoh2013].
-Alignments were trimmed with TrimAl in gappy-out mode [@Capella-Gutiérrez2009].
-To calculate maximum-likelihood trees, we used IQ-tree2 with the LG+G4 model.
+The alignment was trimmed with TrimAl in gappy-out mode [@Capella-Gutierrez2009].
+To calculate a maximum-likelihood tree, we used IQ-tree2 with the LG+G4 model.
 To calculate branch support, we ran 1000 replicates with the aLRT-SH-like and aBayes methods [@Minh2020].
-The sequences used for the phylogenetic analysis are available in **Supplementary File 4**, the trimmed alignment is available in **Supplementary File 5** and the pre-trimmed data in **Supplementary file 6**.
 
@@
 We used a single-cell RNA sequencing dataset from *Platynereis* larvae [@Achim2015] that was further filtered to 107 cells [@williams2017].
-We normalized raw read-count data to TPM and converted to log10.
+We normalized the raw read-count data to TPM and converted to log10.
 To calculate % expression, we used the sum of the expression levels in the 107 cells for each gene.
@@
 Individual cells were identified by suits of well-characterised marker genes and spatial mapping [@Achim2015].
-The total TPM values for each gene and the percentage of expressed genes are available in **Supplementary file 7**.
 
@@
 Probe hybridization buffer, probe wash buffer, amplification buffer, and fluorescent HCR hairpins were purchased from Molecular Instruments (Los Angeles, USA).
-Hairpins associated with the b2 initiator sequence were labeled with Alexa Fluor 647, and the hairpins associated with the b3 initiator sequence were labeled with Alexa Fluor 546.
+The hairpins associated with the b2 initiator sequence were labeled with Alexa Fluor 647, and the hairpins associated with the b3 initiator sequence were labeled with Alexa Fluor 546.
 To design probes for HCR, we used custom software [@Kuehn2022] to create 20 DNA oligo probe pairs specific to *P.
@@
 To remove excess probe, samples were washed 4× with 1 ml hybridization wash buffer for 15 min at 37°C, and subsequently 2× in 1 ml 5× SSCT (5× SSC with 0.1% Tween20) for 5 min at room temperature.
-For the amplification stage, samples were pre-incubated with 100 µl of amplification buffer for 30 min, room temperature, and then incubated with 150 µl amplification buffer containing fluorescently labeled hairpins (40nM concentration (2ul of 3uM stock in 150ul amplification buffer, snap-cooled as described; [@Choi2018]) overnight in the dark at 25°C.
+For the amplification stage, samples were pre-incubated with 100 µl of amplification buffer for 30 min, room temperature, and then incubated with 150 µl amplification buffer containing fluorescently labelled hairpins (40nM concentration (2ul of 3uM stock in 150ul amplification buffer, snap-cooled as described; [@Choi2018]) overnight in the dark at 25°C.
 To remove excess hairpins, samples were washed in 1 ml 5× SSCT at room temperature, twice for 5 min, twice for 30 min, and once for 5 min.
@@
 
-Whole-mount immunostaining of *Platynereis* larvae fixed with 4% paraformaldehyde was carried out using the primary antibodies listed in KEY RESOURCES TABLE.
+Whole-mount immunostaining of *Platynereis* larvae fixed with 4% paraformaldehyde was carried out using the primary antibodies listed in the KEY RESOURCES TABLE.
 Immunostainings were carried out as previously described [@Conzelmann2012].
@@
 Larvae injected with the promoter construct (ca.
-250 ng/ml) were analysed for reporter expression at three days post fertilization on an AxioImager Z.1 fluorescence microscope (Carl Zeiss GmbH, Jena) and if positive fixed for immunostaining.
+250 ng/ml) were analysed for reporter expression at three days post fertilization on an AxioImager Z.1 fluorescence microscope (Carl Zeiss GmbH, Jena) and if positive were fixed for immunostaining.
 The immunostaining for the HA-tagged reporter was done as described [@veraszto2017].
@@
 
-For Ca^2+^ imaging, 49-55 hours post fertilisation larvae were used.
-Experiments were carried out at room temperature and larvae were immobilised in 2.5% agarose in artificial sea water and mounted between a slide and coverslip spaced with adhesive tape.
-GCaMP6s mRNA (1 mg/ml) was injected into zygotes as described previously [@Randel2014a].
-Larvae were imaged on a Zeiss LSM 880 with Airyscan (with a C-Apochromat 63X/1.2 Corr - water) with a frame rate of 1.88 frame/sec and an image size of 512 x 512 pixels.
-The larvae were stimulated in a region of interest (a circle with 50 pixel diameter) with a 405 nm laser controlled by the Bleaching mode of the Zeiss software.
+For Ca^2+^ imaging, GCaMP6s mRNA (1 mg/ml) was injected into zygotes as described previously [@Randel2014a].
+At 49-55 hours post fertilisation, larvae were immobilised in 2.5% agarose in artificial sea water and mounted between a slide and coverslip spaced with adhesive tape.
+Larvae were imaged at room temperature on a Zeiss LSM 880 with Airyscan (with a C-Apochromat 63X/1.2 Corr - water) with a frame rate of 1.88 frame/sec and an image size of 512 x 512 pixels.
+The larvae were stimulated in a region of interest (a circle with 50 pixel diameter) with a 405 nm laser controlled by the bleaching mode of the Zeiss software.
 The imaging laser had a similar intensity than the stimulus laser but covered an area that was 10 times larger than the stimulus ROI.
@@
 Transfections were carried out with 150 ng of each plasmid and 0.3 μl of Lipofectamine 3000 Reagent (invitrogen; Cat. No. L3000001).
-Three hours post-transfection, the culture medium was substituted for fresh DMEM-medium.
+Three hours post-transfection, the culture medium was substituted with fresh DMEM medium.
 For single-wavelength imaging experiments, cells in 35-mm dishes were washed twice and imaged in modified Ringer's buffer (140 mM NaCl, 3.5 mM KCl, 0.5 mM NaH~2~PO~4~ , 0.5 mM S-3 MgSO~4~, 1.5 mM CaCl~2~ , 10 mM HEPES, 2 mM NaHCO~3~ and 5 mM glucose).
-Dishes were mounted on a stage heated to 37 °C and imaging was done on an inverted confocal microscope (LSM880, Carl Zeiss GmbH, Jena) equipped with an oil-immersion objective lens (UApo/340, 40x, NA = 0.17).
+The dishes were mounted on a stage heated to 37 °C and imaged on an inverted confocal microscope (LSM880, Carl Zeiss GmbH, Jena) equipped with an oil-immersion objective lens (UApo/340, 40x, NA = 0.17).
 The exposure time of the EM-CCD camera was controlled by the ZEN software (Carl Zeiss GmbH, Jena).
 Images were acquired every 15 s for 10 min and stimulation was initiated 2 min after starting image acquisition.
-Imaging data analysis was done with ImageJ (National Institutes of Health, Bethesda, MD, USA).
+Imaging data were processed with ImageJ (National Institutes of Health, Bethesda, MD, USA).
 As NO donor, we used S-Nitroso-N-acetyl-D, L-penicillamine (SNAP, Sigma-Aldrich, St.
@@
 The synaptic signal $S(t)$ and all terms depending on it are removed due to the assumed dependence on NIT-GC2.
-When the $S(t)$ input is disregarded, the activity of the INNOS cells ($C_N(t)$ and $N(t)$ variables) can be represented by a constant steady state whose values can subsequently be absorbed into all of the terms depending on these variables.
+When the $S(t)$ input is disregarded, the activity of the INNOS cells ($C_N(t)$ and $N(t)$ variables) can be represented by a constant steady state whose values can subsequently be absorbed into all the terms depending on these variables.
 Hence these variables are eliminated in the reduced model.
@@
 
-To find sets of model parameters that reproduce recordings from the Ca^2+^ imaging experiments, we establish an optimisation problem to minimise the discrepancy between the modelled Ca^2+^ trajectories and the experimental data.
-To obtain the modelled Ca^2+^ trajectories, we simulated the model using the Matlab built-in solvers \texttt{fsolve} to find the steady state in the absence of UV stimulation, and then use the explicit integration algorithm \texttt{ode45} to simulate the model using the computed steady state as the initial condition.
+To find sets of model parameters that reproduce recordings from the Ca^2+^ imaging experiments, we define an optimisation problem to minimise the discrepancy between the modelled Ca^2+^ traces and the experimental data.
+To obtain the modelled Ca^2+^ traces, we run the model using the Matlab built-in solvers \texttt{fsolve} to find the steady state in the absence of UV stimulation, and then use the explicit integration algorithm \texttt{ode45} to simulate the model with the computed steady state as the initial condition.
 
@@
 \end{equation*} where the timepoints $t_i = i \Delta t$ are equispaced with $\Delta = 0.5s$ and where $t_{k_{UV\: start}} = t_{UV \: start}$ (and similar for the other time points).
-Here ${C}_X^{data}$ denotes normalised experimental data and ${C}_X^{model}$ denotes normalised simulated Ca^2+^ trajectory with $X \in \{ P, N, R\}$ indicating cell type.
-Simulated Ca^2+^ trajectories, ${C}_X^{model}$, are normalised in the same way as the experimental data, i.e., we divide values uniformly by their value at $t_{k_{UV\: start}-1}$, a time point immediately prior to the start of the UV stimulation.
+Here ${C}_X^{data}$ denotes normalised experimental data and ${C}_X^{model}$ denotes normalised simulated Ca^2+^ trace with $X \in \{ P, N, R\}$ indicating cell type.
+Simulated Ca^2+^ traces, ${C}_X^{model}$, are normalised in the same way as the experimental data, i.e., we divide values uniformly by their value at $t_{k_{UV\: start}-1}$, a time point immediately prior to the start of the UV stimulation.
 Simulations are sampled with the same sampling rate as the data.
@@
 To solve the optimisation problem with the above objective function, we employ a global optimisation method known as a genetic algorithm (GA).
-GAs are a class of iterative, gradient-free algorithms inspired by processes of natural selection including breeding and mutation, amongst others \cite{GAbook}.
+GAs are a class of iterative, gradient-free algorithms inspired by processes of natural selection including breeding and mutation \cite{GAbook}.
 In reference to natural selection, sets of candidate solutions in GAs are referred to as \`populations'.
-As our algorithm of choice, we use the Matlab implementation of a genetic algorithm for multiobjective optimisation, \texttt{gamultiobj}.
+As our algorithm of choice, we use the Matlab implementation of a genetic algorithm for multiobjective optimisation \texttt{gamultiobj}.
 The \texttt{gamultiobj} algorithm uses a controlled, elitist genetic algorithm (a variant of NSGA-II \cite{GAbook}).
@@
 Parameter values in these children can also be changed by 'mutations', which perturb said values randomly by a small amount.
-Breeding and mutation introduce a level of randomness to the optimisation process that allows for exploration of the objective function (fitting) landscape.
+Breeding and mutation introduce a level of randomness to the optimisation process that allows exploring the objective function (fitting) landscape.
 The stochastic nature of GAs helps them to avoid local minima that can plague gradient-descent-based optimisation methods.
@@
 A subset of the size of the initial population is then selected as a new parent population based on the scores and the \textit{Crowding Distance}, distance to other parameter sets in terms of a sum of element-wise differences between values of the objective function (used to ensure diversity of the parent population)\cite{crowding_distance,GAbook}.
-This iterative process is repeated until set convergence criteria are met, indicating that the GA has found a set of model parameters that best reproduce the data under consideration.
+This iterative process is repeated until the convergence criteria are met, indicating that the GA has found a set of model parameters that best reproduce the data under consideration.
 
@@
 We assume that the inflection is caused by the input from the INNOS cells.
--   We then compute the $AUP$ as an integral over $C_P(t)$ starting at the inflection point and ending 125 seconds later: \$AUP=\int*{t*{start}}\^{t\_{end}} C_P(t) dt \$.
+-   We then compute the $AUP$ as an integral over $C_P(t)$ starting at the inflection point and ending 125 seconds later: \$AUP=\int\*{t\*{start}}\^{t\_{end}} C_P(t) dt \$.
 To compute this, we use numerical integration via the trapezoidal method as implemented in Matlab function \texttt{trapz}.
@@
 -   To capture possible variability between different cells/organisms, we compute the $AUP$ using 47 model parameter sets fitted to individual WT cPRC experimental recordings.
-We remove outliers, defined as being any simulated trajectories for which $C_P(t)$, after normalisation, exits the interval [0.25, 4] at any time.
+We remove outliers, defined as being any simulated traces for which $C_P(t)$, after normalisation, exits the interval [0.25, 4] at any time.
 This $C_P(t)$ range was decided of the basis of values observed in various experiments in the lab.
@@
 
-Figure 6 shows how the results of the simulations of the WT model depend on the amplitude and duration of UV stimulation, as captured by the median values of $AUP$.
+Figure 7 shows how the results of the simulations of the WT model depend on the amplitude and duration of UV stimulation, as captured by the median values of $AUP$.
 Delayed activation occurs only for UV stimulation of sufficient amplitude and duration.
@@
 Maximum-likelihood phylogenetic tree of NOS protein sequences.** Individual branches are coloured by taxonomy.
-Support values are based on 1000 replicates of ultrafast bootstrap with the aLRT-SH-like and aBayes methods."}
+Support values are based on 1000 replicates of ultrafast bootstrap with the aLRT-SH-like and aBayes methods.
+<br> Figure 1 -- figure supplement 1 -- source data 1.
+NOS sequences used for the phylogenetic reconstruction.
+<br> Figure 1 -- figure supplement 1 -- source data 2.
+Aligned and trimmed NOS sequences used for the phylogenetic reconstruction.
+<br> Figure 1 -- figure supplement 1 -- source data 3.
+Tree file of the reconstructed NOS phylogeny.
+"}
 knitr::include_graphics(path = "figures/Fig1_sup1.png")
@@
 
-Maximum-likelihood phylogenetic tree of NOS protein sequences.\*\* Individual branches are coloured by taxonomy.
-Support values are based on 1000 replicates of ultrafast bootstrap with the aLRT-SH-like and aBayes methods.
-
 ```{r echo=FALSE,fig.cap="**Figure 4 -- figure supplement 2.
@@
 Dot size indicates specificity (percent of transcripts expressed in the indicated cell across all cells).
-Dot colour represents the logarithm of the number of transcripts in the expressing cells."}
+Dot colour represents the logarithm of the number of transcripts in the expressing cells.
+<br> Figure 4 -- figure supplement 2 -- source data 1.
+GC sequences used for the phylogenetic reconstruction.
+<br> Figure 4 -- figure supplement 2 -- source data 2.
+Aligned and trimmed GC sequences used for the phylogenetic reconstruction.
+<br> Figure 4 -- figure supplement 3 -- source data 3.
+Tree file of the reconstructed GC phylogeny.
+"}
 knitr::include_graphics(path = "figures/Fig4_sup2.png")
@@
 ```{r echo=FALSE,fig.cap="**Figure 6 -- figure supplement 1.
-Simulated Ca^2+^ trajectories for parameter sets fitted to individual Ca^2+^-recordings collected in wild type, *NOS* knockout, and *NIT-GC2* morphant larvae.** (**A-C**) Simulated Ca^2+^ trajectories in cPRC (A), INNOS (B) and INRGW (C) cells in the wild-type condition.
-(**D-F**) Simulated Ca^2+^ trajectories in cPRC (D), INNOS (E) and INRGW (F) cells in the *NOS*-knockout condition.
-(**G-I**) Simulated Ca^2+^ trajectories in cPRC (G), INNOS (H) and INRGW (I) cells in the *NIT-GC2*-morphant condition.
-Thin coloured curves indicate individual recordings (cPRC - purple, INNOS - blue and INRGW - green), thin grey curves indicate simulated Ca^2+^ trajectories based on model parameters fitted to the individual recording, thick coloured curves indicate averages of the recordings and thick black curves represent the average of the fits.
+Simulated Ca^2+^ traces for parameter sets fitted to individual Ca^2+^-recordings collected in wild type, *NOS* knockout, and *NIT-GC2* morphant larvae.** (**A-C**) Simulated Ca^2+^ traces in cPRC (A), INNOS (B) and INRGW (C) cells in the wild-type condition.
+(**D-F**) Simulated Ca^2+^ traces in cPRC (D), INNOS (E) and INRGW (F) cells in the *NOS*-knockout condition.
+(**G-I**) Simulated Ca^2+^ traces in cPRC (G), INNOS (H) and INRGW (I) cells in the *NIT-GC2*-morphant condition.
+Thin coloured curves indicate individual recordings (cPRC - purple, INNOS - blue and INRGW - green), thin grey curves indicate simulated Ca^2+^ traces based on model parameters fitted to the individual recording, thick coloured curves indicate averages of the recordings and thick black curves represent the average of the fits.
 "}
@@
 ```{r echo=FALSE,fig.cap="**Figure 6 -- figure supplement 2.
-Comparison of fits with different fixed parameter sets when fitting the WT model to individual recordings from the WT INRGW cells.** Panels A-B, D-E, G-H  (column fixed parameters) show simulated trajectories of the Ca^2+^ levels in cPRC and INNOS cells produced by different fixed parameter sets.
+Comparison of fits with different fixed parameter sets when fitting the WT model to individual recordings from the WT INRGW cells.** Panels A-B, D-E, G-H  (column fixed parameters) show simulated traces of the Ca^2+^ levels in cPRC and INNOS cells produced by different fixed parameter sets.
 Panels C, F and I show individual recordings fitted to the data from INRGW cells."}
@@
 ```{r echo=FALSE,fig.cap="**Figure 6 -- figure supplement 3.
-Comparison of fits to the cPRC recordings from NIT-GC2 morphant larvae using the NIT-GC2 morpholino model or the NOS-knockout model.**  Thin coloured curves indicate individual recordings, thin grey curves are simulated Ca^2+^ trajectories, thick coloured curves indicate averages of the recordings and thick black curves represent the mean of the fits."}
+Comparison of fits to the cPRC recordings from NIT-GC2 morphant larvae using the NIT-GC2 morpholino model or the NOS-knockout model.**  Thin coloured curves indicate individual recordings, thin grey curves are simulated Ca^2+^ traces, thick coloured curves indicate averages of the recordings and thick black curves represent the mean of the fits."}
 knitr::include_graphics(path = "figures/Figure6_fig_Suppl3.svg")

### 2023-07-03 Manuscript/Manuscript.Rmd
Ms: added source data refs to all figs, added commit

@@
 Thick lines show the mean value with 0.95 confidence intervals.
-N = 9 larvae for control and 11 for cPRC stimulation."}
+N = 9 larvae for control and 11 for cPRC stimulation.
+<br> Figure 2 -- source data 1.
+DAF-FM fluorescenc reads."}
 
@@
 (**D**) Vertical displacement in 30 sec bins of wild type and mutant (NOSΔ11 and NOSΔ23) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
-(**E**) Vertical displacement in 30 sec bins of control and L-NAME-treated (0.1 and 1 mM) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top."}
+(**E**) Vertical displacement in 30 sec bins of control and L-NAME-treated (0.1 and 1 mM) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
+<br> Figure 3 -- source data 1-5.
+Source data for panels A-E."}
 
@@
 (**M**, **N**) GCaMP6s signals in cPRCs in (M) NIT-GC1 and (N) NIT-GC2 morphant larvae during 405 nm light stimulation.
-Individual responses and their mean with 0.95 confidence interval are shown."}
+Individual responses and their mean with 0.95 confidence interval are shown.
+<br> Figure 4 -- source data 1-8.
+Source data for panels A, B and I-N."}
 
@@
 **(G)** GCaMP6s fluorescence in INRGWa cells in NIT-GC2 morphant larvae during 405 nm stimulation.
-**(H, I)** GCaMP6s fluorescence in (H) Ser-h1 cells and (I) the MC cell in wild type and NOSΔ11/Δ11 mutant larvae during 405 nm stimulation."}
+**(H, I)** GCaMP6s fluorescence in (H) Ser-h1 cells and (I) the MC cell in wild type and NOSΔ11/Δ11 mutant larvae during 405 nm stimulation.
+<br> Figure 5 -- source data 1-6.
+Source data for panels D-I."}
 
@@
 All figures were assembled in R with the cowplot and patchwork packages.
-All scripts are available at <https://github.com/JekelyLab/Joukra_et_al_NOS> (commit ).
+All scripts are available at <https://github.com/JekelyLab/Joukra_et_al_NOS> (commit: c223fe3).
 
@@
 The starting position of each larval trajectory was set to 0.
-(**E-G**) Swimming speed of batches of wild type and mutant two-day old (E) and three-day-old (F) larvae and L-NAME-treated (NOS inhibitor) three-day-old larvae under 395 nm UV stimulation."}
+(**E-G**) Swimming speed of batches of wild type and mutant two-day old (E) and three-day-old (F) larvae and L-NAME-treated (NOS inhibitor) three-day-old larvae under 395 nm UV stimulation.
+<br> Figure 3 -- figure suppmenet 1 -- source data 1-6.
+Source data for panels B-G."}
 knitr::include_graphics(path = "figures/Fig3_sup1.png")

### 2023-07-04 Manuscript/Manuscript.Rmd
Update by Martin comments

@@
 
-**CRISPR-Cas9 Design and Microinjection**
+**CRISPR-Cas9 design and Microinjection**
 
@@
 A mixture of wild-type and deletion alleles in a sample gave double peaks in the sequencing chromatograms, with the relative height of the double peaks reflecting the relative allele ratio in the sample.
+
+**Morpholino design**
+
+Morpholino injections were performed as previously described [@Conzelmann2013].
+We used the following morpholinos to target the NIT-GC1 and NIT-GC2 genes (GeneTools, LLC).
+NIT-GC1 MO1: TGCTTGTCATTATTCAACCAGCAAA, NIT-GC1 MO2: TTCAATTAAACCCTCCAGGTTGCTG, NIT-GC2 MO1: AAATGAAGAGAGGTGTTCTCCTTCG, NIT-GC2 MO2: ATATTCATTATGTGAAGAACTTCCA.
 
@@
 From these sequences, we generated a Hidden Markov Model (HMM) to mine the genomes and transcriptomes of 47 metazoan, 2 choanoflagellate and 2 filasterea species.
-The HMM model was run in HMMR3 with an e-value of 1e−15.
+The HMM model was run in HMMR3 [@Eddy2011] with an e-value of 1e−15.
 We ran CD-Hit [@Fu2012] to eliminate redundant sequences (at a 80% threshold).
-We aligned the sequences with MAFFT version 7, with the iterative refinement method E-INS-i.
+We aligned the sequences with MAFFT version 7, with the iterative refinement method E-INS-i [@Katoh2008].
 The alignment was trimmed with TrimAl in gappy-out mode [@Capella-Gutierrez2009].
 To calculate a maximum-likelihood tree, we used IQ-tree2 with the LG+G4 model.
-To calculate branch support, we ran 1000 replicates with the aLRT-SH-like [@Guindon2010] and aBayes methods [@Minh2020].
+To calculate branch support, we ran 1000 replicates with the ultrafast bootstrap approximation (UFBoot) [@Hoang2018] and aLRT-SH-like values [@Guindon2010] methods.
 
@@
 To analyze the phylogeny of NIT-GCs, the cluster containing these GCs together with membrane-bound guanylate cyclases were parsed and used for tree building.
-We aligned the sequences with MAFFT version 7, with the iterative refinement method E-INS-i[@Katoh2013].
+We aligned the sequences with MAFFT version 7, with the iterative refinement method E-INS-i [@Katoh2008].
 The alignment was trimmed with TrimAl in gappy-out mode [@Capella-Gutierrez2009].
 To calculate a maximum-likelihood tree, we used IQ-tree2 with the LG+G4 model.
-To calculate branch support, we ran 1000 replicates with the aLRT-SH-like and aBayes methods [@Minh2020].
+To calculate branch support, we ran 1000 replicates with the ultrafast bootstrap approximation (UFBoot) [@Hoang2018] and aLRT-SH-like values [@Guindon2010] methods.
 

### 2023-07-06 Manuscript/Manuscript.Rmd
MS: methods edit
Edit Figure 1 – figure supplement 1 legend
Ms: added iqtree command, updated bootstrap explanation

@@
 
-Through this optimisation procedure, we found parameters that gave a fit to our experimental data under all conditions.
-We could fit to mean Ca^2+^ profiles but also individual Ca^2+^ transients (Figure 6B, Figure 6--figure supplement 1).
+Through this optimisation procedure, we found parameters that gave a fit to our experimental data under all conditions (Figure 6B; Figure 6--figure supplement 1-5).
+We could fit to mean Ca^2+^ profiles but also individual Ca^2+^ transients.
 Our model only includes a minimal set of parameters and assumptions of interactions that are required to describe the dynamics of cPRC, INNOS and INRGWa in wild type and loss-of-function conditions (Figure 6A, B).
@@
 
-**Morpholino design**
+**Morpholino-mediated knockdowns**
 
@@
 The alignment was trimmed with TrimAl in gappy-out mode [@Capella-Gutierrez2009].
-To calculate a maximum-likelihood tree, we used IQ-tree2 with the LG+G4 model.
-To calculate branch support, we ran 1000 replicates with the ultrafast bootstrap approximation (UFBoot) [@Hoang2018] and aLRT-SH-like values [@Guindon2010] methods.
+To calculate a maximum-likelihood tree, we used IQ-tree2 with the LG+G4 model with the following options: iqtree2 -s "FILENAME" -mset WAG, LG, Blosum62, Dayhoff, JTT, Poisson -B 1000 -alrt 1000 -T AUTO.
+Branch-support values are based on 1000 replicates with the ultrafast bootstrap approximation (UFBoot) [@Hoang2018] and aLRT-SH-like [@Guindon2010] methods.
 
@@
 Maximum-likelihood phylogenetic tree of NOS protein sequences.** Individual branches are coloured by taxonomy.
-Support values are based on 1000 replicates of ultrafast bootstrap with the aLRT-SH-like and aBayes methods.
+Branch support values indicate UFBoot and aLRT-SH-like values.
 <br> Figure 1 -- figure supplement 1 -- source data 1.
@@
 Guanylate cyclases with NIT domains are found in most animal phyla except Porifera, Ctenophora, Urochordata and Chordata.
+Branch support values indicate UFBoot and aLRT-SH-like values.
 The expression of *Platynereis* NIT-GC genes in cPRC, INNOS and INRGW cells is indicated on the right side of the tree.
-The values represent expression in single-cell sequencing data.
+The values represent expression based on single-cell sequencing data.
 Dot size indicates specificity (percent of transcripts expressed in the indicated cell across all cells).

### 2023-07-10 Manuscript/Manuscript.Rmd
Ms: edits

@@
 Nitric oxide (NO) is one such modulator with unique physical and signalling properties.
-This free radical synthesized from L-arginine by nitric oxide synthase (NOS), is short-lived and can diffuse across biological membranes [@Cudeiro1999; @thomas2015].
+This free radical synthesized from L-arginine by nitric oxide synthase (NOS) is short-lived and can diffuse across biological membranes [@Cudeiro1999; @thomas2015].
 Canonical NO signalling involves the Ca^2+^/calmodulin-dependent activation of NOS, NO production and diffusion, and the NO-dependent activation of soluble guanylate cyclases (sGC) leading to cGMP production [@Bredt1990; @Hoelscher1997].
@@
 Upon UV/violet exposure, the cPRCs show a characteristic biphasic Ca^2+^ response that is c-opsin1 dependent.
-UV/violet avoidance is also *c-opsin1*-dependent and is defective in *c-opsin1* mutants [@Veraszto2018].
-Here we show that *Platynereis* *NOS* is expressed in interneurons of the cPRC circuit and is required for UV/violet-avoidance.
+UV/violet avoidance is also c-opsin1 dependent and is defective in *c-opsin1* mutants [@Veraszto2018].
+Here we show that *Platynereis* NOS is expressed in interneurons of the cPRC circuit and is required for UV/violet-avoidance.
 By combining Ca^2+^ imaging across the fully-mapped cPRC circuit [@Veraszto2018] with genetic perturbations and mathematical modelling, we describe how NO tunes circuit dynamics through non-synaptic retrograde signalling to cPRCs.
@@
 *NOS* was also expressed in the region of the visual eyes (adult eyes) and the pigmented eyespots (Figure 1---figure supplement 2).
-The four apical organ cells, but not the eyes, were also labelled with a *NOS* reporter construct driving palmitoylated-tdTomato (Figure 1E).
+The four apical organ cells, but not the eyes, were also labelled with a *NOS* reporter construct driving palmitoylated tdTomato (Figure 1E).
 This reporter also revealed the axonal projections of these central NOS-expressing neurons.
@@
 The projections of INNOS cells are segregated into input (dendritic) and output (axonal) compartments, with cPRC inputs occurring on the dendritic part in the dense neurosecretory plexus.
-INNOS output synapses form in the more ventral projection region (Figure 1---figure supplement 3).
+INNOS output-synapses form in the more ventral projection region (Figure 1---figure supplement 3).
 INRGWa neurons synapse on the head serotonergic ciliomotor neurons (Ser-h1), which synapse on the cholinergic MC ciliomotor neuron and cells of the prototroch ciliary band (Figure 1C, F) [@veraszto2017].
@@
 ```{r echo=FALSE,fig.cap="**Figure 1.
-Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** **(A)** Scanning electron microscopy image of a three-day-old *Platynereis* larva.
+Identification of *NOS*-expressing interneurons (INNOS) within the cPRC circuit.** **(A)** Scanning electron microscopy image of a three-day-old *Platynereis* larva.
 **(B, C)** Volume rendering of the neuron types (cPRC, INNOS, INRGWa, Ser-h1 and MC) in the cPRC circuit reconstructed from a whole-body transmission electron microscopy volume of a three-day-old larva.
@@
 Antibody staining for acetylated α-tubulin (acTub: green) highlights cPRC cilia and the neuropil.
-**(E)** Expression of a *NOS* reporter (NOSp::palmi-3xHA: magenta) labelled with an anti-HA antibody in a two-day-old larva (anterior view).
-Antibody staining for acetylated α-tubulin (acTub: green) highlights cPRC cilia and the neuropil.
+**(E)** Expression of a *NOS* reporter (NOSp::palmi-3xHA; magenta) labelled with an anti-HA antibody in a two-day-old larva (anterior view).
+Antibody staining for acetylated α-tubulin (acTub; green) highlights cPRC cilia and the neuropil.
 **(F)** Synaptic wiring diagram of the cPRC circuit.
@@
 The cPRCs are marked by thin lines.
-(**B**) Changes in DAF-FM fluorescence before and during 405 nm light stimulation.
-(**C**) Changes in DAF-FM fluorescence over time during 405 nm stimulation of the cPRCs (green) or a control area (ctr: gray).
+(**B**) DAF-FM fluorescence before and during 405 nm light stimulation.
+(**C**) Changes in DAF-FM fluorescence over time during 405 nm stimulation of the cPRCs or a control area (ctr stim.).
 The purple box indicates the duration of 405 nm stimulation.
-Individual traces normalized (ΔF/F0) are shown as thin lines.
+Individual normalized traces (ΔF/F0) are shown as thin lines.
 Thick lines show the mean value with 0.95 confidence intervals.
@@
 <br> Figure 2 -- source data 1.
-DAF-FM fluorescenc reads."}
+DAF-FM fluorescence reads."}
 
@@
 ```{r echo=FALSE,fig.cap="**Figure 3.
-NOS is required for UV avoidance in *Platynereis* larvae.**  (**A**) Swimming trajectories of wild type (WT, n=32) and NOS mutant (NOSΔ11/Δ11, n=26 and NOSΔ23/Δ23, n=47) three-day-old larvae.
+NOS is required for UV avoidance in *Platynereis* larvae.**  (**A**) Swimming trajectories of wild type (WT, n=32) and *NOS* mutant (*NOSΔ11/Δ11*, n=26 and *NOSΔ23/Δ23*, n=47) three-day-old larvae.
 All trajectories start at 0 x and y position and time 0 corresponding to 10 sec after the onset of 395 nm stimulation from the side.
@@
 The starting position of each larval trajectory was set to 0.
-(**D**) Vertical displacement in 30 sec bins of wild type and mutant (NOSΔ11 and NOSΔ23) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
+(**D**) Vertical displacement in 30 sec bins of wild type and mutant (*NOSΔ11/Δ11* and *NOSΔ23/Δ23*) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
 (**E**) Vertical displacement in 30 sec bins of control and L-NAME-treated (0.1 and 1 mM) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
@@
 
-We aimed next to identify the NO receptor in the cPRCs.
+Next we aimed to identify the NO receptor in the cPRCs.
 NO generally acts via soluble guanylate cyclases (sGC), belonging to the guanylate cyclase family with a CYC domain (PFAM domain: PF00211).
@@
 
-To characterise the expression of NIT-GCs, we used previously published spatially mapped single-cell transcriptome data [@Achim2015; @williams2017].
+To characterise the expression of NIT-GCs, we used previously published spatially-mapped single-cell transcriptome data [@Achim2015; @williams2017].
 Among the 15 NIT-GCs, two showed high and specific expression in the cPRCs and one was expressed in the INNOS cells (Figure 4---figure supplement 2).
@@
 ```{r echo=FALSE,fig.cap="**Figure 4.
-NOS and two NIT-GCs shape Ca^2+^ signals during cPRC UV/violet response.** (**A**, **B**) GCaMP6s signals in cPRCs in wild type and *NOS* mutant (A, NOSΔ11/Δ11, B, NOSΔ23/Δ23) larvae during 405 nm light stimulation.
+NOS and two NIT-GCs shape Ca^2+^ signals during cPRC UV/violet response.** (**A**, **B**) GCaMP6s signals in cPRCs in wild type and *NOS* mutant (A, *NOSΔ11/Δ11*, B, *NOSΔ23/Δ23*) larvae during 405 nm light stimulation.
 (**C**, **D**) In situ HCR for (C) *NIT-GC1* and (D) *NIT-GC2* (magenta) in three-day-old *Platynereis* larvae.
@@
 ```{r echo=FALSE,fig.cap="**Figure 5.
-NOS- and NIT-GC2-dependent dynamics of the cPRC circuit.** **(A, B)** GCaMP6s imaging from cPRCs and INNOS cells (left panels) followed by on-slide immunostaining for (A) RYamide to label INNOS and (B) RGWamide+serotonin to label INRGWa and Ser-h1 (red).
+NOS- and NIT-GC2-dependent dynamics of the cPRC circuit.** (**A, B**) GCaMP6s imaging from cPRCs and INNOS cells (left panels) followed by on-slide immunostaining for (A) RYamide to label INNOS and (B) RGWamide+serotonin to label INRGWa and Ser-h1 (red).
 Nuclei are stained with DAPI (cyan).
@@
 Numbers mark the same cells in the GCaMP and immunostaining images matched by position.
-**(C)** Correlation map of neuronal activity of the cPRCs, INNOS, INRGWa, Ser-h1 and MC neurons.
-**(D)** GCaMP6s fluorescence in INNOS cells in wild type (WT) and NOSΔ11/Δ11 mutant larvae during 405 nm stimulation of the cPRC cilia.
-**(E)** GCaMP6s fluorescence in INNOS cells in NIT-GC2 morphant larvae during 405 nm stimulation.
-**(F)** GCaMP6s fluorescence in INRGWa cells in wild type and NOSΔ23/Δ23 mutant larvae during 405 nm stimulation.
-**(G)** GCaMP6s fluorescence in INRGWa cells in NIT-GC2 morphant larvae during 405 nm stimulation.
-**(H, I)** GCaMP6s fluorescence in (H) Ser-h1 cells and (I) the MC cell in wild type and NOSΔ11/Δ11 mutant larvae during 405 nm stimulation.
+(**C**) Correlation map of neuronal activity of the cPRCs, INNOS, INRGWa, Ser-h1 and MC neurons.
+(**D**) GCaMP6s fluorescence in INNOS cells in wild type (WT) and *NOSΔ11/Δ11* mutant larvae during 405 nm stimulation of the cPRC cilia.
+(**E**) GCaMP6s fluorescence in INNOS cells in NIT-GC2 morphant larvae during 405 nm stimulation.
+(**F**) GCaMP6s fluorescence in INRGWa cells in wild type and *NOSΔ23/Δ23* mutant larvae during 405 nm stimulation.
+(**G**) GCaMP6s fluorescence in INRGWa cells in NIT-GC2 morphant larvae during 405 nm stimulation.
+(**H, I**) GCaMP6s fluorescence in (H) Ser-h1 cells and (I) the MC cell in wild type and *NOSΔ11/Δ11* mutant larvae during 405 nm stimulation.
 <br> Figure 5 -- source data 1-6.

### 2023-07-22 Manuscript/Manuscript.tex
Ms: knit

@@
 \usepackage{amsmath,amssymb}
-\usepackage{lmodern}
 \usepackage{setspace}
@@
 \else % if luatex or xetex
-  \usepackage{unicode-math}
+  \usepackage{unicode-math} % this also loads fontspec
   \defaultfontfeatures{Scale=MatchLowercase}
   \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else
+  % xetex/luatex font selection
   \setmainfont[]{Arial}
@@
 \makeatother
+\usepackage{svg}
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
+\urlstyle{same}
 \hypersetup{
@@
 signals retrogradely to cPRCs to induce their sustained post-stimulus
-activation through an unconventional guanylyl cyclase. This late
+activation through an unconventional guanylate cyclase. This late
 activation inhibits serotonergic ciliomotor neurons to induce downward
@@
 biological membranes (Cudeiro and Rivadulla, 1999; Thomas, 2015).
-Canonical NO signalling involves the Ca2+/calmodulin-dependent
-activation of NOS, NO production and diffusion, and the NO-dependent
-activation of soluble guanylate cyclases (sGC). This can occur
-cell-autonomously or in other cells leading to cGMP production (Bredt et
-al., 1990; Hölscher, 1997). Given that NOS activation is calcium
+Canonical NO signalling involves the
+Ca\textsuperscript{2+}/calmodulin-dependent activation of NOS, NO
+production and diffusion, and the NO-dependent activation of soluble
+guanylate cyclases (sGC) leading to cGMP production (Bredt et al., 1990;
+Hölscher, 1997). Given that NOS activation is Ca\textsuperscript{2+}
 dependent and NOS shows neuron-type-specific expression (Aso et al.,
@@
 1999; Jacoby et al., 2018; Wang et al., 2007). For example, defective NO
-signalling in NOS knockouts leads to a decreased sensitivity of retinal
-ganglion cells to light stimulation (Wang et al., 2007). In the retina,
-NO signalling can also involve pathways other than canonical sGC
+signalling in NOS knockout mice leads to a decreased sensitivity of
+retinal ganglion cells to light stimulation (Wang et al., 2007). In the
+retina, NO signalling can also involve pathways other than canonical sGC
 signalling (Jacoby et al., 2018; Tooker et al., 2013; Wei et al., 2012).
@@
 
-Here we uncovered an essential function for NO signalling in larval
+We uncovered an essential function for NO signalling in larval
 UV/violet-light avoidance behaviour. In \emph{Platynereis} larvae,
@@
 maximum around 384 nm (Tsukamoto et al., 2017; Veedin Rajan et al.,
-2021; Verasztó et al., 2018). Upon UV/violet exposure, the cPRC show a
-characteristic biphasic calcium response that is \emph{c-opsin1}
-dependent. UV/violet avoidance is also \emph{c-opsin1}-dependent and is
+2021; Verasztó et al., 2018). Upon UV/violet exposure, the cPRCs show a
+characteristic biphasic Ca\textsuperscript{2+} response that is c-opsin1
+dependent. UV/violet avoidance is also c-opsin1 dependent and is
 defective in \emph{c-opsin1} mutants (Verasztó et al., 2018). Here we
-show that \emph{Platynereis} \emph{NOS} is expressed in interneurons of
-the cPRC circuit and is required for UV/violet-avoidance. By combining
-calcium imaging across the fully-mapped cPRC circuit (Verasztó et al.,
-2018) with genetic perturbations and mathematical modelling, we describe
-how NO tunes circuit dynamics through non-synaptic retrograde signalling
-to cPRCs. This delayed neuroendocrine feedback integrates UV/violet
-exposure to induce a short-term memory manifested in altered circuit
-activity and an aversive behavioural response.
+show that \emph{Platynereis} NOS is expressed in interneurons of the
+cPRC circuit and is required for UV/violet-avoidance. By combining
+Ca\textsuperscript{2+} imaging across the fully-mapped cPRC circuit
+(Verasztó et al., 2018) with genetic perturbations and mathematical
+modelling, we describe how NO tunes circuit dynamics through
+non-synaptic retrograde signalling to cPRCs. This delayed neuroendocrine
+feedback integrates UV/violet exposure to induce a short-term memory
+manifested in altered circuit activity and an aversive behavioural
+response.
 
@@
 1---figure supplement 1). To characterise the expression pattern of
-\emph{NOS} we used in situ hybridization chain reaction (HCR) and
+\emph{NOS}, we used in situ hybridization chain reaction (HCR) and
 transient transgenesis. In two- and three-day-old larvae, we detected
@@
 four apical organ cells, but not the eyes, were also labelled with a
-\emph{NOS} reporter construct driving palmitoylated-tdTomato (Figure
+\emph{NOS} reporter construct driving palmitoylated tdTomato (Figure
 1E). This reporter also revealed the axonal projections of these central
@@
 the UV-sensory cPRCs and presynaptic to the INRGWa interneurons, which
-are also cPRC targets (Figure 1C,F). The projections of INNOS cells are
+are also cPRC targets (Figure 1C, F). The projections of INNOS cells are
 segregated into input (dendritic) and output (axonal) compartments, with
 cPRC inputs occurring on the dendritic part in the dense neurosecretory
-plexus. INNOS output synapses form in the more ventral projection region
+plexus. INNOS output-synapses form in the more ventral projection region
 (Figure 1---figure supplement 3). INRGWa neurons synapse on the head
 serotonergic ciliomotor neurons (Ser-h1), which synapse on the
-prototroch ciliary band and the cholinergic MC ciliomotor neurons
-(Figure 1C,F) (Verasztó et al., 2017).
+cholinergic MC ciliomotor neuron and cells of the prototroch ciliary
+band (Figure 1C, F) (Verasztó et al., 2017).
 
@@
 \includegraphics[width=25in]{figures/Fig1} \caption{**Figure 1.
-Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** **(A)** Scanning electron microscopy image of a three-day-old *Platynereis* larva.
+Identification of *NOS*-expressing interneurons (INNOS) within the cPRC circuit.** **(A)** Scanning electron microscopy image of a three-day-old *Platynereis* larva.
 **(B, C)** Volume rendering of the neuron types (cPRC, INNOS, INRGWa, Ser-h1 and MC) in the cPRC circuit reconstructed from a whole-body transmission electron microscopy volume of a three-day-old larva.
@@
 neurosecretory neuropil, the region of INNOS projections. This increase
-did not occur in larvae where we illuminated a control area (Figure 2).
+did not occur in larvae in which we illuminated a control area (Figure
+2).
 
@@
 \includegraphics[width=29.17in]{figures/Fig2} \caption{**Figure 2.
-NO produced by UV/violet stimulation to cPRCs.** **(A)** DAF-FM fluorescence in the region of the neurosecretory neuropil.
+NO produced by UV/violet stimulation to cPRCs.** (**A**) DAF-FM fluorescence in the region of the neurosecretory neuropil.
 The white line indicates the outline of the larva.
 The dashed line corresponds to the area where fluorescence was quantified.
-Circles indicate the location of cPRC and control stimulation.
-cPRCs are marked by thin lines.
-**(B)** Changes in DAF-FM fluorescence before and during 405 nm light stimulation.
-**(C)** Changes in DAF-FM fluorescence over time during 405 nm stimulation of the cPRCs (green) or a control area (ctr: gray).
+The circles indicate the location of cPRC and control stimulation.
+The cPRCs are marked by thin lines.
+(**B**) DAF-FM fluorescence before and during 405 nm light stimulation.
+(**C**) Changes in DAF-FM fluorescence over time during 405 nm stimulation of the cPRCs or a control area (ctr stim.).
 The purple box indicates the duration of 405 nm stimulation.
-Individual traces normalized (ΔF/F0) are shown as thin lines.
+Individual normalized traces (ΔF/F0) are shown as thin lines.
 Thick lines show the mean value with 0.95 confidence intervals.
-N = 9 larvae for control and 11 for cPRC stimulation.}\label{fig:unnamed-chunk-2}
+N = 9 larvae for control and 11 for cPRC stimulation.
+<br> Figure 2 -- source data 1.
+DAF-FM fluorescence reads.}\label{fig:unnamed-chunk-2}
 \end{figure}
@@
 achieve this, we generated two \emph{Platynereis} \emph{NOS} knockout
-lines with the CRISPR/Cas9 method. We recovered two deletions
+lines using the CRISPR/Cas9 method. We recovered two deletions
 (NOSΔ11/Δ11 and Δ23/Δ23), both frame-shift mutations leading to an early
@@
 \emph{NOS}-mutant larvae showed a strongly diminished UV-avoidance
-response (Figure 3A, B and Figure 3---figure supplement 1B,C). This
+response (Figure 3A, B and Figure 3---figure supplement 1B, C). This
 phenotype is similar to the defective UV-avoidance of \emph{c-opsin1}
@@
 To distinguish between an acute and developmental function of NOS in
-light responses, we next tested larvae exposed to the NOS inhibitor
+light responses, we also tested larvae exposed to the NOS inhibitor
 L-NAME. Larvae incubated for 5 min in 0.1 mM or 1 mM L-NAME showed a
@@
 \includegraphics[width=33.33in]{figures/Fig3} \caption{**Figure 3.
-NOS is required for UV avoidance in *Platynereis* larvae.**  **(B)** Swimming trajectories of wild type (WT, n=32) and NOS mutant (NOSΔ11/Δ11, n=26 and NOSΔ23/Δ23, n=47) three-day-old larvae.
+NOS is required for UV avoidance in *Platynereis* larvae.**  (**A**) Swimming trajectories of wild type (WT, n=32) and *NOS* mutant (*NOSΔ11/Δ11*, n=26 and *NOSΔ23/Δ23*, n=47) three-day-old larvae.
 All trajectories start at 0 x and y position and time 0 corresponding to 10 sec after the onset of 395 nm stimulation from the side.
-**(B)** Vertical position of batches of wild type and mutant three-day-old larvae over time under 395 nm UV stimulation.
+(**B**) Vertical position of batches of wild type and mutant three-day-old larvae over time under 395 nm UV stimulation.
 The starting position of each larval trajectory was set to 0.
-**(C)** Vertical position of batches of control and L-NAME-treated (0.1 and 1 mM) three-day-old larvae over time under 395 nm UV stimulation.
+(**C**) Vertical position of batches of control and L-NAME-treated (0.1 and 1 mM) three-day-old larvae over time under 395 nm UV stimulation.
 The starting position of each larval trajectory was set to 0.
-**(D)** Vertical displacement in 30 sec bins of wild type and mutant (NOSΔ11 and NOSΔ23) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
-**(E)** Vertical displacement in 30 sec bins of control and L-NAME-treated (0.1 and 1 mM) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.}\label{fig:unnamed-chunk-3}
+(**D**) Vertical displacement in 30 sec bins of wild type and mutant (*NOSΔ11/Δ11* and *NOSΔ23/Δ23*) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
+(**E**) Vertical displacement in 30 sec bins of control and L-NAME-treated (0.1 and 1 mM) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
+<br> Figure 3 -- source data 1-5.
+Source data for panels A-E.}\label{fig:unnamed-chunk-3}
 \end{figure}
@@
 To investigate how NO signalling alters the dynamics of the cPRC
-circuit, we carried out calcium imaging experiments. We ubiquitously
-expressed the calcium sensor GCaMP6s in larvae and imaged calcium
-signals during 405 nm stimulation of the cPRCs. As we have shown
-previously, a 20-sec local stimulation of cPRC cilia led to a transient
-increase in cPRC calcium levels, followed by a transient decrease
-(Verasztó et al., 2018). After \textasciitilde20-sec, calcium levels in
-cPRCs were raising again, reaching higher levels than at the start of
-the stimulus -- a response that may involve depolarisation (Figure 4A,
-B). This activation phase occurs after the 20 sec stimulation period and
-is likely due to a delayed neuroendocrine feedback (Verasztó et al.,
-2018). To determine whether NO mediates such a feedback, we repeated the
-experiment in \emph{NOS}-mutant larvae. While we detected the initial
-activation phase followed by inhibition, in homozygous
-\emph{NOS}-mutants for both CRISPR alleles this was not followed by
-delayed activation. Instead, calcium levels dropped to a low
-steady-state level (Figure 4A, B). We thus identified a requirement for
-NO signalling in the late-phase activation of cPRCs.
-
-\textbf{Two unconventional guanylyl cyclases are expressed in the cPRCs}
-
-We aimed next to identify the NO receptor in the cPRCs. NO generally
+circuit, we carried out Ca\textsuperscript{2+} imaging experiments. We
+ubiquitously expressed the Ca\textsuperscript{2+} sensor GCaMP6s in
+larvae and imaged Ca\textsuperscript{2+} signals during 405 nm
+stimulation of the cPRCs. As we have shown previously, a 20 sec local
+stimulation of cPRC cilia led to a transient increase in cPRC
+Ca\textsuperscript{2+} levels, followed by a transient decrease
+(Verasztó et al., 2018). After \textasciitilde20-sec,
+Ca\textsuperscript{2+} levels in cPRCs were raising again, reaching
+higher levels than at the start of the stimulus -- a response that may
+involve cPRC depolarisation (Figure 4A, B). This activation phase occurs
+after the 20 sec stimulation period and is likely due to a delayed
+neuroendocrine feedback (Verasztó et al., 2018). To determine whether NO
+mediates such a feedback, we repeated the experiment in
+\emph{NOS}-mutant larvae. While we detected the initial activation phase
+followed by inhibition, in homozygous \emph{NOS}-mutants for both CRISPR
+alleles this was not followed by delayed activation. Instead,
+Ca\textsuperscript{2+} levels dropped to a low steady-state level
+(Figure 4A, B). We thus identified a requirement for NO signalling in
+the late-phase activation of cPRCs.
+
+\textbf{Two unconventional guanylate cyclases are expressed in the
+cPRCs}
+
+Next we aimed to identify the NO receptor in the cPRCs. NO generally
 acts via soluble guanylate cyclases (sGC), belonging to the guanylate
@@
 these genes are not expressed in any of the cells of the cPRC circuit
-(Verasztó et al., 2017). Recently, Moroz and coworkers reported an
-atypical but widely conserved family of guanylyl cyclases with a NIT
-(nitrite/nitrate sensing) domain (PF08376) (NIT-GC) as potential
-mediators of NO signalling (Moroz et al., 2020). To identify NIT-GCs in
-\emph{Platynereis}, we searched transcriptome resources and retrieved 15
-potential NIT-GC homologs (Figure 4---figure supplement 1 and 2). To
-analyse the relationship of these sequences to metazoan NIT-GCs, we
-retrieved protein sequences with a CYC domain from the transcriptome and
-genome databases of 45 metazoan and 2 choanoflagellate species. We
-carried out cluster analysis and did phylogenetic reconstruction on a
-group of membrane-bound guanylyl cyclases with sGCs as an outgroup. In
-agreement with Moroz et al. (Moroz et al., 2020), we found a group of
-GCs with NIT domains with representatives in placozoans, cnidarians,
-some ecdysozoans, echinoderms, and lophotrochozoans. The 15
+(not shown and (Verasztó et al., 2017)). Recently, Moroz and coworkers
+reported an atypical but widely conserved family of guanylate cyclases
+with a NIT (nitrite/nitrate sensing) domain (PF08376) (NIT-GC) as
+potential mediators of NO signalling (Moroz et al., 2020). To identify
+NIT-GCs in \emph{Platynereis}, we searched transcriptome resources and
+retrieved 15 potential NIT-GC homologs (Figure 4---figure supplement 1
+and 2). To analyse the relationship of these sequences to metazoan
+NIT-GCs, we retrieved protein sequences with a CYC domain from the
+transcriptome and genome databases of 45 metazoan and 2 choanoflagellate
+species. We carried out cluster analysis and phylogenetic reconstruction
+on a group of membrane-bound guanylate cyclases with sGCs as an
+outgroup. In agreement with Moroz et al. (Moroz et al., 2020), we found
+a group of GCs with NIT domains with representatives in placozoans,
+cnidarians, some ecdysozoans, echinoderms, and lophotrochozoans. The 15
 \emph{Platynereis} sequences belonged to several deeply diverged clades
@@
 To characterise the expression of NIT-GCs, we used previously published
-spatially mapped single-cell transcriptome data (Achim et al., 2015;
+spatially-mapped single-cell transcriptome data (Achim et al., 2015;
 Williams et al., 2017). Among the 15 NIT-GCs, two showed high and
@@
 cells (Figure 4---figure supplement 2). In the single-cell data, we
-could identify the cPRCs by the specific expression of \emph{c-opsin1}
-and the \emph{pedal-peptide2 neuropeptide precursor} (\emph{MLD
-proneuropeptide}), previously described cPRC markers (Arendt et al.,
-2004; Williams et al., 2017) (Figure 4---figure supplement 3A). The
-INNOS cells were identified by NOS expression and spatial mapping in the
-brain (Achim et al., 2015). We decided to focus on two NIT-GCs expressed
-in the cPRCs and with a full-length sequence, \emph{NIT-GC1} and
-\emph{NIT-GC2}. To confirm the single-cell data, we first carried out in
-situ hybridisation chain reaction (HCR) with probes for \emph{NIT-GC1}
-and \emph{NIT-GC2} mRNA. Both genes were specifically expression in the
-four cPRCs, as confirmed by co-labeling with an acetylated α-tubulin
-antibody and with an HCR probe against \emph{pedal peptide 2/MLD
-proneuropeptide} (Figure 4C, D and Figure 4---figure supplement 3A-C).
-To analyse the subcellular localisation of NIT-GC1 and NIT-GC2 at the
-protein level, we raised and affinity-purified polyclonal antibodies
-against a specific peptide sequence from both proteins. In
-immunostainings, we found that NIT-GC1 was localise to the region
-corresponding to the axonal projections of the cPRCs in the anterior
-neurosecretory plexus (Figure 4E). Co-immunostaining with the rabbit
-NIT-GC1 antibody and a custom rat antibody raised against
-\emph{Platynereis} NOS revealed the localisation of both proteins in
-dots in the neurosecretory neuropil (Figure 4---figure supplement 3D).
-In contrast, NIT-GC2 specifically labelled the ramified sensory cilia of
-the cPRCs (Figure 4F). These different subcellular localisations suggest
-that the two NIT-GCs are involved in different intracellular signalling
-processes in the ciliary and axonal regions of the cPRCs.
+could identify the cPRCs by the specific expression of the markers
+\emph{c-opsin1} and the \emph{pedal-peptide2 neuropeptide precursor}
+(\emph{MLD proneuropeptide}) (Arendt et al., 2004; Williams et al.,
+2017) (Figure 4---figure supplement 3A). The INNOS cells were identified
+by NOS expression and spatial mapping in the brain (Achim et al., 2015).
+We decided to focus on two NIT-GCs expressed in the cPRCs and with a
+full-length sequence, \emph{NIT-GC1} and \emph{NIT-GC2}. To confirm the
+single-cell data, we first carried out in situ hybridisation chain
+reaction (HCR) with probes for \emph{NIT-GC1} and \emph{NIT-GC2} mRNA.
+Both genes were specifically expressed in the four cPRCs, as confirmed
+by co-labeling with an acetylated α-tubulin antibody and with an HCR
+probe against \emph{pedal peptide 2/MLD proneuropeptide} (Figure 4C, D
+and Figure 4---figure supplement 3A-C). To analyse the subcellular
+localisation of NIT-GC1 and NIT-GC2 at the protein level, we raised and
+affinity-purified polyclonal antibodies against a specific peptide
+sequence from both proteins. In immunostainings, we found that NIT-GC1
+was localise to the region corresponding to the axonal projections of
+the cPRCs in the anterior neurosecretory plexus (Figure 4E).
+Co-immunostaining with the rabbit NIT-GC1 antibody and a custom rat
+antibody raised against \emph{Platynereis} NOS revealed the localisation
+of both proteins in dots in the neurosecretory neuropil (Figure
+4---figure supplement 3D). In contrast, NIT-GC2 specifically labelled
+the ramified sensory cilia of the cPRCs (Figure 4F). These different
+subcellular localisations suggest that the two NIT-GCs are involved in
+different intracellular signalling processes in the ciliary and axonal
+regions of the cPRCs.
 
@@
 
-To further characterise these atypical guanylyl cyclases, we focused on
+To further characterise these atypical guanylate cyclases, we focused on
 NIT-GC1 and carried out in vitro experiments. In bacteria, NIT domains
@@
 To test the in vivo function of NIT-GC1 and NIT-GC2 in cPRC responses,
-we combined calcium imaging with morpholino-mediated knockdowns. We used
-two translation-blocking morpholinos for each \emph{NIT-GC} gene and
-tested knockdown efficiency by immunostaining injected animals with the
-NIT-GC1 and NIT-GC2 antibodies (Figure 4---figure supplement 3E,F). For
-both genes, the morpholinos led to a strong reduction in the respective
-antibody signal, confirming efficient knockdown and antibody
-specificity.
+we combined Ca\textsuperscript{2+} imaging with morpholino-mediated
+knockdowns. We used two translation-blocking morpholinos for each
+\emph{NIT-GC} gene and tested knockdown efficiency by immunostaining
+injected animals with the NIT-GC1 and NIT-GC2 antibodies (Figure
+4---figure supplement 3E,F). For both genes, the morpholinos led to a
+strong reduction in the respective antibody signal, confirming efficient
+knockdown and antibody specificity.
 
@@
 al., 2014). NIT-GC2 morphant larvae, in contrast, showed a step-up
-increase in calcium following light stimulation (Figure 4N). The calcium
-signal decayed during stimulation and was off after light off. These
-data support an essential role for ciliary-localised NIT-GC2 in
-suppressing cPRC calcium following its transient rise at stimulus onset.
-Overall, these knockdown experiments revealed different signalling
-mechanisms for the two NIT-GCs that may be due to their different
-subcellular localisations.
+increase in Ca\textsuperscript{2+} following light stimulation (Figure
+4N). The Ca\textsuperscript{2+} signal decayed during stimulation and
+was off after light off. These data support an essential role for
+ciliary-localised NIT-GC2 in suppressing cPRC Ca\textsuperscript{2+}
+following its transient rise at stimulus onset. Overall, these knockdown
+experiments revealed different signalling mechanisms for the two NIT-GCs
+that may be due to their different subcellular localisations.
 
@@
 \includegraphics[width=38.89in]{figures/Fig4} \caption{**Figure 4.
-NOS and two NIT-GCs shape calcium signals during cPRC UV/violet response.** **(A, B)** GCaMP6s signals in cPRCs in wild type and *NOS* mutant (A, NOSΔ11/Δ11, B, NOSΔ23/Δ23) larvae during 405 nm light stimulation.
-**(C, D)** In situ HCR for (C) *NIT-GC1* and (D) *NIT-GC2* (magenta) in three-day-old *Platynereis* larvae.
+NOS and two NIT-GCs shape Ca^2+^ signals during cPRC UV/violet response.** (**A**, **B**) GCaMP6s signals in cPRCs in wild type and *NOS* mutant (A, *NOSΔ11/Δ11*, B, *NOSΔ23/Δ23*) larvae during 405 nm light stimulation.
+(**C**, **D**) In situ HCR for (C) *NIT-GC1* and (D) *NIT-GC2* (magenta) in three-day-old *Platynereis* larvae.
 Larvae were co-stained with an antibody against acetylated α-tubulin to label cPRC cilia and the neuropil (green).
-**(E, F)** Immunostaining for (E) NIT-GC1 and (F) NIT-GC2 (magenta), co-stained for acetylated α-tubulin (green).
-**(G)** The domain structure of *Platynereis* NIT-GC1 and the truncated NIT-GC1ΔNIT protein lacking the NIT domain.
+(**E**, **F**) Immunostaining for (E) NIT-GC1 and (F) NIT-GC2 (magenta), co-stained for acetylated α-tubulin (green).
+(**G**) The domain structure of *Platynereis* NIT-GC1 and the truncated NIT-GC1ΔNIT protein lacking the NIT domain.
 A predicted transmembrane region (TM) is shown in grey.
-**(H)** Schematic of the cell-based assay to detect cGMP production following the addition of an NO donor SNAP or DMSO as control.
-**(I-L)** Green cGull fluorescence over time for the four conditions tested.
+(**H**) Schematic of the cell-based assay to detect cGMP production following the addition of an NO donor SNAP or DMSO as control.
+(**I**-**L**) Green cGull fluorescence over time for the four conditions tested.
 Individual responses and their mean with 0.95 confidence interval are shown (n > 6 cells).
@@
 The indicated chemicals were added at 2 min after the start of imaging (grey bars).
-**(M, N)** GCaMP6s signals in cPRCs in (M) NIT-GC1 and (N) NIT-GC2 morphant larvae during 405 nm light stimulation.
-Individual responses and their mean with 0.95 confidence interval are shown.}\label{fig:unnamed-chunk-4}
+(**M**, **N**) GCaMP6s signals in cPRCs in (M) NIT-GC1 and (N) NIT-GC2 morphant larvae during 405 nm light stimulation.
+Individual responses and their mean with 0.95 confidence interval are shown.
+<br> Figure 4 -- source data 1-8.
+Source data for panels A, B and I-N.}\label{fig:unnamed-chunk-4}
 \end{figure}
@@
 To investigate how NO and NIT-GC signalling influence the dynamics of
-the cPRC circuit, we imaged calcium signals from postsynaptic neurons in
-wild type, mutant and morphant larvae. We were able to image the
-activity of all neurons in the cPRC circuit (INNOS, INRGWa, Ser-h1 and
-MC). The MC cell was identified based on its position and intrinsic
-activity (Verasztó et al., 2017). To unambiguously identify all other
-cells from which we recorded calcium signals, we developed an on-slide
-immunostaining method (Figure 5---figure supplement 1A). We used the
-cell-specific antibody markers against RYamide (INNOS) (Figure
-5---figure supplement 1B-D), RGWamide (INRGWa) and serotonin (Ser-h1)
-(Conzelmann et al., 2011) to immunostain agar-embedded larvae following
-calcium imaging. Based on the position of the nuclei, we could correlate
-live and fixed samples at a single-cell precision (Figure 5A,B). Due to
-the stereotypy of the larvae, we could also identify neurons based on
-their position and calcium activity in activity-correlation maps (Figure
+the cPRC circuit, we imaged Ca\textsuperscript{2+} signals from
+postsynaptic neurons in wild type, mutant and morphant larvae. We were
+able to image the activity of all neurons in the cPRC circuit (INNOS,
+INRGWa, Ser-h1 and MC). The MC cell was identified based on its position
+and intrinsic activity (Verasztó et al., 2017). To unambiguously
+identify all other cells from which we recorded Ca\textsuperscript{2+}
+signals, we developed an on-slide immunostaining method (Figure
+5---figure supplement 1A). We used the cell-specific antibody markers
+against RYamide (INNOS) (Figure 5---figure supplement 1B-D), RGWamide
+(INRGWa) and serotonin (Ser-h1) (Conzelmann et al., 2011) to immunostain
+agar-embedded larvae following Ca\textsuperscript{2+} imaging. Based on
+the position of the nuclei, we could correlate live and fixed samples at
+a single-cell precision (Figure 5A, B). Due to the stereotypy of the
+larvae, we could also identify neurons based on their position and
+Ca\textsuperscript{2+} activity in activity-correlation maps (Figure
 5C).
@@
 during 405 nm stimulation of the cPRCs. In both wild type and
-\emph{NOS}-mutant larvae, INNOS cells showed an increase in calcium
-during stimulation (Figure 5D). In contrast, the INNOS response was flat
-or slightly negative in NIT-GC2 morphant larvae (Figure 5E) revealing an
-essential role for NIT-GC2-mediated cPRC suppression in INNOS
-activation. INRGWa cells were initially inhibited during cPRC
-stimulation, followed by a delayed activation paralleling the second
-activation phase of cPRCs. This late INRGWa response was lacking in
-\emph{NOS}-mutants (Figure 5F). In NIT-GC2 morphants, INRGWa cells
-showed a transient increase in calcium that decayed after light off and
-a delayed activation was not present (Figure 5G).
-
-Next, we imaged calcium signals from Ser-h1 and MC neurons in wild type
-and \emph{NOS} mutant larvae. Ser-h1 cells showed an activation profile
-that correlated with cPRC activity, including a reduction in calcium
-during stimulation followed by rebound, a response that was defective in
-\emph{NOS} mutants (Figure 5I). MC cells showed sustained activation,
-including a late-phase that was lacking in \emph{NOS} mutants (Figure
-5I). These data suggest that during 405 nm stimulation the Ser-h1 cells
-are inhibited and MC cells are activated, and this regulation is
-NO-dependent. This pattern is expected to inhibit ciliary activity in
-the prototroch but not in the other ciliary bands, triggering
+\emph{NOS}-mutant larvae, INNOS cells showed an increase in
+Ca\textsuperscript{2+} during stimulation (Figure 5D). In contrast, the
+INNOS response was flat or slightly negative in NIT-GC2 morphant larvae
+(Figure 5E) revealing an essential role for NIT-GC2-mediated cPRC
+suppression in INNOS activation. The INRGWa cells were initially
+inhibited during cPRC stimulation, followed by a delayed activation
+paralleling the second activation phase of the cPRCs. This late INRGWa
+response was lacking in \emph{NOS}-mutants (Figure 5F). In NIT-GC2
+morphants, INRGWa cells showed a transient increase in
+Ca\textsuperscript{2+} that decayed after light off and a delayed
+activation was not present (Figure 5G).
+
+Next, we imaged Ca\textsuperscript{2+} signals from Ser-h1 and MC
+neurons in wild type and \emph{NOS} mutant larvae. Ser-h1 cells showed
+an activation profile that correlated with cPRC activity, including a
+reduction in Ca\textsuperscript{2+} during stimulation followed by a
+rebound, a response that was defective in \emph{NOS} mutants (Figure
+5I). The MC cell showed sustained activation, including a late-phase
+that was lacking in \emph{NOS} mutants (Figure 5I). These data suggest
+that during 405 nm stimulation the Ser-h1 cells are inhibited and the MC
+cell is activated, and this regulation is NO-dependent. This pattern is
+expected to inhibit ciliary activity in the prototroch but not in the
+other ciliary bands (with no Ser-h1 or MC synapses), triggering
 NO-dependent downward swimming.
@@
 \includegraphics[width=52.08in]{figures/Fig5} \caption{**Figure 5.
-NOS- and NIT-GC2-dependent dynamics of the cPRC circuit.** **(A, B)** GCaMP6s imaging from cPRCs and INNOS cells (left panels) followed by on-slide immunostaining for (A) RYamide to label INNOS and (B) RGWamide+serotonin to label INRGWa and Ser-h1 (red).
+NOS- and NIT-GC2-dependent dynamics of the cPRC circuit.** (**A, B**) GCaMP6s imaging from cPRCs and INNOS cells (left panels) followed by on-slide immunostaining for (A) RYamide to label INNOS and (B) RGWamide+serotonin to label INRGWa and Ser-h1 (red).
 Nuclei are stained with DAPI (cyan).
@@
 Numbers mark the same cells in the GCaMP and immunostaining images matched by position.
-**(C)** Correlation map of neuronal activity of the cPRCs, INNOS, INRGWa, Ser-h1 and MC neurons.
-**(D)** GCaMP6s fluorescence in INNOS cells in wild type (WT) and NOSΔ11/Δ11 mutant larvae during 405 nm stimulation of the cPRC cilia.
-**(E)** GCaMP6s fluorescence in INNOS cells in NIT-GC2 morphant larvae during 405 nm stimulation.
-**(F)** GCaMP6s fluorescence in INRGWa cells in wild type and NOSΔ23/Δ23 mutant larvae during 405 nm stimulation.
-**(G)** GCaMP6s fluorescence in INRGWa cells in NIT-GC2 morphant larvae during 405 nm stimulation.
-**(H, I)** GCaMP6s fluorescence in (H) Ser-h1 cells and (I) the MC cell in wild type and NOSΔ11/Δ11 mutant larvae during 405 nm stimulation.}\label{fig:unnamed-chunk-5}
+(**C**) Correlation map of neuronal activity of the cPRCs, INNOS, INRGWa, Ser-h1 and MC neurons.
+(**D**) GCaMP6s fluorescence in INNOS cells in wild type (WT) and *NOSΔ11/Δ11* mutant larvae during 405 nm stimulation of the cPRC cilia.
+(**E**) GCaMP6s fluorescence in INNOS cells in NIT-GC2 morphant larvae during 405 nm stimulation.
+(**F**) GCaMP6s fluorescence in INRGWa cells in wild type and *NOSΔ23/Δ23* mutant larvae during 405 nm stimulation.
+(**G**) GCaMP6s fluorescence in INRGWa cells in NIT-GC2 morphant larvae during 405 nm stimulation.
+(**H, I**) GCaMP6s fluorescence in (H) Ser-h1 cells and (I) the MC cell in wild type and *NOSΔ11/Δ11* mutant larvae during 405 nm stimulation.
+<br> Figure 5 -- source data 1-6.
+Source data for panels D-I.}\label{fig:unnamed-chunk-5}
 \end{figure}
@@
 signalling, we developed a mixed cellular-circuit-level mathematical
-model. We used our Ca2+ imagining data of cPRC, INNOS and INRGWa cells
-collected in wild type, \emph{NOS} knockout, and \emph{NIT-GC2} morphant
-larvae to formulate assumptions that are the basis of the proposed
-model.
-
-We model a direct c-opsin1-dependent response to UV leading to the
-initial rise in Ca2+ levels. The mechanism of this signal is not known
-but may include Gα and Gβ signalling, CNGAα channel activation or other
-pathways. We further assume that in cPRC cells, Ca2+ levels increase
-with cGMP. We model a decrease in cGMP and Ca2+ levels due to a
-NIT-GC2-dependent pathway (Figre 6A). The effect of NO in cPRC cells is
-captured by another term that describes a NIT-GC1 and NO-dependent
-increase in cGMP. Based on the synaptic connectome, we infer a
-feedforward coupling between cPRC cells and INNOS and INRGWa cells. The
-proposed synaptic coupling is UV-dependent, decays linearly and is
-suppressed in NIT-GC2 morphants (Figure 6A).
-
-In the INNOS cells, we assume that excitatory synaptic input from cPRC
-(or rebound from tonic inhibition) leads to a rise in Ca2+ leading to NO
-production via NOS. This input is NIT-GC2-dependent. In the INRGWa
-cells, we model a decrease in Ca2+ based on an inhibitory UV-dependent
-synaptic signal from the cPRCs. We further assume the existence of
-direct excitatory coupling between cPRC Ca2+ levels and INRGWa Ca2+ to
-account for the late effects of cPRC on INRGWa (Figure 6A). The
-UV-dependent and Ca2+-dependent signals may be mediated by different
+model. We used our Ca\textsuperscript{2+} imagining data of cPRC, INNOS
+and INRGWa cells collected in wild type, \emph{NOS} knockout and
+\emph{NIT-GC2} morphant larvae to formulate assumptions that are the
+basis of the proposed model.
+
+We modeled a direct c-opsin1-dependent response to UV leading to the
+initial rise in Ca\textsuperscript{2+} levels. The mechanism of this
+signal is not known but may include Gα and Gβ signalling, CNGAα channel
+activation or other pathways. We further assumed that in cPRC cells,
+Ca\textsuperscript{2+} levels increase with cGMP. We modeled a decrease
+in cGMP and Ca\textsuperscript{2+} levels due to a NIT-GC2-dependent
+pathway (Figure 6A). The effect of NO in cPRC cells is captured by
+another term that describes a NIT-GC1 and NO-dependent increase in cGMP.
+Based on the synaptic connectome, we infered a feedforward coupling
+between cPRC cells and INNOS and INRGWa cells. The proposed synaptic
+coupling is UV-dependent, decays linearly and is suppressed in NIT-GC2
+morphants (Figure 6A).
+
+In the INNOS cells, we assumed that excitatory synaptic input from cPRC
+(or rebound from tonic inhibition) leads to a rise in
+Ca\textsuperscript{2+} leading to NO production via NOS. This input is
+NIT-GC2-dependent. In the INRGWa cells, we modeled a decrease in
+Ca\textsuperscript{2+} based on an inhibitory UV-dependent synaptic
+signal from the cPRCs. We further assumed the existence of direct
+excitatory coupling between cPRC Ca\textsuperscript{2+} levels and
+INRGWa Ca\textsuperscript{2+} to account for the late effects of cPRC on
+INRGWa (Figure 6A). The UV-dependent and
+Ca\textsuperscript{2+}-dependent signals may be mediated by different
 transmitters released by cPRCs during distinct phases of their
-activation cycle. In addition, we assume a direct inhibitory coupling
-between INNOS Ca2+ and INRGWa Ca2+. Finally, in all variables, we
-assumed a linear decay and constant production to set a steady state
-(Figure 6A). Since the aim of the model is to capture the normalised
-fluorescence data, the model is nondimensionalised. UV stimulation is
-modelled as a square pulse. To find model parameters producing an output
-fitting the experimental data we employed a global optimisation method
-known as a genetic algorithm (GA).
-
-Through this optimisation procedure, we found parameters that gave a
-very good fit to our experimental data under all conditions. We could
-fit to mean Ca2+ values but also individual calcium transients (Figure
-6B, Figure 6--figure supplement 1). Our model only includes a minimal
-set of parameters and assumptions of interactions that are required to
-describe the dynamics of cPRC, INNOS and INRGWa in wild type and
-loss-of-function conditions (Figure 6A, B). The model highlights that
-cPRC phototransduction employs distinct pathways that operate on
-different time scales and differentially influence cPRC Ca2+ levels. The
-coupling between cPRCs and interneurons also requires different
-signalling mechanisms, either through distinct neurotransmitters or
-receptors expressed in the different cells.
+activation cycle. In addition, we assumed a direct inhibitory coupling
+between INNOS Ca\textsuperscript{2+} and INRGWa Ca\textsuperscript{2+}.
+Finally, in all variables, we assumed a linear decay and constant
+production to set a steady state (Figure 6A). Since the aim of the model
+is to capture the normalised fluorescence data, the model is
+nondimensionalised. UV stimulation is modelled as a square pulse. To
+find model parameters producing an output fitting the experimental data
+we employed a global optimisation method known as a genetic algorithm
+(GA).
+
+Through this optimisation procedure, we found parameters that gave a fit
+to our experimental data under all conditions (Figure 6B; Figure
+6--figure supplement 1-5). We could fit to mean Ca\textsuperscript{2+}
+profiles but also individual Ca\textsuperscript{2+} transients. Our
+model only includes a minimal set of parameters and assumptions of
+interactions that are required to describe the dynamics of cPRC, INNOS
+and INRGWa in wild type and loss-of-function conditions (Figure 6A, B).
+The model highlights that cPRC phototransduction employs distinct
+pathways that operate on different time scales and differentially
+influence cPRC Ca\textsuperscript{2+} levels. The coupling between cPRCs
+and interneurons also requires different signalling mechanisms, either
+through distinct neurotransmitters or receptors expressed in the
+different cells.
 
@@
 
+\begin{figure}
+\includegraphics[width=25in]{figures/Fig6} \caption{**Figure 6.
+Mathematical modelling and signalling mechanisms of the cPRC circuit.** (**A**) Diagram of the mathematical model with the componenets, interactions, parameters and equations used to model Ca^2+^ dynamics.
+(**B**) Average Ca^2+^ traces (upper) and traces from model simulations (lower) fitted to the data.
+(**C**) Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq.
+The size of the dots is expressed in proportion to the percentage of cells expressing that gene relative to all cells.
+The colours represent the normal logarithm of the number of transcripts in the cells expressing the gene.
+(**D**) Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback.
+<br> Figure 6 -- source data 1.
+TPM values for each gene and the percentage of expressed genes.}\label{fig:unnamed-chunk-6}
+\end{figure}
+
 To generate new insights from the model, we asked how the duration and
 intensity of UV stimulation influences signalling output, measured as
-the magnitude, duration and time to maximum of the NO-dependent Ca2+
-signal in the cPRCs. The model allowed us to test a large number of
-conditions that would not have been feasible in single-larva
-experiments. We found that the onset, magnitude and duration of the
-NO-dependent signal scales with stimulus intensity and duration (Figure
-6--figure supplement 2). This suggests an integratory function of NO
-signalling that tunes circuit output to the strength of UV exposure.
+the magnitude (quantified as area under peak, AUP) of the NO-dependent
+Ca\textsuperscript{2+} signal in the cPRCs. The model allowed us to
+systematically vary stimulus conditions that would not have been
+feasible in single-larva experiments. We found that delayed activation
+occurs only for UV stimulation of sufficient amplitude and duration
+(Figure 7). This suggests an integratory function of NO signalling to
+tune circuit output to the strength of UV exposure.
 
 \begin{figure}
-\includegraphics[width=25in]{figures/Fig6} \caption{**Figure 6.
-Mathematical modelling of cPRC circuit dynamics during UV stimulation and potential signalling mechanisms.**  **(A)** Diagram for **(B)** **(C)** Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq.
-The size of the dots is expressed in proportion to the percentage of cells expressing that gene relative to all cells.
-The colours represent the normal logarithm of the number of transcripts in the cells expressing the gene.
-**(D)** Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback.}\label{fig:unnamed-chunk-6}
+\centering
+\includesvg{figures/Fig7.svg}
+\caption{\textbf{Figure 7. Magnitude of cPRC activation as a function of
+stimulus duration and intensity.} Median values of the \(AUP\) are
+presented as a heatmap with dark blue indicating flat \(C_P(t)\) traces
+with median \(AUP\) close to 0 and red indicating that the \(C_P(t)\)
+traces have a clearly defined delayed activation peak.}
 \end{figure}
@@
 NO is a free radical with a millisecond-to-second half life and thus a
-limited signalling range. In neuronal signalling, NOS is often localized
-to neurites (Kuntz et al., 2017) at close proximity to sGC at synapses
-(Burette et al., 2002; Garthwaite, 2015). In the cPRC circuit, NOS is
-localised to the dendritic compartment of INNOS cells where also cPRC
-postsynaptic sites occur. NIT-GC1, the target of NO signalling is
-localised to cPRC projections. NO-mediated retrograde signalling thus
-likely occurs in the neurosecretory plexus where NOS- and
-NIT-GC1-containing projections are in close proximity and where we
-detected NO production following UV stimulation. In contrast, INNOS to
-INRGWa synapses occur outside the neurosecretory plexus in the ventral
-brain neuropil.
+limited signalling range. In neuronal signalling, NOS is often localised
+to neurites (Kuntz et al., 2017) close to sGC at synapses (Burette et
+al., 2002; Garthwaite, 2015). In the cPRC circuit, NOS is localised to
+the dendritic compartment of INNOS cells where also cPRC postsynaptic
+sites occur. NIT-GC1, the target of NO signalling is localised to cPRC
+projections. NO-mediated retrograde signalling thus likely occurs in the
+neurosecretory plexus where NOS- and NIT-GC1-containing projections are
+in close proximity and where we detected NO production after UV
+stimulation. In contrast, INNOS to INRGWa synapses occur outside the
+neurosecretory plexus in the ventral brain neuropil.
 
@@
 peptidergic hub neuron Dp7 links UV-sensory neurons (v'td2) and motor
-circuits. Dp7 expresses an an insulin-like peptide Ilp7 that is required
+circuits. Dp7 expresses an insulin-like peptide Ilp7 that is required
 for acute and sustained UV avoidance. Ilp7 signalling occurs at a
@@
 Bacterial NIT domains regulate cellular functions in response to changes
-in extracellular and intracellular nitrate and nitrite concentrations
+in extracellular and intracellular nitrate and/or nitrite concentrations
 (Camargo et al., 2007). NO is readily converted to nitrate and nitrite
@@
 detailed model of cPRC phototransduction and neurotransmission. The
-cPRCs have high basal Ca2+ and respond to UV/violet light dependent on
-c-opsin1 (Verasztó et al., 2018). c-opsin1 forms a bistable photopigment
-and signals through Gi/oα and Gβγ (Tsukamoto et al., 2017; Tsukamoto and
-Kubo, 2023; Veedin Rajan et al., 2021). In heterologous systems, the Gβγ
-subunits released following c-opsin1 activation open GIRK channels
-inducing K+ efflux (hyperpolarisation) (Tsukamoto et al., 2017;
-Tsukamoto and Kubo, 2023) and close voltage-gated Ca2+ channels, thereby
-reducing intracellular calcium levels (Tsukamoto and Kubo, 2023). A GIRK
-channel is also expressed in the cPRCs (Figure 6A). The pathway for the
-first rapid cPRC activation phase following UV/violet stimulus is not
-known but may involve the activation of a CNGα channel expressed in the
-cPRCs (Tosches et al., 2014). For the second inhibitory phase of
+cPRCs have high basal Ca\textsuperscript{2+} and respond to UV/violet
+light dependent on c-opsin1 (Verasztó et al., 2018). c-opsin1 forms a
+bistable photopigment and signals through Gi/oα and Gβγ (Tsukamoto et
+al., 2017; Tsukamoto and Kubo, 2023; Veedin Rajan et al., 2021). In
+heterologous systems, the Gβγ subunits released following c-opsin1
+activation open GIRK channels inducing K+ efflux (hyperpolarisation)
+(Tsukamoto et al., 2017; Tsukamoto and Kubo, 2023) and close
+voltage-gated Ca\textsuperscript{2+} channels, thereby reducing
+intracellular Ca\textsuperscript{2+} levels (Tsukamoto and Kubo, 2023).
+A GIRK channel is also expressed in the cPRCs (Figure 6A). The pathway
+for the first rapid cPRC activation phase following UV/violet stimulus
+is not known but may involve the activation of a CNGα channel expressed
+in the cPRCs (Tosches et al., 2014). For the second inhibitory phase of
 phototransduction, we identified a key requirement for ciliary-localised
@@
 INNOS leads to NOS activation and NO release, potentially through a
-canonical Ca2+-calmodulin pathway. Our model predicts feedforward
-inhibiton from INNOS to INRGWa, possibly mediated by glycin
+canonical Ca\textsuperscript{2+}-calmodulin pathway. Our model predicts
+feedforward inhibition from INNOS to INRGWa, possibly mediated by glycin
 transmission. NO released by INNOS neurites activates NIT-GC1 in cPRC
 projections that could lead to cGMP production and the opening of CNGAα.
-This late-phase cPRC activation is not directly dependent on the
+This late-phase cPRC activation does not directly depend on the
 UV/violet signal and can happen post-stimulus. The late-phase cPRC
@@
 In addition, the cPRC circuit expresses several neuropeptides and their
-receptors, suggesting further neuromodulatory signals. For example,
-INRGWa express the proneuropeptide RGWamide and cPRCs and INNOS express
-its receptor, suggesting retrograde peptidergic signalling in the
-circuit.
+receptors, suggesting further neuromodulatory signals. For example, the
+INRGWa cells express the proneuropeptide RGWamide and cPRCs and INNOS
+express its receptor, suggesting retrograde peptidergic signalling in
+the circuit.
 
@@
 
-Pressure-induced Ca2+ transients in cPRCs lack an inhibitory phase and
-late activation and resemble UV responses in NIT-GC2 morphants
-(Bezares-Calderón et al., 2023). This indicates that the differentiation
-of sensory cues by the multisensory cPRCs occurs already at the level of
-sensory signal transduction. The different signalling pathways then
-likely result in the differential release of transmitters and modulators
-that are decoded by the postsynaptic interneuron circuit. The complex
-transmitter phenotype of cPRCs could underlie such differential signal
-processing.
+Pressure-induced Ca\textsuperscript{2+} transients in cPRCs lack an
+inhibitory phase and late activation and resemble UV responses in
+NIT-GC2 morphants (Bezares-Calderón et al., 2023). This indicates that
+the differentiation of sensory cues by the multisensory cPRCs occurs
+already at the level of sensory signal transduction. The different
+signalling pathways then likely result in the differential release of
+transmitters and modulators that are decoded by the postsynaptic
+interneuron circuit. The complex transmitter phenotype of cPRCs could
+underlie such differential signal processing.
 
@@
 same axons. This autocrine signal leads to a CNG-dependent temporary
-increase in calcium levels, carrying the working memory trace (Kuntz et
-al., 2017).
+increase in Ca\textsuperscript{2+} levels, carrying the working memory
+trace (Kuntz et al., 2017).
 
 In the \emph{Platynereis} circuit, our mathematical model indicates that
-the onset, magnitude and duration of the NO-dependent signal depends on
-the intensity and duration of the UV/violet stimulus. This suggests that
-the NO-dependent memory trace also encodes the magnitude and duration of
-the stimulus.
+the magnitude of the NO-dependent signal depends on the intensity and
+duration of the UV/violet stimulus. This suggests that the NO-dependent
+memory trace also encodes the intensity and duration of the stimulus.
 
@@
 
-We hypothesise that the differential beating of prototroch versus trunk
+We hypothesize that the differential beating of prototroch versus trunk
 cilia causes a head-up or head-down orientation due to physics alone.
@@
 
-\hypertarget{uv-avoidance-circuits-of-extraocular-photoreceptors}{%
-\subsection{UV-avoidance circuits of extraocular
-photoreceptors}\label{uv-avoidance-circuits-of-extraocular-photoreceptors}}
+\textbf{UV-avoidance circuits of extraocular photoreceptors}
 
@@
 phototaxis with a maximum sensitivity to cyan light (\textasciitilde500
-nm) (Randel et al., 2014; Verasztó et al., 2018). Planarians also have
-pigmented visual eyes mediating phototaxis to cyan light and peripheral
-extraocular photoreceptors mediating UV avoidance behaviour (Shettigar
-et al., 2021). \emph{Drosophila} larvae have cerebral eyes called
-Bolwig's organs involved in phototaxis (Kane et al., 2013) and several
-types of extraocular UV-sensory cells that tile the body wall and
-mediate UV avoidance (Imambocus et al., 2022).
+nm) (Gühmann et al., 2015; Randel et al., 2014; Verasztó et al., 2018).
+Planarians also have pigmented visual eyes mediating phototaxis to cyan
+light and peripheral extraocular photoreceptors mediating UV avoidance
+behaviour (Shettigar et al., 2021). \emph{Drosophila} larvae have
+cerebral eyes called Bolwig's organs involved in phototaxis (Kane et
+al., 2013) and several types of extraocular UV-sensory cells that tile
+the body wall and mediate UV avoidance (Imambocus et al., 2022).
 
@@
 
-We have revealed how NO shapes dynamics in a fully-mapped sensory-motor
-UV avoidance circuit. We identified an unconventional GC as the NO
-receptor in \emph{Platynereis} and measured circuit activity in
-different genetic backgrounds. Finally, we could link circuit activity
-to light-avoidance behaviour. The richly modulated multi-transmitter
-cPRC circuit will serve as a fertile ground for future studies on how
-neuromodulators shape activity and behaviour.
+Overall, we have revealed how NO shapes the dynamics of a fully mapped
+sensory-motor UV avoidance circuit. We identified an unconventional GC
+as the NO receptor and measured circuit activity in different genetic
+backgrounds. Finally, we could link circuit activity to light-avoidance
+behaviour. The richly modulated multi-transmitter cPRC circuit will
+serve as a fertile ground for future studies on how neuromodulators
+shape activity and behaviour.
 
@@
 
-NA
-
 ~Knockout generated by CRISPR/Cas-9-induced gene editing
@@
 
-NA
+\url{https://www.atcc.org/products/crl-1651}
 
@@
 
-Angio-proteomie (CAT no. cAP-0203)??
+ATCC (CRL-1651™)
 
@@
 
-NA
-
 Gene (Platynereis dumerilii)
@@
 
-NA
-
 NOS: Nitric\_Oxide\_Synthase
@@
 
-NA
-
-NA
-
 5'-TAGGGCAATACTGGCTCCACTC-3'
@@
 
-NA
-
-NA
-
 5'-AAACGAGTGGAGCCAGTATTGC-3'
@@
 
-NA
-
 plasmid
@@
 
-NA
-
 Antibody
@@
 
-NA
-
-NIT-GC1 polychronal antibodies
+NIT-GC1 polyclonal antibody
 
@@
 
-NIT-GC2 polychronal antibodies
+NIT-GC2 polyclonal antibody
 
@@
 
-NOS polychronal antibodies
+NOS polyclonal antibody
 
@@
 
-RYamide neuropeptide antibodies
+RYamide neuropeptide antibody
 
@@
 
-RGWamide neuropeptide antibodies
+RGWamide neuropeptide antibody
 
@@
 
-NA
-
 Recombinant DNA reagent
@@
 
-NA
-
 GGTTGAATAATGACAAGCAAGGAGA
@@
 
-NA
-
 GTGCTATCATTTCCAGGTAAATACCC
@@
 
-NA
-
 AATATCTAGCGAAGGAGAACACCTCTCTTC
@@
 
-NA
-
 ATGGCCAGTAATAAACCATCAGTGGTTC
@@
 
-NA
-
 CGTTTAAACTTAAGCTTGGTACCGAG
@@
 
-NA
-
 CCAGCTTGGGTCTCCCTATAGT
@@
 
-NA
-
 TATAGGGAGACCCAAGCTGGGCCACCATGACCCAGATG
@@
 
-NA
-
 GCATGTTAGAAGACTTCCTCTGCCCTCATAATCAAACCCCTCTCT
@@
 
-NA
-
 AGGGCCGGGATTCTCCTCCACGTCACCGCATGTTAGAAGACTTCC
@@
 
-NA
-
 TGCTCACCATAGGGCCGGGATTCTCCTC
@@
 
-NA
-
 TCCCGGCCCTATGGTGAGCAAGGGCGAG
@@
 
-NA
-
 ACCAAGCTTAAGTTTAAACGTTACTTGTACAGCTCGTCCATG
@@
 
-NA
-
 AGCCATCTACGAGTGGTA
@@
 
-NA
-
 TGCCCTTCAGCTCGATG
@@
 
-NA
-
 TGACTGACGAACCCTCC
@@
 
-NA
-
 AGATATCTTGAAACGGACGA
@@
 
-NA
-
 TGACGTGGAGGAGAATCCCGGCCCTATGGTGAGCAAGGGCGAGGAGCTGT
@@
 
-NA
-
 GCAGAGGAAGTCTTCTAACATGCGGTGACGTGGAGGAGAATCCCGGCCCT
@@
 
-NA
-
 TCCTTGCTTGTCATGGTGGCCCAGCTTGGGTCTCCCTATAGTGAGTCGTA
@@
 
-NA
-
 GAGACCCAAGCTGGGCCACCATGACAAGCAAGGAGATGCCTGTACTCATG
@@
 
-NA
-
 TGTTAGAAGACTTCCTCTGCCCTCTATGACTTTTTCTATGCTTTCTTCGG
@@
 
-NA
-
 AGCGTGGAGGTGGGCCTAGACGAAAGAGCTGAAAA
@@
 
-NA
-
 AGCTCTTTCGTCTAGGCCCACCTCCACGCTGAATA
@@
 
-NA
-
 GCCGGTCTTGTCGATCAGGATGATCTGGAC
@@
 
-NA
-
 GTCCAGATCATCCTGATCGACAAGACCGGC
@@
 
-NA
-
 Promoter construct: injected at 250 ng/μl
@@
 
-NA
-
 Used for generating tdTomato-P2A-GCaMP6s mRNA
@@
 
-NA
-
 HCR
@@
 
-NA
-
-NA
-
 HCR
@@
 
-NA
-
-NA
-
 fluorescently labeled hairpins
@@
 
-NA
-
-NA
-
 morpholino
@@
 
-NA
-
 TGCTTGTCATTATTCAACCAGCAAA
@@
 
-NA
-
 TTCAATTAAACCCTCCAGGTTGCTG
@@
 
-NA
-
 AAATGAAGAGAGGTGTTCTCCTTCG
@@
 
-NA
-
 ATATTCATTATGTGAAGAACTTCCA
@@
 
-NA
-
 Plasmid
@@
 
-501 μM
+500 μM
+
+Chemical compound, drug
+
+DMEM, low glucose
+
+Thermofisher
+
+Cat\#:11885084
 
@@
 
-NA
-
-NA
-
 Commercial assay or kit
@@
 
-NA
-
-NA
-
 Software, algorithm
@@
 
-NA
-
 Commercial assay or kit
@@
 
-NA
-
 Software, algorithm
@@
 
-NA
-
 Software, algorithm
 
-PhyML
-
-\url{DOI:10.1093/sysbio/syq010}
-
-RRID:SCR\_014629
-
-NA
+IQ-tree2
+
+DOI: 10.1093/molbev/msaa131
+
+RRID:SCR\_017254
 
@@
 
-Gblocks
-
-\url{DOI:10.1080/10635150701472100}
-
-RRID:SCR\_015945
-
-NA
+TrimAL
+
+DOI: 10.1093/bioinformatics/btp348
+
+RRID:SCR\_017334
+
+Software, algorithm
+
+MAFFT
+
+DOI: 10.1093/molbev/mst010
+
+RRID:SCR\_011811
+
+Software, algorithm
+
+cd-hit
+
+\url{http://cd-hit.org}
+
+Software, algorithm
+
+clans
+
+DOI: 10.1093/bioinformatics/bth444
+
+Software, algorithm
+
+Hmmer
+
+\url{http://hmmer.org/}
+
+RRID:SCR\_005305
 
@@
 
-\textbf{CRISPR-Cas9 Design and Microinjection}
-
-Before designing the small guide RNA (sgRNA) for the sgRNA:Cas9
-nuclease, splice sites and polymorphic sites in our laboratory culture
-were identified to avoid them. The sgRNA targeted the third exon of
-Platynereis dumerilii NOS (target site: 5'-GGGCAATACTGGCTCCACTC-3'). The
-sgRNA was assembled from two annealed oligonucleotides
-(5'-TAGGGCAATACTGGCTCCACTC-3', 5'-AAACGAGTGGAGCCAGTATTGC-3') forming
-overhangs for cloning into a BsaI site of the plasmid pDR27456 (Hwang et
-al.~2013)(42250, Addgene), which contains next to the BsaI site a
-tracrRNA sequence. The plasmid was then used to PCR amplify DNA
-(primers: T7, 5'-AAAAGCACCGACTCGGTGCC-3') for synthesizing the sgRNA.
-The DNA was purified with the QIAquick PCR Purification Kit (Qiagen).
-From the DNA, the sgRNA was synthesized with the MEGAshortscript Kit
-(Thermo Fisher Scientific) and was purified with the MEGAclear Kit
-(Thermo Fisher Scientific). Cas9-mRNA was transcribed, capped, and
-polyA-tailed with the mMessage mMachine Kit and the Poly(A) Tailing Kit
-(both Thermo Fisher Scientific) from a plasmid (pUC57-T7-RPP2-Cas9)
-containing the Cas9 ORF fused to 169 base pair 5' UTR from the
-Platynereis dumerilii 60S acidic ribosomal protein P2. The sgRNA (18
-ng/ml) and the Cas9-mRNA (180 ng/µl) were coinjected into fertilized
-eggs of Platynereis dumerilii wild-type parents according to an
-established injection procedure (Conzelmann et al., 2013). The eggs were
-kept at 18°C for 45 min before injection and were injected at 14.5°C.
-The injected individuals were kept at 18°C for 5 to 8 days in 6-well-
-plates (Nunc multidish no. 150239, Thermo Scientific) and then cultured
-at 22°C until sexual maturity. The mature worms were crossed to
-wild-type worms and the progeny was genotyped, resulting in two founder
-lines, which were bred to homozygosity.
-
-\textbf{NOS sequencing and genotyping}
-
-For genotyping of the NOS locus, genomic DNA was isolated from single
-larvae, groups of 6-20 larvae, or from the tails of adult worms. The DNA
-was amplified by PCR (primers: 5'-GGTTCATTGGTTTCGATAACATTGCGG-3',
-5'-CAGAGTCGATCAGTCTGCATATCTCCA-3') with the dilution protocol of the
-Phusion Human Specimen Direct PCR Kit (Thermo Scientific). The PCR
-product was sequenced directly with a nested sequencing primer
-(5'-GGTGCTCTCCCGGGTACACAA-3'). A mixture of wild-type and deletion
-alleles in a sample gave double peaks in the sequencing chromatograms,
-with the relative height of the double peaks reflecting the relative
-allele ratio in the sample.
-
-\textbf{Vertical column setup for measuring photoresponses}
-
-Photoresponses of larvae of different ages were assayed in a vertical
-Plexiglas column (31 mm x 10 mm x 160 mm water height). The column was
-illuminated from top with light from a monochromator (Polychrome II,
-Till Photonics). The monochromator was controlled by AxioVision 4.8.2.0
-(Carl Zeiss MicroImaging GmbH) via analog voltage. The light passed a
-collimator lens (LAG-65.0-53.0-C with MgF2 Coating, CVI Melles Griot)
-before entering the column. The column was illuminated from both sides
-with light-emitting diodes (LEDs). The LEDs on each side were grouped
-into two strips. One strip contained UV (395 nm) LEDs (SMB1W-395,
-Roithner Lasertechnik) and the other infrared (810 nm) LEDs
-(SMB1W-810NR-I, Roithner Lasertechnik). The UV LEDs were run at 4 V to
-stimulate the larvae in the column from the side. The infrared LEDs were
-run at 8 V (overvoltage) to illuminate the larvae for the camera (DMK
-22BUC03, The Imaging Source), which recorded videos at 15 frames per
-second and was controlled by IC Capture (The Imaging Source).
-
-\textbf{Comparing behavior of wildtype and NOS-knockout 3-day-old
-larvae}
-
-To compare the behavior of wildtype and NOS-knockout larvae at 3 days in
-the vertical column, the larvae were mixed and left in the dark for 5
-min. The larvae were treated with NOS inhibitors for pharmacology. The
-NOS inhibitors were L-NAME. The larvae were treated with different
-concentrations in adjacent columns. The concentrations for the NOS
-inhibitors were control, 1 mM, 0.1 mM. The larvae were recorded for 1
-min in the dark followed by exposure to collimated cyan (480 nm) light
-from the top of the column for 2 min, then 2 min darkness, and finally
-collimated UV (395 nm) light from the top of the column for 2 min.
-Stimulus light was provided by the monochromator (Polychrome II, Till
-Photonics). Scripts are available at
-\url{https://github.com/JekelyLab/NOS}.
-
-\textbf{NOS Identification and Phylogenetic Analysis}
-
-To identify NOS, we obtained a ``seed'' database of oxygenase domain in
-Pfam database, PF02898. From these sequences, we produced a Hidden
-Markov Model (HMM) and used this to mine the 47 metazoan species, 2
-choanoflagellate species and 2 filasterea species investigated. HMM
-models were run in HMMR3 with an e-value of 1e−15. We ran CD-Hit (Fu et
-al., 2012) to eliminate redundant sequences (at a 80\% threshold). We
-aligned the sequences with MAFFT version 7, with the iterative
-refinement method E-INS-i. Alignments were trimmed with TrimAl in
-gappy-out mode (Capella-Gutierrez et al., 2009). To calculate
-maximum-likelihood trees, we used IQ-tree2 with the LG+G4 model. To
-calculate branch support, we ran 1,000 replicates with the aLRT-SH-like
-and aBayes methods (Minh et al., 2020). The sequences used for the
-phylogenetic analysis are available in \textbf{Supplementary File 1},
-the trimmed alignment is available in \textbf{Supplementary File 2} and
-the pre-trimmed data in \textbf{Supplementary file 3}.
-
-\textbf{NIT-GC Identification and Phylogenetic Analysis}
-
-To identify NIT-GCs, we obtained a ``seed'' database of Adenylate and
-Guanylate cyclase catalytic domain in Pfam database, PF00211. From these
-sequences, we produced a Hidden Markov Model (HMM) and used this to mine
-the 45 metazoan species, 2 choanoflagellate species and 2 filasterea
-species investigated. HMM models were run in HMMR3 with an e-value of
+\textbf{CRISPR-Cas9 design and Microinjection}
+
+For the generation of a \emph{NOS} knockout, we used an sgRNA targeting
+the third exon of the \emph{Platynereis dumerilii} \emph{NOS} gene
+(target site: 5'-GGGCAATACTGGCTCCACTC-3'). We selected the site to avoid
+polymorphic sites in our laboratory culture. The sgRNA was assembled
+from two annealed oligonucleotides (5'-TAGGGCAATACTGGCTCCACTC-3',
+5'-AAACGAGTGGAGCCAGTATTGC-3') forming overhangs for cloning into the
+BsaI site of the plasmid pDR27456 \href{42250,\%20Addgene}{@hwang2013},
+which contains next to the BsaI site a tracrRNA sequence. We used this
+plasmid to PCR amplify DNA (primers: T7, 5'-AAAAGCACCGACTCGGTGCC-3') for
+synthesizing the sgRNA. To purify the PCR product, we used the QIAquick
+PCR Purification Kit (Qiagen). We then synthesized sgRNA with the
+MEGAshortscript Kit (Thermo Fisher Scientific) and purified it with the
+MEGAclear Kit (Thermo Fisher Scientific). To produce the Cas9-mRNA, we
+used the plasmid (pUC57-T7-RPP2-Cas9) containing the Cas9 ORF fused to
+169 base pair 5' UTR from the \emph{Platynereis dumerilii} 60S acidic
+ribosomal protein P2. After in vitro transcription with the mMessage
+mMachine Kit (Thermo Fisher Scientific), we capped and polyA-tailed the
+Cas9-mRNA with the Poly(A) Tailing Kit (Thermo Fisher Scientific). We
+coinjected the sgRNA (18 ng/ml) and the Cas9-mRNA (180 ng/µl) into
+fertilized eggs of \emph{Platynereis dumerilii} wild-type parents
+according to an established injection procedure (Conzelmann et al.,
+2013a). Eggs were kept at 18°C for 45 min before injection and were
+injected at 14.5°C. The injected individuals were kept at 18°C for 5 to
+8 days in 6-well-plates (Nunc multidish no. 150239, Thermo Scientific)
+and then cultured at 22°C until sexual maturity. The mature worms were
+crossed to wild-type worms and the progeny was genotyped, resulting in
+two founder lines, which were bred to homozygosity.
+
+\textbf{Genotyping of NOS alleles}
+
+For genotyping of the \emph{NOS} locus, genomic DNA was isolated from
+single larvae, groups of 6-20 larvae, or from the tails of adult worms.
+The DNA was amplified by PCR (primers:
+5'-GGTTCATTGGTTTCGATAACATTGCGG-3', 5'-CAGAGTCGATCAGTCTGCATATCTCCA-3')
+with the dilution protocol of the Phusion Human Specimen Direct PCR Kit
+(Thermo Scientific). The PCR product was sequenced directly with a
+nested sequencing primer (5'-GGTGCTCTCCCGGGTACACAA-3'). A mixture of
+wild-type and deletion alleles in a sample gave double peaks in the
+sequencing chromatograms, with the relative height of the double peaks
+reflecting the relative allele ratio in the sample.
+
+\textbf{Morpholino-mediated knockdowns}
+
+Morpholino injections were performed as previously described (Conzelmann
+et al., 2013b). We used the following morpholinos to target the NIT-GC1
+and NIT-GC2 genes (GeneTools, LLC). NIT-GC1 MO1:
+TGCTTGTCATTATTCAACCAGCAAA, NIT-GC1 MO2: TTCAATTAAACCCTCCAGGTTGCTG,
+NIT-GC2 MO1: AAATGAAGAGAGGTGTTCTCCTTCG, NIT-GC2 MO2:
+ATATTCATTATGTGAAGAACTTCCA.
+
+\textbf{Vertical column setup to measure photoresponses}
+
+We assayed larval photoresponses in a vertical Plexiglas column (31 mm x
+10 mm x 160 mm water height). For stimulation, we illuminated the column
+from above with a monochromator (Polychrome II, Till Photonics)
+controlled by AxioVision 4.8.2.0 (Carl Zeiss GmbH, Jena) via analog
+voltage. The light passed a collimator lens (LAG-65.0-53.0-C with MgF2
+Coating, CVI Melles Griot) before entering the column. The column was
+illuminated from both sides with light-emitting diodes (LEDs). The LEDs
+on each side were grouped into two strips. One strip contained UV (395
+nm) LEDs (SMB1W-395, Roithner Lasertechnik) and the other infrared (810
+nm) LEDs (SMB1W-810NR-I, Roithner Lasertechnik). The UV LEDs were run at
+4 V. The infrared LEDs were run at 8 V (overvoltage) to illuminate the
+larvae for imaging with a DMK camera (DMK 22BUC03, The Imaging Source).
+We recorded videos at 15 frames per second with the software IC Capture
+(The Imaging Source).
+
+We compared the behavior of wildtype and \emph{NOS}-knockout larvae in
+the vertical column. In the inhibitor experiments we compared wildtype
+larvae before and after treatment with 0.1 mM or 1.0 mM NOS inhibitor
+L-NAME (in seawater). All larvae were three days oldt. Before the
+experiment, the larvae were mixed and left in the dark for 5 min. The
+larvae were recorded for 1.5 min in the dark followed by exposure to
+diffuse UV light (395 nm) from the side for 2 min. Then the larvae were
+left for another 2 min in darkness followed by exposue to collimated
+cyan (480 nm) light from above for 2 min, then 2 min darkness, and
+finally collimated UV (395 nm) light from above for 2 min. Scripts are
+available at \url{https://github.com/JekelyLab/Jokura_et_al_NOS}.
+
+\textbf{Identification and Phylogenetic Analysis of NOS sequences}
+
+To identify NOS sequences, we obtained a seed database of oxygenase
+domains from the Pfam database (PF02898). From these sequences, we
+generated a Hidden Markov Model (HMM) to mine the genomes and
+transcriptomes of 47 metazoan, 2 choanoflagellate and 2 filasterea
+species. The HMM model was run in HMMR3 (Eddy, 2011) with an e-value of
 1e−15. We ran CD-Hit (Fu et al., 2012) to eliminate redundant sequences
-(at a 80\% threshold). To identify clusters, we used the
-convex-clustering option with 100 jack-knife replicates. The NIT-GCs are
-extremely well conserved in membrane-bound guanylate cyclases and form
-an easily recognizable cluster. To analyze the phylogeny of NIT-GCs, the
-cluster containing these GCs together with membrane-bound guanylate
-cyclases were parsed and used for tree building. We aligned the
-sequences with MAFFT version 7, with the iterative refinement method
-E-INS-i. Alignments were trimmed with TrimAl in gappy-out mode
-(Capella-Gutierrez et al., 2009). To calculate maximum-likelihood trees,
-we used IQ-tree2 with the LG+G4 model. To calculate branch support, we
-ran 1,000 replicates with the aLRT-SH-like and aBayes methods (Minh et
-al., 2020). The sequences used for the phylogenetic analysis are
-available in \textbf{Supplementary File 1}, the trimmed alignment is
-available in \textbf{Supplementary File 2} and the pre-trimmed data in
-\textbf{Supplementary file 3}.
+(at a 80\% threshold). We aligned the sequences with MAFFT version 7,
+with the iterative refinement method E-INS-i (Katoh and Toh, 2008). The
+alignment was trimmed with TrimAl in gappy-out mode (Capella-Gutierrez
+et al., 2009). To calculate a maximum-likelihood tree, we used IQ-tree2
+with the LG+G4 model with the following options: iqtree2 -s ``FILENAME''
+-mset WAG, LG, Blosum62, Dayhoff, JTT, Poisson -B 1000 -alrt 1000 -T
+AUTO. Branch-support values are based on 1000 replicates with the
+ultrafast bootstrap approximation (UFBoot) (Hoang et al., 2018) and
+aLRT-SH-like (Guindon et al., 2010) methods.
+
+\textbf{Identification and Phylogenetic Analysis of NIT-GC sequences}
+
+To identify NIT-GCs, we obtained a seed database of adenylate and
+guanylate cyclase catalytic domains from the Pfam database (PF00211).
+From these sequences, we generated a Hidden Markov Model (HMM) to mine
+the 45 metazoan, 2 choanoflagellate and 2 filasterea genomes and
+transcriptomes. The HMM model was run in HMMR3 (Eddy, 2011) with an
+e-value of 1e−15. We ran CD-Hit (Fu et al., 2012) to eliminate redundant
+sequences (at a 80\% threshold). To identify clusters, we used
+sequence-similarity-based clustering in Clans (Frickey and Lupas, 2004)
+and the convex-clustering option with 100 jack-knife replicates. The
+NIT-GCs are extremely well conserved within the membrane-bound guanylate
+cyclases and form an easily recognizable cluster. To analyze the
+phylogeny of NIT-GCs, the cluster containing these GCs together with
+membrane-bound guanylate cyclases were parsed and used for tree
+building. We aligned the sequences with MAFFT version 7, with the
+iterative refinement method E-INS-i (Katoh and Toh, 2008). The alignment
+was trimmed with TrimAl in gappy-out mode (Capella-Gutierrez et al.,
+2009). To calculate a maximum-likelihood tree, we used IQ-tree2 with the
+LG+G4 model with the following options: iqtree2 -s ``FILENAME'' -mset
+WAG, LG, Blosum62, Dayhoff, JTT, Poisson -B 1000 -alrt 1000 -T AUTO.
+Branch-support values are based on 1000 replicates with the ultrafast
+bootstrap approximation (UFBoot) (Hoang et al., 2018) and aLRT-SH-like
+(Guindon et al., 2010) methods.
 
@@
 
-We used Achim et al.~for the single-cell data (Achim et al., 2015). In
-Williams et al., they used 107 cells as neurons by removing duplicates
-from Achim et al.~single-cell data, so we used those cells (Williams et
-al., 2017). Since the raw data were read count data, we normalized them
-to TPM using Python. After that we converted them to log10. From the sum
-of the expression levels in 107 cells for each gene, We calculated the
-percentage in each cell. For each cell, we identified them with marker
-genes. After created the data in Python, plotted it using R dot plots.
-RPKM calculates the total number of reads per million bp and then
-divides by the length of each gene, so it is not possible to compare
-between samples. Instead, TPM first divides by the length of the gene
-and then divides by the total number of reads per million bases, which
-allows for more accurate comparisons between samples. In this case we
-wanted to compare between samples, so we used TPM. The total TPM of each
-gene between the samples was used to calculate the percentage of
-expressed genes. The total TPM values for each gene and the percentage
-of expressed genes are available in \textbf{Supplementary file 1}.
+We used a single-cell RNA sequencing dataset from \emph{Platynereis}
+larvae (Achim et al., 2015) that was further filtered to 107 cells
+(Williams et al., 2017). We normalized the raw read-count data to TPM
+and converted to log10. To calculate \% expression, we used the sum of
+the expression levels in the 107 cells for each gene. The total TPM of
+each gene between the samples was used to calculate the percentage of
+expressed genes. We processed the data in Python and plotted in R.
+Individual cells were identified by suits of well-characterised marker
+genes and spatial mapping (Achim et al., 2015).
 
@@
 hairpins were purchased from Molecular Instruments (Los Angeles, USA).
-Hairpins associated with the b2 initiator sequence were labeled with
+The hairpins associated with the b2 initiator sequence were labeled with
 Alexa Fluor 647, and the hairpins associated with the b3 initiator
@@
 used custom software (Kuehn et al., 2021) to create 20 DNA oligo probe
-pairs specific to P. dumerilii NOS, NIT-GC1, NIT-GC2, RYa-pNP (GenBank
-accession: JF811330.1), and MLD/pedal 2-pNP (GenBank accession:
-KF515945.1). The NOS, NIT-GC1 and NIT-GC2 probes were designed to be
-associated with the b2 initiator sequence, while the RYa-pNP and
-MLD/pedal 2-pNP probes were designed to be associated with the b3
-initiator sequence. For the detection stage, samples were pre-hybridized
-in 200 µl of probe hybridization buffer for 1 hr at 37°C, and then
-incubated in 250 µl hybridization buffer containing probe oligos (4
-pmol/ml) overnight at 37°C. To remove excess probe, samples were washed
-4× with 1 ml hybridization wash buffer for 15 min at 37°C, and
-subsequently 2× in 1 ml 5× SSCT (5× SSC with 0.1\% Tween20) for 5 min at
-room temperature. For the amplification stage, samples were
-pre-incubated with 100 µl of amplification buffer for 30 min, room
-temperature, and then incubated with 150 µl amplification buffer
-containing fluorescently labeled hairpins (40nM concentration (2ul of
-3uM stock in 150ul amplification buffer, snap-cooled as described; (Choi
-et al., 2018)) overnight in the dark at 25°C. To remove excess hairpins,
-samples were washed in 1 ml 5× SSCT at room temperature, twice for 5
-min, twice for 30 min, and once for 5 min. During the first 30 min wash,
-samples were counterstained with DAPI (Cat. \#40043, Biotium, USA).
+pairs specific to \emph{P. dumerilii} \emph{NOS}, \emph{NIT-GC1},
+\emph{NIT-GC2}, \emph{RYa proneuropeptide} (GenBank accession:
+JF811330.1), and \emph{MLD/pedal peptide 2 proneuropeptide} (GenBank
+accession: KF515945.1). The \emph{NOS}, \emph{NIT-GC1} and
+\emph{NIT-GC2} probes were designed to be associated with the b2
+initiator sequence, while the \emph{RYa} and \emph{MLD/pedal peptide 2}
+probes were designed to be associated with the b3 initiator sequence.
+For the detection stage, samples were pre-hybridized in 200 µl of probe
+hybridization buffer for 1 hr at 37°C, and then incubated in 250 µl
+hybridization buffer containing probe oligos (4 pmol/ml) overnight at
+37°C. To remove excess probe, samples were washed 4× with 1 ml
+hybridization wash buffer for 15 min at 37°C, and subsequently 2× in 1
+ml 5× SSCT (5× SSC with 0.1\% Tween20) for 5 min at room temperature.
+For the amplification stage, samples were pre-incubated with 100 µl of
+amplification buffer for 30 min, room temperature, and then incubated
+with 150 µl amplification buffer containing fluorescently labelled
+hairpins (40nM concentration (2ul of 3uM stock in 150ul amplification
+buffer, snap-cooled as described; (Choi et al., 2018)) overnight in the
+dark at 25°C. To remove excess hairpins, samples were washed in 1 ml 5×
+SSCT at room temperature, twice for 5 min, twice for 30 min, and once
+for 5 min. During the first 30 min wash, samples were counterstained
+with DAPI (Cat. \#40043, Biotium, USA).
 
@@
 
-Whole-mount immunostaining of 2 day old Platynereis larvae fixed with
-4\% paraformaldehyde were carried out using primary antibodies raised
-against NIT-GC1, NIT-GC2, NOS, RYamide neuropeptide, RGWamide
-neuropeptide in rabbit, plus a commercial antibody raised against
-acetylated tubulin in mouse (Sigma T7451). The synthetic peptides
-contained an N-terminal Cys that was used for coupling during
-purification. Antibodies were affinity purified from sera as previously
-described (Conzelmann and Jékely, 2012). Immunostainings were carried
-out as previously described (Conzelmann and Jékely, 2012). The NOS
-promoter (fragment sizes: 12 Kb) was amplified and cloned upstream of
-3xHA- Palmi-tdTomato. Larvae injected with promoter constructs (ca. 250
-ng/ml) were analysed for reporter expression at 3 days post
-fertilization using an AxioImager Z.1 fluorescence wide-field microscope
-(Carl Zeiss GmbH, Jena) and immediately fixed for immunostainings. The
-protocol followed for immunostaining of HA-tagged reporters was recently
-described (Verasztó et al., 2017). Specimens were imaged with a LSM 780
-NLO or LSM 880 with Airysan Confocal Microscope (Zeiss, Jena).
+Whole-mount immunostaining of \emph{Platynereis} larvae fixed with 4\%
+paraformaldehyde was carried out using the primary antibodies listed in
+the KEY RESOURCES TABLE. Immunostainings were carried out as previously
+described (Conzelmann and Jékely, 2012).
+
+\textbf{Antibody generation and purification}
+
+Antibodies were raised in rats or rabbits against synthetic peptides
+containing an N-term Cys residue (Altabioscience). The same
+Cys-containing synthetic peptides were used for affinity purification.
+Antibodies were affinity purified from sera on a SulfoLink Coupling
+Resin (Thermo Fischer) as previously described (Conzelmann and Jékely,
+2012). View a detailed protocol here:
+\url{https://bio-protocol.org/exchange/preprintdetail?id=2333\&type=3}
+
+\textbf{Transient transgenesis}
+
+A 12 kb genomic region upstream of the start site of the \emph{NOS} gene
+was amplified and cloned upstream of 3xHA-Palmi-tdTomato to yield the
+plasmid \emph{NOS-3xHA-Palmi-tdTomato}. Larvae injected with the
+promoter construct (ca. 250 ng/ml) were analysed for reporter expression
+at three days post fertilization on an AxioImager Z.1 fluorescence
+microscope (Carl Zeiss GmbH, Jena) and if positive were fixed for
+immunostaining. The immunostaining for the HA-tagged reporter was done
+as described (Verasztó et al., 2017). Stained specimens were imaged on
+an LSM 780 NLO or LSM 880 Airysan Confocal Microscope (Carl Zeiss GmbH,
+Jena).
 
@@
 
-For calcium imaging, 49--55 hpf larvae were used. Experiments were
-performed at room temperature and larvae were immobilised by being
-embedded in 2.5\% agarose filtered artificial seawater between a slide
-and coverslip spaced with adhesive tape. GCaMP6s mRNA (1 mg/ml) was
-injected into zygotes as described previously (Randel et al., 2014).
-Larvae were imaged on a Zeiss LSM 880 with Airyscan (with a C-Apochromat
-63X/1.2 Corr - water) with a frame rate of 1.88 frame/sec and an image
-size of 512 x 512 pixels. The larvae were stimulated in a region of
-interest (a circle with 50 pixel diameter) with 405 nm lasers controlled
-by the Bleaching mode. The imaging laser had a similar intensity than
-the stimulus laser but covered an area that was 10 times larger than the
+For Ca\textsuperscript{2+} imaging, GCaMP6s mRNA (1 mg/ml) was injected
+into zygotes as described previously (Randel et al., 2014). At 49-55
+hours post fertilisation, larvae were immobilised in 2.5\% agarose in
+artificial sea water and mounted between a slide and coverslip spaced
+with adhesive tape. Larvae were imaged at room temperature on a Zeiss
+LSM 880 with Airyscan (with a C-Apochromat 63X/1.2 Corr - water) with a
+frame rate of 1.88 frame/sec and an image size of 512 x 512 pixels. The
+larvae were stimulated in a region of interest (a circle with 50 pixel
+diameter) with a 405 nm laser controlled by the bleaching mode of the
+Zeiss software. The imaging laser had a similar intensity than the
+stimulus laser but covered an area that was 10 times larger than the
 stimulus ROI.
 
-\textbf{Cell culture experiment}
-
-Green cGull was used for the cGMP assay (Matsuda et al., 2016). A
-full-length Pdum-NIT-GC1 and -NIT-GC2 coding sequences were amplified by
-PCR starting from a \emph{Platynereis dumerilii} cDNA library and cloned
-into the pcDNA3.1(+) vector using the T2A self-cleaving sequence. Cos-7
-cells with low expression of endogenous soluble guanylate cyclase were
-used as cultured cells for gene expression. This cell line was purchased
-from Angio-proteomie (CAT no. cAP-0203). The Cos-7 cells were maintained
-at 37 °C in 35mm dishes (Nunc™ Glass Bottom Dishes) containing 3 mL of
-DMEM, high glucose glutamax medium (Thermo; Cat. No.~10566016)
-supplemented with 10\% fetal bovine serum (Thermo; Cat. No.~10082147).
-Upon reaching confluency of approximately 85\%, we transfected the cells
-with the plasmid containing Green cGull-T2A-NITGC1. Transfections were
-carried out with 150 ng of each plasmid and 0.3 μl of the transfection
-Lipofectamine 3000 Reagent (invitrogen; Cat. No.~L3000001). Two days
-post-transfection, we removed the culture medium and substituted it for
-fresh DMEM-medium. For single-wavelength imaging experiments, cells in
-35-mm dishes were washed twice and imaged in modified Ringer's buffer
-(140 mM NaCl, 3.5 mM KCl, 0.5mM NaH2PO4 , 0.5mM S-3 MgSO4 , 1.5 mM CaCl2
-, 10 mM HEPES, 2 mM NaHCO3 and 5 mM glucose). Dishes were mounted on a
-stage heated at 37 °C and imaging was performed using an inverted
-microscope (LSM880, Zeiss) equipped with an oil-immersion objective lens
-(UApo/340, 40, NA = 0.17). Images were acquired using a xenon lamp,
-460--495 nm excitation filter, 505-nm dichroic mirror and 510-- 550-nm
-emission filter (Zeiss). S-Nitroso-N-acetyl-D, L-penicillamine (SNAP)
-was purchased from Sigma-Aldrich (St.~Louis, MO, USA) . The exposure
-time of the EM-CCD camera was controlled by the ZEN software (Zeiss).
-Images were acquired every 15 s for 10 min and stimulation was initiated
-2 min after starting image acquisition. Imaging data analysis was
-performed using ImageJ (National Institutes of Health, Bethesda, MD,
-USA).
+\textbf{cGMP production assay}
+
+To measure cGMP production in a cell culture assay we used the Green
+cGull reporter (Matsuda et al., 2016). Full-length \emph{NIT-GC1} and
+\emph{NIT-GC2} coding sequences were amplified by PCR from a
+\emph{Platynereis dumerilii} cDNA library and cloned into the
+pcDNA3.1(+) vector with the T2A self-cleaving sequence. For the assay,
+COS-7 cells (Angio-proteomie, CAT no. cAP-0203) with a low endogenous
+level of soluble guanylate cyclase activity were used. Cells were not
+contaminated with mycoplasma. Cells were maintained at 37 °C in 35 mm
+dishes (Nunc™ Glass Bottom Dishes) containing 3 ml of DMEM, low-glucose
+medium (Thermo; Cat. No.~11885084) supplemented with 10\% fetal bovine
+serum (Thermo; Cat. No.~10082147). Upon reaching approximately 85\%
+confluency, cells were transfected with the
+pcDNA3.1(+)-Green-cGull-T2A-NITGC1 plasmid. Transfections were carried
+out with 150 ng of each plasmid and 0.3 μl of Lipofectamine 3000 Reagent
+(invitrogen; Cat. No.~L3000001). Three hours post-transfection, the
+culture medium was substituted with fresh DMEM medium. For
+single-wavelength imaging experiments, cells in 35-mm dishes were washed
+twice and imaged in modified Ringer's buffer (140 mM NaCl, 3.5 mM KCl,
+0.5 mM NaH\textsubscript{2}PO\textsubscript{4} , 0.5 mM S-3
+MgSO\textsubscript{4}, 1.5 mM CaCl\textsubscript{2} , 10 mM HEPES, 2 mM
+NaHCO\textsubscript{3} and 5 mM glucose). The dishes were mounted on a
+stage heated to 37 °C and imaged on an inverted confocal microscope
+(LSM880, Carl Zeiss GmbH, Jena) equipped with an oil-immersion objective
+lens (UApo/340, 40x, NA = 0.17). The exposure time of the EM-CCD camera
+was controlled by the ZEN software (Carl Zeiss GmbH, Jena). Images were
+acquired every 15 s for 10 min and stimulation was initiated 2 min after
+starting image acquisition. Imaging data were processed with ImageJ
+(National Institutes of Health, Bethesda, MD, USA). As NO donor, we used
+S-Nitroso-N-acetyl-D, L-penicillamine (SNAP, Sigma-Aldrich, St.
+
+\hypertarget{mathematical-modelling}{%
+\subsection{Mathematical modelling}\label{mathematical-modelling}}
+
+\hypertarget{wild-type-model}{%
+\subsubsection{Wild type model}\label{wild-type-model}}
+
+The mathematical model for the wild type data comprises six dynamical
+equations:
+
+\begin{align*}
+\frac{d C_P(t)}{dt} &= 1 - \delta^{C_P} C_P(t) + K_{UV}^{C_P} UV(t) C_O(t)  + K_{G}^{C_P} G(t) , \\
+\frac{dC_O(t)}{dt} &= 1 - \delta^{C_O} C_O(t) - K_{UV}^{C_O} UV(t) C_O(t), \\
+\frac{dG(t)}{dt} &= 1 - \delta^{G} G(t) + K_{GC2}^G \left(1 - UV(t)\right) G(t) + K_{GC1}^G N(t) , \\
+\frac{dS(t)}{dt} &= UV(t) - \frac{S(t)}{\tau_S}, \\
+\frac{dC_N(t)}{dt} &= 1 - \delta^{C_N} C_N(t) + K_{S}^{C_N} S(t), \\
+\frac{dN(t)}{dt} &= 1 - \delta^{N} N(t) + K_{C_N}^{N} C_N(t), \\
+\frac{dC_R(t)}{dt} &= 1 - \delta^{C_R} C_R(t) - K_{S}^{C_R} S(t) + K_{C_P}^{C_R} C_P(t) - K_{C_N}^{C_R} C_N(t) C_R(t), \\
+UV(t) &=
+    \begin{cases}
+        a, & t_{UV\: start} \leq t \leq t_{UV\: end}\\
+        0, & \text{otherwise}
+    \end{cases}   .
+\end{align*}
+
+The \(C_P(t)\) equation describes the dynamics of the
+Ca\textsuperscript{2+} concentration in the cPRC cells. The
+\(K_{UV}^{C_P} UV(t) C_O(t)\) term captures the Ca\textsuperscript{2+}
+increase due to c-opsin1 activity induced by UV stimulation, while the
+\(K_{G}^{C_P} G(t)\) term captures dependence of the
+Ca\textsuperscript{2+} dynamics on cGMP. The \(C_O(t)\) equation is
+added to capture the decrease of the Ca\textsuperscript{2+} levels
+observed during the UV stimulation; the mechanism of this decrease is
+unknown. It could, for example, be related to the opening of the GIRK
+channels. The \(G(t)\) equation describes the dynamics of cGMP in the
+cPRC cells. The \(K_{GC1}^G N(t)\) term describes the NO-dependent
+increase in cGMP by NIT-GC1, while the
+\(K_{GC2}^G \left(1 - UV(t)\right) G(t)\) term captures the cGMP
+decrease mediated by NIT-GC2 during UV stimulation. The \(S(t)\)
+describes synaptic coupling from the cPRC cells, which is assumed to act
+in an excitatory fashion on INNOS cells and an inhibitory fashion on
+INRGW cells. This coupling is assumed to be suppressed in NIT-GC2
+morphants. The \(C_N(t)\) equation describes the Ca\textsuperscript{2+}
+dynamics in the INNOS cells. The \(K_{S}^{C_N} S(t)\) term describes the
+Ca\textsuperscript{2+} increase due to the synaptic signal \(S(t)\). The
+\(N(t)\) variable represents NO produced and secreted by the INNOS
+cells. The \(K_{C_N}^{N} C_N(t)\) term describes the
+Ca\textsuperscript{2+}-dependent NO production within the INNOS cells.
+The \(C_R(t)\) equation describes the dynamics of Ca\textsuperscript{2+}
+in the INRGW cells, which depends on the other equations in a
+feed-forward manner. The \(K_{S}^{C_R} S(t)\) and
+\(K_{C_N}^{C_R} C_N(t) C_R(t)\) terms describe the decrease in
+Ca\textsuperscript{2+} with respect to the synaptic signal \(S(t)\) and
+the Ca\textsuperscript{2+} levels in the INNOS cells, respectively. The
+\(K_{C_P}^{C_R} C_P(t)\) term describes the increase in
+Ca\textsuperscript{2+} levels in the INRGW cells due to increases in the
+Ca\textsuperscript{2+} levels in the cPRC cells. Note that all sources
+of Ca\textsuperscript{2+} variation in the INRGW cells are
+phenomenological in nature. All variables have a constant (unit)
+production term and a linear decay rate set by the parameter
+\(\delta_X\) where \(X \in \{ C_P, C_O, G, S, C_N, C_R\}\). Finally, the
+UV stimulation is described by a square pulse with amplitude \(a\) and
+duration \(t_{UV end} - t_{UV \: start}\).
+
+\hypertarget{nos-knockout-model}{%
+\subsubsection{NOS-knockout model}\label{nos-knockout-model}}
+
+The NOS-knockout model is derived from the WT model by eliminating the
+\(N(t)\) variable, which represents the NO produced by the INNOS cells
+in a NOS-dependent manner, and terms that depend on it. This gives the
+system \begin{align*}
+\frac{C_P(t)}{dt} &= 1 - \delta^{C_P} C_P(t) + K_{UV}^{C_P} UV(t) C_O(t) + K_{G}^{C_P} G(t) , \\
+\frac{C_O(t)}{dt} &= 1 - \delta^{C_O} C_O(t) - K_{UV}^{C_O} UV(t) C_O(t), \\
+\frac{G(t)}{dt} &= 1 - \delta^{G} G(t)+ \left(1 - UV(t)\right)K_{GC2}^G G(t), \\
+\frac{S(t)}{dt} &= UV(t) - \frac{S(t)}{\tau_S}, \\
+\frac{C_N(t)}{dt} &= 1 - \delta^{C_N} C_N(t) + K_{S}^{C_N} S(t), \\
+\frac{C_R(t)}{dt} &= 1 - \delta^{C_R} C_R(t) - K_{S}^{C_R} S(t) + K_{C_P}^{C_R} C_P(t) - K_{C_N}^{C_R} C_N(t) C_R(t), \\
+UV(t) &=
+    \begin{cases}
+        a, & t_{UV\: start} \leq t \leq t_{UV\: end}\\
+        0, & \text{otherwise}
+    \end{cases}   .
+\end{align*}
+
+We remark that the above system of equations can be used to also
+describe the NIT-GC1 morpholino condition (Fig. 4M).
+
+\hypertarget{nit-gc2-morpholino-model}{%
+\subsubsection{NIT-GC2 morpholino
+model}\label{nit-gc2-morpholino-model}}
+
+The NIT-GC2 morpholino model is a further restriction of the WT model:
+
+\begin{align*}
+\frac{C_P(t)}{dt} &= 1 - \delta^{C_p} C_P(t) + K_{UV}^{C_P} UV(t) C_O(t) , \\
+\frac{C_O(t)}{dt} &= 1 - \delta^{C_O} C_O(t) - K_{UV}^{C_O} UV(t) C_O(t), \\
+\frac{C_R(t)}{dt} &= 1 - \delta^{C_R} C_R(t) + K_{C_P}^{C_R} C_P(t), \\
+UV(t) &=
+    \begin{cases}
+        a, & t_{UV\: start} \leq t \leq t_{UV\: end}\\
+        0, & \text{otherwise}
+    \end{cases} .
+\end{align*}
+
+This reduced model eliminates the \(G(t)\), \(S(t)\), \(C_N(t)\) and
+\(N(t)\) variables and hence includes only c-opsin1 activity in the cPRC
+cells and the feedforward coupling to INRGW cells. The synaptic signal
+\(S(t)\) and all terms depending on it are removed due to the assumed
+dependence on NIT-GC2. When the \(S(t)\) input is disregarded, the
+activity of the INNOS cells (\(C_N(t)\) and \(N(t)\) variables) can be
+represented by a constant steady state whose values can subsequently be
+absorbed into all the terms depending on these variables. Hence these
+variables are eliminated in the reduced model. Since the
+\(K_{GC2}^G \left(1 - UV(t)\right) G(t)\) term depends directly on
+NIT-GC2 and when \(N(t)\) is ignored the \(K_{GC1}^G N(t)\) term
+disappears, the \(G(t)\) variable can be eliminated, which consequently
+simplifies the \(C_P(t)\) equation. The only remaining term in
+\(C_R(t)\) equation, \(K_{C_P}^{C_R} C_P(t)\), describes feedforward
+coupling between the cPRC and INRGW cells.
+
+\hypertarget{parameter-fitting-genetic-algorithm}{%
+\subsubsection{Parameter fitting (genetic
+algorithm)}\label{parameter-fitting-genetic-algorithm}}
+
+To find sets of model parameters that reproduce recordings from the
+Ca\textsuperscript{2+} imaging experiments, we define an optimisation
+problem to minimise the discrepancy between the modelled
+Ca\textsuperscript{2+} traces and the experimental data. To obtain the
+modelled Ca\textsuperscript{2+} traces, we run the model using the
+Matlab built-in solvers \texttt{fsolve} to find the steady state in the
+absence of UV stimulation, and then use the explicit integration
+algorithm \texttt{ode45} to simulate the model with the computed steady
+state as the initial condition.
+
+We quantify the difference between the model output and the data using a
+vector-valued objective function. The first component of this function
+measures this difference during the UV stimulation, whilst the second
+component measures the difference from the end of the stimulation until
+the end of the recording. We thus define a vector-valued objective
+function, \(\mathbf{D} \in \mathbb{R}^2_{\geq 0}\), as:
+\begin{equation*}
+ \mathbf{D}\left({C}_X^{data},{C}_X^{model}\right)=\left(D_{k_{UV\: start}}^{k_{UV\: end}}\left({C}_X^{data},{C}_X^{model}\right),D_{k_{UV\: end }+1}^{k_{end}}\left({C}_X^{data},{C}_X^{model}\right) \right),
+\end{equation*} with
+
+\begin{equation*}
+ D_{k_1}^{k_2}\left({C}_X^{data}(t),{C}_X^{model}(t)\right)=\sqrt{\sum_{i=k_1}^{k_2}\left({C}_{X}^{data}(t_i)-{C}_{X}^{model}(t_i)\right)^2} ,   
+\end{equation*} where the timepoints \(t_i = i \Delta t\) are equispaced
+with \(\Delta = 0.5s\) and where
+\(t_{k_{UV\: start}} = t_{UV \: start}\) (and similar for the other time
+points). Here \({C}_X^{data}\) denotes normalised experimental data and
+\({C}_X^{model}\) denotes normalised simulated Ca\textsuperscript{2+}
+trace with \(X \in \{ P, N, R\}\) indicating cell type. Simulated
+Ca\textsuperscript{2+} traces, \({C}_X^{model}\), are normalised in the
+same way as the experimental data, i.e., we divide values uniformly by
+their value at \(t_{k_{UV\: start}-1}\), a time point immediately prior
+to the start of the UV stimulation. Simulations are sampled with the
+same sampling rate as the data. (Recall that \(t_{UV\: start}\) is the
+time of the start of UV stimulation, \(t_{UV\: end}\) is the time of the
+end of UV stimulation, and \(t_{end}\) is the last time sample.)
+
+To solve the optimisation problem with the above objective function, we
+employ a global optimisation method known as a genetic algorithm (GA).
+GAs are a class of iterative, gradient-free algorithms inspired by
+processes of natural selection including breeding and mutation
+\cite{GAbook}. In reference to natural selection, sets of candidate
+solutions in GAs are referred to as `populations'. As our algorithm of
+choice, we use the Matlab implementation of a genetic algorithm for
+multiobjective optimisation \texttt{gamultiobj}. The \texttt{gamultiobj}
+algorithm uses a controlled, elitist genetic algorithm (a variant of
+NSGA-II \cite{GAbook}). We use the default settings for most
+hyperparameters of the algorithm.
+
+Here we outline the main steps of the \texttt{gamultiobj} implementation
+of the GA. The first step is to generate a set of random model parameter
+sets, known as the initial population. Next, we score the fitness of
+each parameter set in the population according to the objective
+function. Using these scores, the \texttt{gamultiobj} algorithm employs
+a binary tournament selection function\cite{Miller1995GeneticAT} to
+select `parents' for the next population. A population of children
+(i.e., new parameter sets) is generated by `breeding' randomly chosen
+pairs of parents (exchanging subsets of parameter values between them).
+Parameter values in these children can also be changed by `mutations',
+which perturb said values randomly by a small amount. Breeding and
+mutation introduce a level of randomness to the optimisation process
+that allows exploring the objective function (fitting) landscape. The
+stochastic nature of GAs helps them to avoid local minima that can
+plague gradient-descent-based optimisation methods.
+
+In the next step, the \texttt{gamultiobj} algorithm scores the parameter
+sets in the children population using the values of the objective
+function. Scores from the parent and child populations are combined. A
+subset of the size of the initial population is then selected as a new
+parent population based on the scores and the
+\textit{Crowding Distance}, distance to other parameter sets in terms of
+a sum of element-wise differences between values of the objective
+function (used to ensure diversity of the parent
+population)\cite{crowding_distance,GAbook}. This iterative process is
+repeated until the convergence criteria are met, indicating that the GA
+has found a set of model parameters that best reproduce the data under
+consideration.
+
+The \texttt{gamultiobj} algorithm returns a set of model parameter
+values on a Pareto front \cite{censor1977pareto,GAbook}. The coordinates
+of points on the front are the values of the components of the objective
+function \(\mathbf{D}\). Points on the Pareto front are nondominated,
+meaning that they are closest to 0 for at least one of the two
+components. To identify a single optimal solution, we select the
+parameter set along the Pareto front whose objective function value is
+the smallest with respect to the Euclidean norm.
+
+Below we present a summary of the steps of our optimisation pipeline for
+fitting the experimental recordings from the cPRC cells.
+
+\begin{itemize}
+\tightlist
+\item
+  When fitting to the cPRC data using WT or NIT-GC2 morpholino models,
+  we eliminate the \(C_R(t)\) variable since it does not have any
+  feedback coupling with the \(C_P(t)\) equation and does not affect its
+  dynamics. When fitting to the cPRC data using the NOS-knockout model,
+  we also eliminate the \(S(t)\), \(C_N(t)\) and \(C_R(t)\) variables
+  since these do not have any effect on the \(C_P(t)\).
+\item
+  Set the size of the initial population to 1,000. This value provides a
+  compromise in terms of the coverage of the parameter space and
+  run-time of the GA.
+\item
+  Set bounds of random distributions of parameters. We draw all the
+  initial parameter values from uniform distributions over the unit
+  interval ({[}0,1{]}). For WT condition, we fit 13 parameters, for
+  NOS-knockout, we fit 7 parameters and for NIT-GC2 morpholino, we fit 4
+  parameters.
+\item
+  Evaluate the objective function on the initial population.
+\item
+  Run the \texttt{gamultiobj} algorithm.
+\item
+  Repeat steps 1 to 4 with bounds of parameters extended to {[}0,5{]}
+  and using the Pareto front returned in step 4 as a part of the initial
+  population. This step was added after observing that, although good
+  fits to most datasets are achieved with parameter values in the
+  interval {[}0,1{]}, fits to some datasets can be improved by extending
+  the admissible parameter range.
+\item
+  Use the Euclidean norm to find the set of model parameters that best
+  fits the data.
+\end{itemize}
+
+Since INNOS cells in the NOS-knockout condition and INRGW cells in all
+conditions are coupled in a feedforward manner and do not affect the
+`up-stream' dynamics, we run a separate optimisation to fit data from
+these cells. This minimises the number of parameters that are being
+fitted simultaneously and ensures that the objective function is as
+simple as possible (using only recordings from a cell type that is being
+fitted). First, we fix the parameters that were found by fitting the
+model to the cPRC data. Next, we run the optimisation in the same way as
+in the case of the cPRC data but only for the parameters that pertain to
+dynamics that is strictly`downstream' of cPRC dynamics.
+
+Fixing a subset of parameters allows us to ensure that model parameters
+inferred from recordings of one cell type stay relevant when fitting the
+model to the data from other cells. This means that the fitted model is
+extended to include other cell types, rather than being completely
+reparamatrised for each cell type. This is particularly important since,
+for technical reasons, we do not have simultaneous recordings from
+multiple cell types.
+
+We note that in the WT condition, model parameters of the INNOS cells
+are fitted without using any INNOS recordings. Since the NO production
+feeds back to the cPRC cells, their parameters are fitted to reproduce
+the recordings from the cPRC cells. Furthermore, in the NIT-GC2
+morpholino condition, we do not fit the INNOS data since we assumed that
+the feedforward coupling between the cPRC cells and INNOS cell depends
+on NIT-GC2 activity (that is, we assume that the Ca\textsuperscript{2+}
+levels in INNOS cells are constant in this condition).
+
+All the simulated curves presented in Fig. 6B are the averages of the
+curves fitted to individual experimental recordings/datasets.
+
+\hypertarget{predictions-from-a-range-of-stimulation-protocols}{%
+\subsubsection{Predictions from a range of stimulation
+protocols}\label{predictions-from-a-range-of-stimulation-protocols}}
+
+To explore possible effects of varying the intensity and duration of the
+UV stimulation on experimentally observed delayed activation in the cPRC
+cells, we perform a series of simulations in which we systematically
+varied amplitude (\(a \in [0.1,1]\)) and width
+(\(t_{UV\: end} \in [0.1,51]\)) of the \(UV(t)\) variable in the WT
+model. To quantify the effects of the stimulation, we measure the area
+under the peak of the delayed activation \(AUP\) in the \(C_P(t)\)
+variable. To compute the \(AUP\):
+
+\begin{itemize}
+\tightlist
+\item
+  We find an inflection point that occurs after the start
+  (\(t_{UV\: start}\)) of the UV stimulation. To this end, we identify
+  the first time point \(t_{AUP\: start}>t_{UV\: start}\) for which
+  \(d^2C_P(t)/dt^2=0\) and
+  \(d^2C_P(t_{k_{AUP\: start}}-1)/dt^2<d^2C_P(t_{k_{AUP\: start}})/dt^2<d^2C_P(t_{k_{AUP\: start}}+1)/dt^2\).
+  We assume that the inflection is caused by the input from the INNOS
+  cells.
+\item
+  We then compute the \(AUP\) as an integral over \(C_P(t)\) starting at
+  the inflection point and ending 125 seconds later:
+  \$AUP=\int*\{t*\{start\}\}\^{}\{t\_\{end\}\} C\_P(t) dt \$. To compute
+  this, we use numerical integration via the trapezoidal method as
+  implemented in Matlab function \texttt{trapz}. We remark that the
+  baseline value of the \(AUP\) for a Ca\textsuperscript{2+} profile of
+  constant, unit value is 125.
+\item
+  To capture possible variability between different cells/organisms, we
+  compute the \(AUP\) using 47 model parameter sets fitted to individual
+  WT cPRC experimental recordings. We remove outliers, defined as being
+  any simulated traces for which \(C_P(t)\), after normalisation, exits
+  the interval {[}0.25, 4{]} at any time. This \(C_P(t)\) range was
+  decided of the basis of values observed in various experiments in the
+  lab.
+\item
+  For analysis, we use the median value of the \(AUP\) at each
+  combination of amplitude and duration.
+\end{itemize}
+
+Figure 7 shows how the results of the simulations of the WT model depend
+on the amplitude and duration of UV stimulation, as captured by the
+median values of \(AUP\). Delayed activation occurs only for UV
+stimulation of sufficient amplitude and duration. Specifically, we
+observe sudden increase in AUP for UV pulse amplitude \(a>0.5\) and a
+linear dependence of AUP on the pulse duration.
+
+\hypertarget{data-analysis-and-figure-preparation}{%
+\subsection{Data analysis and figure
+preparation}\label{data-analysis-and-figure-preparation}}
+
+Data were analysed in ImageJ and R. All figures were assembled in R with
+the cowplot and patchwork packages. All scripts are available at
+\url{https://github.com/JekelyLab/Joukra_et_al_NOS} (commit: c223fe3).
 
@@
 European Union's Horizon 2020 research and innovation programme (grant
-agreement No 101020792). KJ has been supported by a JSPS Postdoctoral
-fellowhip, LAYG by a BBSRC Discovery fellowship (BB/W010305/1).
+agreement No 101020792). KJ has been supported by a JSPS Overseas
+Research Fellowship, LAYG by a BBSRC Discovery fellowship
+(BB/W010305/1), PS by a Wellcome Trust Institutional Strategic Support
+Award (204909/Z/16/Z), and KCAW by the EPSRC Hub for Quantitive
+Modelling in Healthcare (EP/T017856/1). We thank Dr.~Francesca Carlisle
+for the mycoplasma testing of COS-7 cells.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Capella-Gutiuxe9rrez2009}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-Capella-Gutierrez2009}{}}%
 Capella-Gutierrez S, Silla-Martinez JM, Gabaldon T. 2009. trimAl: a tool
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Conzelmann2013a}{}}%
-Conzelmann M, Williams EA, Krug K, Franz-Wachtel M, Macek B, Jékely G.
-2013. The neuropeptide complement of the marine annelid Platynereis
-dumerilii. \emph{BMC Genomics} \textbf{14}:906.
+Conzelmann M, Williams Elizabeth A, Krug K, Franz-Wachtel M, Macek B,
+Jékely G. 2013a. The neuropeptide complement of the marine annelid
+Platynereis dumerilii. \emph{BMC Genomics} \textbf{14}:906.
 doi:\href{https://doi.org/10.1186/1471-2164-14-906}{10.1186/1471-2164-14-906}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Conzelmann2013}{}}%
+Conzelmann M, Williams Elizabeth A, Tunaru S, Randel N, Shahidi R,
+Asadulina A, Berger J, Offermanns S, Jékely G. 2013b. Conserved MIP
+receptor{\textendash}ligand pair regulates {\emph{Platynereis}} larval
+settlement. \emph{Proceedings of the National Academy of Sciences}
+\textbf{110}:8224--8229.
+doi:\href{https://doi.org/10.1073/pnas.1220285110}{10.1073/pnas.1220285110}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Eddy2011}{}}%
+Eddy SR. 2011. Accelerated Profile HMM Searches. \emph{PLOS
+Computational Biology} \textbf{7}:1--16.
+doi:\href{https://doi.org/10.1371/journal.pcbi.1002195}{10.1371/journal.pcbi.1002195}
+
+\leavevmode\vadjust pre{\hypertarget{ref-frickey2004}{}}%
+Frickey T, Lupas A. 2004. CLANS: a Java application for visualizing
+protein families based on pairwise similarity. \emph{Bioinformatics}
+\textbf{20}:3702--3704.
+doi:\href{https://doi.org/10.1093/bioinformatics/bth444}{10.1093/bioinformatics/bth444}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Fu2012}{}}%
@@
 doi:\href{https://doi.org/10.1016/s0896-6273(00)80436-5}{10.1016/s0896-6273(00)80436-5}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Guxfchmann2015}{}}%
+Gühmann M, Jia H, Randel N, Verasztó C, Bezares-Calderón Luis~A,
+Michiels Nico~K, Yokoyama S, Jékely G. 2015. Spectral Tuning of
+Phototaxis by a Go-Opsin in the Rhabdomeric Eyes of Platynereis.
+\emph{Current Biology} \textbf{25}:2265--2271.
+doi:\href{https://doi.org/10.1016/j.cub.2015.07.017}{10.1016/j.cub.2015.07.017}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Guindon2010}{}}%
+Guindon S, Dufayard J-F, Lefort V, Anisimova M, Hordijk W, Gascuel O.
+2010. New Algorithms and Methods to Estimate Maximum-Likelihood
+Phylogenies: Assessing the Performance of PhyML 3.0. \emph{Systematic
+Biology} \textbf{59}:307--321.
+doi:\href{https://doi.org/10.1093/sysbio/syq010}{10.1093/sysbio/syq010}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Hoang2018}{}}%
+Hoang DT, Chernomor O, Haeseler A von, Minh BQ, Vinh LS. 2018. UFBoot2:
+Improving the Ultrafast Bootstrap Approximation. \emph{Molecular Biology
+and Evolution} \textbf{35}:518--522.
+doi:\href{https://doi.org/10.1093/molbev/msx281}{10.1093/molbev/msx281}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Katoh2008}{}}%
+Katoh K, Toh H. 2008. Recent developments in the MAFFT multiple sequence
+alignment program. \emph{Briefings in Bioinformatics}
+\textbf{9}:286--298.
+doi:\href{https://doi.org/10.1093/bib/bbn013}{10.1093/bib/bbn013}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Kuehn2022}{}}%
@@
 doi:\href{https://doi.org/10.1021/acssensors.6b00582}{10.1021/acssensors.6b00582}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Minh2020}{}}%
-Minh BQ, Schmidt HA, Chernomor O, Schrempf D, Woodhams MD, Haeseler A
-von, Lanfear R. 2020. IQ-TREE 2: New Models and Efficient Methods for
-Phylogenetic Inference in the Genomic Era. \emph{Molecular Biology and
-Evolution} \textbf{37}:1530--1534.
-doi:\href{https://doi.org/10.1093/molbev/msaa015}{10.1093/molbev/msaa015}
 
@@
 \begin{figure}
-\includegraphics[width=27.78in]{figures/Fig1_sup1} \caption{**Figure 1 -- figure supplement 1**  Phylogenetic tree of NOS by maximum likelihood.
-Tree robustness was tested with 1000 replicates of ultrafast bootstrap with the aLRT-SH-like and aBayes methods.}\label{fig:unnamed-chunk-8}
+\includegraphics[width=27.78in]{figures/Fig1_sup1} \caption{**Figure 1 -- figure supplement 1.
+Maximum-likelihood phylogenetic tree of NOS protein sequences.** Individual branches are coloured by taxonomy.
+Branch support values indicate UFBoot and aLRT-SH-like values.
+<br> Figure 1 -- figure supplement 1 -- source data 1.
+NOS sequences used for the phylogenetic reconstruction.
+<br> Figure 1 -- figure supplement 1 -- source data 2.
+Aligned and trimmed NOS sequences used for the phylogenetic reconstruction.
+<br> Figure 1 -- figure supplement 1 -- source data 3.
+Tree file of the reconstructed NOS phylogeny.
+}\label{fig:unnamed-chunk-9}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=16.67in]{figures/Fig1_sup2} \caption{**Figure 1 -- figure supplement 2** Expression analysis of the *NOS* gene (magenta) using in situ HCR using the larva at three-day-old in dorsal **(A)** and anterior view (B).
-(B) Insert: showing two NOS cells (INNOS_dl and INNOS_vl) on left side.}\label{fig:unnamed-chunk-9}
+\includegraphics[width=16.67in]{figures/Fig1_sup2} \caption{**Figure 1 -- figure supplement 2.
+Expression of *NOS* in three-day-old *Platynereis* larva.** (**A**) HCR in situ for *NOS* (magenta), dorsal view.
+(**B**) HCR in situ for *NOS* (magenta), ventral view.
+Inset in (B) shows close-up of two *NOS*-positive cells (INNOS_dl and INNOS_vl) on the left side.
+Samples were also stained for DAPI (cyan) to label nuclei.}\label{fig:unnamed-chunk-10}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=43.06in]{figures/Fig1_sup3} \caption{**Figure 1 -- figure supplement 3** ssTEM reconstruction of the INNOS (A), INRGWa (B) and cPRC (C), anterior view.
-NS plexus; neurosecretory plexus}\label{fig:unnamed-chunk-10}
+\includegraphics[width=43.06in]{figures/Fig1_sup3} \caption{**Figure 1 -- figure supplement 3.
+Synapse distribution in cPRCs and possynaptic interneurons** (**A**) Volume rendering of the four INNOS neurons with incoming and outgoing synapses.
+(**B**) Axo-dendritic splitting of the INNOS neurons.
+(**C**) Volume rendering of the four INRGW neurons with incoming and outgoing synapses.
+(**D**) Volume rendering of the four cPRC neurons with incoming and outgoing synapses.
+All images are anterior views.
+Asterisk marks the same position across images for reference.
+Grey speheres indicate the position of cell nuclei.
+NS plexus; neurosecretory plexus.
+The radius of INNOS nuclei is set to 2 µm for scale.}\label{fig:unnamed-chunk-11}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=44.44in]{figures/Fig3_sup1} \caption{**Figure 3 -- figure supplement 1** **(A)** Top: The domain and Exon/Intron structure of NOS.
-Bottom: Close-up region showing the genomic locus of NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9.
-The generated mutants (NOSΔ11, NOSΔ23) are also shown.
-Pink indicates target sites.
-Gray shows PAM sequences, red shows stop codons.
-**(B)** Overlaid trajectories for WT (n=37) and NOS mutant (NOSΔ11/Δ11, n=18 and NOSΔ23/Δ23, n=8) at two-day-old larvae.
-0 sec as the starting point.
-After 10 sec, UV (395 nm) stimulation from the side.
-**(C)** The temporal changes in the vertical position of the WT and mutant two-days-old larvae before and after UV stimulation are shown.
-The starting points of each larval trajectory are set to 0.
-After UV stimulation is indicated by purple squares.
-**(D, E)** The temporal changes in the distance traveled of the WT and mutant in two-day (D) and three-day-old (E) larvae before and after UV stimulation are shown.
-**(F)** The temporal changes in the distance traveled of larvae treated with NOS inhibitors, L-NAME in three-day larvae before and after UV stimulation are shown.
-**(G)** Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23) larvae at two-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top.
-The data are shown in 30 s bins.}\label{fig:unnamed-chunk-11}
+\includegraphics[width=44.44in]{figures/Fig3_sup1} \caption{**Figure 3 -- figure supplement 1.
+Generation and behavioural characterisation of *NOS* CRISPR knockouts.** (**A**) Top: The domain organisation of *Platynereis* NOS protein and exon/intron structure of the *NOS* gene.
+Bottom: The genomic locus of *NOS* with the CRISPR target site, wild-type and knockout (NOSΔ11, NOSΔ23) sequences and predicted protein sequences.
+The PAM sequence is shown in grey, stop codons in red.
+(**B**) Swimming trajectories of wild type (WT, n=37) and *NOS* mutant (NOSΔ11/Δ11, n=18 and NOSΔ23/Δ23, n=8) two-day-old larvae.
+All trajectories start at 0 x and y position and time 0 corresponding to 10 sec after the onset of 395 nm stimulation from the side.
+(**C**)  Vertical displacement in 30 sec bins of wild type and mutant (NOSΔ11 and NOSΔ23) two-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
+(**D**) Vertical position of batches of wild type and mutant two-day-old larvae over time under 395 nm UV stimulation.
+The starting position of each larval trajectory was set to 0.
+(**E-G**) Swimming speed of batches of wild type and mutant two-day old (E) and three-day-old (F) larvae and L-NAME-treated (NOS inhibitor) three-day-old larvae under 395 nm UV stimulation.
+<br> Figure 3 -- figure suppmenet 1 -- source data 1-6.
+Source data for panels B-G.}\label{fig:unnamed-chunk-12}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=33.33in]{figures/Fig4_sup1} \caption{**Figure 4 -- figure supplement 1** Cluster analysis of guanylate and adenylate cyclases.
-Connections are based on blast similarities < 1e-16 as shown on the upper right.
-Animal groups are colour coded as shown on the upper left.
-NIT-GCs, NIT domain containing guanylate cyclases; membrane-bound GCs, Membrane-bound guanylyl cyclases; sGCs, soluble guanylate cyclases; ACs, adenylate cyclases.}\label{fig:unnamed-chunk-12}
+\includegraphics[width=33.33in]{figures/Fig4_sup1} \caption{**Figure 4 -- figure supplement 1.
+Cluster analysis of guanylate and adenylate cyclase sequences.** Each node represents one sequence, colour-coded by taxonomy.
+Connections represent BLAST P-values of <1e-16.
+NIT-GCs, NIT domain containing guanylate cyclases; membrane-bound GCs, membrane-bound guanylate cyclases; sGCs, soluble guanylate cyclases; ACs, adenylate cyclases.}\label{fig:unnamed-chunk-13}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=20.83in]{figures/Fig4_sup2} \caption{**Figure 4 -- figure supplement 2** Phylogenetic tree of guanylate cyclase by maximum likelihood (ML).
-Guanylate cyclase-coupled receptor and soluble guanylate cyclases (sGC) as outgroups.
+\includegraphics[width=20.83in]{figures/Fig4_sup2} \caption{**Figure 4 -- figure supplement 2.
+Maximum-likelihood phylogenetic tree of NIT-domain-containing guanylate cyclases** Membrane-bound and soluble guanylate cyclases (sGC) were included as outgroups.
 Guanylate cyclases with NIT domains are found in most animal phyla except Porifera, Ctenophora, Urochordata and Chordata.
-Dot plot of Platynereis NIT-GC genes (columns) expressed in cPRC, INNOS and INRGWa (rows) using single cell RNA-Seq.
-The size of the dots is expressed in proportion to the percentage of cells expressing that gene relative to all cells.
-The colours represent the normal logarithm of the number of transcripts in the cells expressing the gene.}\label{fig:unnamed-chunk-13}
+Branch support values indicate UFBoot and aLRT-SH-like values.
+The expression of *Platynereis* NIT-GC genes in cPRC, INNOS and INRGW cells is indicated on the right side of the tree.
+The values represent expression based on single-cell sequencing data.
+Dot size indicates specificity (percent of transcripts expressed in the indicated cell across all cells).
+Dot colour represents the logarithm of the number of transcripts in the expressing cells.
+<br> Figure 4 -- figure supplement 2 -- source data 1.
+GC sequences used for the phylogenetic reconstruction.
+<br> Figure 4 -- figure supplement 2 -- source data 2.
+Aligned and trimmed GC sequences used for the phylogenetic reconstruction.
+<br> Figure 4 -- figure supplement 3 -- source data 3.
+Tree file of the reconstructed GC phylogeny.
+}\label{fig:unnamed-chunk-14}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=33.33in]{figures/Fig4_sup3} \caption{**Figure 4 -- figure supplement 3** **(A)** Co-expression analysis image of the NIT-GC1 (magenta) and MLD-pedal2 amide proneuropeptide gene (MLD: green).
-Anterior view of the larva at two-day-old.
-**(B, C)** Expression analysis of the NIT-GC2 gene (magenta) using in situ HCR.
-Anterior (B) and posterior (C) views of the larva at three-day-old.
-**(D)** Co-localisation analysis using NIT-GC1 (magenta) and NOS (green) antibodies.
-Anterior view of the larva at two-day-old.
-**(E, F)** Localisation analysis using NIT-GC1 and NIT-GC2 antibodies for NIT-GC1 (E) and NIT-GC2 (F) morphant.
-Green shows co-staining with acetylated α-tubulin antibody (acTub).
-Anterior view of the larva at two-day-old.}\label{fig:unnamed-chunk-14}
+\includegraphics[width=33.33in]{figures/Fig4_sup3} \caption{**Figure 4 -- figure supplement 3.
+Expression of NIT-GC1 and NIT-GC2 in the cPRCs** (**A**) HCR co-expression analysis of *NIT-GC1* (magenta) and *MLD/pedal-peptide-2 proneuropeptide* (green).
+Anterior view of a two-day-old larva.
+Nuclei were stained with DAPI (cyan).
+**(B, C)** Expression of *NIT-GC2* (magenta) detected by in situ HCR.
+Anterior (B) and ventral (C) views of a three-day-old larva.
+Nuclei were stained with DAPI (cyan).
+(**D**) Immunostaining for NIT-GC1 (magenta) and NOS (green).
+Anterior view of a two-day-old larva.
+Acetylated α-tubulin staining (blue) highlights the neuropil and cPRC cilia.
+**(E, F)** Immunostaining for NIT-GC1 (E) and NIT-GC2 (F) antibodies in larvae injected with NIT-GC1 (E) or NIT-GC2 (F) morpholinos.
+Acetylated α-tubulin staining (green) highlights the neuropil and cPRC cilia.
+Anterior view of a two-day-old larva.}\label{fig:unnamed-chunk-15}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=27.78in]{figures/Fig5_sup1} \caption{**Figure 5 -- figure supplement 1** **(A)** Schematic diagram of the immunostaining procedure after calcium imaging.**(B-D)** Co-expression analysis image of the NOS (magenta) and RY amide proneuropeptide gene (RYa: green).
-Anterior view of the larva at two-day-old.
-}\label{fig:unnamed-chunk-15}
+\includegraphics[width=27.78in]{figures/Fig5_sup1} \caption{**Figure 5 -- figure supplement 1.
+On-slide immunostaining and coexpression analysis of *NOS* and *RYamide proneuropeptide* ** (**A**) Schematic diagram of the on-slide immunostaining procedure after Ca^2+^ imaging.(**B-D**) Co-expression analysis by HCR in situ of *NOS* (magenta) and *RYamide proneuropeptide* (green).
+Anterior view of a two-day-old larva.
+Nuceli were stained by DAPI (cyan).
+}\label{fig:unnamed-chunk-16}
 \end{figure}
 
+\begin{figure}
+\centering
+\includesvg{figures/Figure6_fig_Suppl1.svg}
+\caption{\textbf{Figure 6 -- figure supplement 1. Simulated
+Ca\textsuperscript{2+} traces for parameter sets fitted to individual
+Ca\textsuperscript{2+}-recordings collected in wild type, \emph{NOS}
+knockout, and \emph{NIT-GC2} morphant larvae.} (\textbf{A-C}) Simulated
+Ca\textsuperscript{2+} traces in cPRC (A), INNOS (B) and INRGW (C) cells
+in the wild-type condition. (\textbf{D-F}) Simulated
+Ca\textsuperscript{2+} traces in cPRC (D), INNOS (E) and INRGW (F) cells
+in the \emph{NOS}-knockout condition. (\textbf{G-I}) Simulated
+Ca\textsuperscript{2+} traces in cPRC (G), INNOS (H) and INRGW (I) cells
+in the \emph{NIT-GC2}-morphant condition. Thin coloured curves indicate
+individual recordings (cPRC - purple, INNOS - blue and INRGW - green),
+thin grey curves indicate simulated Ca\textsuperscript{2+} traces based
+on model parameters fitted to the individual recording, thick coloured
+curves indicate averages of the recordings and thick black curves
+represent the average of the fits.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includesvg{figures/Figure6_fig_Suppl2.svg}
+\caption{\textbf{Figure 6 -- figure supplement 2. Comparison of fits
+with different fixed parameter sets when fitting the WT model to
+individual recordings from the WT INRGW cells.} Panels A-B, D-E, G-H
+(column fixed parameters) show simulated traces of the
+Ca\textsuperscript{2+} levels in cPRC and INNOS cells produced by
+different fixed parameter sets. Panels C, F and I show individual
+recordings fitted to the data from INRGW cells.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includesvg{figures/Figure6_fig_Suppl3.svg}
+\caption{\textbf{Figure 6 -- figure supplement 3. Comparison of fits to
+the cPRC recordings from NIT-GC2 morphant larvae using the NIT-GC2
+morpholino model or the NOS-knockout model.} Thin coloured curves
+indicate individual recordings, thin grey curves are simulated
+Ca\textsuperscript{2+} traces, thick coloured curves indicate averages
+of the recordings and thick black curves represent the mean of the
+fits.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includesvg{figures/Figure6_fig_Suppl4.svg}
+\caption{\textbf{Figure 6 -- figure supplement 4. Distributions of the
+parameter values fitted to the cPRC recordings.} Violin plots (grey
+lines) are kernel density estimates of the underlying distributions;
+computed using the Matlab function \emph{ksdensity} with default
+settings, i.e., using normal kernel function, plots are trimmed to the
+observed range of data. Markers represent individual parameter values,
+white markers indicate medians, vertical grey bars indicate
+interquartile range (Q25 to Q75), and the dashed line indicates the
+value 1. Where violin plots are not shown, these parameters are
+associated with terms/equations that are excluded from the fitting
+procedure (see Methods for details).}
+\end{figure}
+
+\begin{figure}
+\centering
+\includesvg{figures/Figure6_fig_Suppl5.svg}
+\caption{\textbf{Figure 6 -- figure supplement 5. Pairwise correlations
+between parameters of the WT model fitted to WT cPRC recordings}
+Parameters associated with the \(C_R\) equation are not included (see
+Methods for details). Panels on the diagonal show kernel density
+estimates of the distributions of the individual parameters; computed
+using Matlab function \emph{ksdensity} with default settings, i.e.,
+using normal kernel function, plots are trimmed to the observed range of
+data. The other panels show kernel density estimates of bivariate
+distributions of pairs of parameters (heatmaps; darker colours indicate
+higher density). Also shown using markers are the values of the
+individual parameter combinations.}
+\end{figure}
+
 \end{document}

### 2023-12-18 Manuscript/Manuscript.Rmd
Update Figure 1~4

@@
 author: 
-  - name: "Kei Jokura<sup>1,`*`</sup>, Nobuo Ueda<sup>2</sup>, Martin Gühmann<sup>3</sup>, Luis Alfonso Yañez-Guerra<sup>1</sup>, Piotr Słowiński<sup>1</sup>, Kyle C.
+  - name: "Kei Jokura<sup>1,`*`</sup>, Nobuo Ueda<sup>2</sup>, Martin Gühmann<sup>3</sup>, Luis Alfonso Yañez-Guerra<sup>1,4</sup>, Piotr Słowiński<sup>1</sup>, Kyle C.
 A.
-Wedgwood<sup>1</sup> and Gáspár Jékely<sup>1,4,`*`</sup>"
-    affiliation: "<sup>1</sup>Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4 4QD, United Kingdom <sup>2</sup>Okinawa Institute of Science and Technology, Okinawa, Japan, <sup>3</sup>School of Biological Sciences, University of Bristol, Bristol, United Kingdom, <sup>4</sup>Centre for Organismal Studies (COS), University of Heidelberg, 69120 Heidelberg, Germany  <br> <sup>*</sup>Correspondence`:` k.jokura @ exeter.ac.uk, gaspar.jekely @ cos.uni-heidelberg.de"
+Wedgwood<sup>1</sup> and Gáspár Jékely<sup>1,5,`*`</sup>"
+    affiliation: "<sup>1</sup>Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4 4QD, United Kingdom <sup>2</sup>Okinawa Institute of Science and Technology, Okinawa, Japan, <sup>3</sup>School of Biological Sciences, University of Bristol, Bristol, United Kingdom, <sup>4</sup>School of Biological sciences, University of Southampton, Southampton, United Kingdom, <sup>5</sup>Centre for Organismal Studies (COS), University of Heidelberg, 69120 Heidelberg, Germany  <br> <sup>*</sup>Correspondence`:` k.jokura @ exeter.ac.uk, gaspar.jekely @ cos.uni-heidelberg.de"
 output:
@@
 To characterise the expression pattern of *NOS*, we used in situ hybridization chain reaction (HCR) and transient transgenesis.
-In two- and three-day-old larvae, we detected *NOS* expression in four cells (two of them weakly expressing) in the apical organ region (Figure 1D and Figure 1---figure supplement 2).
+In two- and three-day-old larvae, we detected *NOS* expression in four cells (two of them weakly expressing) in the apical organ region (Figure 1E and Figure 1---figure supplement 2).
 *NOS* was also expressed in the region of the visual eyes (adult eyes) and the pigmented eyespots (Figure 1---figure supplement 2).
-The four apical organ cells, but not the eyes, were also labelled with a *NOS* reporter construct driving palmitoylated tdTomato (Figure 1E).
+The four apical organ cells, but not the eyes, were also labelled with a *NOS* reporter construct driving palmitoylated tdTomato (Figure 1F).
 This reporter also revealed the axonal projections of these central NOS-expressing neurons.
 The position and morphology of the four *NOS+* cells allowed us to identify the same four cells as four interneurons (INNOS) in our three-day-old whole-body *Platynereis* volume EM data [@williams2017; @veraszto2020] (Figure 1B,C).
-In the synaptic connectome, the INNOS cells are postsynaptic to the UV-sensory cPRCs and presynaptic to the INRGWa interneurons, which are also cPRC targets (Figure 1C, F).
-The projections of INNOS cells are segregated into input (dendritic) and output (axonal) compartments, with cPRC inputs occurring on the dendritic part in the dense neurosecretory plexus.
+In the synaptic connectome, the INNOS cells are postsynaptic to the UV-sensory cPRCs and presynaptic to the INRGWa interneurons, which are also cPRC targets (Figure 1C, H).
+The projections of INNOS cells are segregated into input (dendritic) and output (axonal) compartments, with cPRC inputs occurring on the dendritic part in the dense neurosecretory plexus (Figure 1D).
+Immunostaining revealed that the localization of NOS at the protein level is concentrated in the dendrites of the input region even within INNOS (Figure 1G).
 INNOS output-synapses form in the more ventral projection region (Figure 1---figure supplement 3).
-INRGWa neurons synapse on the head serotonergic ciliomotor neurons (Ser-h1), which synapse on the cholinergic MC ciliomotor neuron and cells of the prototroch ciliary band (Figure 1C, F) [@veraszto2017].
+INRGWa neurons synapse on the head serotonergic ciliomotor neurons (Ser-h1), which synapse on the cholinergic MC ciliomotor neuron and cells of the prototroch ciliary band (Figure 1C, H) [@veraszto2017].
 
@@
 In B, nuclei positions of the prototroch head ciliary band are shown as grey spheres.
-**(D)** Expression of the *NOS* gene detected by in situ HCR (magenta) in a two-day-old larva (anterior view).
+**(D)** Volume rendering of the four INNOS neurons with incoming and outgoing synapses.
+**(E)** Expression of the *NOS* gene detected by in situ HCR (magenta) in a two-day-old larva (anterior view).
 Antibody staining for acetylated α-tubulin (acTub: green) highlights cPRC cilia and the neuropil.
-**(E)** Expression of a *NOS* reporter (NOSp::palmi-3xHA; magenta) labelled with an anti-HA antibody in a two-day-old larva (anterior view).
+**(F)** Expression of a *NOS* reporter (NOSp::palmi-3xHA; magenta) labelled with an anti-HA antibody in a two-day-old larva (anterior view).
 Antibody staining for acetylated α-tubulin (acTub; green) highlights cPRC cilia and the neuropil.
-**(F)** Synaptic wiring diagram of the cPRC circuit.
+**(G)** Immunostaining for NOS (magenta), co-stained for acetylated α-tubulin (acTub; green).
+**(H)** Synaptic wiring diagram of the cPRC circuit.
 Hexagons represent cell groups, with the number of cells per group shown in square brackets.
@@
 To test this, first we asked whether NO is produced during UV/violet stimulation of the larvae.
-We injected the fluorescent NO-reporter DAF-FM into zygotes and imaged two-day-old larvae while exposing the region of cPRC cilia to 405 nm violet light.
+We injected the fluorescent NO-reporter DAF-FM into zygotes and imaged two-day-old larvae while exposing the region of cPRC cilia to 405 nm violet light (Figure 2---figure supplement 1).
 To mark cell outlines, we coinjected mRNA encoding a red-fluorescent reporter (RGECO), allowing the identification of the cPRCs.
@@
 We could establish a homozygous line for both mutations indicating that NOS is not an essential gene in *Platynereis*.
-To quantify UV avoidance, we recorded the trajectories of freely swimming wild type and mutant larvae in vertical columns, illuminated laterally from two opposite sides with 395 nm UV light (Figure 3A and Figure 3---figure supplement 1B).
+To quantify UV avoidance, we recorded the trajectories of freely swimming wild type and mutant larvae in vertical columns, illuminated laterally from two opposite sides with 395 nm UV light (Figure 3A and Figure 3---figure supplement 1B, C).
 As previously shown, wild-type larvae swim downward following non-directional UV/violet light stimulation [@Veraszto2018].
-In contrast, both two- and three-day-old homozygous *NOS*-mutant larvae showed a strongly diminished UV-avoidance response (Figure 3A, B and Figure 3---figure supplement 1B, C).
+In contrast, both two- and three-day-old homozygous *NOS*-mutant larvae showed a significantly diminished UV-avoidance response (Figure 3B, C and Figure 3---figure supplement 1D, E).
 This phenotype is similar to the defective UV-avoidance of *c-opsin1* mutant larvae [@Veraszto2018] and reveals a requirement for NOS in UV-avoidance behaviour.
-Wild type but not mutant larvae also showed an increase in swimming speed under UV light that may be due to downward swimming trajectories (swimming in the direction of gravity) (Figure 3B and Figure 3---figure supplement 1C).
+Wild type but not mutant larvae also showed a significantly increase in swimming speed with an increase in ciliary beat frequency under UV light that may be due to downward swimming trajectories (swimming in the direction of gravity) (Figure 3---figure supplement 2).
 We also tested directional phototaxis, by exposing larvae to 480 nm directional collimated light from the top of the column.
-Three-day-old but not two-day-old *NOS*-mutant larvae also showed reduced phototactic behaviour, suggesting a function for *NOS* in the visual eyes that mediate three-day-old phototaxis [@Randel2014a] (Figure 3D and Figure 3---figure supplement 1G).
+Three-day-old but not two-day-old *NOS*-mutant larvae also showed reduced phototactic behaviour, suggesting a function for *NOS* in the visual eyes that mediate three-day-old phototaxis [@Randel2014a] (Figure 3D and Figure 3---figure supplement 1F).
 
 To distinguish between an acute and developmental function of NOS in light responses, we also tested larvae exposed to the NOS inhibitor L-NAME.
-Larvae incubated for 5 min in 0.1 mM or 1 mM L-NAME showed a dose-dependent inhibition of UV avoidance.
-In contrast, phototaxis was not affected (Figure 3C, E).
+Larvae incubated for 5 min in 0.1 mM or 1 mM L-NAME showed a dose-dependent inhibition of UV avoidance (Figure 3E, F).
+In contrast, phototaxis was not affected (Figure 3G).
 Overall, our results indicate an acute requirement for NOS signalling in UV-avoidance and a possible indirect, developmental role in the visual system, reminiscent of the function of NO signalling in *Drosophila* eye development [@Gibbs1998a].
@@
 NOS is required for UV avoidance in *Platynereis* larvae.**  (**A**) Swimming trajectories of wild type (WT, n=32) and *NOS* mutant (*NOSΔ11/Δ11*, n=26 and *NOSΔ23/Δ23*, n=47) three-day-old larvae.
-All trajectories start at 0 x and y position and time 0 corresponding to 10 sec after the onset of 395 nm stimulation from the side.
+All trajectories start at 0 x and y position and time 0 corresponding to 10 sec before the onset of 395 nm stimulation from the side.
 (**B**) Vertical position of batches of wild type and mutant three-day-old larvae over time under 395 nm UV stimulation.
 The starting position of each larval trajectory was set to 0.
-(**C**) Vertical position of batches of control and L-NAME-treated (0.1 and 1 mM) three-day-old larvae over time under 395 nm UV stimulation.
+(**C**) Comparison of vertical position of wild type and mutant three-day-old larvae after 30 seconds of UV stimulation.
+One-way ANOVA with Dunnett’s multiple-comparison test are shown.
+The wild type (WT, n=32) and *NOS* mutant (*NOSΔ11/Δ11*, n=26 and *NOSΔ23/Δ23*, n=47) larvae.
+(**D**) Vertical displacement in 30 sec bins of wild type (10 batches) and mutant (*NOSΔ11/Δ11*, 5 batches and *NOSΔ23/Δ23*, 7 batches) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
+One-way ANOVA with Dunnett’s multiple-comparison test are shown.
+*P < 0.05.
+(**E**) Vertical position of batches of control and L-NAME-treated (0.1 and 1 mM) three-day-old larvae over time under 395 nm UV stimulation.
 The starting position of each larval trajectory was set to 0.
-(**D**) Vertical displacement in 30 sec bins of wild type and mutant (*NOSΔ11/Δ11* and *NOSΔ23/Δ23*) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
-(**E**) Vertical displacement in 30 sec bins of control and L-NAME-treated (0.1 and 1 mM) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
+(**F**) Comparison of vertical position of control (n=6) and L-NAME-treated (0.1 and 1 mM, n=6 each) three-day-old larvae after 30 seconds of UV stimulation.
+One-way ANOVA with Dunnett’s multiple-comparison test are shown.
+(**G**) Vertical displacement in 30 sec bins of control (15 batches) and L-NAME-treated (0.1 and 1 mM, 15 batches each) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
+One-way ANOVA with Dunnett’s multiple-comparison test are shown.
+*P < 0.05.
 <br> Figure 3 -- source data 1-5.
@@
 The imaging laser had a similar intensity than the stimulus laser but covered an area that was 10 times larger than the stimulus ROI.
+
+**Ciliary beat frequency analysis**
+
+Larvae were imaged at room temperature on a Zeiss AxioObserver Z1 Microscope (Carl Zeiss Microscopy, Germany) with a C-Apochromat 10X lens, 850 nm power LED (OSRAM OSLON® Black, SFH 4716S) and filter sets equipped with a 405/40 ET excitation Bandpass, Beamsplitter dichroic mirror 495LPXR and without an emission filter with a IR camera (UI-3360CP-NIR-GL Rev.2, IDS Imaging Development Systems GmbH, Germany) frame rate of 200 frame/sec with 0.2 msec pulse controlled by the LED stroboscopic illumination system synchronized with the exposure signals from the camera (TYM-002, developed by Dr.Shoji A.
+Baba, Ochanomizu University, Otsuka, Japan) and an image size of 640 x 480 pixels.
+The larvae were stimulated with a 395 nm LED (M395L5, Thorlabs).
+Kymographs of the cilia obtained with a Fiji were used to measure the ciliary beat frequency.
 
@@
 
+We thank Dr.
+Cameron Hird, Adam Johnstone, Rebecca Turner and Jamie Haddon for animal husbandry, Dr.
+Corin Liddle for advice on calcium imaging, and Dr.
+Emelie Brodrick and Dr.
+Elizabeth Williams for critical reading of the manuscript.
 This work was funded by the Wellcome Trust (214337/Z/18/Z).
@@
 ```{r echo=FALSE,fig.cap="**Figure 1 -- figure supplement 3.
-Synapse distribution in cPRCs and possynaptic interneurons** (**A**) Volume rendering of the four INNOS neurons with incoming and outgoing synapses.
-(**B**) Axo-dendritic splitting of the INNOS neurons.
-(**C**) Volume rendering of the four INRGW neurons with incoming and outgoing synapses.
-(**D**) Volume rendering of the four cPRC neurons with incoming and outgoing synapses.
+Synapse distribution in cPRCs and possynaptic interneurons** (**A**) Axo-dendritic splitting of the INNOS neurons.
+(**B**) Volume rendering of the four INRGW neurons with incoming and outgoing synapses.
+(**C**) Volume rendering of the four cPRC neurons with incoming and outgoing synapses.
 All images are anterior views.
@@
 
+```{r echo=FALSE,fig.cap="**Figure 2 -- figure supplement 1.
+Schematic diagrams of the UV stimulation experiment.** (**A**) Left: The larvae were embedded in 2% agarose seawater with the anterior side down.
+Right: Changes in DAF-FM fluorescence intensities were examined when the ciliary region of the cPRC was stimulated with a 405 nm laser."}
+knitr::include_graphics(path = "figures/Fig2_sup1.png")
+
+```
+
 ```{r echo=FALSE,fig.cap="**Figure 3 -- figure supplement 1.
@@
 The PAM sequence is shown in grey, stop codons in red.
-(**B**) Swimming trajectories of wild type (WT, n=37) and *NOS* mutant (NOSΔ11/Δ11, n=18 and NOSΔ23/Δ23, n=8) two-day-old larvae.
-All trajectories start at 0 x and y position and time 0 corresponding to 10 sec after the onset of 395 nm stimulation from the side.
-(**C**)  Vertical displacement in 30 sec bins of wild type and mutant (NOSΔ11 and NOSΔ23) two-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
-(**D**) Vertical position of batches of wild type and mutant two-day-old larvae over time under 395 nm UV stimulation.
+(**B**) Schematic diagram of the set-up of the behavioural experiment.
+(**C**) Swimming trajectories of wild type (WT, n=37) and *NOS* mutant (NOSΔ11/Δ11, n=18 and NOSΔ23/Δ23, n=8) two-day-old larvae.
+All trajectories start at 0 x and y position and time 0 corresponding to 10 sec before the onset of 395 nm stimulation from the side.
+(**D**)  Vertical displacement in 30 sec bins of wild type (WT, n=37) and mutant (NOSΔ11/Δ11, n=18 and NOSΔ23/Δ23, n=8) two-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
+(**E**) Comparison of vertical position of wild type and mutant two-day-old larvae after 30 seconds of UV stimulation.
+One-way ANOVA with Dunnett’s multiple-comparison test are shown.
+The wild type (WT, n=52) and *NOS* mutant (*NOSΔ11/Δ11*, n=30 and *NOSΔ23/Δ23*, n=16) larvae.
+(**F**) Vertical position of batches of wild type (12 batches) and mutant (NOSΔ11/Δ11, 8 batches and NOSΔ23/Δ23, 3 batches) two-day-old larvae over time under 395 nm UV stimulation.
 The starting position of each larval trajectory was set to 0.
-(**E-G**) Swimming speed of batches of wild type and mutant two-day old (E) and three-day-old (F) larvae and L-NAME-treated (NOS inhibitor) three-day-old larvae under 395 nm UV stimulation.
+One-way ANOVA with Dunnett’s multiple-comparison test are shown.
+*P < 0.05.
 <br> Figure 3 -- figure suppmenet 1 -- source data 1-6.
@@
 knitr::include_graphics(path = "figures/Fig3_sup1.png")
+
+```
+
+```{r echo=FALSE,fig.cap="**Figure 3 -- figure supplement 2.
+Generation and behavioural characterisation of *NOS* CRISPR knockouts.** (**A,B**) Swimming speed of batches of wild type and *NOS* mutant two-day-old (WT, n=37, NOSΔ11/Δ11, n=18 and NOSΔ23/Δ23, n=8) (A) and three-day-old (WT, n=32, *NOSΔ11/Δ11*, n=26 and *NOSΔ23/Δ23*, n=47) (B) larvae under 395 nm UV stimulation.
+(**C,D**) Mean swimming speed of wild type and *NOS* mutant two-day-old (WT, n=37, NOSΔ11/Δ11, n=18 and NOSΔ23/Δ23, n=8) (C) and three-day-old (WT, n=32, *NOSΔ11/Δ11*, n=26 and *NOSΔ23/Δ23*, n=47) (D) larvae during 10 seconds before (dark) and 10 seconds after the start of 395 nm UV stimulation (sideUV).
+Data points for the same larva are joined by lines.
+One-tailed paired t-test; p-values (‘***’ <0.001) are shown.
+(**E,F**) Average ciliary beat frequency of wild type and mutant two-day-old (E) and three-day-old (F) single larvae during 10 seconds before (gray) and 10 seconds after the start of 395 nm UV stimulation (purple).
+N=6 larvae each.
+Data points for the same larva are joined by lines.
+One-tailed paired t-test; p-values (‘**’ <0.01 or ‘*’ <0.05) are shown.
+<br> Figure 3 -- figure suppmenet 2 -- source data 1-6.
+Source data for panels B-G."}
+knitr::include_graphics(path = "figures/Fig3_sup2.png")
 
@@
 Nuclei were stained with DAPI (cyan).
-(**D**) Immunostaining for NIT-GC1 (magenta) and NOS (green).
+**(D, E)** Immunostaining for NIT-GC1 (D) and NIT-GC2 (E) antibodies in larvae injected with NIT-GC1 (D) or NIT-GC2 (E) morpholinos.
+Acetylated α-tubulin staining (green) highlights the neuropil and cPRC cilia.
 Anterior view of a two-day-old larva.
-Acetylated α-tubulin staining (blue) highlights the neuropil and cPRC cilia.
-**(E, F)** Immunostaining for NIT-GC1 (E) and NIT-GC2 (F) antibodies in larvae injected with NIT-GC1 (E) or NIT-GC2 (F) morpholinos.
-Acetylated α-tubulin staining (green) highlights the neuropil and cPRC cilia.
-Anterior view of a two-day-old larva."}
+(**F**) The fluorescence intensities of the NIT-GC1 or NIT-GC2 in the control and morphant (MO1 and MO2) larvae (n=6 each) were normalized to the Acetylated α-tubulin signals and plotted as a relative fluorescence.
+One-way ANOVA with Dunnett’s multiple-comparison test; p-values (‘***’ <0.001) are shown."}
 knitr::include_graphics(path = "figures/Fig4_sup3.png")

### 2023-12-18 code/Fig1.R

@@ -333,14 +333,14 @@ panel_NOS3d_trans <- ggdraw() + draw_image(readPNG("pictures/NOS-promotor_3d_acT
   draw_label("INNOS_vl", x = 0.85, y = 0.65, size = 7, color = "white")
 
 
-panel_NOS3d_IHC <- ggdraw() + draw_image(readPNG("pictures/IHC_NOS_3d_XXum_3.png")) +
-  draw_label("IHC", x = 0.3, y = 0.99, size = 10) +
+panel_NOS3d_IHC <- ggdraw() + draw_image(readPNG("pictures/IHC_NOS_3d_48.1um_3.png")) +
+  draw_label("IHC", x = 0.15, y = 0.99, size = 10) +
   draw_label("NOS", x = 0.12, y = 0.9, color="magenta", size = 11, fontface="plain") +
   draw_label("acTub", x = 0.36, y = 0.9, color="green", size = 11, fontface="plain") +
   draw_line(x = c(0.04, 0.4), y = c(0.08, 0.08), color = "white", size = 0.5) +
-  draw_label("cPRC", x = 0.55, y = 0.82, size = 7, color = "white") +
-  draw_line(x = c(0.33, 0.47), y = c(0.79, 0.82), color = "white", size = 0.3) +
-  draw_line(x = c(0.79, 0.63), y = c(0.79, 0.82), color = "white", size = 0.3)
+  draw_label("cPRC", x = 0.49, y = 0.83, size = 7, color = "white") +
+  draw_line(x = c(0.33, 0.42), y = c(0.8, 0.83), color = "white", size = 0.3) +
+  draw_line(x = c(0.67, 0.56), y = c(0.8, 0.83), color = "white", size = 0.3)
 
 
 INNOS_img <- readPNG("pictures/INNOS_synapses.png")
@@ -382,6 +382,6 @@ ggsave("Manuscript/figures/Fig1.png", limitsize = FALSE,
 
 
 ggsave("Manuscript/figures/Fig1.pdf", limitsize = FALSE, 
-       units = c("px"), Fig1, width = 1800, height = 1350) 
+       units = c("px"), Fig1, width = 2500, height = 1400) 
 
 

### 2023-12-18 code/Fig1_sup3.R

@@ -128,26 +128,6 @@ panel_cPRC <- ggdraw() +
   draw_label("cPRC", x = 0.4, y = 0.99, size = 10) +
   draw_label("*", x = 0.5, y = 0.29, color='black',size = 18,fontface='plain')
 
-layout <- "
-A#B#C#D
-"
-
-Fig1_sup3 <- panel_INNOS + panel_INNOS_split + panel_INRGW + panel_cPRC + 
-  patchwork::plot_layout(design = layout, 
-                         widths = c(1, 0.02,1,0.02,1,0.02,1)) + #we can change the heights of the rows in our layout (widths also can be defined)
-  patchwork::plot_annotation(tag_levels = "A") &  #we can change this to 'a' for small caps or 'i' or '1'
-  ggplot2::theme(plot.tag = element_text(size = 12, 
-                                         face='plain', color='black')) #or 'bold', 'italic'
-
-
-ggsave("Manuscript/figures/Fig1_sup3.png", limitsize = FALSE, 
-       units = c("px"), Fig1_sup3, width = 3100, height = 800, bg='white')  
-
-
-ggsave("Manuscript/figures/Fig1_sup3.pdf", limitsize = FALSE, 
-       units = c("px"), Fig1_sup3, width = 3100, height = 800) 
-
-
 
 
 layout <- "
@@ -167,4 +147,4 @@ ggsave("Manuscript/figures/Fig1_sup3.png", limitsize = FALSE,
 
 
 ggsave("Manuscript/figures/Fig1_sup3.pdf", limitsize = FALSE, 
-       units = c("px"), Fig1_sup3, width = 3100, height = 800) 
+       units = c("px"), Fig1_sup3, width = 2400, height = 800) 

### 2023-12-18 code/Fig2.R

@@ -71,25 +71,9 @@ ggsave("pictures/DAFFM.png", limitsize = FALSE,
 # assemble figure ---------------------------------------------------------
 arrow_fluo <- data.frame(x1 = 0.95, x2 = 0.95, y1 = 0.75, y2 = 0.85)
 
-panel_DAF <- ggdraw() + draw_image(readPNG("pictures/55hpf_DAF-FM_134.95um.png")) +
-  draw_label("DAF-FM", x = 0.15, y = 0.88, color="#009E73", size = 11, fontface="plain") +
-  draw_line(x = c(0.04, 0.42), y = c(0.09, 0.09), color = "white", size = 0.5) +
-  draw_label(expression(paste("50 ", mu, "m")), x = 0.23, y = 0.12, color = "white", size = 8) +
-  draw_label("D", x = 0.95, y = 0.88, size = 6, color = "white") +
-  draw_label("V", x = 0.95, y = 0.72, size = 6, color = "white") +
-  draw_label("cPRC stim.", x = 0.67, y = 0.43, size = 4, color = "darkorchid1", fontface='bold') +
-  draw_label("ctr stim.", x = 0.75, y = 0.38, size = 4, color = "darkorchid1", fontface='bold') +
-  geom_segment(aes(x = x1, y = y1, xend = x2, yend = y2), data = arrow_fluo, color = "white", 
-               arrow = arrow(ends = "both", type = "closed", length = unit(0.1,"cm")),
-               lineend = "butt",
-               linejoin = "mitre",
-               arrow.fill = "white", size = 0.2)
+panel_DAF <- ggdraw() + draw_image(readPNG("pictures/55hpf_DAF-FM_134.95um.png"))
 
-panel_DAFFM_intensity  <- ggdraw() + draw_image(readPNG("pictures/55hpf_DAF-FM_intensity.png")) +
-  draw_label("before", x = 0.17, y = 0.9, color='white',size = 9, fontface='plain') +
-  draw_label("405 nm", x = 0.2, y = 0.45, color='white',size = 9, fontface='plain') +
-  draw_line(x = c(0.05, 0.42), y = c(0.08, 0.08), color = "white", size = 0.5) +
-  draw_label(expression(paste("10 ", mu, "m")), x = 0.23, y = 0.11, color = "white", size = 8)
+panel_DAFFM_intensity  <- ggdraw() + draw_image(readPNG("pictures/55hpf_DAF-FM_intensity.png"))
 
[…]

### 2023-12-18 code/Fig3_and_Fig3_figure_Suppl1.R

@@ -32,7 +32,7 @@ df_WT_test <- read_csv("data/220923_Tracking raw data/WT_0701_08_01.csv")
 df_NOS_test <- read_csv("data/220923_Tracking raw data/NOS_0623_00_01.csv")
 
 #read L-NAME 3days tracking csv file
-df_L_NAME_tracking <- read_csv("data/230511_df_L_NAME_2.csv")
+df_L_NAME_tracking <- read_csv("data/230511_df_L_NAME.csv")
 
 }
 
@@ -1655,16 +1655,16 @@ df_2d_nor_stat <- df_2d_nor %>%
 
 df_2d_nor_stat <- df_2d_nor_stat %>%
   filter(!(Genotype %in% "NOS11x23")) %>%
-  filter(sec >= 39.9 & sec < 40)
+  filter(sec >= 39.86 & sec < 40)
 
 #Dunnett's test
 genotype_2d=factor(df_2d_nor_stat$Genotype)
 y_axis_2d=df_2d_nor_stat$y_axis
 summary(glht(aov(y_axis_2d~genotype_2d),linfct=mcp(genotype_2d= "Dunnett")))
 
-#                     Estimate Std. Error t value Pr(>|t|)
-#NOS11x11 - cont == 0    6.926      4.207   1.646    0.194
-#NOS23x23 - cont == 0    9.295      5.707   1.629    0.201
+#                     Estimate Std. Error t value Pr(>|t|)  
+#NOS11x11 - cont == 0    6.923      2.935   2.359   0.0389 *
+#NOS23x23 - cont == 0    9.282      3.982   2.331   0.0417 *
 
 #‘***’ <0.001 ‘**’ <0.01 ‘*’ <0.05
 
@@ -1674,7 +1674,7 @@ ggplot(df_2d_nor_stat) +
   geom_boxplot(size = 0.5, outlier.shape = NA) +
   geom_beeswarm(aes(color = Genotype),
                 size = 2,
-                cex = 3.2,
+                cex = 3,
                 alpha =.3, 
                 color = "gray5") +
   labs(y = "vertical displacement [mm/s]")+
@@ -1690,10 +1690,10 @@ ggplot(df_2d_nor_stat) +
                               "NOS11x11" = expression('NOS'^'Δ11/Δ11'), 
                               "NOS23x23" = expression('NOS'^'Δ23/Δ23'))) +
   geom_signif(comparisons = list(c("cont", "NOS11x11")),
-              annotations = "0.194",
[…]

### 2023-12-18 code/Fig4_sup3.R

@@ -150,7 +150,8 @@ panel_NIT1MO_NIT1 <- ggdraw() + draw_image(readPNG("pictures/IHC_51_AP_NIT1MO_NI
   draw_label("morpholino", x = 0.57, y = 0.995, size = 10) +
   draw_label("NIT-GC1", x = 0.12, y = 0.91, color="magenta", size = 10, fontface="plain") +
   draw_label("acTub", x = 0.33, y = 0.91, color="green", size = 10, fontface="plain") +
-  draw_line(x = c(0.04, 0.392), y = c(0.08, 0.08), color = "white", size = 0.5)
+  draw_line(x = c(0.04, 0.392), y = c(0.08, 0.08), color = "white", size = 0.5) +
+  draw_label(expression(paste("20 ", mu, "m")), x = 0.21, y = 0.11, color = "white", size = 8)
 
 panel_NIT2MO_NIT2 <- ggdraw() + draw_image(readPNG("pictures/IHC_57_AP_NIT2MO_NITGC2_actub_57um.png")) +
   draw_label("IHC:", x = 0.15, y = 0.995, size = 10) +

### 2023-12-19 Manuscript/Manuscript.Rmd
Update_response

@@
 The position and morphology of the four *NOS+* cells allowed us to identify the same four cells as four interneurons (INNOS) in our three-day-old whole-body *Platynereis* volume EM data [@williams2017; @veraszto2020] (Figure 1B,C).
-In the synaptic connectome, the INNOS cells are postsynaptic to the UV-sensory cPRCs and presynaptic to the INRGWa interneurons, which are also cPRC targets (Figure 1C, H).
+In the synaptic connectome, the INNOS cells are postsynaptic to the UV-sensory cPRCs and presynaptic to the INRGWa interneurons (cholinergic neurons that strongly express RGW neuropeptide), which are also cPRC targets (Figure 1C, H).
 The projections of INNOS cells are segregated into input (dendritic) and output (axonal) compartments, with cPRC inputs occurring on the dendritic part in the dense neurosecretory plexus (Figure 1D).
@@
 Inset in (B) shows close-up of two *NOS*-positive cells (INNOS_dl and INNOS_vl) on the left side.
-Samples were also stained for DAPI (cyan) to label nuclei."}
+Samples were also stained for DAPI (cyan) to label nuclei.(**C**) Expression of the *NOS* gene detected by in situ HCR (magenta) in a three-day-old larva (anterior view).
+Antibody staining for acetylated α-tubulin (acTub: green) highlights cPRC cilia and the neuropil."}
 knitr::include_graphics(path = "figures/Fig1_sup2.png")
@@
 The PAM sequence is shown in grey, stop codons in red.
-(**B**) Schematic diagram of the set-up of the behavioural experiment.
-(**C**) Swimming trajectories of wild type (WT, n=37) and *NOS* mutant (NOSΔ11/Δ11, n=18 and NOSΔ23/Δ23, n=8) two-day-old larvae.
+**(B, C)** Immunostaining for NOS antibody in NOS mutant (B, NOSΔ11/Δ11; C, NOSΔ23/Δ23) larvae.
+Acetylated α-tubulin staining (green) highlights the neuropil and cPRC cilia.
+Anterior view of a three-day-old larva."}
+knitr::include_graphics(path = "figures/Fig3_sup1.png")
+
+```
+
+```{r echo=FALSE,fig.cap="**Figure 3 -- figure supplement 2.
+Generation and behavioural characterisation of *NOS* CRISPR knockouts.** (**A**) Schematic diagram of the set-up of the behavioural experiment.
+(**B**) Swimming trajectories of wild type (WT, n=37) and *NOS* mutant (NOSΔ11/Δ11, n=18 and NOSΔ23/Δ23, n=8) two-day-old larvae.
 All trajectories start at 0 x and y position and time 0 corresponding to 10 sec before the onset of 395 nm stimulation from the side.
-(**D**)  Vertical displacement in 30 sec bins of wild type (WT, n=37) and mutant (NOSΔ11/Δ11, n=18 and NOSΔ23/Δ23, n=8) two-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
-(**E**) Comparison of vertical position of wild type and mutant two-day-old larvae after 30 seconds of UV stimulation.
+(**C**)  Vertical displacement in 30 sec bins of wild type (WT, n=37) and mutant (NOSΔ11/Δ11, n=18 and NOSΔ23/Δ23, n=8) two-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
+(**D**) Comparison of vertical position of wild type and mutant two-day-old larvae after 30 seconds of UV stimulation.
 One-way ANOVA with Dunnett’s multiple-comparison test are shown.
 The wild type (WT, n=52) and *NOS* mutant (*NOSΔ11/Δ11*, n=30 and *NOSΔ23/Δ23*, n=16) larvae.
-(**F**) Vertical position of batches of wild type (12 batches) and mutant (NOSΔ11/Δ11, 8 batches and NOSΔ23/Δ23, 3 batches) two-day-old larvae over time under 395 nm UV stimulation.
+(**E**) Vertical position of batches of wild type (12 batches) and mutant (NOSΔ11/Δ11, 8 batches and NOSΔ23/Δ23, 3 batches) two-day-old larvae over time under 395 nm UV stimulation.
 The starting position of each larval trajectory was set to 0.
@@
 Source data for panels B-G."}
-knitr::include_graphics(path = "figures/Fig3_sup1.png")
-
-```
-
-```{r echo=FALSE,fig.cap="**Figure 3 -- figure supplement 2.
+knitr::include_graphics(path = "figures/Fig3_sup2.png")
+
+```
+
+```{r echo=FALSE,fig.cap="**Figure 3 -- figure supplement 3.
 Generation and behavioural characterisation of *NOS* CRISPR knockouts.** (**A,B**) Swimming speed of batches of wild type and *NOS* mutant two-day-old (WT, n=37, NOSΔ11/Δ11, n=18 and NOSΔ23/Δ23, n=8) (A) and three-day-old (WT, n=32, *NOSΔ11/Δ11*, n=26 and *NOSΔ23/Δ23*, n=47) (B) larvae under 395 nm UV stimulation.
@@
 Source data for panels B-G."}
-knitr::include_graphics(path = "figures/Fig3_sup2.png")
+knitr::include_graphics(path = "figures/Fig3_sup3.png")
 

### 2023-12-19 code/Fig1_sup2.R

@@ -58,22 +58,39 @@ panel_NOS3d_ant_HCR <- ggdraw() + draw_image(readPNG("pictures/HCR_72_AP_NOS_89.
   draw_label("INNOS", x = 0.51, y = 0.52, color="white", size = 5, fontface="bold")
 
 
+panel_NOS3d_ant_HCR_IHC <- ggdraw() + draw_image(readPNG("pictures/HCR_IHC_75_NOS_acTub_AP_3_high_46.18um.png")) +
+  draw_label("in situ HCR + IHC", x = 0.1, y = 0.99, size = 10, hjust = 0) +
+  draw_label("NOS", x = 0.12, y = 0.9, color="magenta", size = 11, fontface="italic") +
+  draw_label("acTub", x = 0.33, y = 0.9, color="green", size = 11, fontface="plain") +
+  draw_line(x = c(0.04, 0.26), y = c(0.08, 0.08), color = "white", size = 0.5) +
+  draw_label(expression(paste("10 ", mu, "m")), x = 0.15, y = 0.11, color = "white", size = 8) +
+  draw_label("D", x = 0.95, y = 0.88, size = 6, color = "white") +
+  draw_label("V", x = 0.95, y = 0.72, size = 6, color = "white") +
+  geom_segment(aes(x = x1, y = y1, xend = x2, yend = y2), data = arrow_fluo, color = "white", 
+               arrow = arrow(ends = "both", type = "closed", length = unit(0.1,"cm")),
+               lineend = "butt",
+               linejoin = "mitre",
+               arrow.fill = "white", size = 0.2) +
+  draw_label("cPRC", x = 0.54, y = 0.88, size = 6, color = "white") +
+  draw_line(x = c(0.33, 0.47), y = c(0.81, 0.87), color = "white", size = 0.2) +
+  draw_line(x = c(0.6, 0.71), y = c(0.87, 0.84), color = "white", size = 0.2)
+
 
 
 #combine panels into Figure and save final figure as pdf and png
 #panels of different sizes
 layout <- "
-A#B
+A#B#C
 "
 
-Fig1_sup2 <- panel_NOS3d_ventr_HCR + panel_NOS3d_ant_HCR +
-  patchwork::plot_layout(design = layout, widths = c(1, 0.02, 1)) + #we can change the heights of the rows in our layout (widths also can be defined)
+Fig1_sup2 <- panel_NOS3d_ventr_HCR + panel_NOS3d_ant_HCR + panel_NOS3d_ant_HCR_IHC +
[…]

### 2023-12-19 code/Fig3_and_Fig3_figure_Suppl1.R

@@ -2438,6 +2438,23 @@ ggsave("pictures/CBF_3d_plot.png", limitsize = FALSE,
 panel_Architecture <- ggdraw() + draw_image(readPNG("pictures/NOS-Architecture.png"))
 panel_setup <- ggdraw() + draw_image(readPNG("pictures/vertical_experiment_setup.png"))
 
+#NOS mutant immunostaining
+panel_IHC_3d_NOS11 <- ggdraw() + draw_image(readPNG("pictures/IHC_3d_NOS11_7.png")) +
+  draw_label("IHC:", x = 0.16, y = 0.991, size = 10) +
+  draw_label(expression(paste('NOS'^'Δ11/Δ11')), x = 0.43, y = 0.999, size = 10) +
+  draw_label("NOS", x = 0.12, y = 0.91, color="magenta", size = 10, fontface="plain") +
+  draw_label("acTub", x = 0.33, y = 0.91, color="green", size = 10, fontface="plain") +
+  draw_line(x = c(0.04, 0.392), y = c(0.08, 0.08), color = "white", size = 0.5) +
+  draw_label(expression(paste("20 ", mu, "m")), x = 0.21, y = 0.11, color = "white", size = 8) 
+
+panel_IHC_3d_NOS23 <- ggdraw() + draw_image(readPNG("pictures/IHC_3d_NOS23_8.png")) +
+  draw_label("IHC:", x = 0.16, y = 0.991, size = 10) +
+  draw_label(expression(paste('NOS'^'Δ23/Δ23')), x = 0.43, y = 0.999, size = 10) +
+  draw_label("NOS", x = 0.12, y = 0.91, color="magenta", size = 10, fontface="plain") +
+  draw_label("acTub", x = 0.33, y = 0.91, color="green", size = 10, fontface="plain") +
+  draw_line(x = c(0.04, 0.392), y = c(0.08, 0.08), color = "white", size = 0.5) +
+  draw_label(expression(paste("20 ", mu, "m")), x = 0.21, y = 0.11, color = "white", size = 8) 
+  
 #Martin images
 panel_2d_vd <- ggdraw() + draw_image(readPNG("pictures/vertical_displacement_2dpf_WTvsNOSmix.png"))
 panel_3d_vd <- ggdraw() + draw_image(readPNG("pictures/vertical_displacement_3dpf_WTvsNOSmix.png"))
[…]

### 2023-12-19 Manuscript/Response to the Reviews.Rmd
Response to the Reviews_1

---
title: "Reviews"
author: "Kei Jokura"
date: "2023-09-19"
output: html_document
---

**eLife assessment**

This study reports the discovery of a new circuit mechanism for light-avoidance behavior in the marine annelid, Platynereis dumerilii. Using calcium imaging, molecular perturbations, behavioral measurements, and modeling, the authors provide compelling evidence that nitric oxide is released by postsynaptic neurons onto ciliary photoreceptors to prolong and enhance their response to ultraviolet light. The fundamental new role of nitric oxide described in this study may be conserved across animal phyla and thus will be of broad interests to neuroscientists and neuroendocrinologists.

**Public Reviews:**

**Reviewer #1 (Public Review):**

*Summary: The ciliary photoreceptor cells and its downstream neurons of larval annelid must be orchestrated in a specific pattern to promote downward swimming in response to long duration of UV exposure. The authors first conducted neuroanatomical examination of the circuit to identify NOS-expression neurons (INNOS) that are immediately downstream to the ciliary photoreceptor cells. The INNOS is activated by UV and produces NO. The NOS is required for UV avoidance by Platynereis larvae and neural dynamics of the photoreceptor cells and their downstream circuit. Following up the RNA-seq data with in situ hybridization experiments, the authors found that two unconventional guanylate cyclases, NIT-GC1 and NIT-GC2, are expressed and localized in different subcellular domain of the photoreceptor cells. Experiments using the culture cells and genetically encoded sensors demonstrated that NIT-GC1 can generate cGMP in response to nitric oxide. Finally, authors build a mathematical model that fit the live imaging data and used it to predict how the magnitude of the photoreceptor activation varied by intensity and duration of UV light.*

*Strengths: The authors conducted comprehensive interrogations of the UV avoidance pathway at the molecular and circuit levels, and constructed a mathematical model. The main conclusions are supported by layers of evidence from different assays.*

*Weaknesses: Statistics are missing in both figure legends and methods. The perturbations of genes and molecules were not cell-type-specific and therefore the observed behavioral defect could be attributed to the malfunction of the circuit elsewhere not examined in this study. I suggest adding more explanation about the functions of other NOS-expressing cells and conducting a control experiment to test behavioral response to a non-visual stimulus.*

1.  As you point out, we have performed statistical tests and added explanatory text to figure legends and methods.
2.  Only one NOS homologue has been found in the Platynereis genome. Note that it is specifically expressed in NOS interneurons, which are our target in this study, based on the results of single-cell transcriptome analysis and in situ hybridisation against the NOS gene. However, as you have pointed out, the functions of nitric oxide are diverse and we have revised some of our discussion to include the possibility of its interneurons outside the non-visual system.

**Reviewer #2 (Public Review):**

*Summary: This study is quite thorough, tackling this NO-dependent UV avoidance circuit with both breadth and depth. There are several novel discoveries throughout, but the whole package represents perhaps even more than the sum of these parts.*

*Strengths: The presentation of the work is compelling. The introduction sets up the question and the state of the field very nicely. The discovery of the non-canonical NO receptor pathway in the ciliary photoreceptors is fascinating and will likely open up new avenues for future research into NO-pathways in different species. The use of genetic and pharmacological manipulations of circuit components was well thought-out. The authors applied different experimental techniques expertly throughout the study so that they could develop a comprehensive view from the molecular to the behavioral levels.*

*Weaknesses: While I appreciate the intent of bringing together a large set of measurements from connectomics and calcium imaging in the framework of a model, the model seemed rather poorly constrained. How many parameters are in the model shown in Figure 6A? How many of them are well constrained by experimental measurements? The authors also don't perform sensitivity analysis on the parameters of the model. And ultimately, the conclusion over the model in Figure 7 is somewhat trivial within the unitless construction: larger amplitude and longer duration stimuli lead to increased activation of the downstream neuron thought to lead to the downward swim behavior. I could imagine that a large family of models would arrive at this same result, and without units, there is no way to really test it with new behavioral experiments.*

**Reviewer #3 (Public Review):**

*The transition from planktonic to benthic depends upon several physical and chemical cues. Nitric oxide (NO) is known as a critical player in the induction of larval metamorphosis in several invertebrates. Although NO is a widespread signalling molecule in a broad range of organisms regulating key physiological processes, internal regulatory mechanisms studies are scarce. While the UV sensing in larvae of the annelid Platynereis dumerilii using ciliary photoreceptors has been studied, the neuronal signalling mechanism remains unknown. In this study, Kei Jokura et al. investigated how annelid Platynereis dumerilii larvae detect UV sensing and modulate swimming behaviour through nitric oxide feedback. Using existing resources of Platynereis larval connectome/volume EM data, they identified NOS-expressing interneurons within the ciliary photoreceptors circuit (cPRCs). They demonstrated that NO is produced in cPRCs during UV/violet stimulation by using a fluorescent NO-reporter line. Further, they demonstrated that Nitric oxide signalling mediates UV-avoidance behaviour by using NOS-mutant larvae. Finally, they mapped out the signalled mechanisms of the cPRC circuit using published spatially mapped single-cell transcriptome data of Platynereis larvae, the Ca sensor lines, in situ HCR, and immunostaining. Additionally, by using their findings from Ca imagining data of cPRC, INNOS and INRGWa cells collected in wild-type, NOS knockout and NIT-GC2 morphant larvae, Kei Jokura et al. developed a mixed cellular-circuit-level mathematical model. However, my expertise in mathematical modelling is limited, so I cannot comment on this section.*

*No doubt, the study has been conducted extensively. However, I have a few comments, please see below.*

*Page 4: "In contrast, both two- and three-day-old homozygous NOS-mutant larvae showed a strongly diminished UV avoidance response (Figure 3A, B and Figure 3-figure supplement 1B, C)." Instead of using subjective terms like "strongly," it would be more relevant to provide statistical values. However, I could not locate any means of statistical analysis on larval behaviour. Can the authors indicate the statistical values for all behaviour studies?*

As noted, we have performed statistical analyses and added to the figures in all behavioural experiments. We have also corrected the wording in the text.

*Page 5: "(D) Vertical displacement in 30 sec bins of wild type and mutant (NOSΔ11/Δ11 and NOSΔ23/Δ23) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top." The error bars for WT are too long at the end of the experiment. It is not clear how the authors decided to use this time frame. Did the authors try carrying this out for an extended time period? How did the authors decide on 120 seconds as the time frame for exposure? Authors should provide data on larval behaviour for an extended time.*

The size of the chamber is limited due to the experimental system, so many larvae tend to accumulate at the bottom of the chamber by descending swimming at the end of the experiment. This may have led to variability in the data at the end point and lengthened the error bars.

A 120 s exposure time was most suitable for quantifying larval behaviour from a previous paper by Verasztó et al. 2017.

*Page 13: "During the UV response, prototroch cilia beat slower than trunk cilia, resulting in a head-down stable state ('rear-wheel drive'). In contrast, during the pressure response prototroch cilia beat faster than trunk cilia, leading to a head-up orientation ('front-wheel drive'). Testing this hypothesis will require biophysical experiments and mathematical modelling." Authors should carry out ciliary beating analysis under UV light in the current study with NOS mutant larvae. Since the pressure and UV detection systems are closely related, comparing the difference in ciliary beating is important to demonstrate this hypothesis. Further, did the authors check the Ca sensor GCaMP6s under pressure conditions?*

We have newly analysed the frequency of ciliary beating of larvae by UV stimulation and added Figure 3---figure supplement 3.

For calcium imaging under pressurised conditions, this was done in Bezares-Calderón et al., 2023, see: doi: <https://doi.org/10.1101/2023.02.28.530398>

*Page 18: "strips. One strip contained UV (395 nm) LEDs (SMB1W-395, Roithner Lasertechnik) and the other infrared (810 nm) LEDs (SMB1W-810NR-I, Roithner Lasertechnik)." Authors should test larval swimming behaviour at different wavelengths. Even though they are performed in previous work, the experiment with different wavelengths is necessary to be conducted in NOS mutant larvae in parallel with a control. This will confirm that NOS is principally associated with UV. Further, to demonstrate that this mechanism is associated with ciliary movement, authors need to provide this evidence.*

In this study we have conducted behavioural experiments at different wavelengths (Figure 3D and Figure 3---figure supplement 1F). However, as an explanation was lacking, a more detailed explanation has been added in the text.

------------------------------------------------------------------------

Recommendations for the authors: please note that you control which revisions to undertake from the public reviews and recommendations for the authors

**Reviewer #1 (Recommendations For The Authors):**

1)  *The current introduction focuses on the nitric oxide signaling. It would be helpful for readers to have an introductory section about Platynereis larvae (a total number of neurons, etc) and rationales to use its nervous system as a model to study mechanisms of NO signaling and gating of visual response.*

2)  *The readers would want to know why these larvae swim downward in response to UV and upward to 480nm light. Is that for maintaining the certain depth from the surface of the water?*

3)  *Are there splicing isoforms of NOS in the Platynereis dumerilii? If so, do antibodies and probes for in situ distinguish them? In Drosophila, truncated isoforms can inhibit the function of full length isoform, and therefore it was important to use methods to distinguish spicing isoforms.*

It is not known whether the splicing isoform of NOS is present in Platynereis. It is also not clear how to find out...

4)  *"INRGWs" acronym appears without explanation in the first paragraph of the results.*

We have corrected it.

5)  *In Figure 1C, it is difficult to see the projection patterns of individual cell types. Figure supplement 3 can be combined with the current Figure 1.*

We have moved one figure from Figure 1---supplement 3 to the main figure1D.

6)  *Add more explanation about NOSp::palmi-3xHA reporter. Is it membrane-targeted reporter with the upstream promotor sequence of NOS? Or is it endogenous NOS that was tagged with palmi-3xHA?*

It is a membrane-targeted reporter with an upstream promoter sequence of the NOS gene. We have added a description in the text.

7)  *Show lack of anti-NOS immunostaining in NOS mutant to warrant specificity of the antibody. The subcellular localization of NOS in the dendritic arbors of INNOS is essential for the proposed model. The immunostaining image in Figure 4 -figure supplement 3D can be in the main figure. Is NOS also in the axons of INNOS?*

We redid the immunostaining with NOS antibodies in WT and NOS mutant and added the staining data to the main Figure 1G and Figure 3---supplement 1. NOS was observed to be more strongly localised on the dendrite side of INNOS. We have also added these explanations to the text.

8)  *"that was defective in NOS mutant (Figure 5I)" should be corrected as "Figure 5H".*

We have corrected it.

9)  *Related to Figure 7, how do the larvae respond to a sequence of UV light (e.g. ten times 0.5s ON and 0.5 OFF) or ramping up/down UV light? Can the model make any predictions?*

10) *What are the knock down efficiency of NOS and NIT-GC morpholnio?*

We compared the fluorescence intensity in immunostaining of each morphant larvae and added a graph to Figure 4---supplement 3F.

**Reviewer #2 (Recommendations For The Authors):**

*Why is the behavior of the control animals so different in panels B and C of Figure 3? One group reaches only 15 mm vertical position and the other reaches \~60 mm. Is this just batch variation? Am I missing an experimental variable here? If this is indeed batch variation, then some additional text and statistical analyses might help the reader interpret the behavioral data.*

Our unit notation has been corrected. Additionally we have added statistical tests.

*Really, that's the only, relatively minor issue I could find. This was a pleasure to read, and I learned a lot. Congratulations on an excellent study.*

**Reviewer #3 (Recommendations For The Authors):**

*Page 3, Figure 1 A: The authors reconstructed the cPRC circuit in 3-day-old larvae and detected NOS gene expression in 2-day-old larvae in Figure 1D&E. Can the authors provide a cPRC circuit reconstruction for 2-day-old larvae?*

We have added further data on NOS gene expression in 3-day-old larvae to Figure 1---supplement 2C, although it is difficult for us to show data on the connectome of 2-day-old larvae.

*Page 4, Figure 2: NO produced by UV/violet stimulation to cPRCs: Including a diagram of the larvae would enhance reader understanding.*

We have added a schematic diagram to Figure 2---supplement 1.

*Page 4: Two Platynereis NOS knockout lines (NOSΔ11/Δ11 and Δ23/Δ23) using the CRISPR/Cas9: Could you direct me to the knockout conformation results for NOS knockout lines NOSΔ11/Δ11 and Δ23/Δ23?*

We have shown in Figure 3---supplement 1A.

*Page 4: "Three day-old but not two-day-old NOS-mutant larvae also showed reduced phototactic behaviour, suggesting a function for NOS in the visual eyes that mediate three-day-old phototaxis" This sentence is unclear. Why is it only three days old but not two days old?*

We have performed statistical tests on the data in Figure 3---supplement 1F and added more detailed explanations.

*Page 5: "Figure 3. NOS is required for UV avoidance in Platynereis larvae." Detailed experiment setup, if possible, schematic or real setup images would help to replicate the experiments.*

We have added a schematic diagram to Figure 3---supplement 1B.

*Page 5: "All trajectories start at 0 x and y position and time 0 corresponding to 10 sec after the onset of 395 nm stimulation from the side." How did the authors determine the 10-second duration? Are there any potential reasons for this choice?*

Due to constraints imposed by the experimental set-up and analysis, we found it difficult to track the same larvae for extended periods of time, so we restricted our analysis to a total of 40 s of tracking, starting 10 s before stimulus onset and continuing for 30 s after stimulus onset. A more detailed description has been added in the text.

*"(A) Swimming trajectories of wild type (WT, n=32) and NOS mutant (NOSΔ11/Δ11, n=26 and NOSΔ23/Δ23, n=47) three-day-old larvae." Authors keep shifting between 2-day and 3-day-old larval data in Figures 1, 2, and 3, causing inconsistency.*

Due to sample immobilisation constraints in live imaging, we found it difficult to perform DAF-FM and calcium imaging with 3-day-old larvae. However, we have observed similar patterns of UV-avoidance behaviour and NOS gene expression in 2-day-old larvae as in 3-day-old larvae, and we believe that it is possible to make a correspondence.

*Page 5: "Figure 3. NOS is required for UV avoidance in Platynereis larvae." Why didn't the authors present any statistics on the plots? A statistical test is required to prove that the difference is significant.*

We have added statistical tests to the data in Figure 3.

*Page 5: "Analysis of sGCs in Platynereis indicated that these genes are not expressed in any of the cells of the cPRC circuit (not shown and (Verasztó et al., 2017))." Hence the data is relevant; please provide this data in supplementary.*

We re-analysed previous single-cell data (Achim et al., 2018) and found no sGC homologues detected in cPRC and INNOS. However, a few were detected in INRGW. Additions were made to the SUPPLEMENT data regarding these data and the text was also amended.

*Page 10: "Diagram of the mathematical model with the components, interactions, parameters and equations used to model Ca dynamics." It would be helpful to include a legend explaining the meaning of each term (e.g. what is K, Co, Cp etc) and arrow color. Additionally, the main text should provide detailed descriptions to aid understanding.*

*Page 12: "This activated state is maintained for several tens of second." Can authors point out the data?*

*Page 13: "In the Platynereis circuit, our mathematical model indicates that the magnitude of the NO-dependent signal depends on the intensity and duration of the UV/violet stimulus." Did the authors conduct experiments of various intensity and duration apart from modelling? If not, such experiments should be carried out to validate the mathematical model.*

### 2023-12-20 Manuscript/Response to the Reviews.Rmd
edit
Update response

@@
 
-1.  As you point out, we have performed statistical tests and added explanatory text to figure legends and methods.
+1.  As you point out, we have performed statistical tests and added explanatory text to figure legends.
 2.
@@
 
-We have newly analysed the frequency of ciliary beating of larvae by UV stimulation and added Figure 3---figure supplement 3.
+We have newly analysed the ciliary beat frequency of larvae by UV stimulation and added Figure 3---figure supplement 3.
 
@@
 
+    The 3dpf larvae reveals 90 non-neuronal cell types and 180 neuronal cell types, mapping the complete neuronal connectome.
+This can shed light on the question of how synaptic and non-synaptic networks regulate their various patterns of behaviour.
+
 2)  *The readers would want to know why these larvae swim downward in response to UV and upward to 480nm light. Is that for maintaining the certain depth from the surface of the water?*
+
+    It has been suggested that the ciliary photoreceptor circuit, which senses UV light, and the rhabdomeric photoreceptor circuit, which senses blue light, exchange messages with each other and that the two work together to form a depth gauge.
+By allowing larvae to swim at their preferred depth, the depth gauge influences where they end up when they become adults.
 
@@
 
-We have added further data on NOS gene expression in 3-day-old larvae to Figure 1---supplement 2C, although it is difficult for us to show data on the connectome of 2-day-old larvae.
+We have added further data on NOS gene expression in three-day-old larvae to Figure 1---supplement 2C, although it is difficult for us to show data on the connectome of two-day-old larvae.
 
@@
 
-We have performed statistical tests on the data in Figure 3---supplement 1F and added more detailed explanations.
+The visual eyes (adult eyes) co-express r-opsin1 and Go-opsin1, which are known to be involved in phototaxis, and it is thought that differences in phototaxis behaviour may arise because the visual eyes themselves are still underdeveloped in two-day-old larvae.
 
@@
 Due to constraints imposed by the experimental set-up and analysis, we found it difficult to track the same larvae for extended periods of time, so we restricted our analysis to a total of 40 s of tracking, starting 10 s before stimulus onset and continuing for 30 s after stimulus onset.
-A more detailed description has been added in the text.
 
@@
 
-Due to sample immobilisation constraints in live imaging, we found it difficult to perform DAF-FM and calcium imaging with 3-day-old larvae.
-However, we have observed similar patterns of UV-avoidance behaviour and NOS gene expression in 2-day-old larvae as in 3-day-old larvae, and we believe that it is possible to make a correspondence.
+Due to sample immobilisation constraints in live imaging, we found it difficult to perform DAF-FM and calcium imaging with three-day-old larvae.
+However, we have observed similar patterns of UV-avoidance behaviour and NOS gene expression in two-day-old larvae as in three-day-old larvae, and we believe that it is possible to make a correspondence.
 
@@
 
-We have added statistical tests to the data in Figure 3.
+We have added statistical tests to the data in Figure3.
 
@@
 However, a few were detected in INRGW.
-Additions were made to the SUPPLEMENT data regarding these data and the text was also amended.
+Additions were made to the supplement data regarding these data and the text was also amended.
 

### 2023-12-20 Manuscript/Manuscript.Rmd
Update response

@@
 To characterise the expression pattern of *NOS*, we used in situ hybridization chain reaction (HCR) and transient transgenesis.
-In two- and three-day-old larvae, we detected *NOS* expression in four cells (two of them weakly expressing) in the apical organ region (Figure 1E and Figure 1---figure supplement 2).
-*NOS* was also expressed in the region of the visual eyes (adult eyes) and the pigmented eyespots (Figure 1---figure supplement 2).
-The four apical organ cells, but not the eyes, were also labelled with a *NOS* reporter construct driving palmitoylated tdTomato (Figure 1F).
+In two- and three-day-old larvae, we detected *NOS* expression in four cells (two of them weakly expressing) in the apical organ region (Figure 1E and Figure 1---figure supplement 2A, B).
+*NOS* was also expressed in the region of the visual eyes (adult eyes) and the pigmented eyespots (Figure 1---figure supplement 2A, B).
+The four apical organ cells, but not the eyes, were also labelled with an upstream promoter sequence of the *NOS* gene construct driving membrane-targeted palmitoylated tdTomato reporter (Figure 1F).
 This reporter also revealed the axonal projections of these central NOS-expressing neurons.
-The position and morphology of the four *NOS+* cells allowed us to identify the same four cells as four interneurons (INNOS) in our three-day-old whole-body *Platynereis* volume EM data [@williams2017; @veraszto2020] (Figure 1B,C).
+The position and morphology of the four *NOS+* cells allowed us to identify the same four cells as four interneurons (INNOS) in our three-day-old whole-body *Platynereis* volume EM data [@williams2017; @veraszto2020] (Figure 1C, D).
 In the synaptic connectome, the INNOS cells are postsynaptic to the UV-sensory cPRCs and presynaptic to the INRGWa interneurons (cholinergic neurons that strongly express RGW neuropeptide), which are also cPRC targets (Figure 1C, H).
-The projections of INNOS cells are segregated into input (dendritic) and output (axonal) compartments, with cPRC inputs occurring on the dendritic part in the dense neurosecretory plexus (Figure 1D).
-Immunostaining revealed that the localization of NOS at the protein level is concentrated in the dendrites of the input region even within INNOS (Figure 1G).
-INNOS output-synapses form in the more ventral projection region (Figure 1---figure supplement 3).
+The projections of INNOS cells are segregated into input (dendritic) and output (axonal) compartments, with cPRC inputs occurring on the dendritic part in the dense neurosecretory plexus (Figure 1D and Figure 1---figure supplement 3A).
+Immunostaining revealed that the localization of NOS at the protein level is more strongly concentrated in the dendrites of the input region even within INNOS (Figure 1G and Figure 1---figure supplement 2C).
+INNOS output-synapses form in the more ventral projection region (Figure 1D and Figure 1---figure supplement 3).
 INRGWa neurons synapse on the head serotonergic ciliomotor neurons (Ser-h1), which synapse on the cholinergic MC ciliomotor neuron and cells of the prototroch ciliary band (Figure 1C, H) [@veraszto2017].
@@
 We recovered two deletions (NOSΔ11/Δ11 and Δ23/Δ23), both frame-shift mutations leading to an early stop codon and thus likely representing null alleles (Figure 3---figure supplement 1A).
-We could establish a homozygous line for both mutations indicating that NOS is not an essential gene in *Platynereis*.
-To quantify UV avoidance, we recorded the trajectories of freely swimming wild type and mutant larvae in vertical columns, illuminated laterally from two opposite sides with 395 nm UV light (Figure 3A and Figure 3---figure supplement 1B, C).
+We could establish a homozygous line for both mutations indicating that NOS is not an essential gene in *Platynereis* (Figure 3---figure supplement 1B, C).
+To quantify UV avoidance, we recorded the trajectories of freely swimming wild type and mutant larvae in vertical columns, illuminated laterally from two opposite sides with 395 nm UV light (Figure 3A and Figure 3---figure supplement 2A, B).
 As previously shown, wild-type larvae swim downward following non-directional UV/violet light stimulation [@Veraszto2018].
-In contrast, both two- and three-day-old homozygous *NOS*-mutant larvae showed a significantly diminished UV-avoidance response (Figure 3B, C and Figure 3---figure supplement 1D, E).
+In contrast, both two- and three-day-old homozygous *NOS*-mutant larvae showed a significantly diminished UV-avoidance response (Figure 3B, C and Figure 3---figure supplement 1C, D).
 This phenotype is similar to the defective UV-avoidance of *c-opsin1* mutant larvae [@Veraszto2018] and reveals a requirement for NOS in UV-avoidance behaviour.
-Wild type but not mutant larvae also showed a significantly increase in swimming speed with an increase in ciliary beat frequency under UV light that may be due to downward swimming trajectories (swimming in the direction of gravity) (Figure 3---figure supplement 2).
-We also tested directional phototaxis, by exposing larvae to 480 nm directional collimated light from the top of the column.
-Three-day-old but not two-day-old *NOS*-mutant larvae also showed reduced phototactic behaviour, suggesting a function for *NOS* in the visual eyes that mediate three-day-old phototaxis [@Randel2014a] (Figure 3D and Figure 3---figure supplement 1F).
+Wild type but not mutant larvae also showed a significantly increase in swimming speed with an increase in ciliary beat frequency under UV light that may be due to downward swimming trajectories (swimming in the direction of gravity) (Figure 3---figure supplement 3).
+We also tested directional phototaxis, by exposing larvae to 480 nm directional collimated light from the top of the column (Figure 3---figure supplement 2A).
+Three-day-old but not two-day-old *NOS*-mutant larvae also showed reduced phototactic behaviour, suggesting a function for *NOS* in the visual eyes that mediate three-day-old phototaxis [@Randel2014a] (Figure 3D and Figure 3---figure supplement 1E).
 
@@
 NO binding to the heme group of sGC leads to increased cyclic guanosine monophosphate (cGMP) production.
-Analysis of sGCs in *Platynereis* indicated that these genes are not expressed in any of the cells of the cPRC circuit (not shown and [@veraszto2017]).
+Analysis of sGCs in *Platynereis* showed that these genes were not expressed in cPRC and INNOS in the cells of the cPRC circuit, but only in INRGW [@veraszto2017].
 Recently, Moroz and coworkers reported an atypical but widely conserved family of guanylate cyclases with a NIT (nitrite/nitrate sensing) domain (PF08376) (NIT-GC) as potential mediators of NO signalling [@Moroz2020].
@@
 In immunostainings, we found that NIT-GC1 was localise to the region corresponding to the axonal projections of the cPRCs in the anterior neurosecretory plexus (Figure 4E).
-Co-immunostaining with the rabbit NIT-GC1 antibody and a custom rat antibody raised against *Platynereis* NOS revealed the localisation of both proteins in dots in the neurosecretory neuropil (Figure 4---figure supplement 3D).
 In contrast, NIT-GC2 specifically labelled the ramified sensory cilia of the cPRCs (Figure 4F).
@@
 To test the in vivo function of NIT-GC1 and NIT-GC2 in cPRC responses, we combined Ca^2+^ imaging with morpholino-mediated knockdowns.
-We used two translation-blocking morpholinos for each *NIT-GC* gene and tested knockdown efficiency by immunostaining injected animals with the NIT-GC1 and NIT-GC2 antibodies (Figure 4---figure supplement 3E,F).
-For both genes, the morpholinos led to a strong reduction in the respective antibody signal, confirming efficient knockdown and antibody specificity.
+We used two translation-blocking morpholinos for each *NIT-GC* gene and tested knockdown efficiency by immunostaining injected animals with the NIT-GC1 and NIT-GC2 antibodies (Figure 4---figure supplement 3D,E).
+For both genes, the morpholinos led to a strong reduction in the respective antibody signal, confirming efficient knockdown and antibody specificity (Figure 4---figure supplement 3F).
 

### 2024-01-06 Manuscript/Manuscript.Rmd
New Figure7

@@
 ```{r echo=FALSE,fig.cap="**Figure 7.
-Magnitude of cPRC activation as  a function of stimulus duration and intensity.** Median values of the $AUP$ are presented as a heatmap with dark blue indicating flat $C_P(t)$ traces with median $AUP$ close to 0 and red indicating that the $C_P(t)$ traces have a clearly defined delayed activation peak.
+Comparison of the magnitude of cPRC activation as a function of stimulus duration and intensity.** (**A**) Change in magnitude calculated from mathematical models.
+Median values of the $AUP$ are presented as a heatmap with dark blue indicating flat $C_P(t)$ traces with median $AUP$ close to 0 and red indicating that the $C_P(t)$ traces have a clearly defined delayed activation peak.
+(**B**) Change in magnitude calculated from experiments data.
+The amount of change in cPRC fluorescence intensity (ΔF/F0) when the laser power (%) and time duration (seconds) of the stimulus are varied using calcium imaging.
+Laser power is shown as relative values.
 "}
 
-knitr::include_graphics(path = "figures/Fig7.svg")
+knitr::include_graphics(path = "figures/Fig7.png")
 ```

### 2024-01-06 code/Fig7.R

@@ -1,0 +1,496 @@
+
+#Code to assemble Fig7 of the Jokura et al Platynereis NOS paper
+#2023 December - Kei Jokura, Gaspar Jekely
+
+# source packages ---------------------------------------------------------
+source("code/Packages_to_load.R")
+
+# read data ---------------------------------------------------------------
+
+folder_path <- "data/20231223_405_stimulation/csv"
+csv_files <- list.files(path = folder_path, pattern = "\\.csv$", full.names = TRUE)
+remove_common_part <- function(file_path) {
+  common_part <- "ROI60_PD1.54_scan471.86ms_pin200.7_"
+  file_path <- gsub(common_part, "", file_path)
+  return(file_path)
+}
+combined_df_1 <- map_dfr(csv_files, ~{
+  removed_part <- remove_common_part(.x)
+  read_csv(.x) %>% mutate(removed_part = removed_part)
+})
+
+folder_path <- "data/20231224_405_stimulation/csv"
+csv_files <- list.files(path = folder_path, pattern = "\\.csv$", full.names = TRUE)
+remove_common_part <- function(file_path) {
+  common_part <- "ROI60_PD1.54_scan471.86ms_pin200.7_"
+  file_path <- gsub(common_part, "", file_path)
+  return(file_path)
+}
+combined_df_2 <- map_dfr(csv_files, ~{
+  removed_part <- remove_common_part(.x)
+  read_csv(.x) %>% mutate(removed_part = removed_part)
+})
+
+folder_path <- "data/20231230_405_stimulation/csv"
+csv_files <- list.files(path = folder_path, pattern = "\\.csv$", full.names = TRUE)
+remove_common_part <- function(file_path) {
+  common_part <- "ROI60_PD1.54_scan471.86ms_pin200.7_"
+  file_path <- gsub(common_part, "", file_path)
+  return(file_path)
+}
+combined_df_3 <- map_dfr(csv_files, ~{
+  removed_part <- remove_common_part(.x)
+  read_csv(.x) %>% mutate(removed_part = removed_part)
+})
+
+
+combined_df <- rbind(combined_df_1, combined_df_2, combined_df_3)
+combined_df <- combined_df %>% 
+  separate(removed_part, into = c("data","date","UV","Stim",
+                                  "CSV_1","ID","ID_num","power",
+                                  "time","position","CSV_2"), sep = "[_/.]")
+
+combined_df <- combined_df %>% 
+  rename(frame = ...1, intensity = Mean1)
[…]

### 2024-01-06 Manuscript/Response to the Reviews.Rmd
Edited responses to reviewers

@@
 
+    We performed new experiments, both in the mathematical model and in real calcium imaging, changing the amplitude and duration of UV stimulation. The results are shown in Figure 7A and B.
+
 10) *What are the knock down efficiency of NOS and NIT-GC morpholnio?*
@@
 
+In Figure4A and B, it can be seen that cPRC activity continues after UV stimulation.
+
 *Page 13: "In the Platynereis circuit, our mathematical model indicates that the magnitude of the NO-dependent signal depends on the intensity and duration of the UV/violet stimulus." Did the authors conduct experiments of various intensity and duration apart from modelling?
 If not, such experiments should be carried out to validate the mathematical model.*
+
+We have performed a new calcium imaging experiment: a plot of the change in calcium levels in the cPRC when changing the power and duration of the UV stimulus is added in Figure7B.

### 2024-01-06 Manuscript/Manuscript.Rmd
Edited responses to reviewers
Update source data

@@
 output:
+  pdf_document:
+    latex_engine: lualatex
+    template: default
   html_document:
@@
   word_document: default
-  pdf_document:
-    latex_engine: lualatex
-    template: default
 geometry: left=0.8in,right=0.8in,top=0.8in,bottom=0.5in
@@
 *P < 0.05.
-<br> Figure 3 -- source data 1-5.
-Source data for panels A-E."}
+<br> Figure 3 -- source data 1-7.
+Source data for panels A-G."}
 
@@
 Source data for panels A, B and I-N."}
-
 knitr::include_graphics(path = "figures/Fig4.png")
@@
 The amount of change in cPRC fluorescence intensity (ΔF/F0) when the laser power (%) and time duration (seconds) of the stimulus are varied using calcium imaging.
+The intensity change is indicated by the size of the circle and the colour gradient from blue to red.
 Laser power is shown as relative values.
+<br> Figure 7 -- source data 1.
+Bubble chart readings.
 "}
@@
 *P < 0.05.
-<br> Figure 3 -- figure suppmenet 1 -- source data 1-6.
-Source data for panels B-G."}
+<br> Figure 3 -- figure suppmenet 2 -- source data 1-4.
+Source data for panels B-E."}
 knitr::include_graphics(path = "figures/Fig3_sup2.png")
@@
 One-tailed paired t-test; p-values (‘**’ <0.01 or ‘*’ <0.05) are shown.
-<br> Figure 3 -- figure suppmenet 2 -- source data 1-6.
-Source data for panels B-G."}
+<br> Figure 3 -- figure suppmenet 3 -- source data 1-6.
+Source data for panels A-F."}
 knitr::include_graphics(path = "figures/Fig3_sup3.png")
@@
 (**F**) The fluorescence intensities of the NIT-GC1 or NIT-GC2 in the control and morphant (MO1 and MO2) larvae (n=6 each) were normalized to the Acetylated α-tubulin signals and plotted as a relative fluorescence.
-One-way ANOVA with Dunnett’s multiple-comparison test; p-values (‘***’ <0.001) are shown."}
+One-way ANOVA with Dunnett’s multiple-comparison test; p-values (‘***’ <0.001) are shown.
+<br> Figure 4 -- figure suppmenet 3 -- source data 1.
+Comparison NIT-GCs signals readings."}
 knitr::include_graphics(path = "figures/Fig4_sup3.png")
@@
 "}
-knitr::include_graphics(path = "figures/Figure6_fig_Suppl1.svg")
+knitr::include_graphics(path = "figures/Figure6_fig_Suppl1.png")
 
@@
 Panels C, F and I show individual recordings fitted to the data from INRGW cells."}
-knitr::include_graphics(path = "figures/Figure6_fig_Suppl2.svg")
+knitr::include_graphics(path = "figures/Figure6_fig_Suppl2.png")
 ```
@@
 Comparison of fits to the cPRC recordings from NIT-GC2 morphant larvae using the NIT-GC2 morpholino model or the NOS-knockout model.**  Thin coloured curves indicate individual recordings, thin grey curves are simulated Ca^2+^ traces, thick coloured curves indicate averages of the recordings and thick black curves represent the mean of the fits."}
-knitr::include_graphics(path = "figures/Figure6_fig_Suppl3.svg")
+knitr::include_graphics(path = "figures/Figure6_fig_Suppl3.png")
 ```
@@
 Where violin plots are not shown, these parameters are associated with terms/equations that are excluded from the fitting procedure (see Methods for details)." }
-knitr::include_graphics(path = "figures/Figure6_fig_Suppl4.svg")
+knitr::include_graphics(path = "figures/Figure6_fig_Suppl4.png")
 ```
@@
 Also shown using markers are the values of the individual parameter combinations."}
-knitr::include_graphics(path = "figures/Figure6_fig_Suppl5.svg")
-```
+knitr::include_graphics(path = "figures/Figure6_fig_Suppl5.png")
+```

### 2024-01-06 code/Fig1_sup2.R

@@ -93,4 +93,4 @@ ggsave("Manuscript/figures/Fig1_sup2.png", limitsize = FALSE,
        units = c("px"), Fig1_sup2, width = 1800, height = 688, bg='white') 
 
 ggsave("Manuscript/figures/Fig1_sup2.pdf", limitsize = FALSE, 
-       units = c("px"), Fig1_sup2, width = 1200, height = 688)  
+       units = c("px"), Fig1_sup2, width = 1800, height = 688)  

### 2024-01-06 code/Fig2.R

@@ -22,13 +22,13 @@ DAFFM <- rbind(long_cPRC, long_NC)
 #sample size control
 DAFFM %>%
   filter(stimuli=='NC') %>%
-  select(sample) %>%
+  dplyr::select(sample) %>%
   unique()
 
 #sample size cPRC
 DAFFM %>%
   filter(stimuli=='cPRC') %>%
-  select(sample) %>%
+  dplyr::select(sample) %>%
   unique()
 DAFFM
 

### 2024-01-06 code/Fig3_and_Fig3_figure_Suppl1.R

@@ -613,7 +613,7 @@ WTvsNOS_2d %>%
     "dark_30", "dark_60", "dark_90", "dark_120", "dark_150",
     "dark3_30", "dark3_60", "dark3_90", "dark3_120", "dark3_150", "dark3_180"))) %>%
   filter(genotype %in% c("WT", "NOS11x11", "NOS23x23")) %>%
-  write_csv("source_data/Figure3_supplement1_source_data2.csv")
+  write_csv("source_data/Figure3_supplement2_source_data4.csv")
 
 
 # save plot ---------------------------------------------------------------
@@ -1505,7 +1505,7 @@ L_NAME_3d %>%
   filter(!(condition %in% c(
     "dark_30", "dark_60", "dark_90", "dark_120", "dark_150",
     "dark3_30", "dark3_60", "dark3_90", "dark3_120", "dark3_150", "dark3_180"))) %>%
-  write_csv("source_data/Figure3_source_data5.csv")
+  write_csv("source_data/Figure3_source_data7.csv")
 
 
 # save plot ---------------------------------------------------------------
@@ -1550,7 +1550,7 @@ df_2d_nor %>%
 # save to source data-------------------------------------------------------
 
 df_2d_nor %>%
-  write_csv("source_data/Figure3_supplement1_source_data1.csv")
+  write_csv("source_data/Figure3_supplement2_source_data1.csv")
 
 
 
@@ -1639,7 +1639,7 @@ df_2d_nor %>%
 df_2d_nor %>%
   filter(sec > 0 & sec <= 40) %>%
   filter(Genotype %in% c("WT", "NOS11x11", "NOS23x23")) %>%
-  write_csv("source_data/Figure3_supplement1_source_data3.csv")
+  write_csv("source_data/Figure3_supplement2_source_data2.csv")
 
 
 # save plot ---------------------------------------------------------------
@@ -1650,8 +1650,11 @@ ggsave("pictures/vertical_tracking_position_sideUV_2d.png", limitsize = FALSE,
 # statistics analysis of the 2d tracking mean vertical position (side UV after 30sec)--------------------------------------
 
 #Change the name to perform the Dunnett
-df_2d_nor_stat <- df_2d_nor %>% 
-  mutate(Genotype = ifelse(Genotype == "WT", "cont", Genotype))
+df_2d_nor_stat <- df_2d_nor %>%
+  mutate(Genotype = case_when(
+    Genotype == "WT" ~ "cont",
+    TRUE ~ Genotype
+  ))
 
[…]

### 2024-01-06 code/Fig4_sup3.R

@@ -94,8 +94,8 @@ summary(glht(aov(y_NIT2 ~ vx_NIT2),
 
 # save to source data-------------------------------------------------------
 
-#df_L_NAME_tracking %>%
-#  write_csv("source_data/Figure3_supplement1_source_data6.csv")
+immuno_NIT %>%
+  write_csv("source_data/Figure4_supplement3_source_data1.csv")
 
 
 # save plot ---------------------------------------------------------------

### 2024-02-05 Manuscript/Manuscript.Rmd
MS, version info

@@
 One-way ANOVA with Dunnett’s multiple-comparison test are shown.
-*P < 0.05.
+*P < 0.05, **P < 0.01.
 <br> Figure 3 -- source data 1-7.
@@
 Anterior view of a two-day-old larva.
-(**F**) The fluorescence intensities of the NIT-GC1 or NIT-GC2 in the control and morphant (MO1 and MO2) larvae (n=6 each) were normalized to the Acetylated α-tubulin signals and plotted as a relative fluorescence.
+(**F**) The fluorescence intensities of the NIT-GC1 or NIT-GC2 in the control and morphant (MO1 and MO2) larvae (n=6 each) were normalized to the acetylated α-tubulin signals and plotted as a relative fluorescence.
 One-way ANOVA with Dunnett’s multiple-comparison test; p-values (‘***’ <0.001) are shown.

### 2024-02-05 Manuscript/Manuscript.tex
MS, version info

@@
 \makeatother
-\usepackage{svg}
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
 \emph{NOS} expression in four cells (two of them weakly expressing) in
-the apical organ region (Figure 1D and Figure 1---figure supplement 2).
-\emph{NOS} was also expressed in the region of the visual eyes (adult
-eyes) and the pigmented eyespots (Figure 1---figure supplement 2). The
-four apical organ cells, but not the eyes, were also labelled with a
-\emph{NOS} reporter construct driving palmitoylated tdTomato (Figure
-1E). This reporter also revealed the axonal projections of these central
-NOS-expressing neurons. The position and morphology of the four
+the apical organ region (Figure 1E and Figure 1---figure supplement 2A,
+B). \emph{NOS} was also expressed in the region of the visual eyes
+(adult eyes) and the pigmented eyespots (Figure 1---figure supplement
+2A, B). The four apical organ cells, but not the eyes, were also
+labelled with an upstream promoter sequence of the \emph{NOS} gene
+construct driving membrane-targeted palmitoylated tdTomato reporter
+(Figure 1F). This reporter also revealed the axonal projections of these
+central NOS-expressing neurons. The position and morphology of the four
 \emph{NOS+} cells allowed us to identify the same four cells as four
@@
 volume EM data (Verasztó et al., 2020; Williams et al., 2017) (Figure
-1B,C). In the synaptic connectome, the INNOS cells are postsynaptic to
-the UV-sensory cPRCs and presynaptic to the INRGWa interneurons, which
-are also cPRC targets (Figure 1C, F). The projections of INNOS cells are
+1C, D). In the synaptic connectome, the INNOS cells are postsynaptic to
+the UV-sensory cPRCs and presynaptic to the INRGWa interneurons
+(cholinergic neurons that strongly express RGW neuropeptide), which are
+also cPRC targets (Figure 1C, H). The projections of INNOS cells are
 segregated into input (dendritic) and output (axonal) compartments, with
 cPRC inputs occurring on the dendritic part in the dense neurosecretory
-plexus. INNOS output-synapses form in the more ventral projection region
-(Figure 1---figure supplement 3). INRGWa neurons synapse on the head
+plexus (Figure 1D and Figure 1---figure supplement 3A). Immunostaining
+revealed that the localization of NOS at the protein level is more
+strongly concentrated in the dendrites of the input region even within
+INNOS (Figure 1G and Figure 1---figure supplement 2C). INNOS
+output-synapses form in the more ventral projection region (Figure 1D
+and Figure 1---figure supplement 3). INRGWa neurons synapse on the head
 serotonergic ciliomotor neurons (Ser-h1), which synapse on the
 cholinergic MC ciliomotor neuron and cells of the prototroch ciliary
-band (Figure 1C, F) (Verasztó et al., 2017).
+band (Figure 1C, H) (Verasztó et al., 2017).
 
 \begin{figure}
-\includegraphics[width=25in]{figures/Fig1} \caption{**Figure 1.
+\includegraphics[width=34.72in]{figures/Fig1} \caption{**Figure 1.
 Identification of *NOS*-expressing interneurons (INNOS) within the cPRC circuit.** **(A)** Scanning electron microscopy image of a three-day-old *Platynereis* larva.
@@
 NO-reporter DAF-FM into zygotes and imaged two-day-old larvae while
-exposing the region of cPRC cilia to 405 nm violet light. To mark cell
-outlines, we coinjected mRNA encoding a red-fluorescent reporter
-(RGECO), allowing the identification of the cPRCs. Following light
-stimulation in the region of the ramified cilia of the cPRCs, we
-detected an increase in DAF-FM fluorescence in the anterior
-neurosecretory neuropil, the region of INNOS projections. This increase
-did not occur in larvae in which we illuminated a control area (Figure
-2).
+exposing the region of cPRC cilia to 405 nm violet light (Figure
+2---figure supplement 1). To mark cell outlines, we coinjected mRNA
+encoding a red-fluorescent reporter (RGECO), allowing the identification
+of the cPRCs. Following light stimulation in the region of the ramified
+cilia of the cPRCs, we detected an increase in DAF-FM fluorescence in
+the anterior neurosecretory neuropil, the region of INNOS projections.
+This increase did not occur in larvae in which we illuminated a control
+area (Figure 2).
 
 \begin{figure}
-\includegraphics[width=29.17in]{figures/Fig2} \caption{**Figure 2.
+\includegraphics[width=41.67in]{figures/Fig2} \caption{**Figure 2.
 NO produced by UV/violet stimulation to cPRCs.** (**A**) DAF-FM fluorescence in the region of the neurosecretory neuropil.
@@
 supplement 1A). We could establish a homozygous line for both mutations
-indicating that NOS is not an essential gene in \emph{Platynereis}. To
-quantify UV avoidance, we recorded the trajectories of freely swimming
-wild type and mutant larvae in vertical columns, illuminated laterally
-from two opposite sides with 395 nm UV light (Figure 3A and Figure
-3---figure supplement 1B). As previously shown, wild-type larvae swim
-downward following non-directional UV/violet light stimulation (Verasztó
-et al., 2018). In contrast, both two- and three-day-old homozygous
-\emph{NOS}-mutant larvae showed a strongly diminished UV-avoidance
-response (Figure 3A, B and Figure 3---figure supplement 1B, C). This
-phenotype is similar to the defective UV-avoidance of \emph{c-opsin1}
-mutant larvae (Verasztó et al., 2018) and reveals a requirement for NOS
-in UV-avoidance behaviour. Wild type but not mutant larvae also showed
-an increase in swimming speed under UV light that may be due to downward
-swimming trajectories (swimming in the direction of gravity) (Figure 3B
-and Figure 3---figure supplement 1C). We also tested directional
-phototaxis, by exposing larvae to 480 nm directional collimated light
-from the top of the column. Three-day-old but not two-day-old
+indicating that NOS is not an essential gene in \emph{Platynereis}
+(Figure 3---figure supplement 1B, C). To quantify UV avoidance, we
+recorded the trajectories of freely swimming wild type and mutant larvae
+in vertical columns, illuminated laterally from two opposite sides with
+395 nm UV light (Figure 3A and Figure 3---figure supplement 2A, B). As
+previously shown, wild-type larvae swim downward following
+non-directional UV/violet light stimulation (Verasztó et al., 2018). In
+contrast, both two- and three-day-old homozygous \emph{NOS}-mutant
+larvae showed a significantly diminished UV-avoidance response (Figure
+3B, C and Figure 3---figure supplement 1C, D). This phenotype is similar
+to the defective UV-avoidance of \emph{c-opsin1} mutant larvae (Verasztó
+et al., 2018) and reveals a requirement for NOS in UV-avoidance
+behaviour. Wild type but not mutant larvae also showed a significantly
+increase in swimming speed with an increase in ciliary beat frequency
+under UV light that may be due to downward swimming trajectories
+(swimming in the direction of gravity) (Figure 3---figure supplement 3).
+We also tested directional phototaxis, by exposing larvae to 480 nm
+directional collimated light from the top of the column (Figure
+3---figure supplement 2A). Three-day-old but not two-day-old
 \emph{NOS}-mutant larvae also showed reduced phototactic behaviour,
@@
 three-day-old phototaxis (Randel et al., 2014) (Figure 3D and Figure
-3---figure supplement 1G).
+3---figure supplement 1E).
 
@@
 L-NAME. Larvae incubated for 5 min in 0.1 mM or 1 mM L-NAME showed a
-dose-dependent inhibition of UV avoidance. In contrast, phototaxis was
-not affected (Figure 3C, E). Overall, our results indicate an acute
-requirement for NOS signalling in UV-avoidance and a possible indirect,
-developmental role in the visual system, reminiscent of the function of
-NO signalling in \emph{Drosophila} eye development (Gibbs and Truman,
-1998).
+dose-dependent inhibition of UV avoidance (Figure 3E, F). In contrast,
+phototaxis was not affected (Figure 3G). Overall, our results indicate
+an acute requirement for NOS signalling in UV-avoidance and a possible
+indirect, developmental role in the visual system, reminiscent of the
+function of NO signalling in \emph{Drosophila} eye development (Gibbs
+and Truman, 1998).
 
 \begin{figure}
-\includegraphics[width=33.33in]{figures/Fig3} \caption{**Figure 3.
+\includegraphics[width=44.44in]{figures/Fig3} \caption{**Figure 3.
 NOS is required for UV avoidance in *Platynereis* larvae.**  (**A**) Swimming trajectories of wild type (WT, n=32) and *NOS* mutant (*NOSΔ11/Δ11*, n=26 and *NOSΔ23/Δ23*, n=47) three-day-old larvae.
-All trajectories start at 0 x and y position and time 0 corresponding to 10 sec after the onset of 395 nm stimulation from the side.
+All trajectories start at 0 x and y position and time 0 corresponding to 10 sec before the onset of 395 nm stimulation from the side.
 (**B**) Vertical position of batches of wild type and mutant three-day-old larvae over time under 395 nm UV stimulation.
 The starting position of each larval trajectory was set to 0.
-(**C**) Vertical position of batches of control and L-NAME-treated (0.1 and 1 mM) three-day-old larvae over time under 395 nm UV stimulation.
+(**C**) Comparison of vertical position of wild type and mutant three-day-old larvae after 30 seconds of UV stimulation.
+One-way ANOVA with Dunnett’s multiple-comparison test are shown.
+The wild type (WT, n=32) and *NOS* mutant (*NOSΔ11/Δ11*, n=26 and *NOSΔ23/Δ23*, n=47) larvae.
+(**D**) Vertical displacement in 30 sec bins of wild type (10 batches) and mutant (*NOSΔ11/Δ11*, 5 batches and *NOSΔ23/Δ23*, 7 batches) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
+One-way ANOVA with Dunnett’s multiple-comparison test are shown.
+*P < 0.05.
+(**E**) Vertical position of batches of control and L-NAME-treated (0.1 and 1 mM) three-day-old larvae over time under 395 nm UV stimulation.
 The starting position of each larval trajectory was set to 0.
-(**D**) Vertical displacement in 30 sec bins of wild type and mutant (*NOSΔ11/Δ11* and *NOSΔ23/Δ23*) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
-(**E**) Vertical displacement in 30 sec bins of control and L-NAME-treated (0.1 and 1 mM) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
-<br> Figure 3 -- source data 1-5.
-Source data for panels A-E.}\label{fig:unnamed-chunk-3}
+(**F**) Comparison of vertical position of control (n=6) and L-NAME-treated (0.1 and 1 mM, n=6 each) three-day-old larvae after 30 seconds of UV stimulation.
+One-way ANOVA with Dunnett’s multiple-comparison test are shown.
+(**G**) Vertical displacement in 30 sec bins of control (15 batches) and L-NAME-treated (0.1 and 1 mM, 15 batches each) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
+One-way ANOVA with Dunnett’s multiple-comparison test are shown.
+*P < 0.05.
+<br> Figure 3 -- source data 1-7.
+Source data for panels A-G.}\label{fig:unnamed-chunk-3}
 \end{figure}
@@
 the heme group of sGC leads to increased cyclic guanosine monophosphate
-(cGMP) production. Analysis of sGCs in \emph{Platynereis} indicated that
-these genes are not expressed in any of the cells of the cPRC circuit
-(not shown and (Verasztó et al., 2017)). Recently, Moroz and coworkers
-reported an atypical but widely conserved family of guanylate cyclases
-with a NIT (nitrite/nitrate sensing) domain (PF08376) (NIT-GC) as
-potential mediators of NO signalling (Moroz et al., 2020). To identify
-NIT-GCs in \emph{Platynereis}, we searched transcriptome resources and
-retrieved 15 potential NIT-GC homologs (Figure 4---figure supplement 1
-and 2). To analyse the relationship of these sequences to metazoan
-NIT-GCs, we retrieved protein sequences with a CYC domain from the
-transcriptome and genome databases of 45 metazoan and 2 choanoflagellate
-species. We carried out cluster analysis and phylogenetic reconstruction
-on a group of membrane-bound guanylate cyclases with sGCs as an
-outgroup. In agreement with Moroz et al. (Moroz et al., 2020), we found
-a group of GCs with NIT domains with representatives in placozoans,
-cnidarians, some ecdysozoans, echinoderms, and lophotrochozoans. The 15
-\emph{Platynereis} sequences belonged to several deeply diverged clades
-in the phylogenetic tree (Figure 4---figure supplement 1 and 2).
+(cGMP) production. Analysis of sGCs in \emph{Platynereis} showed that
+these genes were not expressed in cPRC and INNOS in the cells of the
+cPRC circuit, but only in INRGW (Verasztó et al., 2017). Recently, Moroz
+and coworkers reported an atypical but widely conserved family of
+guanylate cyclases with a NIT (nitrite/nitrate sensing) domain (PF08376)
+(NIT-GC) as potential mediators of NO signalling (Moroz et al., 2020).
+To identify NIT-GCs in \emph{Platynereis}, we searched transcriptome
+resources and retrieved 15 potential NIT-GC homologs (Figure 4---figure
+supplement 1 and 2). To analyse the relationship of these sequences to
+metazoan NIT-GCs, we retrieved protein sequences with a CYC domain from
+the transcriptome and genome databases of 45 metazoan and 2
+choanoflagellate species. We carried out cluster analysis and
+phylogenetic reconstruction on a group of membrane-bound guanylate
+cyclases with sGCs as an outgroup. In agreement with Moroz et al. (Moroz
+et al., 2020), we found a group of GCs with NIT domains with
+representatives in placozoans, cnidarians, some ecdysozoans,
+echinoderms, and lophotrochozoans. The 15 \emph{Platynereis} sequences
+belonged to several deeply diverged clades in the phylogenetic tree
+(Figure 4---figure supplement 1 and 2).
 
@@
 was localise to the region corresponding to the axonal projections of
-the cPRCs in the anterior neurosecretory plexus (Figure 4E).
-Co-immunostaining with the rabbit NIT-GC1 antibody and a custom rat
-antibody raised against \emph{Platynereis} NOS revealed the localisation
-of both proteins in dots in the neurosecretory neuropil (Figure
-4---figure supplement 3D). In contrast, NIT-GC2 specifically labelled
-the ramified sensory cilia of the cPRCs (Figure 4F). These different
-subcellular localisations suggest that the two NIT-GCs are involved in
-different intracellular signalling processes in the ciliary and axonal
-regions of the cPRCs.
+the cPRCs in the anterior neurosecretory plexus (Figure 4E). In
+contrast, NIT-GC2 specifically labelled the ramified sensory cilia of
+the cPRCs (Figure 4F). These different subcellular localisations suggest
+that the two NIT-GCs are involved in different intracellular signalling
+processes in the ciliary and axonal regions of the cPRCs.
 
@@
 injected animals with the NIT-GC1 and NIT-GC2 antibodies (Figure
-4---figure supplement 3E,F). For both genes, the morpholinos led to a
+4---figure supplement 3D,E). For both genes, the morpholinos led to a
 strong reduction in the respective antibody signal, confirming efficient
-knockdown and antibody specificity.
+knockdown and antibody specificity (Figure 4---figure supplement 3F).
 
@@
 \begin{figure}
-\centering
-\includesvg{figures/Fig7.svg}
-\caption{\textbf{Figure 7. Magnitude of cPRC activation as a function of
-stimulus duration and intensity.} Median values of the \(AUP\) are
-presented as a heatmap with dark blue indicating flat \(C_P(t)\) traces
-with median \(AUP\) close to 0 and red indicating that the \(C_P(t)\)
-traces have a clearly defined delayed activation peak.}
+\includegraphics[width=52.08in]{figures/Fig7} \caption{**Figure 7.
+Comparison of the magnitude of cPRC activation as a function of stimulus duration and intensity.** (**A**) Change in magnitude calculated from mathematical models.
+Median values of the $AUP$ are presented as a heatmap with dark blue indicating flat $C_P(t)$ traces with median $AUP$ close to 0 and red indicating that the $C_P(t)$ traces have a clearly defined delayed activation peak.
+(**B**) Change in magnitude calculated from experiments data.
+The amount of change in cPRC fluorescence intensity (ΔF/F0) when the laser power (\%) and time duration (seconds) of the stimulus are varied using calcium imaging.
+The intensity change is indicated by the size of the circle and the colour gradient from blue to red.
+Laser power is shown as relative values.
+<br> Figure 7 -- source data 1.
+Bubble chart readings.
+}\label{fig:unnamed-chunk-7}
 \end{figure}
@@
 stimulus ROI.
+
+\textbf{Ciliary beat frequency analysis}
+
+Larvae were imaged at room temperature on a Zeiss AxioObserver Z1
+Microscope (Carl Zeiss Microscopy, Germany) with a C-Apochromat 10X
+lens, 850 nm power LED (OSRAM OSLON® Black, SFH 4716S) and filter sets
+equipped with a 405/40 ET excitation Bandpass, Beamsplitter dichroic
+mirror 495LPXR and without an emission filter with a IR camera
+(UI-3360CP-NIR-GL Rev.2, IDS Imaging Development Systems GmbH, Germany)
+frame rate of 200 frame/sec with 0.2 msec pulse controlled by the LED
+stroboscopic illumination system synchronized with the exposure signals
+from the camera (TYM-002, developed by Dr.Shoji A. Baba, Ochanomizu
+University, Otsuka, Japan) and an image size of 640 x 480 pixels. The
+larvae were stimulated with a 395 nm LED (M395L5, Thorlabs). Kymographs
+of the cilia obtained with a Fiji were used to measure the ciliary beat
+frequency.
 
@@
 
-This work was funded by the Wellcome Trust (214337/Z/18/Z). This project
-has received funding from the European Research Council (ERC) under the
-European Union's Horizon 2020 research and innovation programme (grant
-agreement No 101020792). KJ has been supported by a JSPS Overseas
-Research Fellowship, LAYG by a BBSRC Discovery fellowship
-(BB/W010305/1), PS by a Wellcome Trust Institutional Strategic Support
-Award (204909/Z/16/Z), and KCAW by the EPSRC Hub for Quantitive
-Modelling in Healthcare (EP/T017856/1). We thank Dr.~Francesca Carlisle
-for the mycoplasma testing of COS-7 cells.
+We thank Dr.~Cameron Hird, Adam Johnstone, Rebecca Turner and Jamie
+Haddon for animal husbandry, Dr.~Corin Liddle for advice on calcium
+imaging, and Dr.~Emelie Brodrick and Dr.~Elizabeth Williams for critical
+reading of the manuscript. This work was funded by the Wellcome Trust
+(214337/Z/18/Z). This project has received funding from the European
+Research Council (ERC) under the European Union's Horizon 2020 research
+and innovation programme (grant agreement No 101020792). KJ has been
+supported by a JSPS Overseas Research Fellowship, LAYG by a BBSRC
+Discovery fellowship (BB/W010305/1), PS by a Wellcome Trust
+Institutional Strategic Support Award (204909/Z/16/Z), and KCAW by the
+EPSRC Hub for Quantitive Modelling in Healthcare (EP/T017856/1). We
+thank Dr.~Francesca Carlisle for the mycoplasma testing of COS-7 cells.
 
@@
 \begin{figure}
-\includegraphics[width=16.67in]{figures/Fig1_sup2} \caption{**Figure 1 -- figure supplement 2.
+\includegraphics[width=25in]{figures/Fig1_sup2} \caption{**Figure 1 -- figure supplement 2.
 Expression of *NOS* in three-day-old *Platynereis* larva.** (**A**) HCR in situ for *NOS* (magenta), dorsal view.
@@
 Inset in (B) shows close-up of two *NOS*-positive cells (INNOS_dl and INNOS_vl) on the left side.
-Samples were also stained for DAPI (cyan) to label nuclei.}\label{fig:unnamed-chunk-10}
+Samples were also stained for DAPI (cyan) to label nuclei.(**C**) Expression of the *NOS* gene detected by in situ HCR (magenta) in a three-day-old larva (anterior view).
+Antibody staining for acetylated α-tubulin (acTub: green) highlights cPRC cilia and the neuropil.}\label{fig:unnamed-chunk-10}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=43.06in]{figures/Fig1_sup3} \caption{**Figure 1 -- figure supplement 3.
-Synapse distribution in cPRCs and possynaptic interneurons** (**A**) Volume rendering of the four INNOS neurons with incoming and outgoing synapses.
-(**B**) Axo-dendritic splitting of the INNOS neurons.
-(**C**) Volume rendering of the four INRGW neurons with incoming and outgoing synapses.
-(**D**) Volume rendering of the four cPRC neurons with incoming and outgoing synapses.
+\includegraphics[width=33.33in]{figures/Fig1_sup3} \caption{**Figure 1 -- figure supplement 3.
+Synapse distribution in cPRCs and possynaptic interneurons** (**A**) Axo-dendritic splitting of the INNOS neurons.
+(**B**) Volume rendering of the four INRGW neurons with incoming and outgoing synapses.
+(**C**) Volume rendering of the four cPRC neurons with incoming and outgoing synapses.
 All images are anterior views.
@@
 \begin{figure}
-\includegraphics[width=44.44in]{figures/Fig3_sup1} \caption{**Figure 3 -- figure supplement 1.
+\includegraphics[width=16.67in]{figures/Fig2_sup1} \caption{**Figure 2 -- figure supplement 1.
+Schematic diagrams of the UV stimulation experiment.** (**A**) Left: The larvae were embedded in 2\% agarose seawater with the anterior side down.
+Right: Changes in DAF-FM fluorescence intensities were examined when the ciliary region of the cPRC was stimulated with a 405 nm laser.}\label{fig:unnamed-chunk-12}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=33.33in]{figures/Fig3_sup1} \caption{**Figure 3 -- figure supplement 1.
 Generation and behavioural characterisation of *NOS* CRISPR knockouts.** (**A**) Top: The domain organisation of *Platynereis* NOS protein and exon/intron structure of the *NOS* gene.
@@
 The PAM sequence is shown in grey, stop codons in red.
+**(B, C)** Immunostaining for NOS antibody in NOS mutant (B, NOSΔ11/Δ11; C, NOSΔ23/Δ23) larvae.
+Acetylated α-tubulin staining (green) highlights the neuropil and cPRC cilia.
+Anterior view of a three-day-old larva.}\label{fig:unnamed-chunk-13}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=18.06in]{figures/Fig3_sup2} \caption{**Figure 3 -- figure supplement 2.
+Generation and behavioural characterisation of *NOS* CRISPR knockouts.** (**A**) Schematic diagram of the set-up of the behavioural experiment.
 (**B**) Swimming trajectories of wild type (WT, n=37) and *NOS* mutant (NOSΔ11/Δ11, n=18 and NOSΔ23/Δ23, n=8) two-day-old larvae.
-All trajectories start at 0 x and y position and time 0 corresponding to 10 sec after the onset of 395 nm stimulation from the side.
-(**C**)  Vertical displacement in 30 sec bins of wild type and mutant (NOSΔ11 and NOSΔ23) two-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
-(**D**) Vertical position of batches of wild type and mutant two-day-old larvae over time under 395 nm UV stimulation.
+All trajectories start at 0 x and y position and time 0 corresponding to 10 sec before the onset of 395 nm stimulation from the side.
+(**C**)  Vertical displacement in 30 sec bins of wild type (WT, n=37) and mutant (NOSΔ11/Δ11, n=18 and NOSΔ23/Δ23, n=8) two-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
+(**D**) Comparison of vertical position of wild type and mutant two-day-old larvae after 30 seconds of UV stimulation.
+One-way ANOVA with Dunnett’s multiple-comparison test are shown.
+The wild type (WT, n=52) and *NOS* mutant (*NOSΔ11/Δ11*, n=30 and *NOSΔ23/Δ23*, n=16) larvae.
+(**E**) Vertical position of batches of wild type (12 batches) and mutant (NOSΔ11/Δ11, 8 batches and NOSΔ23/Δ23, 3 batches) two-day-old larvae over time under 395 nm UV stimulation.
 The starting position of each larval trajectory was set to 0.
-(**E-G**) Swimming speed of batches of wild type and mutant two-day old (E) and three-day-old (F) larvae and L-NAME-treated (NOS inhibitor) three-day-old larvae under 395 nm UV stimulation.
-<br> Figure 3 -- figure suppmenet 1 -- source data 1-6.
-Source data for panels B-G.}\label{fig:unnamed-chunk-12}
+One-way ANOVA with Dunnett’s multiple-comparison test are shown.
+*P < 0.05.
+<br> Figure 3 -- figure suppmenet 2 -- source data 1-4.
+Source data for panels B-E.}\label{fig:unnamed-chunk-14}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=41.67in]{figures/Fig3_sup3} \caption{**Figure 3 -- figure supplement 3.
+Generation and behavioural characterisation of *NOS* CRISPR knockouts.** (**A,B**) Swimming speed of batches of wild type and *NOS* mutant two-day-old (WT, n=37, NOSΔ11/Δ11, n=18 and NOSΔ23/Δ23, n=8) (A) and three-day-old (WT, n=32, *NOSΔ11/Δ11*, n=26 and *NOSΔ23/Δ23*, n=47) (B) larvae under 395 nm UV stimulation.
+(**C,D**) Mean swimming speed of wild type and *NOS* mutant two-day-old (WT, n=37, NOSΔ11/Δ11, n=18 and NOSΔ23/Δ23, n=8) (C) and three-day-old (WT, n=32, *NOSΔ11/Δ11*, n=26 and *NOSΔ23/Δ23*, n=47) (D) larvae during 10 seconds before (dark) and 10 seconds after the start of 395 nm UV stimulation (sideUV).
+Data points for the same larva are joined by lines.
+One-tailed paired t-test; p-values (‘***’ <0.001) are shown.
+(**E,F**) Average ciliary beat frequency of wild type and mutant two-day-old (E) and three-day-old (F) single larvae during 10 seconds before (gray) and 10 seconds after the start of 395 nm UV stimulation (purple).
+N=6 larvae each.
+Data points for the same larva are joined by lines.
+One-tailed paired t-test; p-values (‘**’ <0.01 or ‘*’ <0.05) are shown.
+<br> Figure 3 -- figure suppmenet 3 -- source data 1-6.
+Source data for panels A-F.}\label{fig:unnamed-chunk-15}
 \end{figure}
@@
 Connections represent BLAST P-values of <1e-16.
-NIT-GCs, NIT domain containing guanylate cyclases; membrane-bound GCs, membrane-bound guanylate cyclases; sGCs, soluble guanylate cyclases; ACs, adenylate cyclases.}\label{fig:unnamed-chunk-13}
+NIT-GCs, NIT domain containing guanylate cyclases; membrane-bound GCs, membrane-bound guanylate cyclases; sGCs, soluble guanylate cyclases; ACs, adenylate cyclases.}\label{fig:unnamed-chunk-16}
 \end{figure}
@@
 Tree file of the reconstructed GC phylogeny.
-}\label{fig:unnamed-chunk-14}
+}\label{fig:unnamed-chunk-17}
 \end{figure}
@@
 Nuclei were stained with DAPI (cyan).
-(**D**) Immunostaining for NIT-GC1 (magenta) and NOS (green).
+**(D, E)** Immunostaining for NIT-GC1 (D) and NIT-GC2 (E) antibodies in larvae injected with NIT-GC1 (D) or NIT-GC2 (E) morpholinos.
+Acetylated α-tubulin staining (green) highlights the neuropil and cPRC cilia.
 Anterior view of a two-day-old larva.
-Acetylated α-tubulin staining (blue) highlights the neuropil and cPRC cilia.
-**(E, F)** Immunostaining for NIT-GC1 (E) and NIT-GC2 (F) antibodies in larvae injected with NIT-GC1 (E) or NIT-GC2 (F) morpholinos.
-Acetylated α-tubulin staining (green) highlights the neuropil and cPRC cilia.
-Anterior view of a two-day-old larva.}\label{fig:unnamed-chunk-15}
+(**F**) The fluorescence intensities of the NIT-GC1 or NIT-GC2 in the control and morphant (MO1 and MO2) larvae (n=6 each) were normalized to the Acetylated α-tubulin signals and plotted as a relative fluorescence.
+One-way ANOVA with Dunnett’s multiple-comparison test; p-values (‘***’ <0.001) are shown.
+<br> Figure 4 -- figure suppmenet 3 -- source data 1.
+Comparison NIT-GCs signals readings.}\label{fig:unnamed-chunk-18}
 \end{figure}
@@
 Nuceli were stained by DAPI (cyan).
-}\label{fig:unnamed-chunk-16}
+}\label{fig:unnamed-chunk-19}
 \end{figure}
@@
 \begin{figure}
-\centering
-\includesvg{figures/Figure6_fig_Suppl1.svg}
-\caption{\textbf{Figure 6 -- figure supplement 1. Simulated
-Ca\textsuperscript{2+} traces for parameter sets fitted to individual
-Ca\textsuperscript{2+}-recordings collected in wild type, \emph{NOS}
-knockout, and \emph{NIT-GC2} morphant larvae.} (\textbf{A-C}) Simulated
-Ca\textsuperscript{2+} traces in cPRC (A), INNOS (B) and INRGW (C) cells
-in the wild-type condition. (\textbf{D-F}) Simulated
-Ca\textsuperscript{2+} traces in cPRC (D), INNOS (E) and INRGW (F) cells
-in the \emph{NOS}-knockout condition. (\textbf{G-I}) Simulated
-Ca\textsuperscript{2+} traces in cPRC (G), INNOS (H) and INRGW (I) cells
-in the \emph{NIT-GC2}-morphant condition. Thin coloured curves indicate
-individual recordings (cPRC - purple, INNOS - blue and INRGW - green),
-thin grey curves indicate simulated Ca\textsuperscript{2+} traces based
-on model parameters fitted to the individual recording, thick coloured
-curves indicate averages of the recordings and thick black curves
-represent the average of the fits.}
+\includegraphics[width=26.25in]{figures/Figure6_fig_Suppl1} \caption{**Figure 6 -- figure supplement 1.
+Simulated Ca^2+^ traces for parameter sets fitted to individual Ca^2+^-recordings collected in wild type, *NOS* knockout, and *NIT-GC2* morphant larvae.** (**A-C**) Simulated Ca^2+^ traces in cPRC (A), INNOS (B) and INRGW (C) cells in the wild-type condition.
+(**D-F**) Simulated Ca^2+^ traces in cPRC (D), INNOS (E) and INRGW (F) cells in the *NOS*-knockout condition.
+(**G-I**) Simulated Ca^2+^ traces in cPRC (G), INNOS (H) and INRGW (I) cells in the *NIT-GC2*-morphant condition.
+Thin coloured curves indicate individual recordings (cPRC - purple, INNOS - blue and INRGW - green), thin grey curves indicate simulated Ca^2+^ traces based on model parameters fitted to the individual recording, thick coloured curves indicate averages of the recordings and thick black curves represent the average of the fits.
+}\label{fig:unnamed-chunk-20}
 \end{figure}
@@
 \begin{figure}
-\centering
-\includesvg{figures/Figure6_fig_Suppl2.svg}
-\caption{\textbf{Figure 6 -- figure supplement 2. Comparison of fits
-with different fixed parameter sets when fitting the WT model to
-individual recordings from the WT INRGW cells.} Panels A-B, D-E, G-H
-(column fixed parameters) show simulated traces of the
-Ca\textsuperscript{2+} levels in cPRC and INNOS cells produced by
-different fixed parameter sets. Panels C, F and I show individual
-recordings fitted to the data from INRGW cells.}
+\includegraphics[width=26.25in]{figures/Figure6_fig_Suppl2} \caption{**Figure 6 -- figure supplement 2.
+Comparison of fits with different fixed parameter sets when fitting the WT model to individual recordings from the WT INRGW cells.** Panels A-B, D-E, G-H  (column fixed parameters) show simulated traces of the Ca^2+^ levels in cPRC and INNOS cells produced by different fixed parameter sets.
+Panels C, F and I show individual recordings fitted to the data from INRGW cells.}\label{fig:unnamed-chunk-21}
 \end{figure}
@@
 \begin{figure}
-\centering
-\includesvg{figures/Figure6_fig_Suppl3.svg}
-\caption{\textbf{Figure 6 -- figure supplement 3. Comparison of fits to
-the cPRC recordings from NIT-GC2 morphant larvae using the NIT-GC2
-morpholino model or the NOS-knockout model.} Thin coloured curves
-indicate individual recordings, thin grey curves are simulated
-Ca\textsuperscript{2+} traces, thick coloured curves indicate averages
-of the recordings and thick black curves represent the mean of the
-fits.}
+\includegraphics[width=26.25in]{figures/Figure6_fig_Suppl3} \caption{**Figure 6 -- figure supplement 3.
+Comparison of fits to the cPRC recordings from NIT-GC2 morphant larvae using the NIT-GC2 morpholino model or the NOS-knockout model.**  Thin coloured curves indicate individual recordings, thin grey curves are simulated Ca^2+^ traces, thick coloured curves indicate averages of the recordings and thick black curves represent the mean of the fits.}\label{fig:unnamed-chunk-22}
 \end{figure}
@@
 \begin{figure}
-\centering
-\includesvg{figures/Figure6_fig_Suppl4.svg}
-\caption{\textbf{Figure 6 -- figure supplement 4. Distributions of the
-parameter values fitted to the cPRC recordings.} Violin plots (grey
-lines) are kernel density estimates of the underlying distributions;
-computed using the Matlab function \emph{ksdensity} with default
-settings, i.e., using normal kernel function, plots are trimmed to the
-observed range of data. Markers represent individual parameter values,
-white markers indicate medians, vertical grey bars indicate
-interquartile range (Q25 to Q75), and the dashed line indicates the
-value 1. Where violin plots are not shown, these parameters are
-associated with terms/equations that are excluded from the fitting
-procedure (see Methods for details).}
+\includegraphics[width=26.25in]{figures/Figure6_fig_Suppl4} \caption{**Figure 6 -- figure supplement 4.
+Distributions of the parameter values fitted to the cPRC recordings.** Violin plots (grey lines) are kernel density estimates of the underlying distributions; computed using the Matlab function *ksdensity* with default settings, i.e., using normal kernel function, plots are trimmed to the observed range of data.
+Markers represent individual parameter values, white markers indicate medians, vertical grey bars indicate interquartile range (Q25 to Q75), and the dashed line indicates the value 1.
+Where violin plots are not shown, these parameters are associated with terms/equations that are excluded from the fitting procedure (see Methods for details).}\label{fig:unnamed-chunk-23}
 \end{figure}
@@
 \begin{figure}
-\centering
-\includesvg{figures/Figure6_fig_Suppl5.svg}
-\caption{\textbf{Figure 6 -- figure supplement 5. Pairwise correlations
-between parameters of the WT model fitted to WT cPRC recordings}
-Parameters associated with the \(C_R\) equation are not included (see
-Methods for details). Panels on the diagonal show kernel density
-estimates of the distributions of the individual parameters; computed
-using Matlab function \emph{ksdensity} with default settings, i.e.,
-using normal kernel function, plots are trimmed to the observed range of
-data. The other panels show kernel density estimates of bivariate
-distributions of pairs of parameters (heatmaps; darker colours indicate
-higher density). Also shown using markers are the values of the
-individual parameter combinations.}
+\includegraphics[width=26.25in]{figures/Figure6_fig_Suppl5} \caption{**Figure 6 -- figure supplement 5.
+Pairwise correlations between parameters of the WT model fitted to WT cPRC recordings** Parameters associated with the $C_R$ equation are not included (see Methods for details).
+Panels on the diagonal show kernel density estimates of the distributions of the individual parameters; computed using Matlab function *ksdensity* with default settings, i.e., using normal kernel function, plots are trimmed to the observed range of data.
+The other panels show kernel density estimates of bivariate distributions of pairs of parameters (heatmaps; darker colours indicate higher density).
+Also shown using markers are the values of the individual parameter combinations.}\label{fig:unnamed-chunk-24}
 \end{figure}

### 2024-02-20 Manuscript/Manuscript.Rmd
eLife revised 1

--- 2024-01-06 Manuscript/Manuscript.Rmd
@@
 ```{r echo=FALSE,fig.cap="**Figure 1.
-Identification of *NOS*-expressing interneurons (INNOS) within the cPRC circuit.** **(A)** Scanning electron microscopy image of a three-day-old *Platynereis* larva.
-**(B, C)** Volume rendering of the neuron types (cPRC, INNOS, INRGWa, Ser-h1 and MC) in the cPRC circuit reconstructed from a whole-body transmission electron microscopy volume of a three-day-old larva.
+Identification of *NOS*-expressing interneurons (INNOS) within the cPRC circuit.** (**A**) Scanning electron microscopy image of a three-day-old *Platynereis* larva.
+(**B, C**) Volume rendering of the neuron types (cPRC, INNOS, INRGWa, Ser-h1 and MC) in the cPRC circuit reconstructed from a whole-body transmission electron microscopy volume of a three-day-old larva.
 Neurite skeletons are shown with cell-body positions represented by spheres.
@@
 In B, nuclei positions of the prototroch head ciliary band are shown as grey spheres.
-**(D)** Volume rendering of the four INNOS neurons with incoming and outgoing synapses.
-**(E)** Expression of the *NOS* gene detected by in situ HCR (magenta) in a two-day-old larva (anterior view).
+(**D**) Volume rendering of the four INNOS neurons with incoming and outgoing synapses.
+(**E**) Expression of the *NOS* gene detected by in situ HCR (magenta) in a two-day-old larva (anterior view).
 Antibody staining for acetylated α-tubulin (acTub: green) highlights cPRC cilia and the neuropil.
-**(F)** Expression of a *NOS* reporter (NOSp::palmi-3xHA; magenta) labelled with an anti-HA antibody in a two-day-old larva (anterior view).
+(**F**) Expression of a *NOS* reporter (NOSp::palmi-3xHA; magenta) labelled with an anti-HA antibody in a two-day-old larva (anterior view).
 Antibody staining for acetylated α-tubulin (acTub; green) highlights cPRC cilia and the neuropil.
-**(G)** Immunostaining for NOS (magenta), co-stained for acetylated α-tubulin (acTub; green).
-**(H)** Synaptic wiring diagram of the cPRC circuit.
+(**G**) Immunostaining for NOS (magenta), co-stained for acetylated α-tubulin (acTub; green).
+(**H**) Synaptic wiring diagram of the cPRC circuit.
 Hexagons represent cell groups, with the number of cells per group shown in square brackets.
@@
 To test this, first we asked whether NO is produced during UV/violet stimulation of the larvae.
-We injected the fluorescent NO-reporter DAF-FM into zygotes and imaged two-day-old larvae while exposing the region of cPRC cilia to 405 nm violet light (Figure 2---figure supplement 1).
+We injected the fluorescent NO-reporter DAF-FM into zygotes and imaged two-day-old larvae while exposing the region of cPRC cilia to 405 nm violet light (Figure 2A, B).
 To mark cell outlines, we coinjected mRNA encoding a red-fluorescent reporter (RGECO), allowing the identification of the cPRCs.
-Following light stimulation in the region of the ramified cilia of the cPRCs, we detected an increase in DAF-FM fluorescence in the anterior neurosecretory neuropil, the region of INNOS projections.
-This increase did not occur in larvae in which we illuminated a control area (Figure 2).
+Following light stimulation in the region of the ramified cilia of the cPRCs, we detected an increase in DAF-FM fluorescence in the anterior neurosecretory neuropil, the region of INNOS projections (Figure 2C).
+This increase did not occur in larvae in which we illuminated a control area (Figure 2D).
 
 ```{r echo=FALSE,fig.cap="**Figure 2.
-NO produced by UV/violet stimulation to cPRCs.** (**A**) DAF-FM fluorescence in the region of the neurosecretory neuropil.
+NO produced by UV/violet stimulation to cPRCs.** (**A**) Left: The larvae were embedded in 2% agarose seawater with the anterior side down.
+Right: Changes in DAF-FM fluorescence intensities were examined when the ciliary region of the cPRC was stimulated with a 405 nm laser.(**B**) DAF-FM fluorescence in the region of the neurosecretory neuropil.
 The white line indicates the outline of the larva.
@@
 The cPRCs are marked by thin lines.
-(**B**) DAF-FM fluorescence before and during 405 nm light stimulation.
-(**C**) Changes in DAF-FM fluorescence over time during 405 nm stimulation of the cPRCs or a control area (ctr stim.).
+(**C**) DAF-FM fluorescence before and during 405 nm light stimulation.
+(**D**) Changes in DAF-FM fluorescence over time during 405 nm stimulation of the cPRCs or a control area (ctr stim.).
 The purple box indicates the duration of 405 nm stimulation.
@@
 We could establish a homozygous line for both mutations indicating that NOS is not an essential gene in *Platynereis* (Figure 3---figure supplement 1B, C).
+
 To quantify UV avoidance, we recorded the trajectories of freely swimming wild type and mutant larvae in vertical columns, illuminated laterally from two opposite sides with 395 nm UV light (Figure 3A and Figure 3---figure supplement 2A, B).
@@
 
-To distinguish between an acute and developmental function of NOS in light responses, we also tested larvae exposed to the NOS inhibitor L-NAME.
-Larvae incubated for 5 min in 0.1 mM or 1 mM L-NAME showed a dose-dependent inhibition of UV avoidance (Figure 3E, F).
-In contrast, phototaxis was not affected (Figure 3G).
+To distinguish between an acute and developmental function of NOS in light responses, we also tested larvae exposed to the NOS inhibitor L-NAME (Figure 3F).
+Larvae incubated for 5 min in 0.1 mM or 1 mM L-NAME showed a dose-dependent inhibition of UV avoidance (Figure 3G, H).
+In contrast, phototaxis was not affected (Figure 3F).
 Overall, our results indicate an acute requirement for NOS signalling in UV-avoidance and a possible indirect, developmental role in the visual system, reminiscent of the function of NO signalling in *Drosophila* eye development [@Gibbs1998a].
@@
 ```{r echo=FALSE,fig.cap="**Figure 3.
-NOS is required for UV avoidance in *Platynereis* larvae.**  (**A**) Swimming trajectories of wild type (WT, n=32) and *NOS* mutant (*NOSΔ11/Δ11*, n=26 and *NOSΔ23/Δ23*, n=47) three-day-old larvae.
+NOS is required for UV avoidance in *Platynereis* larvae.** (**A**) Schematic diagram of the set-up of the behavioural experiment.
+(**B**) Vertical displacement in 30 sec bins of wild type (10 batches) and mutant (*NOSΔ11/Δ11*, 5 batches and *NOSΔ23/Δ23*, 7 batches) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
+One-way ANOVA with Dunnett’s multiple-comparison test are shown.
+*P < 0.05, **P < 0.01.
+(**C**) Swimming trajectories of wild type (WT, n=32) and *NOS* mutant (*NOSΔ11/Δ11*, n=26 and *NOSΔ23/Δ23*, n=47) three-day-old larvae.
 All trajectories start at 0 x and y position and time 0 corresponding to 10 sec before the onset of 395 nm stimulation from the side.
-(**B**) Vertical position of batches of wild type and mutant three-day-old larvae over time under 395 nm UV stimulation.
+(**D**) Vertical position of batches of wild type and mutant three-day-old larvae over time under 395 nm UV stimulation.
 The starting position of each larval trajectory was set to 0.
-(**C**) Comparison of vertical position of wild type and mutant three-day-old larvae after 30 seconds of UV stimulation.
+(**E**) Comparison of vertical position of wild type and mutant three-day-old larvae after 30 seconds of UV stimulation.
 One-way ANOVA with Dunnett’s multiple-comparison test are shown.
 The wild type (WT, n=32) and *NOS* mutant (*NOSΔ11/Δ11*, n=26 and *NOSΔ23/Δ23*, n=47) larvae.
-(**D**) Vertical displacement in 30 sec bins of wild type (10 batches) and mutant (*NOSΔ11/Δ11*, 5 batches and *NOSΔ23/Δ23*, 7 batches) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
+(**F**) Vertical displacement in 30 sec bins of control (15 batches) and L-NAME-treated (0.1 and 1 mM, 15 batches each) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
 One-way ANOVA with Dunnett’s multiple-comparison test are shown.
 *P < 0.05.
-(**E**) Vertical position of batches of control and L-NAME-treated (0.1 and 1 mM) three-day-old larvae over time under 395 nm UV stimulation.
+(**G**) Vertical position of batches of control and L-NAME-treated (0.1 and 1 mM) three-day-old larvae over time under 395 nm UV stimulation.
 The starting position of each larval trajectory was set to 0.
-(**F**) Comparison of vertical position of control (n=6) and L-NAME-treated (0.1 and 1 mM, n=6 each) three-day-old larvae after 30 seconds of UV stimulation.
+(**H**) Comparison of vertical position of control (n=6) and L-NAME-treated (0.1 and 1 mM, n=6 each) three-day-old larvae after 30 seconds of UV stimulation.
 One-way ANOVA with Dunnett’s multiple-comparison test are shown.
-(**G**) Vertical displacement in 30 sec bins of control (15 batches) and L-NAME-treated (0.1 and 1 mM, 15 batches each) three-day-old larvae stimulated with 395 nm light from the side, 488 nm light from the top and 395 nm light from the top.
-One-way ANOVA with Dunnett’s multiple-comparison test are shown.
-*P < 0.05.
 <br> Figure 3 -- source data 1-7.
@@
 Scripts are available at <https://github.com/JekelyLab/Jokura_et_al_NOS>.
+
+The TrackMate7 plugin [@Ershov2022] in Fiji ImageJ was used to trace the swimming trajectory of the larvae, using the DoG detector for 3 pixel size and the LAP Tracker, filtered at linking max distance: 10.0 pixels, gap closing max distance: 5.0 pixels.
 
@@
 
-```{r echo=FALSE,fig.cap="**Figure 2 -- figure supplement 1.
-Schematic diagrams of the UV stimulation experiment.** (**A**) Left: The larvae were embedded in 2% agarose seawater with the anterior side down.
-Right: Changes in DAF-FM fluorescence intensities were examined when the ciliary region of the cPRC was stimulated with a 405 nm laser."}
+```{r echo=FALSE,fig.cap="**Figure 2 -- figure supplement 1. "}
 knitr::include_graphics(path = "figures/Fig2_sup1.png")
@@
 ```{r echo=FALSE,fig.cap="**Figure 3 -- figure supplement 2.
-Generation and behavioural characterisation of *NOS* CRISPR knockouts.** (**A**) Schematic diagram of the set-up of the behavioural experiment.
-(**B**) Swimming trajectories of wild type (WT, n=37) and *NOS* mutant (NOSΔ11/Δ11, n=18 and NOSΔ23/Δ23, n=8) two-day-old larvae.
+Generation and behavioural characterisation of *NOS* CRISPR knockouts.** (**B**) Swimming trajectories of wild type (WT, n=37) and *NOS* mutant (NOSΔ11/Δ11, n=18 and NOSΔ23/Δ23, n=8) two-day-old larvae.
 All trajectories start at 0 x and y position and time 0 corresponding to 10 sec before the onset of 395 nm stimulation from the side.

### 2024-02-20 code/Fig1.R

@@ -5,7 +5,7 @@
 source("code/Packages_Catmaid_connection.R")
 
 
-# load and plot neurons ---------------------------------------------------
+# load neurons ---------------------------------------------------
 
 cPRC = nlapply(read.neurons.catmaid("^celltype5$", pid=11),
                     function(x) smooth_neuron(x, sigma=20))
@@ -21,6 +21,9 @@ MC = nlapply(read.neurons.catmaid("^celltype9$", pid=11),
                function(x) smooth_neuron(x, sigma=6000))
 prototroch = nlapply(read.neurons.catmaid("^celltype_non_neuronal3$", pid=11),
              function(x) smooth_neuron(x, sigma=6000))
+
+# plot neurons ---------------------------------------------------
+
 
 {
 plot_background()
@@ -49,6 +52,9 @@ clipplanes3d(0, -1, 0, 145000)
 rgl.snapshot("pictures/INNOS_Catmaid.png")
 close3d()
 }
+
+
+
 
 {
 plot_background_ventral2()
@@ -80,6 +86,8 @@ plot3d(outline, WithConnectors = F, WithNodes = F, soma=F, lwd=2,
 rgl.snapshot("pictures/INNOS_Catmaid_ventr.png")
 close3d()
 }
+
+
 
 # get connectivity from CATMAID and plot network --------------------------
 {
@@ -335,6 +343,7 @@ panel_NOS3d_trans <- ggdraw() + draw_image(readPNG("pictures/NOS-promotor_3d_acT
 
 panel_NOS3d_IHC <- ggdraw() + draw_image(readPNG("pictures/IHC_NOS_3d_48.1um_3.png")) +
   draw_label("IHC", x = 0.15, y = 0.99, size = 10) +
+  draw_label("*", x = 0.5, y = 0.35, color='white',size = 18,fontface="plain", alpha = 0.8) +
   draw_label("NOS", x = 0.12, y = 0.9, color="magenta", size = 11, fontface="plain") +
   draw_label("acTub", x = 0.36, y = 0.9, color="green", size = 11, fontface="plain") +
   draw_line(x = c(0.04, 0.4), y = c(0.08, 0.08), color = "white", size = 0.5) +

### 2024-02-20 code/Fig1_sup3.R

@@ -88,7 +88,7 @@ close3d()
 
 
 # assemble figure ---------------------------------------------------------
-INNOS_img <- readPNG("pictures/INNOS_synapses.png")
+
 INNOS_split <- readPNG("pictures/INNOS_axon_dendrite_split.png")
 
 cPRC_img <- readPNG("pictures/cPRC_synapses.png")
@@ -97,31 +97,28 @@ INRGW_img <- readPNG("pictures/INRGW_synapses.png")
 #read png convert to image panel
 arrow <- data.frame(x1 = 0.95, x2 = 0.95, y1 = 0.8, y2 = 0.9)
 
-panel_INNOS <- ggdraw() + 
-  draw_image(INNOS_img) +
+panel_INNOS_split <- ggdraw() + 
+  draw_image(INNOS_split) +
   draw_label("INNOS", x = 0.3, y = 0.99, size = 10) +
-  draw_label("NS plexus", x = 0.485, y = 0.59, size = 8) +
-  draw_label("outgoing", x = 0.9, y = 0.45, size = 10, color='#E69F00') +
-  draw_label("incoming", x = 0.89, y = 0.5, size = 10, color='#0072B2') +
+  draw_label("dendrite", x = 0.8, y = 0.45, size = 10, color='grey40') +
+  draw_label("axon", x = 0.75, y = 0.22, size = 10, color='grey20') +
+  draw_label("*", x = 0.52, y = 0.28, color='black',size = 18,fontface='plain') +
   draw_label("D", x = 0.95, y = 0.93, size = 6) +
   draw_label("V", x = 0.95, y = 0.77, size = 6) +
-  draw_label("*", x = 0.5, y = 0.29, color='black',size = 18,fontface='plain') +
   geom_segment(aes(x = x1, y = y1, xend = x2, yend = y2), data = arrow, 
                arrow = arrow(ends = "both", type = "closed", length = unit(0.1,"cm")),
                lineend = "butt",
                linejoin = "mitre",
                arrow.fill = "black", size = 0.2)
 
-panel_INNOS_split <- ggdraw() + 
-  draw_image(INNOS_split) +
-  draw_label("INNOS", x = 0.3, y = 0.99, size = 10) +
-  draw_label("dendrite", x = 0.8, y = 0.45, size = 10, color='grey40') +
-  draw_label("axon", x = 0.75, y = 0.22, size = 10, color='grey20') +
[…]

### 2024-02-20 code/Fig2.R

@@ -82,36 +82,26 @@ panel_UV_stimulation  <- ggdraw() + draw_image(readPNG("pictures/UV_stimulation.
 #combine panels into Figure and save final figure as pdf and png
 #panels of different sizes
 layout <- "
-A#B#C
+AAAA#B
+######
+C#DDDD
 "
 
-Fig2 <- panel_DAF + panel_DAFFM_intensity + panel_DAFFM + 
+Fig2 <- panel_UV_stimulation + panel_DAF + panel_DAFFM_intensity + panel_DAFFM + 
   patchwork::plot_layout(design = layout, 
-                         widths = c(0.5,0.05,0.28,0.025,0.95)) + #we can change the heights of the rows in our layout (widths also can be defined)
+                         heights = c(1, 0.05, 1), 
+                         widths = c(0.5,0.1,0.5,0.5,0.1,1)) + #we can change the heights of the rows in our layout (widths also can be defined)
   patchwork::plot_annotation(tag_levels = 'A') &  #we can change this to 'a' for small caps or 'i' or '1'
   ggplot2::theme(plot.tag = element_text(size = 12, 
                                          face='plain')) #or 'bold', 'italic'
 
 ggsave("Manuscript/figures/Fig2.png", limitsize = FALSE, 
-       units = c("px"), Fig2, width = 3000, height = 950, bg='white')
+       units = c("px"), Fig2, width = 2000, height = 1700, bg='white')
 
 ggsave("Manuscript/figures/Fig2.pdf", limitsize = FALSE, 
-       units = c("px"), Fig2, width = 3000, height = 950)  
+       units = c("px"), Fig2, width = 2000, height = 1700)  
   
 
 
-
-layout_sup1 <- "
-A
-"
-Fig2_sup1 <- panel_UV_stimulation +
-  patchwork::plot_annotation(tag_levels = 'A') &  #we can change this to 'a' for small caps or 'i' or '1'
-  ggplot2::theme(plot.tag = element_text(size = 12, face='plain')) #or 'bold', 'italic'
-
-ggsave("Manuscript/figures/Fig2_sup1.png", limitsize = FALSE, 
-       units = c("px"), Fig2_sup1, width = 1200, height = 850, bg='white')
-
-ggsave("Manuscript/figures/Fig2_sup1.pdf", limitsize = FALSE, 
-       units = c("px"), Fig2_sup1, width = 1200, height = 850)  
 
 

### 2024-02-20 code/Fig3_and_Fig3_figure_Suppl1.R

@@ -37,6 +37,11 @@ df_L_NAME_tracking <- read_csv("data/230511_df_L_NAME.csv")
 }
 
 
+df_2d_nor$Genotype = factor(df_2d_nor$Genotype, levels=c("WT", "NOS11x11", "NOS23x23", "NOS11x23"))
+df_3d_nor$Genotype = factor(df_3d_nor$Genotype, levels=c("WT", "NOS11x11", "NOS23x23", "NOS11x23"))
+
+label <- c("WT", "NOS^'Δ11/Δ11'", "NOS^'Δ23/Δ23'")
+
 
 # tidying Martin data (30 sec bins, vertical displacement)------------------------------------------------
 
@@ -134,7 +139,7 @@ WTvsNOS_2d %>%
                expression('NOS'^'Δ23/Δ23'))) +
   theme_minimal()+
   theme_plot +
-  labs(fill = "genotype", x = "time [s]", y = "Vertical displacement [mm/s]")+
+  labs(fill = "two-day-old", x = "time [s]", y = "Vertical displacement [mm/s]")+
   scale_y_continuous(limits = c(-0.46, 0.56)) +
   scale_x_discrete(labels=c("30", "60", "90", 
                             "30", "60", "90","120",
@@ -655,7 +660,7 @@ WTvsNOS_3d %>%
                expression('NOS'^'Δ23/Δ23'))) +
   theme_minimal()+
   theme_plot +
-  labs(fill = "genotype", x = "time [s]", y = "Vertical displacement [mm/s]")+
+  labs(fill = "three-day-old", x = "time [s]", y = "Vertical displacement [mm/s]")+
   scale_y_continuous(limits = c(-0.81, 0.5)) +
   scale_x_discrete(labels=c("30", "60", "90", 
                            "30", "60", "90","120",
@@ -663,28 +668,41 @@ WTvsNOS_3d %>%
                            "30", "60", "90","120",
                            "30", "60", "90","120",
                            "30", "60", "90","120")) +
-  annotate("text", x=5.255, y=0.07, label="*", color="black", size = 5, fontface="plain") +
+  annotate("text", x=4, y=-0.09, label="*", color="black", size = 5, fontface="plain") +
[…]

### 2024-02-20 code/Packages_to_load.R

@@ -25,6 +25,7 @@ Sys.setenv('R_MAX_VSIZE'=800000000)
   library(ggpubr)
   library(gridExtra)
   library(ggbeeswarm)
+  library(viridis)
 }
 
 

### 2025-06-27 Manuscript/revision/Response to the Reviews.Rmd
initiate revision

--- 2024-01-06 Manuscript/Response to the Reviews.Rmd
@@
 
-1.  As you point out, we have performed statistical tests and added explanatory text to figure legends.
-2.
-Only one NOS homologue has been found in the Platynereis genome.
-Note that it is specifically expressed in NOS interneurons, which are our target in this study, based on the results of single-cell transcriptome analysis and in situ hybridisation against the NOS gene.
-However, as you have pointed out, the functions of nitric oxide are diverse and we have revised some of our discussion to include the possibility of its interneurons outside the non-visual system.
+Response: Thank you for this assessment of our work. We have now added the results of statistical tests to the figures and included explanatory text in the figure legends.
+Regarding the cell-type-specific effects, in the larval stage NOS is only expressed in the INNOS cells and the pigmented eyes.
+We could previously show that UV avoidance is only mediated by the cPRC circuit (including INNOS) whereas phototaxis is exclusively mediated by the pigmented eyes (Jekely et al 2008, Randel et al 2014 and Veraszto el al.
+2018).
+Due to this behavioural specificity, we are therefore confident that the effects of the NOS mutation on UV avoidance are due to the lack of NOS from the INNOS cells and not the pigmented eyes.
+Besides UV avoidance, we have characterised the speed of ciliary swimming without a light stimulus, as well as phototaxis (Randel et al 2014).
+Likewise, the effects of NIT-GC-knockdown can only be due to cell-type-specific effects in the cPRCs, since the two NIT-GC genes are only expressed in the four cPRCs, as we demonstrated by HCR, immunostainings and the analysis of single-cell RNAseq data.
+Given that we used two morpholinos per gene, it is also unlikely that the observed changes in Ca-dynamics are due to off-target effects.
+
+We have now updated the discussion to better address these points.
 
@@
 I could imagine that a large family of models would arrive at this same result, and without units, there is no way to really test it with new behavioral experiments.*
+
+Response:
+We have now carried out a variance-based sensitivity analysis that is included as Figure...
+
+In the model, most parameters are identifiable. If we fix only to of the parameters (C  and ..) all other parameters can be identified.
+
+The model is unitless, since already our input data from calcium imaging represent normalised data and the model was fit to these data.
+We would need a lot more information to build up a proper ground-up biophysical model (e.g.
+including capacitance, ionic concentrations etc.).
+
+We have created a pipeline to carry out- local and global sensitivity analysis
+
+all but two of our parameters were identifiable.
+
+In terms of the constrains, since we don't have dimensioned quantitites, our constraints are bounds on the parameters.
+However, the structural identifibiality analysis tells us where we can identify parameters
+guard against sloppiness and overfitting
+
+Lots of models: 
+we only fitted the model on a subset of the data and can reproduce dynamics on a larger set of data
+
+A family of models that can capture the same mechanism that we are postulating 
+for important mechanistic interactions (the signs and some interactions that are not possible) are constraint. And the model shows that we don't need them or the signs are constrained.
+
+In the future, we can use this model to generate prediction e.g. about what the response to repeated stimulations can be.
+
+To map out the components 
+we can rule out things - e.g. we could rule out a very simple model of progression from step to step
+
+The modelling also completely changed the way we were thinking about the system.
+We needed to include a time-limited step in the cPRC transduction cascade leading to NOS activation to capture the time-invariance of the peak.
+
+Many of the couplings in our model are constrained (e.g.
+UV leads to INNOS activation) and e.g.
+reversing the sign of some of these couplings would not arrive at this same result.
+Several earlier variants of the model could not be fit to the data, therefore we disagree that the model is poorly constrained and cannot be tested with behavioural experiments.
+
+In the revised version, we included a new series of measurements of calcium dynamics in the cPRCs under varying duration and intensity of UV light.
+These important new results gave us further insights and necessitated a revision of the model.
+
+We can reproduce all the qualitative dynamics
 
@@
 
-As noted, we have performed statistical analyses and added to the figures in all behavioural experiments. We have also corrected the wording in the text.
+Response: We thank the reviewer for these comments. We have changed the wording and also added the results of statistical analyses to the figures and explanations to the figure legends.
 
@@
 
-The size of the chamber is limited due to the experimental system, so many larvae tend to accumulate at the bottom of the chamber by descending swimming at the end of the experiment.
-This may have led to variability in the data at the end point and lengthened the error bars.
-
-A 120 s exposure time was most suitable for quantifying larval behaviour from a previous paper by Verasztó et al. 2017.
+Response: The 120 seconds time frame of exposure takes into account the reaction time and swimming speed of the larvae as well as the size of the assay chamber (160 mm water height).
+By the end of a 120 stimulation many larvae tend to accumulate at the bottom of the chamber due to downward swimming and cannot be further tracked.
+This effect leads to higher variability in the data towards the end of the experiment in the wild-type batches.
+We have showed both continuous vertical data as well as the binned data.
+The 30 sec bin was chosen for convenience and for better comparison with our previous paper on UV avoidance behaviour (Verasztó et al.
+2018)
 
@@
 
-We have newly analysed the ciliary beat frequency of larvae by UV stimulation and added Figure 3---figure supplement 3.
+We thank the reviewer for this suggestion.
+We have carried out further experiments and analysed the ciliary beat frequency (CBF) of larvae exposed to UV stimulation.
+We added these data to Figure 3---figure supplement 3.
+The results (increase of CBF under UV in wild-type but not NOS mutant larvae) were quite surprising to us and falsified our initial hypothesis.
+We have rewritten the discussion to reflect this important new finding.
 
-For calcium imaging under pressurised conditions, this was done in Bezares-Calderón et al., 2023, see: doi: <https://doi.org/10.1101/2023.02.28.530398>
+The response of cPRC cilia to changes in hydrostatic pressure has been extensively documented in our recent paper on the mechanism of barotaxis in the Platynereis larva (see Bezares-Calderón et al., https://doi.org/10.1101/2023.02.28.530398).
 
